### 2018-04-24 summarystats/solanaceaedescriptivev2.Rmd
first organization of data

---
title: "Descriptive analysis of Solanaceae"
author: "Rosana Zenil-Ferguson"
date: "4/20/2018"
output:
  html_document: default
  pdf_document: default
---
```{r,error=FALSE, message=FALSE, warning=FALSE}
library("knitr")
setwd('~/Dropbox/solploidy/basicdata')
library("ape")
library("phytools")
library("treeplyr")

solanaceae.tree<-read.tree("solanaceae2013.tre")

#idnumber, source,fullname, genus, species,cultivarhybrid,csomenumber,lifehistoryclean, ploidy #selfincomp, ploidybyZF
solanaceae.allrecords<-read.csv("solanaceaerecords.csv",header=TRUE,sep=",", stringsAsFactors=FALSE, na.strings="")

species<-unique(solanaceae.allrecords$fullname)
total.species<-length(species)# 2278 total species
genera<-unique(solanaceae.allrecords$genus)
total.genera<-length(genera)#97 genera

Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

mode.csome<-rep(0,total.species)
min.csome<-rep(0,total.species)
mode.ploidy<-rep(0,total.species)
min.ploidy<-rep(0,total.species)
self.incomp<-rep(0,total.species)
life.hist<-rep(0,total.species)

for (i in 1:total.species){
  index<-which(solanaceae.allrecords$fullname==species[i])
  aux1<- solanaceae.allrecords$csomenumber[index]
  if(all(is.na(aux1))){
    mode.csome[i]<-NA
    min.csome[i]<-NA
  }else{
    mode.csome[i]<-min(Mode(aux1))
    min.csome[i]<-min(aux1,na.rm=TRUE)
  }
  aux2<- solanaceae.allrecords$ploidy[index]
  if(all(is.na(aux2))){
    mode.ploidy[i]<-NA
    min.ploidy[i]<-NA
  }else{
    mode.ploidy[i]<-min(Mode(aux2))
    min.ploidy[i]<-min(aux2,na.rm=TRUE)
  }
  self.incomp[i]<-solanaceae.allrecords$selfincomp[index[1]]
  life.hist[i]<-solanaceae.allrecords$lifehistoryclean[index[1]]
}

simplified.solanaceae<-data.frame(species, min.csome, mode.csome,min.ploidy,mode.ploidy,self.incomp, life.hist)

per.genus.counts<-list()
per.genus.table<-list()
per.genus.countploidy<-list()
per.genus.ploidy<-list()
for(j in 1:total.genera){
  index<-which(solanaceae.allrecords$genus==genera[j])
  per.genus.counts[[j]]<-data.frame(species=solanaceae.allrecords$fullname[index], csome.number=solanaceae.allrecords$csomenumber[index])
  per.genus.table[[j]]<-table(per.genus.counts[[j]]$species, per.genus.counts[[j]]$csome.number)
  per.genus.countploidy[[j]]<-data.frame(species=solanaceae.allrecords$fullname[index], ploidy=solanaceae.allrecords$ploidy[index])
   per.genus.ploidy[[j]]<-table(per.genus.countploidy[[j]]$species, per.genus.countploidy[[j]]$ploidy)
}
names(per.genus.counts)<-genera
names(per.genus.countploidy)<-genera
#Bar plots that show the variance of chromosome numbers per genus

for(j in 1:length(genera)){

  aux1<-colnames(per.genus.table[[j]])
  if(length(aux1)>0){
  #print(per.genus.table[[j]])
  #print(kable(per.genus.table[[j]],format = "markdown"))
  #kable(aux2,format = "markdown")
  barplot(per.genus.table[[j]],main=names(per.genus.counts)[j],border="white")
  }
  aux2<-colnames(per.genus.ploidy[[j]])
  if(length(aux2)>0){
  #print(kable(per.genus.countploidy[[j]],format = "markdown"))
  #kable(aux2,format = "markdown")
  barplot(per.genus.ploidy[[j]],main=names(per.genus.countploidy)[j],border="white")
  }
  }

```

### 2018-04-24 summarystats/solanaceaesummarystats.R

@@ -1,0 +1,108 @@
+setwd('~/Dropbox/solploidy/basicdata')
+library("ape")
+library("phytools")
+library("treeplyr")
+solanaceae.tree<-read.tree("solanaceae2013.tre")
+
+
+#idnumber, source,fullname, genus, species,cultivarhybrid,csomenumber,lifehistoryclean, ploidy selfincomp, ploidybyZF
+solanaceae.allrecords<-read.csv("solanaceaerecords.csv",header=TRUE,sep=",", stringsAsFactors=FALSE, na.strings="")
+
+species<-unique(solanaceae.allrecords$fullname)
+total.species<-length(species)# 2278 total species
+genera<-unique(solanaceae.allrecords$genus)
+total.genera<-length(genera)#97 genera
+
+Mode <- function(x) {
+  ux <- unique(x)
+  ux[which.max(tabulate(match(x, ux)))]
+}
+
+mode.csome<-rep(0,total.species)
+min.csome<-rep(0,total.species)
+mode.ploidy<-rep(0,total.species)
+min.ploidy<-rep(0,total.species)
+self.incomp<-rep(0,total.species)
+life.hist<-rep(0,total.species)
+
+for (i in 1:total.species){
+  index<-which(solanaceae.allrecords$fullname==species[i])
+  aux1<- solanaceae.allrecords$csomenumber[index]
+  if(all(is.na(aux1))){
+  	 mode.csome[i]<-NA
+  	 min.csome[i]<-NA  	
+  }else{
+  mode.csome[i]<-min(Mode(aux1))
+  min.csome[i]<-min(aux1,na.rm=TRUE)
+  }
+  aux2<- solanaceae.allrecords$ploidy[index]
+  if(all(is.na(aux2))){
+  	 mode.ploidy[i]<-NA
+  	 min.ploidy[i]<-NA  	
+  }else{
+  mode.ploidy[i]<-min(Mode(aux2))
+  min.ploidy[i]<-min(aux2,na.rm=TRUE)
+  }	
+  self.incomp[i]<-solanaceae.allrecords$selfincomp[index[1]]
+  life.hist[i]<-solanaceae.allrecords$lifehistoryclean[index[1]]
+}
+
+
[…]

### 2018-04-25 basicdata/solanaceaetreecorrection.R
datasets for ploidy classification

@@ -1,0 +1,26 @@
+setwd('~/Dropbox/solploidy/basicdata')
+library("ape")
+library("phytools")
+library("treeplyr")
+
+## Time-tree from Sarkinen et al. 2013 solaneaceae done with Beast
+solanum.sarkinen<-read.tree(file="solanumbeast.tre")
+## Names for the solaneaceae tree, the tree above was listed with numbers which was less than ideal
+solanum.names<-read.csv("speciesnames2.csv",header=FALSE,sep=",",stringsAsFactors=FALSE)
+## Putting the names on the tip labels
+fixingtiplabels<-as.numeric(solanum.sarkinen$tip.label)
+solanum.sarkinen$tip.label<-solanum.names[fixingtiplabels,2] 
+# I double checked with their publication turns out to be exact tree shown in figure 12862_2013_2443_MOESM2_ESM.tiff
+
+#Tips to drop/ repeated species list from EEG
+remove.names<-read.csv("tipstodrop.csv",header=FALSE, sep=",",stringsAsFactors=FALSE)
+length(remove.names[,1])#55 to drop
+solanaceae.correcttips<-drop.tip(solanum.sarkinen,remove.names[,1])
+# This corrected tree has 1023 tips, it dropped 53 out 55
+write.tree(solanaceae.correcttips, file="solcorrecttip.tre")
+
+##Plot
+plotTree(solanaceae.correcttips,part=0.98,fsize=0.05, lwd=0.9)
+# The tree is approximately 30my
+h<-max(nodeHeights(solanaceae.correcttips))
+#axis(1,labels=c(136.9,68.4,0), pos=-0.03*h,at=seq(0,h,by=h/2),lwd=0.2,cex=0.2)

### 2018-04-25 summarystats/solanaceaesummarystats (Autosaved).R

@@ -106,6 +106,8 @@ matched2<-make.treedata(solanaceae.tree, csomefreq2, name_column="taxa")
 csomeseriesdat2<-data.frame(matched2$dat)
 row.names(csomeseriesdat2)<-matched2$phy$tip.label
 
+write.tree(matched2$phy,file="~/Dropbox/solploidy/csomeseries/soltreecsometaxa.tre")
+write.csv(csomeseriesdat2,file="~/Dropbox/solploidy/csomeseries/soldatacsometaxa.csv")
 
 
 ##TRYING TO PLOT THE DATA BUT NOT READY YET

### 2018-05-01 rev_analyses/src/generate_chromo_data.R
generating data file for rev analysis

@@ -1,0 +1,72 @@
+
+#
+# Generate chromosome data input file for RevBayes analysis.
+#
+# Will Freyman
+#
+
+in_file = "../csomeseries/soldataalltaxa.csv"
+out_file = "data/chromo_counts.tsv"
+out_file_notes = "data/chromo_counts_notes.txt"
+
+d = read.csv(in_file, stringsAsFactors=FALSE)
+a = FALSE
+a_notes = FALSE
+max_c = 0
+min_c = 1000
+
+for (i in 1:nrow(d)) {
+
+    counts = vector()
+
+    # chromo counts are in columns 2 thru 16
+    # of original data file
+    for (j in 2:16) {
+        
+        c = d[i,j]
+        if (!is.na(c) & c %% 2 == 0) {
+            
+            c = c/2
+            counts = c(counts, c)
+            if (c > max_c)
+                max_c = c
+            if (c < min_c)
+                min_c = c
+
+        } else if (!is.na(c) & c %% 2 != 0) {
+    
+            write(paste0(d[i,1], ": dropped odd diploid chromo count = ", c),
+                  file=out_file_notes,
+                  append=a_notes)
+            a_notes = TRUE
+
+        }
+
+    }
+
+    if (length(counts) == 0) {
+
+        counts = paste0(d[i,1], "\t", "?")
+
+    } else if (length(counts) == 1) {
+        
+        counts = paste0(d[i,1], "\t", counts)
+
+    } else {
+
+        counts = paste0(counts, collapse=" ")
+        counts = paste0(d[i,1], "\t", "(", counts, ")")
+
+    }
+
+    write(counts, file=out_file, append=a)
+    a = TRUE
+
+}
+
+write(paste0("min haploid chromosome count = ", min_c),
+             file=out_file_notes,
+             append=a_notes)
+write(paste0("max haploid chromosome count = ", max_c),
+             file=out_file_notes,
+             append=a_notes)

### 2018-05-03 manuscript/introduction.tex
a skeleton manuscript structure
I don't know what conventions everyone likes, but Rosana asked for a
starting point so here it is.  Modify as you see fit, of course.

\section{Introduction}

Classic work gets cited \citep{stebbins1957}.
Also, when working with tex and git, it is very helpful if each sentence starts on a new line.
Like this (except that was a fragment).

### 2018-05-03 manuscript/main.tex
Manuscript format basic tex

\documentclass[11pt]{article}
\usepackage{evolution}

\begin{document}

%--------------------------------------------------
% Titlepage
%--------------------------------------------------

\begin{center}
    \textbf{TITLE}
\end{center}

\vfill

\noindent
Someone's Name,%
\endnote{Someone's Address}
%
and
Another Name%
\endnote{Another Address}

\vfill

\theendnotes

\noindent
Author for correspondence: XXX

\vfill

\noindent
\textit{Running head:} XXX

\vfill

\noindent
\textit{Keywords:} 
XXX,
YYY

\vfill

\linenumbers

%--------------------------------------------------
% Abstract
%--------------------------------------------------

\clearpage
\input{abstract}

%--------------------------------------------------
% Main text
%--------------------------------------------------

\clearpage
\input{introduction}
\input{methods}
\input{results}
\input{discussion}

%--------------------------------------------------
\section{Acknowledgements}
%--------------------------------------------------

%--------------------------------------------------
% References
%--------------------------------------------------

\clearpage

\setstretch{1}
\bibliography{refs}
\bibliographystyle{evolution}
\setstretch{\stretchby}

\end{document}

### 2018-05-03 summarystats/solanaceaesummarystats (Autosaved).R
updated file formats

@@ -92,13 +92,15 @@ a<-hist(ploidy.dataset$mode.ploidy)
 
 write.tree(ploidy.tree,file="~/Dropbox/solploidy/basicdata/ploidy.tre")
 write.nexus(ploidy.tree, file="~/Dropbox/solploidy/basicdata/ploidy.nex")
-tip.names<-ploidy.tree$tip.labels
+tip.names<-ploidy.tree$tip.label
 write.table(ploidy.dataset$mode.ploidy,file="~/Dropbox/solploidy/basicdata/ploidy.txt",sep=',',row.names=FALSE, col.names=TRUE)
 aux1<-which(ploidy.dataset$mode.ploidy==2)#360 diploids
 aux<-which(ploidy.dataset$mode.ploidy!=2)#63 polyploids 14.8% of the sample
 binary.var<-rep(1,423) 
 binary.var[aux1]=0
-write.table(binary.var,file="~/Dropbox/solploidy/basicdata/binaryploidy.txt",sep=',',row.names=FALSE, col.names=TRUE)
+binary.var<-data.frame(binary.var)
+binary.table<-cbind(tip.names, binary.var)
+write.table(binary.table,file="~/Dropbox/solploidy/basicdata/binaryploidy.tsv",sep="\t",row.names=FALSE, col.names=FALSE)
 
 ###########
 

### 2018-06-07 summarystats/solanaceaedescriptivev2.Rmd
results SSE polyploidy

@@
 date: "4/20/2018"
+#last modified 06/06/2018
 output:
@@
 library("knitr")
-setwd('~/Dropbox/solploidy/basicdata')
+setwd('~/Dropbox/solploidy')
 library("ape")

### 2018-06-07 twostateSSE/bisseplotresults.R

@@ -1,32 +1,14 @@
-source("https://bioconductor.org/biocLite.R")
-biocLite("ggtree")
-biocLite("treeio")
+#source("https://bioconductor.org/biocLite.R")
+#biocLite("ggtree")
+#biocLite("treeio")
 
 library(RevGadgets)
 library(ggplot2)
 # plot the ancestral states tree generated by RevBayes
 
-p = plot_ancestral_states("~/revbayes-master/projects/cmake/output/anc_states_summaryBiSSE_polydip.tree")
 
-size=0.4,
-
-include_start_states=TRUE, 
-
-summary_statistic="MAPChromosome", 
-
-tip_label_size=3, 
-
-tip_label_offset=0.4,
-
-tip_label_italics=TRUE,
-
-node_label_size=2,
-
-node_label_nudge_x=0.1,
-
-shoulder_label_size=2,
-
-alpha=.3)
+p = plot_ancestral_states("~/revbayes-master/projects/cmake/output/anc_states_summaryBiSSE_polynodip.tree", size=0.4, include_start_states=FALSE, summary_statistic="MAP",tip_label_size=0.5, tip_label_offset=0.4,tip_label_italics=TRUE,node_label_size=2,node_label_nudge_x=0.1,shoulder_label_size=2,alpha=.3,node_size_range=c(2, 6))
+print(p)
 
 
 

### 2018-09-14 manuscript/introduction.tex
Adding first manuscript draft

@@
 \section{Introduction}
+Linking changes in diversification to polyploidy and breeding system has been one of the key botany questions for the past n number of years.
+The prevalence of polyploidy, and the multiple evolutionary changes of breeding systems in important angiosperm clades, leads to the inevitable question of the evolutionary changes.
 
-Classic work gets cited \citep{stebbins1957}.
-Also, when working with tex and git, it is very helpful if each sentence starts on a new line.
-Like this (except that was a fragment).
+- Diversification and polyploidy\newline
+An important debate regarding the diversification of angiosperms has been ongoing since the publication of \citet{mayrose_2011}.
+The authors discovered that using the latest diversification model linked to two states diploid and polyploid, the net diversification of polyploids was much slower than the net diversification rate than polyploids.
+This result was surprising and an sparked a discussion about the long-term evolutionary consequences of polyploidy.
+\citet{soltis_2014} questioned if polyploidy should be regarded as an evolutionary dead-end since the net diversification of polyploids was negative, despite overwhelming evidence about the incidence of polyploidy  especially at the root of highly diverse angiosperm claims (ref here).
+A year later, diversification models were re-tested and corrected, and still found the same pattern \citep{mayrose_2015}, to the disappointment of plenty of botanists there was no denying of the weak trend of diversification that polyploids left behind.
+\newline
+There are two key questions that at the time of the polyploidy debate were difficult to ask.
+The first is if the models used to measure the diversification of polyploids were correct, and the second question is if the models have potentially included more evidence and potential traits that are not polyploidy or other lines of evidence that could be driving the patterns.
+At the time, in a different context \citet{beaulieu_2016} were finding an alternative solution to the first question, coming up with a new model that could represent the broad heterogeneity of the diversification process and parse out the signal between the trait of interest and the noise in diversification.
+Their model, the hidden state speciation and extinction model is a key component to detect whether polyploidy our something else unknown but related to it is driving the speciation and extinction patterns that we see in angiosperms.
+\newline
+
+The second question (talk here about breeding system and polyploidy, and how this could be different across clades and this is one of the reasons why we focused on Solanaceae).
+-Diversification and breeding systems\\
+Goldberg and Igic 2012, different  perspectives\\
+
+-What we did\\
+We investigate the relative importance of both b
+
+Debate of Mayrose

### 2018-09-14 manuscript/main.tex
Adding first manuscript draft

@@
 \usepackage{evolution}
+\usepackage{hyperref}
 
@@
 \begin{center}
-    \textbf{TITLE}
+    \textbf{When your favorite trait is not enough to explain diversification: An example of diversification linked to polyploidy and breeding system in Solanaceae}
 \end{center}
@@
 \noindent
-Someone's Name,%
-\endnote{Someone's Address}
+Rosana Zenil-Ferguson,%
+\endnote{University of Minnesota Twin Cities}
+%
+\noindent
+J. Gordon Burleigh,%
+\endnote{University of Florida}
+%
+\noindent
+William A. Freyman,%
+\endnote{23 and me}
+%
+\noindent
+Boris Igi\'c,%
+\endnote{University of Illinois}
+%
+\noindent
+Itay Mayrose,%
+\endnote{Tel Aviv University}
 %
 and
-Another Name%
-\endnote{Another Address}
+Emma E. Goldberg%
+\endnote{University of Minnesota Twin Cities}
 
@@
 \noindent
-Author for correspondence: XXX
+Author for correspondence: Rosana Zenil-Ferguson
 
@@
 \noindent
-\textit{Running head:} XXX
+\textit{Running head:} Polyploidy and breeding systems Solanaceae
 
@@
 \textit{Keywords:} 
-XXX,
-YYY
+Polyploidy,
+Breeding System,
+Diversification, SSE models
 

### 2018-09-14 manuscript/methods.tex
Adding first manuscript draft

\section{Methods}
\subsection{Data}
Chromosome number data was obtained from the Chromosome Count Database (CCDB, \citet{rice_2015}) for all taxa from Solanaceae family. The records obtained were first cleaned using CCDBcurator R package \citep{zenilferguson_2017}, to those we aggregated independent data of chromosome numbers and breeding systems recorded and curated by Emma Goldberg and Boris Igi\'c  (citations for original sources can be found in the supplementary information). In total we obtained 14,000 records of chromosome number and breeding system than were curated semi automatically by CCDBcurator and by hand to score the chromosome number and the breeding system with the most reconciliation. Furthermore, for the majority of taxa ploidy was assigned according to information from original sources and Kew Royal Botanic Gardens C-value DNA (bennett citation here dataset and the original sources agreggated by Goldberg and Igi\'c. For taxa without ploidy information but with information about chromosome number we assigned ploidy based on multiplicity of chromosomes within the genus. For example, \textit{Solanum betaceum} did not include information about ploidy level but its number of chromosomes is 24, since $x=12$ is the base chromosome number of the \textit{Solanum} genus we assigned  \textit{Solanum betaceum} as diploid. Species with more than one ploidy level were assigned smallest most frequent ploidy level recorded. Domesticated taxa (i.e. \textit{Solanum lycopersicon}) were removed to avoid since they contained multiple ploidy levels due to domestication and not spontaneous polyploidy. The remainder of taxa was then matched to largest time-calibrated  phylogeny of Solanaceae and this process resulted in 595 taxa used in each of the diversification models.

Binary or three state classification of ploidy and breeding system for the 595 taxa  was done according to Table \autoref{tab:binaryclassifications}

\subsection{Models}
We first defined a binary state speciation and extinction model (BiSSE, \citet{maddison_2007})  for polyploidy evolution where taxa was classified as diploid (D=0) and polyploid (P=1) \ref{datatable}. In  a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$) , net diversification rates ($r_D=\lambda_D-\mu_D, r_P=\lambda_P-\mu_P$), and  relative extinction rates ($\nu_D=\frac{\mu_D}{\lambda_D}, \nu_D=\frac{\mu_D}{\lambda_D}$) just as previously explored in \citet{mayrose_2011}. The cladogenetic changes explored in BiSSE were polyploidization rate are represented by parameter $\rho$ and by diploidization parameter rate $\delta$.\newline

As a second step, we fitted a hidden state speciation and extinction model (HiSSE, \citet{beaulieu_2016}) to evaluate whether the differences in diversification rates were found due to a hidden trait associated to polyploidy. As discussed by \citet{beaulieu_2016} BiSSE-like models suffer from a large type I error because they fail to account as part of the null hypothesis heterogeneous diversification rate changes  that do not depend on the trait of interest. By including a hidden trait in a state speciation and extinction  model,  HiSSE-like models address the heterogeneous background diversification changes while also parsing out the possible signal of diversification due to the trait of interest. Therefore, HiSSE model for polyploidy linked to diversification has four states: diploid and polyploid subdivided by a binary hidden trait with states A and B for which we estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\lambda_{D_B}, \lambda_{P_A},\lambda_{P_B}$), extinction rates ($\mu_{D_A},\mu_{D_B}, \mu_{P_A},\mu_{P_B}$),  net diversification rates ($r_{D_A},r_{D_B},r_{P_A},r_{P_B}$), and relative extinction($\nu_{D_A},\nu_{D_B}, \nu_{PA},\nu_{P_B}$). The cladogenetic changes assumed by the fitted model are polyploidization rate $\rho$, diploidization rate $\rho$, and we assumed that the changes between every hidden state are symmetrical with rate $\alpha$. \newline

For breeding system both BiSSE and a HiSSE models were fitted independently by coding data as self-compatible (SC=0) or self-incompatible (SI=1) \ref{datatable}. The BiSSE model fitted to breeding system data was done with the goal of investigating if the effect of self-incompatibility in diversification was similar to the pattern found by  \citet{goldberg_2012}. However, the HiSSE model on breeding system parses out if the effect of breeding system in diversification is significant  by minimizing type I error as mentioned above. For these models we assumed that self-compatibility is irreversible as discussed by \citet{igic_2013}, so we defined the transition rate $q_{IC}$ as the parameter for changes from self-incompatible to self-compatible state.\newline

Next, we proposed a multivariate speciation and extinction model (MuSSE, \citet{fitzjohn_2012}) to investigate the link amongst diversification, breeding system, and polyploidy simultaneously. The MuSSE model is defined using three states self-compatible diploids (SC-D=0), polyploids that are always self-compatible (ref for this??) (P=1), and self-incompatible diploids (SI-D=2, see table \ref{datatable}) and contains ten parameters, six defining diversification ($\lambda_{SD}, \mu_{SD},\lambda_{P},\mu_{P}, \lambda_{ID},\mu_{ID}$) and the other four key cladogenetic changes that are: polyploidization of self-compatible diploids $\rho_{SD}$, diploidization $\delta$, polyploidization of self-incompatible diploids $\rho_{ID}$, and self-incompatible to self-compatible rate $q_{IC}$.\newline

Since the null hypothesis of the MuSSE model is that the diversification is equal and constant for all three states defined, it is possible that MuSSE can also suffer from large type I errors. In order to account for heterogeneity of the diversification rates and parse the signal in diversification coming from breeding system and polyploidy we extended MuSSE model to account for a hidden state, as \citet{beaulieu_2016} did in the bivariate case. The model we proposed is a multivariate and hidden states speciation and extinction stochastic process (MuHiSSE) that allowed us to account simultaneously for the diversification rates linked to breeding system and polyploidy but also the presence of some more heterogeneity in the process.  This model is analogous to the GeoHiSSE model proposed by Caetano et al. (2018) in a biogeographical context. The full MuHiSSE model has 26 parameters, however, our goal was to look for diversification rate differences so we fitted a simplified version of 17 parameters by fixing the rates amongst hidden states to be equal with parameter $\alpha$ and the transition rates amongst breeding system and polyploidy as defined in the MuSSE model  ($\rho_{SD}$,  $\delta$,  $\rho_{ID}$,  $q_{IC}$) despite the hidden state. Using MuHiSSE we estimated twelve speciation and extinction rates ($\lambda_{SD_A}, \mu_{SD_A},\lambda_{P_A},\mu_{P_A}, \lambda_{ID_A},\mu_{ID_B},\lambda_{SD_B}, \mu_{SD_B},\lambda_{P_B},\mu_{P_B}, \lambda_{ID_B},\mu_{ID_B}$), and the net diversification and relative extinction rates associated with them.\newline

All the models were performed using RevBayes \citep{hoehna_2016} software that performs Bayesian inference via MCMC using Metropolis-Hastings algorithm. A correction for sampling bias was done in all models by assuming that Solanaceae family has approximately 3,000 species ($s=595/3000$) as the Solanaceae Source project indicates \citep{source_2011}.  In all models, speciation and extinction parameters used log-normal prior distributions that with mean the expected net diversification rate $ ( \frac{(number of taxa)/2}{root age})$ and standard deviation $0.5$. Prior distributions for parameters defining cladogenetic changes were gamma distributed with parameters $k=0.5$ and $\theta=1$. MCMC was performed for 96 hours in the cluster at Minnesota Supercomputing Instute which allowed for 5,000 generation of burn-in and a minimum of 200,000 generations of MCMC for each of the 6 models. Convergence of the MCMC was tested using R package coda (ref) and software Tracer (ref) to assess convergence and mixing (see supplementary information).

\noindent \begin{tabular}{@{}llccc@{}} \toprule %\label{datatable}
%  \caption{Binary and three state classifications for 595 taxa with ploidy and/or breeding system data. The number of taxa in the t he sample was maximize by including tips with only ploidy or only breeding system and assigned them as uncertain in the unknown character.}
\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
Type & Total of & BiSSE/HiSSE & BiSSE/HiSSE & MuSSE/MuHiSSE\\ 
 &Taxa &  Polyploidy & Breeding System &  \\ \midrule
Diploid Self Compatible & 152 & 0 &  0 & 0 \\
Diploid Self Incompatible& 97 & 0  & 1 & 2\\
Diploid with unknown breeding system & 219 & 0 & (0,1) & (0,2) \\
Polyploid & 81 & 1& 0 & 1 \\
Unknow ploidy and self compatible& 34 & (0,1)& 0 & (0,1) \\ 
Unknown ploidy and self incompatible & 12 & 0 & 1 & 2 \\ \bottomrule
\end{tabular}

### 2018-09-17 manuscript/methods.tex
small tex matters

@@
 \subsection{Models}
-We first defined a binary state speciation and extinction model (BiSSE, \citet{maddison_2007})  for polyploidy evolution where taxa was classified as diploid (D=0) and polyploid (P=1) \ref{datatable}.
+We first defined a binary state speciation and extinction model (BiSSE, \citet{maddison_2007})  for polyploidy evolution where taxa was classified as diploid (D=0) and polyploid (P=1) \cref{table:datatable}.
 In  a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$) , net diversification rates ($r_D=\lambda_D-\mu_D, r_P=\lambda_P-\mu_P$), and  relative extinction rates ($\nu_D=\frac{\mu_D}{\lambda_D}, \nu_D=\frac{\mu_D}{\lambda_D}$) just as previously explored in \citet{mayrose_2011}.
@@
 
-For breeding system both BiSSE and a HiSSE models were fitted independently by coding data as self-compatible (SC=0) or self-incompatible (SI=1) \ref{datatable}.
+For breeding system both BiSSE and a HiSSE models were fitted independently by coding data as self-compatible (SC=0) or self-incompatible (SI=1) \cref{table:datatable}.
 The BiSSE model fitted to breeding system data was done with the goal of investigating if the effect of self-incompatibility in diversification was similar to the pattern found by  \citet{goldberg_2012}.
@@
 Next, we proposed a multivariate speciation and extinction model (MuSSE, \citet{fitzjohn_2012}) to investigate the link amongst diversification, breeding system, and polyploidy simultaneously.
-The MuSSE model is defined using three states self-compatible diploids (SC-D=0), polyploids that are always self-compatible (ref for this??) (P=1), and self-incompatible diploids (SI-D=2, see table \ref{datatable}) and contains ten parameters, six defining diversification ($\lambda_{SD}, \mu_{SD},\lambda_{P},\mu_{P}, \lambda_{ID},\mu_{ID}$) and the other four key cladogenetic changes that are: polyploidization of self-compatible diploids $\rho_{SD}$, diploidization $\delta$, polyploidization of self-incompatible diploids $\rho_{ID}$, and self-incompatible to self-compatible rate $q_{IC}$.\newline
+The MuSSE model is defined using three states self-compatible diploids (SC-D=0), polyploids that are always self-compatible (ref for this??) (P=1), and self-incompatible diploids (SI-D=2, see table \cref{table:datatable}) and contains ten parameters, six defining diversification ($\lambda_{SD}, \mu_{SD},\lambda_{P},\mu_{P}, \lambda_{ID},\mu_{ID}$) and the other four key cladogenetic changes that are: polyploidization of self-compatible diploids $\rho_{SD}$, diploidization $\delta$, polyploidization of self-incompatible diploids $\rho_{ID}$, and self-incompatible to self-compatible rate $q_{IC}$.\newline
 
@@
 
-\noindent \begin{tabular}{@{}llccc@{}} \toprule %\label{datatable}
-%  \caption{Binary and three state classifications for 595 taxa with ploidy and/or breeding system data.
-The number of taxa in the t he sample was maximize by including tips with only ploidy or only breeding system and assigned them as uncertain in the unknown character.}
+\begin{table}
+\begin{tabular}{@{}llccc@{}} \toprule
 \multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
@@
 \end{tabular}
+\caption{Binary and three state classifications for 595 taxa with ploidy and/or breeding system data.
+The number of taxa in the t he sample was maximize by including tips with only ploidy or only breeding system and assigned them as uncertain in the unknown character.}
+\label{table:datatable}
+\end{table}

### 2018-09-17 manuscript/methods.tex
demo of one line per sentence
will be much nicer for git

@@
 \section{Methods}
+
 \subsection{Data}
+
 Chromosome number data was obtained from the Chromosome Count Database (CCDB, \citet{rice_2015}) for all taxa from Solanaceae family.
@@
 Species with more than one ploidy level were assigned smallest most frequent ploidy level recorded.
-Domesticated taxa (i.e.
-\textit{Solanum lycopersicon}) were removed to avoid since they contained multiple ploidy levels due to domestication and not spontaneous polyploidy.
+Domesticated taxa (\ie \textit{Solanum lycopersicon}) were removed to avoid since they contained multiple ploidy levels due to domestication and not spontaneous polyploidy.
 The remainder of taxa was then matched to largest time-calibrated phylogeny of Solanaceae and this process resulted in 595 taxa used in each of the diversification models.

### 2018-09-20 manuscript/methods.tex
notes to self, and linebreaks to aid diff

@@
 
+% E: also need some text about how SI was scored
+
 Binary or three state classification of ploidy and breeding system for the 595 taxa  was done according to Table \autoref{tab:binaryclassifications}
@@
 \subsection{Models}
+
 We first defined a binary state speciation and extinction model (BiSSE, \citet{maddison_2007})  for polyploidy evolution where taxa was classified as diploid (D=0) and polyploid (P=1) \cref{table:datatable}.
 In a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$) , net diversification rates ($r_D=\lambda_D-\mu_D, r_P=\lambda_P-\mu_P$), and  relative extinction rates ($\nu_D=\frac{\mu_D}{\lambda_D}, \nu_D=\frac{\mu_D}{\lambda_D}$) just as previously explored in \citet{mayrose_2011}.
-The cladogenetic changes explored in BiSSE were polyploidization rate are represented by parameter $\rho$ and by diploidization parameter rate $\delta$.\newline
+The cladogenetic changes explored in BiSSE were polyploidization rate are represented by parameter $\rho$ and by diploidization parameter rate $\delta$.
+% E: cladogenetic?
+
+% E: above and below, drop numerical state names ... even in the table
 
@@
 The cladogenetic changes assumed by the fitted model are polyploidization rate $\rho$, diploidization rate $\rho$, and we assumed that the changes between every hidden state are symmetrical with rate $\alpha$.
-\newline
 
@@
 However, the HiSSE model on breeding system parses out if the effect of breeding system in diversification is significant  by minimizing type I error as mentioned above.
-For these models we assumed that self-compatibility is irreversible as discussed by \citet{igic_2013}, so we defined the transition rate $q_{IC}$ as the parameter for changes from self-incompatible to self-compatible state.\newline
+For these models we assumed that self-compatibility is irreversible as discussed by \citet{igic_2013}, so we defined the transition rate $q_{IC}$ as the parameter for changes from self-incompatible to self-compatible state.
+% E: explain irreversibility assumption; fix ancestry instead?
 
 Next, we proposed a multivariate speciation and extinction model (MuSSE, \citet{fitzjohn_2012}) to investigate the link amongst diversification, breeding system, and polyploidy simultaneously.
-The MuSSE model is defined using three states self-compatible diploids (SC-D=0), polyploids that are always self-compatible (ref for this??) (P=1), and self-incompatible diploids (SI-D=2, see table \cref{table:datatable}) and contains ten parameters, six defining diversification ($\lambda_{SD}, \mu_{SD},\lambda_{P},\mu_{P}, \lambda_{ID},\mu_{ID}$) and the other four key cladogenetic changes that are: polyploidization of self-compatible diploids $\rho_{SD}$, diploidization $\delta$, polyploidization of self-incompatible diploids $\rho_{ID}$, and self-incompatible to self-compatible rate $q_{IC}$.\newline
+The MuSSE model is defined using three states self-compatible diploids (SC-D=0), polyploids that are always self-compatible (ref for this??) (P=1), and self-incompatible diploids (SI-D=2, see table \cref{table:datatable}) and contains ten parameters, six defining diversification ($\lambda_{SD}, \mu_{SD},\lambda_{P},\mu_{P}, \lambda_{ID},\mu_{ID}$) and the other four key cladogenetic changes that are: polyploidization of self-compatible diploids $\rho_{SD}$, diploidization $\delta$, polyploidization of self-incompatible diploids $\rho_{ID}$, and self-incompatible to self-compatible rate $q_{IC}$.
+% E: for SC, use C instead of S
+% E: lack of SI-P should maybe be explained in Intro?
 
@@
 The model we proposed is a multivariate and hidden states speciation and extinction stochastic process (MuHiSSE) that allowed us to account simultaneously for the diversification rates linked to breeding system and polyploidy but also the presence of some more heterogeneity in the process.
-This model is analogous to the GeoHiSSE model proposed by Caetano et al.
-(2018) in a biogeographical context.
+This model is analogous to the GeoHiSSE model proposed by Caetano et al. (2018) in a biogeographical context.
 The full MuHiSSE model has 26 parameters, however, our goal was to look for diversification rate differences so we fitted a simplified version of 17 parameters by fixing the rates amongst hidden states to be equal with parameter $\alpha$ and the transition rates amongst breeding system and polyploidy as defined in the MuSSE model  ($\rho_{SD}$,  $\delta$,  $\rho_{ID}$,  $q_{IC}$) despite the hidden state.
-Using MuHiSSE we estimated twelve speciation and extinction rates ($\lambda_{SD_A}, \mu_{SD_A},\lambda_{P_A},\mu_{P_A}, \lambda_{ID_A},\mu_{ID_B},\lambda_{SD_B}, \mu_{SD_B},\lambda_{P_B},\mu_{P_B}, \lambda_{ID_B},\mu_{ID_B}$), and the net diversification and relative extinction rates associated with them.\newline
+Using MuHiSSE we estimated twelve speciation and extinction rates ($\lambda_{SD_A}, \mu_{SD_A},\lambda_{P_A},\mu_{P_A}, \lambda_{ID_A},\mu_{ID_B},\lambda_{SD_B}, \mu_{SD_B},\lambda_{P_B},\mu_{P_B}, \lambda_{ID_B},\mu_{ID_B}$), and the net diversification and relative extinction rates associated with them.
+% E: caetano ref; and huang
 
@@
 Convergence of the MCMC was tested using R package coda (ref) and software Tracer (ref) to assess convergence and mixing (see supplementary information).
+% E: solsource ref
 
@@
 The number of taxa in the t he sample was maximize by including tips with only ploidy or only breeding system and assigned them as uncertain in the unknown character.}
+% E: but P are not SI
 \label{table:datatable}

### 2018-09-20 manuscript/methods.tex
edited the Data section
added some explanation of SI/SC trait, and lots of refs

@@
 
-Chromosome number data was obtained from the Chromosome Count Database (CCDB, \citet{rice_2015}) for all taxa from Solanaceae family.
-The records obtained were first cleaned using CCDBcurator R package \citep{zenilferguson_2017}, to those we aggregated independent data of chromosome numbers and breeding systems recorded and curated by Emma Goldberg and Boris Igi\'c  (citations for original sources can be found in the supplementary information).
-In total we obtained 14,000 records of chromosome number and breeding system than were curated semi automatically by CCDBcurator and by hand to score the chromosome number and the breeding system with the most reconciliation.
-Furthermore, for the majority of taxa ploidy was assigned according to information from original sources and Kew Royal Botanic Gardens C-value DNA (bennett citation here dataset and the original sources agreggated by Goldberg and Igi\'c.
-For taxa without ploidy information but with information about chromosome number we assigned ploidy based on multiplicity of chromosomes within the genus.
-For example, \textit{Solanum betaceum} did not include information about ploidy level but its number of chromosomes is 24, since $x=12$ is the base chromosome number of the \textit{Solanum} genus we assigned  \textit{Solanum betaceum} as diploid.
-Species with more than one ploidy level were assigned smallest most frequent ploidy level recorded.
-Domesticated taxa (\ie \textit{Solanum lycopersicon}) were removed to avoid since they contained multiple ploidy levels due to domestication and not spontaneous polyploidy.
-The remainder of taxa was then matched to largest time-calibrated phylogeny of Solanaceae and this process resulted in 595 taxa used in each of the diversification models.
+Chromosome number data were obtained for all Solanaceae taxa in the the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.\ 14,000 records were cleaned semi-automatically using the CCDBcurator R package \citep{zenilferguson_2017}.
+This large dataset includes the compilation of Solanaceae ploidy states from \citep{robertson_2011}.
+Species were coded as either diploid (D) or polyploid (P).
+For the majority of species, ploidy was assigned according to information from the original publications and the Kew Royal Botanic Gardens C-value DNA resource \citep{bennett_2005}.
+For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus.
+For example, \textit{Solanum betaceum} did not include information about ploidy level but it has 24 chromosomes, so because $x=12$ is the base chromosome number of the \textit{Solanum} genus \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid.
+Species with more than one ploidy level were assigned the smallest most frequent ploidy level recorded.
+% E: Does this apply to the diploid-polyploid levels?
+If so, we need to explain how it is also reconciled against the SI/SC value.
+Or is this sentence just about multiple levels of polyploidy, which wouldn't need to be mentioned here?
 
-% E: also need some text about how SI was scored
+Breeding system was scored as self-incompatible (SI) or self-compatible (SC) based on results hand-curated from the literature (as in \citealt{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}).
+Most species could unambiguously be coded as either SI or SC \citep{raduski_2012}.
+Following previous work, we coded as SI any species with functional SI systems, even if SC or dioecy was also reported.
+Dioecious species without functional SI were coded as SC.
 
-Binary or three state classification of ploidy and breeding system for the 595 taxa  was done according to Table \autoref{tab:binaryclassifications}
+To those existing data sets, we added some additional records for chromosome number and breeding system.
+The Supplementary Information contains citations for the numerous original sources for all of the data. % todo: add number of refs when we have it
+Synonymy followed Solanaceae Source \citep{solsource}.
+Hybrids and cultivars were excluded because ploidy and breeding system are likely to be altered by domestication.
+As in \citet{robertson_2011}, we examined closely the few species for which the merged ploidy and breeding system data indicated the presence of SI polyploids.
+Although SI populations frequently contain some SC individuals and diploid populations frequently contain some polyploid individuals, in no case did we find individuals that were both SI and polyploid.
+Because of this empirical observation and the functional explanation for whole genome duplication disabling gametophytic self-incompatibility \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: SI-D, SC-D, and SC-P.
+
+% E: We should probably include some more details about how many polymorphic species and how the particularly tricky ones were dealt with.
+
+Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 595 species with ploidy and/or breeding system information on the tree.
+Binary or three-state classification of ploidy and breeding system for the 595 taxa is summarized in \cref{table:stateclassifications}. % E: I think you're still updating the state coding?
+We retained all of these species in each of the analyses below, because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
+% E: Though someone might ask why tips with no state data were pruned.  Analyses took long enough on even this-size tree.
 
@@
 
-We first defined a binary state speciation and extinction model (BiSSE, \citet{maddison_2007})  for polyploidy evolution where taxa was classified as diploid (D=0) and polyploid (P=1) \cref{table:datatable}.
+We first defined a binary state speciation and extinction model (BiSSE, \citet{maddison_2007})  for polyploidy evolution where taxa was classified as diploid (D=0) and polyploid (P=1) \cref{table:stateclassifications}.
 In a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$) , net diversification rates ($r_D=\lambda_D-\mu_D, r_P=\lambda_P-\mu_P$), and  relative extinction rates ($\nu_D=\frac{\mu_D}{\lambda_D}, \nu_D=\frac{\mu_D}{\lambda_D}$) just as previously explored in \citet{mayrose_2011}.
@@
 
-For breeding system both BiSSE and a HiSSE models were fitted independently by coding data as self-compatible (SC=0) or self-incompatible (SI=1) \cref{table:datatable}.
+For breeding system both BiSSE and a HiSSE models were fitted independently by coding data as self-compatible (SC=0) or self-incompatible (SI=1) \cref{table:stateclassifications}.
 The BiSSE model fitted to breeding system data was done with the goal of investigating if the effect of self-incompatibility in diversification was similar to the pattern found by \citet{goldberg_2012}.
@@
 Next, we proposed a multivariate speciation and extinction model (MuSSE, \citet{fitzjohn_2012}) to investigate the link amongst diversification, breeding system, and polyploidy simultaneously.
-The MuSSE model is defined using three states self-compatible diploids (SC-D=0), polyploids that are always self-compatible (ref for this??) (P=1), and self-incompatible diploids (SI-D=2, see table \cref{table:datatable}) and contains ten parameters, six defining diversification ($\lambda_{SD}, \mu_{SD},\lambda_{P},\mu_{P}, \lambda_{ID},\mu_{ID}$) and the other four key cladogenetic changes that are: polyploidization of self-compatible diploids $\rho_{SD}$, diploidization $\delta$, polyploidization of self-incompatible diploids $\rho_{ID}$, and self-incompatible to self-compatible rate $q_{IC}$.
+The MuSSE model is defined using three states self-compatible diploids (SC-D=0), polyploids that are always self-compatible (ref for this??) (P=1), and self-incompatible diploids (SI-D=2, see table \cref{table:stateclassifications}) and contains ten parameters, six defining diversification ($\lambda_{SD}, \mu_{SD},\lambda_{P},\mu_{P}, \lambda_{ID},\mu_{ID}$) and the other four key cladogenetic changes that are: polyploidization of self-compatible diploids $\rho_{SD}$, diploidization $\delta$, polyploidization of self-incompatible diploids $\rho_{ID}$, and self-incompatible to self-compatible rate $q_{IC}$.
 % E: for SC, use C instead of S
@@
 All the models were performed using RevBayes \citep{hoehna_2016} software that performs Bayesian inference via MCMC using Metropolis-Hastings algorithm.
-A correction for sampling bias was done in all models by assuming that Solanaceae family has approximately 3,000 species ($s=595/3000$) as the Solanaceae Source project indicates \citep{source_2011}.
+A correction for sampling bias was done in all models by assuming that Solanaceae family has approximately 3,000 species ($s=595/3000$) as the Solanaceae Source project indicates \citep{solsource}.
 In all models, speciation and extinction parameters used log-normal prior distributions that with mean the expected net diversification rate $ ( \frac{(number of taxa)/2}{root age})$ and standard deviation $0.5$.
@@
 Convergence of the MCMC was tested using R package coda (ref) and software Tracer (ref) to assess convergence and mixing (see supplementary information).
-% E: solsource ref
 
@@
 % E: but P are not SI
-\label{table:datatable}
+\label{table:stateclassifications}
 \end{table}

### 2018-09-21 manuscript/methods.tex
thinking about model names
edited model descriptions

@@
 
-We first defined a binary state speciation and extinction model (BiSSE, \citet{maddison_2007})  for polyploidy evolution where taxa was classified as diploid (D=0) and polyploid (P=1) \cref{table:stateclassifications}.
-In a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$) , net diversification rates ($r_D=\lambda_D-\mu_D, r_P=\lambda_P-\mu_P$), and  relative extinction rates ($\nu_D=\frac{\mu_D}{\lambda_D}, \nu_D=\frac{\mu_D}{\lambda_D}$) just as previously explored in \citet{mayrose_2011}.
-The cladogenetic changes explored in BiSSE were polyploidization rate are represented by parameter $\rho$ and by diploidization parameter rate $\delta$.
-% E: cladogenetic?
+% E: I think it would help to start with an "overall strategy" paragraph, maybe something like this.
+To investigate the associations of these traits with lineage diversification, we fit six models from the BiSSE \citep{maddison_2007} family.
+The simplest models consider ploidy only or breeding system only, as previous work has done. % add refs
+We also consider the two traits simultaneously, to investigate whether the effect of one might override the effect of the other.
+Additionally, to allow for the possibility that something other than these traits might affect diversification, we add models that include a hidden trait. % hisse ref
+% E: And say something about including diploidization?  Will the motivation be in the Intro?
+% E: And explain about the SC -> SI rate being fixed to 0, so we don't bury the explanation below? vs fixing ancestry?
 
-% E: above and below, drop numerical state names ... even in the table
+We first considered only the ploidy trait, scoring each species in either the diploid state (D) or the polyploid state (P).
+To allow for the influence of this binary trait on lineage diversification, we used the BiSSE model \citep{maddison_2007}.
+% \cref{table:stateclassifications}
+In a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$), net diversification rates ($r_D=\lambda_D-\mu_D,\ r_P=\lambda_P-\mu_P$), and relative extinction rates ($\nu_D = \mu_D / \lambda_D,\ \nu_D = \mu_D / \lambda_D$) for each state.
+% as previously explored by \citet{mayrose_2011} % E: actually, they used bisse-ness/classe to allow for cladogenetic change
+We also simultaneously estimated the polyploidization rate (the transition rate from D to P, $\rho$), and we allowed for diploidization (the transition rate from P to D, $\delta$).
+We call this the ``ploidy only'' model. % E: not sure that we need to be this explicit, but try for now
 
-As a second step, we fitted a hidden state speciation and extinction model (HiSSE, \citet{beaulieu_2016}) to evaluate whether the differences in diversification rates were found due to a hidden trait associated to polyploidy.
-As discussed by \citet{beaulieu_2016} BiSSE-like models suffer from a large type I error because they fail to account as part of the null hypothesis heterogeneous diversification rate changes  that do not depend on the trait of interest.
-By including a hidden trait in a state speciation and extinction  model,  HiSSE-like models address the heterogeneous background diversification changes while also parsing out the possible signal of diversification due to the trait of interest.
-Therefore, HiSSE model for polyploidy linked to diversification has four states: diploid and polyploid subdivided by a binary hidden trait with states A and B for which we estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\lambda_{D_B}, \lambda_{P_A},\lambda_{P_B}$), extinction rates ($\mu_{D_A},\mu_{D_B}, \mu_{P_A},\mu_{P_B}$),  net diversification rates ($r_{D_A},r_{D_B},r_{P_A},r_{P_B}$), and relative extinction($\nu_{D_A},\nu_{D_B}, \nu_{PA},\nu_{P_B}$).
-The cladogenetic changes assumed by the fitted model are polyploidization rate $\rho$, diploidization rate $\rho$, and we assumed that the changes between every hidden state are symmetrical with rate $\alpha$.
+Second, we considered ploidy and also a unobserved, hidden trait.
+As discussed by \citet{beaulieu_2016}, the previous model might attribute diversification differences to the ploidy trait merely because there are no other explanatory factors in the model.
+% As a second step, we fitted a hidden state speciation and extinction model (HiSSE, \citet{beaulieu_2016}) to evaluate whether the differences in diversification rates were found due to a hidden trait associated to polyploidy.
+% As discussed by \citet{beaulieu_2016} BiSSE-like models suffer from a large type I error because they fail to account as part of the null hypothesis heterogeneous diversification rate changes  that do not depend on the trait of interest.
+By including a hidden trait in a state speciation and extinction model, HiSSE-like models allow for heterogeneous background diversification while also parsing out the possible signal of diversification due to the trait of interest.
+Our second model thus has four states, with D and P subdivided by an unobserved binary trait with states A and B.
+We therefore estimated the posterior probability distributions of four speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\  \lambda_{P_A},\ \lambda_{P_B}$), four extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$), and the corresponding four net diversification rates and relative extinction rates.
+We again include transition due to polyploidization (rate $\rho$) and diploidization (rate $\delta$), and we assume that the changes between hidden states are symmetrical with rate $\alpha$.
+We call this the ``ploidy and hidden trait'' model.
 
-For breeding system both BiSSE and a HiSSE models were fitted independently by coding data as self-compatible (SC=0) or self-incompatible (SI=1) \cref{table:stateclassifications}.
-The BiSSE model fitted to breeding system data was done with the goal of investigating if the effect of self-incompatibility in diversification was similar to the pattern found by \citet{goldberg_2012}.
-However, the HiSSE model on breeding system parses out if the effect of breeding system in diversification is significant  by minimizing type I error as mentioned above.
-For these models we assumed that self-compatibility is irreversible as discussed by \citet{igic_2013}, so we defined the transition rate $q_{IC}$ as the parameter for changes from self-incompatible to self-compatible state.
-% E: explain irreversibility assumption; fix ancestry instead?
+% E: I added the names of the rate parameters to these next models
 
-Next, we proposed a multivariate speciation and extinction model (MuSSE, \citet{fitzjohn_2012}) to investigate the link amongst diversification, breeding system, and polyploidy simultaneously.
-The MuSSE model is defined using three states self-compatible diploids (SC-D=0), polyploids that are always self-compatible (ref for this??) (P=1), and self-incompatible diploids (SI-D=2, see table \cref{table:stateclassifications}) and contains ten parameters, six defining diversification ($\lambda_{SD}, \mu_{SD},\lambda_{P},\mu_{P}, \lambda_{ID},\mu_{ID}$) and the other four key cladogenetic changes that are: polyploidization of self-compatible diploids $\rho_{SD}$, diploidization $\delta$, polyploidization of self-incompatible diploids $\rho_{ID}$, and self-incompatible to self-compatible rate $q_{IC}$.
-% E: for SC, use C instead of S
-% E: lack of SI-P should maybe be explained in Intro?
+In our third, ``breeding system only'' model, each species is scored as either self-incompatible (SI) or self-compatible (SC).
+In this BiSSE analysis, state-dependent speciation rates ($\lambda_I$, $\lambda_C$) and extinction rates ($\mu_I$, $\mu_C$) are estimated, with the corresponding net diversification and relative extinction rates.
+The rate for transitions from SI to SC is $q_{IC}$, and regain of SI is prohibited, as explained above.
+This is the same as the analysis of \citet{goldberg_2010}.
 
-Since the null hypothesis of the MuSSE model is that the diversification is equal and constant for all three states defined, it is possible that MuSSE can also suffer from large type I errors.
-In order to account for heterogeneity of the diversification rates and parse the signal in diversification coming from breeding system and polyploidy we extended MuSSE model to account for a hidden state, as \citet{beaulieu_2016} did in the bivariate case.
-The model we proposed is a multivariate and hidden states speciation and extinction stochastic process (MuHiSSE) that allowed us to account simultaneously for the diversification rates linked to breeding system and polyploidy but also the presence of some more heterogeneity in the process.
-This model is analogous to the GeoHiSSE model proposed by Caetano et al. (2018) in a biogeographical context.
-The full MuHiSSE model has 26 parameters, however, our goal was to look for diversification rate differences so we fitted a simplified version of 17 parameters by fixing the rates amongst hidden states to be equal with parameter $\alpha$ and the transition rates amongst breeding system and polyploidy as defined in the MuSSE model  ($\rho_{SD}$,  $\delta$,  $\rho_{ID}$,  $q_{IC}$) despite the hidden state.
-Using MuHiSSE we estimated twelve speciation and extinction rates ($\lambda_{SD_A}, \mu_{SD_A},\lambda_{P_A},\mu_{P_A}, \lambda_{ID_A},\mu_{ID_B},\lambda_{SD_B}, \mu_{SD_B},\lambda_{P_B},\mu_{P_B}, \lambda_{ID_B},\mu_{ID_B}$), and the net diversification and relative extinction rates associated with them.
-% E: caetano ref; and huang
+The fourth model considers both breeding system and a hidden trait, again using a HiSSE strategy to test whether diversification shifts unrelated to the focal trait might be attributed to it by the previous BiSSE analysis.
+In this ``breeding system and hidden trait'' model, there are four speciation rates ($\lambda_{I_A},\ \lambda_{I_B},\  \lambda_{C_A},\ \lambda_{C_B}$), four extinction rates ($\mu_{I_A},\ \mu_{I_B},\ \mu_{C_A},\ \mu_{C_B}$), and correspondingly four rates of net diversification and relative extinction.
+Transition rates are again $q_{IC}$ for loss of SI and $\alpha$ for changes in the hidden state.
+
+% E: I adjusted the rate names a little.  Does this seem okay?
+Our fifth model investigates the link among diversification, breeding system, and ploidy simultaneously.
+There are three possible states: self-incompatible diploids (ID), self-compatible diploids (CD), and self-compatible polyploids (CP); self-incompatible polyploids are not present in this family, as explained earlier.
+This ``ploidy and breeding system'' MuSSE model \citep{fitzjohn_2012} has six rates for speciation and extinction ($\lambda_{ID},\ \mu_{ID},\ \lambda_{CD},\ \mu_{CD},\ \lambda_{CP},\ \mu_{CP}$).
+It also has four rates for transitions:
+polyploidization of SI diploids (transition from ID to CP at rate $\rho_I$),
+polyploidization of SC diploids (transition from CD to CP at rate $\rho_C$),
+diploidization, which does not restore SI (transition from CP to CD at rate $\delta_C$),
+and loss of SI without a change in ploidy (transition from ID to CD at rate $q_{IC}$).
+
+Finally, the sixth model considers a hidden trait along with ploidy and breeding system.
+This again provides a means to account simultaneously for diversification rates linked to ploidy and breeding system while also allowing for the presence of heterogeneity from unknown sources.
+(A similar approach was used for a different variant of BiSSE by \citealt{huang_2018,caetano_2018}.)
+The full model would have 26 parameters, but our goal is to look for diversification rate differences so we fit a simplified version with 17 parameters by fixing the transition rates among hidden states to be equal with parameter $\alpha$, and the transition rates between breeding system and ploidy states as defined in the previous model ($\rho_I$, $\rho_C$, $\delta_C$,  $q_{IC}$) independent of the hidden state.
+We estimated twelve speciation and extinction rates ($\lambda_{ID_A},\ \mu_{ID_A},\ \lambda_{CD_A},\ \mu_{CD_A},\ \lambda_{CP_A},\ \mu_{CP_A},\ \lambda_{ID_B},\ \mu_{ID_B},\ \lambda_{CD_B},\ \mu_{CD_B}, \lambda_{CP_B},\ \mu_{CP_B}$), and the net diversification and relative extinction rates associated with them.
 

### 2019-02-09 manuscript/abstract.tex
Updated methods and bayes factors table

\section{Abstract}

The effect of polyploidy in diversification remains a contentious issue. On the one hand,  recent studies  that found that polyploids have slower speciation rates and higher extinction rates than diploids left scientist wondering if polyploidy is truly an evolutionary dead-end. On the other hand, botanist have found strong molecular support of multiple polyploidy events at the root of highly diverse clades which challenges the evolutionary dead-end conclusions reached by modeling approaches. We re-investigate the role of polyploidy in speciation and extinction from a new modeling perspective considering that patterns found in diversification models can be misleading and incorrectly attributed to polyploidy when other observed and unobserved plant traits are responsible of shaping diversification.  Using  statistically robust comparative phylogenetic approaches, we show that it is possible to  detect whether  the contribution of polyploidy to speciation and extinction is significant  under the presence of  other potential traits also affect diversification. We use the phylogeny, polyploidy, and breeding system data of 595 Solanaceae species to understand the contribution of polyploidy to diversification. We ask if Solanaceae polyploids are evolutionary dead-ends, and whether breeding system or some other unobserved traits are responsible of the patterns of diversification observed in the phylogeny.

### 2019-02-09 manuscript/introduction.tex
Updated methods and bayes factors table

@@
 \section{Introduction}
-Linking changes in diversification to polyploidy and breeding system has been one of the key botany questions for the past n number of years.
-The prevalence of polyploidy, and the multiple evolutionary changes of breeding systems in important angiosperm clades, leads to the inevitable question of the evolutionary changes.
+-Studying diversification linked to trait evolution\newline
+Here I need help (perhaps will be the last line I write)\newline
+In these context, botanist have been asking how polyploidy shapes patterns of diversification.
+The prevalence of polyploidy and its detection across multiple and highly diverse clades of angiosperms inevitably lead to the hypothesis of the importance of polyploidy in the speciation and extinction patterns observed across the flowering plant phylogeny.
+At the same time, a similar question has been asked in the breeding system world, where self-compatibility has evolved multiple times in flowering plants.
+However, the influence in diversification using both polyploidy and self-compatibility information has not been studied simultaneously.
 
-- Diversification and polyploidy\newline
-An important debate regarding the diversification of angiosperms has been ongoing since the publication of \citet{mayrose_2011}.
+- Why polyploidy and self-compatibility in particular makes for an  interesting study in the context of diversification\newline
+In the polyploidy world, an important debate regarding the diversification of angiosperms has been ongoing since the publication of \citet{mayrose_2011}.
 The authors discovered that using the latest diversification model linked to two states diploid and polyploid, the net diversification of polyploids was much slower than the net diversification rate than polyploids.
@@
 A year later, diversification models were re-tested and corrected, and still found the same pattern \citep{mayrose_2015}, to the disappointment of plenty of botanists there was no denying of the weak trend of diversification that polyploids left behind.
-\newline
+The most recent study by \citet{landis_2018} that used not only the presence of polyploidy in the tips but also the number of whole genome duplications in a taxon lineage found that....
+\newline  Meanwhile, studies focusing on diversification patterns and  breeding system have consistently found that self-incompatible plants often have higher net diversification rates compare to their self-compatible counterparts (Emma and Boris' papers here, what about papers that are not solanaceae?)
+
+- Why other traits need to be considered as well\newline
+
+- What other models and studies have done in the past\newline
+
+- What is lacking from past approaches? \newline
 There are two key questions that at the time of the polyploidy debate were difficult to ask.
@@
 
+-What is our proposal to tackle this problem?\newline
+
+- How this paper is structured \newline
+
 The second question (talk here about breeding system and polyploidy, and how this could be different across clades and this is one of the reasons why we focused on Solanaceae).
@@
 
--What we did\\
-We investigate the relative importance of both b
-
-Debate of Mayrose
+=

### 2019-02-09 manuscript/methods.tex
Updated methods and bayes factors table

--- 2018-09-20 manuscript/methods.tex
@@
 
-Chromosome number data were obtained for all Solanaceae taxa in the the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.\ 14,000 records were cleaned semi-automatically using the CCDBcurator R package \citep{zenilferguson_2017}.
+Chromosome number data were obtained for all Solanaceae taxa in the the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.\ 14,000 records were cleaned semi-automatically using the CCDBcurator R package \citep{rivero_2019}.
 This large dataset includes the compilation of Solanaceae ploidy states from \citep{robertson_2011}.
@@
 For example, \textit{Solanum betaceum} did not include information about ploidy level but it has 24 chromosomes, so because $x=12$ is the base chromosome number of the \textit{Solanum} genus \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid.
-Species with more than one ploidy level were assigned the smallest most frequent ploidy level recorded.
-% E: Does this apply to the diploid-polyploid levels?
+Species with more than one ploidy level were assigned the smallest and most frequent ploidy level recorded.
+We also cross-verified using the breeding system information and found that there were no self-incompatible species possibly listed as polyploid, or listed as polyploid and diploid which increase our confidence in the assignations% E: Does this apply to the diploid-polyploid levels?
+R- Yes there are no self-incompatible polyploids at all.
 If so, we need to explain how it is also reconciled against the SI/SC value.
 Or is this sentence just about multiple levels of polyploidy, which wouldn't need to be mentioned here?
-
-Breeding system was scored as self-incompatible (SI) or self-compatible (SC) based on results hand-curated from the literature (as in \citealt{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}).
+R:  Solved
+Breeding system was scored as self-incompatible (I) or self-compatible (C) based on results hand-curated from the literature (as in \citealt{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}).
 Most species could unambiguously be coded as either SI or SC \citep{raduski_2012}.
-Following previous work, we coded as SI any species with functional SI systems, even if SC or dioecy was also reported.
-Dioecious species without functional SI were coded as SC.
+Following previous work, we coded as I any species with functional I systems, even if C or dioecy was also reported.
+Dioecious species without functional I were coded as C.
 
@@
 Hybrids and cultivars were excluded because ploidy and breeding system are likely to be altered by domestication.
-As in \citet{robertson_2011}, we examined closely the few species for which the merged ploidy and breeding system data indicated the presence of SI polyploids.
-Although SI populations frequently contain some SC individuals and diploid populations frequently contain some polyploid individuals, in no case did we find individuals that were both SI and polyploid.
-Because of this empirical observation and the functional explanation for whole genome duplication disabling gametophytic self-incompatibility \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: SI-D, SC-D, and SC-P.
+As in \citet{robertson_2011}, we examined closely the few species for which the merged ploidy and breeding system data indicated the presence of I polyploids.
+Although SI populations frequently contain some C individuals and diploid populations frequently contain some polyploid individuals, in no case did we find individuals that were both SI and polyploid.
+Because of this empirical observation and the functional explanation for whole genome duplication disabling gametophytic self-incompatibility \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploid (ID), self-compatible diploid (CD), and polyploid (P) which are always self-compatible.
 
@@
 Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 595 species with ploidy and/or breeding system information on the tree.
-Binary or three-state classification of ploidy and breeding system for the 595 taxa is summarized in \cref{table:stateclassifications}. % E: I think you're still updating the state coding?
+Binary or three-state classification of ploidy and breeding system for the 595 taxa is summarized in \cref{table:stateclassifications}.
+% E: I think you're still updating the state coding?
+R=no since models are not comparable there is no need.
 We retained all of these species in each of the analyses below, because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
-% E: Though someone might ask why tips with no state data were pruned.  Analyses took long enough on even this-size tree.
+A total of 405 taxa without any information about breeding system nor polyploidy were pruned.
+Tips without trait data are non informative for  diversification parameters linked to trait change and can significantly slow-down the convergence of inferences in trees with already hundreds of tips.% E: Though someone might ask why tips with no state data were pruned.
+Analyses took long enough on even this-size tree.
+R: I think this point needs to be further investigated in another paper, so I just sort-of justify our choice
 
-\subsection{Models}
+\subsection{Models for Diversification and Polyploidy}
+
+% E: There are so many models, I think we need to try hard to help the reader keep them straight.
+% I would avoid naming any states with numbers, except maybe within the stateclassification table and its caption.
+% I also still find bisse/hisse/muhisse terminology confusing, I think because it both describes a general class of model and applies to a specific analysis here, and because bisse and hisse are run separately on ploidy breeding system.
+% How about a model naming scheme like this?
+%   D/P polyploidy model
+%   D/P-A/B polyploidy hidden state model
+%   I/C breeding system model
+%   I/C-A/B breeding system and hidden state model
+%   ID/CD/CP ploidy and breeding system model
+%   ID/CD/CP-A/B ploidy, breeding system and hidden state model
+% Or something else more concise and descriptive?
 
 We first defined a binary state speciation and extinction model (BiSSE, \citet{maddison_2007})  for polyploidy evolution where taxa was classified as diploid (D=0) and polyploid (P=1) \cref{table:stateclassifications}.
-In a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$) , net diversification rates ($r_D=\lambda_D-\mu_D, r_P=\lambda_P-\mu_P$), and  relative extinction rates ($\nu_D=\frac{\mu_D}{\lambda_D}, \nu_D=\frac{\mu_D}{\lambda_D}$) just as previously explored in \citet{mayrose_2011}.
-The cladogenetic changes explored in BiSSE were polyploidization rate are represented by parameter $\rho$ and by diploidization parameter rate $\delta$.
-% E: cladogenetic?
-
-% E: above and below, drop numerical state names ... even in the table
+We refer to this model as the D/P polyploidy model. 
+In a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$) , net diversification rates ($r_D=\lambda_D-\mu_D, r_P=\lambda_P-\mu_P$), and  relative extinction rates ($\nu_D=\frac{\mu_D}{\lambda_D}, \nu_D=\frac{\mu_D}{\lambda_D}$) just as previously explored in \citet{mayrose_2011} and \citet{mayrose_2015}.
+However, our analyses differ from previous models since we consider not only a polyploidization rate $\rho$ and but also a diploidization parameter rate $\delta$.
 
 As a second step, we fitted a hidden state speciation and extinction model (HiSSE, \citet{beaulieu_2016}) to evaluate whether the differences in diversification rates were found due to a hidden trait associated to polyploidy.
-As discussed by \citet{beaulieu_2016} BiSSE-like models suffer from a large type I error because they fail to account as part of the null hypothesis heterogeneous diversification rate changes  that do not depend on the trait of interest.
-By including a hidden trait in a state speciation and extinction  model,  HiSSE-like models address the heterogeneous background diversification changes while also parsing out the possible signal of diversification due to the trait of interest.
-Therefore, HiSSE model for polyploidy linked to diversification has four states: diploid and polyploid subdivided by a binary hidden trait with states A and B for which we estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\lambda_{D_B}, \lambda_{P_A},\lambda_{P_B}$), extinction rates ($\mu_{D_A},\mu_{D_B}, \mu_{P_A},\mu_{P_B}$),  net diversification rates ($r_{D_A},r_{D_B},r_{P_A},r_{P_B}$), and relative extinction($\nu_{D_A},\nu_{D_B}, \nu_{PA},\nu_{P_B}$).
-The cladogenetic changes assumed by the fitted model are polyploidization rate $\rho$, diploidization rate $\rho$, and we assumed that the changes between every hidden state are symmetrical with rate $\alpha$.
+This model is referred to as the D/P-A/B polyploidy and hidden state model.
+As discussed by \citet{beaulieu_2016} BiSSE-like models suffer from a large type I error since they fail to account as part of the null hypothesis heterogeneous diversification rate changes  that do not directly depend on the trait of interest.
+By expanding the D/P model by using a hidden trait with states A and B,  the polyploidy and hidden state model addressed the signal of diversification due to polyploidy while parsing out the heterogeneity of diversification rates due to unobserved traits.
+The D/P-A/B polyploidy and hidden trait is composed by four states: diploid and polyploid subdivided by a binary hidden trait with states A and B for which we estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\lambda_{D_B}, \lambda_{P_A},\lambda_{P_B}$), extinction rates ($\mu_{D_A},\mu_{D_B}, \mu_{P_A},\mu_{P_B}$),  net diversification rates ($r_{D_A},r_{D_B},r_{P_A},r_{P_B}$), and relative extinction rates($\nu_{D_A},\nu_{D_B}, \nu_{PA},\nu_{P_B}$).
+In this model polyploidization rate $\rho$, diploidization rate $\delta$ are also considered, while changes between every hidden state are symmetrical with the same rate $\alpha$.
 
-For breeding system both BiSSE and a HiSSE models were fitted independently by coding data as self-compatible (SC=0) or self-incompatible (SI=1) \cref{table:stateclassifications}.
-The BiSSE model fitted to breeding system data was done with the goal of investigating if the effect of self-incompatibility in diversification was similar to the pattern found by \citet{goldberg_2012}.
-However, the HiSSE model on breeding system parses out if the effect of breeding system in diversification is significant  by minimizing type I error as mentioned above.
-For these models we assumed that self-compatibility is irreversible as discussed by \citet{igic_2013}, so we defined the transition rate $q_{IC}$ as the parameter for changes from self-incompatible to self-compatible state.
+\subsection{Models for Diversification and Breeding System}
+In order to analyze the effects of breeding system in the diversification process, two models the I/C (I= self-incompatible, C= self-compatible) and the I/C-A/B were performed.
+The I/C breeding system model (BiSSE) was done with the goal of investigating if the effect of self-incompatibility in diversification similar to the analysis performed by \citet{goldberg_2012}.
+The I/C-A/B  breeding system with hidden trait model parses out the effect of breeding system in diversification by incorporating the possibility of heterogenous changes in the diversification rates that can be attributed to a hidden state.
+A hidden state model has not previously been fitted to Solanaceae data.
+For these two models, we assumed that self-compatibility is irreversible as discussed by \citet{igic_2013}, so we defined the transition rate $q_{IC}$ as the parameter for changes from self-incompatible to self-compatible state.
+This is an important assumption because ancestral state reconstruction in the models might differ when polyploidy and breeding system are analyzed as independent  in the Solanaceae tree.
+This irreversibility assumption prevents self-compatible state from evolving into self-incompatible state and as a result, the ancestral state of all taxa in this model has probability one of being self-incompatible.
 % E: explain irreversibility assumption; fix ancestry instead?
+- R: Fix ancestry doesn't prevent you from going back to SI from SC but irreversibility conditions  the ancestral state to SI by default.
+I thought the latter is what we wanted based on Igic et al 2013 correct?
+%The input data for these models was coded as self-compatible (C=0) or self-incompatible (I=1) \cref{table:stateclassifications}.
 
-Next, we proposed a multivariate speciation and extinction model (MuSSE, \citet{fitzjohn_2012}) to investigate the link amongst diversification, breeding system, and polyploidy simultaneously.
-The MuSSE model is defined using three states self-compatible diploids (SC-D=0), polyploids that are always self-compatible (ref for this??) (P=1), and self-incompatible diploids (SI-D=2, see table \cref{table:stateclassifications}) and contains ten parameters, six defining diversification ($\lambda_{SD}, \mu_{SD},\lambda_{P},\mu_{P}, \lambda_{ID},\mu_{ID}$) and the other four key cladogenetic changes that are: polyploidization of self-compatible diploids $\rho_{SD}$, diploidization $\delta$, polyploidization of self-incompatible diploids $\rho_{ID}$, and self-incompatible to self-compatible rate $q_{IC}$.
-% E: for SC, use C instead of S
-% E: lack of SI-P should maybe be explained in Intro?
+\subsection{Models for Polyploidy, Breeding System, and Diversification}
+As a third approach, we analyzed the roles of both polyploidy and breeding system in the process of diversification using a multivariate speciation and extinction model (MuSSE, \citet{fitzjohn_2012}).
+We named our MuSSE the ID/P/CD breeding system and polyploidy model, where the three states are self-incompatible diploids (ID), polyploids (always self-compatible, P=1), and self-compatible diploids (CD).
+The model is defined via ten parameters; six defining diversification ($\lambda_{CD}, \mu_{CD},\lambda_{P},\mu_{P}, \lambda_{ID},\mu_{ID}$) and  four parameters defining key trait changes: polyploidization of self-compatible diploids $\rho_{C}$, diploidization $\delta$, polyploidization of self-incompatible diploids $\rho_{I}$, and self-incompatible to self-compatible rate $q_{IC}$.
 
-Since the null hypothesis of the MuSSE model is that the diversification is equal and constant for all three states defined, it is possible that MuSSE can also suffer from large type I errors.
-In order to account for heterogeneity of the diversification rates and parse the signal in diversification coming from breeding system and polyploidy we extended MuSSE model to account for a hidden state, as \citet{beaulieu_2016} did in the bivariate case.
-The model we proposed is a multivariate and hidden states speciation and extinction stochastic process (MuHiSSE) that allowed us to account simultaneously for the diversification rates linked to breeding system and polyploidy but also the presence of some more heterogeneity in the process.
-This model is analogous to the GeoHiSSE model proposed by Caetano et al. (2018) in a biogeographical context.
-The full MuHiSSE model has 26 parameters, however, our goal was to look for diversification rate differences so we fitted a simplified version of 17 parameters by fixing the rates amongst hidden states to be equal with parameter $\alpha$ and the transition rates amongst breeding system and polyploidy as defined in the MuSSE model  ($\rho_{SD}$,  $\delta$,  $\rho_{ID}$,  $q_{IC}$) despite the hidden state.
-Using MuHiSSE we estimated twelve speciation and extinction rates ($\lambda_{SD_A}, \mu_{SD_A},\lambda_{P_A},\mu_{P_A}, \lambda_{ID_A},\mu_{ID_B},\lambda_{SD_B}, \mu_{SD_B},\lambda_{P_B},\mu_{P_B}, \lambda_{ID_B},\mu_{ID_B}$), and the net diversification and relative extinction rates associated with them.
-% E: caetano ref; and huang
+% E: lack of SI-P is now mentioned in Data and should be explained in Intro: Fixed in data still needs to be mentioned in intro
 
-All the models were performed using RevBayes \citep{hoehna_2016} software that performs Bayesian inference via MCMC using Metropolis-Hastings algorithm.
-A correction for sampling bias was done in all models by assuming that Solanaceae family has approximately 3,000 species ($s=595/3000$) as the Solanaceae Source project indicates \citep{solsource}.
-In all models, speciation and extinction parameters used log-normal prior distributions that with mean the expected net diversification rate $ ( \frac{(number of taxa)/2}{root age})$ and standard deviation $0.5$.
-Prior distributions for parameters defining cladogenetic changes were gamma distributed with parameters $k=0.5$ and $\theta=1$. 
-MCMC was performed for 96 hours in the cluster at Minnesota Supercomputing Instute which allowed for 5,000 generation of burn-in and a minimum of 200,000 generations of MCMC for each of the 6 models.
-Convergence of the MCMC was tested using R package coda (ref) and software Tracer (ref) to assess convergence and mixing (see supplementary information).
+Since the null hypothesis of the MuSSE model is that the diversification is equal and constant for all three states defined, it is possible that MuSSE can also suffer from large type I errors as their BiSSE equivalents.
+Therefore.
+we extended MuSSE model to account for a hidden state like the methodology recently proposed by \citet{caetano_2018} and \citet{herrera_2018}.
+A full fitted MuHiSSE model would have 26 parameters total \citep{herrera_2018}.
+Since our goal was to look for diversification rate differences polyploidy and breeding system rather than differences occurring at  hidden state, we fitted a simplified version, the  ID/P/CD-A/B model, that is defined via16 parameters.
+The reduction in parameter space is achieved by fixing the rates amongst hidden states to be equal with parameter $\alpha$ and fixing the transition rates amongst breeding system to be the same as  the ones defined in the MuSSE model ($\rho_{ID}$,  $\delta$,  $\rho_{CD}$,  $q_{IC}$) independently from the hidden state.
+Therefore, the ID/P/CD- A/B model was used to estimate twelve speciation and extinction rates ($\lambda_{ID_A}, \mu_{ID_A},\lambda_{P_A},\mu_{P_A}, \lambda_{CD_A},\mu_{CD_B},\lambda_{ID_B}, \mu_{ID_B},\lambda_{P_B},\mu_{P_B}, \lambda_{CD_B},\mu_{CD_B}$), and the net diversification and relative extinction rates associated with them.\newline
+
+\subsection{Diploidization as an exploratory hypothesis}
+For all the four models that consider ploidy changes we assumed that diploidization could happen.
+Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like ChromEvol \citep{mayrose_2010}.
+These types of classifications do not allow for a ploidy reversion.
+Where indicated, the classification of ploidy for the data used in our models was based on chromosome multiplicity at the genus level.
+However, the majority of the ploidy classifications were adopted from original studies with alternative sources of information (i.e.
+geographic distribution, genus ploidy distribution) where ploidy was defined by authors that found evidence for it.
+Since it is not clear whether diploidization can be detected under alternative ploidy classifications or even classifications based on chromosome number multiplicity at the genus level, we also fitted the models without diploidization in order to test  whether including an assumption of diploidization is sensitive.
+As discussed by \citet{servedio_2014}  the presence or absence of a hypothesis can have an exploratory goal.
+In our case the diploidization parameter (or the lack of denoted by no $\delta$) in our models  is an opportunity to explore an assumption that might be important but it is not the single definitive process to understand the interactions among polyploidy, breeding system, and diversification.
+
+\subsection{Statistical Inference under the models}
+Parameters for each of the ten diversification models described previously were performed using custom code in RevBayes \citep{hoehna_2016} environment.
+Code and results for all analyses is freely available at https://github.com/roszenil/solploidy .
+For all analyses, a correction for sampling bias is included assuming that Solanaceae family has approximately 3,000 species ($s=595/3000$) as estimated by the Solanaceae Source project \citep{solsource}.
+For all ten models, we assumed that speciation and extinction parameters for all models  had log-normal prior distributions with means equal the expected net diversification rate $ ( \frac{(\textrm{number of taxa})/2}{\textrm{root age}})$ and standard deviation $0.5$.
+Prior distributions for parameters defining trait changes were  assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
+For each model, an MCMC was performed for 96 hours in the high-performance computational cluster at Minnesota Supercomputing Institute which allowed for 5,000 generation of burn-in and a minimum of 200,000 generations of MCMC for each of the 6 models.
+For each model, convergence and mixing of the MCMC was tested using R package coda and software Tracer (see supplementary information for convergence plots).
+
+\subsection{Model selection}
+We calculated the marginal likelihoods  for each of the ten models using custom code in RevBayes \citet{hoehna_2016}.
+Marginal likelihoods were calculated using 50  stepping stone steps using the methodology of \citet{xie_2010}.
+Each stepping stone step was found by calculating 500 generations of burn-in followed by a total of 1000 MCMC steps \cref{table:marginallike}.
+The calculation of each marginal likelihood ran for 24 hours in the high-performance computational cluster at Minnesota Supercomputing Institute.\newline
+
+Using the marginal likelihood values we calculated fourteen different Bayes factors in log-scale.
+The first six Bayes factors compared the four models of polyploidy against each other, followed by one single comparison between the breeding system models, and finalizing with four comparisons between each polyploidy and breeding system models (see \cref{table:bayesfactors} for all comparisons).
+Comparison between polyploidy models vs.
+polyploidy and breeding system models are not valid for two reasons: first, the input data in each model is numerically different (see \cref{table:stateclassifications} for data classifications)
+%R: Emma this is where the figure can be important to show that we have to make different input
+; second, even when it is possible to re-write the polyploidy and diversification models to include a hidden state that would account for breeding system allowing to use the same input data, the re-written model would still not be comparable to the multivariate SSE models proposed for polyploidy and breeding system due to the hierarchical structure imposed by the interaction between breeding system and polyploidy.
+This second point has been recently denoted by \citet{tarasov_2019} as the lack of lumpability in Markov models for discrete traits.
+In our particular case, polyploids cannot be self compatible, and the irreversibility of self-compatible to self-incompatible state makes the three-state Markov models non-lumpable into two-state models ultimately making  multivariate and the bivariate models incomparable.
 
@@
 \multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
-Type & Total of & BiSSE/HiSSE & BiSSE/HiSSE & MuSSE/MuHiSSE\\ 
- &Taxa &  Polyploidy & Breeding System &  \\ \midrule
+Type & Total of & D/P and D/P-A/B& I/C and I/C-A/B & ID/P/CD and CD/P/ID-A/B\\ \midrule
 Diploid Self Compatible & 152 & 0 &  0 & 0 \\
@@
 \end{table}
+
+\begin{table}
+\begin{tabular}{@{}llcccccc@{}} \toprule
+%\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
+Model& Model& Ploidy & Diploidization & Breeding  & Hidden & Parameters & Marginal \\
+& Type & &  &System & State & & Log- Likelihood \\
+1. D/P &BiSSE &	Yes  &	Yes &	No	&No	& 6	& -1182.93 \\
+2. D/P no $\delta$ & BiSSE  &	Yes & 	No	& No	& No & 	5 &	-1193.66\\
+3. D/P- A/B  & HiSSE &	Yes &	Yes	&No &	Yes &	11 &	\textbf{-1145.69}\\
+4. D/P no $\delta$-A/B &	HiSSE &	Yes & No &	No &	Yes &	10	&-1150.99\\
+5. I/C &	BiSSE &	No &	 No	&Yes &	No &	 5 &  -1194.80 \\
+6. I/C-A/B &	HiSSE &	No &	 No	&Yes &	Yes	& 10 & \textbf{-1155.37}\\
+7. ID/P/CD & MuSSE &	Yes & 	Yes &	Yes &	No &	10 & -1344.50\\
+8. ID/P/CD no $\delta$ &	MuSSE &	Yes & 	No &	Yes	&No &	9 &-1345.87\\
+9. ID/P/CD-A/B & MuHiSSE &	Yes 	&Yes &	Yes &	Yes &	16 & \textbf{-1300.35} \\
+10.ID/P/CD no $\delta$-A/B &	 MuHiSSE	 & Yes & 	No	&Yes &	Yes	&15 &-1303.55 \\ \bottomrule
+\end{tabular}
+\caption{Marginal likelihoods for the ten diversification models proposed.}
+% E: but P are not SI
+\label{table:marginallike}
+\end{table}
+
+\begin{table}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|c|c|c|}
+\toprule
+Polyploidy Models & Breeding System Models & Polyploidy and Breeding System Models \\ \midrule
+{\begin{tabular}{lcccc}
+ & 1 & 2 & 3 & 4 \\
+1. D/P & $\cdot$	 &10.72 &	-37.24	&-31.94\\
+2. D/P no $\delta$ &$\cdot$&$\cdot$ &	-47.97	&-42.66\\
+\textbf{3. D/P- A/B}  &$\cdot$  & $\cdot$&	$\cdot$	& 5.30 \\
+4. D/P no $\delta$-A/B &$\cdot$& $\cdot$ & $\cdot$&$\cdot$ \\
+\end{tabular}
+}  & 
+{\begin{tabular}{lcc}
+ & 5 & 6\\
+5. I/C &$\cdot$ & -39.43 \\
+\textbf{6. I/C-A/B} &$\cdot$& $\cdot$ \\
+& & \\
+& & \\
+\end{tabular}
+} & 
+{\begin{tabular}{lcccc}
+& 7 & 8 & 9 & 10\\
+7. CD/P/ID & $\cdot$	&1.36 & -44.15&-40.95 \\
+8. CD/P/ID no $\delta$ & $\cdot$ & $\cdot$ & -45.51 &-42.31 \\
+9. CD/P/ID-A/B& $\cdot$ & $\cdot$ &$\cdot$	& 3.2 \\ 
+10. ID/P/CD no $\delta$-A/B & $\cdot$& $\cdot$&$\cdot$ & $\cdot$\\
+\end{tabular}
+}\\
+\bottomrule
+\end{tabular}
+ \caption{Bayes factors in log scale. We compare every possible pair. Number models as indicated in  \cref{table:marginallike}. }
+% E: but P are not SI
+\label{table:bayesfactors}
+\end{table}
+
+%\begin{tabular}{@{}lcccccccccc@{}} \toprule
+%\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
+%Model& 1& 2 & 3 & 4  &5 & 6& 7 & 8 & 9 & 10 \\
+%Polyploidy & & & & & & & & & \\
+%1. D/P &	-  &	10.727 &	-37.243	&-31.941& -	& - & - & - & - &- \\
+%2. D/P no $\delta$ & -  & -  &	-47.97	&-42.668& -	& - & - & - & - &- \\
+%3. D/P- A/B  & -  & -  &	-	& 5.302 & -	& - & - & - & - &- \\
+%4. D/P no $\delta$-A/B &
+%Breeding System& & & & & & & & & \\
+%5. C/I & -  & -  &	-	& - & -	& -39.434 & - & - & - &- \\
+%6. I/C-A/B &
+%Polyploidy and Breeding System & & & & & & & & & \\
+%7. CD/P/ID & -  & -  &	-	& - & -	& - & - &1.368 & -44.151&-40.951 \\
+%8. CD/P/ID no $\delta$& -  & -  &	-	& - & -	& - & - &-& -45.519&-42.319 \\
+%9. CD/P/ID-A/B& -  & -  &	-	& - & -	& - & - &-& -& 3.2 \\ 
+%10.ID/P/CD no $\delta$-A/B &
+%\bottomrule
+%\end{tabular}

### 2019-02-09 manuscript/results.tex
Updated methods and bayes factors table

\section{Results}
\subsection{Polyploidy and Diversification Models}
Similarly to the results obtained by \citet{mayrose_2011} and \citet{mayrose_2015}, the D/P polyploidy  model we found that net diversification of diploids is significantly higher than the the net diversification of polyploids. This result holds true whether or not the diploidization parameter is present. However, when diploidization is present the net diversification of polyploids is nonnegative with probability 1, whereas in the absence of diploidization the net diversification of polyploids is negative with probability 0.75(HERE VERIFY QUANTILE). In terms of the relative extinction, when the diploidization parameter is present both polyploids and diploids have posterior distributions that overlap, but that pattern changes in the absence of the diploidization parameter leading to a significant difference between relative extinction where polyploids have a significant higher relative extinction rate. Polyploidization rate $rho$ has a credible interval between (a,b) and diploidization rate has a credible interval between (c,d).

For the D/P- A/B model with diploididization the diploid and polyploid net diversification rates are overlapping for both state A and B of the hidden trait. In this model, the differences in net diversification are due to the presence of a hidden trait and not to the differences in ploidy. When diploidization parameter is absent the pattern holds. Rates of polyploidization have a 95\% credible interval of (a,b) and diploidization rate has a 95\% credible interval ()  for the model with diploidization, and polyploidy rate has a 95\% credible interval (c,d) for the model without diploidization.

\subsection{Breeding System and Diversification models }

\subsection{Bre}

### 2019-02-11 manuscript/methods.tex
updated figures and partial results

@@
 
-\begin{table}
-\begin{tabular}{@{}llccc@{}} \toprule
-\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
-Type & Total of & D/P and D/P-A/B& I/C and I/C-A/B & ID/P/CD and CD/P/ID-A/B\\ \midrule
-Diploid Self Compatible & 152 & 0 &  0 & 0 \\
-Diploid Self Incompatible& 97 & 0  & 1 & 2\\
-Diploid with unknown breeding system & 219 & 0 & (0,1) & (0,2) \\
-Polyploid & 81 & 1& 0 & 1 \\
-Unknow ploidy and self compatible& 34 & (0,1)& 0 & (0,1) \\ 
-Unknown ploidy and self incompatible & 12 & 0 & 1 & 2 \\ \bottomrule
-\end{tabular}
-\caption{Binary and three state classifications for 595 taxa with ploidy and/or breeding system data.
-The number of taxa in the t he sample was maximize by including tips with only ploidy or only breeding system and assigned them as uncertain in the unknown character.}
+%\begin{table}
+%\begin{tabular}{@{}llccc@{}} \toprule
+%\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
+%Type & Total of & D/P and D/P-A/B& I/C and I/C-A/B & ID/P/CD and CD/P/ID-A/B\\ \midrule
+%Diploid Self Compatible & 152 & 0 &  0 & 0 \\
+%Diploid Self Incompatible& 97 & 0  & 1 & 2\\
+%Diploid with unknown breeding system & 219 & 0 & (0,1) & (0,2) \\
+%Polyploid & 81 & 1& 0 & 1 \\
+%Unknow ploidy and self compatible& 34 & (0,1)& 0 & (0,1) \\ 
+%Unknown ploidy and self incompatible & 12 & 0 & 1 & 2 \\ \bottomrule
+%\end{tabular}
+%\caption{Binary and three state classifications for 595 taxa with ploidy and/or breeding system data.
+The number of taxa in the the sample was maximized by including tips with only ploidy or only breeding system and assigned them as uncertain in the unknown character.}
 % E: but P are not SI
-\label{table:stateclassifications}
-\end{table}
+%\label{table:stateclassifications}
+%\end{table}
 
@@
 & 7 & 8 & 9 & 10\\
-7. CD/P/ID & $\cdot$	&1.36 & -44.15&-40.95 \\
-8. CD/P/ID no $\delta$ & $\cdot$ & $\cdot$ & -45.51 &-42.31 \\
-9. CD/P/ID-A/B& $\cdot$ & $\cdot$ &$\cdot$	& 3.2 \\ 
+7. ID/P/CD & $\cdot$	&1.36 & -44.15&-40.95 \\
+8. ID/P/CD no $\delta$ & $\cdot$ & $\cdot$ & -45.51 &-42.31 \\
+\textbf{9. ID/P/CD-A/B}& $\cdot$ & $\cdot$ &$\cdot$	& 3.2 \\ 
 10. ID/P/CD no $\delta$-A/B & $\cdot$& $\cdot$&$\cdot$ & $\cdot$\\

### 2019-02-11 manuscript/results.tex
updated figures and partial results

@@
 \subsection{Polyploidy and Diversification Models}
-Similarly to the results obtained by \citet{mayrose_2011} and \citet{mayrose_2015}, the D/P polyploidy  model we found that net diversification of diploids is significantly higher than the the net diversification of polyploids.
+Similarly to the results obtained by \citet{mayrose_2011} and \citet{mayrose_2015}, we found that in the D/P polyploidy model the net diversification of diploids is larger than the the net diversification of polyploids since the net diversification distributions do not overlap (Figure 2(A)).
 This result holds true whether or not the diploidization parameter is present.
-However, when diploidization is present the net diversification of polyploids is nonnegative with probability 1, whereas in the absence of diploidization the net diversification of polyploids is negative with probability 0.75(HERE VERIFY QUANTILE).
-In terms of the relative extinction, when the diploidization parameter is present both polyploids and diploids have posterior distributions that overlap, but that pattern changes in the absence of the diploidization parameter leading to a significant difference between relative extinction where polyploids have a significant higher relative extinction rate.
-Polyploidization rate $rho$ has a credible interval between (a,b) and diploidization rate has a credible interval between (c,d).
+However in the presence of  the diploidization parameter the net diversification rate of polyploids is nonnegative with probability 1 (Figure 2(A)), whereas in the absence of diploidization the net diversification rate of polyploids can be negative with a probability (HERE verify the quantile) (FIGURE 3(A)).
+In terms of the relative extinction, when the diploidization parameter is present both polyploids and diploids have posterior distributions that overlap, but that pattern changes in the absence of the diploidization parameter leading to a significant difference between relative extinction where polyploids have a significant higher relative extinction rate (see Supplementary Information).\newline
 
-For the D/P- A/B model with diploididization the diploid and polyploid net diversification rates are overlapping for both state A and B of the hidden trait.
+For the D/P- A/B model with diploididization the diploid and polyploid net diversification rates are overlapping for both state A and B of the hidden trait (Figure 2(B)).
 In this model, the differences in net diversification are due to the presence of a hidden trait and not to the differences in ploidy.
-When diploidization parameter is absent the pattern holds.
-Rates of polyploidization have a 95\% credible interval of (a,b) and diploidization rate has a 95\% credible interval ()  for the model with diploidization, and polyploidy rate has a 95\% credible interval (c,d) for the model without diploidization.
+When diploidization parameter is absent the hidden state is still driving the differences in diversification rates (Figure 3(B)).
 
 \subsection{Breeding System and Diversification models }
+In the I/C breedyng system model we found that the net diversification rate for self-incompatible state is larger than the net diversification rate for self-compatible state (Figure 2(C)).
+The net diversification rate for self-compatible has a probability distribution centered at zero.
+\newline
 
-\subsection{Bre}
+When a hidden state is added in the I/C-A/B model, we found that under the hidden state A the self-compatible and self-incompatible net diversification rates are different.
+Under the hidden state B, those two rates overlap with a probability (HERE CALCULATE THAT) meaning that they are different with probability  (CALCULATE) (Figure 2(D)).
+These results agree with previous results found by  \citet{goldberg_2012}.
+However, \citet{goldberg_2012} used a ClaSSE  approach since they were interested in anagenetic and cladogenetic changes for self-incompatibility using a smaller subset of the data presented in the current work.
+
+% Describe rate q_IC and hidden state rate
+
+\subsection{Polyploidy and Breeding Sytem models}
+In the ID/P/CD model we found that self-incompatible and diploid state has a significantly larger net diversification rate compare to both self-compatible diploid and polyploid rates.
+Meanwhile, both self-compatible diploid and polyploid posterior distributions of net diversification rates completely overlap (Figure 2(E)).
+When hidden state was added in the ID/P/CD-A/B model, we observed significant differences between self-compatible and self-incompatible diploids for both A and B values of the hidden state.
+Self-incompatible state had a larger net diversification rate than self compatible for both A and B states.
+However, the posterior distribution for the net diversification rate of polyploids overlaps with both the self-compatible and self-incompatible posterior distributions for each value of the hidden state(Figure 2F) meaning that polyploidy state is not significantly different from diploid in net diversification terms.
+The resulting effect of adding the hidden state values is significant
+
+\subsection{Diploidization as an exploratory hypothesis}
+In the D/P model  the diploidization rate $\delta$ and polyploidization rate $\rho$ are different from zero with probability 1.
+Diploidization rate is more uncertain than polyploidization.
+For the D/p no $\delta$ model, the polyploidization rate is still different from zero but has a wider 95\% credible interval (see Supplementary Information).
+
+For the models containing both polyploidy and breeding system traits, we found that the diploidization rate is really uncertain with a MAP that is (CALCULATE VALUE HERE- close to zero).
+The polyploidization rate from self-incompatible diploid $\rho_I$ is slightly faster than the polyploidization rate from self-compatible diploid $\rho_C$
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{states.pdf}
+  \caption{Binary and three state classifications for 595 taxa with ploidy and/or breeding system data.
+The number of taxa in the the sample was maximized by including tips with only ploidy or only breeding system and assigned them as uncertain in some of the models.
+For example, the 34 self-compatible taxa with unknown ploidy were assigned as uncertain (diploid or polyploid) in the D/P models but as fully compatible in the I/C models.}
+\label{figure:stateclassifications}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{Netdiversificationallmodels.pdf}
+  \caption{Net diversification rates for all models that include diploidization.}  
+\label{figure:netdivall}
+\end{figure}

### 2019-02-12 manuscript/results.tex
More results, fixing figure 2

@@
 Diploidization rate is more uncertain than polyploidization.
-For the D/p no $\delta$ model, the polyploidization rate is still different from zero but has a wider 95\% credible interval (see Supplementary Information).
+For the D/P no $\delta$ model, the polyploidization rate is still different from zero but has a wider 95\% credible interval (see Supplementary Information).
+In the D/P-A/B model the diploidization rate remains positive but there rate of polyploidization becomes really uncertain.
+Whereas, in the absence of diploidzation (D/P no $\delta$ A/B) the rate of polyploidy has a small credible interval \newline
 
-For the models containing both polyploidy and breeding system traits, we found that the diploidization rate is really uncertain with a MAP that is (CALCULATE VALUE HERE- close to zero).
-The polyploidization rate from self-incompatible diploid $\rho_I$ is slightly faster than the polyploidization rate from self-compatible diploid $\rho_C$
+For the model ID/P/CD containing both polyploidy and breeding system traits, we found that the diploidization rate is really uncertain with a MAP that is (CALCULATE VALUE HERE- close to zero).
+The polyploidization rate from self-incompatible diploid $\rho_I$ is slightly faster than the polyploidization rate from self-compatible diploid $\rho_C$  and that pattern remains the same for the ID/P no $\delta$/CD model(see supplementary information).
+%Here I'm missing the MuHiSSE plots.
+
+\subsection{Model selection}
+
+In \cref{table:marginallike} we list the marginal likelihood in log scale for each of the models tested.
+In the table we show what are the different components included and excluded for each model as a a summary of the diagrams from \cref{figure:netdivall}.
+From the marginal likelihoods in log scale, the Bayes factors in log-scale were calculated as shown in table \cref{table:bayesfactors}.
+After testing every single pair of polyploidy modes (1-4) we found overwhelming evidence that the best polyploidy model is always  the D/P-A/B, that is the model with hidden state and diploidization.
+\newline
+For the two models following the evolution of breeding system, the I/C-A/B is the best choice between models 3 and 4.\newline
+The models that follow the diversification linked to both polyploidy and breeding system are the last 4 (models 5-8).
+When comparing using Bayes factores every two models we found that the IC/P/CD-A/B is always preferred over the rest, meaning that the model that has a hidden state and diploidization is chosen over the ones that lack either of both of those options.\newline
+Therefore, the models that were chosen were always the ones containing a hidden trait, and in the case of polyploidy models, the ones containing a diploidization parameter $\delta$ are preferable.
 

### 2019-02-13 manuscript/methods.tex
started editing Methods

@@
 
-Chromosome number data were obtained for all Solanaceae taxa in the the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.\ 14,000 records were cleaned semi-automatically using the CCDBcurator R package \citep{rivero_2019}.
-This large dataset includes the compilation of Solanaceae ploidy states from \citep{robertson_2011}.
+Chromosome number data were obtained for all Solanaceae taxa in the the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were cleaned semi-automatically using the CCDBcurator R package \citep{rivero_2019}.
+This large dataset includes the compilation of Solanaceae ploidy states from \citet{robertson_2011}.
 Species were coded as either diploid (D) or polyploid (P).
@@
 Species with more than one ploidy level were assigned the smallest and most frequent ploidy level recorded.
-We also cross-verified using the breeding system information and found that there were no self-incompatible species possibly listed as polyploid, or listed as polyploid and diploid which increase our confidence in the assignations% E: Does this apply to the diploid-polyploid levels?
-R- Yes there are no self-incompatible polyploids at all.
-If so, we need to explain how it is also reconciled against the SI/SC value.
-Or is this sentence just about multiple levels of polyploidy, which wouldn't need to be mentioned here?
-R:  Solved
-Breeding system was scored as self-incompatible (I) or self-compatible (C) based on results hand-curated from the literature (as in \citealt{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}).
-Most species could unambiguously be coded as either SI or SC \citep{raduski_2012}.
+% E: There was more here about cross-verifying that no P species were SI.  I removed it because there were conflicts before, but we resolved them.  This is addressed in the next paragraph.
+%
+Breeding system was scored as self-incompatible (I) or self-compatible (C) based on results hand-curated from the literature \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
+Most species could unambiguously be coded as either I or C \citep{raduski_2012}.
 Following previous work, we coded as I any species with functional I systems, even if C or dioecy was also reported.
 Dioecious species without functional I were coded as C.
+
+% E: I am trying to stick with I/C when talking about the state values, but using SI/SC when talking about the meaning of the trait (because this what people use normally).
 
@@
 Hybrids and cultivars were excluded because ploidy and breeding system are likely to be altered by domestication.
-As in \citet{robertson_2011}, we examined closely the few species for which the merged ploidy and breeding system data indicated the presence of I polyploids.
-Although SI populations frequently contain some C individuals and diploid populations frequently contain some polyploid individuals, in no case did we find individuals that were both SI and polyploid.
-Because of this empirical observation and the functional explanation for whole genome duplication disabling gametophytic self-incompatibility \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploid (ID), self-compatible diploid (CD), and polyploid (P) which are always self-compatible.
+As in \citet{robertson_2011}, we examined closely the few species for which the merged ploidy and breeding system data indicated the presence of self-incompatible polyploids.
+Although SI populations frequently contain some SC individuals and diploid populations frequently contain some polyploid individuals, in no case did we find individuals that were both SI and polyploid.
+Because of this empirical observation and the functional explanation for whole genome duplication disabling gametophytic self-incompatibility \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).
 
-% E: We should probably include some more details about how many polymorphic species and how the particularly tricky ones were dealt with.
+% E: Should we include more details about how many polymorphic species and how the particularly tricky ones were dealt with?
 
 Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 595 species with ploidy and/or breeding system information on the tree.
-Binary or three-state classification of ploidy and breeding system for the 595 taxa is summarized in \cref{table:stateclassifications}.
-% E: I think you're still updating the state coding?
-R=no since models are not comparable there is no need.
-We retained all of these species in each of the analyses below, because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
-A total of 405 taxa without any information about breeding system nor polyploidy were pruned.
-Tips without trait data are non informative for  diversification parameters linked to trait change and can significantly slow-down the convergence of inferences in trees with already hundreds of tips.% E: Though someone might ask why tips with no state data were pruned.
-Analyses took long enough on even this-size tree.
-R: I think this point needs to be further investigated in another paper, so I just sort-of justify our choice
+Binary or three-state classification of ploidy and breeding system for the 595 taxa is summarized in \cref{figure:stateclassifications}.
+We retained all of these species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
+A total of 405 taxa without any information about breeding system or polyploidy were excluded.
+Tips without trait data are much less informative for diversification parameters linked to trait values, and including this many more species would have prohibitively slowed our analyses, expecially for the most complex models.
 
-\subsection{Models for Diversification and Polyploidy}
+\subsection{Models for Ploidy and Diversification}
 
-% E: There are so many models, I think we need to try hard to help the reader keep them straight.
-% I would avoid naming any states with numbers, except maybe within the stateclassification table and its caption.
-% I also still find bisse/hisse/muhisse terminology confusing, I think because it both describes a general class of model and applies to a specific analysis here, and because bisse and hisse are run separately on ploidy breeding system.
-% How about a model naming scheme like this?
 % D/P polyploidy model
@@
 % ID/CD/CP-A/B ploidy, breeding system and hidden state model
-% Or something else more concise and descriptive?
 
-We first defined a binary state speciation and extinction model (BiSSE, \citet{maddison_2007})  for polyploidy evolution where taxa was classified as diploid (D=0) and polyploid (P=1) \cref{table:stateclassifications}.
-We refer to this model as the D/P polyploidy model. 
-In a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$) , net diversification rates ($r_D=\lambda_D-\mu_D, r_P=\lambda_P-\mu_P$), and  relative extinction rates ($\nu_D=\frac{\mu_D}{\lambda_D}, \nu_D=\frac{\mu_D}{\lambda_D}$) just as previously explored in \citet{mayrose_2011} and \citet{mayrose_2015}.
-However, our analyses differ from previous models since we consider not only a polyploidization rate $\rho$ and but also a diploidization parameter rate $\delta$.
+To investigate the association between ploidy level and diversification, we first defined a binary state speciation and extinction model (BiSSE, \citealt{maddison_2007}) in which taxa were classified as diploid (D) or polyploid (P) (\cref{figure:stateclassifications}).
+We call this the $D/P$ ploidy model. 
+In a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$), net diversification rates ($r_D=\lambda_D-\mu_D$, $r_P=\lambda_P-\mu_P$), and relative extinction rates ($\nu_D=\mu_D / \lambda_D$, $\nu_D=\mu_D / \lambda_D$) associated with each state.
+This analysis explores the same question as \citet{mayrose_2011, mayrose_2015}, but our analyses differ because we include not only polyploidization (parameter $\rho$, the transition rate from $D$ to $P$), but also diploidization (parameter $\delta$, the transition rate from $P$ to $D$).
+% E: presumably the motivation will be explained in the Intro
 
-As a second step, we fitted a hidden state speciation and extinction model (HiSSE, \citet{beaulieu_2016}) to evaluate whether the differences in diversification rates were found due to a hidden trait associated to polyploidy.
-This model is referred to as the D/P-A/B polyploidy and hidden state model.
-As discussed by \citet{beaulieu_2016} BiSSE-like models suffer from a large type I error since they fail to account as part of the null hypothesis heterogeneous diversification rate changes  that do not directly depend on the trait of interest.
-By expanding the D/P model by using a hidden trait with states A and B,  the polyploidy and hidden state model addressed the signal of diversification due to polyploidy while parsing out the heterogeneity of diversification rates due to unobserved traits.
-The D/P-A/B polyploidy and hidden trait is composed by four states: diploid and polyploid subdivided by a binary hidden trait with states A and B for which we estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\lambda_{D_B}, \lambda_{P_A},\lambda_{P_B}$), extinction rates ($\mu_{D_A},\mu_{D_B}, \mu_{P_A},\mu_{P_B}$),  net diversification rates ($r_{D_A},r_{D_B},r_{P_A},r_{P_B}$), and relative extinction rates($\nu_{D_A},\nu_{D_B}, \nu_{PA},\nu_{P_B}$).
+To evaluate whether diversification rate differences found by the previous model could be explained by something other than ploidy, we also added a hidden state (HiSSE model; \citet{beaulieu_2016}).
+We call this the $D/P--A/B$ ploidy and hidden state model.
+
+% As a second step, we fit a hidden state speciation and extinction model (HiSSE, \citet{beaulieu_2016}) to evaluate whether the differences in diversification rates were found due to a hidden trait associated to polyploidy.
+% This model is referred to as the D/P-A/B polyploidy and hidden state model.
+As discussed by \citet{rabosky_2015, beaulieu_2016}, BiSSE-like models can suffer from a large false discovery rate because they fail to account for diversification rate changes that do not directly depend on the trait of interest.
+By expanding the $D/P$ model to include a hidden trait with states $A$ and $B$, the ploidy and hidden state model addressed the signal of diversification due to polyploidy while parsing out the heterogeneity of diversification rates due a possible unobserved trait.
+The $D/P--A/B$ model is composed by four states: diploid and polyploid subdivided by a binary hidden trait with states A and B for which we estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\lambda_{D_B}, \lambda_{P_A},\lambda_{P_B}$), extinction rates ($\mu_{D_A},\mu_{D_B}, \mu_{P_A},\mu_{P_B}$),  net diversification rates ($r_{D_A},r_{D_B},r_{P_A},r_{P_B}$), and relative extinction rates($\nu_{D_A},\nu_{D_B}, \nu_{PA},\nu_{P_B}$).
 In this model polyploidization rate $\rho$, diploidization rate $\delta$ are also considered, while changes between every hidden state are symmetrical with the same rate $\alpha$.
 
-\subsection{Models for Diversification and Breeding System}
+\subsection{Models Breeding System and Diversification}
+
 In order to analyze the effects of breeding system in the diversification process, two models the I/C (I= self-incompatible, C= self-compatible) and the I/C-A/B were performed.
@@
 
-\subsection{Models for Polyploidy, Breeding System, and Diversification}
+\subsection{Models for Ploidy, Breeding System, and Diversification}
+
 As a third approach, we analyzed the roles of both polyploidy and breeding system in the process of diversification using a multivariate speciation and extinction model (MuSSE, \citet{fitzjohn_2012}).

### 2019-02-14 manuscript/methods.tex
more editing Methods

@@
 
-\subsection{Models for Ploidy and Diversification}
-
-% D/P polyploidy model
-% D/P-A/B polyploidy hidden state model
+% D/P ploidy model
+% D/P-A/B ploidy hidden state model
 % I/C breeding system model
@@
 % ID/CD/CP ploidy and breeding system model
-% ID/CD/CP-A/B ploidy, breeding system and hidden state model
+% ID/CD/CP-A/B ploidy, breeding system, and hidden state model
+
+\subsection{Models for ploidy and diversification}
 
@@
 
-To evaluate whether diversification rate differences found by the previous model could be explained by something other than ploidy, we also added a hidden state (HiSSE model; \citet{beaulieu_2016}).
-We call this the $D/P--A/B$ ploidy and hidden state model.
+Our second model assesses the the signal of diversification due to ploidy while also parsing out the heterogeneity of diversification rates due a possible unobserved trait.
+BiSSE-like models can suffer from a large false discovery rate because they fail to account for diversification rate changes that do not directly depend on the trait of interest \citep{rabosky_2015, beaulieu_2016}.
+To allow for diversification rate differences explained by something other than ploidy, we added a hidden state (HiSSE model; \citealt{beaulieu_2016}).
+In this model, each of the observed diploid and polyploid states is subdivided by a binary hidden trait with states $A$ and $B$.
+We call this the $D/P-A/B$ ploidy and hidden state model. 
+We estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$), extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$), net diversification rates ($r_{D_A},\ r_{D_B},\ r_{P_A},\ r_{P_B}$), and relative extinction rates ($\nu_{D_A},\ \nu_{D_B},\ \nu_{P_A},\ \nu_{P_B}$).
+In this model polyploidization rate $\rho$ and diploidization rate $\delta$ are also included, and changes between hidden states are symmetrical with rate $\alpha$.
 
-% As a second step, we fit a hidden state speciation and extinction model (HiSSE, \citet{beaulieu_2016}) to evaluate whether the differences in diversification rates were found due to a hidden trait associated to polyploidy.
-% This model is referred to as the D/P-A/B polyploidy and hidden state model.
-As discussed by \citet{rabosky_2015, beaulieu_2016}, BiSSE-like models can suffer from a large false discovery rate because they fail to account for diversification rate changes that do not directly depend on the trait of interest.
-By expanding the $D/P$ model to include a hidden trait with states $A$ and $B$, the ploidy and hidden state model addressed the signal of diversification due to polyploidy while parsing out the heterogeneity of diversification rates due a possible unobserved trait.
-The $D/P--A/B$ model is composed by four states: diploid and polyploid subdivided by a binary hidden trait with states A and B for which we estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\lambda_{D_B}, \lambda_{P_A},\lambda_{P_B}$), extinction rates ($\mu_{D_A},\mu_{D_B}, \mu_{P_A},\mu_{P_B}$),  net diversification rates ($r_{D_A},r_{D_B},r_{P_A},r_{P_B}$), and relative extinction rates($\nu_{D_A},\nu_{D_B}, \nu_{PA},\nu_{P_B}$).
-In this model polyploidization rate $\rho$, diploidization rate $\delta$ are also considered, while changes between every hidden state are symmetrical with the same rate $\alpha$.
+\subsection{Models for breeding system and diversification}
 
-\subsection{Models Breeding System and Diversification}
+To assess the effects of breeding system in the diversification process, we first fit model in which the states are self-incompatible (I) or self-compatible (C).
+This is the same as the analysis of \citet{goldberg_2010}, though with an updated phylogeny \citep{sarkinen_2013}.
+We call this BiSSE model the $I/C$ breeding system model. 
+To parse out the effect of breeding system in diversification while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
+We call this HiSSE model the $I/C-A/B$ breeding system and hidden state model.
 
-In order to analyze the effects of breeding system in the diversification process, two models the I/C (I= self-incompatible, C= self-compatible) and the I/C-A/B were performed.
-The I/C breeding system model (BiSSE) was done with the goal of investigating if the effect of self-incompatibility in diversification similar to the analysis performed by \citet{goldberg_2012}.
-The I/C-A/B  breeding system with hidden trait model parses out the effect of breeding system in diversification by incorporating the possibility of heterogenous changes in the diversification rates that can be attributed to a hidden state.
-A hidden state model has not previously been fitted to Solanaceae data.
-For these two models, we assumed that self-compatibility is irreversible as discussed by \citet{igic_2013}, so we defined the transition rate $q_{IC}$ as the parameter for changes from self-incompatible to self-compatible state.
-This is an important assumption because ancestral state reconstruction in the models might differ when polyploidy and breeding system are analyzed as independent  in the Solanaceae tree.
-This irreversibility assumption prevents self-compatible state from evolving into self-incompatible state and as a result, the ancestral state of all taxa in this model has probability one of being self-incompatible.
-% E: explain irreversibility assumption; fix ancestry instead?
-- R: Fix ancestry doesn't prevent you from going back to SI from SC but irreversibility conditions  the ancestral state to SI by default.
-I thought the latter is what we wanted based on Igic et al 2013 correct?
-%The input data for these models was coded as self-compatible (C=0) or self-incompatible (I=1) \cref{table:stateclassifications}.
+For these and the subsequent models including breeding system, we allow transitions from $I$ to $C$ (at rate $q_{IC}$) but not the reverse.
+Within Solanaceae, self-incompatibility in species that possess it is a homologous form of gametophytic SI \citep[shared even shared even with other dicot families;][]{igic_2001, steinbachs_2002}.
+Furthermore, species that are distantly related within the family carry several closely-related alleles at the locus that controls the SI response \citep{ioerger_1990, igic_2006}.
+This is very strong evidence that the SI mechanism is ancestral in Solanaceae and has not re-evolved within the family.
+% This is an important assumption because ancestral state reconstruction in the models might differ when polyploidy and breeding system are analyzed as independent in the Solanaceae tree.
+% This irreversibility assumption prevents self-compatible state from evolving into self-incompatible state and as a result, the ancestral state of all taxa in this model has probability one of being self-incompatible.
 
-\subsection{Models for Ploidy, Breeding System, and Diversification}
+\subsection{Models for ploidy, breeding system, and diversification}
 
@@
 \subsection{Diploidization as an exploratory hypothesis}
+
 For all the four models that consider ploidy changes we assumed that diploidization could happen.
@@
 \subsection{Statistical Inference under the models}
+
 Parameters for each of the ten diversification models described previously were performed using custom code in RevBayes \citep{hoehna_2016} environment.
@@
 \subsection{Model selection}
+
 We calculated the marginal likelihoods  for each of the ten models using custom code in RevBayes \citet{hoehna_2016}.

### 2019-02-14 manuscript/methods.tex
more editing Methods

@@
 
-As a third approach, we analyzed the roles of both polyploidy and breeding system in the process of diversification using a multivariate speciation and extinction model (MuSSE, \citet{fitzjohn_2012}).
-We named our MuSSE the ID/P/CD breeding system and polyploidy model, where the three states are self-incompatible diploids (ID), polyploids (always self-compatible, P=1), and self-compatible diploids (CD).
-The model is defined via ten parameters; six defining diversification ($\lambda_{CD}, \mu_{CD},\lambda_{P},\mu_{P}, \lambda_{ID},\mu_{ID}$) and  four parameters defining key trait changes: polyploidization of self-compatible diploids $\rho_{C}$, diploidization $\delta$, polyploidization of self-incompatible diploids $\rho_{I}$, and self-incompatible to self-compatible rate $q_{IC}$.
+If both ploidy and breeding system have the potential to influence lineage diversification, we should logically consider their effects jointly.
+We thus fit a model that includes both traits \citep[MuSSE,][]{fitzjohn_2012}.
+The three states in this model are self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids, which are always self-compatible (CP).
+We did not include a state for self-incompatible polyploids because they are not present in the data, and because that trait combination is logically expected not to occur (as explained above).
+We call this the $ID/CD/CP$ ploidy and breeding system model.
+It has ten parameters, six for diversification in each state ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and four for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $\delta$ for diploidization from $CP$ to $CD$; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
+The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
+Diploidization from $CP$ to $ID$ is not allowed because it would be a regain of self-incompatibility.
 
-% E: lack of SI-P is now mentioned in Data and should be explained in Intro: Fixed in data still needs to be mentioned in intro
-
-Since the null hypothesis of the MuSSE model is that the diversification is equal and constant for all three states defined, it is possible that MuSSE can also suffer from large type I errors as their BiSSE equivalents.
-Therefore.
-we extended MuSSE model to account for a hidden state like the methodology recently proposed by \citet{caetano_2018} and \citet{herrera_2018}.
-A full fitted MuHiSSE model would have 26 parameters total \citep{herrera_2018}.
-Since our goal was to look for diversification rate differences polyploidy and breeding system rather than differences occurring at  hidden state, we fitted a simplified version, the  ID/P/CD-A/B model, that is defined via16 parameters.
-The reduction in parameter space is achieved by fixing the rates amongst hidden states to be equal with parameter $\alpha$ and fixing the transition rates amongst breeding system to be the same as  the ones defined in the MuSSE model ($\rho_{ID}$,  $\delta$,  $\rho_{CD}$,  $q_{IC}$) independently from the hidden state.
-Therefore, the ID/P/CD- A/B model was used to estimate twelve speciation and extinction rates ($\lambda_{ID_A}, \mu_{ID_A},\lambda_{P_A},\mu_{P_A}, \lambda_{CD_A},\mu_{CD_B},\lambda_{ID_B}, \mu_{ID_B},\lambda_{P_B},\mu_{P_B}, \lambda_{CD_B},\mu_{CD_B}$), and the net diversification and relative extinction rates associated with them.\newline
+The $ID/CD/CP$ model could potentially capture similar dynamics as earlier models, if the hidden state in $D/P-A/B$ was effectively breeding system, and the hidden state in $I/C-A/B$ was effectively ploidy.
+There is also the potential, however, for a hidden factor to be influencing diversification even beyond both of our focal traits, and this could again mislead inferences.
+We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018, herrera_2018, huang_2018}.
+We refer to this as the $ID/CD/CP-A/B$ model.
+A fully parameterized version of this model would have 26 rate parameters \citep{herrera_2018}. 
+Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fitted a simplified version with 16 parameters.
+The reduction in parameter space is achieved by fixing the rates for transitions among hidden states to be equal with rate $\alpha$, and fixing the transition rates between observed states to be independent of the hidden state (rates $\rho_I,\ \rho_C,\ \delta,\ q_{IC}$ as defined for the $ID/CD/CP$ model).
+There are additionally twelve diversification rate parameters ($\lambda_{ID_A},\ \lambda_{ID_B},\ \lambda_{CD_A},\ \lambda_{CD_B},\ \lambda_{CP_A},\ \lambda_{CP_B},\ \mu_{ID_A},\ \mu_{ID_B},\ \mu_{CD_A},\ \mu_{CD_B},\ \mu_{CP_A},\ \mu_{CP_B}$).
 
@@
 
-\subsection{Statistical Inference under the models}
+\subsection{Statistical inference under the models}
 

### 2019-02-15 manuscript/methods.tex
done editing Methods

@@
 To investigate the association between ploidy level and diversification, we first defined a binary state speciation and extinction model (BiSSE, \citealt{maddison_2007}) in which taxa were classified as diploid (D) or polyploid (P) (\cref{figure:stateclassifications}).
-We call this the $D/P$ ploidy model. 
+We call this the D/P ploidy model. 
 In a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$), net diversification rates ($r_D=\lambda_D-\mu_D$, $r_P=\lambda_P-\mu_P$), and relative extinction rates ($\nu_D=\mu_D / \lambda_D$, $\nu_D=\mu_D / \lambda_D$) associated with each state.
@@
 In this model, each of the observed diploid and polyploid states is subdivided by a binary hidden trait with states $A$ and $B$.
-We call this the $D/P-A/B$ ploidy and hidden state model. 
+We call this the D/P--A/B ploidy and hidden state model. 
 We estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$), extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$), net diversification rates ($r_{D_A},\ r_{D_B},\ r_{P_A},\ r_{P_B}$), and relative extinction rates ($\nu_{D_A},\ \nu_{D_B},\ \nu_{P_A},\ \nu_{P_B}$).
@@
 This is the same as the analysis of \citet{goldberg_2010}, though with an updated phylogeny \citep{sarkinen_2013}.
-We call this BiSSE model the $I/C$ breeding system model. 
+We call this BiSSE model the I/C breeding system model. 
 To parse out the effect of breeding system in diversification while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
-We call this HiSSE model the $I/C-A/B$ breeding system and hidden state model.
+We call this HiSSE model the I/C--A/B breeding system and hidden state model.
 
@@
 We did not include a state for self-incompatible polyploids because they are not present in the data, and because that trait combination is logically expected not to occur (as explained above).
-We call this the $ID/CD/CP$ ploidy and breeding system model.
+We call this the ID/CD/CP ploidy and breeding system model.
 It has ten parameters, six for diversification in each state ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and four for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $\delta$ for diploidization from $CP$ to $CD$; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
@@
 
-The $ID/CD/CP$ model could potentially capture similar dynamics as earlier models, if the hidden state in $D/P-A/B$ was effectively breeding system, and the hidden state in $I/C-A/B$ was effectively ploidy.
+The ID/CD/CP model could potentially capture similar dynamics as earlier models, if the hidden state in D/P--A/B was effectively breeding system, and the hidden state in I/C--A/B was effectively ploidy.
 There is also the potential, however, for a hidden factor to be influencing diversification even beyond both of our focal traits, and this could again mislead inferences.
 We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018, herrera_2018, huang_2018}.
-We refer to this as the $ID/CD/CP-A/B$ model.
+We refer to this as the ID/CD/CP--A/B model.
 A fully parameterized version of this model would have 26 rate parameters \citep{herrera_2018}. 
 Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fitted a simplified version with 16 parameters.
-The reduction in parameter space is achieved by fixing the rates for transitions among hidden states to be equal with rate $\alpha$, and fixing the transition rates between observed states to be independent of the hidden state (rates $\rho_I,\ \rho_C,\ \delta,\ q_{IC}$ as defined for the $ID/CD/CP$ model).
+The reduction in parameter space is achieved by fixing the rates for transitions among hidden states to be equal with rate $\alpha$, and fixing the transition rates between observed states to be independent of the hidden state (rates $\rho_I,\ \rho_C,\ \delta,\ q_{IC}$ as defined for the ID/CD/CP model).
 There are additionally twelve diversification rate parameters ($\lambda_{ID_A},\ \lambda_{ID_B},\ \lambda_{CD_A},\ \lambda_{CD_B},\ \lambda_{CP_A},\ \lambda_{CP_B},\ \mu_{ID_A},\ \mu_{ID_B},\ \mu_{CD_A},\ \mu_{CD_B},\ \mu_{CP_A},\ \mu_{CP_B}$).
@@
 
-For all the four models that consider ploidy changes we assumed that diploidization could happen.
-Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like ChromEvol \citep{mayrose_2010}.
+For all four models that consider ploidy changes, we allowed that diploidization could happen.
+Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010}.
 These types of classifications do not allow for a ploidy reversion.
 Where indicated, the classification of ploidy for the data used in our models was based on chromosome multiplicity at the genus level.
-However, the majority of the ploidy classifications were adopted from original studies with alternative sources of information (i.e.
-geographic distribution, genus ploidy distribution) where ploidy was defined by authors that found evidence for it.
-Since it is not clear whether diploidization can be detected under alternative ploidy classifications or even classifications based on chromosome number multiplicity at the genus level, we also fitted the models without diploidization in order to test  whether including an assumption of diploidization is sensitive.
-As discussed by \citet{servedio_2014}  the presence or absence of a hypothesis can have an exploratory goal.
-In our case the diploidization parameter (or the lack of denoted by no $\delta$) in our models  is an opportunity to explore an assumption that might be important but it is not the single definitive process to understand the interactions among polyploidy, breeding system, and diversification.
+% ref data/sources table?
+However, the majority of the ploidy classifications were adopted from original studies with alternative sources of information (\eg geographic distribution, genus ploidy distribution) where ploidy was defined by authors that found evidence for it.
+Since it is not clear whether diploidization can be detected under alternative ploidy classifications or even classifications based on chromosome number multiplicity at the genus level, we also fit the models without diploidization in order to test  whether the conclusions about diversification are sensitive to including diploidization.
+As discussed by \citet{servedio_2014}, the presence or absence of a hypothesis can have an exploratory goal.
+In our case the diploidization parameter (or its absence, $\delta=0$) in our models is an opportunity to explore an assumption that might be important but that is not the single definitive process to understand the interactions among polyploidy, breeding system, and diversification.
 
@@
 
-Parameters for each of the ten diversification models described previously were performed using custom code in RevBayes \citep{hoehna_2016} environment.
-Code and results for all analyses is freely available at https://github.com/roszenil/solploidy .
-For all analyses, a correction for sampling bias is included assuming that Solanaceae family has approximately 3,000 species ($s=595/3000$) as estimated by the Solanaceae Source project \citep{solsource}.
-For all ten models, we assumed that speciation and extinction parameters for all models  had log-normal prior distributions with means equal the expected net diversification rate $ ( \frac{(\textrm{number of taxa})/2}{\textrm{root age}})$ and standard deviation $0.5$.
-Prior distributions for parameters defining trait changes were  assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
-For each model, an MCMC was performed for 96 hours in the high-performance computational cluster at Minnesota Supercomputing Institute which allowed for 5,000 generation of burn-in and a minimum of 200,000 generations of MCMC for each of the 6 models.
-For each model, convergence and mixing of the MCMC was tested using R package coda and software Tracer (see supplementary information for convergence plots).
+Parameters for each of the ten diversification models were performed using custom code in the RevBayes \citep{hoehna_2016} environment.
+Code and results for all analyses is available at \url{https://github.com/roszenil/solploidy}.
+For all analyses, a correction for incomplete sampling was included, based on assuming that the Solanaceae family has approximately 3,000 species ($s=595/3000$) as estimated by the Solanaceae Source project \citep{solsource}.
+For all ten models, we assumed that speciation and extinction parameters  had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
+Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
+For each model, an MCMC chain was run for 96 hours in the high-performance computational cluster at Minnesota Supercomputing Institute, which allowed for 5,000 generation of burn-in and a minimum of 200,000 generations of MCMC for each of the 6 models.
+For each model, convergence and mixing of the MCMC was tested using the R package coda and the software Tracer (see supplementary information for convergence plots).
+% eventually, add supp info fig ref
 
@@
 
-We calculated the marginal likelihoods  for each of the ten models using custom code in RevBayes \citet{hoehna_2016}.
-Marginal likelihoods were calculated using 50  stepping stone steps using the methodology of \citet{xie_2010}.
-Each stepping stone step was found by calculating 500 generations of burn-in followed by a total of 1000 MCMC steps \cref{table:marginallike}.
-The calculation of each marginal likelihood ran for 24 hours in the high-performance computational cluster at Minnesota Supercomputing Institute.\newline
+We calculated the marginal likelihood for each of the ten models in RevBayes \citep{hoehna_2016}.
+Marginal likelihoods were calculated using 50 stepping stone steps under the methodology of \citet{xie_2010}.
+Each stepping stone step was found by calculating 500 generations of burn-in followed by a total of 1000 MCMC steps (\cref{table:marginallike}).
+The calculation of each marginal likelihood ran for 24 hours on a high-performance computational cluster.
 
-Using the marginal likelihood values we calculated fourteen different Bayes factors in log-scale.
-The first six Bayes factors compared the four models of polyploidy against each other, followed by one single comparison between the breeding system models, and finalizing with four comparisons between each polyploidy and breeding system models (see \cref{table:bayesfactors} for all comparisons).
-Comparison between polyploidy models vs.
-polyploidy and breeding system models are not valid for two reasons: first, the input data in each model is numerically different (see \cref{table:stateclassifications} for data classifications)
+Using the marginal likelihood values, we calculated thirteen different Bayes factors. % E: it said "fourteen" but I only count thirteen?
+Six compared the models of polyploidy against each other (D/P and D/P--A/B, each with or without diploidization), one compared the breeding system models (I/C and I/C--A/B), and six compared the models with both traits (ID/CD/CP and ID/CD/CP--A/B, each with or without diploidization) (\cref{table:bayesfactors}).
+%
+Other comparisons between these models are not valid because the input data are different under the different state space codings (\cref{figure:stateclassifications}).
+In essence, the D/P, I/C, and ID/CD/CP state spaces are not lumpable with respect to one another \citep{tarasov_2019}.
 %R: Emma this is where the figure can be important to show that we have to make different input
-; second, even when it is possible to re-write the polyploidy and diversification models to include a hidden state that would account for breeding system allowing to use the same input data, the re-written model would still not be comparable to the multivariate SSE models proposed for polyploidy and breeding system due to the hierarchical structure imposed by the interaction between breeding system and polyploidy.
-This second point has been recently denoted by \citet{tarasov_2019} as the lack of lumpability in Markov models for discrete traits.
-In our particular case, polyploids cannot be self compatible, and the irreversibility of self-compatible to self-incompatible state makes the three-state Markov models non-lumpable into two-state models ultimately making  multivariate and the bivariate models incomparable.
+% E: I think this short explanation is clear enough, especially with the figure.  I'm not sure whether the hierarchical concept is exactly applicable here.
 
@@
 The number of taxa in the the sample was maximized by including tips with only ploidy or only breeding system and assigned them as uncertain in the unknown character.}
-% E: but P are not SI
 %\label{table:stateclassifications}
 %\end{table}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{states.pdf}
+\caption{
+Character states used for each of the models.
+Each species retained on the tree belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known.
+The number of species in each is shown under the corresponding circles in the top row.
+These categories were then grouped in a manner appropriate to the states of each model.
+For example, there are 34 species that are self-compatible and of unknown ploidy; these are coded as either $D$ or $P$ in the D/P models (uncertain, or consistent with either state), as $C$ in the I/C models, and as either $CD$ or $CP$ in the ID/CD/CP models.
+In all cases, species were coded as either $A$ or $B$ in the hidden state models.
+}
+\label{figure:stateclassifications}
+\end{figure}
 
@@
 \caption{Marginal likelihoods for the ten diversification models proposed.}
-% E: but P are not SI
 \label{table:marginallike}
@@
  \caption{Bayes factors in log scale. We compare every possible pair. Number models as indicated in  \cref{table:marginallike}. }
-% E: but P are not SI
 \label{table:bayesfactors}

### 2019-02-15 manuscript/introduction.tex
- minor edits while reading methods - comments in introduction

@@
 \section{Introduction}
--Studying diversification linked to trait evolution\newline
-Here I need help (perhaps will be the last line I write)\newline
-In these context, botanist have been asking how polyploidy shapes patterns of diversification.
+Studying diversification linked to trait evolution\newline
+% Here I need help (perhaps will be the last line I write)
+%In these context, botanist have been asking how polyploidy shapes patterns of diversification. 
+% B: I would start this off with:
+%(1) general intro to macroevolutionary inference, 
+%(2) intro to two traits that we reasonably perceive as having the potential to influence speciation and extinction rates, and 
+%(3) the causal relationship between D/P and SI/SC transitions, which influences their mutual dependence and possibly conflates inference of individual effects on speciation and extinction rates.
 The prevalence of polyploidy and its detection across multiple and highly diverse clades of angiosperms inevitably lead to the hypothesis of the importance of polyploidy in the speciation and extinction patterns observed across the flowering plant phylogeny.

### 2019-02-15 manuscript/methods.tex
- minor edits while reading methods - comments in introduction

--- 2019-02-14 manuscript/methods.tex
@@
 Chromosome number data were obtained for all Solanaceae taxa in the the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were cleaned semi-automatically using the CCDBcurator R package \citep{rivero_2019}.
+% what does "cleaned" mean?
 This large dataset includes the compilation of Solanaceae ploidy states from \citet{robertson_2011}.
@@
 %
-Breeding system was scored as self-incompatible (I) or self-compatible (C) based on results hand-curated from the literature \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
+Breeding system was scored as self-incompatible (I) or self-compatible (C) based on results curated from the literature and  original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
 Most species could unambiguously be coded as either I or C \citep{raduski_2012}.
@@
 
+% B: It would be nice to clarify what was added--what is new--and separate the subsections of ploidy and breeding system data.
+%
 % E: I am trying to stick with I/C when talking about the state values, but using SI/SC when talking about the meaning of the trait (because this what people use normally).
@@
 To those existing data sets, we added some additional records for chromosome number and breeding system.
-The Supplementary Information contains citations for the numerous original sources for all of the data. % todo: add number of refs when we have it
-Synonymy followed Solanaceae Source \citep{solsource}.
-Hybrids and cultivars were excluded because ploidy and breeding system are likely to be altered by domestication.
-As in \citet{robertson_2011}, we examined closely the few species for which the merged ploidy and breeding system data indicated the presence of self-incompatible polyploids.
-Although SI populations frequently contain some SC individuals and diploid populations frequently contain some polyploid individuals, in no case did we find individuals that were both SI and polyploid.
+The Supplementary Information contains citations for the numerous sources for the added data. % todo: add number of refs when we have it
+Resolution of taxonomic synonymy followed the conventions provided in Solanaceae Source \citep{solsource}. %B: was Synonymy followed Solanaceae Source \citep{solsource}.
+Hybrids and cultivars were excluded because ploidy and breeding system can be affected by artificial selection during domestication.
+%B: was: Hybrids and cultivars were excluded because ploidy and breeding system are likely to be altered by domestication.
+Following the reasoning outlined in \citet{robertson_2011}, we examined closely the few species for which the merged ploidy and breeding system data indicated the presence of self-incompatible polyploids.
+Although SI populations frequently contain some SC individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find a convincing case of a naturally occurring SI and polyploid population.
+%B: Thus qualified, as above, I think this should be inserted for maximum Truth.
+The single instance of an SI and polyploid individual appears to be an allopentaploid hybrid of {\em Solanum oplocense} Hawkes x {\em Solanum gourlayii} Hawkes, reported by \citealt{camadro1981}.
+
 Because of this empirical observation and the functional explanation for whole genome duplication disabling gametophytic self-incompatibility \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).

### 2019-02-15 manuscript/results.tex

@@
 \begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{states.pdf}
-  \caption{Binary and three state classifications for 595 taxa with ploidy and/or breeding system data.
-The number of taxa in the the sample was maximized by including tips with only ploidy or only breeding system and assigned them as uncertain in some of the models.
-For example, the 34 self-compatible taxa with unknown ploidy were assigned as uncertain (diploid or polyploid) in the D/P models but as fully compatible in the I/C models.}
-\label{figure:stateclassifications}
-\end{figure}
-
-\begin{figure}
 \includegraphics[width=\textwidth]{Netdiversificationallmodels.pdf}

### 2019-02-15 manuscript/methods.tex
- something was messed up with references (minimum needed fix; hoehna_2016, soltis_2014 missing fields) - if someone wishes to git remove main.pdf, that would be okay

@@
 %B: Thus qualified, as above, I think this should be inserted for maximum Truth.
-The single instance of an SI and polyploid individual appears to be an allopentaploid hybrid of {\em Solanum oplocense} Hawkes x {\em Solanum gourlayii} Hawkes, reported by \citealt{camadro1981}.
+The single instance of an SI and polyploid individual appears to be an allopentaploid hybrid of {\em Solanum oplocense} Hawkes x {\em Solanum gourlayii} Hawkes, reported by \citealt{camadro_1981}.
 

### 2019-02-15 manuscript/methods.tex
- methods.tex: Data section edits

--- 2019-02-15 manuscript/methods.tex
@@
 
-% B: It would be nice to clarify what was added--what is new--and separate the subsections of ploidy and breeding system data.
+%B: It would be nice to clarify what was added--what is new--and separate the subsections of ploidy and breeding system data.
 %
@@
 Although SI populations frequently contain some SC individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find a convincing case of a naturally occurring SI and polyploid population.
-%B: Thus qualified, as above, I think this should be inserted for maximum Truth.
-The single instance of an SI and polyploid individual appears to be an allopentaploid hybrid of {\em Solanum oplocense} Hawkes x {\em Solanum gourlayii} Hawkes, reported by \citealt{camadro1981}.
-
-Because of this empirical observation and the functional explanation for whole genome duplication disabling gametophytic self-incompatibility \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).
-
-% E: Should we include more details about how many polymorphic species and how the particularly tricky ones were dealt with?
+%B: Thus qualified, as above, I think this should be inserted.
+The single instance of an SI and polyploid individual appears to be an allopentaploid hybrid of {\em Solanum oplocense} Hawkes x {\em Solanum gourlayii} Hawkes, reported by \citet{camadro_1981}.
+Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain SI, so long as they express a single allele at the S-locus (discussed in \citealt{robertson_2011}).
+Because of the resulting absence of SI and polyploid populations, as well as the linked functional explanation for disabling of gametophytic self-incompatibility systems with non-self recognition, following whole genome duplication \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).
+%B: I added the non-self recognition bit, because in Prunus, heteroallelic pollen may not cause SC.
+% E: Should we include more details about how many polymorphic species and how the particularly tricky ones were dealt with? %B: No and yes, maybe. Interested parties can go to data for part 1?
 
@@
 A total of 405 taxa without any information about breeding system or polyploidy were excluded.
-Tips without trait data are much less informative for diversification parameters linked to trait values, and including this many more species would have prohibitively slowed our analyses, expecially for the most complex models.
+Tips without trait data are much less informative for diversification parameters linked to trait values.
+Including this many more species would have prohibitively slowed our analyses, especially those implementing the most complex models.
 

### 2019-02-16 manuscript/methods.tex
- methods.tex edits down to Diploidization subsection

@@
 
-Chromosome number data were obtained for all Solanaceae taxa in the the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were cleaned semi-automatically using the CCDBcurator R package \citep{rivero_2019}.
-% what does "cleaned" mean?
+%B: IMO: The data subsection would be better organized with parts, (1) ploidy, (2) breeding system, (3) tree, (4) models, (5) model selection and statistical inference.
+
+Chromosome number data were obtained for all Solanaceae taxa in the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were cleaned semi-automatically using the CCDBcurator R package \citep{rivero_2019}.
+%B: perhaps "cleaned" should be explained, unless it's hard to identically repeat as-is.
 This large dataset includes the compilation of Solanaceae ploidy states from \citet{robertson_2011}.
@@
 The Supplementary Information contains citations for the numerous sources for the added data. % todo: add number of refs when we have it
-Resolution of taxonomic synonymy followed the conventions provided in Solanaceae Source \citep{solsource}. %B: was Synonymy followed Solanaceae Source \citep{solsource}.
+Resolution of taxonomic synonymy followed the conventions provided in Solanaceae Source \citep{solsource}. 
 Hybrids and cultivars were excluded because ploidy and breeding system can be affected by artificial selection during domestication.
-%B: was: Hybrids and cultivars were excluded because ploidy and breeding system are likely to be altered by domestication.
 Following the reasoning outlined in \citet{robertson_2011}, we examined closely the few species for which the merged ploidy and breeding system data indicated the presence of self-incompatible polyploids.
@@
 To investigate the association between ploidy level and diversification, we first defined a binary state speciation and extinction model (BiSSE, \citealt{maddison_2007}) in which taxa were classified as diploid (D) or polyploid (P) (\cref{figure:stateclassifications}).
-We call this the D/P ploidy model. 
+We call this the \textit{D/P ploidy} model. %B: emphasis on introduced term (quotes awkward)
 In a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$), net diversification rates ($r_D=\lambda_D-\mu_D$, $r_P=\lambda_P-\mu_P$), and relative extinction rates ($\nu_D=\mu_D / \lambda_D$, $\nu_D=\mu_D / \lambda_D$) associated with each state.
@@
 % E: presumably the motivation will be explained in the Intro
+%B: should the term above be "re-diploidization" or "diploidization" (as now)?
 
-Our second model assesses the the signal of diversification due to ploidy while also parsing out the heterogeneity of diversification rates due a possible unobserved trait.
+Our second model assesses the signal of diversification due to ploidy differences while also parsing out the heterogeneity of diversification rates due a possible unobserved trait.
 BiSSE-like models can suffer from a large false discovery rate because they fail to account for diversification rate changes that do not directly depend on the trait of interest \citep{rabosky_2015, beaulieu_2016}.
-To allow for diversification rate differences explained by something other than ploidy, we added a hidden state (HiSSE model; \citealt{beaulieu_2016}).
+Diversification rate differences explained by something (trait) other than ploidy, are accommodated by adding a hidden state (HiSSE model; \citealt{beaulieu_2016}).
+%B: inserted "(trait)", because that seems to be the goal - a biological explanation?
 In this model, each of the observed diploid and polyploid states is subdivided by a binary hidden trait with states $A$ and $B$.
-We call this the D/P--A/B ploidy and hidden state model. 
+We call this the \textit{D/P--A/B ploidy and hidden state} model. 
 We estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$), extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$), net diversification rates ($r_{D_A},\ r_{D_B},\ r_{P_A},\ r_{P_B}$), and relative extinction rates ($\nu_{D_A},\ \nu_{D_B},\ \nu_{P_A},\ \nu_{P_B}$).
 In this model polyploidization rate $\rho$ and diploidization rate $\delta$ are also included, and changes between hidden states are symmetrical with rate $\alpha$.
+%B: same dip vs.
+re-dip question, as above.
 
@@
 To assess the effects of breeding system in the diversification process, we first fit model in which the states are self-incompatible (I) or self-compatible (C).
-This is the same as the analysis of \citet{goldberg_2010}, though with an updated phylogeny \citep{sarkinen_2013}.
-We call this BiSSE model the I/C breeding system model. 
-To parse out the effect of breeding system in diversification while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
-We call this HiSSE model the I/C--A/B breeding system and hidden state model.
+This is the same as the analysis of \citet{goldberg_2010}, save for an updated phylogeny \citep{sarkinen_2013}.
+We call this BiSSE model the \textit{I/C breeding system} model. 
+To parse out the effect of breeding system on diversification, while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
+We call this HiSSE model the \textit{I/C--A/B breeding system and hidden state model}.
 
-For these and the subsequent models including breeding system, we allow transitions from $I$ to $C$ (at rate $q_{IC}$) but not the reverse.
-Within Solanaceae, self-incompatibility in species that possess it is a homologous form of gametophytic SI \citep[shared even shared even with other dicot families;][]{igic_2001, steinbachs_2002}.
-Furthermore, species that are distantly related within the family carry several closely-related alleles at the locus that controls the SI response \citep{ioerger_1990, igic_2006}.
-This is very strong evidence that the SI mechanism is ancestral in Solanaceae and has not re-evolved within the family.
+%B: we moved back from "SI" to "self-incompatibility" below. Maybe a no-no.
+For all breeding system models, we allow transitions from $I$ to $C$ (at rate $q_{IC}$) but not the reverse. %B: was "these and the subsequent models including breeding systems"
+Within Solanaceae, self-incompatibility is homologous in all species in which S-alleles were cloned, and controlled crosses performed.  
+All species sampled to date, possess a non-self recognition, RNase-based, gametophytic self-incompatibility \citep[shared even shared even with other euasterid families;][]{ramanauskas2017}.
+%B: this now makes a narrower claim than igic_2001/steinbachs_2002, which includes Rosaceae and a bag of avoidable troubles.
+Furthermore, species that are distantly related within this family carry closely-related alleles, with deep trans-specific polymorphism, at the S-locus, which controls the SI response \citep{ioerger_1990, igic_2006}.
+%B: may need to introduce "S-locus" earlier.
+This represents very strong evidence that the SI mechanism, and our $I$ state is ancestral to the Solanaceae, and did not arise independently within the family ($q_{CI}=0$).
+%B: I think I have this right <- someone please double-check.
 % This is an important assumption because ancestral state reconstruction in the models might differ when polyploidy and breeding system are analyzed as independent in the Solanaceae tree.
@@
 
+%B: What happens if SC->SI is not fixed? Asking for a friend.
+
 \subsection{Models for ploidy, breeding system, and diversification}
 
-If both ploidy and breeding system have the potential to influence lineage diversification, we should logically consider their effects jointly.
-We thus fit a model that includes both traits \citep[MuSSE,][]{fitzjohn_2012}.
+If ploidy and breeding system each influence lineage diversification individually, it is logical to examine their possible joint effects.
+%B: unsure about edit, which was: "If both ploidy and breeding system have the potential to influence lineage diversification, we should logically consider their effects jointly."
+We thus fit a multi-state model that includes both traits \citep[MuSSE,][]{fitzjohn_2012}.
 The three states in this model are self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids, which are always self-compatible (CP).
-We did not include a state for self-incompatible polyploids because they are not present in the data, and because that trait combination is logically expected not to occur (as explained above).
-We call this the ID/CD/CP ploidy and breeding system model.
-It has ten parameters, six for diversification in each state ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and four for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $\delta$ for diploidization from $CP$ to $CD$; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
+As explained above, we did not include a state for self-incompatible polyploids because they are not observed in the data, and   that trait combination state is mechanistically predicted not to occur.
+We call this the \textit{ID/CD/CP ploidy and breeding system} model.
+The model has ten parameters, six for diversification in each state ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and four for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $\delta$ for diploidization from $CP$ to $CD$; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
 The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
-Diploidization from $CP$ to $ID$ is not allowed because it would be a regain of self-incompatibility.
+Diploidization from $CP$ to $ID$ is not allowed because it would represent a simultaneous regain of SI.
 
-The ID/CD/CP model could potentially capture similar dynamics as earlier models, if the hidden state in D/P--A/B was effectively breeding system, and the hidden state in I/C--A/B was effectively ploidy.
-There is also the potential, however, for a hidden factor to be influencing diversification even beyond both of our focal traits, and this could again mislead inferences.
+The ID/CD/CP model could potentially capture similar dynamics as earlier models, if the effects of the hidden state in D/P--A/B were effectively caused by breeding system (or its correlates), and the hidden state in I/C--A/B was effectively caused by ploidy.
+%B: you may wish to get rid of "(or its correlates)", which I inserted; along with "caused".
+I know that causation is not strict here, but it clarifies the intended biological explanation.
+%B: was: "if the hidden state in D/P--A/B was effectively breeding system, and the hidden state in I/C--A/B was effectively ploidy."
+There is also the potential, however, for a hidden factor to be influencing diversification beyond both of our focal traits, and this could again mislead inferences.
 We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018, herrera_2018, huang_2018}.
-We refer to this as the ID/CD/CP--A/B model.
+We refer to this as the \textit{ID/CD/CP--A/B} model.
 A fully parameterized version of this model would have 26 rate parameters \citep{herrera_2018}. 
@@
 There are additionally twelve diversification rate parameters ($\lambda_{ID_A},\ \lambda_{ID_B},\ \lambda_{CD_A},\ \lambda_{CD_B},\ \lambda_{CP_A},\ \lambda_{CP_B},\ \mu_{ID_A},\ \mu_{ID_B},\ \mu_{CD_A},\ \mu_{CD_B},\ \mu_{CP_A},\ \mu_{CP_B}$).
+%B: I can't tell, but it seems to me that this 26->16 parameter reduction makes it possible to underestimate the effect of these hidden states.
+Perhaps this complication is covered in the Discussion section.
 
@@
 %\caption{Binary and three state classifications for 595 taxa with ploidy and/or breeding system data.
-The number of taxa in the the sample was maximized by including tips with only ploidy or only breeding system and assigned them as uncertain in the unknown character.}
+The number of taxa in the sample was maximized by including tips with only ploidy or only breeding system and assigned them as uncertain in the unknown character.}
 %\label{table:stateclassifications}

### 2019-02-16 manuscript/methods.tex
- %B done with this iteration of methods.tex

@@
 We call this the \textit{ID/CD/CP ploidy and breeding system} model.
-The model has ten parameters, six for diversification in each state ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and four for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $\delta$ for diploidization from $CP$ to $CD$; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
+The model has 10 parameters, six for diversification in each state ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and four for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $\delta$ for diploidization from $CP$ to $CD$; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
 The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
@@
 
-For all four models that consider ploidy changes, we allowed that diploidization could happen.
+%B: this section needs a motivation.
+Right now, that appears to be hung on an argument from a citation (mayrose_2010), which is not clarified.
+I think you guys may have the paleo-WGD literature in mind here.
+If so, perhaps simply stating something ought to be written about the expectations for capturing some underlying process we think is ongoing.
+For all four models that consider ploidy changes, we allowed diploidization.
 Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010}.
@@
 
-Parameters for each of the ten diversification models were performed using custom code in the RevBayes \citep{hoehna_2016} environment.
-Code and results for all analyses is available at \url{https://github.com/roszenil/solploidy}.
-For all analyses, a correction for incomplete sampling was included, based on assuming that the Solanaceae family has approximately 3,000 species ($s=595/3000$) as estimated by the Solanaceae Source project \citep{solsource}.
-For all ten models, we assumed that speciation and extinction parameters  had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
+Parameters for each of the 10 diversification models were performed using custom code in the RevBayes \citep{hoehna_2016} environment.
+%B: fixme: I think the intent here is  "Parameter estimation [...] was peformed"
+Code for analyses and key results is available at \url{https://github.com/roszenil/solploidy}.
+We included a correction for incomplete sampling in all analyses, based on assuming that the Solanaceae family has approximately 3,000 species ($s=595/3000$) as estimated by the Solanaceae Source project \citep{solsource}.
+For all 10 models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
 Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
-For each model, an MCMC chain was run for 96 hours in the high-performance computational cluster at Minnesota Supercomputing Institute, which allowed for 5,000 generation of burn-in and a minimum of 200,000 generations of MCMC for each of the 6 models.
-For each model, convergence and mixing of the MCMC was tested using the R package coda and the software Tracer (see supplementary information for convergence plots).
-% eventually, add supp info fig ref
+For each model, an MCMC chain was run for 96 hours in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the 6 models.
+%B: the number of models changes here?
+For each model, convergence and mixing of the MCMC was tested using the R library \texttt{coda} and the software package Tracer (see supplementary information for convergence plots).
+% fixme: eventually, add supp info fig ref
 
@@
 
-We calculated the marginal likelihood for each of the ten models in RevBayes \citep{hoehna_2016}.
+We calculated the marginal likelihood for each of the 10 models in RevBayes \citep{hoehna_2016}.
 Marginal likelihoods were calculated using 50 stepping stone steps under the methodology of \citet{xie_2010}.
-Each stepping stone step was found by calculating 500 generations of burn-in followed by a total of 1000 MCMC steps (\cref{table:marginallike}).
+Each stepping stone step was found by calculating 500 generations of burn-in followed by a total of 1,000 MCMC steps (\cref{table:marginallike}).
 The calculation of each marginal likelihood ran for 24 hours on a high-performance computational cluster.
 
-Using the marginal likelihood values, we calculated thirteen different Bayes factors. % E: it said "fourteen" but I only count thirteen?
+Using the marginal likelihood values, we calculated thirteen different Bayes factors. % E: it said "fourteen" but I only count thirteen? %B: fixme: my count too.
 Six compared the models of polyploidy against each other (D/P and D/P--A/B, each with or without diploidization), one compared the breeding system models (I/C and I/C--A/B), and six compared the models with both traits (ID/CD/CP and ID/CD/CP--A/B, each with or without diploidization) (\cref{table:bayesfactors}).
@@
 % E: I think this short explanation is clear enough, especially with the figure.  I'm not sure whether the hierarchical concept is exactly applicable here.
+%B: my TeX GUI (and my brain) underlined "lumpable", especially in the context of "with respect to one another." I think I get the gist, but I am not sure that I agree others will easily.
+I find it useful to imagine Spencer reading this, for example.
 
@@
 \end{tabular}
-\caption{Marginal likelihoods for the ten diversification models proposed.}
+\caption{Marginal likelihoods for the 10 diversification models proposed.}
 \label{table:marginallike}

### 2019-02-16 manuscript/methods.tex
- minor changes (typos) noted at compilation - added ramanauskas2017 to replace igic2001

@@
 Within Solanaceae, self-incompatibility is homologous in all species in which S-alleles were cloned, and controlled crosses performed.  
-All species sampled to date, possess a non-self recognition, RNase-based, gametophytic self-incompatibility \citep[shared even shared even with other euasterid families;][]{ramanauskas2017}.
+All species sampled to date, possess a non-self recognition, RNase-based, gametophytic self-incompatibility \citep[shared even shared even with other euasterid families;][]{ramanauskas_2017}.
 %B: this now makes a narrower claim than igic_2001/steinbachs_2002, which includes Rosaceae and a bag of avoidable troubles.
@@
 \end{tabular}
-\caption{Marginal likelihoods for the 10 diversification models proposed.}
+\caption{Marginal likelihoods for the ten models examined.
+States encoded with letter D: diploid, P: polyploid, C: self-compatible, I: self-incompatible, A: first state of hidden trait, B: second state of hidden trait, $\delta$: re-diploidization.}
 \label{table:marginallike}
@@
 \toprule
-Polyploidy Models & Breeding System Models & Polyploidy and Breeding System Models \\ \midrule
+Ploidy Models & Breeding System Models & Ploidy and Breeding System Models \\ \midrule
 {\begin{tabular}{lcccc}

### 2019-02-17 manuscript/results.tex
latex code for supplemental figures
updated main figures
- starting paragraph minor comments/changes.
many comments, while talking with Boris
outline for Intro
table skeleton for median rate estimates
initial thoughts for Discussion
notes to include pathways ala robertson2011
misc elsewhere

@@
 
+% E & B:  We just spent awhile poring over the great supp figures of rate estimates, which are essential to grasp the uncertainties.
+But to facilitate quick comparisons across models, could we have a table of the median rate estimates for all models?
+Possibly like the format I've laid out below?
+Or maybe there's a better way.
+
+\begin{supptable}
+    \begin{center}
+    \begin{tabular}{lcccccc}
+                       & D/P & D/P+A/B & I/C & I/C+A/B & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
+        %
+        $r_D$          &     &         & --- & ---     & ---      & ---          \\
+        $r_P$          &     &         & --- & ---     & ---      & ---          \\
+        $r_I$          & --- & ---     &     &         & ---      & ---          \\
+        $r_C$          & --- & ---     &     &         & ---      & ---          \\
+        $r_{ID}$       & --- & ---     & --- & ---     &          &              \\
+        $r_{CD}$       & --- & ---     & --- & ---     &          &              \\
+        $r_{CP}$       & --- & ---     & --- & ---     &          &              \\
+        %
+        $\rho$         &     &         & --- & ---     & ---      & ---          \\
+        $\rho_I$       & --- & ---     & --- & ---     &          &              \\
+        $\rho_C$       & --- & ---     & --- & ---     &          &              \\
+        $\delta$       &     &         &     &         &          &              \\
+        $q_{IC}$       & --- & ---     &     &         &          &              
+    \end{tabular}
+
+    \vspace{12pt}
+    \begin{tabular}{lcccccc}
+                       & D/P & D/P+A/B & I/C & I/C+A/B & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
+        %
+        $r_D$          &     &         & --- & ---     & ---      & ---          \\
+        $r_P$          &     &         & --- & ---     & ---      & ---          \\
+        $r_I$          & --- & ---     &     &         & ---      & ---          \\
+        $r_C$          & --- & ---     &     &         & ---      & ---          \\
+        $r_{ID}$       & --- & ---     & --- & ---     &          &              \\
+        $r_{CD}$       & --- & ---     & --- & ---     &          &              \\
+        $r_{CP}$       & --- & ---     & --- & ---     &          &              \\
+        %
+        $\rho$         &     &         & --- & ---     & ---      & ---          \\
+        $\rho_I$       & --- & ---     & --- & ---     &          &              \\
+        $\rho_C$       & --- & ---     & --- & ---     &          &              \\
+        $q_{IC}$       & --- & ---     &     &         &          &              
+    \end{tabular}
+    \end{center}
+    \caption{median rate estimates for all models (see supp figs for more details and uncertainties); when there are two numbers in one cell, they're for the A and B hidden states; --- means that parameter was not present in the model; upper section is for models with $\delta$ and lower section is for models without}
+    \label{table:estimates}
+\end{supptable}
+
 \begin{figure}
@@
 \end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{netdiversificationallnodip.pdf}
+  \caption{Net diversification rates for all models that do not include diploidization.}  
+\label{figure:netdivnodip}
+\end{figure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{bisseDPposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the D/P polyploidy model} % XXX
+\label{suppfigure:DP}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the D/P no $\delta$ polyploidy model} % XXX
+\label{suppfigure:DPnodip}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{hisseDPposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the D/P-A/B polyploidy model} % XXX
+% fixme: offset plotting error in "relative extinction" panel
+\label{suppfigure:DPAB}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{hisseDPnodipposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the D/P no $\delta$-A/B polyploidy model} % XXX
+\label{suppfigure:DPnodipAB}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{bisseSIposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the I/C breeding system model} % XXX
+\label{suppfigure:IC}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{hisseSInoretposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the I/C-A/B breeding system model} % XXX
+\label{suppfigure:ICAB}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{musseDPSIposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the ID/CD/CP polyploidy and breeding system model} % XXX
+\label{suppfigure:IDCDCP}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the ID/CD/CP no $\delta$ polyploidy and breeding system model} % XXX
+\label{suppfigure:IDCDCPnodip}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{muhisseDPSIposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the ID/CD/CP-A/B polyploidy and breeding system model} % XXX
+\label{suppfigure:IDCDCPAB}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{muhisseDPSInodipposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the ID/CD/CP no $\delta$- A/B polyploidy and breeding system model} % XXX
+\label{suppfigure:IDCDCPnodipAB}
+\end{suppfigure}
+
+% E or R todo: which pathway from SI-D to SC-P is more common, based on rate estimates?  apply procedure from robertson2011 to rates from ID/CD/CP model

### 2019-02-18 manuscript/discussion.tex

\section{Discussion}

% diversification
%     main finding about our traits
%     more generally, not okay to approximate musse with bisse when states are correlated (cf Pyron)

% diploidization: comment
%     effect on inferred diversification

% pathways: compare/contrast with robertson2011
%     \rho_I > \rho_C in all!

% alpha is always low
%     new supp fig showing how A & B are inferred on the tree?  at least for ID/CD/CP+A/B model
%     maybe we can hypothesize factors that underlie the hidden state (geography?)

% issues
%     meaning of diploidization parameter
%     irreversibility SI -> SC assumption

### 2019-02-18 manuscript/main.tex
small tex matters
notes to self, and linebreaks to aid diff
done editing Methods
- minor changes (typos) noted at compilation - added ramanauskas2017 to replace igic2001

@@
 \usepackage{evolution}
-\usepackage{hyperref}
 
@@
 
+% Accumulate ideas for titles here...
+%
+% When your favorite trait is not enough to explain diversification: An example of diversification linked to polyploidy and breeding system in Solanaceae % E: Fun, but I think we need something more formal
+%
+% New comparative approaches for the study of breeding system, polyploidy, and lineage diversification % E: the tentative title given to New Phyt
+%
+% Comparative study of breeding system, ploidy, and lineage diversification % E: too boring?
+%
+% Interacting effects of ploidy and breeding system on lineage diversification % E: but "effects" is too causal
+
 \begin{center}
-    \textbf{When your favorite trait is not enough to explain diversification: An example of diversification linked to polyploidy and breeding system in Solanaceae}
+    \textbf{Interacting associations between ploidy, breeding system, and lineage diversification}
 \end{center}
@@
 \noindent
-J. Gordon Burleigh,%
+J.~Gordon Burleigh,%
 \endnote{University of Florida}
@@
 \noindent
-William A. Freyman,%
+William A.~Freyman,%
 \endnote{23 and me}
@@
 and
-Emma E. Goldberg%
+Emma E.~Goldberg%
 \endnote{University of Minnesota Twin Cities}
@@
 \noindent
-\textit{Running head:} Polyploidy and breeding systems Solanaceae
+\textit{Running head:} Ploidy and breeding systems in Solanaceae
+% E: I've been changing "polyploidy" to "ploidy" in various places because "polyploidy" is one state of the "ploidy" trait, just as "SI" is one state of the "breeding system" trait
 
@@
 
+NSF DEB-1655478.
+The Minnesota Supercomputing Institute (MSI) at the University of Minnesota provided computing resources for this project.
+
 %--------------------------------------------------
@@
 \bibliography{refs}
-\bibliographystyle{evolution}
+\bibliographystyle{evolution} % E: remove URLs
 \setstretch{\stretchby}

### 2019-02-18 manuscript/introduction.tex
- I took over the start of the Intro for my part - (it's right at the top, so you can't miss it) - this is a backup, TBC tomorrow morning (2/18/2019)

--- 2019-02-09 manuscript/introduction.tex
@@
 \section{Introduction}
--Studying diversification linked to trait evolution\newline
-Here I need help (perhaps will be the last line I write)\newline
-In these context, botanist have been asking how polyploidy shapes patterns of diversification.
+
+% Outline, discussed by Boris and Emma:
+%
+% we are interested in the evolution of traits and their influence on lineage diversification
+%     ref Barrett work 
+% 		no diversification?, but trait shifts:
+%		barrett_1996 (mating system + LH)
+%		barrett_2013
+%		barrett_2008: "Because polyploidy affects the entire genome, it is perhaps not surprising that it influences many as- pects of the phenotype, including the mating system.
+However, although it has long been recognized that the evolutionary tran sition from diploidy to polyploidy may result in correlated changes in mating patterns, the theoretical and empirical evidence is limited and often contradictory." (p.5)
+%
+%     any one study focuses on one/few traits
+%     but any real organisms have many, so need to enlarge our view
+%     mini "here, we..."
+% we have tools to try to answer these questions
+%     sse models
+%     issues
+%     but a way forward is to allow for multiple factors that could influence diversification
+% ploidy
+%     Stebbins
+%     Mayrose, Soltis
+%     recent paleo WGD, diploidization
+% breeding system
+%     directionality
+%     diversification in Solanaceae
+%     correlation with ploidy
+%     pathways from SI-D to SC-P: direct or via SC-D
+% here, we...
+%     diversification: interaction between ploidy and breeding system
+%     also: diploidization, pathways
+
+%B: I placed this out of order, so you can see it
+The prevalence and significance of polyploidy was broadly considered a salient feature of flowering plants \citep{}, even before the advent of genomic tools dramatically escalated the stakes, so that nearly flowering plants are thought to have undergone at least one round of polyploidization \citep{}.
+It is therefore perhaps surprising that there appears to be little agreement regarding the evolutionary consequences of polyploidy, especially how it affects diversification rates. %species richness.
+% I think Rosana wants this part here (above)
+% next, continue with ploidy narrative (mayrose/soltis)
+
+Polyploidization changes the genomic content of all cells, and consequently has the potential to affect a wide variety, if not all traits. 
+Among the many of changes associated with polyploidization, perhaps the most prominent is the association between polyploidy and propensity for self-fertilization \citep{stebbins_1950, grant_1956, barrett_1988}.
+Polyploidization is not only suspected of a mere correlated change in breeding systems, but of a causal link.
+By disrupting certain genetic self-incompatibility systems, polyploidy is also associated with the ability to self-fertilize in many plant groups (Miller and Venable, 2000; Mable, 2004; Robertson et al., 2011).
+% It's not Spencer's best line, but it's not bad for our work: Barrett 2008: Because polyploidy affects the entire genome, it is perhaps not surprising that it influences many aspects of the phenotype, including the mating system.
+However, although it has long been recognized that the evolutionary transition from diploidy to polyploidy may result in correlated changes in mating patterns, the theoretical and empirical evidence is limited and often contradictory.
+%
+% how polyploidy breaks down SI
+% how that generates associations (and variable associations of mating system and ploidy! husband2008, others cited by barrett_2013
+
+% Breeding systems also affect diversification rate and so there could be an interaction
+% we also have this great data for SI loci and the transition could be even more common, providing power to resolve the great diversification questions
+% set up pathways
+
+Studying diversification linked to trait evolution\newline
+% Here I need help (perhaps will be the last line I write)
+%In these context, botanist have been asking how polyploidy shapes patterns of diversification. 
 The prevalence of polyploidy and its detection across multiple and highly diverse clades of angiosperms inevitably lead to the hypothesis of the importance of polyploidy in the speciation and extinction patterns observed across the flowering plant phylogeny.
@@
 \newline
-
--What is our proposal to tackle this problem?\newline
-
-- How this paper is structured \newline
+What is our proposal to tackle this problem?\newline
 
@@
 Goldberg and Igic 2012, different  perspectives\\
-
-=

### 2019-02-18 manuscript/introduction.tex
- new paragraph(s) in introduction, and matching citations - new title ideas or modifications

@@
 
-Polyploidization changes the genomic content of all cells, and consequently has the potential to affect a wide variety, if not all traits. 
-Among the many of changes associated with polyploidization, perhaps the most prominent is the association between polyploidy and propensity for self-fertilization \citep{stebbins_1950, grant_1956, barrett_1988}.
-Polyploidization is not only suspected of a mere correlated change in breeding systems, but of a causal link.
-By disrupting certain genetic self-incompatibility systems, polyploidy is also associated with the ability to self-fertilize in many plant groups (Miller and Venable, 2000; Mable, 2004; Robertson et al., 2011).
-% It's not Spencer's best line, but it's not bad for our work: Barrett 2008: Because polyploidy affects the entire genome, it is perhaps not surprising that it influences many aspects of the phenotype, including the mating system.
-However, although it has long been recognized that the evolutionary transition from diploidy to polyploidy may result in correlated changes in mating patterns, the theoretical and empirical evidence is limited and often contradictory.
+Polyploidization changes the genomic content of all cells, and consequently has the potential to affect a wide variety, if not all, traits. 
+Among the many of changes associated with polyploidization, perhaps the most prominent is the association between polyploidy and propensity for self-fertilization \citep{stebbins1950, barrett1988}.
+Polyploidy is not only a suspected correlate of breeding systems but a causal link \citep{stout1942, lewis1947}.
+Doubled number of alleles in pollen is strongly suspected to effect disruption of the genetic mechanisms in gametophytic self-incompatibility systems \citep{entani1999, tsukamoto2005, kubo2010}. 
+Polyploidy is also thought to be associated with changes in the rate of self-fertilization, although the evidence for this correlated shift in mating system appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
+% It's not Spencer's best line, but it's not bad for our work: Barrett 2008: "Because polyploidy affects the entire genome, it is perhaps not surprising that it influences many aspects of the phenotype, including the mating system.
+However, although it has long been recognized that the evolutionary transition from diploidy to polyploidy may result in correlated changes in mating patterns, the theoretical and empirical evidence is limited and often contradictory."
+% how that generates associations (and variable associations of mating system and ploidy! husband2008, others cited by barrett2013
+Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974,barrett2013}.
+They are by themselves thought to have a broad effect on other traits. 
+The frequent transition from self-incompatibility (SI) to self-compatibility (SC) is strongly associated with net diversification rate changes \citep{goldberg_2010,devos2014}.
+Given that  changes in ploidy and breeding systems may be causally related and have profound profound affects on the fate of lineages, it seems particularly profitable to examine whether and how their macroevolutionary effects interact.
 %
-% how polyploidy breaks down SI
-% how that generates associations (and variable associations of mating system and ploidy! husband2008, others cited by barrett_2013
-
-% Breeding systems also affect diversification rate and so there could be an interaction
 % we also have this great data for SI loci and the transition could be even more common, providing power to resolve the great diversification questions
-% set up pathways
+%
+% set up pathway dependence
 

### 2019-02-18 manuscript/methods.tex
ref typo

--- 2019-02-16 manuscript/methods.tex
@@
 
-%B: IMO: The data subsection would be better organized with parts, (1) ploidy, (2) breeding system, (3) tree, (4) models, (5) model selection and statistical inference.
-
-Chromosome number data were obtained for all Solanaceae taxa in the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were cleaned semi-automatically using the CCDBcurator R package \citep{rivero_2019}.
-%B: perhaps "cleaned" should be explained, unless it's hard to identically repeat as-is.
+Chromosome number data were obtained for all Solanaceae taxa in the the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were cleaned semi-automatically using the CCDBcurator R package \citep{rivero_2019}.
+% what does "cleaned" mean?
 This large dataset includes the compilation of Solanaceae ploidy states from \citet{robertson_2011}.
@@
 
-%B: It would be nice to clarify what was added--what is new--and separate the subsections of ploidy and breeding system data.
+% B: It would be nice to clarify what was added--what is new--and separate the subsections of ploidy and breeding system data.
 %
@@
 The Supplementary Information contains citations for the numerous sources for the added data. % todo: add number of refs when we have it
-Resolution of taxonomic synonymy followed the conventions provided in Solanaceae Source \citep{solsource}. 
+Resolution of taxonomic synonymy followed the conventions provided in Solanaceae Source \citep{solsource}. %B: was Synonymy followed Solanaceae Source \citep{solsource}.
 Hybrids and cultivars were excluded because ploidy and breeding system can be affected by artificial selection during domestication.
+%B: was: Hybrids and cultivars were excluded because ploidy and breeding system are likely to be altered by domestication.
 Following the reasoning outlined in \citet{robertson_2011}, we examined closely the few species for which the merged ploidy and breeding system data indicated the presence of self-incompatible polyploids.
 Although SI populations frequently contain some SC individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find a convincing case of a naturally occurring SI and polyploid population.
-%B: Thus qualified, as above, I think this should be inserted.
+%B: Thus qualified, as above, I think this should be inserted for maximum Truth.
 The single instance of an SI and polyploid individual appears to be an allopentaploid hybrid of {\em Solanum oplocense} Hawkes x {\em Solanum gourlayii} Hawkes, reported by \citet{camadro_1981}.
-Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain SI, so long as they express a single allele at the S-locus (discussed in \citealt{robertson_2011}).
-Because of the resulting absence of SI and polyploid populations, as well as the linked functional explanation for disabling of gametophytic self-incompatibility systems with non-self recognition, following whole genome duplication \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).
-%B: I added the non-self recognition bit, because in Prunus, heteroallelic pollen may not cause SC.
-% E: Should we include more details about how many polymorphic species and how the particularly tricky ones were dealt with? %B: No and yes, maybe. Interested parties can go to data for part 1?
+
+Because of this empirical observation and the functional explanation for whole genome duplication disabling gametophytic self-incompatibility \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).
+
+% E: Should we include more details about how many polymorphic species and how the particularly tricky ones were dealt with?
 
@@
 A total of 405 taxa without any information about breeding system or polyploidy were excluded.
-Tips without trait data are much less informative for diversification parameters linked to trait values.
-Including this many more species would have prohibitively slowed our analyses, especially those implementing the most complex models.
+Tips without trait data are much less informative for diversification parameters linked to trait values, and including this many more species would have prohibitively slowed our analyses, expecially for the most complex models.
 
@@
 To investigate the association between ploidy level and diversification, we first defined a binary state speciation and extinction model (BiSSE, \citealt{maddison_2007}) in which taxa were classified as diploid (D) or polyploid (P) (\cref{figure:stateclassifications}).
-We call this the \textit{D/P ploidy} model. %B: emphasis on introduced term (quotes awkward)
+We call this the D/P ploidy model. 
 In a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$), net diversification rates ($r_D=\lambda_D-\mu_D$, $r_P=\lambda_P-\mu_P$), and relative extinction rates ($\nu_D=\mu_D / \lambda_D$, $\nu_D=\mu_D / \lambda_D$) associated with each state.
@@
 % E: presumably the motivation will be explained in the Intro
-%B: should the term above be "re-diploidization" or "diploidization" (as now)?
 
-Our second model assesses the signal of diversification due to ploidy differences while also parsing out the heterogeneity of diversification rates due a possible unobserved trait.
+Our second model assesses the the signal of diversification due to ploidy while also parsing out the heterogeneity of diversification rates due a possible unobserved trait.
 BiSSE-like models can suffer from a large false discovery rate because they fail to account for diversification rate changes that do not directly depend on the trait of interest \citep{rabosky_2015, beaulieu_2016}.
-Diversification rate differences explained by something (trait) other than ploidy, are accommodated by adding a hidden state (HiSSE model; \citealt{beaulieu_2016}).
-%B: inserted "(trait)", because that seems to be the goal - a biological explanation?
+To allow for diversification rate differences explained by something other than ploidy, we added a hidden state (HiSSE model; \citealt{beaulieu_2016}).
 In this model, each of the observed diploid and polyploid states is subdivided by a binary hidden trait with states $A$ and $B$.
-We call this the \textit{D/P--A/B ploidy and hidden state} model. 
+We call this the D/P--A/B ploidy and hidden state model. 
 We estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$), extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$), net diversification rates ($r_{D_A},\ r_{D_B},\ r_{P_A},\ r_{P_B}$), and relative extinction rates ($\nu_{D_A},\ \nu_{D_B},\ \nu_{P_A},\ \nu_{P_B}$).
 In this model polyploidization rate $\rho$ and diploidization rate $\delta$ are also included, and changes between hidden states are symmetrical with rate $\alpha$.
-%B: same dip vs.
-re-dip question, as above.
 
@@
 To assess the effects of breeding system in the diversification process, we first fit model in which the states are self-incompatible (I) or self-compatible (C).
-This is the same as the analysis of \citet{goldberg_2010}, save for an updated phylogeny \citep{sarkinen_2013}.
-We call this BiSSE model the \textit{I/C breeding system} model. 
-To parse out the effect of breeding system on diversification, while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
-We call this HiSSE model the \textit{I/C--A/B breeding system and hidden state model}.
+This is the same as the analysis of \citet{goldberg_2010}, though with an updated phylogeny \citep{sarkinen_2013}.
+We call this BiSSE model the I/C breeding system model. 
+To parse out the effect of breeding system in diversification while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
+We call this HiSSE model the I/C--A/B breeding system and hidden state model.
 
-%B: we moved back from "SI" to "self-incompatibility" below. Maybe a no-no.
-For all breeding system models, we allow transitions from $I$ to $C$ (at rate $q_{IC}$) but not the reverse. %B: was "these and the subsequent models including breeding systems"
-Within Solanaceae, self-incompatibility is homologous in all species in which S-alleles were cloned, and controlled crosses performed.  
-All species sampled to date, possess a non-self recognition, RNase-based, gametophytic self-incompatibility \citep[shared even shared even with other euasterid families;][]{ramanauskas2017}.
-%B: this now makes a narrower claim than igic_2001/steinbachs_2002, which includes Rosaceae and a bag of avoidable troubles.
-Furthermore, species that are distantly related within this family carry closely-related alleles, with deep trans-specific polymorphism, at the S-locus, which controls the SI response \citep{ioerger_1990, igic_2006}.
-%B: may need to introduce "S-locus" earlier.
-This represents very strong evidence that the SI mechanism, and our $I$ state is ancestral to the Solanaceae, and did not arise independently within the family ($q_{CI}=0$).
-%B: I think I have this right <- someone please double-check.
+For these and the subsequent models including breeding system, we allow transitions from $I$ to $C$ (at rate $q_{IC}$) but not the reverse.
+Within Solanaceae, self-incompatibility in species that possess it is a homologous form of gametophytic SI \citep[shared even shared even with other dicot families;][]{igic_2001, steinbachs_2002}.
+Furthermore, species that are distantly related within the family carry several closely-related alleles at the locus that controls the SI response \citep{ioerger_1990, igic_2006}.
+This is very strong evidence that the SI mechanism is ancestral in Solanaceae and has not re-evolved within the family.
 % This is an important assumption because ancestral state reconstruction in the models might differ when polyploidy and breeding system are analyzed as independent in the Solanaceae tree.
@@
 
-%B: What happens if SC->SI is not fixed? Asking for a friend.
-
 \subsection{Models for ploidy, breeding system, and diversification}
 
-If ploidy and breeding system each influence lineage diversification individually, it is logical to examine their possible joint effects.
-%B: unsure about edit, which was: "If both ploidy and breeding system have the potential to influence lineage diversification, we should logically consider their effects jointly."
-We thus fit a multi-state model that includes both traits \citep[MuSSE,][]{fitzjohn_2012}.
+If both ploidy and breeding system have the potential to influence lineage diversification, we should logically consider their effects jointly.
+We thus fit a model that includes both traits \citep[MuSSE,][]{fitzjohn_2012}.
 The three states in this model are self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids, which are always self-compatible (CP).
-As explained above, we did not include a state for self-incompatible polyploids because they are not observed in the data, and   that trait combination state is mechanistically predicted not to occur.
-We call this the \textit{ID/CD/CP ploidy and breeding system} model.
-The model has ten parameters, six for diversification in each state ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and four for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $\delta$ for diploidization from $CP$ to $CD$; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
+We did not include a state for self-incompatible polyploids because they are not present in the data, and because that trait combination is logically expected not to occur (as explained above).
+We call this the ID/CD/CP ploidy and breeding system model.
+It has ten parameters, six for diversification in each state ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and four for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $\delta$ for diploidization from $CP$ to $CD$; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
 The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
-Diploidization from $CP$ to $ID$ is not allowed because it would represent a simultaneous regain of SI.
+Diploidization from $CP$ to $ID$ is not allowed because it would be a regain of self-incompatibility.
 
-The ID/CD/CP model could potentially capture similar dynamics as earlier models, if the effects of the hidden state in D/P--A/B were effectively caused by breeding system (or its correlates), and the hidden state in I/C--A/B was effectively caused by ploidy.
-%B: you may wish to get rid of "(or its correlates)", which I inserted; along with "caused".
-I know that causation is not strict here, but it clarifies the intended biological explanation.
-%B: was: "if the hidden state in D/P--A/B was effectively breeding system, and the hidden state in I/C--A/B was effectively ploidy."
-There is also the potential, however, for a hidden factor to be influencing diversification beyond both of our focal traits, and this could again mislead inferences.
+The ID/CD/CP model could potentially capture similar dynamics as earlier models, if the hidden state in D/P--A/B was effectively breeding system, and the hidden state in I/C--A/B was effectively ploidy.
+There is also the potential, however, for a hidden factor to be influencing diversification even beyond both of our focal traits, and this could again mislead inferences.
 We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018, herrera_2018, huang_2018}.
-We refer to this as the \textit{ID/CD/CP--A/B} model.
+We refer to this as the ID/CD/CP--A/B model.
 A fully parameterized version of this model would have 26 rate parameters \citep{herrera_2018}. 
@@
 There are additionally twelve diversification rate parameters ($\lambda_{ID_A},\ \lambda_{ID_B},\ \lambda_{CD_A},\ \lambda_{CD_B},\ \lambda_{CP_A},\ \lambda_{CP_B},\ \mu_{ID_A},\ \mu_{ID_B},\ \mu_{CD_A},\ \mu_{CD_B},\ \mu_{CP_A},\ \mu_{CP_B}$).
-%B: I can't tell, but it seems to me that this 26->16 parameter reduction makes it possible to underestimate the effect of these hidden states.
-Perhaps this complication is covered in the Discussion section.
 
@@
 %\caption{Binary and three state classifications for 595 taxa with ploidy and/or breeding system data.
-The number of taxa in the sample was maximized by including tips with only ploidy or only breeding system and assigned them as uncertain in the unknown character.}
+The number of taxa in the the sample was maximized by including tips with only ploidy or only breeding system and assigned them as uncertain in the unknown character.}
 %\label{table:stateclassifications}

### 2019-02-18 manuscript/introduction.tex

@@
 The prevalence and significance of polyploidy was broadly considered a salient feature of flowering plants \citep{}, even before the advent of genomic tools dramatically escalated the stakes, so that nearly flowering plants are thought to have undergone at least one round of polyploidization \citep{}.
-It is therefore perhaps surprising that there appears to be little agreement regarding the evolutionary consequences of polyploidy, especially how it affects diversification rates. %species richness.
+It is therefore surprising that there is little agreement regarding the evolutionary consequences of polyploidy, especially how it affects diversification rates. %species richness.
 % I think Rosana wants this part here (above)

### 2019-02-18 manuscript/methods.tex

--- 2019-02-16 manuscript/methods.tex
@@
 %B: Thus qualified, as above, I think this should be inserted.
-The single instance of an SI and polyploid individual appears to be an allopentaploid hybrid of {\em Solanum oplocense} Hawkes x {\em Solanum gourlayii} Hawkes, reported by \citet{camadro_1981}.
+The single instance of an SI and polyploid individual appears to be an allopentaploid hybrid of {\em Solanum oplocense} Hawkes x {\em Solanum gourlayii} Hawkes, reported by \citealt{camadro_1981}.
 Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain SI, so long as they express a single allele at the S-locus (discussed in \citealt{robertson_2011}).
@@
 %B: should the term above be "re-diploidization" or "diploidization" (as now)?
+
+% B (via E): need some explanation of what \delta means, biologically, when ploidy is scored like this
 
@@
 Perhaps this complication is covered in the Discussion section.
+
+% E todo: explain "pathway" method of robertson2011
 
@@
 %\end{tabular}
+
+% E: Could we include a figure of the traits on the tree?  At least as supp info. TODO

### 2019-02-19 manuscript/introduction.tex

@@
 % 		no diversification?, but trait shifts:
-%		barrett_1996 (mating system + LH)
-%		barrett_2013
-%		barrett_2008: "Because polyploidy affects the entire genome, it is perhaps not surprising that it influences many as- pects of the phenotype, including the mating system.
-However, although it has long been recognized that the evolutionary tran sition from diploidy to polyploidy may result in correlated changes in mating patterns, the theoretical and empirical evidence is limited and often contradictory." (p.5)
+%		barrett_2008: "Because polyploidy affects the entire genome, it is perhaps not surprising that it influences many aspects of the phenotype, including the mating system.
+However, although it has long been recognized that the evolutionary transition from diploidy to polyploidy may result in correlated changes in mating patterns, the theoretical and empirical evidence is limited and often contradictory." (p.5)
 %
@@
 
-%B: I placed this out of order, so you can see it
-The prevalence and significance of polyploidy was broadly considered a salient feature of flowering plants \citep{}, even before the advent of genomic tools dramatically escalated the stakes, so that nearly flowering plants are thought to have undergone at least one round of polyploidization \citep{}.
-It is therefore surprising that there is little agreement regarding the evolutionary consequences of polyploidy, especially how it affects diversification rates. %species richness.
-% I think Rosana wants this part here (above)
-% next, continue with ploidy narrative (mayrose/soltis)
+%Studying diversification linked to trait evolution\newline
+%Here I need help (perhaps will be the last line I write).
+In these context, botanist have been asking how polyploidy shapes patterns of diversification.
+The prevalence of polyploidy and its detection across multiple and highly diverse clades of angiosperms inevitably lead to the hypothesis of the importance of polyploidy in the speciation and extinction patterns observed across the flowering plant phylogeny.
+At the same time, a similar question has been asked in the breeding system world, where self-compatibility has evolved multiple times in flowering plants.
+However, the influence in diversification using both polyploidy and self-compatibility information has not been studied simultaneously.
 
-Polyploidization changes the genomic content of all cells, and consequently has the potential to affect a wide variety, if not all, traits. 
+%Why polyploidy and self-compatibility in particular makes for an  interesting study in the context of diversification\newline In the polyploidy world, an important debate regarding the diversification of angiosperms has been ongoing since the publication of \citet{mayrose_2011}.
+The authors discovered that using the latest diversification model linked to two states diploid and polyploid, the net diversification of polyploids was much slower than the net diversification rate than polyploids.
+This result was surprising and an sparked a discussion about the long-term evolutionary consequences of polyploidy.
+\citet{soltis_2014} questioned if polyploidy should be regarded as an evolutionary dead-end since the net diversification of polyploids was negative, despite overwhelming evidence about the incidence of polyploidy  especially at the root of highly diverse angiosperm claims (ref here).
+A year later, diversification models were re-tested and corrected, and still found the same pattern \citep{mayrose_2015}, to the disappointment of plenty of botanists there was no denying of the weak trend of diversification that polyploids left behind.
+The most recent study by \citet{landis_2018} that used not only the presence of polyploidy in the tips but also the number of whole genome duplications in a taxon lineage found that....
+\newline  Meanwhile, studies focusing on diversification patterns and  breeding system have consistently found that self-incompatible plants often have higher net diversification rates compare to their self-compatible counterparts (Emma and Boris' papers here, what about papers that are not solanaceae?)
+
+%- Why other traits need to be considered as well\newline
+%- What other models and studies have done in the past\newline
+%- What is lacking from past approaches? \newline
+%There are two key questions that at the time of the polyploidy debate were difficult to ask.
+The first is if the models used to measure the diversification of polyploids were correct, and the second question is if the models have potentially included more evidence and potential traits that are not polyploidy or other lines of evidence that could be driving the patterns.
+At the time, in a different context \citet{beaulieu_2016} were finding an alternative solution to the first question, coming up with a new model that could represent the broad heterogeneity of the diversification process and parse out the signal between the trait of interest and the noise in diversification.
+Their model, the hidden state speciation and extinction model is a key component to detect whether polyploidy our something else unknown but related to it is driving the speciation and extinction patterns that we see in angiosperms.
+\newline
+
+The prevalence of variation in chromosome number, and especially ploidy, has been broadly considered a salient feature of flowering plants for nearly a century.
+%\citep{stebbins1938}; instead of "ploidy" here, we may wish to use some stand-in for karyotype multiples, and stick to that language throughout.
+The recent dramatic increase in the scale of available genome sequences uncovered ancient rounds of whole-genome duplications, and subsequent re-diploidization, across angiosperms \citep{lynch2000,vision2000}.
+Nearly all lineages of flowering plants are thought to have undergone at least one or two rounds of polyploidization. %cite
+It is therefore surprising that there is little agreement regarding the evolutionary consequences of polyploidy, especially how it affects diversification rates.
+%species richness; cite: mayrose2011, soltis2014, mayrose2015, tank2015, landis2018?
+
+Polyploidization alters the genomic content of all cells, and consequently has the potential to affect a wide variety, if not all, traits. 
+From the very beginning, the study of this cytogenetic property considered correlations with other traits \citep{stebbins1938}. 
 Among the many of changes associated with polyploidization, perhaps the most prominent is the association between polyploidy and propensity for self-fertilization \citep{stebbins1950, barrett1988}.
@@
 Polyploidy is also thought to be associated with changes in the rate of self-fertilization, although the evidence for this correlated shift in mating system appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
-% It's not Spencer's best line, but it's not bad for our work: Barrett 2008: "Because polyploidy affects the entire genome, it is perhaps not surprising that it influences many aspects of the phenotype, including the mating system.
-However, although it has long been recognized that the evolutionary transition from diploidy to polyploidy may result in correlated changes in mating patterns, the theoretical and empirical evidence is limited and often contradictory."
-% how that generates associations (and variable associations of mating system and ploidy! husband2008, others cited by barrett2013
 Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974,barrett2013}.
@@
 
-Studying diversification linked to trait evolution\newline
-% Here I need help (perhaps will be the last line I write)
-%In these context, botanist have been asking how polyploidy shapes patterns of diversification. 
-The prevalence of polyploidy and its detection across multiple and highly diverse clades of angiosperms inevitably lead to the hypothesis of the importance of polyploidy in the speciation and extinction patterns observed across the flowering plant phylogeny.
-At the same time, a similar question has been asked in the breeding system world, where self-compatibility has evolved multiple times in flowering plants. 
-However, the influence in diversification using both polyploidy and self-compatibility information has not been studied simultaneously.
-
-- Why polyploidy and self-compatibility in particular makes for an  interesting study in the context of diversification\newline
-In the polyploidy world, an important debate regarding the diversification of angiosperms has been ongoing since the publication of \citet{mayrose_2011}.
-The authors discovered that using the latest diversification model linked to two states diploid and polyploid, the net diversification of polyploids was much slower than the net diversification rate than polyploids.
-This result was surprising and an sparked a discussion about the long-term evolutionary consequences of polyploidy.
-\citet{soltis_2014} questioned if polyploidy should be regarded as an evolutionary dead-end since the net diversification of polyploids was negative, despite overwhelming evidence about the incidence of polyploidy  especially at the root of highly diverse angiosperm claims (ref here).
-A year later, diversification models were re-tested and corrected, and still found the same pattern \citep{mayrose_2015}, to the disappointment of plenty of botanists there was no denying of the weak trend of diversification that polyploids left behind.
-The most recent study by \citet{landis_2018} that used not only the presence of polyploidy in the tips but also the number of whole genome duplications in a taxon lineage found that....
-\newline  Meanwhile, studies focusing on diversification patterns and  breeding system have consistently found that self-incompatible plants often have higher net diversification rates compare to their self-compatible counterparts (Emma and Boris' papers here, what about papers that are not solanaceae?)
-
-- Why other traits need to be considered as well\newline
-
-- What other models and studies have done in the past\newline
-
-- What is lacking from past approaches? \newline
-There are two key questions that at the time of the polyploidy debate were difficult to ask.
-The first is if the models used to measure the diversification of polyploids were correct, and the second question is if the models have potentially included more evidence and potential traits that are not polyploidy or other lines of evidence that could be driving the patterns.
-At the time, in a different context \citet{beaulieu_2016} were finding an alternative solution to the first question, coming up with a new model that could represent the broad heterogeneity of the diversification process and parse out the signal between the trait of interest and the noise in diversification.
-Their model, the hidden state speciation and extinction model is a key component to detect whether polyploidy our something else unknown but related to it is driving the speciation and extinction patterns that we see in angiosperms.
-\newline
-What is our proposal to tackle this problem?\newline
-
-The second question (talk here about breeding system and polyploidy, and how this could be different across clades and this is one of the reasons why we focused on Solanaceae).
--Diversification and breeding systems\\
-Goldberg and Igic 2012, different  perspectives\\
+%The second question (talk here about breeding system and polyploidy, and how this could be different across clades and this is one of the reasons why we focused on Solanaceae).
+%-Diversification and breeding systems\\
+%Goldberg and Igic 2012, different  perspectives\\

### 2019-02-20 manuscript/introduction.tex
- small intro change
- draft for Emma with everything but "Here, we..." paragraph

@@
 
+%B: this is highly preliminary, for your editing consideration.
+The systematic search for traits which influence the rates at which species accumulate around the tree of life is at the productive confluence of drastic increases in the size of datasets and equally dramatic improvements in analytical methods \citep{maddison_2007,fitzjohn_2009,goldberg_2012,beaulieu_2015}.
+% some papers with big trees.
+Many limitations impede our progress in identifying estimating trait-dependent differences in speciation and extinction rates \citep{maddison_2015, rabosky_2015}. % add Moore/Hoehna?
+Nevertheless, the field has made steady advances in relaxing assumptions and addressing conceptual difficulties. %maybe add beaulieu_2015 here, instead
+One limitation is the pervasive consideration of traits in isolation. 
+For example, studies commonly aim to identify the effects of alternate states of a single trait on the diversification process. 
+This is problematic, because the context in which traits occur can lead to complex interactions, perhaps with other traits that make up an organism or its environment.
+Consequently, an obvious direction for improvement of trait-dependent diversification models ought to involve development of approaches that allow investigation of individual and joint effects of traits on diversification, as well as each other's state transitions.
+Ideally, the initial application of such approaches ought to rely on well-sampled and studied traits, whose interactions are relatively clear, and which show a wide phylogenetic distribution. 
+%B: We can then have a good prior, reliable data, and replicate the studies.
+In this paper, we use two of the best studied traits in flowering plants, ploidy and breeding system, and demonstrate an approach to consider their individual and joint effects on diversification.
+
 The prevalence of variation in chromosome number, and especially ploidy, has been broadly considered a salient feature of flowering plants for nearly a century.
@@
 Polyploidy is also thought to be associated with changes in the rate of self-fertilization, although the evidence for this correlated shift in mating system appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
+
 Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974,barrett2013}.
@@
 The frequent transition from self-incompatibility (SI) to self-compatibility (SC) is strongly associated with net diversification rate changes \citep{goldberg_2010,devos2014}.
-Given that  changes in ploidy and breeding systems may be causally related and have profound profound affects on the fate of lineages, it seems particularly profitable to examine whether and how their macroevolutionary effects interact.
+Given that changes in ploidy and breeding systems may be causally related and have  profound affects on the fate of lineages, it seems particularly profitable to examine whether and how their macroevolutionary effects interact.
+Moreover,  increases in ploidy are known to directly affect breeding system states and present a rare opportunity for the study of paths and path-dependence on trait evolution and diversification.
+For example, we could ask if the ordinary mutation (and fixation) causes more losses of SI than polyploidization, whether more SC polyploids are generated directly, after polyploidization of diploid SI lineages, or through a two-step pathway, involving loss of SC and subsequent polyploidization of a SC lineage.
+And, are diversification rates of polyploid SC lineages different if one or other path was taken? \citet{charlesworth1985} 
 %
-% we also have this great data for SI loci and the transition could be even more common, providing power to resolve the great diversification questions
+%trans-polymorphism: We also have this great data for SI loci and the transition could be even more common, providing power to resolve the great diversification questions
 %
-% set up pathway dependence
+% set up pathway dependence from SI-D to SC-P: direct or via SC-D
+
+Here, we [...]
+% Someone else should do this, once we have the full data.
+% diversification: interaction between ploidy and breeding system
+% 	examine individual effects
+% 	examine joint effects
+% 	with and without HMMs
+% 	allow for re-diploidization or not
+% also: diploidization, pathways
+% 	examine pathways from SI-D to SC-P: direct or via SC-D
 

### 2019-02-20 manuscript/introduction.tex
- fixing introduced bibtex missing/errors - compiled pdf

@@
 %B: this is highly preliminary, for your editing consideration.
-The systematic search for traits which influence the rates at which species accumulate around the tree of life is at the productive confluence of drastic increases in the size of datasets and equally dramatic improvements in analytical methods \citep{maddison_2007,fitzjohn_2009,goldberg_2012,beaulieu_2015}.
+The systematic search for traits which influence the rates at which species accumulate around the tree of life is at the productive confluence of drastic increases in the size of datasets and equally dramatic improvements in analytical methods \citep{maddison_2007,fitzjohn_2009,goldberg_2012,beaulieu_2016}.
 % some papers with big trees.
-Many limitations impede our progress in identifying estimating trait-dependent differences in speciation and extinction rates \citep{maddison_2015, rabosky_2015}. % add Moore/Hoehna?
-Nevertheless, the field has made steady advances in relaxing assumptions and addressing conceptual difficulties. %maybe add beaulieu_2015 here, instead
+Many limitations impede our progress in identifying estimating trait-dependent differences in speciation and extinction rates \citep{maddison_2014, rabosky_2015}. % add Moore/Hoehna?
+Nevertheless, the field has made steady advances in relaxing assumptions and addressing conceptual difficulties. %maybe add beaulieu_2016 here, instead
 One limitation is the pervasive consideration of traits in isolation. 

### 2019-02-22 manuscript/introduction.tex
revised first paragraph of Intro

--- 2019-02-19 manuscript/introduction.tex
@@
 
+Species accumulate around the tree of life at different rates.
+The search for traits that explain these differences has been accelerated by dramatic increases in phylogenetic data and, despite some setbacks \citep{maddison_2015, rabosky_2015}, advances in analytical methods \citep{maddison_2007, fitzjohn_2009, goldberg_2012, beaulieu_2016}.
+% todo more refs: papers with big trees; quasse, bisse-ness, fisse, abc-quasse, maybe hoehna-vs-rabosky
+In these studies, it is common to identify a single focal trait and investigate its association with rates of speciation and extinction.
+% (but see...
+?) % refs: there must be some papers that musse on multiple traits?
+especially for geographic range?
+This is problematic for two reasons.
+On the biology side, the context in which traits occur can lead to complex interactions, perhaps with other traits that make up an organism or its environment.
+On the methods side, excluding other traits that additionally influence diversification can lead to incorrect conclusions about the focal trait \citep{rabosky_2015, beaulieu_2016}.
+One approach to this problem is to allow for a `hidden' or unspecified trait in the analysis \citep{beaulieu_2016}. % and an earlier beaulieu? applications of hisse?
+This can be effective, but it leaves behind the question of whether the hidden trait corresponds to a real trait that should be sought, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
+Alternatively, we can simultaneously consider the effects of more than one known trait, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
+% B: We can then have a good prior, reliable data, and replicate the studies.
+E: Agree.
+But when I tried to add it, I realized hidden models can also be hypothesis-generating, with the purported hidden trait identified in the next clade.
+In this paper, we focus on two of the best studied traits in flowering plants, ploidy and breeding system, and investigate their individual and joint effects on diversification.
+
 The prevalence of variation in chromosome number, and especially ploidy, has been broadly considered a salient feature of flowering plants for nearly a century.

### 2019-02-22 manuscript/introduction.tex
revised polyploidy paragraphs of Intro

@@
 However, although it has long been recognized that the evolutionary transition from diploidy to polyploidy may result in correlated changes in mating patterns, the theoretical and empirical evidence is limited and often contradictory." (p.5)
-%
 %     any one study focuses on one/few traits
@@
 
-%Studying diversification linked to trait evolution\newline
-%Here I need help (perhaps will be the last line I write).
-In these context, botanist have been asking how polyploidy shapes patterns of diversification.
-The prevalence of polyploidy and its detection across multiple and highly diverse clades of angiosperms inevitably lead to the hypothesis of the importance of polyploidy in the speciation and extinction patterns observed across the flowering plant phylogeny.
-At the same time, a similar question has been asked in the breeding system world, where self-compatibility has evolved multiple times in flowering plants.
-However, the influence in diversification using both polyploidy and self-compatibility information has not been studied simultaneously.
-
-%Why polyploidy and self-compatibility in particular makes for an  interesting study in the context of diversification\newline In the polyploidy world, an important debate regarding the diversification of angiosperms has been ongoing since the publication of \citet{mayrose_2011}.
-The authors discovered that using the latest diversification model linked to two states diploid and polyploid, the net diversification of polyploids was much slower than the net diversification rate than polyploids.
-This result was surprising and an sparked a discussion about the long-term evolutionary consequences of polyploidy.
-\citet{soltis_2014} questioned if polyploidy should be regarded as an evolutionary dead-end since the net diversification of polyploids was negative, despite overwhelming evidence about the incidence of polyploidy  especially at the root of highly diverse angiosperm claims (ref here).
-A year later, diversification models were re-tested and corrected, and still found the same pattern \citep{mayrose_2015}, to the disappointment of plenty of botanists there was no denying of the weak trend of diversification that polyploids left behind.
-The most recent study by \citet{landis_2018} that used not only the presence of polyploidy in the tips but also the number of whole genome duplications in a taxon lineage found that....
-\newline  Meanwhile, studies focusing on diversification patterns and  breeding system have consistently found that self-incompatible plants often have higher net diversification rates compare to their self-compatible counterparts (Emma and Boris' papers here, what about papers that are not solanaceae?)
-
-%- Why other traits need to be considered as well\newline
-%- What other models and studies have done in the past\newline
-%- What is lacking from past approaches? \newline
-%There are two key questions that at the time of the polyploidy debate were difficult to ask.
-The first is if the models used to measure the diversification of polyploids were correct, and the second question is if the models have potentially included more evidence and potential traits that are not polyploidy or other lines of evidence that could be driving the patterns.
-At the time, in a different context \citet{beaulieu_2016} were finding an alternative solution to the first question, coming up with a new model that could represent the broad heterogeneity of the diversification process and parse out the signal between the trait of interest and the noise in diversification.
-Their model, the hidden state speciation and extinction model is a key component to detect whether polyploidy our something else unknown but related to it is driving the speciation and extinction patterns that we see in angiosperms.
-\newline
-
 Species accumulate around the tree of life at different rates.
@@
 
-The prevalence of variation in chromosome number, and especially ploidy, has been broadly considered a salient feature of flowering plants for nearly a century.
-%\citep{stebbins1938}; instead of "ploidy" here, we may wish to use some stand-in for karyotype multiples, and stick to that language throughout.
-The recent dramatic increase in the scale of available genome sequences uncovered ancient rounds of whole-genome duplications, and subsequent re-diploidization, across angiosperms \citep{lynch2000,vision2000}.
-Nearly all lineages of flowering plants are thought to have undergone at least one or two rounds of polyploidization. %cite
-It is therefore surprising that there is little agreement regarding the evolutionary consequences of polyploidy, especially how it affects diversification rates.
-%species richness; cite: mayrose2011, soltis2014, mayrose2015, tank2015, landis2018?
+% E plan, merging the text from R and B:
+%
+% polyploidy is potentially important, widespread
+% link with diversification; include diploidization hypothesis
+%
+% link with other traits -- correlated (geography), more directly connected (selfing, si)
+%
+% si
+% si and ploidy
+% pathways
 
-Polyploidization alters the genomic content of all cells, and consequently has the potential to affect a wide variety, if not all, traits. 
-From the very beginning, the study of this cytogenetic property considered correlations with other traits \citep{stebbins1938}. 
+Polyploidization multiplies the genomic content of cells, and it consequently has the potential to affect many or all traits and a great variety of evolutionary processes.
+It is also a mutation that occurs commonly in plants, at it is widespread at both population and lineage scales.
+The prevalence of variation in chromosome number, and especially ploidy, has been broadly considered a salient feature of flowering plants for nearly a century \citep{stebbins1938}.
+% B: instead of "ploidy" here, we may wish to use some stand-in for karyotype multiples, and stick to that language throughout.
+E: What word would that be?
+I was taking "ploidy" to mean "whatever multiple," but is that a terminological abuse?
+The recent dramatic increase in the scale of available genome sequences uncovered ancient rounds of whole-genome duplications, and subsequent diploidization, across angiosperms \citep{lynch2000, vision2000}.
+% E: I am deferring to R's use of "diploidization" instead of "re-diploidization"
+Nearly all lineages of flowering plants are thought to have undergone at least one or two rounds of polyploidization. % todo refs, jiao etc
+%
+The prevalence of polyploidy, its variation across clades, and its large effects on genotypes and phenotypes raises the hypothesis that ploidy plays an important role in shaping rates of speciation and extinction.
+This has led to a debate: \citet{mayrose_2011, mayrose_2015} found a lower rate of lineage diversification for polyploids, while \citet{soltis_2014} argued that polyploidy is not an evolutionary dead end because whole genome duplications occurred at the base of many successful angiosperm clades, and \citet{landis_2018} reported higher diversification rates following some whole genome duplications.
+% and tank2015?
+Our study revisits this question, additionally incorporating the process of diploidization.
+% E:  Rosana, could you add a little about the importance of including diploidization?  todo
+Furthermore, we consider the interaction between polyploidy and other traits.
+
+From the very beginning of polyploidy research, the study of this cytogenetic property considered correlations with other traits \citep{stebbins1938}. 
+Some of these associations are driven by indirect or ecological effects, such as the tendency for polyploids, and other traits, to be found at higher latitudes or more marginal habitats.
+% todo refs: husband etc
+Other correlations may have a more direct, causal connection.
 Among the many of changes associated with polyploidization, perhaps the most prominent is the association between polyploidy and propensity for self-fertilization \citep{stebbins1950, barrett1988}.
-Polyploidy is not only a suspected correlate of breeding systems but a causal link \citep{stout1942, lewis1947}.
-Doubled number of alleles in pollen is strongly suspected to effect disruption of the genetic mechanisms in gametophytic self-incompatibility systems \citep{entani1999, tsukamoto2005, kubo2010}. 
-Polyploidy is also thought to be associated with changes in the rate of self-fertilization, although the evidence for this correlated shift in mating system appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
+In some cases, the evidence for a correlated shift in mating system along with polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
+In other cases, however, polyploidy is not only a suspected correlate of breeding systems but indeed a causal link \citep{stout1942, lewis1947}.
+Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic self-incompatibility systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}.
+In clades with these systems, it is thus natural to consider the simultanous macroevolution of ploidy and breeding system.
 
+% E note to self: edit this next...
 Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974,barrett2013}.

### 2019-02-22 manuscript/introduction.tex
revision of SI intro paragraph

@@
 \section{Introduction}
-
-% Outline, discussed by Boris and Emma:
-%
-% we are interested in the evolution of traits and their influence on lineage diversification
-%     ref Barrett work 
-% 		no diversification?, but trait shifts:
-%		barrett_2008: "Because polyploidy affects the entire genome, it is perhaps not surprising that it influences many aspects of the phenotype, including the mating system.
-However, although it has long been recognized that the evolutionary transition from diploidy to polyploidy may result in correlated changes in mating patterns, the theoretical and empirical evidence is limited and often contradictory." (p.5)
-%     any one study focuses on one/few traits
-%     but any real organisms have many, so need to enlarge our view
-%     mini "here, we..."
-% we have tools to try to answer these questions
-%     sse models
-%     issues
-%     but a way forward is to allow for multiple factors that could influence diversification
-% ploidy
-%     Stebbins
-%     Mayrose, Soltis
-%     recent paleo WGD, diploidization
-% breeding system
-%     directionality
-%     diversification in Solanaceae
-%     correlation with ploidy
-%     pathways from SI-D to SC-P: direct or via SC-D
-% here, we...
-%     diversification: interaction between ploidy and breeding system
-%     also: diploidization, pathways
 
@@
 
-% E plan, merging the text from R and B:
-%
-% polyploidy is potentially important, widespread
-% link with diversification; include diploidization hypothesis
-%
-% link with other traits -- correlated (geography), more directly connected (selfing, si)
-%
-% si
-% si and ploidy
-% pathways
-
 Polyploidization multiplies the genomic content of cells, and it consequently has the potential to affect many or all traits and a great variety of evolutionary processes.
@@
 From the very beginning of polyploidy research, the study of this cytogenetic property considered correlations with other traits \citep{stebbins1938}. 
-Some of these associations are driven by indirect or ecological effects, such as the tendency for polyploids, and other traits, to be found at higher latitudes or more marginal habitats.
+Some of these associations are driven by indirect or ecological effects, such as the tendency for polyploidy, and other traits, to be found at higher latitudes or marginal habitats.
 % todo refs: husband etc
 Other correlations may have a more direct, causal connection.
-Among the many of changes associated with polyploidization, perhaps the most prominent is the association between polyploidy and propensity for self-fertilization \citep{stebbins1950, barrett1988}.
+Among the many changes associated with polyploidization, perhaps the most prominent is the association between polyploidy and propensity for self-fertilization \citep{stebbins1950, barrett1988}.
 In some cases, the evidence for a correlated shift in mating system along with polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
+% barrett2008: "Because polyploidy affects the entire genome, it is perhaps not surprising that it influences many aspects of the phenotype, including the mating system.
+However, although it has long been recognized that the evolutionary transition from diploidy to polyploidy may result in correlated changes in mating patterns, the theoretical and empirical evidence is limited and often contradictory." (p.5)
 In other cases, however, polyploidy is not only a suspected correlate of breeding systems but indeed a causal link \citep{stout1942, lewis1947}.
 Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic self-incompatibility systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}.
+This creates a correlation between polyploidy and self-compatibility by preventing the existence of self-incompatible polyploids.
 In clades with these systems, it is thus natural to consider the simultanous macroevolution of ploidy and breeding system.
@@
 % E note to self: edit this next...
-Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974,barrett2013}.
-They are by themselves thought to have a broad effect on other traits. 
-The frequent transition from self-incompatibility (SI) to self-compatibility (SC) is strongly associated with net diversification rate changes \citep{goldberg_2010,devos2014}.
-Given that changes in ploidy and breeding systems may be causally related and have profound affects on the fate of lineages, it seems particularly profitable to examine whether and how their macroevolutionary effects interact.
-Moreover, increases in ploidy are known to directly affect breeding system states and present a rare opportunity for the study of paths and path-dependence on trait evolution and diversification.
-For example, we could ask if the ordinary mutation (and fixation) causes more losses of SI than polyploidization, whether more SC polyploids are generated directly, after polyploidization of diploid SI lineages, or through a two-step pathway, involving loss of SC and subsequent polyploidization of a SC lineage.
-And, are diversification rates of polyploid SC lineages different if one or other path was taken? \citet{charlesworth1985} 
-%
-%trans-polymorphism: We also have this great data for SI loci and the transition could be even more common, providing power to resolve the great diversification questions
-%
-% set up pathway dependence from SI-D to SC-P: direct or via SC-D
+Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
+In particular, self-incompatibility (SI) systems cause a plant to reject its own pollen, and their loss, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution.
+% refs: stebbins, igic
+Previous phylogenetic analyses have reported higher rates of diversification for SI than for SC lineages \citep{goldberg_2010, devos2014}, but they have not considered the possibility of other correlated traits driving this pattern.
+Given that changes in ploidy and breeding systems may be causally related and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
+This includes their joint influence on lineage diversification, and also potential patterns in the order of their transitions.
+For example, do losses of SI more commonly occur tied to polyploidization, or without a ploidy shift?
+Do polyploids arise more commonly from SI or SC diploids?
+\Citet{robertson_2011} found that the pathway from SI diploids to SC polyploids is dominated by loss of SI followed later by polyploidization over long timescales, but proceeds in one step via polyploidization of SI species over short timescales.
+We revisit this question with a greatly improved phylogeny and methods that allow for diversification rate differences.
+% E: Technically, we allow for div effects in the rate estimates, but not in the pathways.
+If SC-D have so much extinction that the step to SC-P never happens, the old pathway methods won't see that.
+I'll think about if this is easy to fix.
+% And, are diversification rates of polyploid SC lineages different if one or other path was taken?
+\citet{charlesworth1985} % E: This is a great question.
+We don't answer it with this round of analyses.
+Maybe when we have more time for the revision?
 
+% note: saving trans-specific polymorphism for Methods
+% B: trans-polymorphism: We also have this great data for SI loci and the transition could be even more common, providing power to resolve the great diversification questions
+
+% here, we...
+%     diversification: interaction between ploidy and breeding system
+%     also: diploidization, pathways
 Here, we [...]
@@
 % 	examine pathways from SI-D to SC-P: direct or via SC-D
-
-%The second question (talk here about breeding system and polyploidy, and how this could be different across clades and this is one of the reasons why we focused on Solanaceae).
-%-Diversification and breeding systems\\
-%Goldberg and Igic 2012, different  perspectives\\

### 2019-02-23 manuscript/introduction.tex
rough draft of final Intro paragraph

@@
 Maybe when we have more time for the revision?
+% E: Relatedly, I'm wondering if pathways should be a separate paper.
+There is the robertson2011 question, the modification that allows for diversification, and the charlesworth1985 question.
+And dioecy.
 
@@
 
-% here, we...
-%     diversification: interaction between ploidy and breeding system
-%     also: diploidization, pathways
-Here, we [...]
-% Someone else should do this, once we have the full data.
-% diversification: interaction between ploidy and breeding system
-% 	examine individual effects
-% 	examine joint effects
-% 	with and without HMMs
-% 	allow for re-diploidization or not
-% also: diploidization, pathways
-% 	examine pathways from SI-D to SC-P: direct or via SC-D
+Here, we use extensive data on ploidy and breeding system in Solanaceae to investigate the associations of these two traits with lineage diversification.
+Considering each trait separately indicates that each is connected to diversification differences.
+Considering them jointly, however, reveals that the ploidy connection is removed by incorporating breeding system.
+We further show that the general results are robust to allowing for diploidization and a hidden trait, and something about pathways. % FIXME
+Our results emphasize the importance of considering traits not only in isolation, especially when there are strong correlations between them.

### 2019-02-23 manuscript/results.tex
filled in the table of median rate estimates

@@
 \section{Results}
+
+%B: IMO each instance of "(Figure 2(X))" ought be replaced with "(Figure 2x)", so that (Figure 2(A)) becomes (Figure 2a).
+Double-parentheses should be reserved for equations and we already use caps, for example C and D, for traits.
+I am happy to make the change in the text, which is trivial, but I am not sure how the PDF conversion through illustrator would fare.
+
 \subsection{Polyploidy and Diversification Models}
@@
 
-% E & B:  We just spent awhile poring over the great supp figures of rate estimates, which are essential to grasp the uncertainties.
-But to facilitate quick comparisons across models, could we have a table of the median rate estimates for all models?
-Possibly like the format I've laid out below?
-Or maybe there's a better way.
-
 \begin{supptable}
@@
         %                                                                                      
-        $r_D$          &     &         & --- & ---     & ---      & ---          \\
-        $r_P$          &     &         & --- & ---     & ---      & ---          \\
-        $r_I$          & --- & ---     &     &         & ---      & ---          \\
-        $r_C$          & --- & ---     &     &         & ---      & ---          \\
-        $r_{ID}$       & --- & ---     & --- & ---     &          &              \\
-        $r_{CD}$       & --- & ---     & --- & ---     &          &              \\
-        $r_{CP}$       & --- & ---     & --- & ---     &          &              \\
+        $r_D$          & 0.382 & 0.698, 0.100 & ---    & ---          & ---      & ---          \\
+        $r_P$          & 0.109 & 0.587, 0.182 & ---    & ---          & ---      & ---          \\
+        $r_I$          & ---   & ---          & 0.550  & 0.386, 0.877 & ---      & ---          \\
+        $r_C$          & ---   & ---          &-0.001  &-0.059, 0.606 & ---      & ---          \\
+        $r_{ID}$       & ---   & ---          & ---    & ---          & 0.449    & 0.318, 0.789 \\
+        $r_{CD}$       & ---   & ---          & ---    & ---          & 0.050    &-0.248, 0.494 \\
+        $r_{CP}$       & ---   & ---          & ---    & ---          &-0.027    & 0.110, 0.634 \\
         %                                                                                      
-        $\rho$         &     &         & --- & ---     & ---      & ---          \\
-        $\rho_I$       & --- & ---     & --- & ---     &          &              \\
-        $\rho_C$       & --- & ---     & --- & ---     &          &              \\
-        $\delta$       &     &         &     &         &          &              \\
-        $q_{IC}$       & --- & ---     &     &         &          &              
+        $\rho$         & 0.033 & 0.026        & ---    & ---          & ---      & ---          \\
+        $\rho_I$       & ---   & ---          & ---    & ---          & 0.063    & 0.047        \\
+        $\rho_C$       & ---   & ---          & ---    & ---          & 0.024    & 0.011        \\
+        $\delta$       & 0.050 & 0.162        & ---    & ---          & 0.022    & 0.107        \\
+        $q_{IC}$       & ---   & ---          & 0.364  &              & 0.194    & 0.164        
     \end{tabular}
@@
     \vspace{12pt}
-    \begin{tabular}{lcccccc}
-                       & D/P & D/P+A/B & I/C & I/C+A/B & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
+    \begin{tabular}{lcccc}
+                       & D/P   & D/P+A/B      & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
         %                                                              
-        $r_D$          &     &         & --- & ---     & ---      & ---          \\
-        $r_P$          &     &         & --- & ---     & ---      & ---          \\
-        $r_I$          & --- & ---     &     &         & ---      & ---          \\
-        $r_C$          & --- & ---     &     &         & ---      & ---          \\
-        $r_{ID}$       & --- & ---     & --- & ---     &          &              \\
-        $r_{CD}$       & --- & ---     & --- & ---     &          &              \\
-        $r_{CP}$       & --- & ---     & --- & ---     &          &              \\
+        $r_D$          & 0.260 & 0.193, 0.658 & ---      & ---          \\
+        $r_P$          &-0.056 & 0.030, 0.187 & ---      & ---          \\
+        $r_I$          & ---   & ---          & ---      & ---          \\
+        $r_C$          & ---   & ---          & ---      & ---          \\
+        $r_{ID}$       & ---   & ---          & 0.455    & 0.309, 0.797 \\
+        $r_{CD}$       & ---   & ---          & 0.065    &-0.006, 0.587 \\
+        $r_{CP}$       & ---   & ---          &-0.088    &-0.074, 0.403 \\
         %                                                              
-        $\rho$         &     &         & --- & ---     & ---      & ---          \\
-        $\rho_I$       & --- & ---     & --- & ---     &          &              \\
-        $\rho_C$       & --- & ---     & --- & ---     &          &              \\
-        $q_{IC}$       & --- & ---     &     &         &          &              
+        $\rho$         & 0.047 & 0.047        & ---      & ---          \\
+        $\rho_I$       & ---   & ---          & 0.067    & 0.053        \\
+        $\rho_C$       & ---   & ---          & 0.033    & 0.032        \\
+        $q_{IC}$       & ---   & ---          & 0.198    & 0.145        
     \end{tabular}
     \end{center}
-    \caption{median rate estimates for all models (see supp figs for more details and uncertainties); when there are two numbers in one cell, they're for the A and B hidden states; --- means that parameter was not present in the model; upper section is for models with $\delta$ and lower section is for models without}
+    \caption{
+        Median rate estimates for all models.
+        Units are per million years.
+        Two comma-separated numbers refer to the $A$ and $B$ hidden states, and --- means the parameter was not present in the model.
+        The upper section is for models with diploidization, and the lower section is for models without.
+        The supplemental figures show the corresponding distributions of parameter estimates.
+    }
     \label{table:estimates}

### 2019-02-23 summarystats/pathways.R
first pass at pathways calculations/plots
without or with diversification

@@ -1,0 +1,115 @@
+#--------------------------------------------------
+# method 1: P = exp(Q t)
+#-------------------------------------------------- 
+library(rmutil)  # contains mexp()
+path.probs1 <- function(t, Q)
+{
+    P <- mexp(Q * t)
+    
+    ans <- c(P[1,3], P[1,4])
+    names(ans) <- c("Wa", "Wb")
+
+    return(ans)
+}
+
+# Note: The matrix exponential function is also available in the msm library as
+# MatrixExp(), but that seems to perform poorly for large t.  So use mexp here, 
+# but also check with method 2 when assuming irreversibility.
+
+#--------------------------------------------------
+# method 2: P_14a and P_14b
+#-------------------------------------------------- 
+path.probs2 <- function(t, q13, q14, q34)
+{
+    q134 <- q13 + q14
+    e1t <- exp(-q134 * t)
+
+    ppa <- q14 / q134 * (1 - e1t)
+
+    if (abs(q34 - q13 - q14) > 1e-10)
+    {
+        ppb <- q13 / q134 * 
+              (1 - 1/(q134-q34) * (q134 * exp(-q34*t) - q34 * e1t) )
+    } else
+    {
+        ppb <- q13 / q34 * (1 - (1 + q34 * t) * exp(-q34 * t))
+    }
+
+    ans <- c(ppa, ppb)
+    names(ans) <- c("Wa", "Wb")
+
+    return(ans)
+}
+
+#--------------------------------------------------
+# usage
+#-------------------------------------------------- 
+
+# 1 = SI, D
+# 3 = SC, D
+# 4 = SC, P
+# Wa = 1 -> 4
+# Wb = 1 -> 3 -> 4
+
+# MAP values
+q14 <- 0.0681          # ID -> CP, rho_I
+q13 <- 0.2066          # ID -> CD, q_IC
+q34 <- 0.0370          # CD -> CP, rho_C
+q41 <- q31 <- q43 <- 0 # irreversible
+
+# Q is only needed for method 1
+Q <- matrix(c(
+        -q13 - q14,  q13,        q14,  0,
+        q31,        -q31 - q34,  0,    q34,
+        q41,         0,         -q41,  0,
+        0,           q43,        0,   -q43
+        ), byrow=TRUE, nrow=4)
+
+tmax <- 50
+times <- seq(0, tmax, tmax/1000)
[…]

### 2019-02-23 manuscript/introduction.tex
- adding .gitignore to remove nuisance files (e.g. accidental Rhistory editing) - modified introduction - modified addresses to conform to NP journal standards

@@
 especially for geographic range?
-This is problematic for two reasons.
-On the biology side, the context in which traits occur can lead to complex interactions, perhaps with other traits that make up an organism or its environment.
-On the methods side, excluding other traits that additionally influence diversification can lead to incorrect conclusions about the focal trait \citep{rabosky_2015, beaulieu_2016}.
-One approach to this problem is to allow for a `hidden' or unspecified trait in the analysis \citep{beaulieu_2016}. % and an earlier beaulieu? applications of hisse?
+
+%B: I understand the apparent contrast, but biology vs.
+methods is not a properly framed Venn diagram, IMO.
+%B: was: On the biology side, the context in which traits occur can lead to complex interactions, perhaps with other traits that make up an organism or its environment.
+%This is problematic for two reasons.
+This is problematic because the context in which traits occur can lead to complex interactions, perhaps with other traits that make up an organism or its environment.
+%B: was: On the methods side, excluding other traits that additionally influence diversification can lead to incorrect conclusions about the focal trait \citep{rabosky_2015, beaulieu_2016}.
+Additionally, a lack of methods that allow for simultaneous inference of effects of other traits, which also influence diversification, can lead to incorrect conclusions about the focal trait \citep{rabosky_2015, beaulieu_2016}.
+One approach to this problem is to allow for a `hidden' or unspecified trait in the analysis \citep{beaulieu_2016}.
+% and an earlier beaulieu?
+applications of hisse?
+%B: perhaps above can begin with: "One proposed approach to accommodate..."
 This can be effective, but it leaves behind the question of whether the hidden trait corresponds to a real trait that should be sought, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
@@
 But when I tried to add it, I realized hidden models can also be hypothesis-generating, with the purported hidden trait identified in the next clade.
+%B: I don't get "the next clade"?
 In this paper, we focus on two of the best studied traits in flowering plants, ploidy and breeding system, and investigate their individual and joint effects on diversification.
@@
 I was taking "ploidy" to mean "whatever multiple," but is that a terminological abuse?
+%B: sorry, that was supposed to complain about the use of "polyploidy" instead of "ploidy" to designate the trait.
+The trait is ploidy, and a state is polyploidy.
+Also, we should clarify that polyploidy is any of tetrapoidy+.
+I am not sure what we did with 3x here.
+It's rather rare, but...?
 The recent dramatic increase in the scale of available genome sequences uncovered ancient rounds of whole-genome duplications, and subsequent diploidization, across angiosperms \citep{lynch2000, vision2000}.
-% E: I am deferring to R's use of "diploidization" instead of "re-diploidization"
 Nearly all lineages of flowering plants are thought to have undergone at least one or two rounds of polyploidization. % todo refs, jiao etc
@@
 
-From the very beginning of polyploidy research, the study of this cytogenetic property considered correlations with other traits \citep{stebbins1938}. 
+From the beginning, study of this variable cytogenetic property considered correlations with other traits \citep{stebbins1938}.
+%B: ploidy research, or nothing (as now); was: "From the very beginning of polyploidy research," polyploidy is a state of the trait.
 Some of these associations are driven by indirect or ecological effects, such as the tendency for polyploidy, and other traits, to be found at higher latitudes or marginal habitats.
@@
 Other correlations may have a more direct, causal connection.
-Among the many changes associated with polyploidization, perhaps the most prominent is the association between polyploidy and propensity for self-fertilization \citep{stebbins1950, barrett1988}.
+Among the many changes associated with polyploidization, perhaps the most prominent is the association between polyploidy and propensity for self-fertilization \citep{stebbins1950}.
+%B: removed barrett1988, because he highlights correlation between polyploidy and self-compatibility, not self-fertilization.
 In some cases, the evidence for a correlated shift in mating system along with polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
@@
 Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic self-incompatibility systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}.
-This creates a correlation between polyploidy and self-compatibility by preventing the existence of self-incompatible polyploids.
+This creates a correlation between polyploidy and self-compatibility by precluding the existence of self-incompatible polyploids. %B: prevent twice in a row; replaced with preclude.
 In clades with these systems, it is thus natural to consider the simultanous macroevolution of ploidy and breeding system.

### 2019-02-23 manuscript/main.tex
- new paragraph(s) in introduction, and matching citations - new title ideas or modifications
- adding .gitignore to remove nuisance files (e.g. accidental Rhistory editing) - modified introduction - modified addresses to conform to NP journal standards

@@
 %
+%% When the favorite trait is not enough: the link of polyploidy and breeding system in diversification %B: <- eh?
+%
 % New comparative approaches for the study of breeding system, polyploidy, and lineage diversification % E: the tentative title given to New Phyt
@@
 % Interacting effects of ploidy and breeding system on lineage diversification % E: but "effects" is too causal
+%
+%% Interaction of ploidy and breeding system in lineage diversification %B: <- how 'bout that?
 
@@
 
+%B: I included extended affiliations necessary at submission time. Check, please! 
+%B->R you may wish to add new affiliation at Manoa, but I left the original state alone.
 \noindent
-Rosana Zenil-Ferguson,%
-\endnote{University of Minnesota Twin Cities}
+Rosana Zenil-Ferguson\endnote{Department of Ecology, Evolution, and Behavior, University of Minnesota, Saint Paul, MN 55108, U.S.A.}$^{\dagger}$,
 %
 \noindent
-J.~Gordon Burleigh,%
-\endnote{University of Florida}
+J.~Gordon Burleigh\endnote{Department of Biology, University of Florida, Gainesville, FL 32611, U.S.A.},
 %
 \noindent
-William A.~Freyman,%
-\endnote{23 and me}
+William A.~Freyman\endnote{23andMe, Inc., 899 W Evelyn Ave, Mountain View, CA 94041, U.S.A.},
 %
 \noindent
-Boris Igi\'c,%
-\endnote{University of Illinois}
+Boris Igi\'c\endnote{Department of Biological Sciences, University of Illinois at Chicago, Chicago, IL 60607, U.S.A. },
 %
 \noindent
-Itay Mayrose,%
-\endnote{Tel Aviv University}
+Itay Mayrose\endnote{Department of Molecular Biology and Ecology of Plants, Tel Aviv University, Tel Aviv, Israel},
 %
 and
-Emma E.~Goldberg%
-\endnote{University of Minnesota Twin Cities}
-
+Emma E.~Goldberg\endnote{Department of Ecology, Evolution, and Behavior, University of Minnesota, Saint Paul, MN 55108, U.S.A.}
+%
 \vfill
@@
 \noindent
-Author for correspondence: Rosana Zenil-Ferguson
-
+$\dagger$Author for correspondence.
+%Rosana Zenil-Ferguson
 \vfill

### 2019-02-24 manuscript/results.tex
Edited results, meat comes earlier

@@
 This result holds true whether or not the diploidization parameter is present.
-However in the presence of  the diploidization parameter the net diversification rate of polyploids is nonnegative with probability 1 (Figure 2(A)), whereas in the absence of diploidization the net diversification rate of polyploids can be negative with a probability (HERE verify the quantile) (FIGURE 3(A)).
+However, in the presence of  the diploidization parameter the net diversification rate of polyploids is nonnegative with probability 1 (Figure 2(A)), whereas in the absence of diploidization the net diversification rate of polyploids can be negative with a probability 0.87 (FIGURE 3(A)).
 In terms of the relative extinction, when the diploidization parameter is present both polyploids and diploids have posterior distributions that overlap, but that pattern changes in the absence of the diploidization parameter leading to a significant difference between relative extinction where polyploids have a significant higher relative extinction rate (see Supplementary Information).\newline
 
-For the D/P- A/B model with diploididization the diploid and polyploid net diversification rates are overlapping for both state A and B of the hidden trait (Figure 2(B)).
+For the D/P+A/B model with diploididization the diploid and polyploid net diversification rates are overlapping for both state A and B of the hidden trait (Figure 2(B)).
 In this model, the differences in net diversification are due to the presence of a hidden trait and not to the differences in ploidy.
@@
 
-\subsection{Breeding System and Diversification models }
-In the I/C breedyng system model we found that the net diversification rate for self-incompatible state is larger than the net diversification rate for self-compatible state (Figure 2(C)).
+\subsection{Polyploidy and Breeding System models}
+When breeding system is considered simultaneously from polyploidy, we found that in the ID/CD/CP model self-incompatible and diploid state (ID) has significantly larger net diversification rate compare to both self-compatible diploid (SD) and polyploid rates(CP).
+Meanwhile, both self-compatible diploid and polyploid net diversification rates have overlapping posterior distributions (Figure 2(E)).
+When hidden state was added, the ID/P/CD+A/B model, significant differences between self-compatible and self-incompatible diploids continue to be the truth for both A and B values of the hidden state.
+However, the posterior distribution for the net diversification rate of polyploids overlaps with both the self-compatible and self-incompatible posterior distributions for each value of the hidden state (Figure 2F).
+This result indicated that polyploid state was not significantly different from diploid from the net diversification perspective.
+In the absence of diploidization, models ID/CD/CP no $/delta$ and ID/P/CD no $delta$ +A/B showed that the net diversification of polyploids is equivalent to the net diversification of self-compatible diploids.
+In the absence of diploidization, the net diversification rates of self-incompatible diploids are always faster than their self-compatible diploid and polyploid counterparts (Figures 3C and 3D).
+\newline
+Polyploidization rates from self-compatible diploid $\rho_{CD}$ and from self-incompatible diploid $\rho_{ID}$ are not significantly different (see supplementary information figures).
+However, the distribution of $\rho_{ID}$ seems to always consider faster values compare to the distribution of $\rho_{CD}$.
+This result is highlighted in ancestral reconstructions of the $ID/CD/CP$ model where transitions from ID to CP are more probable than transitions from CD to CP.
+% Emma here we put the figure?
+
+\subsection{Diploidization as an exploratory hypothesis}
+ For the two models that only consider diploid and polyploid state (D/P and D/P+A/B) the effect of removing parameter $\delta$ was that net diversification rate of polyploids can be negative with high probability (Figures 3A and 3B) .
+However, in the models that simultaneously considered polyploidy and breeding system, removing the $\delta$ parameter did not have much of an effect on the location of the posterior distribution of polyploid net diversification rate but on the credible interval width and overlap with net diversification of diploids.
+When diploidization was absent, the net diversification of polyploids overlaps with the net diversification of self-compatible diploids but not with the one of self-incompatible diploids (Figures 3C and 3D).
+For all models, the posterior distribution of diploidization rate is more uncertain than the one of polyploidization rates (See supplementary information figures).
+
+\subsection{Modeling breeding system by itself }
+In the I/C breeding system model we found that the net diversification rate for self-incompatible state is larger than the net diversification rate for self-compatible state (Figure 2(C)).
 The net diversification rate for self-compatible has a probability distribution centered at zero.
 \newline
-
-When a hidden state is added in the I/C-A/B model, we found that under the hidden state A the self-compatible and self-incompatible net diversification rates are different.
-Under the hidden state B, those two rates overlap with a probability (HERE CALCULATE THAT) meaning that they are different with probability  (CALCULATE) (Figure 2(D)).
+When a hidden state is added in the I/C+A/B model, we found that under the hidden state A the self-compatible and self-incompatible net diversification rates are different.
+Under the hidden state B, those two rates overlap.
+Furthermore, the net diversification of self-compatible taxa on hidden state B overlaps with the hidden state rate of self-incompatible taxa on state A  (Figure 2(D)).
 These results agree with previous results found by  \citet{goldberg_2012}.
 However, \citet{goldberg_2012} used a ClaSSE  approach since they were interested in anagenetic and cladogenetic changes for self-incompatibility using a smaller subset of the data presented in the current work.
-
-% Describe rate q_IC and hidden state rate
-
-\subsection{Polyploidy and Breeding Sytem models}
-In the ID/P/CD model we found that self-incompatible and diploid state has a significantly larger net diversification rate compare to both self-compatible diploid and polyploid rates.
-Meanwhile, both self-compatible diploid and polyploid posterior distributions of net diversification rates completely overlap (Figure 2(E)).
-When hidden state was added in the ID/P/CD-A/B model, we observed significant differences between self-compatible and self-incompatible diploids for both A and B values of the hidden state.
-Self-incompatible state had a larger net diversification rate than self compatible for both A and B states.
-However, the posterior distribution for the net diversification rate of polyploids overlaps with both the self-compatible and self-incompatible posterior distributions for each value of the hidden state(Figure 2F) meaning that polyploidy state is not significantly different from diploid in net diversification terms.
-The resulting effect of adding the hidden state values is significant
-
-\subsection{Diploidization as an exploratory hypothesis}
-In the D/P model  the diploidization rate $\delta$ and polyploidization rate $\rho$ are different from zero with probability 1.
-Diploidization rate is more uncertain than polyploidization.
-For the D/P no $\delta$ model, the polyploidization rate is still different from zero but has a wider 95\% credible interval (see Supplementary Information).
-In the D/P-A/B model the diploidization rate remains positive but there rate of polyploidization becomes really uncertain.
-Whereas, in the absence of diploidzation (D/P no $\delta$ A/B) the rate of polyploidy has a small credible interval \newline
-
-For the model ID/P/CD containing both polyploidy and breeding system traits, we found that the diploidization rate is really uncertain with a MAP that is (CALCULATE VALUE HERE- close to zero).
-The polyploidization rate from self-incompatible diploid $\rho_I$ is slightly faster than the polyploidization rate from self-compatible diploid $\rho_C$  and that pattern remains the same for the ID/P no $\delta$/CD model(see supplementary information).
-%Here I'm missing the MuHiSSE plots.
+Transition rate between incompatible and compatible state are frequent with a rate $q_{IC}=0.3$ for both the  I/C and the I/C+A/B models.
 
@@
 
-In \cref{table:marginallike} we list the marginal likelihood in log scale for each of the models tested.
-In the table we show what are the different components included and excluded for each model as a a summary of the diagrams from \cref{figure:netdivall}.
-From the marginal likelihoods in log scale, the Bayes factors in log-scale were calculated as shown in table \cref{table:bayesfactors}.
-After testing every single pair of polyploidy modes (1-4) we found overwhelming evidence that the best polyploidy model is always  the D/P-A/B, that is the model with hidden state and diploidization.
+In \cref{table:marginallike} we listed the marginal likelihood values calculated in log scale ($log(P(\mathbf{X}|M)$) for each of the models tested.
+The different components included and excluded for each model as a a summary of the diagrams from \cref{figure:netdivall} are indicated in each model.
+From the marginal likelihoods in log scale, the Bayes factors in log-scale($BF(M_0,M_1)=log(P(\mathbf{X}|M_0)-P(\mathbf{X}|M_0)$) were calculated as shown in table \cref{table:bayesfactors}.
+After testing every single pair of polyploidy models (1-4) we found overwhelming evidence that the best polyploidy model is always  the D/P+A/B, that is the model with hidden state and diploidization, that is Bayes factors were always  greater than 2 when D/P+A/B was the first model in the Bayes factor equation or negative when D/P+A/B was the second model in Bayes factor equation.
 \newline
-For the two models following the evolution of breeding system, the I/C-A/B is the best choice between models 3 and 4.\newline
-The models that follow the diversification linked to both polyploidy and breeding system are the last 4 (models 5-8).
-When comparing using Bayes factores every two models we found that the IC/P/CD-A/B is always preferred over the rest, meaning that the model that has a hidden state and diploidization is chosen over the ones that lack either of both of those options.\newline
-Therefore, the models that were chosen were always the ones containing a hidden trait, and in the case of polyploidy models, the ones containing a diploidization parameter $\delta$ are preferable.
+For the two models following the evolution of breeding system, the I/C+A/B is the best choice between models 3 and 4 (BF(IC, IC+A/B)=-39.5<0).\newline
+The four models that follow the diversification linked to both polyploidy and breeding system (models 5-8) where compared in pairs.
+Bayes factors show decisive evidence in favor of the  IC/P/CD+A/B over the rest, meaning that the model that has a hidden state and diploidization is chosen over the ones that lack either of both of those options \cref{table:bayesfactors} .\newline
+Therefore, the models that were chosen were always the ones containing the hidden trait, or diploidization parameter $\delta$ when ploidy state was considered.
 

### 2019-02-25 manuscript/methods.tex
some changes from "model-A/B" to "model+A/B"

@@
 % D/P ploidy model
-% D/P-A/B ploidy hidden state model
+% D/P+A/B ploidy hidden state model
 % I/C breeding system model
-% I/C-A/B breeding system and hidden state model
+% I/C+A/B breeding system and hidden state model
 % ID/CD/CP ploidy and breeding system model
-% ID/CD/CP-A/B ploidy, breeding system, and hidden state model
+% ID/CD/CP+A/B ploidy, breeding system, and hidden state model
 
@@
 In this model, each of the observed diploid and polyploid states is subdivided by a binary hidden trait with states $A$ and $B$.
-We call this the \textit{D/P--A/B ploidy and hidden state} model. 
+We call this the \textit{D/P+A/B ploidy and hidden state} model. 
 We estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$), extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$), net diversification rates ($r_{D_A},\ r_{D_B},\ r_{P_A},\ r_{P_B}$), and relative extinction rates ($\nu_{D_A},\ \nu_{D_B},\ \nu_{P_A},\ \nu_{P_B}$).
@@
 To parse out the effect of breeding system on diversification, while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
-We call this HiSSE model the \textit{I/C--A/B breeding system and hidden state model}.
+We call this HiSSE model the \textit{I/C+A/B breeding system and hidden state model}.
 
@@
 
-%B: What happens if SC->SI is not fixed? Asking for a friend.
+%B: What happens if SC->SI is not fixed? Asking for a friend.  E: Surprisingly, Rosana reports, results are not turned on their head.
 
@@
 
-The ID/CD/CP model could potentially capture similar dynamics as earlier models, if the effects of the hidden state in D/P--A/B were effectively caused by breeding system (or its correlates), and the hidden state in I/C--A/B was effectively caused by ploidy.
-%B: you may wish to get rid of "(or its correlates)", which I inserted; along with "caused".
-I know that causation is not strict here, but it clarifies the intended biological explanation.
-%B: was: "if the hidden state in D/P--A/B was effectively breeding system, and the hidden state in I/C--A/B was effectively ploidy."
+The ID/CD/CP model could potentially capture similar dynamics as earlier models, if the effects of the hidden state in D/P+A/B were effectively caused by breeding system (or its correlates), and the hidden state in I/C+A/B was effectively caused by ploidy.
+% B: I know that causation is not strict here, but it clarifies the intended biological explanation.
 There is also the potential, however, for a hidden factor to be influencing diversification beyond both of our focal traits, and this could again mislead inferences.
 We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018, herrera_2018, huang_2018}.
-We refer to this as the \textit{ID/CD/CP--A/B} model.
+We refer to this as the \textit{ID/CD/CP+A/B} model.
 A fully parameterized version of this model would have 26 rate parameters \citep{herrera_2018}. 
@@
 Perhaps this complication is covered in the Discussion section.
-
-% E todo: explain "pathway" method of robertson2011
 
@@
 
+% \subsection{Pathways to polyploidy} % E todo
+% robertson2011 method, and then with diversification
+
 \subsection{Statistical inference under the models}
 
-Parameters for each of the 10 diversification models were performed using custom code in the RevBayes \citep{hoehna_2016} environment.
-%B: fixme: I think the intent here is  "Parameter estimation [...] was peformed"
+Parameters for each of the 10 diversification models were estimated using custom code in the RevBayes \citep{hoehna_2016} environment.
 Code for analyses and key results is available at \url{https://github.com/roszenil/solploidy}.
@@
 
-Using the marginal likelihood values, we calculated thirteen different Bayes factors. % E: it said "fourteen" but I only count thirteen? %B: fixme: my count too.
-Six compared the models of polyploidy against each other (D/P and D/P--A/B, each with or without diploidization), one compared the breeding system models (I/C and I/C--A/B), and six compared the models with both traits (ID/CD/CP and ID/CD/CP--A/B, each with or without diploidization) (\cref{table:bayesfactors}).
+Using the marginal likelihood values, we calculated thirteen different Bayes factors.
+Six compared the models of polyploidy against each other (D/P and D/P+A/B, each with or without diploidization), one compared the breeding system models (I/C and I/C+A/B), and six compared the models with both traits (ID/CD/CP and ID/CD/CP+A/B, each with or without diploidization) (\cref{table:bayesfactors}).
 %
@@
 I find it useful to imagine Spencer reading this, for example.
+% E: "Lumpable" is the math word for a particular relation among state codings, used in that Tarasov paper.
+I don't think we want to get into defining it precisely.
+Instead, we want Fig 1 to make it intuitively clear, to Spencer and everyone else.
+Probably need to add to its caption.
 

### 2019-02-25 manuscript/results.tex
serious editing of the first part of Results

@@
 
-%B: IMO each instance of "(Figure 2(X))" ought be replaced with "(Figure 2x)", so that (Figure 2(A)) becomes (Figure 2a).
-Double-parentheses should be reserved for equations and we already use caps, for example C and D, for traits.
-I am happy to make the change in the text, which is trivial, but I am not sure how the PDF conversion through illustrator would fare.
+\subsection{Traits and diversification}
 
-\subsection{Polyploidy and Diversification Models}
-Similarly to the results obtained by \citet{mayrose_2011} and \citet{mayrose_2015}, we found that in the D/P polyploidy model the net diversification of diploids is larger than the the net diversification of polyploids since the net diversification distributions do not overlap (Figure 2(A)).
-This result holds true whether or not the diploidization parameter is present.
-However, in the presence of  the diploidization parameter the net diversification rate of polyploids is nonnegative with probability 1 (Figure 2(A)), whereas in the absence of diploidization the net diversification rate of polyploids can be negative with a probability 0.87 (FIGURE 3(A)).
-In terms of the relative extinction, when the diploidization parameter is present both polyploids and diploids have posterior distributions that overlap, but that pattern changes in the absence of the diploidization parameter leading to a significant difference between relative extinction where polyploids have a significant higher relative extinction rate (see Supplementary Information).\newline
+% E: I brutally shortened the text here and hence put all the models into one section for the "diversification" question.  But I think I retained all the ideas that were written out before.
+% Much of what I removed was descriptions about which distributions are/aren't overlapping with one another.
+I think that's easier for the reader to see in the figures than to read in the text.
+But now the text says things like if some estimates are "different", so perhaps in Methods we should explain different = non-overlapping?
 
-For the D/P+A/B model with diploididization the diploid and polyploid net diversification rates are overlapping for both state A and B of the hidden trait (Figure 2(B)).
-In this model, the differences in net diversification are due to the presence of a hidden trait and not to the differences in ploidy.
-When diploidization parameter is absent the hidden state is still driving the differences in diversification rates (Figure 3(B)).
+We found that when considered alone, ploidy and breeding system each showed significant associations with net diversification differences.
+When considered together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well.
 
-\subsection{Polyploidy and Breeding System models}
-When breeding system is considered simultaneously from polyploidy, we found that in the ID/CD/CP model self-incompatible and diploid state (ID) has significantly larger net diversification rate compare to both self-compatible diploid (SD) and polyploid rates(CP).
-Meanwhile, both self-compatible diploid and polyploid net diversification rates have overlapping posterior distributions (Figure 2(E)).
-When hidden state was added, the ID/P/CD+A/B model, significant differences between self-compatible and self-incompatible diploids continue to be the truth for both A and B values of the hidden state.
-However, the posterior distribution for the net diversification rate of polyploids overlaps with both the self-compatible and self-incompatible posterior distributions for each value of the hidden state (Figure 2F).
-This result indicated that polyploid state was not significantly different from diploid from the net diversification perspective.
-In the absence of diploidization, models ID/CD/CP no $/delta$ and ID/P/CD no $delta$ +A/B showed that the net diversification of polyploids is equivalent to the net diversification of self-compatible diploids.
-In the absence of diploidization, the net diversification rates of self-incompatible diploids are always faster than their self-compatible diploid and polyploid counterparts (Figures 3C and 3D).
-\newline
-Polyploidization rates from self-compatible diploid $\rho_{CD}$ and from self-incompatible diploid $\rho_{ID}$ are not significantly different (see supplementary information figures).
-However, the distribution of $\rho_{ID}$ seems to always consider faster values compare to the distribution of $\rho_{CD}$.
-This result is highlighted in ancestral reconstructions of the $ID/CD/CP$ model where transitions from ID to CP are more probable than transitions from CD to CP.
-% Emma here we put the figure?
+For ploidy alone (D/P model), we found a larger net diversification rate for diploids than for polyploids, in agreement with \citep{mayrose_2011, mayrose_2015}.
+This result holds with (\cref{figure:netdivall}A) or without (\cref{figure:netdivnodip}A) the diploidization parameter, though including diploidization shifts the net diversification rate of polyploids to be non-negative.
+Incorporating a hidden state, however, removes the clear separation in diversification between diploids and polyploids (D/P+A/B model, \cref{figure:netdivall}B, \cref{figure:netdivnodip}B).
+Thus, differences in net diversification are better explained by an unknown factor than by ploidy.
+% E: I removed the description of relative extinction results.
+We don't discuss it for other models or show it in the main text, and I never really know what to make of that compound parameter.
+But it's good that it's in the supp info.
 
-\subsection{Diploidization as an exploratory hypothesis}
- For the two models that only consider diploid and polyploid state (D/P and D/P+A/B) the effect of removing parameter $\delta$ was that net diversification rate of polyploids can be negative with high probability (Figures 3A and 3B) .
-However, in the models that simultaneously considered polyploidy and breeding system, removing the $\delta$ parameter did not have much of an effect on the location of the posterior distribution of polyploid net diversification rate but on the credible interval width and overlap with net diversification of diploids.
-When diploidization was absent, the net diversification of polyploids overlaps with the net diversification of self-compatible diploids but not with the one of self-incompatible diploids (Figures 3C and 3D).
-For all models, the posterior distribution of diploidization rate is more uncertain than the one of polyploidization rates (See supplementary information figures).
+For breeding system alone, we found a larger net diversification rate for SI and for SC species, in agreement with \citep{goldberg_2010} (I/C model, \cref{figure:netdivall}C).
+When a hidden state is added to the model (I/C+A/B), the large net diversification difference persists for one of the hidden states but is removed for the other (\cref{figure:netdivall}D).
+Thus, differences in net diversification are best explained by both breeding system and an unknown factor.
+The transition rate from SI to SC is $q_{IC}=0.3$ for both the I/C and the I/C+A/B models.
 
-\subsection{Modeling breeding system by itself }
-In the I/C breeding system model we found that the net diversification rate for self-incompatible state is larger than the net diversification rate for self-compatible state (Figure 2(C)).
-The net diversification rate for self-compatible has a probability distribution centered at zero.
-\newline
-When a hidden state is added in the I/C+A/B model, we found that under the hidden state A the self-compatible and self-incompatible net diversification rates are different.
-Under the hidden state B, those two rates overlap.
-Furthermore, the net diversification of self-compatible taxa on hidden state B overlaps with the hidden state rate of self-incompatible taxa on state A  (Figure 2(D)).
-These results agree with previous results found by  \citet{goldberg_2012}.
-However, \citet{goldberg_2012} used a ClaSSE  approach since they were interested in anagenetic and cladogenetic changes for self-incompatibility using a smaller subset of the data presented in the current work.
-Transition rate between incompatible and compatible state are frequent with a rate $q_{IC}=0.3$ for both the  I/C and the I/C+A/B models.
+For ploidy and breeding system together (ID/CD/CP model), the net diversification rate for SI diploids (ID) was greater than for either SC diploids or SC polyploids, with or without diploidization (\cref{figure:netdivall}E, \cref{figure:netdivnodip}C).
+It thus appears that the difference in net diversification with breeding system persists when ploidy is included in the model, but not the reverse.
+The association of ploidy with net diversification (\cref{figure:netdivall}A, \cref{figure:netdivnodip}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids are similar.
+%
+When a hidden state is included (IC/CD/CP+A/B model), the same general pattern remains when diploidization is excluded (\cref{figure:netdivnodip}D), although the higher net diversification rate of ID is less clear within one of the hidden states.
+With diploidization, the net diversification rate of ID is still greater than CD within each hidden state, but diversification for P is highly uncertain and perhaps bimodal.
+
+% E: I moved the discussion of \rho_I and \rho_C values into the Pathways section.
+
+% E note to self: keep editing from here XXX
 
 \subsection{Model selection}
+% E: Maybe incorporate this into the previous subsection?  It could make more sense to organize by question rather than by statistical method (param estimates vs model comparisons).
 
@@
 After testing every single pair of polyploidy models (1-4) we found overwhelming evidence that the best polyploidy model is always the D/P+A/B, that is the model with hidden state and diploidization, that is Bayes factors were always greater than 2 when D/P+A/B was the first model in the Bayes factor equation or negative when D/P+A/B was the second model in Bayes factor equation.
-\newline
-For the two models following the evolution of breeding system, the I/C+A/B is the best choice between models 3 and 4 (BF(IC, IC+A/B)=-39.5<0).\newline
+
+For the two models following the evolution of breeding system, the I/C+A/B is the best choice between models 3 and 4 (BF(IC, IC+A/B)=-39.5<0).
+%
 The four models that follow the diversification linked to both polyploidy and breeding system (models 5-8) where compared in pairs.
-Bayes factors show decisive evidence in favor of the  IC/P/CD+A/B over the rest, meaning that the model that has a hidden state and diploidization is chosen over the ones that lack either of both of those options \cref{table:bayesfactors} .\newline
+Bayes factors show decisive evidence in favor of the IC/P/CD+A/B over the rest, meaning that the model that has a hidden state and diploidization is chosen over the ones that lack either of both of those options \cref{table:bayesfactors}.
+%
 Therefore, the models that were chosen were always the ones containing the hidden trait, or diploidization parameter $\delta$ when ploidy state was considered.
+
+\subsection{Diploidization as an exploratory hypothesis}
+
+For the two models that only consider diploid and polyploid state (D/P and D/P+A/B) the effect of removing parameter $\delta$ was that net diversification rate of polyploids can be negative with high probability (\cref{figure:netdivnodip}AB).
+However, in the models that simultaneously considered polyploidy and breeding system, removing the $\delta$ parameter did not have much of an effect on the location of the posterior distribution of polyploid net diversification rate but on the credible interval width and overlap with net diversification of diploids.
+When diploidization was absent, the net diversification of polyploids overlaps with the net diversification of self-compatible diploids but not with the one of self-incompatible diploids (\cref{figure:netdivnodip}CD).
+For all models, the posterior distribution of diploidization rate is more uncertain than the one of polyploidization rates (see supplementary information figures).
+
+\subsection{Pathways to polyploidy}
+
+% E: If we're going to include pathway results for real, I need to write more in this section and incorporate the figure.  TODO
+
+Polyploidization rates from self-compatible diploid $\rho_{CD}$ and from self-incompatible diploid $\rho_{ID}$ are not significantly different (see supplementary information figures).
+However, the distribution of $\rho_{ID}$ seems to always consider faster values compare to the distribution of $\rho_{CD}$.
+This result is highlighted in ancestral reconstructions of the $ID/CD/CP$ model where transitions from ID to CP are more probable than transitions from CD to CP. % Emma here we put the figure?
 

### 2019-02-25 manuscript/results.tex
- stashing minor suggestions for Results.

@@
 We found that when considered alone, ploidy and breeding system each showed significant associations with net diversification differences.
-When considered together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well.
+%B: check and deleteme: I think this should have "states" added, so "...ploidy and breeding system states showed..."
+When considered together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well. %B: same here.
 
-For ploidy alone (D/P model), we found a larger net diversification rate for diploids than for polyploids, in agreement with \citep{mayrose_2011, mayrose_2015}.
+%B: The "For X (states model), we found" template doesn't sound quite right to me.
+Maybe, "Fitting" or "Considering" or "When" (as in second portion with hidden-state model.
+In latter case, possibly (1) When considering X alone..., (2) Addition of...
+/ Incorporating a hidden state...
+%
+For ploidy alone (D/P model), we found a greater net diversification rate for diploids than for polyploids, in agreement with \citep{mayrose_2011, mayrose_2015}.
 This result holds with (\cref{figure:netdivall}A) or without (\cref{figure:netdivnodip}A) the diploidization parameter, though including diploidization shifts the net diversification rate of polyploids to be non-negative.
@@
 % E: I moved the discussion of \rho_I and \rho_C values into the Pathways section.
-
 % E note to self: keep editing from here XXX

### 2019-02-25 manuscript/methods.tex
editing through model comparisons in Results

@@
 Using the marginal likelihood values, we calculated thirteen different Bayes factors.
-Six compared the models of polyploidy against each other (D/P and D/P+A/B, each with or without diploidization), one compared the breeding system models (I/C and I/C+A/B), and six compared the models with both traits (ID/CD/CP and ID/CD/CP+A/B, each with or without diploidization) (\cref{table:bayesfactors}).
+Six compared the models of ploidy against one other (D/P and D/P+A/B, each with or without diploidization), one compared the breeding system models (I/C and I/C+A/B), and six compared the models with both traits (ID/CD/CP and ID/CD/CP+A/B, each with or without diploidization) (\cref{table:bayesfactors}).
 %
 Other comparisons between these models are not valid because the input data are different under the different state space codings (\cref{figure:stateclassifications}).
-In essence, the D/P, I/C, and ID/CD/CP state spaces are not lumpable with respect to one another \citep{tarasov_2019}.
+In mathematical terms, the D/P, I/C, and ID/CD/CP state spaces are not `lumpable' with respect to one another \citep{tarasov_2019}.
 %R: Emma this is where the figure can be important to show that we have to make different input
@@
 Instead, we want Fig 1 to make it intuitively clear, to Spencer and everyone else.
-Probably need to add to its caption.
+I added to the caption and the text above.
+Okay, or more work needed?
+Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $BF(M_0,\ M_1) = \ln[ P(\mathbf{X} | M_0) - P(\mathbf{X} | M_1]$.
+There is `positive' support for $M_0$ when this value is more than 2, `strong' support when it is more than 6, and `very strong' support when it is more than 10 \citep{kass1995}.
 
@@
 Each species retained on the tree belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known.
-The number of species in each is shown under the corresponding circles in the top row.
+The number of species in each is shown under the corresponding circles in the top row. % FIXME: Rosana, what are the numbers for CP and ?P ?
 These categories were then grouped in a manner appropriate to the states of each model.
@@
 In all cases, species were coded as either $A$ or $B$ in the hidden state models.
+The models depicted in the different rows use state spaces that are not comparable with one another.
+For example, we cannot test whether the D/P model fits better than the I/C model because they use states that are not the same and are not `lumpable' \citep{tarasov_2019}.
 }
@@
 \begin{table}
-\begin{tabular}{@{}llcccccc@{}} \toprule
+\begin{tabular}{lcccccc} \toprule
+% \begin{tabular}{@{}lcccccc@{}} \toprule
 %\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
-Model& Model& Ploidy & Diploidization & Breeding  & Hidden & Parameters & Marginal \\
-& Type & &  &System & State & & Log- Likelihood \\
-1. D/P &BiSSE &	Yes  &	Yes &	No	&No	& 6	& -1182.93 \\
-2. D/P no $\delta$ & BiSSE  &	Yes & 	No	& No	& No & 	5 &	-1193.66\\
-3. D/P- A/B  & HiSSE &	Yes &	Yes	&No &	Yes &	11 &	\textbf{-1145.69}\\
-4. D/P no $\delta$-A/B &	HiSSE &	Yes & No &	No &	Yes &	10	&-1150.99\\
-5. I/C &	BiSSE &	No &	 No	&Yes &	No &	 5 &  -1194.80 \\
-6. I/C-A/B &	HiSSE &	No &	 No	&Yes &	Yes	& 10 & \textbf{-1155.37}\\
-7. ID/P/CD & MuSSE &	Yes & 	Yes &	Yes &	No &	10 & -1344.50\\
-8. ID/P/CD no $\delta$ &	MuSSE &	Yes & 	No &	Yes	&No &	9 &-1345.87\\
-9. ID/P/CD-A/B & MuHiSSE &	Yes 	&Yes &	Yes &	Yes &	16 & \textbf{-1300.35} \\
-10.ID/P/CD no $\delta$-A/B &	 MuHiSSE	 & Yes & 	No	&Yes &	Yes	&15 &-1303.55 \\ \bottomrule
+Model                        & Ploidy & Diploidization & Breeding  & Hidden & Num & Marginal \\
+                             & &  &System & State & Parameters & Log-Likelihood \\ \midrule
+1. D/P                       &	Yes  &	Yes &	No	&No	& 6	& -1182.93 \\
+2. D/P no $\delta$           &	Yes & 	No	& No	& No & 	5 &	-1193.66\\
+3. D/P+A/B                   &	Yes &	Yes	&No &	Yes &	11 &	\textbf{-1145.69}\\
+4. D/P+A/B no $\delta$       &	Yes & No &	No &	Yes &	10	&-1150.99\\
+5. I/C                       &	No & No	&Yes &	No &	 5 &  -1194.80 \\
+6. I/C+A/B                   &	No &	 No	&Yes &	Yes	& 10 & \textbf{-1155.37}\\
+7. ID/CD/CP                  &	Yes & 	Yes &	Yes &	No &	10 & -1344.50\\
+8. ID/CD/CP no $\delta$      &	Yes & 	No &	Yes	&No &	9 &-1345.87\\
+9. ID/CD/CP+A/B              &	Yes 	&Yes &	Yes &	Yes &	16 & \textbf{-1300.35} \\
+10. ID/CD/CP+A/B no $\delta$ & Yes & 	No	&Yes &	Yes	&15 &-1303.55 \\ \bottomrule
 \end{tabular}
-\caption{Marginal likelihoods for the ten models examined.
-States encoded with letter D: diploid, P: polyploid, C: self-compatible, I: self-incompatible, A: first state of hidden trait, B: second state of hidden trait, $\delta$: re-diploidization.}
+\caption{
+    The ten models and their marginal likelihoods.
+    Values in bold are for the best models within each class that are comparable (see \cref{table:bayesfactors}).
+    Abbreviations are D: diploid, P: polyploid, I: self-incompatible, C: self-compatible, A: one state of hidden trait, B: other state of hidden trait, $\delta$: diploidization.}
 \label{table:marginallike}
@@
 2. D/P no $\delta$ &$\cdot$&$\cdot$ &	-47.97	&-42.66\\
-\textbf{3. D/P- A/B}  &$\cdot$  & $\cdot$&	$\cdot$	& 5.30 \\
-4. D/P no $\delta$-A/B &$\cdot$& $\cdot$ & $\cdot$&$\cdot$ \\
+\textbf{3. D/P+A/B}  &$\cdot$  & $\cdot$&	$\cdot$	& 5.30 \\
+4. D/P+A/B no $\delta$ &$\cdot$& $\cdot$ & $\cdot$&$\cdot$ \\
 \end{tabular}
@@
 5. I/C &$\cdot$ & -39.43 \\
-\textbf{6. I/C-A/B} &$\cdot$& $\cdot$ \\
+\textbf{6. I/C+A/B} &$\cdot$& $\cdot$ \\
 & & \\
@@
 & 7 & 8 & 9 & 10\\
-7. ID/P/CD & $\cdot$	&1.36 & -44.15&-40.95 \\
-8. ID/P/CD no $\delta$ & $\cdot$ & $\cdot$ & -45.51 &-42.31 \\
-\textbf{9. ID/P/CD-A/B}& $\cdot$ & $\cdot$ &$\cdot$	& 3.2 \\ 
+7. ID/CD/CP & $\cdot$	&1.36 & -44.15&-40.95 \\
+8. ID/CD/CP no $\delta$ & $\cdot$ & $\cdot$ & -45.51 &-42.31 \\
+\textbf{9. ID/CD/CP+A/B}& $\cdot$ & $\cdot$ &$\cdot$	& 3.2 \\ 
 10. ID/P/CD no $\delta$-A/B & $\cdot$& $\cdot$&$\cdot$ & $\cdot$\\
@@
 \end{tabular}
- \caption{Bayes factors in log scale. We compare every possible pair. Number models as indicated in  \cref{table:marginallike}. }
+% E: We will probably have to change the layout here because it's not technically a table.  But good enough for now.
+\caption{
+    Bayes factors for model comparisons.
+    Each of the three boxes contains models that can be compared with one another, based on the character states they include (see \cref{figure:stateclassifications}).
+    Models are numbered in as \cref{table:marginallike}.
+    Bayes factors are reported on the natural log scale, so numbers greater than $+2$ mean that the model in the row label has `positive' support relative to the model in the column label; numbers less than $-2$ mean that model in the column label is the preferred one.
+    Conventional thresholds for `strong' and `very strong' support are 6 and 10, respectively.
+    The best model in each set is written in bold.
+    In each case, it is the most complex model of the set.
+}
 \label{table:bayesfactors}

### 2019-02-25 manuscript/results.tex
editing through model comparisons in Results

--- 2019-02-25 manuscript/results.tex
@@
 But now the text says things like if some estimates are "different", so perhaps in Methods we should explain different = non-overlapping?
+% For the former model comparison section, I moved the key conclusions into the paragraphs here and the mechanical details (bayes factor definition) into Methods.
 
@@
 
-For ploidy alone (D/P model), we found a larger net diversification rate for diploids than for polyploids, in agreement with \citep{mayrose_2011, mayrose_2015}.
+For ploidy alone (D/P model), we found a larger net diversification rate for diploids than for polyploids, in agreement with \citet{mayrose_2011, mayrose_2015}.
 This result holds with (\cref{figure:netdivall}A) or without (\cref{figure:netdivnodip}A) the diploidization parameter, though including diploidization shifts the net diversification rate of polyploids to be non-negative.
-Incorporating a hidden state, however, removes the clear separation in diversification between diploids and polyploids (D/P+A/B model, \cref{figure:netdivall}B, \cref{figure:netdivnodip}B).
+Incorporating a hidden state, however, removes the clear separation in diversification between diploids and polyploids (D/P+A/B model; \cref{figure:netdivall}B, \cref{figure:netdivnodip}B).
 Thus, differences in net diversification are better explained by an unknown factor than by ploidy.
+Statistical model comparisons show very strong support for including the hidden state and strong support for including diploidization (\cref{table:bayesfactors}).
 % E: I removed the description of relative extinction results.
@@
 
-For breeding system alone, we found a larger net diversification rate for SI and for SC species, in agreement with \citep{goldberg_2010} (I/C model, \cref{figure:netdivall}C).
-When a hidden state is added to the model (I/C+A/B), the large net diversification difference persists for one of the hidden states but is removed for the other (\cref{figure:netdivall}D).
+For breeding system alone, we found a larger net diversification rate for SI than for SC species, in agreement with \citet{goldberg_2010} (I/C model, \cref{figure:netdivall}C).
+When a hidden state is included (I/C+A/B model), the large net diversification difference persists for one of the hidden states but is removed for the other (\cref{figure:netdivall}D).
 Thus, differences in net diversification are best explained by both breeding system and an unknown factor.
-The transition rate from SI to SC is $q_{IC}=0.3$ for both the I/C and the I/C+A/B models.
+% The transition rate from SI to SC is $q_{IC}=0.3$ for both the I/C and the I/C+A/B models.
+The statistical model comparison shows very strong support for including the hidden state (\cref{table:bayesfactors}).
 
-For ploidy and breeding system together (ID/CD/CP model), the net diversification rate for SI diploids (ID) was greater than for either SC diploids or SC polyploids, with or without diploidization (\cref{figure:netdivall}E, \cref{figure:netdivnodip}C).
+For ploidy and breeding system together (ID/CD/CP model), the net diversification rate for SI diploids was greater than for either SC diploids or SC polyploids, with or without diploidization (\cref{figure:netdivall}E, \cref{figure:netdivnodip}C).
 It thus appears that the difference in net diversification with breeding system persists when ploidy is included in the model, but not the reverse.
-The association of ploidy with net diversification (\cref{figure:netdivall}A, \cref{figure:netdivnodip}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids are similar.
+The association of ploidy with net diversification in the D/P model (\cref{figure:netdivall}A, \cref{figure:netdivnodip}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids are similar.
 %
-When a hidden state is included (IC/CD/CP+A/B model), the same general pattern remains when diploidization is excluded (\cref{figure:netdivnodip}D), although the higher net diversification rate of ID is less clear within one of the hidden states.
+When a hidden state is included (IC/CD/CP+A/B model), the same general pattern remains when diploidization is prevented (\cref{figure:netdivnodip}D), although the higher net diversification rate of ID is less clear within one of the hidden states.
 With diploidization, the net diversification rate of ID is still greater than CD within each hidden state, but diversification for P is highly uncertain and perhaps bimodal.
+Statistical model comparisons show very strong support for including the hidden state and at most positive support for including diploidization (\cref{table:bayesfactors}).
 
@@
 % E note to self: keep editing from here XXX
-
-\subsection{Model selection}
-% E: Maybe incorporate this into the previous subsection?  It could make more sense to organize by question rather than by statistical method (param estimates vs model comparisons).
-
-In \cref{table:marginallike} we listed the marginal likelihood values calculated in log scale ($log(P(\mathbf{X}|M)$) for each of the models tested.
-The different components included and excluded for each model as a a summary of the diagrams from \cref{figure:netdivall} are indicated in each model.
-From the marginal likelihoods in log scale, the Bayes factors in log-scale($BF(M_0,M_1)=log(P(\mathbf{X}|M_0)-P(\mathbf{X}|M_0)$) were calculated as shown in table \cref{table:bayesfactors}.
-After testing every single pair of polyploidy models (1-4) we found overwhelming evidence that the best polyploidy model is always the D/P+A/B, that is the model with hidden state and diploidization, that is Bayes factors were always greater than 2 when D/P+A/B was the first model in the Bayes factor equation or negative when D/P+A/B was the second model in Bayes factor equation.
-
-For the two models following the evolution of breeding system, the I/C+A/B is the best choice between models 3 and 4 (BF(IC, IC+A/B)=-39.5<0).
-%
-The four models that follow the diversification linked to both polyploidy and breeding system (models 5-8) where compared in pairs.
-Bayes factors show decisive evidence in favor of the IC/P/CD+A/B over the rest, meaning that the model that has a hidden state and diploidization is chosen over the ones that lack either of both of those options \cref{table:bayesfactors}.
-%
-Therefore, the models that were chosen were always the ones containing the hidden trait, or diploidization parameter $\delta$ when ploidy state was considered.
 
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{Netdiversificationallmodels.pdf}
+\includegraphics[width=\textwidth]{Netdiversificationallmodels.pdf} % FIXME: The panel labels should use "+A/B" instead of "-A/B".
+Rosana, if you add the .svg file for this figure, I can make small changes like this.
 \caption{Net diversification rates for all models that include diploidization.}  
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{netdiversificationallnodip.pdf}
+\includegraphics[width=\textwidth]{netdiversificationallnodip.pdf} % FIXME: same note as above about "+A/B"
 \caption{Net diversification rates for all models that do not include diploidization.}  
@@
 \end{suppfigure}
-
-% E or R todo: which pathway from SI-D to SC-P is more common, based on rate estimates?  apply procedure from robertson2011 to rates from ID/CD/CP model

### 2019-02-25 manuscript/results.tex

@@
 We found that when considered alone, ploidy and breeding system each showed significant associations with net diversification differences.
-When considered together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well.
+%B: check and deleteme: I think this should have "states" added, so "...ploidy and breeding system states showed..."
+When considered together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well. %B: same here.
 
-For ploidy alone (D/P model), we found a larger net diversification rate for diploids than for polyploids, in agreement with \citet{mayrose_2011, mayrose_2015}.
+%B: The "For X (states model), we found" template doesn't sound quite right to me.
+Maybe, "Fitting" or "Considering" or "When" (as in second portion with hidden-state model.
+In latter case, possibly (1) When considering X alone..., (2) Addition of...
+/ Incorporating a hidden state...
+%
+For ploidy alone (D/P model), we found a greater net diversification rate for diploids than for polyploids, in agreement with \citep{mayrose_2011, mayrose_2015}.
 This result holds with (\cref{figure:netdivall}A) or without (\cref{figure:netdivnodip}A) the diploidization parameter, though including diploidization shifts the net diversification rate of polyploids to be non-negative.

### 2019-02-25 manuscript/results.tex
wording edit
Fixing a request from Boris that snuck in when I wasn't looking.

@@
 
-%B: The "For X (states model), we found" template doesn't sound quite right to me.
-Maybe, "Fitting" or "Considering" or "When" (as in second portion with hidden-state model.
-In latter case, possibly (1) When considering X alone..., (2) Addition of...
-/ Incorporating a hidden state...
-%
-For ploidy alone (D/P model), we found a greater net diversification rate for diploids than for polyploids, in agreement with \citep{mayrose_2011, mayrose_2015}.
+When considering ploidy alone (D/P model), we found a greater net diversification rate for diploids than for polyploids, in agreement with \citep{mayrose_2011, mayrose_2015}.
 This result holds with (\cref{figure:netdivall}A) or without (\cref{figure:netdivnodip}A) the diploidization parameter, though including diploidization shifts the net diversification rate of polyploids to be non-negative.
@@
 
-For breeding system alone, we found a larger net diversification rate for SI than for SC species, in agreement with \citet{goldberg_2010} (I/C model, \cref{figure:netdivall}C).
+When considering breeding system alone, we found a larger net diversification rate for SI than for SC species, in agreement with \citet{goldberg_2010} (I/C model, \cref{figure:netdivall}C).
 When a hidden state is included (I/C+A/B model), the large net diversification difference persists for one of the hidden states but is removed for the other (\cref{figure:netdivall}D).
@@
 
-For ploidy and breeding system together (ID/CD/CP model), the net diversification rate for SI diploids was greater than for either SC diploids or SC polyploids, with or without diploidization (\cref{figure:netdivall}E, \cref{figure:netdivnodip}C).
+When considering ploidy and breeding system together (ID/CD/CP model), the net diversification rate for SI diploids was greater than for either SC diploids or SC polyploids, with or without diploidization (\cref{figure:netdivall}E, \cref{figure:netdivnodip}C).
 It thus appears that the difference in net diversification with breeding system persists when ploidy is included in the model, but not the reverse.
@@
 % E: I moved the discussion of \rho_I and \rho_C values into the Pathways section.
+
 % E note to self: keep editing from here XXX

### 2019-02-25 manuscript/results.tex
more editing in Results (diploidization)

@@
 
-% E note to self: keep editing from here XXX
-
 \subsection{Diploidization as an exploratory hypothesis}
 
-For the two models that only consider diploid and polyploid state (D/P and D/P+A/B) the effect of removing parameter $\delta$ was that net diversification rate of polyploids can be negative with high probability (\cref{figure:netdivnodip}AB).
-However, in the models that simultaneously considered polyploidy and breeding system, removing the $\delta$ parameter did not have much of an effect on the location of the posterior distribution of polyploid net diversification rate but on the credible interval width and overlap with net diversification of diploids.
-When diploidization was absent, the net diversification of polyploids overlaps with the net diversification of self-compatible diploids but not with the one of self-incompatible diploids (\cref{figure:netdivnodip}CD).
-For all models, the posterior distribution of diploidization rate is more uncertain than the one of polyploidization rates (see supplementary information figures).
+We considered models both with or without diploidization in order to explore the effects of this process on the estimates of state-dependent diversification.
+For the two models that only include diploid and polyploid states (D/P and D/P+A/B), the net diversification rate of polyploid lineages is likely negative when diploidization (parameter $\delta$) is excluded but positive when $\delta$ is included (\cref{figure:netdivall}A versus \cref{figure:netdivnodip}A).
+For the two models that also included breeding system (ID/CD/CP and ID/CD/CP+A/B), the main effect of including diploidization is increasing the uncertainty of the estimate of polyploid net diversification rate (\cref{figure:netdivall}EF versus \cref{figure:netdivnodip}CD).
+For all models, there is greater uncertainty in the estimate of diploidization rate than polyploidization rate, as judged by the width of the credibility intervals (see supplementary information figures).
+% todo: precise figure refs here
+% E: Rosana, would this also be a good place to say something about what we learned regarding the meaningfulness of diploidization in these data?
 

### 2019-02-25 manuscript/results.tex
- suggestions in Results - small modification to supp table legend in Results - check author info note!

@@
 We found that when considered alone, ploidy and breeding system each showed significant associations with net diversification differences. 
-%B: check and deleteme: I think this should have "states" added, so "...ploidy and breeding system states showed..."
-When considered together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well. %B: same here.
+When considered together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well.
 
@@
 
-When considering breeding system alone, we found a larger net diversification rate for SI than for SC species, in agreement with \citet{goldberg_2010} (I/C model, \cref{figure:netdivall}C).
+When considering breeding system alone (I/C model, \cref{figure:netdivall}C), we found a larger net diversification rate for SI than for SC species, in agreement with \citet{goldberg_2010}.
+%B: moved \cref{} earlier; avoids seeming related to \citep{}
 When a hidden state is included (I/C+A/B model), the large net diversification difference persists for one of the hidden states but is removed for the other (\cref{figure:netdivall}D).
@@
 % E: Rosana, would this also be a good place to say something about what we learned regarding the meaningfulness of diploidization in these data?
+%B: maybe meaning of the data could go to a dedicated Discussion subsection?
+The idea would be to reconcile variation in $\delta$ across (the best-fitting) models considered here, and how the implied values do (or do not) agree with chromosome number changes and WGD inference from tree reconciliation and Ks approaches.
+(I can help here, if needed.)
 
@@
 This result is highlighted in ancestral reconstructions of the $ID/CD/CP$ model where transitions from ID to CP are more probable than transitions from CD to CP. % Emma here we put the figure?
+%B: I am not sure I understand the "However, " sentence. TODO.
 
@@
     \begin{center}
+
+    \begin{tabular}{lcccc}
+                       & D/P   & D/P+A/B      & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
+        %                                                              
+        $r_D$          & 0.260 & 0.193, 0.658 & ---      & ---          \\
+        $r_P$          &-0.056 & 0.030, 0.187 & ---      & ---          \\
+        $r_I$          & ---   & ---          & ---      & ---          \\
+        $r_C$          & ---   & ---          & ---      & ---          \\
+        $r_{ID}$       & ---   & ---          & 0.455    & 0.309, 0.797 \\
+        $r_{CD}$       & ---   & ---          & 0.065    &-0.006, 0.587 \\
+        $r_{CP}$       & ---   & ---          &-0.088    &-0.074, 0.403 \\
+        %                                                              
+        $\rho$         & 0.047 & 0.047        & ---      & ---          \\
+        $\rho_I$       & ---   & ---          & 0.067    & 0.053        \\
+        $\rho_C$       & ---   & ---          & 0.033    & 0.032        \\
+        $q_{IC}$       & ---   & ---          & 0.198    & 0.145        
+    \end{tabular}
+
+            \vspace{12pt}
+
     \begin{tabular}{lcccccc}
@@
 
-    \vspace{12pt}
-    \begin{tabular}{lcccc}
-                       & D/P   & D/P+A/B      & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
-        %                                                              
-        $r_D$          & 0.260 & 0.193, 0.658 & ---      & ---          \\
-        $r_P$          &-0.056 & 0.030, 0.187 & ---      & ---          \\
-        $r_I$          & ---   & ---          & ---      & ---          \\
-        $r_C$          & ---   & ---          & ---      & ---          \\
-        $r_{ID}$       & ---   & ---          & 0.455    & 0.309, 0.797 \\
-        $r_{CD}$       & ---   & ---          & 0.065    &-0.006, 0.587 \\
-        $r_{CP}$       & ---   & ---          &-0.088    &-0.074, 0.403 \\
-        %                                                              
-        $\rho$         & 0.047 & 0.047        & ---      & ---          \\
-        $\rho_I$       & ---   & ---          & 0.067    & 0.053        \\
-        $\rho_C$       & ---   & ---          & 0.033    & 0.032        \\
-        $q_{IC}$       & ---   & ---          & 0.198    & 0.145        
-    \end{tabular}
     \end{center}
     \caption{
-        Median rate estimates for all models.
+        Median rate estimates for all fitted models. %B: "...all fitted/considered/evaluated models."
         Units are per million years.
         Two comma-separated numbers refer to the $A$ and $B$ hidden states, and --- means the parameter was not present in the model.
-        The upper section is for models with diploidization, and the lower section is for models without.
+    %B: perhaps include this in the table so lazies don't have to go seeking subscripts?
+    Net diversification rates ($r$) are subscripted with trait state initials (Diploid, Polyploid, Incompatible, Compatible).
+Transition rates are $\rho$ (polyploidization), subscripted with background breeding system state; $\delta$ (diploidization); and  $q_{IC}$ (loss of self-incompatibility).
+    The upper section is for models without diploidization, and the lower section is for models with diploidization.
     The supplemental figures show the corresponding distributions of parameter estimates.
@@
 \end{supptable}
+%B: The I/C and I/C+A/B columns should be in the top panel, no (they don't have rho)? I'll do this in a separate doc first. %FIXME
 
@@
 \includegraphics[width=\textwidth]{bisseDPposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P polyploidy model} % XXX
+\caption{Posterior distribution for each of the parameters in the D/P, polyploidy model} % XXX
 \label{suppfigure:DP}
@@
 \includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P no $\delta$ polyploidy model} % XXX
+\caption{Posterior distribution for each of the parameters in the D/P no $\delta$, polyploidy model} % XXX
 \label{suppfigure:DPnodip}
@@
 \includegraphics[width=\textwidth]{hisseDPposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P-A/B polyploidy model} % XXX
+\caption{Posterior distribution for each of the parameters in the D/P+A/B, polyploidy model} % XXX
 % fixme: offset plotting error in "relative extinction" panel
@@
 \includegraphics[width=\textwidth]{hisseDPnodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P no $\delta$-A/B polyploidy model} % XXX
+\caption{Posterior distribution for each of the parameters in the D/P no $\delta$+A/B, polyploidy model} % XXX
 \label{suppfigure:DPnodipAB}
@@
 \includegraphics[width=\textwidth]{bisseSIposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the I/C breeding system model} % XXX
+\caption{Posterior distribution for each of the parameters in the I/C, breeding system model} % XXX
 \label{suppfigure:IC}
@@
 \includegraphics[width=\textwidth]{hisseSInoretposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the I/C-A/B breeding system model} % XXX
+\caption{Posterior distribution for each of the parameters in the I/C+A/B, breeding system model} % XXX
 \label{suppfigure:ICAB}
@@
 \includegraphics[width=\textwidth]{musseDPSIposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the ID/CD/CP polyploidy and breeding system model} % XXX
+\caption{Posterior distribution for each of the parameters in the ID/CD/CP, polyploidy and breeding system model} % XXX
 \label{suppfigure:IDCDCP}
@@
 \includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the ID/CD/CP no $\delta$ polyploidy and breeding system model} % XXX
+\caption{Posterior distribution for each of the parameters in the ID/CD/CP no $\delta$, polyploidy and breeding system model} % XXX
 \label{suppfigure:IDCDCPnodip}
@@
 \includegraphics[width=\textwidth]{muhisseDPSIposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the ID/CD/CP-A/B polyploidy and breeding system model} % XXX
+\caption{Posterior distribution for each of the parameters in the ID/CD/CP+A/B, polyploidy and breeding system model} % XXX
 \label{suppfigure:IDCDCPAB}
@@
 \includegraphics[width=\textwidth]{muhisseDPSInodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the ID/CD/CP no $\delta$- A/B polyploidy and breeding system model} % XXX
+\caption{Posterior distribution for each of the parameters in the ID/CD/CP no $\delta$+A/B polyploidy and breeding system model} % XXX
 \label{suppfigure:IDCDCPnodipAB}

### 2019-02-26 manuscript/results.tex
- B&E phone conversation notes in Discussion and Methods

@@
 
-We found that when considered alone, ploidy and breeding system each showed significant associations with net diversification differences. 
+%B: how about below; was: We found that when considered alone, ploidy and breeding system each showed significant associations with net diversification differences. 
+Individual trait-dependent diversification inference, of ploidy and breeding system considered separately, each showed significant associations with net diversification differences. 
 When considered together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well.
@@
 \end{supptable}
-%B: The I/C and I/C+A/B columns should be in the top panel, no (they don't have rho)? I'll do this in a separate doc first. %FIXME
+
+%B: FIXME:
+%B: Table S2 (above) The I/C and I/C+A/B columns should maybe appear in the top panel, no?
+%B: I did not flip the panels, because I realized just now that we'll first discuss no-delta, then delta. Is that okay?
+%
+% check values of delta and rho D/P+A/B with(out) delta
+% i.e. parameter values in figure S1 & S4 vs Table S2
 

### 2019-02-27 manuscript/discussion.tex
- B&E phone conversation notes in Discussion and Methods
a small part of the discussion

@@
 
-% diversification
+% - value of approach
+The present work shows the importance of considering the trait linked diversification patterns under a multivariate approach.
+Species are created and go extinct based un multiple and often highly correlated phenotypes, understanding the speciation and extinction processes requires understanding of the evolutionary consequences that those trait correlations produce in organisms.
+In the present work we show how considering both polyploidy and breeding system can disentangle the importance (or the lack of) of polyploidy when confronted with the evidence brought by breeding system.\newline
+
+% - what happens if PD only, Hidden states
+Using the most complete dataset for polyploidy in a phylogenetic tree in Solanaceae, we were able to replicate the results found by \citet{mayrose_2011}, polyploids have a slower net diversification compare to diploids.
+Furthermore,  we also found polyploids had a high probability of having negative diversification which implies that polyploids can become a macroevolutionary dead-end, a result that was also found in the two large angiosperm diversification studies  \citet{mayrose_2011} and \citet{mayrose_2015}.
+However, we expanded this study to accomodate  background heterogeneity in the diversification process.
+When adding heterogeneity we found that it was more likely that an unobserved trait linked to diploid state was the one leading the net diversification patterns, and that there were some "second-class"  diploids that were not different from  polyploids  in diversification terms (Figure 2A).
+This result lead us to our central question: \textit{what is that other trait linked diploids that makes them different in the diversification process?
+}.\newline
+
 %     main finding about our traits
+In Solanaceae, our immediate intuition was to look into breeding system.
+Previous studies shown that self-incompatible Solanaceae species have also higher rates of diversification compared to their self-compatible counterparts\ctiep{goldberg_2012}.
+Self-incompatible species are diploid in our sample and also expected to be diploid due to ...
+(citation?).
+By considering both polyploidy and breeding system simultaneously for every species in our sample, it was possible to disentangle why some diploids were quantitatively different than polyploids.
+In the three-state diversification model ID/CD/CP, we found that self-incompatible diploids have faster and positive rates than self-compatible diploids and polyploids, and that the difference between the  rates of net diversification of self-compatible diploids is not as large (Figures 2C) as first found by binary trait diversification models (Figures 2A).
+This result is important, since it aligns with the net diversification results of the  D/P+A/B model, where a "hidden-trati" seem to be dictating the diversification pattern.
+By adding breeding system, we were able to hint at which that hidden-trait possibly is.
+Therefore, we consider that finding a heterogenous result in the hidden trait approaches should be be treated as evidence of a second trait that is necessary to consider.
+Pursuing knowledge of  such trait can result on a clearer picture of the importance of trait linked diversification patterns, but also on a better reconstruction on past events in phylogenies.
+
+% - diversification  (frame as no-delta first)
+
 %     more generally, not okay to approximate musse with bisse when states are correlated (cf Pyron)
-
-% diploidization: comment
-%     effect on inferred diversification
-
-% pathways: compare/contrast with robertson2011
-%     \rho_I > \rho_C in all!
 
@@
 
+% pathways: compare/contrast with robertson2011
+%     \rho_I > \rho_C in all!
+
+% where WGD is in Solanaceae--motivation and delta rate analyses
+% diploidization: comment
+%     effect on inferred diversification
+
+% generality of approach
+%
 % issues

### 2019-02-28 manuscript/discussion.tex
- line breaks in Discussion for easy merging, and small typo edits

--- 2019-02-18 manuscript/discussion.tex
@@
 
-% diversification
+% - value of approach
+The present work shows the importance of considering the trait linked diversification patterns under a multivariate approach.
+Species are created and go extinct based un multiple and often highly correlated phenotypes, understanding the speciation and extinction processes requires understanding of the evolutionary consequences that those trait correlations produce in organisms
+In the present work we show how considering both polyploidy and breeding system can disentangle the importance (or the lack of) of polyploidy when confronted with the evidence brought by breeding system.\newline
+
+% - what happens if PD only, Hidden states
+Using the most complete dataset for polyploidy in a phylogenetic tree in Solanaceae, we were able to replicate the results found by \citet{mayrose_2011}, polyploids have a slower net diversification compare to diploids
+Furthermore,  we also found polyploids had a high probability of having negative diversification which implies that polyploids can become a macroevolutionary dead-end, a result that was also found in the two large angiosperm diversification studies  \citet{mayrose_2011} and \citet{mayrose_2015}
+However, we expanded this study to accomodate  background heterogeneity in the diversification process
+When adding heterogeneity we found that it was more likely that an unobserved trait linked to diploid state was the one leading the net diversification patterns, and that there were some ``second-class" diploids that were not different from polyploids in diversification terms (Figure 2A)
+This result lead us to our central question: \textit{what is that other trait linked diploids that makes them different in the diversification process? }.\newline
+
 %     main finding about our traits
+In Solanaceae, our immediate intuition was to look into breeding system
+Previous studies shown that self-incompatible Solanaceae species have also higher rates of diversification compared to their self-compatible counterparts\ctiep{goldberg_2012}
+Self-incompatible species are diploid in our sample and also expected to be diploid due to ..
+(citation?)
+By considering both polyploidy and breeding system simultaneously for every species in our sample, it was possible to disentangle why some diploids were quantitatively different than polyploids
+In the three-state diversification model ID/CD/CP, we found that self-incompatible diploids have faster and positive rates than self-compatible diploids and polyploids, and that the difference between the  rates of net diversification of self-compatible diploids is not as large (Figures 2C) as first found by binary trait diversification models (Figures 2A)
+This result is important, since it aligns with the net diversification results of the  D/P+A/B model, where a ``hidden-trait" seem to be dictating the diversification pattern
+By adding breeding system, we were able to hint at which that hidden-trait possibly is
+Therefore, we consider that finding a heterogenous result in the hidden trait approaches should be be treated as evidence of a second trait that is necessary to consider
+Pursuing knowledge of  such trait can result on a clearer picture of the importance of trait linked diversification patterns, but also on a better reconstruction on past events in phylogenies.
+
+% - diversification  (frame as no-delta first)
+
 %     more generally, not okay to approximate musse with bisse when states are correlated (cf Pyron)
-
-% diploidization: comment
-%     effect on inferred diversification
-
-% pathways: compare/contrast with robertson2011
-%     \rho_I > \rho_C in all!
 

### 2019-02-28 manuscript/discussion.tex
- a variety of small comment deletions, comments and changes (mainly formatting, etc.)

@@
 
+% - diversification  (frame as no-delta first)
 % - what happens if PD only, Hidden states
@@
 In Solanaceae, our immediate intuition was to look into breeding system
-Previous studies shown that self-incompatible Solanaceae species have also higher rates of diversification compared to their self-compatible counterparts\ctiep{goldberg_2012}
+Previous studies shown that self-incompatible Solanaceae species have also higher rates of diversification compared to their self-compatible counterparts\citep{goldberg_2012}
 Self-incompatible species are diploid in our sample and also expected to be diploid due to ..
@@
 Pursuing knowledge of  such trait can result on a clearer picture of the importance of trait linked diversification patterns, but also on a better reconstruction on past events in phylogenies.
-
-% - diversification  (frame as no-delta first)
 

### 2019-02-28 manuscript/main.tex
- suggestions in Results - small modification to supp table legend in Results - check author info note!
- a variety of small comment deletions, comments and changes (mainly formatting, etc.)

@@
 
-%B: I included extended affiliations necessary at submission time. Check, please! 
+%B: I included extended affiliations necessary at submission time. Check, please! TODO
 %B->R you may wish to add new affiliation at Manoa, but I left the original state alone.
 \noindent
-Rosana Zenil-Ferguson\endnote{Department of Ecology, Evolution, and Behavior, University of Minnesota, Saint Paul, MN 55108, U.S.A.}$^{\dagger}$,
+Rosana Zenil-Ferguson\endnote{Department of Ecology, Evolution, and Behavior, University of Minnesota, Saint Paul, MN 55108, U.S.A.}$^{,\dagger}$,
 %
@@
 The Minnesota Supercomputing Institute (MSI) at the University of Minnesota provided computing resources for this project.
+% Perhaps other's funding sources for max BSF/NSF payoff.
 

### 2019-02-28 manuscript/methods.tex
notes while talking to Rosana
- a variety of small comment deletions, comments and changes (mainly formatting, etc.)

@@
 Breeding system was scored as self-incompatible (I) or self-compatible (C) based on results curated from the literature and  original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
-Most species could unambiguously be coded as either I or C \citep{raduski_2012}.
+Most species could unambiguously be coded as either I or C \citep{raduski_2012}. %B: most? Any exceptions?
 Following previous work, we coded as I any species with functional I systems, even if C or dioecy was also reported.
@@
 
-%B: It would be nice to clarify what was added--what is new--and separate the subsections of ploidy and breeding system data.
-%
 % E: I am trying to stick with I/C when talking about the state values, but using SI/SC when talking about the meaning of the trait (because this what people use normally).
@@
 %B: I added the non-self recognition bit, because in Prunus, heteroallelic pollen may not cause SC.
-% E: Should we include more details about how many polymorphic species and how the particularly tricky ones were dealt with? %B: No and yes, maybe. Interested parties can go to data for part 1?
+% E: Should we include more details about how many polymorphic species and how the particularly tricky ones were dealt with? %B: No and yes, maybe.
 
@@
 % ID/CD/CP+A/B ploidy, breeding system, and hidden state model
+
+% NOTE: Should subtly rephrase so that the default models lack diploidization.
 
@@
 %B: we moved back from "SI" to "self-incompatibility" below. Maybe a no-no.
-For all breeding system models, we allow transitions from $I$ to $C$ (at rate $q_{IC}$) but not the reverse. %B: was "these and the subsequent models including breeding systems"
+For all breeding system models, we allow transitions from $I$ to $C$ (at rate $q_{IC}$) but not the reverse.
 Within Solanaceae, self-incompatibility is homologous in all species in which S-alleles were cloned, and controlled crosses performed.  
-All species sampled to date, possess a non-self recognition, RNase-based, gametophytic self-incompatibility \citep[shared even shared even with other euasterid families;][]{ramanauskas_2017}.
-%B: this now makes a narrower claim than igic_2001/steinbachs_2002, which includes Rosaceae and a bag of avoidable troubles.
+All species sampled to date, possess a non-self recognition, RNase-based, gametophytic self-incompatibility \citep[shared even with other euasterid families;][]{ramanauskas_2017}.
 Furthermore, species that are distantly related within this family carry closely-related alleles, with deep trans-specific polymorphism, at the S-locus, which controls the SI response \citep{ioerger_1990, igic_2006}.
@@
 % This irreversibility assumption prevents self-compatible state from evolving into self-incompatible state and as a result, the ancestral state of all taxa in this model has probability one of being self-incompatible.
-
+%
 %B: What happens if SC->SI is not fixed? Asking for a friend.  E: Surprisingly, Rosana reports, results are not turned on their head. 
+%B: Whew, I was worried. Er, "friend" says thanks.
 
@@
 If ploidy and breeding system each influence lineage diversification individually, it is logical to examine their possible joint effects.
-%B: unsure about edit, which was: "If both ploidy and breeding system have the potential to influence lineage diversification, we should logically consider their effects jointly."
+%B: unsure about edit, which was: "If both ploidy and breeding system have the potential to influence lineage diversification, we should logically consider their effects jointly."  E: okay; also try to add that their transitions interact
 We thus fit a multi-state model that includes both traits \citep[MuSSE,][]{fitzjohn_2012}.
@@
 The ID/CD/CP model could potentially capture similar dynamics as earlier models, if the effects of the hidden state in D/P+A/B were effectively caused by breeding system (or its correlates), and the hidden state in I/C+A/B was effectively caused by ploidy.
-% B: I know that causation is not strict here, but it clarifies the intended biological explanation.
 There is also the potential, however, for a hidden factor to be influencing diversification beyond both of our focal traits, and this could again mislead inferences.
@@
 
-%B: this section needs a motivation.
-Right now, that appears to be hung on an argument from a citation (mayrose_2010), which is not clarified.
-I think you guys may have the paleo-WGD literature in mind here.
-If so, perhaps simply stating something ought to be written about the expectations for capturing some underlying process we think is ongoing.
 For all four models that consider ploidy changes, we allowed diploidization.
@@
 % robertson2011 method, and then with diversification
+% propogation matrix, stable stage structure
+% avoid "probability" and "transition matrix"
+% analogous to Leslie matrix?
 
@@
 % fixme: eventually, add supp info fig ref
+
+% R will add a bit about the ASR methods
 
@@
 % E: I think this short explanation is clear enough, especially with the figure.  I'm not sure whether the hierarchical concept is exactly applicable here.
-%B: my TeX GUI (and my brain) underlined "lumpable", especially in the context of "with respect to one another." I think I get the gist, but I am not sure that I agree others will easily.
-I find it useful to imagine Spencer reading this, for example.
-% E: "Lumpable" is the math word for a particular relation among state codings, used in that Tarasov paper.
-I don't think we want to get into defining it precisely.
-Instead, we want Fig 1 to make it intuitively clear, to Spencer and everyone else.
-I added to the caption and the text above.
-Okay, or more work needed?
 Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $BF(M_0,\ M_1) = \ln[ P(\mathbf{X} | M_0) - P(\mathbf{X} | M_1]$.
@@
 For example, we cannot test whether the D/P model fits better than the I/C model because they use states that are not the same and are not `lumpable' \citep{tarasov_2019}.
+%B: Hmmm.
+The presentation is economical, true, and I think it is retrospectively clear, and valuable, now that I've been thinking about the paper for a while.
+But it will be very difficult as a source of clarity for someone reading the paper the first time.
+The top graphical layout also implicitly communicates the idea that all "I" species must be "D", and that all "P" species must be also "C", which is not intuited easily.
+(Are they encoded like that?) Moreover, it's not made clear that the 'species' in question are, in fact, often polymorphic, which makes DPIC possible.
+The subsequent models require a bit of thought, which is not too bad, but the treatment of the state coding and specifically the sentence "In all cases, species were coded as either A or B in the hidden state models." is a bit terse for someone not familiar with HiSSE.
 }

### 2019-02-28 manuscript/discussion.tex
- re-do discussion.tex break removal - Discussion is back to 012641f1b15c6d5c24e3e96251b04fe67fac0128 - typos and minor changes to Discussion (only)

--- 2019-02-27 manuscript/discussion.tex
@@
 Species are created and go extinct based un multiple and often highly correlated phenotypes, understanding the speciation and extinction processes requires understanding of the evolutionary consequences that those trait correlations produce in organisms.
-In the present work we show how considering both polyploidy and breeding system can disentangle the importance (or the lack of) of polyploidy when confronted with the evidence brought by breeding system.\newline
+In the present work we show how considering both polyploidy and breeding system can disentangle the importance (or the lack of) of polyploidy when confronted with the evidence brought by breeding system.
 
@@
 Furthermore,  we also found polyploids had a high probability of having negative diversification which implies that polyploids can become a macroevolutionary dead-end, a result that was also found in the two large angiosperm diversification studies  \citet{mayrose_2011} and \citet{mayrose_2015}.
-However, we expanded this study to accomodate  background heterogeneity in the diversification process.
-When adding heterogeneity we found that it was more likely that an unobserved trait linked to diploid state was the one leading the net diversification patterns, and that there were some "second-class"  diploids that were not different from  polyploids  in diversification terms (Figure 2A).
-This result lead us to our central question: \textit{what is that other trait linked diploids that makes them different in the diversification process?
-}.\newline
+However, we expanded this study to accommodate  background heterogeneity in the diversification process.
+When adding heterogeneity we found that it was more likely that an unobserved trait linked to diploid state was the one leading the net diversification patterns, and that there were some ``second-class"  diploids that were not different from  polyploids  in diversification terms (Figure 2A).
+This result lead us to our central question: \textit{what is that other trait linked diploids that makes them different in the diversification process? }.
 
@@
 In Solanaceae, our immediate intuition was to look into breeding system.
-Previous studies shown that self-incompatible Solanaceae species have also higher rates of diversification compared to their self-compatible counterparts\ctiep{goldberg_2012}.
-Self-incompatible species are diploid in our sample and also expected to be diploid due to ...
-(citation?).
+Previous studies shown that self-incompatible Solanaceae species have also higher rates of diversification compared to their self-compatible counterparts \citep{goldberg_2012}.
+Self-incompatible species are diploid in our sample and also expected to be diploid due to ... (citation?).
 By considering both polyploidy and breeding system simultaneously for every species in our sample, it was possible to disentangle why some diploids were quantitatively different than polyploids.
 In the three-state diversification model ID/CD/CP, we found that self-incompatible diploids have faster and positive rates than self-compatible diploids and polyploids, and that the difference between the  rates of net diversification of self-compatible diploids is not as large (Figures 2C) as first found by binary trait diversification models (Figures 2A).
-This result is important, since it aligns with the net diversification results of the  D/P+A/B model, where a "hidden-trati" seem to be dictating the diversification pattern.
+This result is important, since it aligns with the net diversification results of the  D/P+A/B model, where a `	`hidden-trait" seem to be dictating the diversification pattern.
 By adding breeding system, we were able to hint at which that hidden-trait possibly is.

### 2019-02-28 manuscript/methods.tex
first draft of methods and results for pathways

@@
 
+\subsection{Pathways to polyploidy}
+
+Considering ploidy and breeding system together, there are two evolutionary pathways from SI diploid to SC polyploid \citep{brunet2001, robertson_2011}.
+In the one-step pathway, the CP state is produced directly from the ID state when whole genome duplication disables SI.
+In the two-step pathway, the CD state is an intermediate: SI is first lost, and later the SC diploid undergoes polyploidization.
+We quantify the relative contribution of these pathways to polyploidy in two ways, each using the MAP rate estimates from the IC/CD/CP model.
+
+Both of our methods are based on a propogation matrix that describes flow from ID to CP, as in \citet{robertson_2011}.
+We insert an artificial division in the CP state, so that one substate contains the CP species that arrived via the one-step pathway and the other substate contains the CP species that arrived via the two-step pathway.
+We consider only unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and not strong support for diploidization.
+First, we consider only the rates of transitions between these states, placing them in the propogation matrix \myvec{Q}.
+The matrix $\myvec{P} = \exp(\myvec{Q} t)$ then provides the probabilities of changing from one state to any other state after time $t$.
+Closed-form solutions for the two pathway probabilities are provided in \citet{robertson_2011}.
+Our results will differ from those of \citet{robertson_2011} because our transition rate estimates come from a dated phylogeny and a model that allows for state-dependent diversification.
+%
+Second, we consider not only transitions between states but also diversification within each state.
+State-dependent diversification can change the relative contributions of the two pathways.
+In particular, if the net diversification rate is small for CD, the two-step pathway will contribute relatively less.
+We therefore include the difference between speciation and extinction along the diagonal elements of the propogation matrix.
+As before matrix exponentiation provides the relative chance of changing from one state to any other state after time $t$.
+In this case, however, these are not probabilities because diversification changes the number of lineages as time passes.
+We can still use their ratios to consider the relative contribution of each pathway, though, analogous to the normalized age structure in a growing population.
+
+% E todo: for the case with net div, write out the propogation matrix and equations for pathway contributions
+
 \subsection{Diploidization as an exploratory hypothesis}
@@
 In our case the diploidization parameter (or its absence, $\delta=0$) in our models is an opportunity to explore an assumption that might be important but that is not the single definitive process to understand the interactions among polyploidy, breeding system, and diversification.
-
-% \subsection{Pathways to polyploidy} % E todo
-% robertson2011 method, and then with diversification
-% propogation matrix, stable stage structure
-% avoid "probability" and "transition matrix"
-% analogous to Leslie matrix?
 

### 2019-02-28 manuscript/results.tex
notes while talking to Rosana
- a variety of small comment deletions, comments and changes (mainly formatting, etc.)
first draft of methods and results for pathways

@@
 \subsection{Traits and diversification}
+
+% NOTE: will need to update figure/panel refs if making no-delta models primary
 
@@
 
-%B: how about below; was: We found that when considered alone, ploidy and breeding system each showed significant associations with net diversification differences. 
 Individual trait-dependent diversification inference, of ploidy and breeding system considered separately, each showed significant associations with net diversification differences. 
@@
 
-% E: I moved the discussion of \rho_I and \rho_C values into the Pathways section.
+\subsection{Pathways to polyploidy}
+
+Evolution from SI diploids to SC polyploids can proceed through two different pathways.
+Determining the relative contribution of these pathways based on our estimated transition rates from the ID/CD/CP model, we find that the one-step pathway is more likely on short timescales and the two-step pathway is more likely on long timescales (\cref{figure:pathways}, left panels).
+When not much time has elapsed, the one-step pathway is more likely because only one event need happen.
+When more time has elapsed, the two-step pathway is more likely because the rate of loss of SI within diploids, $q_{IC}$, is greater than the rate of polyploidization for SI species, $\rho_I$ (\cref{suppfigure:IDCDCPnodip}).
+These conclusions are the same as those of \citet{robertson_2011} under one of their branch length approximations.
+
+When rates of lineage diversification are considered, however, the conclusions change.
+Even over long timescales, the two-step pathway contributes less (\cref{figure:pathways}, right panels).
+The lower rate of net diversification in the CD state, relative to ID, means that relatively fewer lineages are available to complete the second step of the two-step pathway.
 
@@
 
-\subsection{Pathways to polyploidy}
-
-% E: If we're going to include pathway results for real, I need to write more in this section and incorporate the figure.  TODO
-
-Polyploidization rates from self-compatible diploid $\rho_{CD}$ and from self-incompatible diploid $\rho_{ID}$ are not significantly different (see supplementary information figures).
-However, the distribution of $\rho_{ID}$ seems to always consider faster values compare to the distribution of $\rho_{CD}$.
-This result is highlighted in ancestral reconstructions of the $ID/CD/CP$ model where transitions from ID to CP are more probable than transitions from CD to CP. % Emma here we put the figure?
-%B: I am not sure I understand the "However, " sentence. TODO.
+% R will add a bit about param confidence
 
@@
     \caption{
-        Median rate estimates for all fitted models. %B: "...all fitted/considered/evaluated models."
+        Median rate estimates for all fitted models.
         Units are per million years.
@@
 
+% TODO: swap panels between the figures so that the first is -delta and the second is +delta
+
 \begin{figure}
@@
 \label{figure:netdivnodip}
+\end{figure}
+
+% E todo: could make this look a little less klunky
+\begin{figure}
+    \centering \includegraphics[width=0.9\textwidth]{pathways}
+    \caption{
+        Contributions of the two pathways to polyploidy.
+        When considering only rates of transitions among the states (left), the one-step pathway dominates on short timescales and the two-step on long timescales.
+        When also considering diversification within each state (right), the one-step pathway dominates over any timescale.
+        The lower panels plot the ratio of the two pathway contributions in the upper panels.
+    }
+    \label{figure:pathways}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{hisseDPposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P+A/B, polyploidy model} % XXX
-% fixme: offset plotting error in "relative extinction" panel
+\caption{Posterior distribution for each of the parameters in the D/P+A/B, polyploidy model.  The axis is offset in one location so that the two overlapping distributions can be seen.} % XXX
 \label{suppfigure:DPAB}

### 2019-02-28 summarystats/pathways.R

@@ -51,7 +51,7 @@ path.probs2 <- function(t, q13, q14, q34)
 # Wa = 1 -> 4
 # Wb = 1 -> 3 -> 4
 
-# MAP values
+# MAP values from EstimatesIDCDPnodelta.csv
 q14 <- 0.0681          # ID -> CP, rho_I
 q13 <- 0.2066          # ID -> CD, q_IC
 q34 <- 0.0370          # CD -> CP, rho_C
@@ -74,9 +74,12 @@ pp <- as.data.frame(t(sapply(times, path.probs2, q13, q14, q34)))
 pp <- as.data.frame(t(sapply(times, path.probs1, Q)))
 
 # plot
-plot(times, pp$Wa, type="l", lwd=3, xlab="elapsed time", ylab="probability", ylim=c(0,1))
+plot(times, pp$Wa, type="l", lwd=3, xlab="elapsed time (My)", ylab="probability", ylim=c(0,1))
 lines(times, pp$Wb, lwd=3, lty=2)
-legend("topleft", c("SI/D to SC/P", "SI/D to SC/D to SC/P"), lty=c(1, 2), lwd=3)
+legend("topleft", c("ID to CP", "ID to CD to CP"), lty=c(1, 2), lwd=3)
+
+plot(times, pp$Wa/pp$Wb, log="y", type="l", lwd=3, xlab="elapsed time (My)", ylab="one-step : two-step", ylim=c(0.3, 100))
+abline(h=1, lty=3)
 
 ### Try with diversification ###
 
@@ -97,19 +100,36 @@ plot(times, ppd$Wa, type="l", lwd=3, xlab="elapsed time", ylab="relative 'probab
 lines(times, ppd$Wb, lwd=3, lty=2)
 legend("topleft", c("SI/D to SC/P", "SI/D to SC/D to SC/P"), lty=c(1, 2), lwd=3)
 
+plot(times, ppd$Wa/ppd$Wb, log="y", type="l", lwd=3, xlab="elapsed time (My)", ylab="one-step : two-step", ylim=c(0.3, 100))
+abline(h=1, lty=3)
+
 #--------------------------------------------------
 # Final plot
 #--------------------------------------------------
 
-pdf(file="../manuscript/pathways.pdf", width=10, height=6)
-par(mfrow=c(1,2))
+pdf(file="../manuscript/pathways.pdf", width=8, height=7)
+par(mfrow=c(2,2), mar=c(4,4,0,2), oma=c(0,0,3,0))
 
[…]

### 2019-02-28 manuscript/abstract.tex
new abstract draft

@@
 
-The effect of polyploidy in diversification remains a contentious issue.
+% E: Here's the previous abstract.  The new one below is aligned more with the current Introduction and Results.
+%
+% The effect of polyploidy in diversification remains a contentious issue.
 On the one hand,  recent studies  that found that polyploids have slower speciation rates and higher extinction rates than diploids left scientist wondering if polyploidy is truly an evolutionary dead-end.
@@
 We ask if Solanaceae polyploids are evolutionary dead-ends, and whether breeding system or some other unobserved traits are responsible of the patterns of diversification observed in the phylogeny.
+
+% E:  Help making this less bland?  Except for the first sentence, which may be too grandiose?
+If particular traits consistently affect rates of speciation and extinction, broad macroevolutionary patterns can be understood as consequences of selection at high levels of the biological hierarchy.
+Identifying traits associated with diversification rate differences is tricky, though, because of the many traits available to consider and the statistical challenge of testing for associations from phylogenetic data.
+Two traits that have been repeatedly suggested as drivers of differential diversification are whether a lineage is diploid or polyploid, and whether it is self-incompatible or self-compatible.
+We investigate the role of each of these traits, and their interaction, on speciation and extinction rates in Solanaceae.
+We find that the effect of ploidy can largely be explained by its correlation with breeding system, and that additional unknown factors work with breeding system to determine diversification rates.
+These results are largely robust to assumptions about whether diploidization occurs.
+Finally, we show that allowing for state-dependent diversification affects conclusions about the relative contribution of different evolutionary pathways to self-compatible polyploids.

### 2019-02-28 manuscript/results.tex
- pathway figure legend mods

@@
     \caption{
-        Contributions of the two pathways to polyploidy.
-        When considering only rates of transitions among the states (left), the one-step pathway dominates on short timescales and the two-step on long timescales.
-        When also considering diversification within each state (right), the one-step pathway dominates over any timescale.
-        The lower panels plot the ratio of the two pathway contributions in the upper panels.
+        %B: FIXME: R&E check this re-wording to see if it is to your liking--I think I clarified it a bit, although with length increase.
+        %Contributions of the two pathways to polyploidy.
+        %When considering only rates of transitions among the states (left), the one-step pathway dominates on short timescales and the two-step on long timescales.
+        %When also considering diversification within each state (right), the one-step pathway dominates over any timescale.
+        %The lower panels plot the ratio of the two pathway contributions in the upper panels.
+        Contributions of the two pathways to polyploidy, the one-step ID$\rightarrow$CP transitions, and the two-step, ID$\rightarrow$CD$\rightarrow$CP transitions.
+        Panel (a) shows the contributions when considering only transition rates among the states (ignoring the diversification rate parameters). 
+        The one-step pathway dominates on short timescales and the two-step on long timescales.
+        Panels (b) shows the contributions when including diversification of lineages in each state. 
+        The one-step pathway, in which polyploidy breaks down SI, dominates over any timescale.
+        Panels (c) and (d) plot the ratio of the pathway contributions in the upper panels, (a) and (b), respectively.
     }
+
+    the one-step pathway, from ID to CP
+
     \label{figure:pathways}

### 2019-03-02 manuscript/discussion.tex
- added commented-out bones of WGD-diploidization paragraph
Re-arranging introduction

@@
 % diploidization: comment
+Diploidization needs to be considered in diversification models because higher rates of net diversification (speciation minus extinction) in diploids can be obtained from models that ignore the possibility of a polyploid lineage that has diploidized being the diversification enhancer.
+Hence, in a diploidization lacking model, diploids will show higher rates of diversification when in fact it is a polyploidy event and its subsequent diploidization that generated higher speciation and less extinction.
+By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with stochastic models.
+
 %     effect on inferred diversification
+% B: notes to self.
+% Many flowering plants are thought to have experienced a WGD. (cite)
+% \citet{ku_2000} and \citet{blanc2004} posited that the lineage leading to tomato, \textit{Solanum lycopersicum}, may have experienced one or more paleopolyploid events.
+% A subsequent analysis of synteny between grape and \textit{Solanum} genomes, as well as the distribution between inferred paralogs within \textit{Solanum} (tomato and potato) genomes both suggest that this lineage experienced a likely round of ancient genome duplication or triplication \citep{tomato_2012}.
+%The age of this bump of paralogs was approximated using Ks distances to be around 71 million years \citep{tomato_2012}.
+% If this is the case, then the genomes are re-diploidized, because low and common chromosome numbers in related families (coffee n=11, tomato n=12, etc.). 
+%Not unlikely, because Arabidopsis n=5.
+% On the other hand, \citet{wu_2008} find no evidence for recent polyploidization-diploidization, based on map-based synteny. 
+% Simple genome re-arrangements appear sufficient to explain chromosomal evolution of x=12 (tomato-tobacco) group.
+% (Add what this has to do with re-diploidization and how to interpret $\delta$.
 

### 2019-03-02 manuscript/introduction.tex
notes while talking to Rosana
Re-arranging introduction

--- 2019-02-23 manuscript/introduction.tex
@@
 Species accumulate around the tree of life at different rates.
-The search for traits that explain these differences has been accelerated by dramatic increases in phylogenetic data and, despite some setbacks \citep{maddison_2015, rabosky_2015}, advances in analytical methods \citep{maddison_2007, fitzjohn_2009, goldberg_2012, beaulieu_2016}.
-% todo more refs: papers with big trees; quasse, bisse-ness, fisse, abc-quasse, maybe hoehna-vs-rabosky
+The search for traits that explain these differences has been accelerated by dramatic increases in phylogenetic data and, despite some setbacks \citep{maddison_2015, rabosky_2015, moore_2016}, advances in analytical methods \citep{maddison_2007, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017} are finding interesting lineages where potential interesting biological processes are changing the  diversification patterns.
+% todo more refs: papers with big trees; quasse, bisse-ness, fisse, abc-quasse, maybe hoehna-vs-rabosky- R: DONE, won't mention all or it will get really annoying.
+
 In these studies, it is common to identify a single focal trait and investigate its association with rates of speciation and extinction.
+This is problematic because the context in which traits occur can lead to complex interactions.
+Scientists are well aware of the contribution of complex interactions to the diversification process, thus, investigating multivariate traits linked to diversification is becoming the next best step in the development of new models (see \citet{caetano_2018, herrera_2018}).
 % (but see...
 ?) % refs: there must be some papers that musse on multiple traits?
-especially for geographic range?
-This is problematic for two reasons.
-On the biology side, the context in which traits occur can lead to complex interactions, perhaps with other traits that make up an organism or its environment.
-On the methods side, excluding other traits that additionally influence diversification can lead to incorrect conclusions about the focal trait \citep{rabosky_2015, beaulieu_2016}.
-One approach to this problem is to allow for a `hidden' or unspecified trait in the analysis \citep{beaulieu_2016}. % and an earlier beaulieu? applications of hisse?
-This can be effective, but it leaves behind the question of whether the hidden trait corresponds to a real trait that should be sought, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
-Alternatively, we can simultaneously consider the effects of more than one known trait, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
-% B: We can then have a good prior, reliable data, and replicate the studies.
-E: Agree.
-But when I tried to add it, I realized hidden models can also be hypothesis-generating, with the purported hidden trait identified in the next clade.
-In this paper, we focus on two of the best studied traits in flowering plants, ploidy and breeding system, and investigate their individual and joint effects on diversification.
+especially for geographic range?: R-FIXED
 
-Polyploidization multiplies the genomic content of cells, and it consequently has the potential to affect many or all traits and a great variety of evolutionary processes.
-It is also a mutation that occurs commonly in plants, at it is widespread at both population and lineage scales.
-The prevalence of variation in chromosome number, and especially ploidy, has been broadly considered a salient feature of flowering plants for nearly a century \citep{stebbins1938}.
+In this paper, we focus on two of the best studied traits in flowering plants, polyploidy and breeding system and their contributions to diversification with the goal of investigating how the complex interactions between these two traits can lead to problematic diversification patterns when considering them as independent.
+% R-This transition is still a little rough
+
+Polyploidy events multiply the genomic content of cells, and they consequently have the potential to affect many or all traits and a great variety of evolutionary \citep{ramsey_2002} and ecological processes\citep{sessa_2019}.
+It is also a mutation that occurs commonly in plants, at it is widespread at both population and lineage scales\citep{husband_2013, zenil_2017}.
+The prevalence of variation in chromosome number, and especially ploidy change, has been broadly considered a salient feature of flowering plants for nearly a century \citep{stebbins1938}.
 % B: instead of "ploidy" here, we may wish to use some stand-in for karyotype multiples, and stick to that language throughout.
@@
 I was taking "ploidy" to mean "whatever multiple," but is that a terminological abuse?
-The recent dramatic increase in the scale of available genome sequences uncovered ancient rounds of whole-genome duplications, and subsequent diploidization, across angiosperms \citep{lynch2000, vision2000}.
-% E: I am deferring to R's use of "diploidization" instead of "re-diploidization"
-Nearly all lineages of flowering plants are thought to have undergone at least one or two rounds of polyploidization. % todo refs, jiao etc
-%
-The prevalence of polyploidy, its variation across clades, and its large effects on genotypes and phenotypes raises the hypothesis that ploidy plays an important role in shaping rates of speciation and extinction.
-This has led to a debate: \citet{mayrose_2011, mayrose_2015} found a lower rate of lineage diversification for polyploids, while \citet{soltis_2014} argued that polyploidy is not an evolutionary dead end because whole genome duplications occurred at the base of many successful angiosperm clades, and \citet{landis_2018} reported higher diversification rates following some whole genome duplications.
-% and tank2015?
-Our study revisits this question, additionally incorporating the process of diploidization.
-% E:  Rosana, could you add a little about the importance of including diploidization?  todo
-Furthermore, we consider the interaction between polyploidy and other traits.
+%B: sorry, that was supposed to complain about the use of "polyploidy" instead of "ploidy" to designate the trait.
+The trait is ploidy, and a state is polyploidy.
+Also, we should clarify that polyploidy is any of tetraploidy+.
+I am not sure what we did with 3x here.
+It's rather rare, but...?% R- I think the trait is polyploidy and the state is ploidy (as 2, or +2).
+I will go with everyone else in the literature and use polyploidy
+The prevalence of polyploidy, its variation across clades, and its large effects on genotypes and phenotypes raise the hypothesis that polyploidy plays an important role in shaping rates of speciation and extinction.
+However,  previous stochastic diversification models found a lower and negative rate of lineage diversification for polyploids \citet{mayrose_2011, mayrose_2015} when applied to 49 clades from the angiosperm phylogeny.
+This key finding lead to a the renaissance of the ``polyploidy is an evolutionary dead-end" idea and initiated a discussion about the location and the diversification effect of  whole genome duplications in flowering plants.
+Genomic evidence has shown that at the base of highly diverse  angiosperm clades (including  the root of all angiosperms \citep{jiao_2011}) it is possible to find at least a polyploidy event  \citet{soltis_2014}.
+Most recently, the genomic evidence  from the 1KP make it possible for scientist to locate 106  whole genome duplications in the angiosperm phylogeny, and using a state depended diversification model that accounted for  the frequency of whole genome duplications in a lineage  they found that approximately 60-\%  of these  whole genome duplications are enhancing diversification\citep{landis_2018}.
+% and tank2015?- tank is more about the lag-time between polyploidy and diversification, I prefer not to go into the leg of lags or a reviewer might ask us to try to incorporate that which is not possible since we don't have the location of WGD.
+This recent dramatic increases in the scale of available genome sequences uncovering ancient rounds of whole-genome duplications like\citet{landis_2018} work bring into question  the role of genome downsizing and diploidization in the diversification process\citep{soltis_2015diploidization, dodsworth_2015}.
+The presence of  a significant  and positive net diversification rate for diploids can be  the result of a polyploid lineage that has been diploidized.
+Under this scenario, polyploidy could not longer be considered a macroevolutionary dead-end but instead, polyploidy would be the initiator of the diversification process.
+In the present work, we explore this polyploidy-diploidization scenario by adding diploidization rates in our proposed stochastic models.
 
-From the very beginning of polyploidy research, the study of this cytogenetic property considered correlations with other traits \citep{stebbins1938}. 
-Some of these associations are driven by indirect or ecological effects, such as the tendency for polyploidy, and other traits, to be found at higher latitudes or marginal habitats.
+%Furthermore, we consider the interaction between polyploidy and other traits.
+
+%From the beginning, study of this variable cytogenetic property considered correlations with other traits \citep{stebbins1938}.
+%B: ploidy research, or nothing (as now); was: "From the very beginning of polyploidy research," polyploidy is a state of the trait.
+%Some of these associations are driven by indirect or ecological effects, such as the tendency for polyploidy, and other traits, to be found at higher latitudes or marginal habitats.
 % todo refs: husband etc
-Other correlations may have a more direct, causal connection.
-Among the many changes associated with polyploidization, perhaps the most prominent is the association between polyploidy and propensity for self-fertilization \citep{stebbins1950, barrett1988}.
+%Other correlations may have a more direct, causal connection.
+
+Another approach useful to understand the differences between the net diversification processes of diploids and polyploids is the work of  \citet{beaulieu_2016}.
+Beaulieu et al.
+proposed to  allow for a `hidden' or unspecified state linked to the trait of interest for models of trait linked to diversification.
+The presence of a hidden state can point out at sudden changes in the diversification rate under a value of a trait that are due to some unobserved source of heterogeneity.
+% and an earlier beaulieu?
+applications of hisse?
+%B: perhaps above can begin with: "One proposed approach to accommodate..." - R.
+This is the earliest HiSSE, that paper took forever for them to publish, people did not get it on the first submissions, before that it was only MuSSE in terms of multivariate.
+
+The hidden-state  models have effectively shown that traits that were believed to be drivers of diversification, in reality are not, and that it is an unobserved but associated trait responsible for differences that are initially found.
+When the hidden state and not the focal trait, is responsible for changes in the diversification process, it leaves behind the question of whether the hidden state corresponds to a real state of a trait that should be sought, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
+Alternatively, one can simultaneously consider the effects of more than one known trait, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
+% B: We can then have a good prior, reliable data, and replicate the studies.
+E: Agree.
+But when I tried to add it, I realized hidden trait models can also be hypothesis-generating, with the purported hidden trait identified in the next clade.
+%B: I don't get "the next clade"?
+E: I study trait X in clade A and based on the hidden state results, I hypothesize that trait Y may be the important hidden factor.
+So in clade B I study both traits X and Y.
+
+In the case of polyploidy,  one of the most prominent complex interactions known is the association between ploidy and propensity for self-fertilization \citep{stebbins1950}.
+%B: removed barrett1988, because he highlights correlation between polyploidy and self-compatibility, not self-fertilization.
 In some cases, the evidence for a correlated shift in mating system along with polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
@@
 Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic self-incompatibility systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}.
-This creates a correlation between polyploidy and self-compatibility by preventing the existence of self-incompatible polyploids.
-In clades with these systems, it is thus natural to consider the simultanous macroevolution of ploidy and breeding system.
+This creates a correlation between polyploidy and self-compatibility by precluding the existence of self-incompatible polyploids.
+In clades with these systems, it is thus natural to consider the simultaneous macroevolution of polyploidy and breeding system.
 
-% E note to self: edit this next...
 Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
@@
 
-Here, we use extensive data on ploidy and breeding system in Solanaceae to investigate the associations of these two traits with lineage diversification.
+Here, we use extensive data on ploidy and breeding system in 595 Solanaceae taxa to investigate the associations of these two traits with lineage diversification.
 Considering each trait separately indicates that each is connected to diversification differences.
@@
 Our results emphasize the importance of considering traits not only in isolation, especially when there are strong correlations between them.
+
+%We investigate their individual and joint effects on diversification, and whether or not adding hidden states at different steps of modeling can help us generate hypothesis about the benefit of searching for a trait that can pass as a hidden state in a simple model.
+%R- I added this to link the paragraph before
+
+%Additionally, a lack of methods that allow for simultaneous inference of effects of other traits, which also influence diversification, can lead to incorrect conclusions about the focal trait \citep{rabosky_2015, beaulieu_2016}.

### 2019-03-02 manuscript/introduction.tex
Major introduction changes, main figure changes and revised methods

@@
 Species accumulate around the tree of life at different rates.
-The search for traits that explain these differences has been accelerated by dramatic increases in phylogenetic data and, despite some setbacks \citep{maddison_2015, rabosky_2015, moore_2016}, advances in analytical methods \citep{maddison_2007, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017} are finding interesting lineages where potential interesting biological processes are changing the  diversification patterns.
+The search for traits that explain these differences has been accelerated by dramatic increases in phylogenetic data and, despite some setbacks \citep{maddison_2015, rabosky_2015, moore_2016}, advances in analytical methods \citep{maddison_2007, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017} are finding lineages where potential interesting biological processes are changing the  diversification patterns.
 % todo more refs: papers with big trees; quasse, bisse-ness, fisse, abc-quasse, maybe hoehna-vs-rabosky- R: DONE, won't mention all or it will get really annoying.
@@
 In these studies, it is common to identify a single focal trait and investigate its association with rates of speciation and extinction.
-This is problematic because the context in which traits occur can lead to complex interactions.
-Scientists are well aware of the contribution of complex interactions to the diversification process, thus, investigating multivariate traits linked to diversification is becoming the next best step in the development of new models (see \citet{caetano_2018, herrera_2018}).
+This is problematic because the context in which traits occur can lead to complex interactions amongst them.
+Scientists are well aware of the contribution of complex interactions to the diversification process, thus, proposing models where multivariate traits linked to diversification is necessary (see \citet{caetano_2018, herrera_2018}).
 % (but see...
@@
 
-In this paper, we focus on two of the best studied traits in flowering plants, polyploidy and breeding system and their contributions to diversification with the goal of investigating how the complex interactions between these two traits can lead to problematic diversification patterns when considering them as independent.
+In the present work,  we focus on two of the best studied traits in flowering plants, polyploidy and breeding system and their contributions to speciation and extinction with the goal of investigating how the interaction between these two traits can illuminate the process of diversification.
 % R-This transition is still a little rough
 
-Polyploidy events multiply the genomic content of cells, and they consequently have the potential to affect many or all traits and a great variety of evolutionary \citep{ramsey_2002} and ecological processes\citep{sessa_2019}.
-It is also a mutation that occurs commonly in plants, at it is widespread at both population and lineage scales\citep{husband_2013, zenil_2017}.
-The prevalence of variation in chromosome number, and especially ploidy change, has been broadly considered a salient feature of flowering plants for nearly a century \citep{stebbins1938}.
+Polyploidy events multiply the genomic content of cells, and have the potential to affect many other phenotypes as well  as a variety of evolutionary \citep{ramsey_2002} and ecological processes\citep{sessa_2019}.
+Polyploidy is also a mutation that occurs commonly in plants, and it is widespread at both population and lineage scales\citep{husband_2013, zenilferguson_2017}.
+The prevalence of chromosome number variation and ploidy change has been broadly considered as a salient feature of flowering plants for nearly a century \citep{stebbins1938}.
 % B: instead of "ploidy" here, we may wish to use some stand-in for karyotype multiples, and stick to that language throughout.
@@
 I will go with everyone else in the literature and use polyploidy
-The prevalence of polyploidy, its variation across clades, and its large effects on genotypes and phenotypes raise the hypothesis that polyploidy plays an important role in shaping rates of speciation and extinction.
-However,  previous stochastic diversification models found a lower and negative rate of lineage diversification for polyploids \citet{mayrose_2011, mayrose_2015} when applied to 49 clades from the angiosperm phylogeny.
-This key finding lead to a the renaissance of the ``polyploidy is an evolutionary dead-end" idea and initiated a discussion about the location and the diversification effect of  whole genome duplications in flowering plants.
-Genomic evidence has shown that at the base of highly diverse  angiosperm clades (including  the root of all angiosperms \citep{jiao_2011}) it is possible to find at least a polyploidy event  \citet{soltis_2014}.
-Most recently, the genomic evidence  from the 1KP make it possible for scientist to locate 106  whole genome duplications in the angiosperm phylogeny, and using a state depended diversification model that accounted for  the frequency of whole genome duplications in a lineage  they found that approximately 60-\%  of these  whole genome duplications are enhancing diversification\citep{landis_2018}.
+The combination of ploidy changes with changes on genotypes and phenotypes raise the hypothesis that polyploidy has played an important role in shaping rates of speciation and extinction.
+However, when modeling polyploidy linked to diversification, statistical inferences have concluded that polyploids have slower (or even negative) rates of net diversification (speciation minus extinction) than diploids \citet{mayrose_2011, mayrose_2015}.
+This key result was overwhelmingly consistent across 49 clades from the angiosperm phylogeny and lead to a the renaissance of the ``polyploidy as an evolutionary dead-end" idea and initiated a necessary discussion about the phylogenetic location and the role of whole genome duplications in flowering plants.
+Genomic evidence has shown that at the base of highly diverse  angiosperm clades (including  the root of all angiosperms \citep{jiao_2011}) it is possible to find at least one polyploidy event  \citet{soltis_2014}.
+Most recently, using the genomic evidence  from the 1KP project \citep{landis_2018} found 106  whole genome duplications in the angiosperm phylogeny that combined with a state dependent diversification model resulted in approximately 60\%  of  whole genome duplication events potentially enhancing net diversification\citep{landis_2018}.
 % and tank2015?- tank is more about the lag-time between polyploidy and diversification, I prefer not to go into the leg of lags or a reviewer might ask us to try to incorporate that which is not possible since we don't have the location of WGD.
-This recent dramatic increases in the scale of available genome sequences uncovering ancient rounds of whole-genome duplications like\citet{landis_2018} work bring into question  the role of genome downsizing and diploidization in the diversification process\citep{soltis_2015diploidization, dodsworth_2015}.
-The presence of  a significant  and positive net diversification rate for diploids can be  the result of a polyploid lineage that has been diploidized.
+Uncovering ancient rounds of whole-genome duplications like\citet{landis_2018}  brings into question  the role of genome downsizing and diploidization in the diversification process\citep{soltis_2015diploidization, dodsworth_2015}.
+The presence of  a significant  and positive net diversification rate for (some) diploids can be  the result of a polyploid lineage that has been diploidized.
 Under this scenario, polyploidy could not longer be considered a macroevolutionary dead-end but instead, polyploidy would be the initiator of the diversification process.
-In the present work, we explore this polyploidy-diploidization scenario by adding diploidization rates in our proposed stochastic models.
+In the present work, we explore this polyploidy-diploidization scenario by adding diploidization rates in our proposed diversification models.
 
@@
 
-Another approach useful to understand the differences between the net diversification processes of diploids and polyploids is the work of  \citet{beaulieu_2016}.
-Beaulieu et al.
-proposed to  allow for a `hidden' or unspecified state linked to the trait of interest for models of trait linked to diversification.
-The presence of a hidden state can point out at sudden changes in the diversification rate under a value of a trait that are due to some unobserved source of heterogeneity.
+Another approach necessary to do a comprehensive evaluation of the role of polyploidy in diversification processes is the work proposed by \citet{beaulieu_2016}.
+The authors proposed models to  allow for a `hidden' or unspecified state linked to the trait of interest for models of trait linked to diversification.
+The presence of a hidden state can point out at sudden changes in the diversification rate under a value of a trait that are due to some unobserved source of heterogeneity and not the focal trait per se.
 % and an earlier beaulieu?
@@
 
-The hidden-state  models have effectively shown that traits that were believed to be drivers of diversification, in reality are not, and that it is an unobserved but associated trait responsible for differences that are initially found.
+These hidden-state  models have effectively shown that traits that were believed to be drivers of diversification, in reality are not, and that it is an unobserved but associated trait responsible for differences that are initially found.
 When the hidden state and not the focal trait, is responsible for changes in the diversification process, it leaves behind the question of whether the hidden state corresponds to a real state of a trait that should be sought, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
+
 Alternatively, one can simultaneously consider the effects of more than one known trait, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
@@
 
-In the case of polyploidy,  one of the most prominent complex interactions known is the association between ploidy and propensity for self-fertilization \citep{stebbins1950}.
+In the particular case of polyploidy,  one of the most prominent complex interactions known is the association between ploidy and propensity for self-fertilization \citep{stebbins1950}.
 %B: removed barrett1988, because he highlights correlation between polyploidy and self-compatibility, not self-fertilization.
@@
 Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
-In particular, self-incompatibility (SI) systems cause a plant to reject its own pollen, and their loss, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution.
+In particular, self-incompatibility systems cause a plant to reject its own pollen, and their loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution.
 % refs: stebbins, igic
-Previous phylogenetic analyses have reported higher rates of diversification for SI than for SC lineages \citep{goldberg_2010, devos2014}, but they have not considered the possibility of other correlated traits driving this pattern.
+Previous phylogenetic analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, but they have not considered the possibility of other correlated traits driving this pattern.
 Given that changes in ploidy and breeding systems may be causally related and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
@@
 For example, do losses of SI more commonly occur tied to polyploidization, or without a ploidy shift?
-Do polyploids arise more commonly from SI or SC diploids?
-\Citet{robertson_2011} found that the pathway from SI diploids to SC polyploids is dominated by loss of SI followed later by polyploidization over long timescales, but proceeds in one step via polyploidization of SI species over short timescales.
+Do polyploids arise more commonly from self-incompatible or self-compatible diploids?
+\Citet{robertson_2011} found that the pathway from self-incompatible diploids to self-compatible polyploids is dominated by loss of self-incompatibility followed later by polyploidization over long timescales, but proceeds in one step via polyploidization of SI species over short timescales.
 We revisit this question with a greatly improved phylogeny and methods that allow for diversification rate differences.
@@
 
-Here, we use extensive data on ploidy and breeding system in 595 Solanaceae taxa to investigate the associations of these two traits with lineage diversification.
-Considering each trait separately indicates that each is connected to diversification differences.
-Considering them jointly, however, reveals that the ploidy connection is removed by incorporating breeding system.
-We further show that the general results are robust to allowing for diploidization and a hidden trait, and something about pathways. % FIXME
-Our results emphasize the importance of considering traits not only in isolation, especially when there are strong correlations between them.
+We present the most comprehensive data on ploidy and breeding system at species level in a phylogenetic context used in any trait linked diversification model.
+Our dataset consists of  a 595 dated Solanaceae phylogeny, with ploidy and/or breeding system information for the tips of the phylogeny to investigate the associations of these two traits with lineage diversification.
+
+In the present work, we propose a series of models that represent a comprehensive pathway of analyses that researchers need to follow when investigating complex interactions of traits in a diversification context.
+First, we present models where polyploidy and breeding system are modeled separatedly, and compare the resulting inferences to results of previous published models.
+Second, we add a hidden state to each of the models, in order to investigate if the focal trait maintains differences in diversification for focal states or if there is an unobserved but correlated hidden state that can drive the diversification pattern.
+Third, we ask if the hidden state is actually related to the complex interaction between breeding system and polyploidy and propose a model to consider both traits linked to diversification.
+In step four, we  ask if on top of the complex interaction there is a second unknown trait that can be driving differences in diversication, so we add to our bivariate diversification model a hidden trait.\newline
+
+Finallly, we ask if diploidization can be playing a role changing patterns of diversication.
+We perform model selection in a Bayesian framework to identify which models are fitting better the evidence of the Solanaceae data.
+%Considering them jointly, however, reveals that the ploidy connection is removed by incorporating breeding system.
+%We further show that the general results are robust to allowing for diploidization and a hidden trait, and something about pathways. % FIXME
+%Our results emphasize the importance of considering traits not only in isolation, especially when there are strong correlations between them.
 

### 2019-03-02 manuscript/methods.tex
Major introduction changes, main figure changes and revised methods

@@
 
-Chromosome number data were obtained for all Solanaceae taxa in the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were cleaned semi-automatically using the CCDBcurator R package \citep{rivero_2019}.
-%B: perhaps "cleaned" should be explained, unless it's hard to identically repeat as-is.
-This large dataset includes the compilation of Solanaceae ploidy states from \citet{robertson_2011}.
-Species were coded as either diploid (D) or polyploid (P).
-For the majority of species, ploidy was assigned according to information from the original publications and the Kew Royal Botanic Gardens C-value DNA resource \citep{bennett_2005}.
-For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus.
-For example, \textit{Solanum betaceum} did not include information about ploidy level but it has 24 chromosomes, so because $x=12$ is the base chromosome number of the \textit{Solanum} genus \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid.
-Species with more than one ploidy level were assigned the smallest and most frequent ploidy level recorded.
+Chromosome number data were obtained for all Solanaceae taxa in the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were curated semi-automatically using the CCDBcurator R package \citep{rivero_2019}.
+CCDB contains records from original sources that have multiple complex symbol patterns denoting multivalence, or irregularites of chromosome counts.
+After a first full round of curation via CCDBcurator, we contrasted results by hand and corrected chromosomes where CCDBcurator's output was not able to curate.
+Our hand curated records were also contrasted against the large dataset Solanaceae ploidy states from \citet{robertson_2011} and against ploidy data from the C-value DNA dataset from Kew royal botanic gardens \citep{bennett_2005}.
+  %B: perhaps "cleaned" should be explained, unless it's hard to identically repeat as-is. R- I added more (it only took 4months to hand-curate 14,000 entries
+By contrasting three sources of information, we were able to code taxa as diploid (D) or polyploid (P).
+For the majority of species, ploidy was assigned according to information from the original publications and the Kew royal botanic gardens C-value DNA resource \citep{bennett_2005}.
+For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus or based on information of incompatibility.
+If a species was incompatible without any chromosomal information it was assigned as diploid.
+For example, \textit{Solanum betaceum} did not include information about ploidy level but it has 24 chromosomes, since $x=12$ is the base chromosome number of the \textit{Solanum} genus \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid.
+Species with more than one ploidy level were assigned the smallest of most frequent ploidy level recorded.
 % E: There was more here about cross-verifying that no P species were SI.  I removed it because there were conflicts before, but we resolved them.  This is addressed in the next paragraph.
@@
 Breeding system was scored as self-incompatible (I) or self-compatible (C) based on results curated from the literature and  original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
-Most species could unambiguously be coded as either I or C \citep{raduski_2012}. %B: most? Any exceptions?
+Most species could unambiguously be coded as either I or C \citep{raduski_2012}.  But for diploids without breeding system information taxa was coded as both I or C.%B: most? Any exceptions?
 Following previous work, we coded as I any species with functional I systems, even if C or dioecy was also reported.
@@
 Following the reasoning outlined in \citet{robertson_2011}, we examined closely the few species for which the merged ploidy and breeding system data indicated the presence of self-incompatible polyploids.
-Although SI populations frequently contain some SC individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find a convincing case of a naturally occurring SI and polyploid population.
+Although I populations frequently contain some C individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find a convincing case of a naturally occurring I and polyploid population.
 %B: Thus qualified, as above, I think this should be inserted.
-The single instance of an SI and polyploid individual appears to be an allopentaploid hybrid of {\em Solanum oplocense} Hawkes x {\em Solanum gourlayii} Hawkes, reported by \citealt{camadro_1981}.
-Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain SI, so long as they express a single allele at the S-locus (discussed in \citealt{robertson_2011}).
-Because of the resulting absence of SI and polyploid populations, as well as the linked functional explanation for disabling of gametophytic self-incompatibility systems with non-self recognition, following whole genome duplication \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).
+The single instance of an I and polyploid individual appears to be an allopentaploid hybrid of {\em Solanum oplocense} Hawkes x {\em Solanum gourlayii} Hawkes, reported by \citealt{camadro_1981}.
+Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain I, so long as they express a single allele at the S-locus (discussed in \citealt{robertson_2011}).
+Because of the resulting absence of I and polyploid populations, as well as the linked functional explanation for disabling of gametophytic self-incompatibility systems with non-self recognition, following whole genome duplication \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).
 %B: I added the non-self recognition bit, because in Prunus, heteroallelic pollen may not cause SC.
@@
 In a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$), net diversification rates ($r_D=\lambda_D-\mu_D$, $r_P=\lambda_P-\mu_P$), and relative extinction rates ($\nu_D=\mu_D / \lambda_D$, $\nu_D=\mu_D / \lambda_D$) associated with each state.
-This analysis explores the same question as \citet{mayrose_2011, mayrose_2015}, but our analyses differ because we include not only polyploidization (parameter $\rho$, the transition rate from $D$ to $P$), but also diploidization (parameter $\delta$, the transition rate from $P$ to $D$).
+This analysis explores the same question as \citet{mayrose_2011, mayrose_2015}, using the polyploidization (parameter $\rho$, the transition rate from $D$ to $P$).
 % E: presumably the motivation will be explained in the Intro
@@
 BiSSE-like models can suffer from a large false discovery rate because they fail to account for diversification rate changes that do not directly depend on the trait of interest \citep{rabosky_2015, beaulieu_2016}.
-Diversification rate differences explained by something (trait) other than ploidy, are accommodated by adding a hidden state (HiSSE model; \citealt{beaulieu_2016}).
+Diversification rate differences explained by some trait other than ploidy, are accommodated by adding a hidden state (HiSSE model; \citealt{beaulieu_2016}).
 %B: inserted "(trait)", because that seems to be the goal - a biological explanation?
@@
 We estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$), extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$), net diversification rates ($r_{D_A},\ r_{D_B},\ r_{P_A},\ r_{P_B}$), and relative extinction rates ($\nu_{D_A},\ \nu_{D_B},\ \nu_{P_A},\ \nu_{P_B}$).
-In this model polyploidization rate $\rho$ and diploidization rate $\delta$ are also included, and changes between hidden states are symmetrical with rate $\alpha$.
-%B: same dip vs.
-re-dip question, as above.
+In this model polyploidization rate  is again represented by $\rho$  and hidden states are symmetrical with rate $\alpha$. %B: same dip vs. re-dip question, as above.
 
@@
 %B: may need to introduce "S-locus" earlier.
-This represents very strong evidence that the SI mechanism, and our $I$ state is ancestral to the Solanaceae, and did not arise independently within the family ($q_{CI}=0$).
+This represents very strong evidence that the self-incompatible mechanism, and our $I$ state is ancestral to the Solanaceae, and did not arise independently within the family ($q_{CI}=0$).
 %B: I think I have this right <- someone please double-check.
@@
 %B: What happens if SC->SI is not fixed? Asking for a friend.  E: Surprisingly, Rosana reports, results are not turned on their head. 
-%B: Whew, I was worried. Er, "friend" says thanks.
+%B: Whew, I was worried.
+Er, "friend" says thanks.
+R- Your welcome, by mistake I did that model with reversibility from C to I.
+I didn't run it for as long as the rest but preliminary ancestral reconstructions seem almost identical when you add that parameter than when you don't.
 
@@
 
-If ploidy and breeding system each influence lineage diversification individually, it is logical to examine their possible joint effects.
+If polyploidy and breeding system each influence lineage diversification individually, it is logical to examine their possible joint effects.
 %B: unsure about edit, which was: "If both ploidy and breeding system have the potential to influence lineage diversification, we should logically consider their effects jointly."  E: okay; also try to add that their transitions interact
@@
 We call this the \textit{ID/CD/CP ploidy and breeding system} model.
-The model has 10 parameters, six for diversification in each state ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and four for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $\delta$ for diploidization from $CP$ to $CD$; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
+The model has 9 parameters, six for diversification in each state ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and three for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
 The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
@@
 A fully parameterized version of this model would have 26 rate parameters \citep{herrera_2018}. 
-Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fitted a simplified version with 16 parameters.
+Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fitted a simplified version with 15 parameters.
 The reduction in parameter space is achieved by fixing the rates for transitions among hidden states to be equal with rate $\alpha$, and fixing the transition rates between observed states to be independent of the hidden state (rates $\rho_I,\ \rho_C,\ \delta,\ q_{IC}$ as defined for the ID/CD/CP model).
 There are additionally twelve diversification rate parameters ($\lambda_{ID_A},\ \lambda_{ID_B},\ \lambda_{CD_A},\ \lambda_{CD_B},\ \lambda_{CP_A},\ \lambda_{CP_B},\ \mu_{ID_A},\ \mu_{ID_B},\ \mu_{CD_A},\ \mu_{CD_B},\ \mu_{CP_A},\ \mu_{CP_B}$).
-%B: I can't tell, but it seems to me that this 26->16 parameter reduction makes it possible to underestimate the effect of these hidden states.
-Perhaps this complication is covered in the Discussion section.
+%B: I can't tell, but it seems to me that this 26->16 parameter reduction makes it possible to underestimate the effect of these hidden states.Perhaps this complication is covered in the Discussion section.
+%R- It is possible, but adding all the parameters can also result in an underestimation of the hidden states contribution because it is not enough sample to accurately estimate transitions.
+In my simulations for diversification-free models, accurately estimating 10 parameters can only be achieved with trees with more than 500 tips.
+Because of that experience I think 16 parameters is already pushing really the inference machine hard.
+This happens because the phylogeny is not an independent sample so your ESS is much much smaller and you might end up over-parameterizing.
 
@@
 
-For all four models that consider ploidy changes, we allowed diploidization.
+For all four models that consider ploidy changes, we allowed diploidization by adding a parameter $\delta$ connecting polyploidy state back to diploid or to self-compatible diploid in the mult-state models (denoted by $+/delta$ in the models of  \cref{table:marginallike}).
 Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010}.
@@
 Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
-For each model, an MCMC chain was run for 96 hours in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the 6 models.
+For each model, an MCMC chain was run for 96 hours in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the 10 models.
 %B: the number of models changes here?
+R-my mistake!
+fixed it
 For each model, convergence and mixing of the MCMC was tested using the R library \texttt{coda} and the software package Tracer (see supplementary information for convergence plots).
@@
 % R will add a bit about the ASR methods
+In the RevBayes code we calculated the probability of ancestral states at each node as performed chromosome number change models proposed in  \citet{freyman_2017}.
+Ancestral reconstructions for all models can be found in the supplementary information and show the maximum a posteriori of the marginal probability distributions for each of the 594 internal nodes for each of the 10 models.
 
@@
 % E: I think this short explanation is clear enough, especially with the figure.  I'm not sure whether the hierarchical concept is exactly applicable here.
+%R: Boris in the introduction mentions this possible causal relationship would that be equivalent to say hierarchical?
 Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $BF(M_0,\ M_1) = \ln[ P(\mathbf{X} | M_0) - P(\mathbf{X} | M_1]$.
@@
 
+% R: I changed this models to go now in the new order
+
 \begin{table}
@@
                              & &  &System & State & Parameters & Log-Likelihood \\ \midrule
-1. D/P                       &	Yes  &	Yes &	No	&No	& 6	& -1182.93 \\
-2. D/P no $\delta$           &	Yes & 	No	& No	& No & 	5 &	-1193.66\\
-3. D/P+A/B                   &	Yes &	Yes	&No &	Yes &	11 &	\textbf{-1145.69}\\
-4. D/P+A/B no $\delta$       &	Yes & No &	No &	Yes &	10	&-1150.99\\
+1. D/P           &	Yes & 	No	& No	& No & 	5 &	-1193.66\\
+2. D/P +$\delta$                      &	Yes  &	Yes &	No	&No	& 6	& -1182.93 \\
+3. D/P+A/B        &	Yes & No &	No &	Yes &	10	&-1150.99\\
+4. D/P+$\delta$+A/B                   &	Yes &	Yes	&No &	Yes &	11 &	\textbf{-1145.69}\\
 5. I/C                       &	No & No	&Yes &	No &	 5 &  -1194.80 \\
 6. I/C+A/B                   &	No &	 No	&Yes &	Yes	& 10 & \textbf{-1155.37}\\
-7. ID/CD/CP                  &	Yes & 	Yes &	Yes &	No &	10 & -1344.50\\
-8. ID/CD/CP no $\delta$      &	Yes & 	No &	Yes	&No &	9 &-1345.87\\
-9. ID/CD/CP+A/B              &	Yes 	&Yes &	Yes &	Yes &	16 & \textbf{-1300.35} \\
-10. ID/CD/CP+A/B no $\delta$ & Yes & 	No	&Yes &	Yes	&15 &-1303.55 \\ \bottomrule
+7. ID/CD/CP      &	Yes & 	No &	Yes	&No &	9 &-1345.87\\
+8. ID/CD/CP +$\delta$                 &	Yes & 	Yes &	Yes &	No &	10 & -1344.50\\
+9. ID/CD/CP+A/B  & Yes & 	No	&Yes &	Yes	&15 &-1303.55 \\ 
+10. ID/CD/CP+$\delta$+A/B              &	Yes 	&Yes &	Yes &	Yes &	16 & \textbf{-1300.35} \\
+\bottomrule
 \end{tabular}

### 2019-03-02 manuscript/results.tex
Major introduction changes, main figure changes and revised methods

@@
 % NOTE: will need to update figure/panel refs if making no-delta models primary
+% R: I am changing no delta as just D/P and prevous D/P as +delta
 
@@
 
-Evolution from SI diploids to SC polyploids can proceed through two different pathways.
+Evolution from self-incompatible diploids to self-compatible polyploids can proceed through two different pathways.
 Determining the relative contribution of these pathways based on our estimated transition rates from the ID/CD/CP model, we find that the one-step pathway is more likely on short timescales and the two-step pathway is more likely on long timescales (\cref{figure:pathways}, left panels).
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{Netdiversificationallmodels.pdf} % FIXME: The panel labels should use "+A/B" instead of "-A/B".
+\includegraphics[width=\textwidth]{allnetdivmodelswithoutdelta.pdf} % FIXME: The panel labels should use "+A/B" instead of "-A/B".
 Rosana, if you add the .svg file for this figure, I can make small changes like this.
-\caption{Net diversification rates for all models that include diploidization.}  
+\caption{Net diversification rates for all models.}  
 \label{figure:netdivall}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{netdiversificationallnodip.pdf} % FIXME: same note as above about "+A/B"
-\caption{Net diversification rates for all models that do not include diploidization.}  
+\includegraphics[width=\textwidth]{allnetdivmodelsdiploidization.pdf} % FIXME: same note as above about "+A/B"
+\caption{Net diversification rates for all models that include diploidization.}  
 \label{figure:netdivnodip}
@@
 \begin{suppfigure}
+\includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the D/P, polyploidy model} % XXX
+\label{suppfigure:DPnodip}
+\end{suppfigure}
+
+\begin{suppfigure}
 \includegraphics[width=\textwidth]{bisseDPposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P, polyploidy model} % XXX
+\caption{Posterior distribution for each of the parameters in the D/P+ $\delta$,polyploidy model} % XXX
 \label{suppfigure:DP}
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P no $\delta$, polyploidy model} % XXX
-\label{suppfigure:DPnodip}
+\includegraphics[width=\textwidth]{hisseDPnodipposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the D/P+A/B, polyploidy model} % XXX
+\label{suppfigure:DPnodipAB}
 \end{suppfigure}
@@
 \includegraphics[width=\textwidth]{hisseDPposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P+A/B, polyploidy model.  The axis is offset in one location so that the two overlapping distributions can be seen.} % XXX
+\caption{Posterior distribution for each of the parameters in the D/P+$\delta$+A/B, polyploidy model.  The axis is offset in one location so that the two overlapping distributions can be seen.} % XXX
 \label{suppfigure:DPAB}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{hisseDPnodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P no $\delta$+A/B, polyploidy model} % XXX
-\label{suppfigure:DPnodipAB}
 \end{suppfigure}
@@
 \begin{suppfigure}
+\includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the ID/CD/CP, polyploidy and breeding system model} % XXX
+\label{suppfigure:IDCDCPnodip}
+\end{suppfigure}
+
+\begin{suppfigure}
 \includegraphics[width=\textwidth]{musseDPSIposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the ID/CD/CP, polyploidy and breeding system model} % XXX
+\caption{Posterior distribution for each of the parameters in the ID/CD/CP+$\delta$, polyploidy and breeding system model} % XXX
 \label{suppfigure:IDCDCP}
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the ID/CD/CP no $\delta$, polyploidy and breeding system model} % XXX
-\label{suppfigure:IDCDCPnodip}
+\includegraphics[width=\textwidth]{muhisseDPSInodipposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the ID/CD/CP+A/B polyploidy and breeding system model} % XXX
+\label{suppfigure:IDCDCPnodipAB}
 \end{suppfigure}
@@
 \includegraphics[width=\textwidth]{muhisseDPSIposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the ID/CD/CP+A/B, polyploidy and breeding system model} % XXX
+\caption{Posterior distribution for each of the parameters in the ID/CD/CP+$\delta$+A/B, polyploidy and breeding system model} % XXX
 \label{suppfigure:IDCDCPAB}
 \end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{muhisseDPSInodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the ID/CD/CP no $\delta$+A/B polyploidy and breeding system model} % XXX
-\label{suppfigure:IDCDCPnodipAB}
-\end{suppfigure}

### 2019-03-02 manuscript/discussion.tex
- addition of a promised diploidization justification and fodder for discussion - requisite references - fixed spilled text on pathways table

--- 2019-02-28 manuscript/discussion.tex
@@
 % diploidization: comment
+Diploidization needs to be considered in diversification models because higher rates of net diversification (speciation minus extinction) in diploids can be obtained from models that ignore the possibility of a polyploid lineage that has diploidized being the diversification enhancer.
+Hence, in a diploidization lacking model, diploids will show higher rates of diversification when in fact it is a polyploidy event and its subsequent diploidization that generated higher speciation and less extinction.
+By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with stochastic models.
+
+% B: This is my pitch for motivation of the \delta paragraph and/or wording for how to reconcile estimates and genomic data (providing summary of data--to date)
+Diploidization is widely considered to be relatively rare, compared with polyploidization, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history.
+As a consequence, it is possible that nearly all extant species classified as ``diploid" in our analyses are possibly secondary diploids, having undergone both polyploidization and re-diploidization.
+Genomic evidence indicates that such an event may have occurred prior to the origin of the family Solanaceae. 
+\citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \textit{Solanum lycopersicum}, may have experienced one or more paleopolyploid events.
+A subsequent analysis of synteny between grape and \textit{Solanum} genomes, as well as the distribution between inferred paralogs within \textit{Solanum} (tomato and potato) genomes both suggest that this lineage experienced a likely round of ancient genome duplication or triplication \citep{tomato2012}.
+The age of the peak of paralog Ks distances, is approximately 71 million years \citep{tomato2012}. 
+If this is the case, then all of the genomes may have been subsequently re-diploidized, because common base chromosome numbers this and related families are n=11-12. 
+%Not unlikely, because Arabidopsis n=5.
+On the other hand, studies comparing map-based genome synteny within the family find no evidence for recent polyploidization-diploidization \citep{wu_2010a}. 
+Simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species, including all of those in the relatively cytogenetically conserved `x=12' group, which includes tomato, potato, eggplant, pepper, and tobacco.
+%B: Once this is re-arranged, and made to fit with the rest of the discussion, add what this has to do with re-diploidization and how to interpret $\delta$.
+
 %     effect on inferred diversification

### 2019-03-03 manuscript/introduction.tex
- Reworked Introduction
- If you read these, please note: separate each line with a line break!

@@
 
+%NOTE: please separate each line with a line break!
+
+%B: the key alteration below is an attempt at easing the transitions of the complicated existing logical structure, which intermixes advances with setbacks and problems.
+%B: the "Here, we" part should maybe come at the bottom of introduction, and it's a little early for here, in the first paragraph.
 Species accumulate around the tree of life at different rates.
-The search for traits that explain these differences has been accelerated by dramatic increases in phylogenetic data and, despite some setbacks \citep{maddison_2015, rabosky_2015, moore_2016}, advances in analytical methods \citep{maddison_2007, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017} are finding lineages where potential interesting biological processes are changing the  diversification patterns.
-% todo more refs: papers with big trees; quasse, bisse-ness, fisse, abc-quasse, maybe hoehna-vs-rabosky- R: DONE, won't mention all or it will get really annoying.
+A possible explanation for this phenomenon is that species possess traits that cause unequal rates of diversification. %B: new.
+The search for influential traits has been accelerated by a dramatic increases in data and an advance in methods, which now make use of far more information from phylogenetic trees \citep{maddison_2007}.
+Despite some setbacks \citep{maddison_2015, rabosky_2015, moore_2016}, advances in analytical methods \citep{fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017} are enabling the discovery of traits that alter the diversification patterns.
+%B: the following is now orphaned, because it's inclusion merits addition of non-trait-focused methods; "potentially interesting biological processes"
+Nevertheless, the aim of identifying the focal trait and investigating its association with rates of speciation and extinction remains complicated. 
+This is, in part, because the context in which traits occur can lead to complex interactions with the diversification process (see \citet{caetano_2018, herrera_2018}). 
+Trait combinations provide examples of such context, and may be of particular interest, given the rapid accumulation of information about traits. %cite?
+Two exceedingly well-studied traits, ploidy and breeding system, are of particular interest. 
+Their individual state-dependent contributions to the process of diversification, as well as influence on each others' trait transitions, have long been the focus of speculation and study \citep{stebbins1950}.
 
-In these studies, it is common to identify a single focal trait and investigate its association with rates of speciation and extinction.
-This is problematic because the context in which traits occur can lead to complex interactions amongst them.
-Scientists are well aware of the contribution of complex interactions to the diversification process, thus, proposing models where multivariate traits linked to diversification is necessary (see \citet{caetano_2018, herrera_2018}).
-% (but see...
-?) % refs: there must be some papers that musse on multiple traits?
-especially for geographic range?: R-FIXED
+%B: IMO: ploidy = trait, di-, tri-, tetra-, hexa-, etc. (including "poly-") = trait states; the uncertainty we were discussing is whether "tri-" < "poly-" or "tri-" = "poly-"
+%(sorry about the confusion: I accidentally mis-wrote in the commented-out banter with %E.)
+% was [B&E banter] and % R- I think the trait is polyploidy and the state is ploidy (as 2, or +2). I will go with everyone else in the literature and use polyploidy
+Whole genome duplication is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
+Such changes is ploidy have the potential to affect many other phenotypes, and therefore alter a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
+The prevalence of ploidy variation has been broadly considered as a salient feature of flowering plants for nearly a century \citep{stebbins1938}. 
+Polyploids have long been thought to have lower (and even negative) rates of net diversification than diploids \citet{mayrose_2011, mayrose_2015}. 
+A number of recent studies, however, find common paleo-polyploidization, including at the base of many highly diverse angiosperm clades \citep{soltis_2014}. 
+%B: explanation for this edit, rm: "Using the genomic evidence  from the 1KP project" : this is the data, not the approach; Ks and tree reconciliation are particularly troubling and weak modes of inference.
+I am okay with having this point to what others have found, but I would prefer to avoid endorsing what I see as a deeply problematic set of methods.
+\citet{landis_2018}, for example, found 106 whole genome duplications within angiosperms that, combined with a state-dependent diversification model, resulted in approximately 60\% of whole genome duplication events potentially enhancing net diversification.
+The mere occurrence of such ancestral whole genome duplications points to an possibly important role of genome downsizing and diploidization in the diversification process \citep{soltis_2015diploidization, dodsworth_2015}.
+The presence of a significant and positive net diversification rate for (some) diploids can be  the result of a polyploid lineage that has been diploidized. 
+Under this scenario, rather than being a macroevolutionary dead-end, polyploidy may instead be positively associated with the diversification process.
 
-In the present work,  we focus on two of the best studied traits in flowering plants, polyploidy and breeding system and their contributions to speciation and extinction with the goal of investigating how the interaction between these two traits can illuminate the process of diversification.
-% R-This transition is still a little rough
-
-Polyploidy events multiply the genomic content of cells, and have the potential to affect many other phenotypes as well  as a variety of evolutionary \citep{ramsey_2002} and ecological processes\citep{sessa_2019}.
-Polyploidy is also a mutation that occurs commonly in plants, and it is widespread at both population and lineage scales\citep{husband_2013, zenilferguson_2017}.
-The prevalence of chromosome number variation and ploidy change has been broadly considered as a salient feature of flowering plants for nearly a century \citep{stebbins1938}.
-% B: instead of "ploidy" here, we may wish to use some stand-in for karyotype multiples, and stick to that language throughout.
-E: What word would that be?
-I was taking "ploidy" to mean "whatever multiple," but is that a terminological abuse?
-%B: sorry, that was supposed to complain about the use of "polyploidy" instead of "ploidy" to designate the trait.
-The trait is ploidy, and a state is polyploidy.
-Also, we should clarify that polyploidy is any of tetraploidy+.
-I am not sure what we did with 3x here.
-It's rather rare, but...?% R- I think the trait is polyploidy and the state is ploidy (as 2, or +2).
-I will go with everyone else in the literature and use polyploidy
-The combination of ploidy changes with changes on genotypes and phenotypes raise the hypothesis that polyploidy has played an important role in shaping rates of speciation and extinction.
-However, when modeling polyploidy linked to diversification, statistical inferences have concluded that polyploids have slower (or even negative) rates of net diversification (speciation minus extinction) than diploids \citet{mayrose_2011, mayrose_2015}.
-This key result was overwhelmingly consistent across 49 clades from the angiosperm phylogeny and lead to a the renaissance of the ``polyploidy as an evolutionary dead-end" idea and initiated a necessary discussion about the phylogenetic location and the role of whole genome duplications in flowering plants.
-Genomic evidence has shown that at the base of highly diverse  angiosperm clades (including  the root of all angiosperms \citep{jiao_2011}) it is possible to find at least one polyploidy event  \citet{soltis_2014}.
-Most recently, using the genomic evidence  from the 1KP project \citep{landis_2018} found 106  whole genome duplications in the angiosperm phylogeny that combined with a state dependent diversification model resulted in approximately 60\%  of  whole genome duplication events potentially enhancing net diversification\citep{landis_2018}.
-% and tank2015?- tank is more about the lag-time between polyploidy and diversification, I prefer not to go into the leg of lags or a reviewer might ask us to try to incorporate that which is not possible since we don't have the location of WGD.
-Uncovering ancient rounds of whole-genome duplications like\citet{landis_2018}  brings into question  the role of genome downsizing and diploidization in the diversification process\citep{soltis_2015diploidization, dodsworth_2015}.
-The presence of  a significant  and positive net diversification rate for (some) diploids can be  the result of a polyploid lineage that has been diploidized.
-Under this scenario, polyploidy could not longer be considered a macroevolutionary dead-end but instead, polyploidy would be the initiator of the diversification process.
-In the present work, we explore this polyploidy-diploidization scenario by adding diploidization rates in our proposed diversification models.
-
-%Furthermore, we consider the interaction between polyploidy and other traits.
-
-%From the beginning, study of this variable cytogenetic property considered correlations with other traits \citep{stebbins1938}.
-%B: ploidy research, or nothing (as now); was: "From the very beginning of polyploidy research," polyploidy is a state of the trait.
-%Some of these associations are driven by indirect or ecological effects, such as the tendency for polyploidy, and other traits, to be found at higher latitudes or marginal habitats.
-% todo refs: husband etc
-%Other correlations may have a more direct, causal connection.
-
-Another approach necessary to do a comprehensive evaluation of the role of polyploidy in diversification processes is the work proposed by \citet{beaulieu_2016}.
-The authors proposed models to  allow for a `hidden' or unspecified state linked to the trait of interest for models of trait linked to diversification.
-The presence of a hidden state can point out at sudden changes in the diversification rate under a value of a trait that are due to some unobserved source of heterogeneity and not the focal trait per se.
-% and an earlier beaulieu?
-applications of hisse?
-%B: perhaps above can begin with: "One proposed approach to accommodate..." - R.
-This is the earliest HiSSE, that paper took forever for them to publish, people did not get it on the first submissions, before that it was only MuSSE in terms of multivariate.
-
-These hidden-state  models have effectively shown that traits that were believed to be drivers of diversification, in reality are not, and that it is an unobserved but associated trait responsible for differences that are initially found.
-When the hidden state and not the focal trait, is responsible for changes in the diversification process, it leaves behind the question of whether the hidden state corresponds to a real state of a trait that should be sought, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
-
-Alternatively, one can simultaneously consider the effects of more than one known trait, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
-% B: We can then have a good prior, reliable data, and replicate the studies.
-E: Agree.
-But when I tried to add it, I realized hidden trait models can also be hypothesis-generating, with the purported hidden trait identified in the next clade.
-%B: I don't get "the next clade"?
-E: I study trait X in clade A and based on the hidden state results, I hypothesize that trait Y may be the important hidden factor.
-So in clade B I study both traits X and Y.
-
-In the particular case of polyploidy,  one of the most prominent complex interactions known is the association between ploidy and propensity for self-fertilization \citep{stebbins1950}.
-%B: removed barrett1988, because he highlights correlation between polyploidy and self-compatibility, not self-fertilization.
-In some cases, the evidence for a correlated shift in mating system along with polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
-% barrett2008: "Because polyploidy affects the entire genome, it is perhaps not surprising that it influences many aspects of the phenotype, including the mating system.
-However, although it has long been recognized that the evolutionary transition from diploidy to polyploidy may result in correlated changes in mating patterns, the theoretical and empirical evidence is limited and often contradictory." (p.5)
-In other cases, however, polyploidy is not only a suspected correlate of breeding systems but indeed a causal link \citep{stout1942, lewis1947}.
-Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic self-incompatibility systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}.
-This creates a correlation between polyploidy and self-compatibility by precluding the existence of self-incompatible polyploids.
+Ploidy has long been thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
+The evidence for correlated shifts in mating system, from outcrossing to selfing, after polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
+In other cases, however, polyploidy is not only suspected to correlate with breeding system but causes the transition from self-incompatibility (SI) to self-compatibility(SC) \citep{stout1942, lewis1947}.
+Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic SI systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
+This creates a correlation between polyploidy and SC by precluding the existence of SI polyploids.
 In clades with these systems, it is thus natural to consider the simultaneous macroevolution of polyploidy and breeding system.
@@
 And dioecy.
+%B: Yes, let's.
 
-% note: saving trans-specific polymorphism for Methods
-% B: trans-polymorphism: We also have this great data for SI loci and the transition could be even more common, providing power to resolve the great diversification questions
+%B: STOPPED HERE.
+this is orphaned; we should not go to HiSSE before breeding systems.
+This yields an odd methods, traits, methods, traits, structure, where the methods are not related to the trait discussed.
+ I will work on it in the next 24 hours. It logically flows from ploidy/BS and other trait existence that hidden state models ought to be useful 
+Another approach necessary to do a comprehensive evaluation of the role of polyploidy in diversification processes is the work proposed by \citet{beaulieu_2016}. 
+The authors proposed models to  allow for a `hidden' or unspecified state linked to the trait of interest for models of trait linked to diversification. 
+The presence of a hidden state can point out at sudden changes in the diversification rate under a value of a trait that are due to some unobserved source of heterogeneity and not the focal trait per se.
+These hidden-state  models have effectively shown that traits that were believed to be drivers of diversification, in reality are not, and that it is an unobserved but associated trait responsible for differences that are initially found.
+When the hidden state and not the focal trait, is responsible for changes in the diversification process, it leaves behind the question of whether the hidden state corresponds to a real state of a trait that should be sought, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
+Alternatively, one can simultaneously consider the effects of more than one known trait, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
 
@@
 Our dataset consists of  a 595 dated Solanaceae phylogeny, with ploidy and/or breeding system information for the tips of the phylogeny to investigate the associations of these two traits with lineage diversification.
-
 In the present work, we propose a series of models that represent a comprehensive pathway of analyses that researchers need to follow when investigating complex interactions of traits in a diversification context.
@@
 In step four, we  ask if on top of the complex interaction there is a second unknown trait that can be driving differences in diversication, so we add to our bivariate diversification model a hidden trait.\newline
-
 Finallly, we ask if diploidization can be playing a role changing patterns of diversication.
@@
 %Our results emphasize the importance of considering traits not only in isolation, especially when there are strong correlations between them.
-
 %We investigate their individual and joint effects on diversification, and whether or not adding hidden states at different steps of modeling can help us generate hypothesis about the benefit of searching for a trait that can pass as a hidden state in a simple model.
 %R- I added this to link the paragraph before
-
 %Additionally, a lack of methods that allow for simultaneous inference of effects of other traits, which also influence diversification, can lead to incorrect conclusions about the focal trait \citep{rabosky_2015, beaulieu_2016}.

### 2019-03-05 manuscript/introduction.tex
%B's final edits to Introduction - moved trait intro together, then musse/hisse

@@
 \Citet{robertson_2011} found that the pathway from self-incompatible diploids to self-compatible polyploids is dominated by loss of self-incompatibility followed later by polyploidization over long timescales, but proceeds in one step via polyploidization of SI species over short timescales.
-We revisit this question with a greatly improved phylogeny and methods that allow for diversification rate differences.
+Greatly improved phylogeny and methods that allow for diversification rate differences allow far more powerful approach to such problems.
 % E: Technically, we allow for div effects in the rate estimates, but not in the pathways.
@@
 And dioecy.
-%B: Yes, let's.
+%B: Yes, agreed.
 
 %B: STOPPED HERE. 
-this is orphaned; we should not go to HiSSE before breeding systems.
-This yields an odd methods, traits, methods, traits, structure, where the methods are not related to the trait discussed.
- I will work on it in the next 24 hours. It logically flows from ploidy/BS and other trait existence that hidden state models ought to be useful 
-Another approach necessary to do a comprehensive evaluation of the role of polyploidy in diversification processes is the work proposed by \citet{beaulieu_2016}. 
-The authors proposed models to  allow for a `hidden' or unspecified state linked to the trait of interest for models of trait linked to diversification. 
-The presence of a hidden state can point out at sudden changes in the diversification rate under a value of a trait that are due to some unobserved source of heterogeneity and not the focal trait per se.
+Two methodological developments are critical for disentangling the relative importance of two or more traits in diversification.
+First, \citet{fitzjohn_2012} extended the state-dependent diversification ``*SSE" modeling framework \citep{maddison2007}, to allow estimating the effect of an arbitrary number of states (enabling multiple states, MuSSE, instead of binary states, BiSSE).
+This allows multiple states to represent trait combinations, particularly useful if one is interested in explicitly testing whether combinations of traits non-additively affect speciation or extinction.
+Second, \citet{beaulieu_2016} proposed an extension of the state-dependent diversification ``*SSE" model framework \citep{maddison2007}, which allows modelling `hidden' or unspecified traits, linked to the trait of interest.
+The presence of a hidden trait can uncover changes in the diversification rate due to unobserved sources of heterogeneity, not attributable to a focal trait.
 These hidden-state  models have effectively shown that traits that were believed to be drivers of diversification, in reality are not, and that it is an unobserved but associated trait responsible for differences that are initially found.
@@
 
-We present the most comprehensive data on ploidy and breeding system at species level in a phylogenetic context used in any trait linked diversification model.
-Our dataset consists of  a 595 dated Solanaceae phylogeny, with ploidy and/or breeding system information for the tips of the phylogeny to investigate the associations of these two traits with lineage diversification.
-In the present work, we propose a series of models that represent a comprehensive pathway of analyses that researchers need to follow when investigating complex interactions of traits in a diversification context.
-First, we present models where polyploidy and breeding system are modeled separatedly, and compare the resulting inferences to results of previous published models.
-Second, we add a hidden state to each of the models, in order to investigate if the focal trait maintains differences in diversification for focal states or if there is an unobserved but correlated hidden state that can drive the diversification pattern.
-Third, we ask if the hidden state is actually related to the complex interaction between breeding system and polyploidy and propose a model to consider both traits linked to diversification.
-In step four, we  ask if on top of the complex interaction there is a second unknown trait that can be driving differences in diversication, so we add to our bivariate diversification model a hidden trait.\newline
-Finallly, we ask if diploidization can be playing a role changing patterns of diversication.
-We perform model selection in a Bayesian framework to identify which models are fitting better the evidence of the Solanaceae data.
+% this paragraph had some line numbers and ligature-pasting problems?
+Here, we outline a broad strategy for the study of complex interactions among traits, in the context of diversification.
+The focus of our study is the individual role of ploidy and breeding systems, as well as their interaction, on the diversification process in Solanaceae. %sans pathway
+We rely on a comprehensive assembled dataset of ploidy and breeding system in nightshade species, and a large dated phylogeny, to evaluate a series of models and assess the influence of changes in these traits on speciation and extinction rates.
+%investigate
+First, we present models in which the effect of ploidy and breeding system are considered separately, and compare the inferences to previously published results. 
+Second, we consider the addition of a hidden trait to each of the models that examine traits individually, in order to investigate whether the focal trait directly affects differences in diversification, or if such differences are attributable to unobserved but correlated hidden states.
+Third, we ask whether the complex interaction between breeding system and ploidy may be explained by the hidden trait states. 
+We evaluate a multi-state model to consider both traits linked to diversification, and then consider whether an additional unobserved trait may be driving differences in diversification.
+Finally, we investigate the effects of diploidization in generating the patterns of diversification. 
+Our results highlight the importance of considering non-additive effects of traits on diversification, especially when there are strong biologically driven correlations between them.
+
+%BWe perform model selection in a Bayesian framework to identify which models are fitting better the evidence of the Solanaceae data.
 %Considering them jointly, however, reveals that the ploidy connection is removed by incorporating breeding system.
@@
 %We investigate their individual and joint effects on diversification, and whether or not adding hidden states at different steps of modeling can help us generate hypothesis about the benefit of searching for a trait that can pass as a hidden state in a simple model.
-%R- I added this to link the paragraph before
 %Additionally, a lack of methods that allow for simultaneous inference of effects of other traits, which also influence diversification, can lead to incorrect conclusions about the focal trait \citep{rabosky_2015, beaulieu_2016}.

### 2019-03-07 manuscript/main.tex
- fixing affiliations as requested by Will and Itay - fixed maddison_2007 citation

--- 2019-02-18 manuscript/main.tex
@@
 % Interacting effects of ploidy and breeding system on lineage diversification % E: but "effects" is too causal
-
+%
+%B: associations isn't exactly right; I think it somewhat undermines the causality from DP to IC; anyway, it's not particularly bad.
+%
+%% Interaction of ploidy and breeding system in lineage diversification %B: <- how 'bout that?
+%
 \begin{center}
-    \textbf{Interacting associations between ploidy, breeding system, and lineage diversification}
-\end{center}
+    \textsc{Interacting associations between ploidy, breeding system, and lineage diversification}
 
@@
 \noindent
-Rosana Zenil-Ferguson,%
-\endnote{University of Minnesota Twin Cities}
+Rosana Zenil-Ferguson$^{1,\dagger}$,
 %
 \noindent
-J.~Gordon Burleigh,%
-\endnote{University of Florida}
+J.~Gordon Burleigh$^{2}$,
 %
 \noindent
-William A.~Freyman,%
-\endnote{23 and me}
+William A.~Freyman$^{1}$,
 %
 \noindent
-Boris Igi\'c,%
-\endnote{University of Illinois}
+Boris Igi\'c$^{3}$,
 %
 \noindent
-Itay Mayrose,%
-\endnote{Tel Aviv University}
+Itay Mayrose$^{4}$,
 %
 and
-Emma E.~Goldberg%
-\endnote{University of Minnesota Twin Cities}
+Emma E.~Goldberg$^{1}$
+%
+\end{center}
 
+%
+\noindent$^{1}$Department of Ecology, Evolution, and Behavior, University of Minnesota, Saint Paul, MN 55108, U.S.A.\\
+\noindent$^{2}$Department of Biology, University of Florida, Gainesville, FL 32611, U.S.A.\\
+\noindent$^{3}$Department of Biological Sciences, University of Illinois at Chicago, Chicago, IL 60607, U.S.A. \\
+\noindent$^{4}$School of Plant Sciences and Food Security, Tel Aviv University, Tel Aviv 6997801, Israel.\\
+\noindent$^\dagger$\textit{Author for correspondence}.\\
 \vfill
-
-\theendnotes
-
-\noindent
-Author for correspondence: Rosana Zenil-Ferguson
-
-\vfill
-
-\noindent
-\textit{Running head:} Ploidy and breeding systems in Solanaceae
-% E: I've been changing "polyploidy" to "ploidy" in various places because "polyploidy" is one state of the "ploidy" trait, just as "SI" is one state of the "breeding system" trait
-
-\vfill
+\noindent\textit{Running head:} Ploidy and Breeding Systems in Solanaceae
+%\vfill
 
@@
 
-NSF DEB-1655478.
+%B: Perhaps other's funding sources for max BSF/NSF payoff? I added ours; no harm, I presume.
+NSF DEB-1655692 to BI, and DEB-1655478 to EEG.
 The Minnesota Supercomputing Institute (MSI) at the University of Minnesota provided computing resources for this project.

### 2019-03-07 manuscript/introduction.tex
polyploidy in the introduction revised

@@
 Species accumulate around the tree of life at different rates.
-A possible explanation for this phenomenon is that species possess traits that cause unequal rates of diversification. %B: new.
-The search for influential traits has been accelerated by a dramatic increases in data and an advance in methods, which now make use of far more information from phylogenetic trees \citep{maddison_2007}.
-Despite some setbacks \citep{maddison_2015, rabosky_2015, moore_2016}, advances in analytical methods \citep{fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017} are enabling the discovery of traits that alter the diversification patterns.
+A possible explanation for this phenomenon is that species possess various traits or character states that may differentially affect rates of diversification. %B: new.
+Dramatic increases in available phylogenetic and phenotypic data %citation
+as well as methodological advances %citation
+have greatly accelerated the search for such traits that influence diversification traits.
+Nevertheless, identifying focal traits associated with rates of speciation and extinction remains as a difficult statistical problem (see \citep{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}) .
 %B: the following is now orphaned, because it's inclusion merits addition of non-trait-focused methods; "potentially interesting biological processes"
-Nevertheless, the aim of identifying the focal trait and investigating its association with rates of speciation and extinction remains complicated. 
-This is, in part, because the context in which traits occur can lead to complex interactions with the diversification process (see \citet{caetano_2018, herrera_2018}). 
-Trait combinations provide examples of such context, and may be of particular interest, given the rapid accumulation of information about traits. %cite?
-Two exceedingly well-studied traits, ploidy and breeding system, are of particular interest. 
-Their individual state-dependent contributions to the process of diversification, as well as influence on each others' trait transitions, have long been the focus of speculation and study \citep{stebbins1950}.
+In part, this is a difficult problem because speciation and extinction rarely depend on a single trait; the context  in which traits occurs can lead to complex interactions resulting in a heterogenous speciation and extinction rates (\citep{beaulieu_2016, caetano_2018, herrera_2018}).
+Consequently, examining the association of only one trait when multiple complex interactions are changing diversification patterns can be misleading.
 
+In flowering plants, both ploidy-level and breeding system are well-studied traits that have been linked to diversification but may also affect each other \citep{stebbins1950}. 
 %B: IMO: ploidy = trait, di-, tri-, tetra-, hexa-, etc. (including "poly-") = trait states; the uncertainty we were discussing is whether "tri-" < "poly-" or "tri-" = "poly-"
@@
 % was [B&E banter] and % R- I think the trait is polyploidy and the state is ploidy (as 2, or +2). I will go with everyone else in the literature and use polyploidy
-Whole genome duplication is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
-Such changes is ploidy have the potential to affect many other phenotypes, and therefore alter a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
+Whole genome duplication (WGD) is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}, resulting in many plant species with a variety of ploidy levels.
 The prevalence of ploidy variation has been broadly considered as a salient feature of flowering plants for nearly a century \citep{stebbins1938}.
-Polyploids have long been thought to have lower (and even negative) rates of net diversification than diploids \citet{mayrose_2011, mayrose_2015}. 
-A number of recent studies, however, find common paleo-polyploidization, including at the base of many highly diverse angiosperm clades \citep{soltis_2014}. 
+
+WGDs have the potential to affect many phentoypes and impact a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
+Polyploids have long been thought to have lower (and even negative) rates of net diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
+Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that WGDS have played an important macroevolutionary role driving innovation and diversification in plants.
 %B: explanation for this edit, rm: "Using the genomic evidence  from the 1KP project" : this is the data, not the approach; Ks and tree reconciliation are particularly troubling and weak modes of inference.
 I am okay with having this point to what others have found, but I would prefer to avoid endorsing what I see as a deeply problematic set of methods.
-\citet{landis_2018}, for example, found 106 whole genome duplications within angiosperms that, combined with a state-dependent diversification model, resulted in approximately 60\% of whole genome duplication events potentially enhancing net diversification.
-The mere occurrence of such ancestral whole genome duplications points to an possibly important role of genome downsizing and diploidization in the diversification process \citep{soltis_2015diploidization, dodsworth_2015}.
-The presence of a significant and positive net diversification rate for (some) diploids can be  the result of a polyploid lineage that has been diploidized. 
-Under this scenario, rather than being a macroevolutionary dead-end, polyploidy may instead be positively associated with the diversification process.
+% R- I agree with your point.
+Tiley, Barker, and Burleigh just published last week a paper about Ks plots and the mixed normal model adjustments to it that "detect" WGDs.
+They found that in many cases those models lead to overestimating the number of WGDs plus the other issue is using the presence of WGD and then doing something like BAMM that quantitatively does not link via parameters is problematic.
+Evidence of paleo-polyploidizations within the genomes of diploid plants also revels the pervasiveness of diploidization, the return of a polyploid to a diploid state throughout the angiosperm phylogeny \citep{soltis_2015diploidization, dodsworth_2015}.
+Testing for the presence of diploidization in a diversification context is important because the presence of a significant and positive net diversification rate for diploids can be  the result of a polyploid lineage that has been diploidized.
+%closing this sentence to wrap how previous results would make sense even when polyploidy is the initiator of diversification
 

### 2019-03-07 manuscript/introduction.tex

@@
 Two methodological developments are critical for disentangling the relative importance of two or more traits in diversification.
-First, \citet{fitzjohn_2012} extended the state-dependent diversification ``*SSE" modeling framework \citep{maddison2007}, to allow estimating the effect of an arbitrary number of states (enabling multiple states, MuSSE, instead of binary states, BiSSE).
+First, \citet{fitzjohn_2012} extended the state-dependent diversification ``*SSE" modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multiple states, MuSSE, instead of binary states, BiSSE).
 This allows multiple states to represent trait combinations, particularly useful if one is interested in explicitly testing whether combinations of traits non-additively affect speciation or extinction.
-Second, \citet{beaulieu_2016} proposed an extension of the state-dependent diversification ``*SSE" model framework \citep{maddison2007}, which allows modelling `hidden' or unspecified traits, linked to the trait of interest.
+Second, \citet{beaulieu_2016} proposed an extension of the state-dependent diversification ``*SSE" model framework \citep{maddison_2007}, which allows modelling `hidden' or unspecified traits, linked to the trait of interest.
 The presence of a hidden trait can uncover changes in the diversification rate due to unobserved sources of heterogeneity, not attributable to a focal trait.

### 2019-03-07 manuscript/introduction.tex
various minor manuscript edits

@@
 have greatly accelerated the search for such traits that influence diversification traits.
-Nevertheless, identifying focal traits associated with rates of speciation and extinction remains as a difficult statistical problem (see \citep{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}) .
+Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a difficult statistical problem (see \citep{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}) .
 %B: the following is now orphaned, because it's inclusion merits addition of non-trait-focused methods; "potentially interesting biological processes"
@@
 % refs: stebbins, igic
-Previous phylogenetic analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, but they have not considered the possibility of other correlated traits driving this pattern.
+Previous phylogenetic analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014, freyman_2018}, but they did not test other correlated traits driving this pattern.
 Given that changes in ploidy and breeding systems may be causally related and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.

### 2019-03-07 manuscript/methods.tex
pathway note
various minor manuscript edits

@@
 We call this HiSSE model the \textit{I/C+A/B breeding system and hidden state model}.
+This model was previously used in \citet{freyman_2018} to examine the effect of
+breeding system on diversification in the plant family Onagraceae.
 
@@
 For all four models that consider ploidy changes, we allowed diploidization by adding a parameter $\delta$ connecting polyploidy state back to diploid or to self-compatible diploid in the mult-state models (denoted by $+/delta$ in the models of  \cref{table:marginallike}).
-Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010}.
+Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models \citep{mayrose_2010, glick2014chromevol, freyman_2017}.
 These types of classifications do not allow for a ploidy reversion.
@@
 
-Parameters for each of the 10 diversification models were estimated using custom code in the RevBayes \citep{hoehna_2016} environment.
-Code for analyses and key results is available at \url{https://github.com/roszenil/solploidy}.
-We included a correction for incomplete sampling in all analyses, based on assuming that the Solanaceae family has approximately 3,000 species ($s=595/3000$) as estimated by the Solanaceae Source project \citep{solsource}.
+Parameters for each of the 10 diversification models were estimated using RevBayes \citep{hoehna_2016}.
+Code for analyses and key results are available at \url{https://github.com/roszenil/solploidy}.
+We accounted for incomplete sampling in all analyses by fixing the probability of sampling a species at the present $s$ to $595/3000$ since the Solanaceae family has approximately 3,000 species as estimated by the Solanaceae Source project \citep{solsource}.
 For all 10 models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
 Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
-For each model, an MCMC chain was run for 96 hours in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the 10 models.
+For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run for 96 hours in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the 10 models.
 %B: the number of models changes here?
@@
 fixed it
-For each model, convergence and mixing of the MCMC was tested using the R library \texttt{coda} and the software package Tracer (see supplementary information for convergence plots).
-% fixme: eventually, add supp info fig ref
+Convergence and mixing of each MCMC 
+was determined by ensuring the effective sample size of each parameter was over 200.
 
-% R will add a bit about the ASR methods
-In the RevBayes code we calculated the probability of ancestral states at each node as performed chromosome number change models proposed in  \citet{freyman_2017}.
-Ancestral reconstructions for all models can be found in the supplementary information and show the maximum a posteriori of the marginal probability distributions for each of the 594 internal nodes for each of the 10 models.
+For each model, ancestral states at each node in the phylogeny were sampled jointly during the MCMC analyses.
+Ancestral state reconstructions for all models can be found in the supplementary information and show the maximum a posteriori of the marginal probability distributions for each of the 594 internal nodes for each of the 10 models.
 

### 2019-03-07 manuscript/introduction.tex
Included all coauthors comments

--- 2019-03-07 manuscript/introduction.tex
@@
 %B: the "Here, we" part should maybe come at the bottom of introduction, and it's a little early for here, in the first paragraph.
-Species accumulate around the tree of life at different rates.
+Species accumulate within the tree of life at different rates.
 A possible explanation for this phenomenon is that species possess various traits or character states that may differentially affect rates of diversification. %B: new.
@@
 %B: the following is now orphaned, because it's inclusion merits addition of non-trait-focused methods; "potentially interesting biological processes"
-In part, this is a difficult problem because speciation and extinction rarely depend on a single trait; the context  in which traits occurs can lead to complex interactions resulting in a heterogenous speciation and extinction rates (\citep{beaulieu_2016, caetano_2018, herrera_2018}).
+In part, this is a difficult problem because speciation and extinction rarely depend on a single trait; the context  in which traits occurs can lead to complex interactions resulting in a heterogenous speciation and extinction rates \citep{beaulieu_2016, caetano_2018, herrera_2018}.
 Consequently, examining the association of only one trait when multiple complex interactions are changing diversification patterns can be misleading.
 
-In flowering plants, both ploidy-level and breeding system are well-studied traits that have been linked to diversification but may also affect each other \citep{stebbins1950}. 
+In flowering plants, both ploidy and breeding system are well-studied traits that have been linked to diversification but may also affect each other \citep{stebbins1950}. 
 %B: IMO: ploidy = trait, di-, tri-, tetra-, hexa-, etc. (including "poly-") = trait states; the uncertainty we were discussing is whether "tri-" < "poly-" or "tri-" = "poly-"
@@
 WGDs have the potential to affect many phentoypes and impact a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
-Polyploids have long been thought to have lower (and even negative) rates of net diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
+Polyploids have long been thought to have lower rates of net diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
 Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that WGDS have played an important macroevolutionary role driving innovation and diversification in plants.
@@
 Evidence of paleo-polyploidizations within the genomes of diploid plants also revels the pervasiveness of diploidization, the return of a polyploid to a diploid state throughout the angiosperm phylogeny \citep{soltis_2015diploidization, dodsworth_2015}.
-Testing for the presence of diploidization in a diversification context is important because the presence of a significant and positive net diversification rate for diploids can be  the result of a polyploid lineage that has been diploidized.
-%closing this sentence to wrap how previous results would make sense even when polyploidy is the initiator of diversification
+Testing for the presence of diploidization in a diversification context is important because the presence of a positive net diversification rate for diploids can be  the result of a polyploid lineage that has been diploidized.
+Under this scenario, polyploidy should no longer be considered to slow down diversification, but instead, polyploidy would be the initiator of the diversification process.
+%R- This last phrase is from Itay and I rather like it.
 
-Ploidy has long been thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
+Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
+In particular, self-incompatibility systems cause a plant to reject its own pollen, and their loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution.
+% refs: stebbins, igic
+% Gordon comments here: Really?
+I find this hard to believe.There are a number of transitions among general selfing and outcrossing mating systems, but I think of SI as only occurring in a limiteed number of angiosperm families.
+Assigned to Boris
+Previous phylogenetic analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}.
+Faster net diversification of self-incompatible state was also found in the family  \textit{Onagraceae} \citep{freyman_2017} , where authors assumed the possibility of hidden states potentially driving diversification patterns.
+However, studying breeding system with other correlated traits and unobserved states driving diversification patterns simultaneously has not yet being explored.
+% Gordon comments here: First, it's a little confusing if you are talking about more SC or SI species more generally, or if you are talking specifically about genetic SI systems.
+I don't think the de Vos paper is dealing with SI systems.
+If your'e talking more generally about breeding system evolution there's a vast literature here, and I think  this is oversimplifying.
+For example, there are a lot of papers trying to link dioecy/hermaphroditism with diversification.
+(Vamosi papers on dioecy and diversification)
+% R- Boris what do you think is the best way to proceed here?
+Narrow it down to SI systems (deleting de Vos paper) or expanding it to talk about breeding systems in general and then saying that we will only focus on SI systems because of Solanaceae
+
+Ploidy might also affect the propensity for self-fertilization \citep{stebbins1950}. 
 The evidence for correlated shifts in mating system, from outcrossing to selfing, after polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
@@
 In clades with these systems, it is thus natural to consider the simultaneous macroevolution of polyploidy and breeding system.
+% Gordon comments: Here again is confused about breeding systems in general vs. SI genetic systems
 
-Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
-In particular, self-incompatibility systems cause a plant to reject its own pollen, and their loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution.
-% refs: stebbins, igic
-Previous phylogenetic analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, but they have not considered the possibility of other correlated traits driving this pattern.
 Given that changes in ploidy and breeding systems may be causally related and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
 This includes their joint influence on lineage diversification, and also potential patterns in the order of their transitions.
-For example, do losses of SI more commonly occur tied to polyploidization, or without a ploidy shift?
-Do polyploids arise more commonly from self-incompatible or self-compatible diploids?
+For example, it is possible to ask whether losses of  SI more commonly occurred tied to polyploidization, or without a ploidy shift.
+Or whether polyploids have arised more commonly from self-incompatible than from self-compatible diploids/
 \Citet{robertson_2011} found that the pathway from self-incompatible diploids to self-compatible polyploids is dominated by loss of self-incompatibility followed later by polyploidization over long timescales, but proceeds in one step via polyploidization of SI species over short timescales.
@@
 %B: STOPPED HERE. 
-Two methodological developments are critical for disentangling the relative importance of two or more traits in diversification.
-First, \citet{fitzjohn_2012} extended the state-dependent diversification ``*SSE" modeling framework \citep{maddison2007}, to allow estimating the effect of an arbitrary number of states (enabling multiple states, MuSSE, instead of binary states, BiSSE).
+Two methodological developments are critical for disentangling the complex interactions between two or more traits and their link to diversification.
+First, the model proposed by \citet{fitzjohn_2012} that extended the state-dependent diversification ``*SSE" modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multiple states, MuSSE, instead of binary states, BiSSE).
 This allows multiple states to represent trait combinations, particularly useful if one is interested in explicitly testing whether combinations of traits non-additively affect speciation or extinction.
-Second, \citet{beaulieu_2016} proposed an extension of the state-dependent diversification ``*SSE" model framework \citep{maddison2007}, which allows modelling `hidden' or unspecified traits, linked to the trait of interest.
-The presence of a hidden trait can uncover changes in the diversification rate due to unobserved sources of heterogeneity, not attributable to a focal trait.
-These hidden-state  models have effectively shown that traits that were believed to be drivers of diversification, in reality are not, and that it is an unobserved but associated trait responsible for differences that are initially found.
-When the hidden state and not the focal trait, is responsible for changes in the diversification process, it leaves behind the question of whether the hidden state corresponds to a real state of a trait that should be sought, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
-Alternatively, one can simultaneously consider the effects of more than one known trait, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
+Second, the model by \citet{beaulieu_2016} HiSSE proposed a first solution to the problem of  elevated false positive type I errors found often in the BiSSE framework \citep{goldberg_2012}.
+The authors argued that it is unlikely that the variation in speciation and extinction rates is only due to a focal trait but by considering heterogeneity in the diversification process as the null hypothesis, it is possible to parse out the importanceof the focal trait in diversification while other unobserved processes can happen.
+HiSSE models have effectively shown that some traits associated with diversification in BiSSE analyses are actually not associated with the diversification process, and instead a hidden state is linked to differences in the diversification rate.
+Such results raise the question of the identity of the hidden state, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
+Third, one can simultaneously consider the effects of multiple known traits using MuSSE, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
+However, MuSSE might also suffer from hight type I errors if there are other processes shaping diversification other than the traits observed \citep{caetano_2018}
+Therefore, a fourth option exists: considering both the presence of multiple known traits  linked to the diversification process, but also, some unknown or hidden state changing speciation and extinction rates.
+These four modeling strategies are proposed here in the context of studying complex interactions among traits, linked diversification.
 
-% this paragraph had some line numbers and ligature-pasting problems?
-Here, we outline a broad strategy for the study of complex interactions among traits, in the context of diversification.
 The focus of our study is the individual role of ploidy and breeding systems, as well as their interaction, on the diversification process in Solanaceae. %sans pathway 
-We rely on a comprehensive assembled dataset of ploidy and breeding system in nightshade species, and a large dated phylogeny, to evaluate a series of models and assess the influence of changes in these traits on speciation and extinction rates.
-%investigate
+Using ploidy and breeding system data from species in Solanaceae, and considering possible heterogeneity in the diversification process, we assess  the influence of these traits on speciation and extinction rates.
 First, we present models in which the effect of ploidy and breeding system are considered separately, and compare the inferences to previously published results. 
@@
 We evaluate a multi-state model to consider both traits linked to diversification, and then consider whether an additional unobserved trait may be driving differences in diversification.
-Finally, we investigate the effects of diploidization in generating the patterns of diversification. 
-Our results highlight the importance of considering non-additive effects of traits on diversification, especially when there are strong biologically driven correlations between them.
+In all these models we also investigate the potential effects of diploidization to explore whether the reversion of polyploidy changes diversification interpretations.
+Our results highlight the importance of considering non-additive effects of traits on net diversification rates, especially when there are strong biologically driven correlations amongst them.
 

### 2019-03-07 manuscript/methods.tex
Included all coauthors comments

--- 2019-03-02 manuscript/methods.tex
@@
 CCDB contains records from original sources that have multiple complex symbol patterns denoting multivalence, or irregularites of chromosome counts.
-After a first full round of curation via CCDBcurator, we contrasted results by hand and corrected chromosomes where CCDBcurator's output was not able to curate.
-Our hand curated records were also contrasted against the large dataset Solanaceae ploidy states from \citet{robertson_2011} and against ploidy data from the C-value DNA dataset from Kew royal botanic gardens \citep{bennett_2005}.
+After a first full round of curation via CCDBcurator, we contrasted results by hand and corrected chromosomes where CCDBcurator's output was not able to adequately correct records.
+Our hand-curated records were also contrasted against the ploidy dataset from \citet{robertson_2011} and against ploidy data in the  C-value DNA dataset from \citet{bennett_2005}.
   %B: perhaps "cleaned" should be explained, unless it's hard to identically repeat as-is. R- I added more (it only took 4months to hand-curate 14,000 entries
-By contrasting three sources of information, we were able to code taxa as diploid (D) or polyploid (P).
-For the majority of species, ploidy was assigned according to information from the original publications and the Kew royal botanic gardens C-value DNA resource \citep{bennett_2005}.
-For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus or based on information of incompatibility.
-If a species was incompatible without any chromosomal information it was assigned as diploid.
-For example, \textit{Solanum betaceum} did not include information about ploidy level but it has 24 chromosomes, since $x=12$ is the base chromosome number of the \textit{Solanum} genus \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid.
-Species with more than one ploidy level were assigned the smallest of most frequent ploidy level recorded.
+By comparing three different sources of information, we were able to code taxa as diploid (D) or polyploid (P).
+For the majority of species, ploidy was assigned according to information from the original publications included in the  C-value DNA dataset \citep{bennett_2005}.
+For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus or based on SI/SC classification.
+For example, if a species was incompatible without any chromosomal information it was assigned as diploid.
+This was the case of  \textit{Solanum betaceum} that did not include information about ploidy level but it has 24 chromosomes, since $x=12$ is the base chromosome number of the \textit{Solanum} genus \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid.
+Species with more than one ploidy level were assigned the most frequent ploidy level recorded or the smallest ploidy in case of frequency ties.
 % E: There was more here about cross-verifying that no P species were SI.  I removed it because there were conflicts before, but we resolved them.  This is addressed in the next paragraph.
@@
 Breeding system was scored as self-incompatible (I) or self-compatible (C) based on results curated from the literature and  original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
-Most species could unambiguously be coded as either I or C \citep{raduski_2012}.  But for diploids without breeding system information taxa was coded as both I or C.%B: most? Any exceptions?
+Many species could unambiguously be coded as either I or C \citep{raduski_2012}.  But for diploids without breeding system information taxa was coded as both I or C. %B: most? Any exceptions? R- many 
 Following previous work, we coded as I any species with functional I systems, even if C or dioecy was also reported.
 Dioecious species without functional I were coded as C.
-
 % E: I am trying to stick with I/C when talking about the state values, but using SI/SC when talking about the meaning of the trait (because this what people use normally). 
-
+%R- Okay- I was getting confused as well thanks for clarifying
 To those existing data sets, we added some additional records for chromosome number and breeding system.
@@
 A total of 405 taxa without any information about breeding system or polyploidy were excluded.
-Tips without trait data are much less informative for diversification parameters linked to trait values.
+%Tips without trait data might not improve point estimation of diversification parameters while increasing uncertainty in the estimates of trait linked diversification
+% R-This is very important but I have removed it because I think it is a full on paper by its own
 Including this many more species would have prohibitively slowed our analyses, especially those implementing the most complex models.
@@
 
-To investigate the association between ploidy level and diversification, we first defined a binary state speciation and extinction model (BiSSE, \citealt{maddison_2007}) in which taxa were classified as diploid (D) or polyploid (P) (\cref{figure:stateclassifications}).
-We call this the \textit{D/P ploidy} model. %B: emphasis on introduced term (quotes awkward)
-In a Bayesian framework, we obtained posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$), net diversification rates ($r_D=\lambda_D-\mu_D$, $r_P=\lambda_P-\mu_P$), and relative extinction rates ($\nu_D=\mu_D / \lambda_D$, $\nu_D=\mu_D / \lambda_D$) associated with each state.
-This analysis explores the same question as \citet{mayrose_2011, mayrose_2015}, using the polyploidization (parameter $\rho$, the transition rate from $D$ to $P$).
+In order to investigate the association between ploidy level and diversification, we first defined a binary state speciation and extinction model (BiSSE, \citealt{maddison_2007}) in which taxa were classified as diploid (D) or polyploid (P) (\cref{figure:stateclassifications}).
+We call this the \textit{D/P ploidy} model. This model is identical to the model proposed by \citet{mayrose_2011, mayrose_2015}, but inferences were performed in a Bayesian framework.
+ %B: emphasis on introduced term (quotes awkward)
+We calculated the posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$), net diversification rates ($r_D=\lambda_D-\mu_D$, $r_P=\lambda_P-\mu_P$), and relative extinction rates ($\nu_D=\mu_D / \lambda_D$, $\nu_D=\mu_D / \lambda_D$) associated with each state.
+Parameter $\rho$ represents the transition rate from $D$ to $P$.
 % E: presumably the motivation will be explained in the Intro
 %B: should the term above be "re-diploidization" or "diploidization" (as now)?
+% B (via E): need some explanation of what \delta means, biologically, when ploidy is scored like this
+% R- the explanation will happen in the discussion
 
-% B (via E): need some explanation of what \delta means, biologically, when ploidy is scored like this
-
-Our second model assesses the signal of diversification due to ploidy differences while also parsing out the heterogeneity of diversification rates due a possible unobserved trait.
-BiSSE-like models can suffer from a large false discovery rate because they fail to account for diversification rate changes that do not directly depend on the trait of interest \citep{rabosky_2015, beaulieu_2016}.
-Diversification rate differences explained by some trait other than ploidy, are accommodated by adding a hidden state (HiSSE model; \citealt{beaulieu_2016}).
+Our second model addressed the diversification due to ploidy differences while also parsing out the heterogeneity of diversification rates due a possible unobserved trait.
+Diversification rate differences explained by some trait other than ploidy were accommodated by adding a hidden state (HiSSE model; \citealt{beaulieu_2016}).
 %B: inserted "(trait)", because that seems to be the goal - a biological explanation?
-In this model, each of the observed diploid and polyploid states is subdivided by a binary hidden trait with states $A$ and $B$.
-We call this the \textit{D/P+A/B ploidy and hidden state} model. 
-We estimated the posterior probability distributions of speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$), extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$), net diversification rates ($r_{D_A},\ r_{D_B},\ r_{P_A},\ r_{P_B}$), and relative extinction rates ($\nu_{D_A},\ \nu_{D_B},\ \nu_{P_A},\ \nu_{P_B}$).
-In this model polyploidization rate  is again represented by $\rho$  and hidden states are symmetrical with rate $\alpha$. %B: same dip vs. re-dip question, as above.
+% R- Now this is mentioned in the intro so that  makes it easier to write here
+In this model, each of the observed diploid and polyploid states were subdivided by a binary hidden trait with states $A$ and $B$.
+This model was called the \textit{D/P+A/B ploidy and hidden state} model. 
+Posterior probability distributions of speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$), extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$), net diversification rates ($r_{D_A},\ r_{D_B},\ r_{P_A},\ r_{P_B}$), and relative extinction rates ($\nu_{D_A},\ \nu_{D_B},\ \nu_{P_A},\ \nu_{P_B}$) were estimated.
+In this model polyploidization rate  is again represented by $\rho$  and hidden state rates are symmetrical with parameter  $\alpha$. %B: same dip vs. re-dip question, as above.
 
@@
 
-To assess the effects of breeding system in the diversification process, we first fit model in which the states are self-incompatible (I) or self-compatible (C).
-This is the same as the analysis of \citet{goldberg_2010}, save for an updated phylogeny \citep{sarkinen_2013}.
-We call this BiSSE model the \textit{I/C breeding system} model. 
-To parse out the effect of breeding system on diversification, while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
-We call this HiSSE model the \textit{I/C+A/B breeding system and hidden state model}.
+The effects of breeding system in the diversification process were modeled using the self-incompatible (I) or self-compatible (C) states on the tips of the phylogeny.
+The proposed model is similar to the  analysis proposed by  \citet{goldberg_2010} but applied to the much larger dataset.
+The proposed model is also BiSSE-like and we refer to it as the \textit{I/C breeding system} model. 
+In order to parse out the effect of breeding system on diversification, while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
+We call this HiSSE model the \textit{I/C+A/B breeding system and hidden state model} .
+The I/C+A/B model is equivalent to the model used in \citet{freyman_2017} but we restricted the hidden state transition rates to be symmetrical.
 
 %B: we moved back from "SI" to "self-incompatibility" below. Maybe a no-no.
-For all breeding system models, we allow transitions from $I$ to $C$ (at rate $q_{IC}$) but not the reverse.
-Within Solanaceae, self-incompatibility is homologous in all species in which S-alleles were cloned, and controlled crosses performed.  
+Self-incompatibility is homologous in all species in which S-alleles are cloned, and controlled crosses are performed. % R- Boris this phrase reads weird but I don't know how to fix it  
 All species sampled to date, possess a non-self recognition, RNase-based, gametophytic self-incompatibility \citep[shared even with other euasterid families;][]{ramanauskas_2017}.
@@
 %B: may need to introduce "S-locus" earlier.
-This represents very strong evidence that the self-incompatible mechanism, and our $I$ state is ancestral to the Solanaceae, and did not arise independently within the family ($q_{CI}=0$).
+This represents very strong evidence that the self-incompatible mechanism, and our $I$ state is ancestral to the Solanaceae, and did not arise independently within the family.Therefore, for all breeding system models, we estimated a transition rate from $I$ to $C$  but not the reverse ($q_{CI}=0$).
  %B: I think I have this right <- someone please double-check.
@@
 
-If polyploidy and breeding system each influence lineage diversification individually, it is logical to examine their possible joint effects.
+Polyploidy and breeding system might have changed lineage diversification individually but these two traits also have complex interactions in Solanaceae species.
+Therefore, a multi-state model that investigates the contribution of both traits and the possible transitions between them was proposed.
 %B: unsure about edit, which was: "If both ploidy and breeding system have the potential to influence lineage diversification, we should logically consider their effects jointly."  E: okay; also try to add that their transitions interact
-We thus fit a multi-state model that includes both traits \citep[MuSSE,][]{fitzjohn_2012}.
-The three states in this model are self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids, which are always self-compatible (CP).
-As explained above, we did not include a state for self-incompatible polyploids because they are not observed in the data, and  that trait combination state is mechanistically predicted not to occur.
-We call this the \textit{ID/CD/CP ploidy and breeding system} model.
-The model has 9 parameters, six for diversification in each state ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and three for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
+% R- Is the above better?
+Self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids, which are always self-compatible (CP) were used as input for a MuSSE model \citep{fitzjohn_2012}.
+We did not include a state for self-incompatible polyploids because they are not observed in the data, and  that trait combination state is mechanistically predicted not to occur.
+The proposed model is called the \textit{ID/CD/CP ploidy and breeding system} and has six parameters for diversification in each state: ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction), and three for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
 The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
@@
 
-The ID/CD/CP model could potentially capture similar dynamics as earlier models, if the effects of the hidden state in D/P+A/B were effectively caused by breeding system (or its correlates), and the hidden state in I/C+A/B was effectively caused by ploidy.
-There is also the potential, however, for a hidden factor to be influencing diversification beyond both of our focal traits, and this could again mislead inferences.
-We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018, herrera_2018, huang_2018}.
+The ID/CD/CP model could also suffer from a hight type I error as the regular two state BiSSE model. 
+A hidden factor might alter diversification beyond both ploidy and breeding system misleading inferences \citep{rabosky_2015}.
+We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018,  huang_2018}.
 We refer to this as the \textit{ID/CD/CP+A/B} model.
-A fully parameterized version of this model would have 26 rate parameters \citep{herrera_2018}. 
+A fully parameterized version of this model would have 26 rate parameters. 
 Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fitted a simplified version with 15 parameters.
@@
 In the two-step pathway, the CD state is an intermediate: SI is first lost, and later the SC diploid undergoes polyploidization.
-We quantify the relative contribution of these pathways to polyploidy in two ways, each using the MAP rate estimates from the IC/CD/CP model.
+We quantify the relative contribution of these pathways to polyploidy in two ways, each using the maximum a posteriori (MAP) estimates of rates from the IC/CD/CP model.
 
-Both of our methods are based on a propogation matrix that describes flow from ID to CP, as in \citet{robertson_2011}.
+Both of our methods are based on a propagation matrix that describes flow from ID to CP, as in \citet{robertson_2011}.
 We insert an artificial division in the CP state, so that one substate contains the CP species that arrived via the one-step pathway and the other substate contains the CP species that arrived via the two-step pathway.
@@
 We therefore include the difference between speciation and extinction along the diagonal elements of the propogation matrix.
-As before matrix exponentiation provides the relative chance of changing from one state to any other state after time $t$.
+As before, matrix exponentiation provides the relative chance of changing from one state to any other state after time $t$.
 In this case, however, these are not probabilities because diversification changes the number of lineages as time passes.
-We can still use their ratios to consider the relative contribution of each pathway, though, analogous to the normalized age structure in a growing population.
+We can still use their ratios to consider the relative contribution of each pathway, though, analogous to the normalized age structure in a growing population. % R- here a reference would be nice
 
@@
 For all four models that consider ploidy changes, we allowed diploidization by adding a parameter $\delta$ connecting polyploidy state back to diploid or to self-compatible diploid in the mult-state models (denoted by $+/delta$ in the models of  \cref{table:marginallike}).
-Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010}.
+Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010, mayrose_2015}.
 These types of classifications do not allow for a ploidy reversion.
-Where indicated, the classification of ploidy for the data used in our models was based on chromosome multiplicity at the genus level.
+Where indicated, the classification of ploidy for the data used in our models was based on chromosome multiplicity at the genus level (data available at https://www.github.com/roszenil/solploidy/datasets).
 % ref data/sources table?
+%github
 However, the majority of the ploidy classifications were adopted from original studies with alternative sources of information (\eg geographic distribution, genus ploidy distribution) where ploidy was defined by authors that found evidence for it.

### 2019-03-07 manuscript/methods.tex

@@
 
-The effects of breeding system in the diversification process were modeled using the self-incompatible (I) or self-compatible (C) states on the tips of the phylogeny.
-The proposed model is similar to the  analysis proposed by  \citet{goldberg_2010} but applied to the much larger dataset.
-The proposed model is also BiSSE-like and we refer to it as the \textit{I/C breeding system} model. 
-In order to parse out the effect of breeding system on diversification, while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
-We call this HiSSE model the \textit{I/C+A/B breeding system and hidden state model} .
-The I/C+A/B model is equivalent to the model used in \citet{freyman_2017} but we restricted the hidden state transition rates to be symmetrical.
+To assess the effects of breeding system in the diversification process, we first fit model in which the states are self-incompatible (I) or self-compatible (C).
+This is the same as the analysis of \citet{goldberg_2010}, but with an updated phylogeny \citep{sarkinen_2013}.
+We call this BiSSE model the \textit{I/C breeding system} model. 
+To parse out the effect of breeding system on diversification, while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
+We call this HiSSE model the \textit{I/C+A/B breeding system and hidden state model}.
+The I/C+A/B model is equivalent to the model \citet{freyman_2017} used to study diversification in Onagraceae, but we restricted the hidden state transition rates to be symmetrical.
 
@@
 For all four models that consider ploidy changes, we allowed diploidization by adding a parameter $\delta$ connecting polyploidy state back to diploid or to self-compatible diploid in the mult-state models (denoted by $+/delta$ in the models of  \cref{table:marginallike}).
-Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010, mayrose_2015}.
+Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010, glick2014, mayrose_2015, freyman_2017}.
 These types of classifications do not allow for a ploidy reversion.

### 2019-03-07 manuscript/discussion.tex

@@
 The present work shows the importance of considering the trait linked diversification patterns under a multivariate approach.
-Species are created and go extinct based un multiple and often highly correlated phenotypes, understanding the speciation and extinction processes requires understanding of the evolutionary consequences that those trait correlations produce in organisms.
+Species are created and go extinct based on multiple and often highly correlated phenotypes, understanding the speciation and extinction processes requires understanding of the evolutionary consequences that those trait correlations produce in organisms.
 In the present work we show how considering both polyploidy and breeding system can disentangle the importance (or the lack of) of polyploidy when confronted with the evidence brought by breeding system.
@@
 
+% Will: I really like the pathways analyses and have a couple comments
+%
+% 1) Were the "without diversification" and "with diversification" analyses computed
+%    using the same MAP transition rate estimates? Transition rates among states will
+%    be very different if estimated using a model that is diversification-independent 
+%    compared to when estimated with a trait-dependent diversification model. Since the point of 
+%    showing these side-by-side seems to be emphasizing the importance of considering trait-dependent
+%    diversification, I wonder how different the "without diversification" contributions would be
+%    if calculated from transition rates estimated using a diversification-independent model.
+%
+% 2) I think stochastic mapping would be another useful perspective on the pathways question
+%    (and yeah, its a bit late in the game to propose analyses, so maybe next time!). With stochastic 
+%    mapping we could use the entire posterior estimate of the transition rates rather than a point
+%    estimate. But more importantly we'd get the relative contributions of the pathways estimated 
+%    over the actual Solanceae tree (which has a lot of short interdependent branches rather than 
+%    hypothetical long branches). 
+%
+% 3) I'll bet allowing cladogenetic transitions would make the one-step ID/CP even more dominant.
+%    I added the paragraph below regarding cladogenetic transitions, but it doesn't really fit with
+%    the rest of the discussion, so modify or discard as needed:
+Our work shows that a diploid Solanaceae lineage is much more likely to take a
+one-step ID/CP pathway to polyploidy rather than a two-step ID/CD/CP pathway. 
+We modeled both of these pathways as anagenetic character changes occurring within a species.
+However \citet{goldberg_2012} showed that I/C transitions are often associated with speciation events, and similarly \citet{freyman_2017} demonstrated that D/P transitions may also be associated with speciation events.
+Models that fail to consider transitions that occur at speciation (cladogenetic changes) may
+make misleading estimates of anagenetic transition rates.
+The HiSSE-based models introduced in our work here could be extended to incorporate cladogenetic transitions. 
+It is possible that allowing for cladogenetic transitions may enable the one-step pathway
+to even further dominate over the two-step pathway to polyploidy, but this remains to be tested in future work.
+
 % where WGD is in Solanaceae--motivation and delta rate analyses

### 2019-03-07 manuscript/introduction.tex

@@
 have greatly accelerated the search for such traits that influence diversification traits.
-Nevertheless, identifying focal traits associated with rates of speciation and extinction remains as a difficult statistical problem (see \citep{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}) .
-%B: the following is now orphaned, because it's inclusion merits addition of non-trait-focused methods; "potentially interesting biological processes"
+Nevertheless, identifying focal traits associated with rates of speciation and extinction remains as a difficult statistical problem (see \citep{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}).
+%B: the following is now orphaned, because its inclusion merits addition of non-trait-focused methods; "potentially interesting biological processes"
 In part, this is a difficult problem because speciation and extinction rarely depend on a single trait; the context in which traits occurs can lead to complex interactions resulting in a heterogenous speciation and extinction rates \citep{beaulieu_2016, caetano_2018, herrera_2018}.
@@
 Assigned to Boris
-Previous phylogenetic analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}.
-Faster net diversification of self-incompatible state was also found in the family  \textit{Onagraceae} \citep{freyman_2017} , where authors assumed the possibility of hidden states potentially driving diversification patterns.
+Previous phylogenetic analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, even when allowing the possibility that hidden states could drive diversification patterns \citep{freyman_2018}.
 However, studying breeding system with other correlated traits and unobserved states driving diversification patterns simultaneously has not yet being explored.

### 2019-03-07 manuscript/main.tex

@@
 %
-\noindent
 J.~Gordon Burleigh$^{2}$,
 %
-\noindent
-William A.~Freyman$^{1}$,
+William A.~Freyman$^{3}$,
 %
-\noindent
-Boris Igi\'c$^{3}$,
+Boris Igi\'c$^{4}$,
 %
-\noindent
-Itay Mayrose$^{4}$,
+Itay Mayrose$^{5}$,
 %
-and
-Emma E.~Goldberg$^{1}$
+and Emma E.~Goldberg$^{3}$
 %
 \end{center}
+%
+\noindent$^{2}$Department of Biology, University of Hawai'i M\={a}noa, Honolulu, HI 96822, U.S.A.\\
+%
+\noindent$^{2}$Department of Biology, University of Florida, Gainesville, FL 32611, U.S.A.\\
+%
+\noindent$^{3}$Department of Ecology, Evolution, and Behavior, University of Minnesota, Saint Paul, MN 55108, U.S.A.\\
+%
+\noindent$^{4}$Department of Biological Sciences, University of Illinois at Chicago, Chicago, IL 60607, U.S.A. \\
+%
+\noindent$^{5}$School of Plant Sciences and Food Security, Tel Aviv University, Tel Aviv 6997801, Israel.\\
+%
+\noindent$^\dagger$\textit{Author for correspondence:}.\\
 
-%
-\noindent$^{1}$Department of Ecology, Evolution, and Behavior, University of Minnesota, Saint Paul, MN 55108, U.S.A.\\
-\noindent$^{2}$Department of Biology, University of Florida, Gainesville, FL 32611, U.S.A.\\
-\noindent$^{3}$Department of Biological Sciences, University of Illinois at Chicago, Chicago, IL 60607, U.S.A. \\
-\noindent$^{4}$School of Plant Sciences and Food Security, Tel Aviv University, Tel Aviv 6997801, Israel.\\
-\noindent$^\dagger$\textit{Author for correspondence}.\\
 \vfill
-\noindent\textit{Running head:} Ploidy and Breeding Systems in Solanaceae
+
+\noindent\textit{Running head:} 
+Ploidy and Breeding Systems in Solanaceae
 %\vfill
 
-\noindent
-\textit{Keywords:} 
+\noindent \textit{Keywords:} 
 Polyploidy,

### 2019-03-07 manuscript/results.tex
- addition of a promised diploidization justification and fodder for discussion - requisite references - fixed spilled text on pathways table
various minor manuscript edits
pathways using proportion instead of ratio

@@
 When a hidden state is included (I/C+A/B model), the large net diversification difference persists for one of the hidden states but is removed for the other (\cref{figure:netdivall}D).
-Thus, differences in net diversification are best explained by both breeding system and an unknown factor.
+Thus, differences in net diversification are best explained by both breeding system and an unknown factor, similar to results found by \citet{freyman_2018}.
 % The transition rate from SI to SC is $q_{IC}=0.3$ for both the I/C and the I/C+A/B models.
@@
     \caption{
-        %B: FIXME: R&E check this re-wording to see if it is to your liking--I think I clarified it a bit, although with length increase.
-        %Contributions of the two pathways to polyploidy.
-        %When considering only rates of transitions among the states (left), the one-step pathway dominates on short timescales and the two-step on long timescales.
-        %When also considering diversification within each state (right), the one-step pathway dominates over any timescale.
-        %The lower panels plot the ratio of the two pathway contributions in the upper panels.
-        Contributions of the two pathways to polyploidy, the one-step ID$\rightarrow$CP transitions, and the two-step, ID$\rightarrow$CD$\rightarrow$CP transitions.
-        Panel (a) shows the contributions when considering only transition rates among the states (ignoring the diversification rate parameters). 
-        The one-step pathway dominates on short timescales and the two-step on long timescales.
-        Panels (b) shows the contributions when including diversification of lineages in each state. 
-        The one-step pathway, in which polyploidy breaks down SI, dominates over any timescale.
-        Panels (c) and (d) plot the ratio of the pathway contributions in the upper panels, (a) and (b), respectively.
+        Contributions of the two pathways to polyploidy.
+        The one-step pathway is direct ID$\rightarrow$CP transitions.
+        The two-step pathway consists of ID$\rightarrow$CD$\rightarrow$CP transitions.
+        When considering only rates of transitions among the states (ignoring the diversification rate parameters), the one-step pathway dominates on short timescales and the two-step on long timescales (left panels).
+        When also considering diversification within each state, the one-step pathway, in which polyploidization breaks down SI, dominates over any timescale (right panels).
+        The top panels show the separate contributions of each pathway.
+        The bottom panels show the proportional contribution of the one-step pathway (\ie one-step / [one-step + two-step]).
     }
-
-    the one-step pathway, from ID to CP
-
     \label{figure:pathways}

### 2019-03-07 summarystats/pathways.R

@@ -81,6 +81,9 @@ legend("topleft", c("ID to CP", "ID to CD to CP"), lty=c(1, 2), lwd=3)
 plot(times, pp$Wa/pp$Wb, log="y", type="l", lwd=3, xlab="elapsed time (My)", ylab="one-step : two-step", ylim=c(0.3, 100))
 abline(h=1, lty=3)
 
+plot(times, pp$Wa / (pp$Wa + pp$Wb), type="l", lwd=3, xlab="elapsed time (My)", ylab="proportion one-step", ylim=c(0,1))
+abline(h=0.5, lty=3)
+
 ### Try with diversification ###
 
 r1 <-  0.4607
@@ -103,6 +106,9 @@ legend("topleft", c("SI/D to SC/P", "SI/D to SC/D to SC/P"), lty=c(1, 2), lwd=3)
 plot(times, ppd$Wa/ppd$Wb, log="y", type="l", lwd=3, xlab="elapsed time (My)", ylab="one-step : two-step", ylim=c(0.3, 100))
 abline(h=1, lty=3)
 
+plot(times, ppd$Wa / (ppd$Wa + ppd$Wb), type="l", lwd=3, xlab="elapsed time (My)", ylab="proportion one-step", ylim=c(0,1))
+abline(h=0.5, lty=3)
+
 #--------------------------------------------------
 # Final plot
 #--------------------------------------------------
@@ -123,13 +129,11 @@ legend("topleft", c("one-step: ID to CP", "two-step: ID to CD to CP"), lty=c(2, 
 mtext("with diversification", line=1.5, font=3)
 
 # trans only
-plot(times, pp$Wa/pp$Wb, log="y", type="l", lwd=3, xlab="elapsed time (My)", ylab="contribution ratio", ylim=c(0.3, 100))
-abline(h=1, lty=3)
-legend("topright", c("one-step / two-step"), lty=1, lwd=3)
+plot(times, pp$Wa / (pp$Wa + pp$Wb), type="l", lwd=3, xlab="elapsed time (My)", ylab="proportion of one-step", ylim=c(0,1))
+abline(h=0.5, lty=3)
 
 # with div
-plot(times, ppd$Wa/ppd$Wb, log="y", type="l", lwd=3, xlab="elapsed time (My)", ylab="contribution ratio", ylim=c(0.3, 100))
-abline(h=1, lty=3)
[…]

### 2019-03-08 manuscript/discussion.tex
First part of the discussion completed, missing some of it

@@
 % - value of approach
-The present work shows the importance of considering the trait linked diversification patterns under a multivariate approach.
-Species are created and go extinct based on multiple and often highly correlated phenotypes, understanding the speciation and extinction processes requires understanding of the evolutionary consequences that those trait correlations produce in organisms.
-In the present work we show how considering both polyploidy and breeding system can disentangle the importance (or the lack of) of polyploidy when confronted with the evidence brought by breeding system.
+Disentangling the patterns of diversification linked to  multiple traits allowed us the opportunity to model and test complex hypotheses about the heterogenous diversification process in Solanaceae.
+Species are created and go extinct based on multiple and often highly correlated phenotypes.
+However, estimating the speciation and extinction rates remains challenging.
+Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as disscussed in  \citet{rabosky_2010, beaulieu_2015}),but also, a problem of inadequate model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focus trait first used to model diversification.
+We presented a roadmap with a series of analyses that can help scientists to identify if an unknown or unobservable trait is worth pursuing.
+By carefully aggregating and curating ploidy and breeding system data for Solanaceae, we showed how  the complex interaction between ploidy and breeding system  can lead to diversification conclusions that differ from the diversification conclusions of models that focus on one trait at a time.
 
 % - what happens if PD only, Hidden states
-Using the most complete dataset for polyploidy in a phylogenetic tree in Solanaceae, we were able to replicate the results found by \citet{mayrose_2011}, polyploids have a slower net diversification compare to diploids.
-Furthermore,  we also found polyploids had a high probability of having negative diversification which implies that polyploids can become a macroevolutionary dead-end, a result that was also found in the two large angiosperm diversification studies  \citet{mayrose_2011} and \citet{mayrose_2015}.
+Previous analyses of polyploidy linked to diversification showed how diploids had greater net diversification rates than polyploids when exploring across multiple clades of the angiosperm phylogeny \citet{mayrose_2011, mayrose_2015}.
+Using the most complete dataset for ploidy in a phylogenetic tree for Solanaceae, we were able to replicate the results found by \citet{mayrose_2011} where polyploids have a slower net diversification compare to diploids.
 However, we expanded this study to accommodate  background heterogeneity in the diversification process.
 When adding heterogeneity we found that it was more likely that an unobserved trait linked to diploid state was the one leading the net diversification patterns, and that there were some ``second-class"  diploids that were not different from  polyploids  in diversification terms (Figure 2A).
-This result lead us to our central question: \textit{what is that other trait linked diploids that makes them different in the diversification process? }.
+This result made us question whether there was a second trait linked to diploidy that could make a differential in the diversification progress.
 
 %     main finding about our traits
-In Solanaceae, our immediate intuition was to look into breeding system.
-Previous studies shown that self-incompatible Solanaceae species have also higher rates of diversification compared to their self-compatible counterparts \citep{goldberg_2012}.
-Self-incompatible species are diploid in our sample and also expected to be diploid due to ... (citation?).
-By considering both polyploidy and breeding system simultaneously for every species in our sample, it was possible to disentangle why some diploids were quantitatively different than polyploids.
-In the three-state diversification model ID/CD/CP, we found that self-incompatible diploids have faster and positive rates than self-compatible diploids and polyploids, and that the difference between the  rates of net diversification of self-compatible diploids is not as large (Figures 2C) as first found by binary trait diversification models (Figures 2A).
-This result is important, since it aligns with the net diversification results of the  D/P+A/B model, where a `	`hidden-trait" seem to be dictating the diversification pattern.
-By adding breeding system, we were able to hint at which that hidden-trait possibly is.
-Therefore, we consider that finding a heterogenous result in the hidden trait approaches should be be treated as evidence of a second trait that is necessary to consider.
+For Solanaceae, looking into self-incompatibility was a logical step.
+Previous studies have shown that self-incompatible Solanaceae species have also higher rates of diversification compared to their self-compatible counterparts \citep{goldberg_2012}. 
+Again, in our sample when looking only into breeding system linked to the diversification process we were able to replicate the slower net diversification for self-compatible taxa (Figure 2C).
+When this model was expanded to incorporate hidden states we found again that some self-incompatible taxa have faster net diversification rates than self-compatible taxa.
+However, there is again a  different class of taxa where those differences are not as striking (Figure 2D).
+However, investigating both  both polyploidy and breeding system simultaneously for every species in our sample, allowed us to further investigate why some diploids were quantitatively different than polyploids, or why some self-incompatible taxa would be different in diversification terms from some self-compatible when a second class is not.
+In the three-state diversification model ID/CD/CP, we represented the complex interactions between ploidy and breeding system.
+We found that self-incompatible diploids have faster and positive rates than self-compatible diploids and polyploids, and that the difference between the  rates of net diversification of self-compatible diploids is not as large (Figure 2E) as first found by binary trait diversification models (Figure 2A).
+This result is key, it not only aligned with the net diversification results from the D/P+A/B model, where a `	`hidden-trait" seem to be dictating the diversification pattern but it also aligned with the hidden state model I/C+A/B where some class of self-compatible net diversification rate might be overlapping with two net diversification rates of self-incompatible.
+% Removed by R. Therefore, we consider that finding a heterogenous result in the hidden trait approaches should be be treated as evidence of a second trait that is necessary to consider.
+% Removed by R.
 Pursuing knowledge of  such trait can result on a clearer picture of the importance of trait linked diversification patterns, but also on a better reconstruction on past events in phylogenies.
-
-% - diversification  (frame as no-delta first)
 
@@
 %    if calculated from transition rates estimated using a diversification-independent model.
+%  R- Good point. Emma will clarify
 %
@@
 %    hypothetical long branches). 
-%
+%R- It will, it was actually on my code I think for the ID/CD/CP I have the stochastic mapping trees.
+We can do it for the revision.
+I think the only model where it could not finish was the ID/CD/CP+A/B (was taking forever and we had the restriction of time from MSI)
+
 % 3) I'll bet allowing cladogenetic transitions would make the one-step ID/CP even more dominant.
@@
 By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with stochastic models.
+In past studies by \citet{mayrose_2011, mayrose_2015} diploidization could not be formally  considered in models because ploidy classification was done via chromosome change models (i.e.
+Chromevol \citep{glick2014}) that do not allow for reversion of polyploidy.
+However, in our dataset ploidy was assigned based on data aggregated from independent studies (see supplementary information and citations in archived dataset) that come from field or cytogenetic studies.
+In the cases where ploidy level was assigned by us based chromosome multiplicity at the genus level, estimating diploidization  might be a potential issue due to ploidy misclassification.
 
@@
 If this is the case, then all of the genomes may have been subsequently re-diploidized, because common base chromosome numbers this and related families are n=11-12.
+
 %Not unlikely, because Arabidopsis n=5.
@@
 %B: Once this is re-arranged, and made to fit with the rest of the discussion, add what this has to do with re-diploidization and how to interpret $\delta$.
+Therefore, it is not surprising that for all models diploidization  had wide credible intervals for parameter $\delta$.
+Despite the classification and genomic limitations to assign ploidy levels and diploidization, we found support for models that include pararmeter $\delta$, meaning that further studying this process is worth pursuing.
 
@@
 %     irreversibility SI -> SC assumption
+
+% Missing a good concluding paragraph.

### 2019-03-08 manuscript/introduction.tex
- re-working Introduction

@@
+\begin{quote}
+\em{``Among life history traits, reproductive characters that determine mating system patterns are perhaps the most influential in governing macroevolution." 
+}
+\\
+\hspace*{\fill}\rm{ Barrett et al. (1996)}
+\end{quote}
+
 \section{Introduction}
@@
 
-%B: the key alteration below is an attempt at easing the transitions of the complicated existing logical structure, which intermixes advances with setbacks and problems.
-%B: the "Here, we" part should maybe come at the bottom of introduction, and it's a little early for here, in the first paragraph.
+%B->%E: instead of a major reorganization, I added a breeding system quote from Barrett that presages our results. How's that? It seemed daunting to reorganize ploidy-first to BS-first.
+
 Species accumulate within the tree of life at different rates.
-A possible explanation for this phenomenon is that species possess various traits or character states that may differentially affect rates of diversification. %B: new.
-Dramatic increases in available phylogenetic and phenotypic data %citation
-as well as methodological advances %citation
-have greatly accelerated the search for such traits that influence diversification traits.
-Nevertheless, identifying focal traits associated with rates of speciation and extinction remains as a difficult statistical problem (see \citep{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}).
-%B: the following is now orphaned, because its inclusion merits addition of non-trait-focused methods; "potentially interesting biological processes"
-In part, this is a difficult problem because speciation and extinction rarely depend on a single trait; the context in which traits occurs can lead to complex interactions resulting in a heterogenous speciation and extinction rates \citep{beaulieu_2016, caetano_2018, herrera_2018}.
-Consequently, examining the association of only one trait when multiple complex interactions are changing diversification patterns can be misleading.
+A possible explanation for this phenomenon is that species possess various traits or character states that may differentially affect rates of diversification. 
+Dramatic increases in available phylogenetic and phenotypic data as well as methodological advances have greatly accelerated the search for traits that influence diversification traits.
+Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a difficult statistical problem (e.g., \citealt{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}).
+%is it only a "statistical" problem?
+In part, the difficulties arise because speciation and extinction may not depend on a single trait, and the context in which traits occur can lead to complex interactions resulting in a heterogenous speciation and extinction rates \citep{beaulieu_2016, caetano_2018, herrera_2018}.
+Consequently, examining the association of only one trait with diversification patterns can be misleading. %B rm: "when multiple complex interactions are changing" <- explained above
 
-In flowering plants, both ploidy and breeding system are well-studied traits that have been linked to diversification but may also affect each other \citep{stebbins1950}. 
-%B: IMO: ploidy = trait, di-, tri-, tetra-, hexa-, etc. (including "poly-") = trait states; the uncertainty we were discussing is whether "tri-" < "poly-" or "tri-" = "poly-"
-%(sorry about the confusion: I accidentally mis-wrote in the commented-out banter with %E.)
-% was [B&E banter] and % R- I think the trait is polyploidy and the state is ploidy (as 2, or +2). I will go with everyone else in the literature and use polyploidy
-Whole genome duplication (WGD) is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}, resulting in many plant species with a variety of ploidy levels.
-The prevalence of ploidy variation has been broadly considered as a salient feature of flowering plants for nearly a century \citep{stebbins1938}.
-
+%Breeding system and ploidy are perhaps the best-studied traits linked to diversification \citep{stebbins1950}. %B rm: no mention of breeding system in this paragraph?
+Whole genome duplication (WGD) is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
+The widespread prevalence of variation in ploidy has long been considered a salient feature of flowering plant lineages \citep{stebbins1938}. 
 WGDs have the potential to affect many phentoypes and impact a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
 Polyploids have long been thought to have lower rates of net diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
-Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that WGDS have played an important macroevolutionary role driving innovation and diversification in plants.
-%B: explanation for this edit, rm: "Using the genomic evidence  from the 1KP project" : this is the data, not the approach; Ks and tree reconciliation are particularly troubling and weak modes of inference.
-I am okay with having this point to what others have found, but I would prefer to avoid endorsing what I see as a deeply problematic set of methods.
-% R- I agree with your point.
-Tiley, Barker, and Burleigh just published last week a paper about Ks plots and the mixed normal model adjustments to it that "detect" WGDs.
-They found that in many cases those models lead to overestimating the number of WGDs plus the other issue is using the presence of WGD and then doing something like BAMM that quantitatively does not link via parameters is problematic.
-Evidence of paleo-polyploidizations within the genomes of diploid plants also revels the pervasiveness of diploidization, the return of a polyploid to a diploid state throughout the angiosperm phylogeny \citep{soltis_2015diploidization, dodsworth_2015}.
-Testing for the presence of diploidization in a diversification context is important because the presence of a positive net diversification rate for diploids can be  the result of a polyploid lineage that has been diploidized.
-Under this scenario, polyploidy should no longer be considered to slow down diversification, but instead, polyploidy would be the initiator of the diversification process.
+Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that WGDs have played an important macroevolutionary role driving innovation and diversification in plants.
+Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies the pervasiveness of diploidization, the return of polyploids to the diploid state throughout the angiosperm phylogeny \citep{soltis_2015diploidization, dodsworth_2015}.
+Testing for the presence of diploidization in the context of diversification is important because the presence of a positive net diversification rate for diploids can be the result of an ancestrally polyploid lineage.
+Under this scenario, polyploidy would no longer be considered to reliably slow down diversification, but instead, polyploidy would be the initiator of the diversification process.
 %R- This last phrase is from Itay and I rather like it.
+<-- %B: Hmmmm.
+How it would be considered an "initiator of the diversification process?" I mean, if it is associated with lower diversification.
+I think this should read "...polyploidy is no longer be considered an evolutionary dead end." In other words, I also agree with both %R and Itay, but not with what the current phrasing is asserting.
 
-Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
-In particular, self-incompatibility systems cause a plant to reject its own pollen, and their loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution.
-% refs: stebbins, igic
+Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and they crucially affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
+In particular, self-incompatibility systems cause a plant to reject its own pollen, and their loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974,igic_2008}.
 % Gordon comments here: Really?
-I find this hard to believe.There are a number of transitions among general selfing and outcrossing mating systems, but I think of SI as only occurring in a limiteed number of angiosperm families.
-Assigned to Boris
-Previous phylogenetic analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, even when allowing the possibility that hidden states could drive diversification patterns \citep{freyman_2018}.
+I find this hard to believe...
+I think of SI as only occurring in a limited number of angiosperm families.
+%B: Ha.
+I won't take it personally!
+True, it is present in a "limited" number of families, that number being about 150, including all of the largest families except for Lamiaceae (Apiaceae, Gesneriaceae, and Cucurbitaceae are the next in size).
+So, I am not sure how to show that it is "the most" common, but even this paper produces data to indicate that it is at the very least ca.
+twice as common as WGD, which probably qualifies it as "one of the most replicated transitions." %B: Also, "The evolutionary pathway from obligate outcrossing based upon self-incompatibility to predominant self-fertilization has probably been followed by more different lines of evolution in flowering plants than has any other." (Stebbins 1974)
+Previous analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, even when allowing the possibility that hidden states could drive diversification patterns \citep{freyman_2018}.
 However, studying breeding system with other correlated traits and unobserved states driving diversification patterns simultaneously has not yet being explored.
@@
 
-Ploidy might also affect the propensity for self-fertilization \citep{stebbins1950}. 
-The evidence for correlated shifts in mating system, from outcrossing to selfing, after polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
+Ploidy may also affect the propensity for self-fertilization \citep{stebbins1950}. 
+The evidence for correlated shifts in mating system, from outcrossing to selfing, following WGD, appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
 In other cases, however, polyploidy is not only suspected to correlate with breeding system but causes the transition from self-incompatibility (SI) to self-compatibility(SC) \citep{stout1942, lewis1947}.
@@
 Maybe when we have more time for the revision?
-% E: Relatedly, I'm wondering if pathways should be a separate paper.
-There is the robertson2011 question, the modification that allows for diversification, and the charlesworth1985 question.
-And dioecy.
-%B: Yes, agreed.
 
-%B: STOPPED HERE. 
 Two methodological developments are critical for disentangling the complex interactions between two or more traits and their link to diversification.
-First, the model proposed by \citet{fitzjohn_2012} that extended the state-dependent diversification ``*SSE" modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multiple states, MuSSE, instead of binary states, BiSSE).
+First, the model proposed by \citet{fitzjohn_2012} that extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multiple states, MuSSE, instead of binary states, BiSSE).
 This allows multiple states to represent trait combinations, particularly useful if one is interested in explicitly testing whether combinations of traits non-additively affect speciation or extinction.
 Second, the model by \citet{beaulieu_2016} HiSSE proposed a first solution to the problem of  elevated false positive type I errors found often in the BiSSE framework \citep{goldberg_2012}.
-The authors argued that it is unlikely that the variation in speciation and extinction rates is only due to a focal trait but by considering heterogeneity in the diversification process as the null hypothesis, it is possible to parse out the importanceof the focal trait in diversification while other unobserved processes can happen.
+The authors argued that it is unlikely that the variation in speciation and extinction rates is only due to a focal trait but by considering heterogeneity in the diversification process as the null hypothesis, it is possible to parse out the importance of the focal trait in diversification while other unobserved processes can happen.
 HiSSE models have effectively shown that some traits associated with diversification in BiSSE analyses are actually not associated with the diversification process, and instead a hidden state is linked to differences in the diversification rate.

### 2019-03-08 manuscript/introduction.tex
- first 2/3 of Introduction

--- 2019-03-07 manuscript/introduction.tex
@@
 %B: the "Here, we" part should maybe come at the bottom of introduction, and it's a little early for here, in the first paragraph.
-Species accumulate within the tree of life at different rates.
-A possible explanation for this phenomenon is that species possess various traits or character states that may differentially affect rates of diversification. %B: new.
-Dramatic increases in available phylogenetic and phenotypic data %citation
-as well as methodological advances %citation
-have greatly accelerated the search for such traits that influence diversification traits.
-Nevertheless, identifying focal traits associated with rates of speciation and extinction remains as a difficult statistical problem (see \citep{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}).
-%B: the following is now orphaned, because its inclusion merits addition of non-trait-focused methods; "potentially interesting biological processes"
-In part, this is a difficult problem because speciation and extinction rarely depend on a single trait; the context in which traits occurs can lead to complex interactions resulting in a heterogenous speciation and extinction rates \citep{beaulieu_2016, caetano_2018, herrera_2018}.
-Consequently, examining the association of only one trait when multiple complex interactions are changing diversification patterns can be misleading.
+Species accumulate around the tree of life at different rates.
+A possible explanation for this phenomenon is that species possess traits that cause unequal rates of diversification. %B: new.
+The search for influential traits has been accelerated by a dramatic increases in data and an advance in methods, which now make use of far more information from phylogenetic trees \citep{maddison_2007}.
+Despite some setbacks \citep{maddison_2015, rabosky_2015, moore_2016}, advances in analytical methods \citep{fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017} are enabling the discovery of traits that alter the diversification patterns.
+%B: the following is now orphaned, because it's inclusion merits addition of non-trait-focused methods; "potentially interesting biological processes"
+Nevertheless, the aim of identifying the focal trait and investigating its association with rates of speciation and extinction remains complicated. 
+This is, in part, because the context in which traits occur can lead to complex interactions with the diversification process (see \citet{caetano_2018, herrera_2018}). 
+Trait combinations provide examples of such context, and may be of particular interest, given the rapid accumulation of information about traits. %cite?
+Two exceedingly well-studied traits, ploidy and breeding system, are of particular interest. 
+Their individual state-dependent contributions to the process of diversification, as well as influence on each others' trait transitions, have long been the focus of speculation and study \citep{stebbins1950}.
 
-In flowering plants, both ploidy and breeding system are well-studied traits that have been linked to diversification but may also affect each other \citep{stebbins1950}. 
 %B: IMO: ploidy = trait, di-, tri-, tetra-, hexa-, etc. (including "poly-") = trait states; the uncertainty we were discussing is whether "tri-" < "poly-" or "tri-" = "poly-"
@@
 % was [B&E banter] and % R- I think the trait is polyploidy and the state is ploidy (as 2, or +2). I will go with everyone else in the literature and use polyploidy
-Whole genome duplication (WGD) is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}, resulting in many plant species with a variety of ploidy levels.
+Whole genome duplication is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
+Such changes is ploidy have the potential to affect many other phenotypes, and therefore alter a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
 The prevalence of ploidy variation has been broadly considered as a salient feature of flowering plants for nearly a century \citep{stebbins1938}. 
-
-WGDs have the potential to affect many phentoypes and impact a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
-Polyploids have long been thought to have lower rates of net diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
-Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that WGDS have played an important macroevolutionary role driving innovation and diversification in plants.
+Polyploids have long been thought to have lower (and even negative) rates of net diversification than diploids \citet{mayrose_2011, mayrose_2015}. 
+A number of recent studies, however, find common paleo-polyploidization, including at the base of many highly diverse angiosperm clades \citep{soltis_2014}. 
 %B: explanation for this edit, rm: "Using the genomic evidence  from the 1KP project" : this is the data, not the approach; Ks and tree reconciliation are particularly troubling and weak modes of inference.
 I am okay with having this point to what others have found, but I would prefer to avoid endorsing what I see as a deeply problematic set of methods.
-% R- I agree with your point.
-Tiley, Barker, and Burleigh just published last week a paper about Ks plots and the mixed normal model adjustments to it that "detect" WGDs.
-They found that in many cases those models lead to overestimating the number of WGDs plus the other issue is using the presence of WGD and then doing something like BAMM that quantitatively does not link via parameters is problematic.
-Evidence of paleo-polyploidizations within the genomes of diploid plants also revels the pervasiveness of diploidization, the return of a polyploid to a diploid state throughout the angiosperm phylogeny \citep{soltis_2015diploidization, dodsworth_2015}.
-Testing for the presence of diploidization in a diversification context is important because the presence of a positive net diversification rate for diploids can be  the result of a polyploid lineage that has been diploidized.
-Under this scenario, polyploidy should no longer be considered to slow down diversification, but instead, polyploidy would be the initiator of the diversification process.
-%R- This last phrase is from Itay and I rather like it.
+\citet{landis_2018}, for example, found 106 whole genome duplications within angiosperms that, combined with a state-dependent diversification model, resulted in approximately 60\% of whole genome duplication events potentially enhancing net diversification.
+The mere occurrence of such ancestral whole genome duplications points to an possibly important role of genome downsizing and diploidization in the diversification process \citep{soltis_2015diploidization, dodsworth_2015}.
+The presence of a significant and positive net diversification rate for (some) diploids can be  the result of a polyploid lineage that has been diploidized. 
+Under this scenario, rather than being a macroevolutionary dead-end, polyploidy may instead be positively associated with the diversification process.
 
-Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
-In particular, self-incompatibility systems cause a plant to reject its own pollen, and their loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution.
-% refs: stebbins, igic
-% Gordon comments here: Really?
-I find this hard to believe.There are a number of transitions among general selfing and outcrossing mating systems, but I think of SI as only occurring in a limiteed number of angiosperm families.
-Assigned to Boris
-Previous phylogenetic analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, even when allowing the possibility that hidden states could drive diversification patterns \citep{freyman_2018}.
-However, studying breeding system with other correlated traits and unobserved states driving diversification patterns simultaneously has not yet being explored.
-% Gordon comments here: First, it's a little confusing if you are talking about more SC or SI species more generally, or if you are talking specifically about genetic SI systems.
-I don't think the de Vos paper is dealing with SI systems.
-If your'e talking more generally about breeding system evolution there's a vast literature here, and I think  this is oversimplifying.
-For example, there are a lot of papers trying to link dioecy/hermaphroditism with diversification.
-(Vamosi papers on dioecy and diversification)
-% R- Boris what do you think is the best way to proceed here?
-Narrow it down to SI systems (deleting de Vos paper) or expanding it to talk about breeding systems in general and then saying that we will only focus on SI systems because of Solanaceae
-
-Ploidy might also affect the propensity for self-fertilization \citep{stebbins1950}. 
+Ploidy has long been thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
 The evidence for correlated shifts in mating system, from outcrossing to selfing, after polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
@@
 In clades with these systems, it is thus natural to consider the simultaneous macroevolution of polyploidy and breeding system.
-% Gordon comments: Here again is confused about breeding systems in general vs. SI genetic systems
 
+Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
+In particular, self-incompatibility systems cause a plant to reject its own pollen, and their loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution.
+% refs: stebbins, igic
+Previous phylogenetic analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, but they have not considered the possibility of other correlated traits driving this pattern.
 Given that changes in ploidy and breeding systems may be causally related and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
 This includes their joint influence on lineage diversification, and also potential patterns in the order of their transitions.
-For example, it is possible to ask whether losses of  SI more commonly occurred tied to polyploidization, or without a ploidy shift.
-Or whether polyploids have arised more commonly from self-incompatible than from self-compatible diploids/
+For example, do losses of SI more commonly occur tied to polyploidization, or without a ploidy shift?
+Do polyploids arise more commonly from self-incompatible or self-compatible diploids?
 \Citet{robertson_2011} found that the pathway from self-incompatible diploids to self-compatible polyploids is dominated by loss of self-incompatibility followed later by polyploidization over long timescales, but proceeds in one step via polyploidization of SI species over short timescales.
@@
 %B: STOPPED HERE. 
-Two methodological developments are critical for disentangling the complex interactions between two or more traits and their link to diversification.
-First, the model proposed by \citet{fitzjohn_2012} that extended the state-dependent diversification ``*SSE" modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multiple states, MuSSE, instead of binary states, BiSSE).
+Two methodological developments are critical for disentangling the relative importance of two or more traits in diversification.
+First, \citet{fitzjohn_2012} extended the state-dependent diversification ``*SSE" modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multiple states, MuSSE, instead of binary states, BiSSE).
 This allows multiple states to represent trait combinations, particularly useful if one is interested in explicitly testing whether combinations of traits non-additively affect speciation or extinction.
-Second, the model by \citet{beaulieu_2016} HiSSE proposed a first solution to the problem of  elevated false positive type I errors found often in the BiSSE framework \citep{goldberg_2012}.
-The authors argued that it is unlikely that the variation in speciation and extinction rates is only due to a focal trait but by considering heterogeneity in the diversification process as the null hypothesis, it is possible to parse out the importanceof the focal trait in diversification while other unobserved processes can happen.
-HiSSE models have effectively shown that some traits associated with diversification in BiSSE analyses are actually not associated with the diversification process, and instead a hidden state is linked to differences in the diversification rate.
-Such results raise the question of the identity of the hidden state, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
-Third, one can simultaneously consider the effects of multiple known traits using MuSSE, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
-However, MuSSE might also suffer from hight type I errors if there are other processes shaping diversification other than the traits observed \citep{caetano_2018}
-Therefore, a fourth option exists: considering both the presence of multiple known traits  linked to the diversification process, but also, some unknown or hidden state changing speciation and extinction rates.
-These four modeling strategies are proposed here in the context of studying complex interactions among traits, linked diversification.
+Second, \citet{beaulieu_2016} proposed an extension of the state-dependent diversification ``*SSE" model framework \citep{maddison_2007}, which allows modelling `hidden' or unspecified traits, linked to the trait of interest.
+The presence of a hidden trait can uncover changes in the diversification rate due to unobserved sources of heterogeneity, not attributable to a focal trait.
+These hidden-state  models have effectively shown that traits that were believed to be drivers of diversification, in reality are not, and that it is an unobserved but associated trait responsible for differences that are initially found.
+When the hidden state and not the focal trait, is responsible for changes in the diversification process, it leaves behind the question of whether the hidden state corresponds to a real state of a trait that should be sought, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
+Alternatively, one can simultaneously consider the effects of more than one known trait, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
 
+% this paragraph had some line numbers and ligature-pasting problems?
+Here, we outline a broad strategy for the study of complex interactions among traits, in the context of diversification.
 The focus of our study is the individual role of ploidy and breeding systems, as well as their interaction, on the diversification process in Solanaceae. %sans pathway
-Using ploidy and breeding system data from species in Solanaceae, and considering possible heterogeneity in the diversification process, we assess  the influence of these traits on speciation and extinction rates.
+We rely on a comprehensive assembled dataset of ploidy and breeding system in nightshade species, and a large dated phylogeny, to evaluate a series of models and assess the influence of changes in these traits on speciation and extinction rates.
+%investigate
 First, we present models in which the effect of ploidy and breeding system are considered separately, and compare the inferences to previously published results. 
@@
 We evaluate a multi-state model to consider both traits linked to diversification, and then consider whether an additional unobserved trait may be driving differences in diversification.
-In all these models we also investigate the potential effects of diploidization to explore whether the reversion of polyploidy changes diversification interpretations.
-Our results highlight the importance of considering non-additive effects of traits on net diversification rates, especially when there are strong biologically driven correlations amongst them.
+Finally, we investigate the effects of diploidization in generating the patterns of diversification. 
+Our results highlight the importance of considering non-additive effects of traits on diversification, especially when there are strong biologically driven correlations between them.
 

### 2019-03-09 manuscript/introduction.tex
tex typos

@@
 Nevertheless, the aim of identifying the focal trait and investigating its association with rates of speciation and extinction remains complicated. 
-This is, in part, because the context in which traits occur can lead to complex interactions with the diversification process (see \citet{caetano_2018, herrera_2018}). 
+This is, in part, because the context in which traits occur can lead to complex interactions with the diversification process \citep[see][]{caetano_2018, herrera_2018}.
 Trait combinations provide examples of such context, and may be of particular interest, given the rapid accumulation of information about traits. %cite?
@@
 The prevalence of ploidy variation has been broadly considered as a salient feature of flowering plants for nearly a century \citep{stebbins1938}. 
-Polyploids have long been thought to have lower (and even negative) rates of net diversification than diploids \citet{mayrose_2011, mayrose_2015}. 
+Polyploids have long been thought to have lower (and even negative) rates of net diversification than diploids \citep{mayrose_2011, mayrose_2015}. 
 A number of recent studies, however, find common paleo-polyploidization, including at the base of many highly diverse angiosperm clades \citep{soltis_2014}. 
@@
 The evidence for correlated shifts in mating system, from outcrossing to selfing, after polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
-In other cases, however, polyploidy is not only suspected to correlate with breeding system but causes the transition from self-incompatibility (SI) to self-compatibility(SC) \citep{stout1942, lewis1947}.
+In other cases, however, polyploidy is not only suspected to correlate with breeding system but causes the transition from self-incompatibility (SI) to self-compatibility (SC) \citep{stout1942, lewis1947}.
 Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic SI systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 

### 2019-03-09 manuscript/introduction.tex

--- 2019-03-08 manuscript/introduction.tex
@@
 
-%B->%E: instead of a major reorganization, I added a breeding system quote from Barrett that presages our results. How's that? It seemed daunting to reorganize ploidy-first to BS-first.
+%B I got a strange conflict here (first few paragraphs) with a seemingly long-ago committed repo version of introduction? The excised text is pasted at the bottom, and commented out, just in case.
 
-Species accumulate within the tree of life at different rates.
-A possible explanation for this phenomenon is that species possess various traits or character states that may differentially affect rates of diversification. 
-Dramatic increases in available phylogenetic and phenotypic data as well as methodological advances have greatly accelerated the search for traits that influence diversification traits.
+Species accumulate across the tree of life at different rates. 
+A possible explanation for this phenomenon is that species possess various traits or character states that differentially affect rates of diversification. 
+Dramatic increases in the available phylogenetic and phenotypic data, as well as methodological advances, have greatly accelerated the search for traits that influence diversification traits.
 Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a difficult statistical problem (e.g., \citealt{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}).
@@
 In part, the difficulties arise because speciation and extinction may not depend on a single trait, and the context in which traits occur can lead to complex interactions resulting in a heterogenous speciation and extinction rates \citep{beaulieu_2016, caetano_2018, herrera_2018}.
-Consequently, examining the association of only one trait with diversification patterns can be misleading. %B rm: "when multiple complex interactions are changing" <- explained above
+Consequently, examining the association of only one trait with diversification patterns can be misleading.
 
@@
 Under this scenario, polyploidy would no longer be considered to reliably slow down diversification, but instead, polyploidy would be the initiator of the diversification process.
-%R- This last phrase is from Itay and I rather like it.
-<-- %B: Hmmmm.
-How it would be considered an "initiator of the diversification process?" I mean, if it is associated with lower diversification.
-I think this should read "...polyploidy is no longer be considered an evolutionary dead end." In other words, I also agree with both %R and Itay, but not with what the current phrasing is asserting.
+%B: I think this should read "...polyploidy is no longer be considered an evolutionary dead end."  %R- This last phrase is from Itay and I rather like it.
+<-- %B: Hmm.
+But how would it be considered an "initiator of the diversification process?" It seems to be associated with lower diversification.
+In other words, I also agree with both %R and Itay, but not with what the current phrasing is asserting.
 
 Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and they crucially affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
-In particular, self-incompatibility systems cause a plant to reject its own pollen, and their loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974,igic_2008}.
-% Gordon comments here: Really?
+In particular, self-incompatibility causes plants to reject their own pollen, and its loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974,igic_2008}.
+% Gordon: Really?
 I find this hard to believe...
-I think of SI as only occurring in a limited number of angiosperm families.
-%B: Ha.
-I won't take it personally!
-True, it is present in a "limited" number of families, that number being about 150, including all of the largest families except for Lamiaceae (Apiaceae, Gesneriaceae, and Cucurbitaceae are the next in size).
-So, I am not sure how to show that it is "the most" common, but even this paper produces data to indicate that it is at the very least ca.
-twice as common as WGD, which probably qualifies it as "one of the most replicated transitions." %B: Also, "The evolutionary pathway from obligate outcrossing based upon self-incompatibility to predominant self-fertilization has probably been followed by more different lines of evolution in flowering plants than has any other." (Stebbins 1974)
-Previous analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, even when allowing the possibility that hidden states could drive diversification patterns \citep{freyman_2018}.
-However, studying breeding system with other correlated traits and unobserved states driving diversification patterns simultaneously has not yet being explored.
+%B:  sigh.
+Previous analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, even when allowing the possibility that unobserved states drive diversification patterns \citep{freyman_2018}.
+Despite their apparent macroevolutionary role in diversification, breeding systems are clearly not the sole trait determining diversification rates.
+Ploidy, for example, is widely thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
+%B: I did not find this confusing, but it worries me that Gordon did. Consequently, I've ejected this sentence, now focusing on SI/SC only.
+%The evidence for correlated shifts in mating system, from outcrossing to selfing, following WGD, appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
+Moreover, polyploidy is not only a mere correlate of breeding system, but may directly cause the loss of self-incompatibility (SI) \citep{stout1942, lewis1947}, and facilitate the transition to selfing.
+
+Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic SI systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
+This precludes the existence of SI polyploids in the most widespread system of SI, and creates a correlation between polyploidy and SC.
+% Gordon comments: Here again is confused about breeding systems in general vs. SI genetic systems.
+%B: not confused here; some SI is broken down by ploidy. Therefore, there is a correlation with coefficient between 0 and 1, and not 1.0 
 % Gordon comments here: First, it's a little confusing if you are talking about more SC or SI species more generally, or if you are talking specifically about genetic SI systems.
@@
 If your'e talking more generally about breeding system evolution there's a vast literature here, and I think  this is oversimplifying.
-For example, there are a lot of papers trying to link dioecy/hermaphroditism with diversification.
-(Vamosi papers on dioecy and diversification)
 % R- Boris what do you think is the best way to proceed here?
 Narrow it down to SI systems (deleting de Vos paper) or expanding it to talk about breeding systems in general and then saying that we will only focus on SI systems because of Solanaceae
-
-Ploidy may also affect the propensity for self-fertilization \citep{stebbins1950}. 
-The evidence for correlated shifts in mating system, from outcrossing to selfing, following WGD, appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
-In other cases, however, polyploidy is not only suspected to correlate with breeding system but causes the transition from self-incompatibility (SI) to self-compatibility(SC) \citep{stout1942, lewis1947}.
-Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic SI systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
-This creates a correlation between polyploidy and SC by precluding the existence of SI polyploids.
-In clades with these systems, it is thus natural to consider the simultaneous macroevolution of polyploidy and breeding system.
-% Gordon comments: Here again is confused about breeding systems in general vs. SI genetic systems
+%B: The apparent problem for Gordon is that devos2014 estimates *SSE parameters for heterostyly and homostyly, and not homomorphic SI.
+That's not a concern, I think, they are all "SI systems." And Dioecy etc.
+is not classically considered a "breeding system", but a "sexual system".
+That may be a part of the confusion.
+It's really not Gordon's problem, and we could do something to yet again re-define these terms?
+Emma?
 
 Given that changes in ploidy and breeding systems may be causally related and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
-This includes their joint influence on lineage diversification, and also potential patterns in the order of their transitions.
-For example, it is possible to ask whether losses of  SI more commonly occurred tied to polyploidization, or without a ploidy shift.
-Or whether polyploids have arised more commonly from self-incompatible than from self-compatible diploids/
-\Citet{robertson_2011} found that the pathway from self-incompatible diploids to self-compatible polyploids is dominated by loss of self-incompatibility followed later by polyploidization over long timescales, but proceeds in one step via polyploidization of SI species over short timescales.
-Greatly improved phylogeny and methods that allow for diversification rate differences allow far more powerful approach to such problems.
+First, it may be of particular interest to establish their joint influence on lineage diversification.
+It is possible that combinations of breeding system and ploidy do not result in predictable additive effects on speciation and extinction rates.
+Second, the two traits may affect each other's state transitions, including their magnitude and ordering of transitions.
+For example, losses of SI in diploids could be commonly caused by polyploidization, resulting in a one-step pathway to polyploid SC species. 
+Alternatively, the path from SI diploids to SC polyploids may involve a transition to SC, without in increase in ploidy, and then a subsequent diploidization. 
+%Or whether polyploids have arised more commonly from self-incompatible than from self-compatible diploids
+\citet{robertson_2011} found that the pathway from SI diploids to SC polyploids is dominated by the one-step pathway over long timescales, but generally proceeds in one-step via polyploidization of SI species over short timescales.
+A vast accumulation of data and methods that allow for diversification rate differences now allow far more powerful approach to such problems.
 % E: Technically, we allow for div effects in the rate estimates, but not in the pathways.
@@
 Maybe when we have more time for the revision?
+<- FIXME?
 
-Two methodological developments are critical for disentangling the complex interactions between two or more traits and their link to diversification.
-First, the model proposed by \citet{fitzjohn_2012} that extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multiple states, MuSSE, instead of binary states, BiSSE).
+Two recent methodological developments are critical for disentangling the complex interactions between two or more traits and their link to diversification.
+%First, the model proposed by \citet{fitzjohn_2012} that extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multiple states, MuSSE, instead of binary states, BiSSE).
+First, \citet{fitzjohn_2012} extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multi-state, MuSSE, instead of binary state, BiSSE, speciation and extinction models).
 This allows multiple states to represent trait combinations, particularly useful if one is interested in explicitly testing whether combinations of traits non-additively affect speciation or extinction.
-Second, the model by \citet{beaulieu_2016} HiSSE proposed a first solution to the problem of  elevated false positive type I errors found often in the BiSSE framework \citep{goldberg_2012}.
-The authors argued that it is unlikely that the variation in speciation and extinction rates is only due to a focal trait but by considering heterogeneity in the diversification process as the null hypothesis, it is possible to parse out the importance of the focal trait in diversification while other unobserved processes can happen.
-HiSSE models have effectively shown that some traits associated with diversification in BiSSE analyses are actually not associated with the diversification process, and instead a hidden state is linked to differences in the diversification rate.
-Such results raise the question of the identity of the hidden state, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
-Third, one can simultaneously consider the effects of multiple known traits using MuSSE, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
-However, MuSSE might also suffer from hight type I errors if there are other processes shaping diversification other than the traits observed \citep{caetano_2018}
-Therefore, a fourth option exists: considering both the presence of multiple known traits  linked to the diversification process, but also, some unknown or hidden state changing speciation and extinction rates.
-These four modeling strategies are proposed here in the context of studying complex interactions among traits, linked diversification.
+%Second, the model by \citet{beaulieu_2016} HiSSE proposed a first solution to the problem of  elevated false positive type I errors found often in the BiSSE framework \citep{goldberg_2012}. 
+Second, \citet{beaulieu_2016}  proposed an extension of the *SSE framework by introducing a model with `hidden' or unspecified traits (HiSSE), which can affect the diversification process.
+%The authors argued that it is unlikely that the variation in speciation and extinction rates is only due to a focal trait but by considering heterogeneity in the diversification process as the null hypothesis, it is possible to parse out the importance of the focal trait in diversification while other unobserved processes can happen.
+Because variation in speciation and extinction rates may not be due solely to the focal trait, HiSSE simultaneously enables such variation to be explained by unobserved states, and reduces the elevated false-positive error rates \citep{rabosky_2015}.
+%HiSSE models have effectively shown that some traits associated with diversification in BiSSE analyses are actually not associated with the diversification process, and instead a hidden state is linked to differences in the diversification rate.
+%Such results raise the question of the identity of the hidden state, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
+%Third, one can simultaneously consider the effects of multiple known traits using MuSSE, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
+%However, MuSSE might also suffer from hight type I errors if there are other processes shaping diversification other than the traits observed \citep{caetano_2018}
+%Therefore, a fourth option exists: considering both the presence of multiple known traits  linked to the diversification process, but also, some unknown or hidden state changing speciation and extinction rates.
+A joint application of these advances allows simultaneous consideration of the effects of multiple known traits on the diversification process, along with the presence of other unobserved states that can affect speciation and extinction rates.
+%These four modeling strategies are proposed here in the context of studying complex interactions among traits, linked diversification.
 
-The focus of our study is the individual role of ploidy and breeding systems, as well as their interaction, on the diversification process in Solanaceae. %sans pathway 
-Using ploidy and breeding system data from species in Solanaceae, and considering possible heterogeneity in the diversification process, we assess  the influence of these traits on speciation and extinction rates.
+Here, we examine the roles of ploidy and breeding system, as well as their interaction, on the diversification process in Solanaceae.
+Using extensive trait data, and considering other possible sources of heterogeneity in the diversification process, we assess the influence of these traits on speciation and extinction rates.
 First, we present models in which the effect of ploidy and breeding system are considered separately, and compare the inferences to previously published results. 
-Second, we consider the addition of a hidden trait to each of the models that examine traits individually, in order to investigate whether the focal trait directly affects differences in diversification, or if such differences are attributable to unobserved but correlated hidden states.
-Third, we ask whether the complex interaction between breeding system and ploidy may be explained by the hidden trait states. 
-We evaluate a multi-state model to consider both traits linked to diversification, and then consider whether an additional unobserved trait may be driving differences in diversification.
-In all these models we also investigate the potential effects of diploidization to explore whether the reversion of polyploidy changes diversification interpretations.
-Our results highlight the importance of considering non-additive effects of traits on net diversification rates, especially when there are strong biologically driven correlations amongst them.
+Second, we consider the effects of addition of a hidden trait to each of the models that examine traits individually, and investigate whether the focal trait directly affects differences in diversification, or if such differences are attributable to unobserved but correlated states.
+Third, we also ask whether the complex interaction between breeding system and ploidy may be explained by the unobserved trait states. 
+Fourth, we evaluate a multi-state model, which simultaneously considers the effect of both traits on diversification, and then consider whether an additional unobserved trait may be driving differences in diversification.
+For all of the above models, we investigate the potential effects of diploidization. 
+Our results highlight the importance of considering non-additive effects of traits on net diversification rates, especially when there are strong biologically driven correlations among them.
 
@@
 %Additionally, a lack of methods that allow for simultaneous inference of effects of other traits, which also influence diversification, can lead to incorrect conclusions about the focal trait \citep{rabosky_2015, beaulieu_2016}.
+
+%>>>>>>> aa786e0237f5a090b69a64b415f317a49723f88c
+%Species accumulate around the tree of life at different rates.
+%A possible explanation for this phenomenon is that species possess traits that cause unequal rates of diversification. %B: new.
+%The search for influential traits has been accelerated by a dramatic increases in data and an advance in methods, which now make use of far more information from phylogenetic trees \citep{maddison_2007}.
+%Despite some setbacks \citep{maddison_2015, rabosky_2015, moore_2016}, advances in analytical methods \citep{fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017} are enabling the discovery of traits that alter the diversification patterns.
+%Nevertheless, the aim of identifying the focal trait and investigating its association with rates of speciation and extinction remains complicated. 
+%This is, in part, because the context in which traits occur can lead to complex interactions with the diversification process \citep[see][]{caetano_2018, herrera_2018}.
+%Trait combinations provide examples of such context, and may be of particular interest, given the rapid accumulation of information about traits. %cite?
+%Two exceedingly well-studied traits, ploidy and breeding system, are of particular interest. 
+%Their individual state-dependent contributions to the process of diversification, as well as influence on each others' trait transitions, have long been the focus of speculation and study \citep{stebbins1950}.
+%Whole genome duplication is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
+%Such changes is ploidy have the potential to affect many other phenotypes, and therefore alter a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
+%The prevalence of ploidy variation has been broadly considered as a salient feature of flowering plants for nearly a century \citep{stebbins1938}. 
+%Polyploids have long been thought to have lower (and even negative) rates of net diversification than diploids \citep{mayrose_2011, mayrose_2015}. 
+%A number of recent studies, however, find common paleo-polyploidization, including at the base of many highly diverse angiosperm clades \citep{soltis_2014}. 
+%\citet{landis_2018}, for example, found 106 whole genome duplications within angiosperms that, combined with a state-dependent diversification model, resulted in approximately 60\% of whole genome duplication events potentially enhancing net diversification.
+%The mere occurrence of such ancestral whole genome duplications points to an possibly important role of genome downsizing and diploidization in the diversification process \citep{soltis_2015diploidization, dodsworth_2015}.
+%The presence of a significant and positive net diversification rate for (some) diploids can be  the result of a polyploid lineage that has been diploidized. 
+%Under this scenario, rather than being a macroevolutionary dead-end, polyploidy may instead be positively associated with the diversification process.
+
+%Ploidy has long been thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
+%The evidence for correlated shifts in mating system, from outcrossing to selfing, after polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
+%In other cases, however, polyploidy is not only suspected to correlate with breeding system but causes the transition from self-incompatibility (SI) to self-compatibility (SC) \citep{stout1942, lewis1947}.

### 2019-03-09 manuscript/methods.tex
pathways using proportion instead of ratio

@@
 
-\subsection{Data}
+\subsection{Ploidy and breeding system data}
 
@@
 For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus or based on SI/SC classification.
-For example, if a species was incompatible without any chromosomal information it was assigned as diploid.
-This was the case of  \textit{Solanum betaceum} that did not include information about ploidy level but it has 24 chromosomes, since $x=12$ is the base chromosome number of the \textit{Solanum} genus \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid.
+For example, \textit{Solanum betaceum} did not have information about ploidy level but it has 24 chromosomes, so since $x=12$ is the base chromosome number of the \textit{Solanum} genus \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid.
+Additionally, because of the absence of SI polyploids (discussed further below), species known to be SI could be scored as diploid.
 Species with more than one ploidy level were assigned the most frequent ploidy level recorded or the smallest ploidy in case of frequency ties.
-% E: There was more here about cross-verifying that no P species were SI.  I removed it because there were conflicts before, but we resolved them.  This is addressed in the next paragraph.
-%
+
 Breeding system was scored as self-incompatible (I) or self-compatible (C) based on results curated from the literature and  original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
-Many species could unambiguously be coded as either I or C \citep{raduski_2012}.  But for diploids without breeding system information taxa was coded as both I or C. %B: most? Any exceptions? R- many 
-Following previous work, we coded as I any species with functional I systems, even if C or dioecy was also reported.
-Dioecious species without functional I were coded as C.
+Many species could unambiguously be coded as either I or C \citep{raduski_2012}.
+Following previous work, we coded as I any species with a functional SI system, even if SC or dioecy was also reported.
+Dioecious species without functional SI were coded as C.
 % E: I am trying to stick with I/C when talking about the state values, but using SI/SC when talking about the meaning of the trait (because this what people use normally). 
 %R- Okay- I was getting confused as well thanks for clarifying
-To those existing data sets, we added some additional records for chromosome number and breeding system.
-The Supplementary Information contains citations for the numerous sources for the added data. % todo: add number of refs when we have it
+
+The Supplementary Information contains citations for the numerous original data sources. % todo: say how many refs?
 Resolution of taxonomic synonymy followed the conventions provided in Solanaceae Source \citep{solsource}. 
@@
 Following the reasoning outlined in \citet{robertson_2011}, we examined closely the few species for which the merged ploidy and breeding system data indicated the presence of self-incompatible polyploids.
-Although I populations frequently contain some C individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find a convincing case of a naturally occurring I and polyploid population.
-%B: Thus qualified, as above, I think this should be inserted.
-The single instance of an I and polyploid individual appears to be an allopentaploid hybrid of {\em Solanum oplocense} Hawkes x {\em Solanum gourlayii} Hawkes, reported by \citealt{camadro_1981}.
-Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain I, so long as they express a single allele at the S-locus (discussed in \citealt{robertson_2011}).
-Because of the resulting absence of I and polyploid populations, as well as the linked functional explanation for disabling of gametophytic self-incompatibility systems with non-self recognition, following whole genome duplication \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).
+Although SI populations frequently contain some SC individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find a convincing case of a naturally occurring SI and polyploid population.
+The single instance of an SI polyploid individual appears to be an allopentaploid hybrid of \textit{Solanum oplocense} Hawkes x \textit{Solanum gourlayii} Hawkes, reported by \citet{camadro_1981}.
+Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain SI, so long as they express a single allele at the S-locus \citep[discussed in][]{robertson_2011}.
+Because of the resulting absence of SI and polyploid populations, as well as the linked functional explanation for disabling of gametophytic self-incompatibility systems with non-self recognition, following whole genome duplication \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).
 %B: I added the non-self recognition bit, because in Prunus, heteroallelic pollen may not cause SC.
-% E: Should we include more details about how many polymorphic species and how the particularly tricky ones were dealt with? %B: No and yes, maybe.
 
@@
 
-% NOTE: Should subtly rephrase so that the default models lack diploidization.
-
 \subsection{Models for ploidy and diversification}
 
-In order to investigate the association between ploidy level and diversification, we first defined a binary state speciation and extinction model (BiSSE, \citealt{maddison_2007}) in which taxa were classified as diploid (D) or polyploid (P) (\cref{figure:stateclassifications}).
-We call this the \textit{D/P ploidy} model. This model is identical to the model proposed by \citet{mayrose_2011, mayrose_2015}, but inferences were performed in a Bayesian framework.
- %B: emphasis on introduced term (quotes awkward)
-We calculated the posterior probability distributions of speciation rates ($\lambda_D$, $\lambda_P$), extinction rates ($\mu_D$, $\mu_P$), net diversification rates ($r_D=\lambda_D-\mu_D$, $r_P=\lambda_P-\mu_P$), and relative extinction rates ($\nu_D=\mu_D / \lambda_D$, $\nu_D=\mu_D / \lambda_D$) associated with each state.
-Parameter $\rho$ represents the transition rate from $D$ to $P$.
-% E: presumably the motivation will be explained in the Intro
-%B: should the term above be "re-diploidization" or "diploidization" (as now)?
-% B (via E): need some explanation of what \delta means, biologically, when ploidy is scored like this
-% R- the explanation will happen in the discussion
+In order to investigate the association between ploidy level and diversification, we first defined a binary state speciation and extinction model (BiSSE, \citealt{maddison_2007}) in which taxa were classified as diploid (D) or polyploid (P).
+We call this the D/P ploidy model.
+The model parameters are speciation rates ($\lambda_D$, $\lambda_P$) and extinction rates ($\mu_D$, $\mu_P$) for each state, and a rate of transition from D to P ($\rho$).
 
-Our second model addressed the diversification due to ploidy differences while also parsing out the heterogeneity of diversification rates due a possible unobserved trait.
+% E: I moved r and nu to later, under statistical inference.  It made it seem like there were too many parameters here.
+% I also removed italics around the model names because we don't actually use those names later on (more just the state codes).
+
+Our second model addressed diversification due to ploidy differences while also parsing out the heterogeneity of diversification rates due a possible unobserved trait.
 Diversification rate differences explained by some trait other than ploidy were accommodated by adding a hidden state (HiSSE model; \citealt{beaulieu_2016}).
@@
 In this model, each of the observed diploid and polyploid states were subdivided by a binary hidden trait with states $A$ and $B$.
-This model was called the \textit{D/P+A/B ploidy and hidden state} model. 
-Posterior probability distributions of speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$), extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$), net diversification rates ($r_{D_A},\ r_{D_B},\ r_{P_A},\ r_{P_B}$), and relative extinction rates ($\nu_{D_A},\ \nu_{D_B},\ \nu_{P_A},\ \nu_{P_B}$) were estimated.
-In this model polyploidization rate  is again represented by $\rho$  and hidden state rates are symmetrical with parameter  $\alpha$. %B: same dip vs. re-dip question, as above.
+This model was called the D/P+A/B ploidy and hidden state model. 
+Model parameters include speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$) and extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$) for each subdivided state.
+The polyploidization rate is again represented by $\rho$, and the rates of transition between hidden states are symmetrical with parameter $\alpha$.
 
@@
 
-To assess the effects of breeding system in the diversification process, we first fit model in which the states are self-incompatible (I) or self-compatible (C).
+To assess the effects of breeding system in the diversification process, we first fit a model in which the states are self-incompatible (I) or self-compatible (C).
 This is the same as the analysis of \citet{goldberg_2010}, but with an updated phylogeny \citep{sarkinen_2013}.
-We call this BiSSE model the \textit{I/C breeding system} model. 
+We call this BiSSE model the I/C breeding system model. 
+It has parameters for state-specific speciation ($\lambda_I,\ \lambda_C$) and extinction ($\mu_I,\ \mu_C$), and for transitions from I to C ($q_{IC}$).
+
 To parse out the effect of breeding system on diversification, while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
-We call this HiSSE model the \textit{I/C+A/B breeding system and hidden state model}.
-The I/C+A/B model is equivalent to the model \citet{freyman_2017} used to study diversification in Onagraceae, but we restricted the hidden state transition rates to be symmetrical.
+We call this HiSSE model the I/C+A/B breeding system and hidden state model.
+It also has parameters for state-specific speciation ($\lambda_{I_A},\ \lambda_{I_B},\ \lambda_{C_A},\ \lambda_{C_B}$) and extinction ($\mu_{I_A},\ \mu_{I_B},\ \mu_{C_A},\ \mu_{C_B}$), for loss of SI ($q_{IC}$), and for transitions between hidden states ($\alpha$).
+An I/C+A/B model was used by \citet{freyman_2017} used to study diversification in Onagraceae, but here we restricted the hidden state transitions to be symmetrical.
 
-%B: we moved back from "SI" to "self-incompatibility" below. Maybe a no-no.
-Self-incompatibility is homologous in all species in which S-alleles are cloned, and controlled crosses are performed. % R- Boris this phrase reads weird but I don't know how to fix it  
-All species sampled to date, possess a non-self recognition, RNase-based, gametophytic self-incompatibility \citep[shared even with other euasterid families;][]{ramanauskas_2017}.
-Furthermore, species that are distantly related within this family carry closely-related alleles, with deep trans-specific polymorphism, at the S-locus, which controls the SI response \citep{ioerger_1990, igic_2006}.
-%B: may need to introduce "S-locus" earlier.
-This represents very strong evidence that the self-incompatible mechanism, and our $I$ state is ancestral to the Solanaceae, and did not arise independently within the family.Therefore, for all breeding system models, we estimated a transition rate from $I$ to $C$  but not the reverse ($q_{CI}=0$).
- %B: I think I have this right <- someone please double-check.
+Self-incompatibility is homologous in all species in which S-alleles have been cloned and controlled crosses performed.
+All species sampled to date possess a non-self recognition, RNase-based gametophytic self-incompatibility \citep[shared even with other euasterid families;][]{ramanauskas_2017}.
+Furthermore, species that are distantly related within this family carry closely-related alleles, with deep trans-specific polymorphism at the locus that controls the SI response \citep{ioerger_1990, igic_2006}.
+This represents very strong evidence that the $I$ state is ancestral in Solanaceae, and that the SI mechanism was not regained within this family.
+Therefore, for all breeding system models, we estimated a transition rate from $I$ to $C$  but not the reverse ($q_{CI}=0$).
 % This is an important assumption because ancestral state reconstruction in the models might differ when polyploidy and breeding system are analyzed as independent in the Solanaceae tree.
-% This irreversibility assumption prevents self-compatible state from evolving into self-incompatible state and as a result, the ancestral state of all taxa in this model has probability one of being self-incompatible.
 %
@@
 
-Polyploidy and breeding system might have changed lineage diversification individually but these two traits also have complex interactions in Solanaceae species.
-Therefore, a multi-state model that investigates the contribution of both traits and the possible transitions between them was proposed.
-%B: unsure about edit, which was: "If both ploidy and breeding system have the potential to influence lineage diversification, we should logically consider their effects jointly."  E: okay; also try to add that their transitions interact
-% R- Is the above better?
-Self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids, which are always self-compatible (CP) were used as input for a MuSSE model \citep{fitzjohn_2012}.
+Polyploidy and breeding system might influence lineage diversification individually, but these two traits also have complex interactions in Solanaceae species.
+Therefore, we considered a multi-state model that investigates the contribution of both traits and the possible transitions between them.
+Self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids, which are always self-compatible (CP) were used in a MuSSE model \citep{fitzjohn_2012}.
 We did not include a state for self-incompatible polyploids because they are not observed in the data, and  that trait combination state is mechanistically predicted not to occur.
-The proposed model is called the \textit{ID/CD/CP ploidy and breeding system} and has six parameters for diversification in each state: ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction), and three for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
+We call this the ID/CD/CP ploidy and breeding system model.
+It has six parameters for diversification in each state ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction), and three for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
 The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
-Diploidization from $CP$ to $ID$ is not allowed because it would represent a simultaneous regain of SI.
+% Diploidization from $CP$ to $ID$ is not allowed because it would represent a simultaneous regain of SI. % E: moved into the diploidization section
 
-The ID/CD/CP model could also suffer from a hight type I error as the regular two state BiSSE model. 
-A hidden factor might alter diversification beyond both ploidy and breeding system misleading inferences \citep{rabosky_2015}.
+The ID/CD/CP model could also suffer from a high type I error, just like a two state BiSSE model. 
 We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018,  huang_2018}.
-We refer to this as the \textit{ID/CD/CP+A/B} model.
+We refer to this as the ID/CD/CP+A/B model.
 A fully parameterized version of this model would have 26 rate parameters. 
-Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fitted a simplified version with 15 parameters.
-The reduction in parameter space is achieved by fixing the rates for transitions among hidden states to be equal with rate $\alpha$, and fixing the transition rates between observed states to be independent of the hidden state (rates $\rho_I,\ \rho_C,\ \delta,\ q_{IC}$ as defined for the ID/CD/CP model).
+Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fit a simplified version with 15 parameters.
+The reduction in parameter space was achieved by fixing the rates for transitions among hidden states to be equal with rate $\alpha$, and fixing the transition rates between observed states to be independent of the hidden state (rates $\rho_I,\ \rho_C,\ \delta,\ q_{IC}$ as defined for the ID/CD/CP model).
 There are additionally twelve diversification rate parameters ($\lambda_{ID_A},\ \lambda_{ID_B},\ \lambda_{CD_A},\ \lambda_{CD_B},\ \lambda_{CP_A},\ \lambda_{CP_B},\ \mu_{ID_A},\ \mu_{ID_B},\ \mu_{CD_A},\ \mu_{CD_B},\ \mu_{CP_A},\ \mu_{CP_B}$).
@@
 We insert an artificial division in the CP state, so that one substate contains the CP species that arrived via the one-step pathway and the other substate contains the CP species that arrived via the two-step pathway.
-We consider only unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and not strong support for diploidization.
+We consider only unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and not strong support for diploidization (see below).
 First, we consider only the rates of transitions between these states, placing them in the propogation matrix \myvec{Q}.
@@
 
-% E todo: for the case with net div, write out the propogation matrix and equations for pathway contributions
+% "one-step predictions for P_I and P_C over time"
+% E & R todo later: for the case with net div, write out the propogation matrix and equations for pathway contributions
 
@@
 
-For all four models that consider ploidy changes, we allowed diploidization by adding a parameter $\delta$ connecting polyploidy state back to diploid or to self-compatible diploid in the mult-state models (denoted by $+/delta$ in the models of  \cref{table:marginallike}).
+For the models with ploidy changes, we allowed diploidization by adding a parameter $\delta$ for transitions from P back to D (in the D/P and D/P+A/B models) or from CP back to CD (in the I/C and I/C+A/B models).
+We did not allow transitions from CP back to ID because that would include regain of SI.
+The models are denoted with $+\delta$ in \cref{table:marginallike}.
 Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010, glick2014, mayrose_2015, freyman_2017}.
 These types of classifications do not allow for a ploidy reversion.
-Where indicated, the classification of ploidy for the data used in our models was based on chromosome multiplicity at the genus level (data available at https://www.github.com/roszenil/solploidy/datasets).
-% ref data/sources table?
-%github
+Where indicated, the classification of ploidy for the data used in our models was based on chromosome multiplicity at the genus level (see Supplementary Information).
+% (data available at https://www.github.com/roszenil/solploidy/datasets).
 However, the majority of the ploidy classifications were adopted from original studies with alternative sources of information (\eg geographic distribution, genus ploidy distribution) where ploidy was defined by authors that found evidence for it.
-Since it is not clear whether diploidization can be detected under alternative ploidy classifications or even classifications based on chromosome number multiplicity at the genus level, we also fit the models without diploidization in order to test  whether the conclusions about diversification are sensitive to including diploidization.
-As discussed by \citet{servedio_2014}, the presence or absence of a hypothesis can have an exploratory goal.
-In our case the diploidization parameter (or its absence, $\delta=0$) in our models is an opportunity to explore an assumption that might be important but that is not the single definitive process to understand the interactions among polyploidy, breeding system, and diversification.
+Since it is not clear whether diploidization can be detected under alternative ploidy classifications or even classifications based on chromosome number multiplicity at the genus level, we fit models with and without diploidization in order to test  whether the conclusions about diversification are sensitive to including diploidization.
+As discussed by \citet{servedio_2014}, the presence or absence of a process can have an exploratory goal.
+In our case the diploidization parameter (or its absence, $\delta=0$) in our models is an opportunity to explore an assumption that might be important but that is not the most essential process to understand the interactions among polyploidy, breeding system, and diversification.
 
@@
 
-Parameters for each of the 10 diversification models were estimated using RevBayes \citep{hoehna_2016}.
+Parameters for each of the ten models described were estimated using RevBayes \citep{hoehna_2016}.
 Code for analyses and key results are available at \url{https://github.com/roszenil/solploidy}.
-We accounted for incomplete sampling in all analyses by fixing the probability of sampling a species at the present $s$ to $595/3000$ since the Solanaceae family has approximately 3,000 species as estimated by the Solanaceae Source project \citep{solsource}.
-For all 10 models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
+We used the method of \citet{fitzjohn_2009} to account for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $595/3000$ since the Solanaceae family has approximately 3,000 species as estimated by the Solanaceae Source project \citep{solsource}.
+For all models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
 Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
@@
 fixed it
-Convergence and mixing of each MCMC 
-was determined by ensuring the effective sample size of each parameter was over 200.
+Convergence and mixing of each MCMC chain was determined by ensuring the effective sample size of each parameter was over 200.
 
-For each model, ancestral states at each node in the phylogeny were sampled jointly during the MCMC analyses.
-Ancestral state reconstructions for all models can be found in the supplementary information and show the maximum a posteriori of the marginal probability distributions for each of the 594 internal nodes for each of the 10 models.
+We report posterior distributions for the model parameters, and also for the compound parameters of net diversification ($\lambda - \mu$) and extinction fraction ($\mu / \lambda$) for each state.
+Additionally, ancestral states at each node in the phylogeny were sampled jointly during the MCMC analyses.
+Ancestral state reconstructions for all models can be found in the Supplementary Information and show the maximum a posteriori of the marginal probability distributions for each of the 594 internal nodes for each of the 10 models.
+% TODO: add fig refs (or remove the mention of them)
 
@@
 
-We calculated the marginal likelihood for each of the 10 models in RevBayes \citep{hoehna_2016}.
+We calculated the marginal likelihood for each of the ten models in RevBayes \citep{hoehna_2016}.
 Marginal likelihoods were calculated using 50 stepping stone steps under the methodology of \citet{xie_2010}.
@@
 In mathematical terms, the D/P, I/C, and ID/CD/CP state spaces are not `lumpable' with respect to one another \citep{tarasov_2019}.
-%R: Emma this is where the figure can be important to show that we have to make different input
 % E: I think this short explanation is clear enough, especially with the figure.  I'm not sure whether the hierarchical concept is exactly applicable here.
@@
 There is `positive' support for $M_0$ when this value is more than 2, `strong' support when it is more than 6, and `very strong' support when it is more than 10 \citep{kass1995}.
-
-%\begin{table}
-%\begin{tabular}{@{}llccc@{}} \toprule
-%\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
-%Type & Total of & D/P and D/P-A/B& I/C and I/C-A/B & ID/P/CD and CD/P/ID-A/B\\ \midrule
-%Diploid Self Compatible & 152 & 0 &  0 & 0 \\
-%Diploid Self Incompatible& 97 & 0  & 1 & 2\\
-%Diploid with unknown breeding system & 219 & 0 & (0,1) & (0,2) \\
-%Polyploid & 81 & 1& 0 & 1 \\
-%Unknow ploidy and self compatible& 34 & (0,1)& 0 & (0,1) \\ 
-%Unknown ploidy and self incompatible & 12 & 0 & 1 & 2 \\ \bottomrule
-%\end{tabular}
-%\caption{Binary and three state classifications for 595 taxa with ploidy and/or breeding system data.
-The number of taxa in the sample was maximized by including tips with only ploidy or only breeding system and assigned them as uncertain in the unknown character.}
-%\label{table:stateclassifications}
-%\end{table}
 
@@
 The subsequent models require a bit of thought, which is not too bad, but the treatment of the state coding and specifically the sentence "In all cases, species were coded as either A or B in the hidden state models." is a bit terse for someone not familiar with HiSSE.
+% E: Yes, all I species are D and all P species are C (no SI polyploids).
+I don't think there are polymorphic species because species with some SI and some SC pops are called I, and the most common ploidy level was used.
 }
@@
 \end{figure}
-
-% R: I changed this models to go now in the new order
 
@@
 Model                     & Ploidy & Diploidization & Breeding  & Hidden & Num        & Marginal       \\
-                             & &  &System & State & Parameters & Log-Likelihood \\ \midrule
+                          &        &                & System    & State  & Parameters & Log-Likelihood \\ \midrule
 1. D/P                    & Yes    & No             & No        & No     & 5          & -1193.66\\
-2. D/P +$\delta$                      &	Yes  &	Yes &	No	&No	& 6	& -1182.93 \\
-3. D/P+A/B        &	Yes & No &	No &	Yes &	10	&-1150.99\\
-4. D/P+$\delta$+A/B                   &	Yes &	Yes	&No &	Yes &	11 &	\textbf{-1145.69}\\
-5. I/C                       &	No & No	&Yes &	No &	 5 &  -1194.80 \\
-6. I/C+A/B                   &	No &	 No	&Yes &	Yes	& 10 & \textbf{-1155.37}\\
-7. ID/CD/CP      &	Yes & 	No &	Yes	&No &	9 &-1345.87\\
-8. ID/CD/CP +$\delta$                 &	Yes & 	Yes &	Yes &	No &	10 & -1344.50\\
-9. ID/CD/CP+A/B  & Yes & 	No	&Yes &	Yes	&15 &-1303.55 \\ 
-10. ID/CD/CP+$\delta$+A/B              &	Yes 	&Yes &	Yes &	Yes &	16 & \textbf{-1300.35} \\
+2. D/P +$\delta$          & Yes    & Yes            & No        & No     & 6          & -1182.93 \\
+3. D/P+A/B                & Yes    & No             & No        & Yes    & 10         & -1150.99\\
+4. D/P+A/B+$\delta$       & Yes    & Yes            & No        & Yes    & 11         & \textbf{-1145.69}\\
+5. I/C                    & No     & No             & Yes       & No     & 5          & -1194.80 \\
+6. I/C+A/B                & No     & No             & Yes       & Yes    & 10         & \textbf{-1155.37}\\
+7. ID/CD/CP               & Yes    & No             & Yes       & No     & 9          & -1345.87\\
+8. ID/CD/CP+$\delta$      & Yes    & Yes            & Yes       & No     & 10         & -1344.50\\
+9. ID/CD/CP+A/B           & Yes    & No             & Yes       & Yes    & 15         & -1303.55 \\ 
+10. ID/CD/CP+A/B+$\delta$ & Yes    & Yes            & Yes       & Yes    & 16         & \textbf{-1300.35} \\
 \bottomrule
@@
             & 1 & 2 & 3 & 4 \\
-1. D/P & $\cdot$	 &10.72 &	-37.24	&-31.94\\
-2. D/P no $\delta$ &$\cdot$&$\cdot$ &	-47.97	&-42.66\\
-\textbf{3. D/P+A/B}  &$\cdot$  & $\cdot$&	$\cdot$	& 5.30 \\
-4. D/P+A/B no $\delta$ &$\cdot$& $\cdot$ & $\cdot$&$\cdot$ \\
+1. D/P$+\delta$ & $\cdot$     &10.72 &    -37.24    &-31.94\\
+2. D/P  &$\cdot$&$\cdot$ &    -47.97    &-42.66\\
+\textbf{3. D/P+A/B$+\delta$}  &$\cdot$  & $\cdot$&    $\cdot$    & 5.30 \\
+4. D/P+A/B  &$\cdot$& $\cdot$ & $\cdot$&$\cdot$ \\
 \end{tabular}
@@
 & 7 & 8 & 9 & 10\\
-7. ID/CD/CP & $\cdot$	&1.36 & -44.15&-40.95 \\
-8. ID/CD/CP no $\delta$ & $\cdot$ & $\cdot$ & -45.51 &-42.31 \\
-\textbf{9. ID/CD/CP+A/B}& $\cdot$ & $\cdot$ &$\cdot$	& 3.2 \\ 
-10. ID/P/CD no $\delta$-A/B & $\cdot$& $\cdot$&$\cdot$ & $\cdot$\\
+7. ID/CD/CP$+\delta$ & $\cdot$    &1.36 & -44.15&-40.95 \\
+8. ID/CD/CP & $\cdot$ & $\cdot$ & -45.51 &-42.31 \\
+\textbf{9. ID/CD/CP+A/B+$\delta$}& $\cdot$ & $\cdot$ &$\cdot$    & 3.2 \\ 
+10. ID/P/CD+A/B & $\cdot$& $\cdot$&$\cdot$ & $\cdot$\\
 \end{tabular}
@@
     Each of the three boxes contains models that can be compared with one another, based on the character states they include (see \cref{figure:stateclassifications}).
-    Models are numbered in as \cref{table:marginallike}.
+    Models are numbered in as \cref{table:marginallike}. % FIXME: they aren't, but it's a pain to change; change in prev table instead?
     Bayes factors are reported on the natural log scale, so numbers greater than $+2$ mean that the model in the row label has `positive' support relative to the model in the column label; numbers less than $-2$ mean that model in the column label is the preferred one.
@@
 
-%\begin{tabular}{@{}lcccccccccc@{}} \toprule
-%\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
-%Model& 1& 2 & 3 & 4  &5 & 6& 7 & 8 & 9 & 10 \\
-%Polyploidy & & & & & & & & & \\
-%1. D/P &	-  &	10.727 &	-37.243	&-31.941& -	& - & - & - & - &- \\
-%2. D/P no $\delta$ & -  & -  &	-47.97	&-42.668& -	& - & - & - & - &- \\
-%3. D/P- A/B  & -  & -  &	-	& 5.302 & -	& - & - & - & - &- \\
-%4. D/P no $\delta$-A/B &
-%Breeding System& & & & & & & & & \\
-%5. C/I & -  & -  &	-	& - & -	& -39.434 & - & - & - &- \\
-%6. I/C-A/B &
-%Polyploidy and Breeding System & & & & & & & & & \\
-%7. CD/P/ID & -  & -  &	-	& - & -	& - & - &1.368 & -44.151&-40.951 \\
-%8. CD/P/ID no $\delta$& -  & -  &	-	& - & -	& - & - &-& -45.519&-42.319 \\
-%9. CD/P/ID-A/B& -  & -  &	-	& - & -	& - & - &-& -& 3.2 \\ 
-%10.ID/P/CD no $\delta$-A/B &
-%\bottomrule
-%\end{tabular}
-
-% E: Could we include a figure of the traits on the tree?  At least as supp info. TODO
+% E: Could we include a figure of the traits on the tree?  At least as supp info. FIXME

### 2019-03-09 manuscript/introduction.tex
- restoring introduction after bad commit
read through Results

@@
 
-%B I got a strange conflict here (first few paragraphs) with a seemingly long-ago committed repo version of introduction? The excised text is pasted at the bottom, and commented out, just in case.
+%B->%E: instead of a major reorganization, I added a breeding system quote from Barrett that presages our results. How's that? It seemed daunting to reorganize ploidy-first to BS-first.
 
@@
 Moreover, polyploidy is not only a mere correlate of breeding system, but may directly cause the loss of self-incompatibility (SI) \citep{stout1942, lewis1947}, and facilitate the transition to selfing.
-
 Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic SI systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
@@
 %Or whether polyploids have arised more commonly from self-incompatible than from self-compatible diploids
-\citet{robertson_2011} found that the pathway from SI diploids to SC polyploids is dominated by the one-step pathway over long timescales, but generally proceeds in one-step via polyploidization of SI species over short timescales.
+\citet{robertson_2011} found that the pathway from SI diploids to SC polyploids is dominated by the two-step pathway over long timescales, but generally proceeds in one step via polyploidization of SI species over short timescales.
 A vast accumulation of data and methods that allow for diversification rate differences now allow far more powerful approach to such problems.
-% E: Technically, we allow for div effects in the rate estimates, but not in the pathways.
-If SC-D have so much extinction that the step to SC-P never happens, the old pathway methods won't see that.
-I'll think about if this is easy to fix.
-% And, are diversification rates of polyploid SC lineages different if one or other path was taken?
-\citet{charlesworth1985} % E: This is a great question.
-We don't answer it with this round of analyses.
-Maybe when we have more time for the revision?
-<- FIXME?
 
@@
 %Additionally, a lack of methods that allow for simultaneous inference of effects of other traits, which also influence diversification, can lead to incorrect conclusions about the focal trait \citep{rabosky_2015, beaulieu_2016}.
-
-%>>>>>>> aa786e0237f5a090b69a64b415f317a49723f88c
-%Species accumulate around the tree of life at different rates.
-%A possible explanation for this phenomenon is that species possess traits that cause unequal rates of diversification. %B: new.
-%The search for influential traits has been accelerated by a dramatic increases in data and an advance in methods, which now make use of far more information from phylogenetic trees \citep{maddison_2007}.
-%Despite some setbacks \citep{maddison_2015, rabosky_2015, moore_2016}, advances in analytical methods \citep{fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017} are enabling the discovery of traits that alter the diversification patterns.
-%Nevertheless, the aim of identifying the focal trait and investigating its association with rates of speciation and extinction remains complicated. 
-%This is, in part, because the context in which traits occur can lead to complex interactions with the diversification process \citep[see][]{caetano_2018, herrera_2018}.
-%Trait combinations provide examples of such context, and may be of particular interest, given the rapid accumulation of information about traits. %cite?
-%Two exceedingly well-studied traits, ploidy and breeding system, are of particular interest. 
-%Their individual state-dependent contributions to the process of diversification, as well as influence on each others' trait transitions, have long been the focus of speculation and study \citep{stebbins1950}.
-%Whole genome duplication is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
-%Such changes is ploidy have the potential to affect many other phenotypes, and therefore alter a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
-%The prevalence of ploidy variation has been broadly considered as a salient feature of flowering plants for nearly a century \citep{stebbins1938}. 
-%Polyploids have long been thought to have lower (and even negative) rates of net diversification than diploids \citep{mayrose_2011, mayrose_2015}. 
-%A number of recent studies, however, find common paleo-polyploidization, including at the base of many highly diverse angiosperm clades \citep{soltis_2014}. 
-%\citet{landis_2018}, for example, found 106 whole genome duplications within angiosperms that, combined with a state-dependent diversification model, resulted in approximately 60\% of whole genome duplication events potentially enhancing net diversification.
-%The mere occurrence of such ancestral whole genome duplications points to an possibly important role of genome downsizing and diploidization in the diversification process \citep{soltis_2015diploidization, dodsworth_2015}.
-%The presence of a significant and positive net diversification rate for (some) diploids can be  the result of a polyploid lineage that has been diploidized. 
-%Under this scenario, rather than being a macroevolutionary dead-end, polyploidy may instead be positively associated with the diversification process.
-
-%Ploidy has long been thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
-%The evidence for correlated shifts in mating system, from outcrossing to selfing, after polyploidization appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
-%In other cases, however, polyploidy is not only suspected to correlate with breeding system but causes the transition from self-incompatibility (SI) to self-compatibility (SC) \citep{stout1942, lewis1947}.

### 2019-03-10 manuscript/results.tex

@@
 
-% NOTE: will need to update figure/panel refs if making no-delta models primary
-% R: I am changing no delta as just D/P and prevous D/P as +delta
-
-% E: I brutally shortened the text here and hence put all the models into one section for the "diversification" question.  But I think I retained all the ideas that were written out before.
-% Much of what I removed was descriptions about which distributions are/aren't overlapping with one another.
-I think that's easier for the reader to see in the figures than to read in the text.
-But now the text says things like if some estimates are "different", so perhaps in Methods we should explain different = non-overlapping?
-% For the former model comparison section, I moved the key conclusions into the paragraphs here and the mechanical details (bayes factor definition) into Methods.
-
-Individual trait-dependent diversification inference, of ploidy and breeding system considered separately, each showed significant associations with net diversification differences. 
+When considered separately, ploidy and breeding system each showed significant associations with net diversification differences in Solanaceae.
 When considered together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well.
@@
 Statistical model comparisons show very strong support for including the hidden state and strong support for including diploidization (\cref{table:bayesfactors}).
-% E: I removed the description of relative extinction results.
-We don't discuss it for other models or show it in the main text, and I never really know what to make of that compound parameter.
-But it's good that it's in the supp info.
 
 When considering breeding system alone (I/C model, \cref{figure:netdivall}C), we found a larger net diversification rate for SI than for SC species, in agreement with \citet{goldberg_2010}.
-%B: moved \cref{} earlier; avoids seeming related to \citep{}
 When a hidden state is included (I/C+A/B model), the large net diversification difference persists for one of the hidden states but is removed for the other (\cref{figure:netdivall}D).
-Thus, differences in net diversification are best explained by both breeding system and an unknown factor, similar to results found by \citet{freyman_2018}.
+Thus, differences in net diversification are best explained by both breeding system and an unknown factor \citep[as in Onagraceae;][]{freyman_2018}.
 % The transition rate from SI to SC is $q_{IC}=0.3$ for both the I/C and the I/C+A/B models.
@@
 It thus appears that the difference in net diversification with breeding system persists when ploidy is included in the model, but not the reverse.
-The association of ploidy with net diversification in the D/P model (\cref{figure:netdivall}A, \cref{figure:netdivnodip}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids are similar.
+That is, the association of ploidy with net diversification in the D/P model (\cref{figure:netdivall}A, \cref{figure:netdivnodip}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids are similar.
 %
@@
 Evolution from self-incompatible diploids to self-compatible polyploids can proceed through two different pathways.
+In the one-step pathway, polyploidization directly disable SI.
+In the two-step pathway, SI is first lost within the diploid state, followed by polyploidization.
 Determining the relative contribution of these pathways based on our estimated transition rates from the ID/CD/CP model, we find that the one-step pathway is more likely on short timescales and the two-step pathway is more likely on long timescales (\cref{figure:pathways}, left panels).
@@
 % todo: precise figure refs here
-% E: Rosana, would this also be a good place to say something about what we learned regarding the meaningfulness of diploidization in these data?
-%B: maybe meaning of the data could go to a dedicated Discussion subsection?
-The idea would be to reconcile variation in $\delta$ across (the best-fitting) models considered here, and how the implied values do (or do not) agree with chromosome number changes and WGD inference from tree reconciliation and Ks approaches.
-(I can help here, if needed.)
 
@@
 
-    \begin{tabular}{lcccc}
-                       & D/P   & D/P+A/B      & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
+    \begin{tabular}{lcccccc}
+                       & D/P   & D/P+A/B      & I/C    & I/C+A/B      & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
         %                                                                                      
-        $r_D$          & 0.260 & 0.193, 0.658 & ---      & ---          \\
-        $r_P$          &-0.056 & 0.030, 0.187 & ---      & ---          \\
-        $r_I$          & ---   & ---          & ---      & ---          \\
-        $r_C$          & ---   & ---          & ---      & ---          \\
-        $r_{ID}$       & ---   & ---          & 0.455    & 0.309, 0.797 \\
-        $r_{CD}$       & ---   & ---          & 0.065    &-0.006, 0.587 \\
-        $r_{CP}$       & ---   & ---          &-0.088    &-0.074, 0.403 \\
+        $r_D$          & 0.260 & 0.193, 0.658 & ---    & ---          & ---      & ---          \\
+        $r_P$          &-0.056 & 0.030, 0.187 & ---    & ---          & ---      & ---          \\
+        $r_I$          & ---   & ---          & 0.550  & 0.386, 0.877 & ---      & ---          \\
+        $r_C$          & ---   & ---          &-0.001  &-0.059, 0.606 & ---      & ---          \\
+        $r_{ID}$       & ---   & ---          & ---    & ---          & 0.455    & 0.309, 0.797 \\
+        $r_{CD}$       & ---   & ---          & ---    & ---          & 0.065    &-0.006, 0.587 \\
+        $r_{CP}$       & ---   & ---          & ---    & ---          &-0.088    &-0.074, 0.403 \\
         %                                                                                      
-        $\rho$         & 0.047 & 0.047        & ---      & ---          \\
-        $\rho_I$       & ---   & ---          & 0.067    & 0.053        \\
-        $\rho_C$       & ---   & ---          & 0.033    & 0.032        \\
-        $q_{IC}$       & ---   & ---          & 0.198    & 0.145        
+        $\rho$         & 0.047 & 0.047        & ---    & ---          & ---      & ---          \\
+        $\rho_I$       & ---   & ---          & ---    & ---          & 0.067    & 0.053        \\
+        $\rho_C$       & ---   & ---          & ---    & ---          & 0.033    & 0.032        \\
+        $q_{IC}$       & ---   & ---          & 0.364  & 0.261        & 0.198    & 0.145        
     \end{tabular}
@@
 
-    \begin{tabular}{lcccccc}
-                       & D/P   & D/P+A/B      & I/C    & I/C+A/B      & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
+    \begin{tabular}{lcccc}
+                       & D/P   & D/P+A/B      & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
         %                                                              
-        $r_D$          & 0.382 & 0.698, 0.100 & ---    & ---          & ---      & ---          \\
-        $r_P$          & 0.109 & 0.587, 0.182 & ---    & ---          & ---      & ---          \\
-        $r_I$          & ---   & ---          & 0.550  & 0.386, 0.877 & ---      & ---          \\
-        $r_C$          & ---   & ---          &-0.001  &-0.059, 0.606 & ---      & ---          \\
-        $r_{ID}$       & ---   & ---          & ---    & ---          & 0.449    & 0.318, 0.789 \\
-        $r_{CD}$       & ---   & ---          & ---    & ---          & 0.050    &-0.248, 0.494 \\
-        $r_{CP}$       & ---   & ---          & ---    & ---          &-0.027    & 0.110, 0.634 \\
+        $r_D$          & 0.382 & 0.698, 0.100 & ---      & ---          \\
+        $r_P$          & 0.109 & 0.587, 0.182 & ---      & ---          \\
+        $r_I$          & ---   & ---          & ---      & ---          \\
+        $r_C$          & ---   & ---          & ---      & ---          \\
+        $r_{ID}$       & ---   & ---          & 0.449    & 0.318, 0.789 \\
+        $r_{CD}$       & ---   & ---          & 0.050    &-0.248, 0.494 \\
+        $r_{CP}$       & ---   & ---          &-0.027    & 0.110, 0.634 \\
         %                                                              
-        $\rho$         & 0.033 & 0.026        & ---    & ---          & ---      & ---          \\
-        $\rho_I$       & ---   & ---          & ---    & ---          & 0.063    & 0.047        \\
-        $\rho_C$       & ---   & ---          & ---    & ---          & 0.024    & 0.011        \\
-        $\delta$       & 0.050 & 0.162        & ---    & ---          & 0.022    & 0.107        \\
-        $q_{IC}$       & ---   & ---          & 0.364  &              & 0.194    & 0.164        
+        $\rho$         & 0.033 & 0.026        & ---      & ---          \\
+        $\rho_I$       & ---   & ---          & 0.063    & 0.047        \\
+        $\rho_C$       & ---   & ---          & 0.024    & 0.011        \\
+        $\delta$       & 0.050 & 0.162        & 0.022    & 0.107        \\
+        $q_{IC}$       & ---   & ---          & 0.194    & 0.164        
     \end{tabular}
@@
         Two comma-separated numbers refer to the $A$ and $B$ hidden states, and --- means the parameter was not present in the model.
-    %B: perhaps include this in the table so lazies don't have to go seeking subscripts?
     Net diversification rates ($r$) are subscripted with trait state initials (Diploid, Polyploid, Incompatible, Compatible).
@@
 
-%B: FIXME:
-%B: Table S2 (above) The I/C and I/C+A/B columns should maybe appear in the top panel, no?
-%B: I did not flip the panels, because I realized just now that we'll first discuss no-delta, then delta. Is that okay?
-%
 % check values of delta and rho D/P+A/B with(out) delta
@@
 
-% TODO: swap panels between the figures so that the first is -delta and the second is +delta
-
 \begin{figure}
-\includegraphics[width=\textwidth]{allnetdivmodelswithoutdelta.pdf} % FIXME: The panel labels should use "+A/B" instead of "-A/B".
-Rosana, if you add the .svg file for this figure, I can make small changes like this.
-\caption{Net diversification rates for all models.}  
+\includegraphics[width=\textwidth]{allnetdivmodelswithoutdelta.pdf}
+\caption{Net diversification rates for all models that exclude diploidization.}
 \label{figure:netdivall}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{allnetdivmodelsdiploidization.pdf} % FIXME: same note as above about "+A/B"
+\includegraphics[width=\textwidth]{allnetdivmodelsdiploidization.pdf}
 \caption{Net diversification rates for all models that include diploidization.}  

### 2019-03-10 manuscript/results.tex
added all ASR plots and the states plot data missing

@@
 \includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P, polyploidy model} % XXX
+\caption{Posterior distribution for each of the parameters in the D/P ploidy model} % XXX
+\label{suppfigure:DPnodip}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrDP.pdf}
+\caption{Ancestral state reconstruction showing the maximum a posteriori for each node of the D/P ploidy model} % XXX
 \label{suppfigure:DPnodip}
@@
 \includegraphics[width=\textwidth]{bisseDPposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P+ $\delta$,polyploidy model} % XXX
+\caption{Posterior distribution for each of the parameters in the D/P+ $\delta$ ploidy model} % XXX
 \label{suppfigure:DP}
@@
 \begin{suppfigure}
+\includegraphics[width=\textwidth]{asrDPdelta.pdf}
+\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the D/P+ $\delta$, ploidy model} % XXX
+\label{suppfigure:DPasr}
+\end{suppfigure}
+
+\begin{suppfigure}
 \includegraphics[width=\textwidth]{hisseDPnodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P+A/B, polyploidy model} % XXX
+\caption{Posterior distribution for each of the parameters in the D/P+A/B ploidy model} % XXX
 \label{suppfigure:DPnodipAB}
@@
 \begin{suppfigure}
+\includegraphics[width=\textwidth]{asrDPAB.pdf}
+\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the D/P D/P+A/B ploidy model} % XXX
+\label{suppfigure:DPnodipABasr}
+\end{suppfigure}
+
+\begin{suppfigure}
 \includegraphics[width=\textwidth]{hisseDPposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P+$\delta$+A/B, polyploidy model.  The axis is offset in one location so that the two overlapping distributions can be seen.} % XXX
+\caption{Posterior distribution for each of the parameters in the D/P+$\delta$+A/B ploidy model.  The axis is offset in one location so that the two overlapping distributions can be seen.} % XXX
 \label{suppfigure:DPAB}
@@
 \begin{suppfigure}
+\includegraphics[width=\textwidth]{asrDPABdelta.pdf}
+\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the D/P+$\delta$+A/B ploidy model.} % XXX
+\label{suppfigure:DPABasr}
+\end{suppfigure}
+
+\begin{suppfigure}
 \includegraphics[width=\textwidth]{bisseSIposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the I/C, breeding system model} % XXX
+\caption{Posterior distribution for each of the parameters in the I/C breeding system model} % XXX
 \label{suppfigure:IC}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIC.pdf}
+\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the I/C breeding system model} % XXX
+\label{suppfigure:ICasr}
 \end{suppfigure}
@@
 \begin{suppfigure}
+\includegraphics[width=\textwidth]{asrICAB.pdf}
+\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the I/C+A/B breeding system model} % XXX
+\label{suppfigure:ICABasr}
+\end{suppfigure}
+
+\begin{suppfigure}
 \includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the ID/CD/CP, polyploidy and breeding system model} % XXX
+\caption{Posterior distribution for each of the parameters in the ID/CD/CP ploidy and breeding system model} % XXX
 \label{suppfigure:IDCDCPnodip}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIDCDCP.pdf}
+\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP ploidy and breeding system model} % XXX
+\label{suppfigure:IDCDCPnodipasr}
 \end{suppfigure}
@@
 \begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIDCDCPdelta.pdf}
+\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP+$\delta$ ploidy and breeding system model} % XXX
+\label{suppfigure:IDCDCPasr}
+\end{suppfigure}
+
+\begin{suppfigure}
 \includegraphics[width=\textwidth]{muhisseDPSInodipposteriordist.pdf}
@@
 \label{suppfigure:IDCDCPnodipAB}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIDCDCPAB.pdf}
+\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP+A/B polyploidy and breeding system model} % XXX
+\label{suppfigure:IDCDCPnodipABasr}
 \end{suppfigure}
@@
 \end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIDCDCPdeltaAB.pdf}
+\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP+$\delta$+A/B, polyploidy and breeding system model} % XXX
+\label{suppfigure:IDCDCPABasr}
+\end{suppfigure}

### 2019-03-10 summarystats/solanaceaesummarystats.R

@@ -67,14 +67,16 @@ norploidysi.solanaceae<-simplified.solanaceae[-c(nor.ploidysi),]
 ##Just checking here that all data has either ploidy or breeding system info
 ##nor.ploidysi2<-which(is.na(norploidysi.solanaceae[,5])==TRUE & is.na(norploidysi.solanaceae[,6])==TRUE)
 
-matched<-make.treedata(solanaceae.tree,norploidysi.solanaceae) #613
+matched<-make.treedata(solanaceae.tree,norploidysi.solanaceae) #595
 write.nexus(matched$phy,file="~/Dropbox/solploidy/basicdata/fullmatchtree.nex")
 
 
-diploidsi<-which(matched$dat[,4]==2 & matched$dat[,5]=="SI")
-diploidsc<-which(matched$dat[,4]==2 & matched$dat[,5]=="SC")
-diploidsna<-which(matched$dat[,4]==2 & is.na(matched$dat[,5])==TRUE)
-polyploids<-which(matched$dat[,4]>2)
+diploidsi<-which(matched$dat[,4]==2 & matched$dat[,5]=="SI")# 97
+diploidsc<-which(matched$dat[,4]==2 & matched$dat[,5]=="SC")#152
+diploidsna<-which(matched$dat[,4]==2 & is.na(matched$dat[,5])==TRUE) #223
+polyploids<-which(matched$dat[,4]>2) #86
+polyploidsc<-which(matched$dat[,4]>2 & matched$dat[,5]=="SC")#43
+polyploidsna<-which(matched$dat[,4]>2 & is.na(matched$dat[,5])==TRUE)#43
 si<-which(is.na(matched$dat[,4])==TRUE & matched$dat[,5]=="SI")
 sc<-which(is.na(matched$dat[,4])==TRUE & matched$dat[,5]=="SC")
 threestate<- rep(0,dim(matched$dat)[1])

### 2019-03-10 manuscript/results.tex
fixing supplementary info crefs

@@
 For the two models that also included breeding system (ID/CD/CP and ID/CD/CP+A/B), the main effect of including diploidization is increasing the uncertainty of the estimate of polyploid net diversification rate (\cref{figure:netdivall}EF versus \cref{figure:netdivnodip}CD).
-For all models, there is greater uncertainty in the estimate of diploidization rate than polyploidization rate, as judged by the width of the credibility intervals (see supplementary information figures).
-% todo: precise figure refs here
+For all models, there is greater uncertainty in the estimate of diploidization rate than polyploidization rate, as judged by the width of the credibility intervals (see \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}).
 

### 2019-03-10 manuscript/discussion.tex
read through Results
- outline in discussion, removed errant mark in main
- first paragraph of Discussion

@@
 \section{Discussion}
+
+% outline
+% - disentangling effects
+% - general value of approach
+%  		another kind of confounding processes, not only transition and diversification, but "other diversifications" and "other transitions", esp. correlated
+%
+% - our study key findings and details
+% - effects of Ploidy (as %R)
+% - trumped by BS
+% - no difference on background of SC 
+% - discuss data/models and their long/shortcomings: ClaSSE, will's idea for stochastic mapping
+% - how to improve / what next
+%  (may be useful to study in SC families)
+%
+% - pathways
+% 
+%- diploidization
+% - rate, effect
+% - history in solanaceae
+%
+% - summary
+
+We investigated the individual and joint effects of ploidy and breeding system states in diversification of Solanaceae.
+Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are trumped by breeding systems when the two traits are considered jointly.
+Interestingly, the difference in diversification rates between the diploids and polyploids is negligible on the background of self-compatibility.
+%B: should we mention here that the hidden states presage these results in each case?
+Maybe we don't want to make too much of that, but it is fairly interesting.
+Also possibly hit that in your paragraph(s), below.
+In the same vein, the difference in diversification rates between self-incompatible and self-compatible lineages is smaller in the diploid genomic background.
+In our final analyses, motivated by the widespread findings of paleopolyploid ancestry across angiosperms, we did not recover strong support for inclusion of diploidization process.
+Below, we discuss the key findings, and outline a broad strategy for applying phylogenetic comparative methods in the study of complex interactions among traits, in the context of diversification.
+%  Surprising outcomes: (1)  D and P on background of C, and (2) SC not as dead of an end as it was in goldberg2010, if P is partitioned out (can be framed as kind of path-dependence, i.e.
+qIC path different than rho path, or same as (1), so that I and C are less different on the background of D.
 
@@
 
+% And, are diversification rates of polyploid SC lineages different if one or other path was taken?
+\citet{charlesworth1985} % E: This is a great question.
+We don't answer it with this round of analyses.
+But discuss?
+
 % where WGD is in Solanaceae--motivation and delta rate analyses

### 2019-03-10 manuscript/discussion.tex
- minor fixes to the previous version of first Discussion paragraph

@@
 
+%B:  at least two surprising outcomes: (1)  D and P on background of C, and (2) SC not as dead of an end as it was in goldberg2010, if P is partitioned out (can be framed as kind of path-dependence, i.e.
+qIC path different than rho path, or same as (1), so that I and C are less different on the background of D.
 We investigated the individual and joint effects of ploidy and breeding system states in diversification of Solanaceae.
@@
 Also possibly hit that in your paragraph(s), below.
-In the same vein, the difference in diversification rates between self-incompatible and self-compatible lineages is smaller in the diploid genomic background.
-In our final analyses, motivated by the widespread findings of paleopolyploid ancestry across angiosperms, we did not recover strong support for inclusion of diploidization process.
-Below, we discuss the key findings, and outline a broad strategy for applying phylogenetic comparative methods in the study of complex interactions among traits, in the context of diversification.
-%  Surprising outcomes: (1)  D and P on background of C, and (2) SC not as dead of an end as it was in goldberg2010, if P is partitioned out (can be framed as kind of path-dependence, i.e.
-qIC path different than rho path, or same as (1), so that I and C are less different on the background of D.
+In the same vein, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against diploid genomic background.
+%B: "in diploids" may work better?
+%B: need help with pathway single-sentence summary
+Motivated by the widespread findings of paleopolyploid ancestry across angiosperms, we also examined support for diploidization in the history of the family, which appears to be weak.
+%B: "mixed" "not strong" what's the right word here?
+Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests complex hypotheses regarding a heterogenous diversification process.
+Below, we discuss these key findings, and outline a broad approach for application of phylogenetic comparative methods in the context of interactions among traits and diversification process.
 

### 2019-03-10 manuscript/discussion.tex
- diploidization paragraph merge in Discussion

@@
 % - no difference on background of SC 
-% - discuss data/models and their long/shortcomings: ClaSSE, will's idea for stochastic mapping
+% - discuss data/models and their long/shortcomings: 
+%	ClaSSE, will's idea for stochastic mapping
+%	irreversibility SI -> SC assumption
 % - how to improve / what next
@@
 Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests complex hypotheses regarding a heterogenous diversification process.
-Below, we discuss these key findings, and outline a broad approach for application of phylogenetic comparative methods in the context of interactions among traits and diversification process.
+Below, we discuss these key findings and outline a broad approach for application of phylogenetic comparative methods in the context of interactions among traits and diversification process.
 
@@
 
-% where WGD is in Solanaceae--motivation and delta rate analyses
-% diploidization: comment
 Diploidization needs to be considered in diversification models because higher rates of net diversification (speciation minus extinction) in diploids can be obtained from models that ignore the possibility of a polyploid lineage that has diploidized being the diversification enhancer.
@@
 By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with stochastic models. 
-In past studies by \citet{mayrose_2011, mayrose_2015} diploidization could not be formally  considered in models because ploidy classification was done via chromosome change models (i.e.
-Chromevol \citep{glick2014}) that do not allow for reversion of polyploidy.
-However, in our dataset ploidy was assigned based on data aggregated from independent studies (see supplementary information and citations in archived dataset) that come from field or cytogenetic studies.
 In the cases where ploidy level was assigned by us based chromosome multiplicity at the genus level, estimating diploidization  might be a potential issue due to ploidy misclassification.
 
-% B: This is my pitch for motivation of the \delta paragraph and/or wording for how to reconcile estimates and genomic data (providing summary of data--to date)
+%B: v2.0
 Diploidization is widely considered to be relatively rare, compared with polyploidization, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history.
-As a consequence, it is possible that nearly all extant species classified as ``diploid" in our analyses are possibly secondary diploids, having undergone both polyploidization and re-diploidization.
-Genomic evidence indicates that such an event may have occurred prior to the origin of the family Solanaceae. 
-\citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \textit{Solanum lycopersicum}, may have experienced one or more paleopolyploid events.
-A subsequent analysis of synteny between grape and \textit{Solanum} genomes, as well as the distribution between inferred paralogs within \textit{Solanum} (tomato and potato) genomes both suggest that this lineage experienced a likely round of ancient genome duplication or triplication \citep{tomato2012}.
+%cite
+As a consequence, it is possible that nearly all extant species classified as ``diploid" in our analyses, based on cytogenetic data, are possibly secondary diploids, having undergone both polyploidization and re-diploidization.
+Unduly ignoring secondary diploidization would clearly underestimate the rate of transitions from polyploids to diploids, and would likely cause inflated net diversification rates for diploids.
+We find weak support for inclusion of diploidization. %cite fig and expand (FIXME).
+
+Genomic evidence indicates that a paleo-polyploidization may have occurred prior to the origin of the family Solanaceae, rendering all current `diploids' secondarily derived. 
+First, \citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \textit{Solanum lycopersicum}, may have experienced one or more WGDs.
+A subsequent analysis of synteny between grape and \textit{Solanum} genomes, as well as the distribution between inferred paralogs within genomes of \textit{Solanum} (tomato and potato) each suggested that this lineage experienced a likely round of ancient genome duplication or triplication \citep{tomato2012}.
 The age of the peak of paralog Ks distances, is approximately 71 million years \citep{tomato2012}. 
-If this is the case, then all of the genomes may have been subsequently re-diploidized, because common base chromosome numbers this and related families are n=11-12.
+If this is the case, then all of the genomes may have been subsequently re-diploidized, yielding the common extant base chromosome numbers in this and related families, n=11-12, considered diploid \citep{robertson2011}.
+%Not unlikely, because Arabidopsis n=5.
 
-%Not unlikely, because Arabidopsis n=5.
-On the other hand, studies comparing map-based genome synteny within the family find no evidence for recent polyploidization-diploidization \citep{wu_2010a}. 
-Simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species, including all of those in the relatively cytogenetically conserved `x=12' group, which includes tomato, potato, eggplant, pepper, and tobacco.
+On the other hand, there is little evidence for occurrence of re-diploidization within the nightshades.
+First, the Ks-inferred duplication long pre-dates the origin of the family \citep{sarkinen_2013}. 
+The best-supported, most-recent WGD, which may represent a reversion from polyploidy to diploidy, is older than the lineage under consideration in this study.
+At best, the species we consider are ancestrally diploid, having re-diploidized after a suspected WGD.
+Second, studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
+Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species, including all of those in the relatively cytogenetically conserved `x=12' group, which includes tomato, potato, eggplant, pepper, and tobacco.
 %B: Once this is re-arranged, and made to fit with the rest of the discussion, add what this has to do with re-diploidization and how to interpret $\delta$.
-Therefore, it is not surprising that for all models diploidization  had wide credible intervals for parameter $\delta$.
-Despite the classification and genomic limitations to assign ploidy levels and diploidization, we found support for models that include pararmeter $\delta$, meaning that further studying this process is worth pursuing.
-
-%     effect on inferred diversification
+It is perhaps not surprising that models diploidization in our study recovered wide credible intervals for the diploidization parameter, $\delta$. 
+Despite the limitations of assignment of ploidy levels, we found some support for models that include pararmeter $\delta$, meaning that further study may be warranted, especially in other groups.
+%Moreover, rare state changes may not be easily recovered. Emma, you had a clever way to summarize this. 
+%     meaning of diploidization parameter
 
@@
 % issues
-%     meaning of diploidization parameter
 %     irreversibility SI -> SC assumption
-
-% Missing a good concluding paragraph.

### 2019-03-11 manuscript/introduction.tex
- cleanup of Intro, title alteration, couple of new refs

@@
 Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and they crucially affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
-In particular, self-incompatibility causes plants to reject their own pollen, and its loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974,igic_2008}.
+In particular, self-incompatibility (SI) causes plants to reject their own pollen, and loss of self-incompatibility, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974,igic_2008}.
 % Gordon: Really?
@@
 %B:  sigh.
-Previous analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, even when allowing the possibility that unobserved states drive diversification patterns \citep{freyman_2018}.
-Despite their apparent macroevolutionary role in diversification, breeding systems are clearly not the sole trait determining diversification rates.
+Previous analyses reported higher rates of diversification for SI than for SC lineages in Solanaceae \citep{goldberg_2010}. 
+Similarly, heterostylous SI lineages in Primulaceae seem to diversify faster \citep{devos2014}, as do outcrossing lineages in Onagraceae \citep{freyman_2018}.
+Despite consistent inferences of their macroevolutionary role, breeding systems are clearly not the sole trait determining diversification rates.
+
 Ploidy, for example, is widely thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
@@
 %The evidence for correlated shifts in mating system, from outcrossing to selfing, following WGD, appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
-Moreover, polyploidy is not only a mere correlate of breeding system, but may directly cause the loss of self-incompatibility (SI) \citep{stout1942, lewis1947}, and facilitate the transition to selfing.
+Polyploidy is not a mere correlate of transition in breeding system, but may directly cause the loss of SI \citep{stout1942, lewis1947}, as well as facilitate the transition to selfing \citep{barringer2007, barrett2008, husband2008}.
 Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic SI systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
-This precludes the existence of SI polyploids in the most widespread system of SI, and creates a correlation between polyploidy and SC.
+With exceedingly few exceptions \citep{hauck_2002,nunes_2006}, this generally precludes the existence of SI polyploids in the most widespread system of SI, which is regarded as ancestral in eudicots \citep{igic_2001,steinbachs_2002}.
+This system, RNase-based SI, is expressed in all SI species of Solanaceae examined to date, and creates a strong correlation between polyploidy and SC \citep{robertson_2011}.
 % Gordon comments: Here again is confused about breeding systems in general vs. SI genetic systems.
@@
 
-Given that changes in ploidy and breeding systems may be causally related and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
-First, it may be of particular interest to establish their joint influence on lineage diversification.
-It is possible that combinations of breeding system and ploidy do not result in predictable additive effects on speciation and extinction rates.
-Second, the two traits may affect each other's state transitions, including their magnitude and ordering of transitions.
+Given that changes in ploidy and breeding systems are causally related, and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
+First, it may be of interest to establish their joint influence on lineage diversification.
+Combinations of breeding system and ploidy may not have individually predicted, additive effects on speciation and extinction rates.
+Second, the two traits may affect each other's state transitions, including both magnitudes and ordering.
 For example, losses of SI in diploids could be commonly caused by polyploidization, resulting in a one-step pathway to polyploid SC species. 
@@
 
-Two recent methodological developments are critical for disentangling the complex interactions between two or more traits and their link to diversification.
+Two recent methodological developments enable disentangling the complex interactions between two or more traits and diversification.
 %First, the model proposed by \citet{fitzjohn_2012} that extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multiple states, MuSSE, instead of binary states, BiSSE).
@@
 %Second, the model by \citet{beaulieu_2016} HiSSE proposed a first solution to the problem of  elevated false positive type I errors found often in the BiSSE framework \citep{goldberg_2012}. 
-Second, \citet{beaulieu_2016}  proposed an extension of the *SSE framework by introducing a model with `hidden' or unspecified traits (HiSSE), which can affect the diversification process.
+Second, \citet{beaulieu_2016}  proposed an extension of the *SSE framework by introducing a model with unspecified, or `hidden' traits (HiSSE model), which can affect the diversification process.
 %The authors argued that it is unlikely that the variation in speciation and extinction rates is only due to a focal trait but by considering heterogeneity in the diversification process as the null hypothesis, it is possible to parse out the importance of the focal trait in diversification while other unobserved processes can happen.
-Because variation in speciation and extinction rates may not be due solely to the focal trait, HiSSE simultaneously enables such variation to be explained by unobserved states, and reduces the elevated false-positive error rates \citep{rabosky_2015}.
+Because variation in speciation and extinction rates may not be due solely to the focal trait, HiSSE simultaneously enables hidden states to explain such variation, and reduces the elevated false-positive error rates \citep{rabosky_2015}.
+%B: This may be a good portion of Discussion--faults and qualities of methods?
 %HiSSE models have effectively shown that some traits associated with diversification in BiSSE analyses are actually not associated with the diversification process, and instead a hidden state is linked to differences in the diversification rate.

### 2019-03-11 manuscript/main.tex
First part of the discussion completed, missing some of it
- outline in discussion, removed errant mark in main
fixing a plot
- cleanup of Intro, title alteration, couple of new refs

@@
 
-% Accumulate ideas for titles here...
-%
-% When your favorite trait is not enough to explain diversification: An example of diversification linked to polyploidy and breeding system in Solanaceae % E: Fun, but I think we need something more formal
-%
-%% When the favorite trait is not enough: the link of polyploidy and breeding system in diversification %B: <- eh?
-%
 % New comparative approaches for the study of breeding system, polyploidy, and lineage diversification % E: the tentative title given to New Phyt
@@
 %
-%B: associations isn't exactly right; I think it somewhat undermines the causality from DP to IC; anyway, it's not particularly bad.
-%
-%% Interaction of ploidy and breeding system in lineage diversification %B: <- how 'bout that?
-%
 \begin{center}
-    \textsc{Interacting associations between ploidy, breeding system, and lineage diversification}
+    \textsc{Interaction between ploidy, breeding system, and lineage diversification}
 
@@
 %
-\noindent$^{2}$Department of Biology, University of Hawai'i M\={a}noa, Honolulu, HI 96822, U.S.A.\\
+\noindent$^{1}$Department of Biology, University of Hawai'i M\={a}noa, Honolulu, HI 96822, U.S.A.\\
 %
@@
 %
-\noindent$^\dagger$\textit{Author for correspondence:}.\\
+\noindent$^\dagger$\textit{Author for correspondence}.\\
 

### 2019-03-11 manuscript/main.tex
- acknowledgments addition
- funding sources wording addition

@@
 
-%B: Perhaps other's funding sources for max BSF/NSF payoff? I added ours; no harm, I presume. 
-NSF DEB-1655692 to BI, and DEB-1655478 to EEG.
-The Minnesota Supercomputing Institute (MSI) at the University of Minnesota provided computing resources for this project.
+This work was supported by the National Science Foundation (DEB-1655692 to BI, and DEB-1655478 to EEG) and
+the Minnesota Supercomputing Institute (MSI) at the University of Minnesota, which provided computing resources for this project.
+%BI would like to thank SCHB for inspiration, sparring, and encouragement; maybe EEG, as well?
 

### 2019-03-11 manuscript/results.tex
Results edits, from Boris' comments

@@
 
-When considered separately, ploidy and breeding system each showed significant associations with net diversification differences in Solanaceae.
-When considered together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well.
+When analyzed separately, ploidy and breeding system each showed significant associations with net diversification differences in Solanaceae.
+When analyzed together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well.
 
 When considering ploidy alone (D/P model), we found a greater net diversification rate for diploids than for polyploids, in agreement with \citep{mayrose_2011, mayrose_2015}.
-This result holds with (\cref{figure:netdivall}A) or without (\cref{figure:netdivnodip}A) the diploidization parameter, though including diploidization shifts the net diversification rate of polyploids to be non-negative.
-Incorporating a hidden state, however, removes the clear separation in diversification between diploids and polyploids (D/P+A/B model; \cref{figure:netdivall}B, \cref{figure:netdivnodip}B).
+This result holds with (\cref{figure:netdivall}A) or without (\cref{figure:netdivnodip}A) the diploidization parameter, although including diploidization shifts the net diversification rate of polyploids to be non-negative.
+Incorporating a hidden state in this model, however, removes the clear separation in diversification between diploids and polyploids (D/P+A/B model; \cref{figure:netdivall}B, \cref{figure:netdivnodip}B).
 Thus, differences in net diversification are better explained by an unknown factor than by ploidy.
@@
 When considering ploidy and breeding system together (ID/CD/CP model), the net diversification rate for SI diploids was greater than for either SC diploids or SC polyploids, with or without diploidization (\cref{figure:netdivall}E, \cref{figure:netdivnodip}C).
-It thus appears that the difference in net diversification with breeding system persists when ploidy is included in the model, but not the reverse.
+The difference in net diversification with breeding system persists when ploidy is included in the model, but not the reverse.
 That is, the association of ploidy with net diversification in the D/P model (\cref{figure:netdivall}A, \cref{figure:netdivnodip}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids are similar.
@@
 
-Evolution from self-incompatible diploids to self-compatible polyploids can proceed through two different pathways.
-In the one-step pathway, polyploidization directly disable SI.
+There are two pathways by which a SI diploid lineages will eventually---given enough time---become a SC polyploid.
+In the one-step pathway, polyploidization directly disables SI.
 In the two-step pathway, SI is first lost within the diploid state, followed by polyploidization.
@@
 When more time has elapsed, the two-step pathway is more likely because the rate of loss of SI within diploids, $q_{IC}$, is greater than the rate of polyploidization for SI species, $\rho_I$ (\cref{suppfigure:IDCDCPnodip}).
-These conclusions are the same as those of \citet{robertson_2011} under one of their branch length approximations.
+% These conclusions are agree with those of \citet{robertson_2011}.
 
-When rates of lineage diversification are considered, however, the conclusions change.
-Even over long timescales, the two-step pathway contributes less (\cref{figure:pathways}, right panels).
+When rates of lineage diversification are incorporated, however, the conclusions change.
+Beginning again with a single ID lineage, it is now necessary to consider its possible proliferation in the ID state, and also in the CD state for the two-step pathway, and to track the progression of the full resulting clade to the CP state.
+In this situation, even over long timescales, the two-step pathway contributes less (\cref{figure:pathways}, right panels).
 The lower rate of net diversification in the CD state, relative to ID, means that relatively fewer lineages are available to complete the second step of the two-step pathway.
@@
 
-We considered models both with or without diploidization in order to explore the effects of this process on the estimates of state-dependent diversification.
+We considered models both with or without diploidization in order to explore its effects on the estimates of state-dependent diversification.
 For the two models that only include diploid and polyploid states (D/P and D/P+A/B), the net diversification rate of polyploid lineages is likely negative when diploidization (parameter $\delta$) is excluded but positive when $\delta$ is included (\cref{figure:netdivall}A versus \cref{figure:netdivnodip}A).
-For the two models that also included breeding system (ID/CD/CP and ID/CD/CP+A/B), the main effect of including diploidization is increasing the uncertainty of the estimate of polyploid net diversification rate (\cref{figure:netdivall}EF versus \cref{figure:netdivnodip}CD).
+For the two models that also included breeding system (ID/CD/CP and ID/CD/CP+A/B), the main effect of including diploidization is greatly increasing the uncertainty of the estimate of polyploid net diversification rate (\cref{figure:netdivall}EF versus \cref{figure:netdivnodip}CD).
 For all models, there is greater uncertainty in the estimate of diploidization rate than polyploidization rate, as judged by the width of the credibility intervals (see \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}).

### 2019-03-11 manuscript/discussion.tex
- minor change to Discussion, diploidization part
removed some tex warnings

@@
 %B: v2.0
+\subsection*{Diploidization}
+
 Diploidization is widely considered to be relatively rare, compared with polyploidization, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history.
@@
 The age of the peak of paralog Ks distances, is approximately 71 million years \citep{tomato2012}. 
-If this is the case, then all of the genomes may have been subsequently re-diploidized, yielding the common extant base chromosome numbers in this and related families, n=11-12, considered diploid \citep{robertson2011}.
+If this is the case, then all of the genomes may have been subsequently re-diploidized, yielding the common extant base chromosome numbers in this and related families, n=11-12, considered diploid \citep{robertson_2011}.
 %Not unlikely, because Arabidopsis n=5.
 
-On the other hand, there is little evidence for occurrence of re-diploidization within the nightshades.
+On the other hand, there is little evidence for occurrence of re-diploidization in the time frame we considered, within the nightshades.
 First, the Ks-inferred duplication long pre-dates the origin of the family \citep{sarkinen_2013}. 

### 2019-03-11 manuscript/introduction.tex
final read of Intro

@@
 
-%B->%E: instead of a major reorganization, I added a breeding system quote from Barrett that presages our results. How's that? It seemed daunting to reorganize ploidy-first to BS-first.
+%B->%E: instead of a major reorganization, I added a breeding system quote from Barrett that presages our results. How's that? It seemed daunting to reorganize ploidy-first to BS-first.  E: nice!
 
@@
 Dramatic increases in the available phylogenetic and phenotypic data, as well as methodological advances, have greatly accelerated the search for traits that influence diversification traits.
-Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a difficult statistical problem (e.g., \citealt{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}).
+Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a difficult statistical problem \citep[\eg][]{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}.
 %is it only a "statistical" problem?
@@
 
-%Breeding system and ploidy are perhaps the best-studied traits linked to diversification \citep{stebbins1950}. %B rm: no mention of breeding system in this paragraph?
 Whole genome duplication (WGD) is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
 The widespread prevalence of variation in ploidy has long been considered a salient feature of flowering plant lineages \citep{stebbins1938}. 
-WGDs have the potential to affect many phentoypes and impact a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
-Polyploids have long been thought to have lower rates of net diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
+WGDs have the potential to affect many phentoypes and a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
+Polyploids have been thought to have lower rates of net diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
 Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that WGDs have played an important macroevolutionary role driving innovation and diversification in plants.
-Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies the pervasiveness of diploidization, the return of polyploids to the diploid state throughout the angiosperm phylogeny \citep{soltis_2015diploidization, dodsworth_2015}.
-Testing for the presence of diploidization in the context of diversification is important because the presence of a positive net diversification rate for diploids can be the result of an ancestrally polyploid lineage.
-Under this scenario, polyploidy would no longer be considered to reliably slow down diversification, but instead, polyploidy would be the initiator of the diversification process.
+Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies the pervasiveness of diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
+% Testing for the presence of diploidization in the context of diversification is important because the presence of a positive net diversification rate for diploids can be the result of an ancestrally polyploid lineage.
+% Under this scenario, polyploidy would no longer be considered to reliably slow down diversification, but instead, polyploidy would be the initiator of the diversification process.
 %B: I think this should read "...polyploidy is no longer be considered an evolutionary dead end."  %R- This last phrase is from Itay and I rather like it.
@@
 In other words, I also agree with both %R and Itay, but not with what the current phrasing is asserting.
+% E: I wrote the two sentences below to try to reconcile the text and comments below.  But then I realized we don't actually test this hypothesis.
+% One hypothesis is that diploids have a higher net diversification rate if they were ancestrally polyploid than if they had no polyploid ancestry.
+% Testing this hypothesis requires incorporating diploidization in analyses of diversification.
+% E: so try again -- okay?
+This could affect interpretation of different net diversification rates for diploids and polyploids, especially if diploids have a higher net diversification rate if they were ancestrally polyploid than if they had no polyploid ancestry.
 
-Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and they crucially affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
+Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and crucially affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
 In particular, self-incompatibility (SI) causes plants to reject their own pollen, and loss of self-incompatibility, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974,igic_2008}.
@@
 %The evidence for correlated shifts in mating system, from outcrossing to selfing, following WGD, appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
-Polyploidy is not a mere correlate of transition in breeding system, but may directly cause the loss of SI \citep{stout1942, lewis1947}, as well as facilitate the transition to selfing \citep{barringer2007, barrett2008, husband2008}.
+Polyploidy may facilitate the transition to selfing \citep{barringer2007, barrett2008, husband2008}, or it may directly cause the loss of SI \citep{stout1942, lewis1947}.
+% Polyploidy is not a mere correlate of transition in breeding system, but may directly cause the loss of SI \citep{stout1942, lewis1947}, as well as facilitate the transition to selfing \citep{barringer2007, barrett2008, husband2008}.
+E: reworded because no one would think the only thing P does is follow BS?
 Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic SI systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
 With exceedingly few exceptions \citep{hauck_2002,nunes_2006}, this generally precludes the existence of SI polyploids in the most widespread system of SI, which is regarded as ancestral in eudicots \citep{igic_2001,steinbachs_2002}.
-This system, RNase-based SI, is expressed in all SI species of Solanaceae examined to date, and creates a strong correlation between polyploidy and SC \citep{robertson_2011}.
+This system, RNase-based SI, is expressed in all SI species of Solanaceae examined to date.
+There is thus a strong correlation between polyploidy and SC in this family \citep{robertson_2011}.
 % Gordon comments: Here again is confused about breeding systems in general vs. SI genetic systems.
@@
 Emma?
+% E: I vote for using terms correctly without providing a glossary---lead by example?
 
 Given that changes in ploidy and breeding systems are causally related, and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
-First, it may be of interest to establish their joint influence on lineage diversification.
-Combinations of breeding system and ploidy may not have individually predicted, additive effects on speciation and extinction rates.
-Second, the two traits may affect each other's state transitions, including both magnitudes and ordering.
-For example, losses of SI in diploids could be commonly caused by polyploidization, resulting in a one-step pathway to polyploid SC species. 
-Alternatively, the path from SI diploids to SC polyploids may involve a transition to SC, without in increase in ploidy, and then a subsequent diploidization. 
-%Or whether polyploids have arised more commonly from self-incompatible than from self-compatible diploids
+First, it may be of interest to establish their joint influence on speciation and extinction rates.
+If the effects of ploidy and breeding system on lineage diversification are not additive, studying each trait separately may not reveal the effects of their combinations.
+Second, the two traits may affect each other's state transitions.
+For example, losses of SI in diploids could be caused by polyploidization, resulting in a one-step pathway to polyploid SC species. 
+Alternatively, the path from SI diploids to SC polyploids may involve a transition to SC, without an increase in ploidy, and then a subsequent polyploidization. 
 \citet{robertson_2011} found that the pathway from SI diploids to SC polyploids is dominated by the two-step pathway over long timescales, but generally proceeds in one step via polyploidization of SI species over short timescales.
-A vast accumulation of data and methods that allow for diversification rate differences now allow far more powerful approach to such problems.
+It remains to be seen whether these results hold true when allowing for the influence of these traits on lineage diversification.
 
 Two recent methodological developments enable disentangling the complex interactions between two or more traits and diversification.
-%First, the model proposed by \citet{fitzjohn_2012} that extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multiple states, MuSSE, instead of binary states, BiSSE).
-First, \citet{fitzjohn_2012} extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007}, to allow estimating the effect of an arbitrary number of states (enabling multi-state, MuSSE, instead of binary state, BiSSE, speciation and extinction models).
+First, \citet{fitzjohn_2012} extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007} to allow estimating the effect of an arbitrary number of states (enabling multi-state, MuSSE, instead of binary state, BiSSE, speciation and extinction models).
 This allows multiple states to represent trait combinations, particularly useful if one is interested in explicitly testing whether combinations of traits non-additively affect speciation or extinction.
-%Second, the model by \citet{beaulieu_2016} HiSSE proposed a first solution to the problem of  elevated false positive type I errors found often in the BiSSE framework \citep{goldberg_2012}. 
 Second, \citet{beaulieu_2016}  proposed an extension of the *SSE framework by introducing a model with unspecified, or `hidden' traits (HiSSE model), which can affect the diversification process.
-%The authors argued that it is unlikely that the variation in speciation and extinction rates is only due to a focal trait but by considering heterogeneity in the diversification process as the null hypothesis, it is possible to parse out the importance of the focal trait in diversification while other unobserved processes can happen.
-Because variation in speciation and extinction rates may not be due solely to the focal trait, HiSSE simultaneously enables hidden states to explain such variation, and reduces the elevated false-positive error rates \citep{rabosky_2015}.
+Because variation in speciation and extinction rates may not be due solely to the focal trait, HiSSE simultaneously enables hidden states to explain such variation, and it reduces the elevated false-positive error rates \citep{rabosky_2017}.
 %B: This may be a good portion of Discussion--faults and qualities of methods?
@@
 Using extensive trait data, and considering other possible sources of heterogeneity in the diversification process, we assess the influence of these traits on speciation and extinction rates.
-First, we present models in which the effect of ploidy and breeding system are considered separately, and compare the inferences to previously published results. 
-Second, we consider the effects of addition of a hidden trait to each of the models that examine traits individually, and investigate whether the focal trait directly affects differences in diversification, or if such differences are attributable to unobserved but correlated states.
-Third, we also ask whether the complex interaction between breeding system and ploidy may be explained by the unobserved trait states. 
-Fourth, we evaluate a multi-state model, which simultaneously considers the effect of both traits on diversification, and then consider whether an additional unobserved trait may be driving differences in diversification.
-For all of the above models, we investigate the potential effects of diploidization. 
+First, we present models in which the effect of ploidy and breeding system are considered separately, and we compare the inferences to previously published results. 
+Second, we add a hidden trait to each of those models and investigate whether the focal trait still explains differences in diversification.
+Third, we model ploidy and breeding system jointly to assess their combined relationship to diversification, with or without an additional hidden trait.
+With this joint trait model, we also quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
+For all of the ploidy models, we investigate the potential effects of including diploidization.
 Our results highlight the importance of considering non-additive effects of traits on net diversification rates, especially when there are strong biologically driven correlations among them.
-
-%BWe perform model selection in a Bayesian framework to identify which models are fitting better the evidence of the Solanaceae data.
-%Considering them jointly, however, reveals that the ploidy connection is removed by incorporating breeding system.
-%We further show that the general results are robust to allowing for diploidization and a hidden trait, and something about pathways. % FIXME
-%Our results emphasize the importance of considering traits not only in isolation, especially when there are strong correlations between them.
-%We investigate their individual and joint effects on diversification, and whether or not adding hidden states at different steps of modeling can help us generate hypothesis about the benefit of searching for a trait that can pass as a hidden state in a simple model.
-%Additionally, a lack of methods that allow for simultaneous inference of effects of other traits, which also influence diversification, can lead to incorrect conclusions about the focal trait \citep{rabosky_2015, beaulieu_2016}.

### 2019-03-11 manuscript/discussion.tex
read new parts of Discussion

@@
 Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are trumped by breeding systems when the two traits are considered jointly.
+% E: is "trumped" too much like gloating?
 Interestingly, the difference in diversification rates between the diploids and polyploids is negligible on the background of self-compatibility.
@@
 %B: "mixed" "not strong" what's the right word here?
-Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests complex hypotheses regarding a heterogenous diversification process.
+E: not "weak", maybe "not strong" or "not nearly as strong as anything else"
+Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests of complex hypotheses regarding a heterogenous diversification process.
 Below, we discuss these key findings and outline a broad approach for application of phylogenetic comparative methods in the context of interactions among traits and diversification process.
@@
 %  R- Good point. Emma will clarify
+%  E: Yes, good point.  Rosana, do you have results for state-independent diversification?  We could compare with using those rates.
 %
@@
 %    the rest of the discussion, so modify or discard as needed:
+% E: A small defense of excluding clado is that it didn't affect the diversification results in goldberg2010 vs goldberg2012
 Our work shows that a diploid Solanaceae lineage is much more likely to take a
@@
 But discuss?
+% E: To elaborate: In the IC/CD/CP model, the CP state would be subdivided into two hidded values, corresponding to the influx from the two pathways.
+Those two states, CP_X and CP_Y, could have different lambda and mu.
 
@@
 Unduly ignoring secondary diploidization would clearly underestimate the rate of transitions from polyploids to diploids, and would likely cause inflated net diversification rates for diploids.
+% E: first clause seems like a tautology?
+second clause does not seem obvious, but could at least say that it could affect div rate estimates
 We find weak support for inclusion of diploidization. %cite fig and expand (FIXME).
@@
 At best, the species we consider are ancestrally diploid, having re-diploidized after a suspected WGD.
+% E: But then how is there SI??
 Second, studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.

### 2019-03-11 manuscript/discussion.tex
minor edits across everything, added ideas to discussion for Boris

@@
 We investigated the individual and joint effects of ploidy and breeding system states in diversification of Solanaceae.
-Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are trumped by breeding systems when the two traits are considered jointly.
+Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are trumped by hidden states and or/ breeding system when added to models.
 % E: is "trumped" too much like gloating?
+- R-Haha
+%R- I added hidden states because that hasn't been discussed ever in polyploidy
 Interestingly, the difference in diversification rates between the diploids and polyploids is negligible on the background of self-compatibility.
@@
 Disentangling the patterns of diversification linked to  multiple traits allowed us the opportunity to model and test complex hypotheses about the heterogenous diversification process in Solanaceae.
-Species are created and go extinct based on multiple and often highly correlated phenotypes.
-However, estimating the speciation and extinction rates remains challenging.
-Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as disscussed in  \citet{rabosky_2010, beaulieu_2015}),but also, a problem of inadequate model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focus trait first used to model diversification.
-We presented a roadmap with a series of analyses that can help scientists to identify if an unknown or unobservable trait is worth pursuing.
+Species are created and go extinct based on multiple and often highly correlated phenotypes. However, estimating the speciation and extinction rates remains challenging.
+
 By carefully aggregating and curating ploidy and breeding system data for Solanaceae, we showed how  the complex interaction between ploidy and breeding system  can lead to diversification conclusions that differ from the diversification conclusions of models that focus on one trait at a time.
@@
 second clause does not seem obvious, but could at least say that it could affect div rate estimates
-We find weak support for inclusion of diploidization. %cite fig and expand (FIXME).
+We find strong support in models that included diploidization \cref{table:bayesfactors}.
+However, estimating the rate at which it occurs based on ploidy level classifications is hard and highly uncertain (see parameter $\delta$ in \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}) and can change the pattern of ancestral reconstructions (\cref{suppfigure:DPnodipasr,suppfigure:DPasr,suppfigure:DPnodipABasr,suppfigure:DPABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPasr,suppfigure:IDCDCPnodipABasr,suppfigure:IDCDCPABasr}).
+%We find weak support for inclusion of diploidization. %cite fig and expand (FIXME).
+% R- I corrected the above phrase, we found strong support for the inclusion of diploidization, models that have it are always preferred over models that don't however that parameter is hard to estimate
 
@@
 %     irreversibility SI -> SC assumption
+
+% R-Here I copied the limitations of the models idea but needs to be reorganized to be cohesive % Fix this a lot more please
+% This is the idea on how thorough analyses should be done.
+Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as disscussed in  \citet{rabosky_2010, beaulieu_2015}),but also, a problem of inadequate  ``*SSE'' model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focus trait first used to model diversification.
+We presented a roadmap with a series of analyses that can help scientists to identify if an unknown or unobservable trait is worth pursuing.
+In this roadmap, we have proposed fitting HiSSE models that have effectively shown that some binary states associated with diversification under BiSSE analyses are actually not different in net diversification terms, and instead a hidden state is linked to differences in the diversification rate \citep{beaulieu_2016}.
+If that is the case on a system, raising the question of the identity of the hidden state should be considered as the next step in the modeling process.
+When considering  a second candidate trait, one can model the complex interactions via a MuSSE model, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
+Since MuSSE might also suffer from hight type I errors considering more heterogeneity using a hidden state model on top of the multistate SSE is necessary to adequately infer the effect of the complex interactions in the speciation and extinction process.

### 2019-03-11 manuscript/introduction.tex
minor edits across everything, added ideas to discussion for Boris

@@
 
-Whole genome duplication (WGD) is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
+Polyploidy, the acquisition of extra sets of chromosomes, is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
+%R- Polyploid people are careful at not using interchangeably WGD so I switched back to polyploidy terminology. Since we are not using anything molecular I'd rather keep it as polyploidy
 The widespread prevalence of variation in ploidy has long been considered a salient feature of flowering plant lineages \citep{stebbins1938}. 
-WGDs have the potential to affect many phentoypes and a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
+Polyploidy events have the potential to affect many phentoypes and a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
 Polyploids have been thought to have lower rates of net diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
-Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that WGDs have played an important macroevolutionary role driving innovation and diversification in plants.
+Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that polyploidy events have played an important macroevolutionary role driving innovation and diversification in plants.
 Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies the pervasiveness of diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
@@
 % E: so try again -- okay? 
+%R- yup- good resolution
 This could affect interpretation of different net diversification rates for diploids and polyploids, especially if diploids have a higher net diversification rate if they were ancestrally polyploid than if they had no polyploid ancestry.
@@
 Second, \citet{beaulieu_2016}  proposed an extension of the *SSE framework by introducing a model with unspecified, or `hidden' traits (HiSSE model), which can affect the diversification process.
-Because variation in speciation and extinction rates may not be due solely to the focal trait, HiSSE simultaneously enables hidden states to explain such variation, and it reduces the elevated false-positive error rates \citep{rabosky_2017}.
+Because variation in speciation and extinction rates may not be due solely to the focal trait, HiSSE simultaneously enables hidden states to explain such variation, and it reduces the elevated false-positive error rates.
 %B: This may be a good portion of Discussion--faults and qualities of methods?
@@
 Our results highlight the importance of considering non-additive effects of traits on net diversification rates, especially when there are strong biologically driven correlations among them.
+
+% R- I'm good with this intro. It is simple and reads well.

### 2019-03-11 manuscript/methods.tex
minor edits across everything, added ideas to discussion for Boris

@@
 % R-This is very important but I have removed it because I think it is a full on paper by its own
-Including this many more species would have prohibitively slowed our analyses, especially those implementing the most complex models.
+%Including this many more species would have prohibitively slowed our analyses, especially those implementing the most complex models.
 
@@
 The model parameters are speciation rates ($\lambda_D$, $\lambda_P$) and extinction rates ($\mu_D$, $\mu_P$) for each state, and a rate of transition from D to P ($\rho$).
+This is the same model fitted by \citet{mayrose_2011} but applied to the larger Solanaceae dataset.
+% R- Itay wanted to hightlight that is the same model. Definitely a BiSSE. He did the ClaSSE later in the 2014 response.
 
@@
 A fully parameterized version of this model would have 26 rate parameters. 
-Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fit a simplified version with 15 parameters.
-The reduction in parameter space was achieved by fixing the rates for transitions among hidden states to be equal with rate $\alpha$, and fixing the transition rates between observed states to be independent of the hidden state (rates $\rho_I,\ \rho_C,\ \delta,\ q_{IC}$ as defined for the ID/CD/CP model).
-There are additionally twelve diversification rate parameters ($\lambda_{ID_A},\ \lambda_{ID_B},\ \lambda_{CD_A},\ \lambda_{CD_B},\ \lambda_{CP_A},\ \lambda_{CP_B},\ \mu_{ID_A},\ \mu_{ID_B},\ \mu_{CD_A},\ \mu_{CD_B},\ \mu_{CP_A},\ \mu_{CP_B}$).
+Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fit a simplified version with 16 parameters.
+% You were missing counting the alpha
+The reduction in parameter space was achieved by fixing the rates for transitions among hidden states to be equal with rate $\alpha$, and fixing the transition rates between observed states to be independent of the hidden state (rates $\rho_I,\rho_C,\ q_{IC}$ as defined for the ID/CD/CP model).
+There are additionally twelve diversification rate parameters ($\lambda_{ID_A},  \lambda_{ID_B},  \lambda_{CD_A}, \lambda_{CD_B}, \lambda_{CP_A}, \lambda_{CP_B}, \mu_{ID_A}, \mu_{ID_B}, \mu_{CD_A}, \mu_{CD_B},  \mu_{CP_A}, \mu_{CP_B}$).
 %B: I can't tell, but it seems to me that this 26->16 parameter reduction makes it possible to underestimate the effect of these hidden states.Perhaps this complication is covered in the Discussion section.
@@
 These types of classifications do not allow for a ploidy reversion.
-Where indicated, the classification of ploidy for the data used in our models was based on chromosome multiplicity at the genus level (see Supplementary Information).
+Where indicated, the classification of ploidy for the data used in our models was based on chromosome multiplicity at the genus level (see \url{https://github.com/roszenil/solploidy} for archived dataset).
 % (data available at https://www.github.com/roszenil/solploidy/datasets).
@@
 %R: Boris in the introduction mentions this possible causal relationship would that be equivalent to say hierarchical?
-Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $BF(M_0,\ M_1) = \ln[ P(\mathbf{X} | M_0) - P(\mathbf{X} | M_1]$.
+Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $BF(M_0,\ M_1) = \ln[ P(\mathbf{X} | M_0) - P(\mathbf{X} | M_1)]$.
 There is `positive' support for $M_0$ when this value is more than 2, `strong' support when it is more than 6, and `very strong' support when it is more than 10 \citep{kass1995}.

### 2019-03-12 manuscript/discussion.tex
- backup of work on manuscript--all sections; extensive comment deletions and edits.

@@
 \section{Discussion}
-
-% outline
-% - disentangling effects
-% - general value of approach
-%  		another kind of confounding processes, not only transition and diversification, but "other diversifications" and "other transitions", esp. correlated
-%
-% - our study key findings and details
-% - effects of Ploidy (as %R)
-% - trumped by BS
-% - no difference on background of SC 
-% - discuss data/models and their long/shortcomings: 
-%	ClaSSE, will's idea for stochastic mapping
-%	irreversibility SI -> SC assumption
-% - how to improve / what next
-%  (may be useful to study in SC families)
-%
-% - pathways
-% 
-%- diploidization
-% - rate, effect
-% - history in solanaceae
-%
-% - summary
 
@@
 We investigated the individual and joint effects of ploidy and breeding system states in diversification of Solanaceae.
-Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are trumped by hidden states and or/ breeding system when added to models.
-% E: is "trumped" too much like gloating?
-- R-Haha
-%R- I added hidden states because that hasn't been discussed ever in polyploidy
+Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are overshadowed by hidden traits, or breeding system when added to models.
 Interestingly, the difference in diversification rates between the diploids and polyploids is negligible on the background of self-compatibility.
-%B: should we mention here that the hidden states presage these results in each case?
-Maybe we don't want to make too much of that, but it is fairly interesting.
-Also possibly hit that in your paragraph(s), below.
-In the same vein, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against diploid genomic background.
-%B: "in diploids" may work better?
-%B: need help with pathway single-sentence summary
-Motivated by the widespread findings of paleopolyploid ancestry across angiosperms, we also examined support for diploidization in the history of the family, which appears to be weak.
+In the same vein, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against the diploid genomic background.
+%Inference of transition rates can appreciably across 
+Motivated by the widespread findings of paleopolyploid ancestry across angiosperms, we also examined and found modest support for diploidization in the history of the family.
 %B: "mixed" "not strong" what's the right word here?
 E: not "weak", maybe "not strong" or "not nearly as strong as anything else"
-Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests of complex hypotheses regarding a heterogenous diversification process.
-Below, we discuss these key findings and outline a broad approach for application of phylogenetic comparative methods in the context of interactions among traits and diversification process.
+Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests of intricate hypotheses regarding a heterogenous diversification process.
+Below, we discuss these key findings and detail their sometimes confusing and complex implications.
+We additionally provide an outline for broader application of phylogenetic comparative methods in the context of interactions among traits and diversification process.
 
@@
 Since MuSSE might also suffer from hight type I errors considering more heterogeneity using a hidden state model on top of the multistate SSE is necessary to adequately infer the effect of the complex interactions in the speciation and extinction process.
+
+ %%%%%%%%%%%
+ %% Notes to Self %%
+ %%%%%%%%%%%
+
+% outline
+% - disentangling effects
+% - general value of approach
+%  		another kind of confounding processes, not only transition and diversification, but "other diversifications" and "other transitions", esp. correlated
+%
+% - our study key findings and details
+% - effects of Ploidy (as %R)
+% - trumped by BS
+% - no difference on background of SC 
+% - discuss data/models and their long/shortcomings: 
+%	ClaSSE, will's idea for stochastic mapping
+%	irreversibility SI -> SC assumption
+% - how to improve / what next
+%  (may be useful to study in SC families)
+%
+% - pathways
+% 
+%- diploidization
+% - rate, effect
+% - history in solanaceae
+%
+% - summary
+%
+% B: NEW REORG COMMENTS:
+
+% First, include all straightforward Results and related observations
+%
+% Then try to square the oval of history of DP and IC.
+Specifically, it is difficult to know what to make of the delta models.
+Their inclusion is 'positively' supported, but yields results that are hard to reconcile with other data.
+This is one of the more interesting tensions of the paper, in the sense that it binds all of the discussion point and something here has to come to a head.
+%Either what we think we know about SI/SC and reversals is wrong, or something about diploidization inference is wrong.
+Why?
+It is hard to square that we don't see IP state, almost ever!, but it's supposed to be the ancestral state, and not just in Solanaceae, across eudicots!
+Doesn't demand a full treatment here, but it's worth pointing out that if the SI/SC inference is wrong, then we must include reversals, qCI, here and elsewhere and it will change all kinds of inferences.
+Then, we'll have to explain maintenance of trans-generic polymorphism, ha, ha!
+Otherwise, something is off with the inference delta >> rho in some cases.
+If so, we'll have to explain paleo-WGD inference, etc.
+Either way, we fundamentally misunderstand the evolution of one of these two traits.
+
+% including delta makes for some nonsense: ancestral states are all P with rampant recent transitions to D (e.g.
+Fig.
+S4/S8.
+Not backed by any data, including synteny, S-allele trans-polymorphism, or anecdotally bearing experimental data about rates of spontaneous polyploidization and diploidization \citep{husband_2012}.
+
+% PATHWAYS -- mutational opportunity vs lineage transition rate reveals something about selection 
+%
+% Starting with SI-D, selection favoring establishment of SC-D seems more common than that of SC-P.
+% This is because the mutation rates are likely higher to SC-P, but rate 3x lower. "
+%Husband2012: The prevalence of polyploidy within plants has led to debate about the influence of genome duplication on net species diversification (Otto and Whitton 2000).
+Researchers have argued, on one hand, that polyploidy may simply accumulate as a unidirectional mutation with little influence on the rate or trajectory of evolutionary divergence (Stebbins 1971; Meyers and Levin 2006).
+This idea is based on the premise that ploidy is more likely to increase than decrease in time.
+Although reversals do occur, the resulting haploids are rela- tively rare and often of reduced fitness and high sterility (Stebbins 1980; Ramsey and Schemske 2002).
+Alternatively, polyploidy itself may enhance evolutionary diversification leading to increased divergence and species richness (Otto and Whitton 2000).
+Elevated diversification may occur if polyploidy accelerates adaptive divergence and evolution of reproductive isolation (Otto and Whitton 2000) or reduces species extinction.
+%The presence of intraspecific ploidy diversity has three main causes.
+First, genome duplication is an ongoing genomic mutation within many natural populations and ploidy polymorphism is therefore expected when poly- ploid formation is common.
+In fact, Ramsey and Schemske (1998) estimated that polyploids form at the rate of 10 5 per individual per generation, well above the genic mutation rate, and in some cases it may be even higher than that (Husband 2004).
+Second, depending on the frequency and pathways of polyploid formation, diploids and polyploids may be maintained as genetically homogeneous populations, which are justifiably treated as single taxonomic species.
+Finally, ploidy polymorphism may arise as an artefact of our imperfect taxonomic approach.
+Since morphological divergence is often used as a criterion for species delinea- tion, polyploid lineages may not be recognized at the species level, especially in autopolyploids, which often have a close resemblance to their diploid progenitors (Soltis et al.
+2007).

### 2019-03-12 manuscript/introduction.tex
- backup of work on manuscript--all sections; extensive comment deletions and edits.

--- 2019-03-11 manuscript/introduction.tex
@@
 
-%NOTE: please separate each line with a line break!
-
-%B->%E: instead of a major reorganization, I added a breeding system quote from Barrett that presages our results. How's that? It seemed daunting to reorganize ploidy-first to BS-first.  E: nice!
-
 Species accumulate across the tree of life at different rates. 
@@
 Dramatic increases in the available phylogenetic and phenotypic data, as well as methodological advances, have greatly accelerated the search for traits that influence diversification traits.
-Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a difficult statistical problem \citep[\eg][]{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}.
-%is it only a "statistical" problem?
-In part, the difficulties arise because speciation and extinction may not depend on a single trait, and the context in which traits occur can lead to complex interactions resulting in a heterogenous speciation and extinction rates \citep{beaulieu_2016, caetano_2018, herrera_2018}.
+Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a difficult problem \citep[\eg][]{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}.
+Problems arise at least in part because speciation and extinction may not depend on a single trait, and the context in which traits occur can lead to complex interactions resulting in a heterogenous speciation and extinction rates \citep{beaulieu_2016, caetano_2018, herrera_2018}.
 Consequently, examining the association of only one trait with diversification patterns can be misleading.
 
-Whole genome duplication (WGD) is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
+Polyploidization is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
 The widespread prevalence of variation in ploidy has long been considered a salient feature of flowering plant lineages \citep{stebbins1938}. 
-WGDs have the potential to affect many phentoypes and a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
-Polyploids have been thought to have lower rates of net diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
-Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that WGDs have played an important macroevolutionary role driving innovation and diversification in plants.
-Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies the pervasiveness of diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
-% Testing for the presence of diploidization in the context of diversification is important because the presence of a positive net diversification rate for diploids can be the result of an ancestrally polyploid lineage.
-% Under this scenario, polyploidy would no longer be considered to reliably slow down diversification, but instead, polyploidy would be the initiator of the diversification process.
-%B: I think this should read "...polyploidy is no longer be considered an evolutionary dead end."  %R- This last phrase is from Itay and I rather like it.
-<-- %B: Hmm.
-But how would it be considered an "initiator of the diversification process?" It seems to be associated with lower diversification.
-In other words, I also agree with both %R and Itay, but not with what the current phrasing is asserting.
-% E: I wrote the two sentences below to try to reconcile the text and comments below.  But then I realized we don't actually test this hypothesis.
-% One hypothesis is that diploids have a higher net diversification rate if they were ancestrally polyploid than if they had no polyploid ancestry.
-% Testing this hypothesis requires incorporating diploidization in analyses of diversification.
-% E: so try again -- okay?
-This could affect interpretation of different net diversification rates for diploids and polyploids, especially if diploids have a higher net diversification rate if they were ancestrally polyploid than if they had no polyploid ancestry.
+Increase in ploidy can alter many traits and impact a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
+%
+Polyploids are thought to have lower diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
+Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications have played an important macroevolutionary role driving innovation and diversification in plants.
+Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies pervasiveness of diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
+This could affect the interpretation of different net diversification rates for diploids and polyploids, especially if diploids have a relatively higher net diversification rate if they were ancestrally polyploid than if they had no polyploid ancestry.
 
@@
 In particular, self-incompatibility (SI) causes plants to reject their own pollen, and loss of self-incompatibility, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974,igic_2008}.
-% Gordon: Really?
-I find this hard to believe...
-%B:  sigh.
 Previous analyses reported higher rates of diversification for SI than for SC lineages in Solanaceae \citep{goldberg_2010}. 
 Similarly, heterostylous SI lineages in Primulaceae seem to diversify faster \citep{devos2014}, as do outcrossing lineages in Onagraceae \citep{freyman_2018}.
-Despite consistent inferences of their macroevolutionary role, breeding systems are clearly not the sole trait determining diversification rates.
+Despite consistent inferences of their macroevolutionary role, breeding systems are clearly not the sole trait determining diversification rates, and may exhibit complex interactions with other traits.
 
-Ploidy, for example, is widely thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
-%B: I did not find this confusing, but it worries me that Gordon did. Consequently, I've ejected this sentence, now focusing on SI/SC only.
-%The evidence for correlated shifts in mating system, from outcrossing to selfing, following WGD, appears limited and sometimes contradictory \citep{barringer2007, barrett2008, husband2008}.
-Polyploidy may facilitate the transition to selfing \citep{barringer2007, barrett2008, husband2008}, or it may directly cause the loss of SI \citep{stout1942, lewis1947}.
-% Polyploidy is not a mere correlate of transition in breeding system, but may directly cause the loss of SI \citep{stout1942, lewis1947}, as well as facilitate the transition to selfing \citep{barringer2007, barrett2008, husband2008}.
-E: reworded because no one would think the only thing P does is follow BS?
+Ploidy, for instance, is widely thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
+Whole genome duplication may help facilitate the loss of SI and the transition to selfing \citep{barringer2007, barrett2008, husband2008}, or it may directly cause the loss of SI \citep{stout1942, lewis1947}.
 Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic SI systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
@@
 This system, RNase-based SI, is expressed in all SI species of Solanaceae examined to date.
-There is thus a strong correlation between polyploidy and SC in this family \citep{robertson_2011}.
-% Gordon comments: Here again is confused about breeding systems in general vs. SI genetic systems.
-%B: not confused here; some SI is broken down by ploidy. Therefore, there is a correlation with coefficient between 0 and 1, and not 1.0 
-% Gordon comments here: First, it's a little confusing if you are talking about more SC or SI species more generally, or if you are talking specifically about genetic SI systems.
-I don't think the de Vos paper is dealing with SI systems.
-If your'e talking more generally about breeding system evolution there's a vast literature here, and I think  this is oversimplifying.
-% R- Boris what do you think is the best way to proceed here?
-Narrow it down to SI systems (deleting de Vos paper) or expanding it to talk about breeding systems in general and then saying that we will only focus on SI systems because of Solanaceae
-%B: The apparent problem for Gordon is that devos2014 estimates *SSE parameters for heterostyly and homostyly, and not homomorphic SI.
-That's not a concern, I think, they are all "SI systems." And Dioecy etc.
-is not classically considered a "breeding system", but a "sexual system".
-That may be a part of the confusion.
-It's really not Gordon's problem, and we could do something to yet again re-define these terms?
-Emma?
-% E: I vote for using terms correctly without providing a glossary---lead by example?
+There is consequently a strong correlation between polyploidy and SC in this family \citep{robertson_2011}.
 
-Given that changes in ploidy and breeding systems are causally related, and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
-First, it may be of interest to establish their joint influence on speciation and extinction rates.
+%B: removed First, ... Second, ... because we had three consecutive counting paragraphs.
+Given that changes in ploidy and breeding systems are causally related, and may have profound effects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
 If the effects of ploidy and breeding system on lineage diversification are not additive, studying each trait separately may not reveal the effects of their combinations.
-Second, the two traits may affect each other's state transitions.
+As a result, it may be of interest to establish their joint influence on speciation and extinction rates.
+The two traits may also affect each other's state transitions.
 For example, losses of SI in diploids could be caused by polyploidization, resulting in a one-step pathway to polyploid SC species. 
@@
 
+%B: removed First, ... Second, ... because we had three consecutive counting paragraphs.
 Two recent methodological developments enable disentangling the complex interactions between two or more traits and diversification.
-First, \citet{fitzjohn_2012} extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007} to allow estimating the effect of an arbitrary number of states (enabling multi-state, MuSSE, instead of binary state, BiSSE, speciation and extinction models).
+\citet{fitzjohn_2012} extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007} to allow estimating the effect of an arbitrary number of states (enabling multi-state, MuSSE, instead of binary state, BiSSE, speciation and extinction models).
 This allows multiple states to represent trait combinations, particularly useful if one is interested in explicitly testing whether combinations of traits non-additively affect speciation or extinction.
-Second, \citet{beaulieu_2016}  proposed an extension of the *SSE framework by introducing a model with unspecified, or `hidden' traits (HiSSE model), which can affect the diversification process.
-Because variation in speciation and extinction rates may not be due solely to the focal trait, HiSSE simultaneously enables hidden states to explain such variation, and it reduces the elevated false-positive error rates \citep{rabosky_2017}.
-%B: This may be a good portion of Discussion--faults and qualities of methods?
-%HiSSE models have effectively shown that some traits associated with diversification in BiSSE analyses are actually not associated with the diversification process, and instead a hidden state is linked to differences in the diversification rate.
-%Such results raise the question of the identity of the hidden state, or whether it is an approximation of some unknowable heterogeneity, as well as whether interactions between the known and hidden trait were modeled appropriately.
-%Third, one can simultaneously consider the effects of multiple known traits using MuSSE, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
-%However, MuSSE might also suffer from hight type I errors if there are other processes shaping diversification other than the traits observed \citep{caetano_2018}
-%Therefore, a fourth option exists: considering both the presence of multiple known traits  linked to the diversification process, but also, some unknown or hidden state changing speciation and extinction rates.
+Soon thereafter, \citet{beaulieu_2016} proposed an extension of the *SSE framework by introducing a model with unspecified, or `hidden' traits (HiSSE model), which can affect the diversification process.
+Because variation in speciation and extinction rates may not be solely due to the focal trait, HiSSE simultaneously enables hidden states to explain such variation, and it reduces the elevated false-positive error rates.
 A joint application of these advances allows simultaneous consideration of the effects of multiple known traits on the diversification process, along with the presence of other unobserved states that can affect speciation and extinction rates.
-%These four modeling strategies are proposed here in the context of studying complex interactions among traits, linked diversification.
 
@@
 With this joint trait model, we also quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
-For all of the ploidy models, we investigate the potential effects of including diploidization.
+For all of the models involving ploidy, we investigate the potential effects of including diploidization.
 Our results highlight the importance of considering non-additive effects of traits on net diversification rates, especially when there are strong biologically driven correlations among them.

### 2019-03-12 manuscript/main.tex
- backup of work on manuscript--all sections; extensive comment deletions and edits.

@@
 
-% New comparative approaches for the study of breeding system, polyploidy, and lineage diversification % E: the tentative title given to New Phyt
-%
-% Comparative study of breeding system, ploidy, and lineage diversification % E: too boring?
+% New comparative approaches for the study of breeding system, polyploidy, and lineage diversification % E: the tentative title given to New Phytologist
 %
@@
 %--------------------------------------------------
-% Main text
+% Main Text
 %--------------------------------------------------
@@
 
-This work was supported by the National Science Foundation (DEB-1655692 to BI, and DEB-1655478 to EEG) and
-the Minnesota Supercomputing Institute (MSI) at the University of Minnesota, which provided computing resources for this project.
+This work was supported by the National Science Foundation (NSF DEB-1655478 to EEG and NSF DEB-1655692 to BI).
+The computing resources were provided by the Minnesota Supercomputing Institute (MSI) at the University of Minnesota.
 %BI would like to thank SCHB for inspiration, sparring, and encouragement; maybe EEG, as well?

### 2019-03-12 manuscript/methods.tex
- backup of work on manuscript--all sections; extensive comment deletions and edits.

@@
 Each species retained on the tree belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known.
-The number of species in each is shown under the corresponding circles in the top row. % FIXME: Rosana, what are the numbers for CP and ?P ?
+The number of species in each is shown under the corresponding circles in the top row.
 These categories were then grouped in a manner appropriate to the states of each model.
-For example, there are 34 species that are self-compatible and of unknown ploidy; these are coded as either $D$ or $P$ in the D/P models (uncertain, or consistent with either state), as $C$ in the I/C models, and as either $CD$ or $CP$ in the ID/CD/CP models.
-In all cases, species were coded as either $A$ or $B$ in the hidden state models.
-The models depicted in the different rows use state spaces that are not comparable with one another.
-For example, we cannot test whether the D/P model fits better than the I/C model because they use states that are not the same and are not `lumpable' \citep{tarasov_2019}.
-%B: Hmmm.
-The presentation is economical, true, and I think it is retrospectively clear, and valuable, now that I've been thinking about the paper for a while.
-But it will be very difficult as a source of clarity for someone reading the paper the first time.
-The top graphical layout also implicitly communicates the idea that all "I" species must be "D", and that all "P" species must be also "C", which is not intuited easily.
-(Are they encoded like that?) Moreover, it's not made clear that the 'species' in question are, in fact, often polymorphic, which makes DPIC possible.
-The subsequent models require a bit of thought, which is not too bad, but the treatment of the state coding and specifically the sentence "In all cases, species were coded as either A or B in the hidden state models." is a bit terse for someone not familiar with HiSSE.
-% E: Yes, all I species are D and all P species are C (no SI polyploids).
+For example, there are 34 species that are self-compatible ($C$) with unknown ploidy, and 40 species known to be polyploid($P$), but without breeding system data; the latter can be inferred, because all polyploid species in the dataset are self-compatible (see \citealt{robertson_2011}, and text).
+%B: I added stuff here
+In the D/P models, these are coded as either $D$ or $P$ (uncertain/consistent with either state); in the I/C  models, they are coded as $C$; and in the ID/CD/CP models, they are coded as either $CD$ or $CP$.
+In the hidden state models, species were coded as either $A$ or $B$.
+The model-specific coding is shown in different rows. 
+%B: I think that these two sentences are more appropriate for the model comparison tables and Methods text. If we are admonished to cut text, they could go.
+They each use state spaces that are not directly comparable with one another.
+For example, we cannot test whether the D/P model fits better than the I/C model because they use states that are not the same, and are therefore not `lumpable' \citep{tarasov_2019}.
+%B: [This] will be very difficult as a source of clarity for someone reading the paper the first time.
+%E: Yes, all I species are D and all P species are C (no SI polyploids).
 I don't think there are polymorphic species because species with some SI and some SC pops are called I, and the most common ploidy level was used.

### 2019-03-12 manuscript/results.tex
removed some tex warnings
- backup of work on manuscript--all sections; extensive comment deletions and edits.

@@
 For the two models that only include diploid and polyploid states (D/P and D/P+A/B), the net diversification rate of polyploid lineages is likely negative when diploidization (parameter $\delta$) is excluded but positive when $\delta$ is included (\cref{figure:netdivall}A versus \cref{figure:netdivnodip}A).
-For the two models that also included breeding system (ID/CD/CP and ID/CD/CP+A/B), the main effect of including diploidization is greatly increasing the uncertainty of the estimate of polyploid net diversification rate (\cref{figure:netdivall}EF versus \cref{figure:netdivnodip}CD).
+For the two models that also included breeding system (ID/CD/CP and ID/CD/CP+A/B), the main effect of including diploidization is greatly increasing the uncertainty of the estimate of polyploid net diversification rate (\cref{figure:netdivall}E,F versus \cref{figure:netdivnodip}C,D).
 For all models, there is greater uncertainty in the estimate of diploidization rate than polyploidization rate, as judged by the width of the credibility intervals (see \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}).
-
 % R will add a bit about param confidence
@@
 
+%B: heads up: I added two clarifying descriptive sentences in the two figures below. Maybe something about noting x-axis scale differences would be useful.
 \begin{figure}
 \includegraphics[width=\textwidth]{allnetdivmodelswithoutdelta.pdf}
-\caption{Net diversification rates for all models that exclude diploidization.}
+\caption{Net diversification rates for all models without diploidization rate estimate. 
+Each panel contains a graphical summary of estimated model parameters and displays posterior distributions for net diversification estimates.}
 \label{figure:netdivall}
@@
 \includegraphics[width=\textwidth]{allnetdivmodelsdiploidization.pdf}
-\caption{Net diversification rates for all models that include diploidization.}  
+\caption{Net diversification rates for all models that include diploidization ($\delta$). 
+Each panel contains a graphical summary of estimated model parameters and displays posterior distributions for net diversification estimates.}  
 \label{figure:netdivnodip}
@@
 \caption{Ancestral state reconstruction showing the maximum a posteriori for each node of the D/P ploidy model} % XXX
-\label{suppfigure:DPnodip}
+\label{suppfigure:DPnodipasr}
 \end{suppfigure}

### 2019-03-12 manuscript/discussion.tex
- backups of all the changes -- mainly to discussion

@@
 Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests of intricate hypotheses regarding a heterogenous diversification process.
-Below, we discuss these key findings and detail their sometimes confusing and complex implications.
+Below, we discuss these key findings and highlight the most interesting and paradoxical implications.
 We additionally provide an outline for broader application of phylogenetic comparative methods in the context of interactions among traits and diversification process.
+
+\subsection*{Disentangling trait effects on diversification}
+
+% confounding effects of other traits
 
@@
 Disentangling the patterns of diversification linked to  multiple traits allowed us the opportunity to model and test complex hypotheses about the heterogenous diversification process in Solanaceae.
-Species are created and go extinct based on multiple and often highly correlated phenotypes. However, estimating the speciation and extinction rates remains challenging.
-
+Species are created and go extinct based on multiple and often highly correlated phenotypes. 
+However, estimating the speciation and extinction rates remains challenging. 
 By carefully aggregating and curating ploidy and breeding system data for Solanaceae, we showed how  the complex interaction between ploidy and breeding system  can lead to diversification conclusions that differ from the diversification conclusions of models that focus on one trait at a time.
@@
 
-% 3) I'll bet allowing cladogenetic transitions would make the one-step ID/CP even more dominant.
-%    I added the paragraph below regarding cladogenetic transitions, but it doesn't really fit with
-%    the rest of the discussion, so modify or discard as needed:
-% E: A small defense of excluding clado is that it didn't affect the diversification results in goldberg2010 vs goldberg2012
-Our work shows that a diploid Solanaceae lineage is much more likely to take a
-one-step ID/CP pathway to polyploidy rather than a two-step ID/CD/CP pathway. 
-We modeled both of these pathways as anagenetic character changes occurring within a species.
-However \citet{goldberg_2012} showed that I/C transitions are often associated with speciation events, and similarly \citet{freyman_2017} demonstrated that D/P transitions may also be associated with speciation events.
-Models that fail to consider transitions that occur at speciation (cladogenetic changes) may
-make misleading estimates of anagenetic transition rates.
-The HiSSE-based models introduced in our work here could be extended to incorporate cladogenetic transitions. 
-It is possible that allowing for cladogenetic transitions may enable the one-step pathway
-to even further dominate over the two-step pathway to polyploidy, but this remains to be tested in future work.
+\subsection*{Pathways to polyploidy}
+
+% examine SC-only families
+% no causal correlation in SSI systems
+% paths to dioecy
+% path-dependent net-diversification
+
+We find that, in Solanaceae, a diploid lineage is much more likely to take a
+one-step  ID$\rightarrow$CP pathway to polyploidy rather than a two-step ID$\rightarrow$CD$\rightarrow$CP pathway.
+This result is significant because it has a number of implications for the inference of history of breeding and sexual systems in flowering plants.
+First, support for the direct pathway lends credence to the idea that breakdown of SI by WGD---by generation of diploid `heteroalleleic' pollen---may often trigger the evolution of gender dimorphism across angiosperms \citep{miller_2000,roberston_2011}.
+
+%missing deeper discussion of with/without SDD/SID estimates and comparison with Robertson et al. 2011.
+
+% qualifying paragraph for pathways
+% "This finding is not without a hitch..."
+We implicitly modeled trait changes in both of these pathways as anagenetic, transitions occurring within lineages, without associated speciation events.
+However, losses of SI (I$\rightarrow$C transitions) are often associated with cladogenesis in Solanaceae \citep{goldberg_2012}, and polyploidization (D$\rightarrow$P transition) also seem to co-occur with speciation events in a number of other families \citep{zhan_2016,freyman_2017}.
+Models that fail to consider such cladogenetic changes may make misleading estimates of anagenetic transition rates.
+In the only available data bearing on this problem, \citet{goldberg_2012} found that extending the *SSE framework to allow cladogenetic changes did not substantially affect inference of net diversification rates.
+Nevertheless, a distinct possibility remains that allowing for cladogenetic transitions could tip the scales further in favor of the one-step pathway to SC polyploidy (CP), but this remains to be tested in future work.
+Specifically, the multi/hidden state SSE-based framework we employed here could be extended to incorporate cladogenetic transitions.
+% We could also list %W+%R idea above, for stochastic mapping?
+Another useful line of inquiry could examine the path-dependent diversification rates, asking whether diversification rates of polyploid SC lineages differ depending on the path taken.
 
@@
 But discuss?
-% E: To elaborate: In the IC/CD/CP model, the CP state would be subdivided into two hidded values, corresponding to the influx from the two pathways.
+% E: To elaborate: In the IC/CD/CP model, the CP state would be subdivided into two hidden values, corresponding to the influx from the two pathways.
 Those two states, CP_X and CP_Y, could have different lambda and mu.
 
-Diploidization needs to be considered in diversification models because higher rates of net diversification (speciation minus extinction) in diploids can be obtained from models that ignore the possibility of a polyploid lineage that has diploidized being the diversification enhancer.
-Hence, in a diploidization lacking model, diploids will show higher rates of diversification when in fact it is a polyploidy event and its subsequent diploidization that generated higher speciation and less extinction.
-By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with stochastic models. 
-In the cases where ploidy level was assigned by us based chromosome multiplicity at the genus level, estimating diploidization  might be a potential issue due to ploidy misclassification.
+%W: 3) I'll bet allowing cladogenetic transitions would make the one-step ID/CP even more dominant.
+%    I added the paragraph below regarding cladogenetic transitions, but it doesn't really fit with
+%    the rest of the discussion, so modify or discard as needed:
+% E: A small defense of excluding clado is that it didn't affect the diversification results in goldberg2010 vs goldberg2012
 
-%B: v2.0
 \subsection*{Diploidization}
 
-Diploidization is widely considered to be relatively rare, compared with polyploidization, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history.
-%cite
+%R: By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with stochastic models. 
+%In the cases where ploidy level was assigned by us based chromosome multiplicity at the genus level, estimating diploidization  might be a potential issue due to ploidy misclassification.
+
+Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.
 As a consequence, it is possible that nearly all extant species classified as ``diploid" in our analyses, based on cytogenetic data, are possibly secondary diploids, having undergone both polyploidization and re-diploidization.
@@
 
+%B: incorporate ancestral states in absence of I/C consideration to rag on \delta
+%Natural and induced tetraploids from a variety of families, including Fabaceae, Onagraceae, Rosaceae, and Solanaceae, almost invariably lose gametophytic SI (reviewed in \citealt{ramsey_1998,stone_2002,mcclure_2009}).
 On the other hand, there is little evidence for occurrence of re-diploidization in the time frame we considered, within the nightshades.
@@
 At best, the species we consider are ancestrally diploid, having re-diploidized after a suspected WGD.
-% E: But then how is there SI??
+% E: But then how is there SI?! Agreed, I'll cover it.
 Second, studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
@@
 %     irreversibility SI -> SC assumption
+
+\bigskip
+
+\subsection*{Conclusions}
+% problems: this system has a big set of advantages - carefully curated (co-)occurrences of two traits, genetic basis, etc. and everything is still contradictory and opaque.
 

### 2019-03-12 manuscript/introduction.tex
- backups of all the changes -- mainly to discussion

@@
 Polyploids are thought to have lower diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
-Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications have played an important macroevolutionary role driving innovation and diversification in plants.
+Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications (WGDs) have played an important macroevolutionary role driving innovation and diversification in plants.
 Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies pervasiveness of diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.

### 2019-03-12 manuscript/results.tex
- backups of all the changes -- mainly to discussion

@@
 Determining the relative contribution of these pathways based on our estimated transition rates from the ID/CD/CP model, we find that the one-step pathway is more likely on short timescales and the two-step pathway is more likely on long timescales (\cref{figure:pathways}, left panels).
-When not much time has elapsed, the one-step pathway is more likely because only one event need happen.
+Beginning with a single SI diploid (ID) lineage, when not much time has elapsed, the one-step pathway is more likely because it only necessitates a single event to effect transition to SC polyploid state.
 When more time has elapsed, the two-step pathway is more likely because the rate of loss of SI within diploids, $q_{IC}$, is greater than the rate of polyploidization for SI species, $\rho_I$ (\cref{suppfigure:IDCDCPnodip}).
-% These conclusions are agree with those of \citet{robertson_2011}.
+These conclusions, however, ignore the changes in numbers of lineages in a particular state from speciation and extinction events.
+By analogy, frequent sinking of a middle stepping stone in a path would lower the value of such a path.
+We therefore examined the effects of diversification rate on relative importance of each pathway.
+%B: revisit this possibly terrible analogy.
 
-When rates of lineage diversification are incorporated, however, the conclusions change.
-Beginning again with a single ID lineage, it is now necessary to consider its possible proliferation in the ID state, and also in the CD state for the two-step pathway, and to track the progression of the full resulting clade to the CP state.
+Indeed, incorporation of lineage diversification rates changes conclusions regarding the relative importance of each pathway.
+Beginning again with a single ID lineage, it is now necessary to consider its possible proliferation in the ID state, and also in the CD state for the two-step pathway, as well as to track the progression of the full resulting clade to the CP state.
 In this situation, even over long timescales, the two-step pathway contributes less (\cref{figure:pathways}, right panels).
-The lower rate of net diversification in the CD state, relative to ID, means that relatively fewer lineages are available to complete the second step of the two-step pathway.
+The lower rate of net diversification in the CD state, relative to ID, means that relatively fewer lineages are available to complete the second step of the two-step pathway, ending in CP.
 

### 2019-03-13 manuscript/discussion.tex
- backup of mostly done first 2/3 Discussion sections - addition of refs and minor text changes in Results

@@
 qIC path different than rho path, or same as (1), so that I and C are less different on the background of D.
-We investigated the individual and joint effects of ploidy and breeding system states in diversification of Solanaceae.
+%
+Although species are composed of vast assemblages of variable traits, many traits are heritable, and they may affect the propensity of species to perish and multiply \citep{lewontin_1970}, examining such macroevolutionary effects of many traits is rare, and remains challenging.
+We investigated the influence of two traits, ploidy and breeding system states, in diversification of Solanaceae.
 Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are overshadowed by hidden traits, or breeding system when added to models.
+%R: added hidden trait.
 Interestingly, the difference in diversification rates between the diploids and polyploids is negligible on the background of self-compatibility.
 In the same vein, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against the diploid genomic background.
-%Inference of transition rates can appreciably across 
 Motivated by the widespread findings of paleopolyploid ancestry across angiosperms, we also examined and found modest support for diploidization in the history of the family.
-%B: "mixed" "not strong" what's the right word here?
+%B: how about "modest"?
+It's 3-ish, but "positive" seems like a really idiosyncratic usage to me, despite kass & raftery.
+was: %B: "mixed" "not strong" what's the right word here?
 E: not "weak", maybe "not strong" or "not nearly as strong as anything else"
 Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests of intricate hypotheses regarding a heterogenous diversification process.
-Below, we discuss these key findings and highlight the most interesting and paradoxical implications.
+Below, we discuss key findings and highlight the most interesting and paradoxical implications.
 We additionally provide an outline for broader application of phylogenetic comparative methods in the context of interactions among traits and diversification process.
 
-\subsection*{Disentangling trait effects on diversification}
+\subsection*{Disentangling trait effects on diversification} %B: Emma may not like "effects,"?
 
-% confounding effects of other traits
+Previous analyses of the effects of ploidy on diversification showed that diploids had greater net diversification rates than polyploids when exploring across multiple clades of the angiosperm phylogeny \citet{mayrose_2011, mayrose_2015}.
+We obtain a broadly similar outcome, when considering only the effects of ploidy, but the results change significantly with joint consideration of hidden states, and breeding systems. 
+The addition of a hidden trait results in blurring of differences in diversification rates between diploids and polyploids (Figure 2A). %B: fix fig ref
+Simultaneous consideration of breeding system goes a long way in explaining variation in the inferred net diversification rates among ploidy states.
+SI, which is itself associated with a high diversification rate, is the breeding system context in which many diploids occur.
+Among SC species, diploids and polyploids are associated with similar diversification rates (Figure 2F). %B: fix fig ref
+The best interpretation of these results is that some of the effects of a hidden trait seem to be dictated by  each of the two traits under consideration.
+The finding is unique, and implications may be of wide interest, because it evaluates the relative influence of these two exceptionally interesting traits in a comparative phylogenetic framework.
 
-% - value of approach
-Disentangling the patterns of diversification linked to  multiple traits allowed us the opportunity to model and test complex hypotheses about the heterogenous diversification process in Solanaceae.
-Species are created and go extinct based on multiple and often highly correlated phenotypes. 
-However, estimating the speciation and extinction rates remains challenging. 
-By carefully aggregating and curating ploidy and breeding system data for Solanaceae, we showed how  the complex interaction between ploidy and breeding system  can lead to diversification conclusions that differ from the diversification conclusions of models that focus on one trait at a time.
-
-% - what happens if PD only, Hidden states
-Previous analyses of polyploidy linked to diversification showed how diploids had greater net diversification rates than polyploids when exploring across multiple clades of the angiosperm phylogeny \citet{mayrose_2011, mayrose_2015}.
-Using the most complete dataset for ploidy in a phylogenetic tree for Solanaceae, we were able to replicate the results found by \citet{mayrose_2011} where polyploids have a slower net diversification compare to diploids.
-However, we expanded this study to accommodate  background heterogeneity in the diversification process.
-When adding heterogeneity we found that it was more likely that an unobserved trait linked to diploid state was the one leading the net diversification patterns, and that there were some ``second-class"  diploids that were not different from  polyploids  in diversification terms (Figure 2A).
-This result made us question whether there was a second trait linked to diploidy that could make a differential in the diversification progress.
-
-%     main finding about our traits
-For Solanaceae, looking into self-incompatibility was a logical step.
-Previous studies have shown that self-incompatible Solanaceae species have also higher rates of diversification compared to their self-compatible counterparts \citep{goldberg_2012}. 
-Again, in our sample when looking only into breeding system linked to the diversification process we were able to replicate the slower net diversification for self-compatible taxa (Figure 2C).
-When this model was expanded to incorporate hidden states we found again that some self-incompatible taxa have faster net diversification rates than self-compatible taxa.
-However, there is again a  different class of taxa where those differences are not as striking (Figure 2D).
-However, investigating both  both polyploidy and breeding system simultaneously for every species in our sample, allowed us to further investigate why some diploids were quantitatively different than polyploids, or why some self-incompatible taxa would be different in diversification terms from some self-compatible when a second class is not.
-In the three-state diversification model ID/CD/CP, we represented the complex interactions between ploidy and breeding system.
-We found that self-incompatible diploids have faster and positive rates than self-compatible diploids and polyploids, and that the difference between the  rates of net diversification of self-compatible diploids is not as large (Figure 2E) as first found by binary trait diversification models (Figure 2A).
-This result is key, it not only aligned with the net diversification results from the D/P+A/B model, where a `	`hidden-trait" seem to be dictating the diversification pattern but it also aligned with the hidden state model I/C+A/B where some class of self-compatible net diversification rate might be overlapping with two net diversification rates of self-incompatible.
-% Removed by R. Therefore, we consider that finding a heterogenous result in the hidden trait approaches should be be treated as evidence of a second trait that is necessary to consider.
-% Removed by R.
-Pursuing knowledge of  such trait can result on a clearer picture of the importance of trait linked diversification patterns, but also on a better reconstruction on past events in phylogenies.
+Previous work has shown that SI species in Solanaceae are associated with higher diversification rates compared to their SC counterparts \citep{goldberg_2010}. 
+When we inferred only the role of breeding system in the diversification process, we were able to replicate the slower net diversification for SC (Figure 2C).  %B: fix fig ref
+Incorporating hidden states again resulted in SI associated with greater net diversification rates than SC in both hidden states (Figure 2D). 
+However, the overall picture is a bit more complex.
+Depending on the background hidden state, it is possible that a class of SC species may experience higher diversification rates than SI species.
+Moreover, these SC lineages may not constitute a dead end (with negative diversification rates).
 
@@
 
-We find that, in Solanaceae, a diploid lineage is much more likely to take a
+% qualifying paragraph for pathways
+We find that, in Solanaceae, SI diploid lineages are much more likely to take a
 one-step  ID$\rightarrow$CP pathway to polyploidy rather than a two-step ID$\rightarrow$CD$\rightarrow$CP pathway.
 This result is significant because it has a number of implications for the inference of history of breeding and sexual systems in flowering plants.
-First, support for the direct pathway lends credence to the idea that breakdown of SI by WGD---by generation of diploid `heteroalleleic' pollen---may often trigger the evolution of gender dimorphism across angiosperms \citep{miller_2000,roberston_2011}.
-
-%missing deeper discussion of with/without SDD/SID estimates and comparison with Robertson et al. 2011.
+First, support for the direct pathway lends credence to the idea that breakdown of SI by WGD---by generation of diploid `heteroalleleic' pollen---may often trigger the evolution of gender dimorphism across angiosperms \citep{miller_2000}.
+A more direct and definitive test of this hypothesis would additionally examine the propensity of polyploids generated through either pathway to become dioecious \citep{robertson_2011}.
+% -- mutational opportunity vs lineage transition rate reveals something about selection 
+Second, statistical phylogenetic approaches measure lineage transition rates.
+Transitions depend on a combination of availability of individual mutation rates, which generate trait variation, and selection, which may cause disproportionate loss or fixation of these mutants.
+Estimates of mutation rates are highly uncertain, but the total rate of breakdown of SI within diploids ranges is estimated to be on the order of  $\mu_{IC} = 10^{-5}$ per pollen grain, and includes breakdown by tetraploidization \citep{lewis1979}, which is itself estimated to occur on the order of $\mu_{DP} = 10^{-5}$ \citep{ramsey_1998}.
+Seemingly, then, the simple genic mutation rate that leads to loss of SI is at best equal to tetraploidization mutation rate, and possibly far lower.
+We nevertheless recover transition rates from ID to CD ($q_{IC}$) threefold greater than $\rho_{I}$, indicating that selection restricts the fixation rate of new polyploids compared with new SC mutants \citep{robertson_2011}.
 
 % qualifying paragraph for pathways
-% "This finding is not without a hitch..."
 We implicitly modeled trait changes in both of these pathways as anagenetic, transitions occurring within lineages, without associated speciation events.
@@
 Nevertheless, a distinct possibility remains that allowing for cladogenetic transitions could tip the scales further in favor of the one-step pathway to SC polyploidy (CP), but this remains to be tested in future work.
-Specifically, the multi/hidden state SSE-based framework we employed here could be extended to incorporate cladogenetic transitions.
-% We could also list %W+%R idea above, for stochastic mapping?
-Another useful line of inquiry could examine the path-dependent diversification rates, asking whether diversification rates of polyploid SC lineages differ depending on the path taken.
+%B Will's comment alluded to, but not directly.
+Specifically, the multi/hidden state SSE-based framework we employed here could be extended to incorporate cladogenetic transitions. %B: We could also list %W+%R idea above, for stochastic mapping?
+Another useful line of inquiry would examine the path-dependent diversification rates, asking whether diversification rates of polyploid SC lineages differ depending on whether they were generated via a one- or two-step pathway.
+%B: could elaborate if you see necessary.
+TODO
+%B: also missing deeper discussion of with/without SDD/SID estimates and comparison with Robertson et al. 2011.
 
-% And, are diversification rates of polyploid SC lineages different if one or other path was taken?
-\citet{charlesworth1985} % E: This is a great question.
-We don't answer it with this round of analyses.
-But discuss?
+%B: And, are diversification rates of polyploid SC lineages different if one or other path was taken? % E: This is a great question.  We don't answer it with this round of analyses.  But discuss?
 % E: To elaborate: In the IC/CD/CP model, the CP state would be subdivided into two hidden values, corresponding to the influx from the two pathways.
 Those two states, CP_X and CP_Y, could have different lambda and mu.
-
+%
 %W: 3) I'll bet allowing cladogenetic transitions would make the one-step ID/CP even more dominant.
@@
 
-%R: By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with stochastic models. 
+%B: TODO: add below to future goals.
+%R: and I: polyploid lineage that has re-diploidized may enhance diversification.
+%R: By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with comparative phylogenetic models.
 %In the cases where ploidy level was assigned by us based chromosome multiplicity at the genus level, estimating diploidization  might be a potential issue due to ploidy misclassification.
@@
 The age of the peak of paralog Ks distances, is approximately 71 million years \citep{tomato2012}. 
-If this is the case, then all of the genomes may have been subsequently re-diploidized, yielding the common extant base chromosome numbers in this and related families, n=11-12, considered diploid \citep{robertson_2011}.
+If this is the case, then all of the genomes may have been subsequently re-diploidized, yielding the common extant base chromosome numbers in this and related families, n=11-12, presently considered to be diploid \citep{robertson_2011}.
 %Not unlikely, because Arabidopsis n=5.
@@
 This idea is based on the premise that ploidy is more likely to increase than decrease in time.
-Although reversals do occur, the resulting haploids are rela- tively rare and often of reduced fitness and high sterility (Stebbins 1980; Ramsey and Schemske 2002).
+Although reversals do occur, the resulting haploids are relaively rare and often of reduced fitness and high sterility (Stebbins 1980; Ramsey and Schemske 2002).
 Alternatively, polyploidy itself may enhance evolutionary diversification leading to increased divergence and species richness (Otto and Whitton 2000).

### 2019-03-13 manuscript/results.tex
- backup of mostly done first 2/3 Discussion sections - addition of refs and minor text changes in Results

@@
 
-Indeed, incorporation of lineage diversification rates changes conclusions regarding the relative importance of each pathway.
-Beginning again with a single ID lineage, it is now necessary to consider its possible proliferation in the ID state, and also in the CD state for the two-step pathway, as well as to track the progression of the full resulting clade to the CP state.
-In this situation, even over long timescales, the two-step pathway contributes less (\cref{figure:pathways}, right panels).
+Indeed, incorporation of lineage diversification rates changes conclusions regarding the relative contribution of each pathway.
+Beginning with a single ID lineage, assessment of relative pathway importance necessitates consideration of proliferation in the ID state, as well as the CD state for the two-step pathway, and to track the transitions to the CP state.
+In this situation, even over long timescales, the two-step pathway contributes less to generation of CP (\cref{figure:pathways}, right panels).
 The lower rate of net diversification in the CD state, relative to ID, means that relatively fewer lineages are available to complete the second step of the two-step pathway, ending in CP.

### 2019-03-13 manuscript/discussion.tex
Addressing some polyploidy and diploidization wording

@@
 Interestingly, the difference in diversification rates between the diploids and polyploids is negligible on the background of self-compatibility.
-In the same vein, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against the diploid genomic background.
-Motivated by the widespread findings of paleopolyploid ancestry across angiosperms, we also examined and found modest support for diploidization in the history of the family.
+Furthermore, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against the diploid genomic background.
+Motivated by the widespread findings of polyploidy events in the ancestry of angiosperms, we also examined and found modest support for diploidization in the history of the family.
 %B: how about "modest"?
@@
 E: not "weak", maybe "not strong" or "not nearly as strong as anything else"
+%R- I think modest is a good way to express it because it is the best we can do with chromosome numbers.
 Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests of intricate hypotheses regarding a heterogenous diversification process.
@@
 
-\subsection*{Disentangling trait effects on diversification} %B: Emma may not like "effects,"?
+\subsection*{Disentangling trait interactions in a diversification context} %B: Emma may not like "effects,"?
+% R- changed the section name to be more accurate
 
 Previous analyses of the effects of ploidy on diversification showed that diploids had greater net diversification rates than polyploids when exploring across multiple clades of the angiosperm phylogeny \citet{mayrose_2011, mayrose_2015}.
-We obtain a broadly similar outcome, when considering only the effects of ploidy, but the results change significantly with joint consideration of hidden states, and breeding systems. 
-The addition of a hidden trait results in blurring of differences in diversification rates between diploids and polyploids (Figure 2A). %B: fix fig ref
-Simultaneous consideration of breeding system goes a long way in explaining variation in the inferred net diversification rates among ploidy states.
-SI, which is itself associated with a high diversification rate, is the breeding system context in which many diploids occur.
-Among SC species, diploids and polyploids are associated with similar diversification rates (Figure 2F). %B: fix fig ref
-The best interpretation of these results is that some of the effects of a hidden trait seem to be dictated by  each of the two traits under consideration.
+We obtain a broadly similar outcome, when considering only the effects of ploidy, but the results change significantly with joint consideration of hidden states, and breeding systems(\cref{figure:netdivall}A and B).
+The addition of a hidden trait results in blurring of differences in diversification rates between diploids and polyploids. %B: fix fig ref 
+%R- added the figure reference
+Simultaneous consideration of breeding system further explains the variation of the inferred net diversification rates among ploidy states.
+SI, which is itself associated with a faster diversification rate, is the breeding system context in which many diploids occur.
+Among SC species, diploids and polyploids are associated with similar diversification rates (\cref{figure:netdivall}E) %B: fix fig ref
+% R- Figure reference is fixed
+These results indicate that some of the effects of a hidden trait can be dictated by  each of the two traits under consideration.
+% R- reworded above not to
 The finding is unique, and implications may be of wide interest, because it evaluates the relative influence of these two exceptionally interesting traits in a comparative phylogenetic framework.
@@
 Previous work has shown that SI species in Solanaceae are associated with higher diversification rates compared to their SC counterparts \citep{goldberg_2010}. 
-When we inferred only the role of breeding system in the diversification process, we were able to replicate the slower net diversification for SC (Figure 2C).  %B: fix fig ref
-Incorporating hidden states again resulted in SI associated with greater net diversification rates than SC in both hidden states (Figure 2D). 
+When we inferred only the role of breeding system in the diversification process, we were able to replicate the slower net diversification for SC  (\cref{figure:netdivall}C) .  %B: fix fig ref
+Incorporating hidden states again resulted in SI associated with greater net diversification rates than SC in both hidden states  (\cref{figure:netdivall}D) . 
 However, the overall picture is a bit more complex.
@@
 %  E: Yes, good point.  Rosana, do you have results for state-independent diversification?  We could compare with using those rates.
+% R- I didn't fit it without diversification but I can and I can definitely do stochastic map of both (do you want it for the revision?)
 %
@@
 
-Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.
+%Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.
+% R- I commented out the rare statement above because most of the community does not agree with this statement anymore, diploidization is now considered the rule rather than the exception.
+Change it for the phrase below
+Flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.
+However, after a polyploidy event occurs genome reorganization, downsizing and loss has been observed \citep{dodsworth_2015, zenil_2016, mandakova_2018}.
 As a consequence, it is possible that nearly all extant species classified as ``diploid" in our analyses, based on cytogenetic data, are possibly secondary diploids, having undergone both polyploidization and re-diploidization.
@@
 
-Genomic evidence indicates that a paleo-polyploidization may have occurred prior to the origin of the family Solanaceae, rendering all current `diploids' secondarily derived. 
+Genomic evidence indicates that a polyploidy event may have occurred prior to the origin of the family Solanaceae, rendering all current `diploids' secondarily derived. 
 First, \citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \textit{Solanum lycopersicum}, may have experienced one or more WGDs.

### 2019-03-13 manuscript/abstract.tex
- up to pardox resolution and conclusion (Igic lab review version)

@@
 
+%B: I think it's fine.
 % E:  Help making this less bland?  Except for the first sentence, which may be too grandiose?
 If particular traits consistently affect rates of speciation and extinction, broad macroevolutionary patterns can be understood as consequences of selection at high levels of the biological hierarchy.
-Identifying traits associated with diversification rate differences is tricky, though, because of the many traits available to consider and the statistical challenge of testing for associations from phylogenetic data.
-Two traits that have been repeatedly suggested as drivers of differential diversification are whether a lineage is diploid or polyploid, and whether it is self-incompatible or self-compatible.
-We investigate the role of each of these traits, and their interaction, on speciation and extinction rates in Solanaceae.
+Identifying traits associated with diversification rate differences is tricky, however, because of the the possibly large selection of traits under consideration and the resulting statistical challenge of testing for associations from comparative phylogenetic data.
+%B: grammatical error in "traits ... are whether". Rephrased.
+%Two traits that have been repeatedly suggested as drivers of differential diversification are whether a lineage is diploid or polyploid, and whether it is self-incompatible or self-compatible.
+Ploidy (diploid vs. polyploid states) and breeding system (self-incompatible vs self-compatible states) have been repeatedly suggested as possibly drivers of differential diversification
+We investigate the role of these traits, including their interaction, on speciation and extinction rates in Solanaceae.
 We find that the effect of ploidy can largely be explained by its correlation with breeding system, and that additional unknown factors work with breeding system to determine diversification rates.

### 2019-03-13 manuscript/discussion.tex
- up to pardox resolution and conclusion (Igic lab review version)

--- 2019-03-13 manuscript/discussion.tex
@@
 In the same vein, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against the diploid genomic background.
-Motivated by the widespread findings of paleopolyploid ancestry across angiosperms, we also examined and found modest support for diploidization in the history of the family.
+Motivated by the widespread findings of paleopolyploid ancestry across angiosperms, we also examined and found modest support for diploidization in the history of the family, but without significant effects on inferences of diversification dynamics.
 %B: how about "modest"?
@@
 % qualifying paragraph for pathways
-We find that, in Solanaceae, SI diploid lineages are much more likely to take a
-one-step  ID$\rightarrow$CP pathway to polyploidy rather than a two-step ID$\rightarrow$CD$\rightarrow$CP pathway.
+We find that the SI diploid lineages are much more likely to take a
+one-step  ID$\rightarrow$CP pathway to polyploidy rather than a two-step ID$\rightarrow$CD$\rightarrow$CP pathway, in Solanaceae.
 This result is significant because it has a number of implications for the inference of history of breeding and sexual systems in flowering plants.
@@
 We nevertheless recover transition rates from ID to CD ($q_{IC}$) threefold greater than $\rho_{I}$, indicating that selection restricts the fixation rate of new polyploids compared with new SC mutants \citep{robertson_2011}.
+Overall, our findings suggest that the effect of ploidy is largely explained explained by its correlation with breeding system, and that other unknown factors work with breeding system to determine diversification rates.
 
@@
 As a consequence, it is possible that nearly all extant species classified as ``diploid" in our analyses, based on cytogenetic data, are possibly secondary diploids, having undergone both polyploidization and re-diploidization.
-Unduly ignoring secondary diploidization would clearly underestimate the rate of transitions from polyploids to diploids, and would likely cause inflated net diversification rates for diploids.
-% E: first clause seems like a tautology?
-second clause does not seem obvious, but could at least say that it could affect div rate estimates
-We find strong support in models that included diploidization \cref{table:bayesfactors}.
-However, estimating the rate at which it occurs based on ploidy level classifications is hard and highly uncertain (see parameter $\delta$ in \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}) and can change the pattern of ancestral reconstructions (\cref{suppfigure:DPnodipasr,suppfigure:DPasr,suppfigure:DPnodipABasr,suppfigure:DPABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPasr,suppfigure:IDCDCPnodipABasr,suppfigure:IDCDCPABasr}).
-%We find weak support for inclusion of diploidization. %cite fig and expand (FIXME).
-% R- I corrected the above phrase, we found strong support for the inclusion of diploidization, models that have it are always preferred over models that don't however that parameter is hard to estimate
+Unduly ignoring secondary diploidization would clearly underestimate the rate of transitions from polyploids to diploids, and would then likely cause inflated net diversification rates for diploids, because species considered in the diploid state, may have been ancestrally polyploid, instead.
+%B: This should be to mutual liking.
+IMO, the support is STRONG in the apparently meager analyses, one with only D/P), which yields highly unreasonable ancestral states.
+But it is decidedly not strong, i.e.
+"positive," in the apparently more informative joint trait analyses.
+We find strong support for including the diploidization rate parameter ($\delta$) in analyses that include only ploidy \cref{table:bayesfactors}. %B <- %Rosana, inspect this please. ->
+However, we find only modest support for diploidization in the joint model, and conclusions regarding the relative effects of ploidy and breeding system on diversification are robust to its inclusion.
+Also, estimating the rate at which it occurs, based on ploidy level classifications, is highly uncertain (see parameter $\delta$ in \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}), and it can dramatically change ancestral reconstructions (\cref{suppfigure:DPnodipasr,suppfigure:DPasr,suppfigure:DPnodipABasr,suppfigure:DPABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPasr,suppfigure:IDCDCPnodipABasr,suppfigure:IDCDCPABasr}).
 
-Genomic evidence indicates that a paleo-polyploidization may have occurred prior to the origin of the family Solanaceae, rendering all current `diploids' secondarily derived. 
-First, \citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \textit{Solanum lycopersicum}, may have experienced one or more WGDs.
+% ***
+%B: FIXME/TODO Then try to square the oval of history of DP and IC.
+Specifically, it is difficult to know what to make of the delta models.
+Their inclusion is 'positively' supported, but yields results that are hard to reconcile with other data.
+This is one of the more interesting tensions of the paper, in the sense that it binds all of the discussion point and something here has to come to a head.
+%Either what we think we know about SI/SC and reversals is wrong, or something about diploidization inference is wrong.
+Why?
+It is hard to square that we don't see IP state, almost ever!, but it's supposed to be the ancestral state, and not just in Solanaceae, across eudicots!
+Doesn't demand a full treatment here, but it's worth pointing out that if the SI/SC inference is wrong, then we must include reversals, qCI, here and elsewhere and it will change all kinds of inferences.
+Then, we'll have to explain maintenance of trans-generic polymorphism, ha, ha!
+Otherwise, something is off with the inference delta >> rho in some cases.
+If so, we'll have to explain paleo-WGD inference, etc.
+Either way, we fundamentally misunderstand the evolution of one of these two traits.
+% ***
+
+%B: "Genomic evidence" reads as if the genomes support this.
+They do not--the data used in studies, with requisite flaws and problems, are derived from genomes.
+But I don't think that it's okay to call it "genomic evidence".
+It would be a bit like referring to evidence in this paper as "evolutionary evidence".
+A few lines of evidence indicates that a paleo-polyploidization may have occurred prior to the origin of Solanaceae, rendering all current `diploids' secondarily derived. 
+First, \citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \textit{Solanum lycopersicum}, may have experienced one or more rounds of WGDs.
 A subsequent analysis of synteny between grape and \textit{Solanum} genomes, as well as the distribution between inferred paralogs within genomes of \textit{Solanum} (tomato and potato) each suggested that this lineage experienced a likely round of ancient genome duplication or triplication \citep{tomato2012}.
@@
 %Natural and induced tetraploids from a variety of families, including Fabaceae, Onagraceae, Rosaceae, and Solanaceae, almost invariably lose gametophytic SI (reviewed in \citealt{ramsey_1998,stone_2002,mcclure_2009}).
-On the other hand, there is little evidence for occurrence of re-diploidization in the time frame we considered, within the nightshades.
+On the other hand, there is little evidence for the occurrence of re-diploidization in the time frame we considered, within the nightshades.
 First, the Ks-inferred duplication long pre-dates the origin of the family \citep{sarkinen_2013}. 
@@
 At best, the species we consider are ancestrally diploid, having re-diploidized after a suspected WGD.
-% E: But then how is there SI?! Agreed, I'll cover it.
+% E: But then how is there SI?! %B: Agreed, we'll cover it.
 Second, studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
@@
 \bigskip
+
+% B: STOPPED HERE
 
@@
 %
-% Then try to square the oval of history of DP and IC.
-Specifically, it is difficult to know what to make of the delta models.
-Their inclusion is 'positively' supported, but yields results that are hard to reconcile with other data.
-This is one of the more interesting tensions of the paper, in the sense that it binds all of the discussion point and something here has to come to a head.
-%Either what we think we know about SI/SC and reversals is wrong, or something about diploidization inference is wrong.
-Why?
-It is hard to square that we don't see IP state, almost ever!, but it's supposed to be the ancestral state, and not just in Solanaceae, across eudicots!
-Doesn't demand a full treatment here, but it's worth pointing out that if the SI/SC inference is wrong, then we must include reversals, qCI, here and elsewhere and it will change all kinds of inferences.
-Then, we'll have to explain maintenance of trans-generic polymorphism, ha, ha!
-Otherwise, something is off with the inference delta >> rho in some cases.
-If so, we'll have to explain paleo-WGD inference, etc.
-Either way, we fundamentally misunderstand the evolution of one of these two traits.
-
 % including delta makes for some nonsense: ancestral states are all P with rampant recent transitions to D (e.g.

### 2019-03-13 manuscript/discussion.tex

@@
 Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are overshadowed by hidden traits, or breeding system when added to models.
-%R: added hidden trait.
 Interestingly, the difference in diversification rates between the diploids and polyploids is negligible on the background of self-compatibility.
-In the same vein, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against the diploid genomic background.
-Motivated by the widespread findings of paleopolyploid ancestry across angiosperms, we also examined and found modest support for diploidization in the history of the family, but without significant effects on inferences of diversification dynamics.
+Furthermore, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against the diploid genomic background.
+%R: furthermore <- in the same vein
+Motivated by the widespread recent findings of polyploidy in the ancestry of angiosperms, we also examined and found modest support for diploidization in the history of the family, but without significant effects on inferences of diversification dynamics.
 %B: how about "modest"?
-It's 3-ish, but "positive" seems like a really idiosyncratic usage to me, despite kass & raftery.
-was: %B: "mixed" "not strong" what's the right word here?
-E: not "weak", maybe "not strong" or "not nearly as strong as anything else"
+%R- I think modest is a good way to express it because it is the best we can do with chromosome numbers.
+%B: changed "polyploidy events" (polyploidy is a noun, only.
+polyploid either noun or adjective).
+
 Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests of intricate hypotheses regarding a heterogenous diversification process.
@@
 
-\subsection*{Disentangling trait effects on diversification} %B: Emma may not like "effects,"?
+\subsection*{Disentangling trait interactions in a diversification context} %B: Emma may not like "effects,"?
+% R- changed the section name to be more accurate
 
 Previous analyses of the effects of ploidy on diversification showed that diploids had greater net diversification rates than polyploids when exploring across multiple clades of the angiosperm phylogeny \citet{mayrose_2011, mayrose_2015}.
-We obtain a broadly similar outcome, when considering only the effects of ploidy, but the results change significantly with joint consideration of hidden states, and breeding systems. 
-The addition of a hidden trait results in blurring of differences in diversification rates between diploids and polyploids (Figure 2A). %B: fix fig ref
-Simultaneous consideration of breeding system goes a long way in explaining variation in the inferred net diversification rates among ploidy states.
-SI, which is itself associated with a high diversification rate, is the breeding system context in which many diploids occur.
-Among SC species, diploids and polyploids are associated with similar diversification rates (Figure 2F). %B: fix fig ref
-The best interpretation of these results is that some of the effects of a hidden trait seem to be dictated by  each of the two traits under consideration.
+We obtain a broadly similar outcome, when considering only the effects of ploidy, but the results change significantly with joint consideration of hidden states, and breeding systems (\cref{figure:netdivall}A and B).
+The addition of a hidden trait results in blurring of differences in diversification rates between diploids and polyploids.
+Simultaneous consideration of breeding system better explains the variation of the inferred net diversification rates among ploidy states. %RB: better <- further <- long way.
+SI, which is itself associated with a faster diversification rate, is the breeding system context in which many diploids occur.
+Among SC species, diploids and polyploids are associated with similar diversification rates (\cref{figure:netdivall}E).
+These results indicate that some of the effects of a hidden trait can be explained by each of the two traits under consideration. %RB: explained <- dictated
 The finding is unique, and implications may be of wide interest, because it evaluates the relative influence of these two exceptionally interesting traits in a comparative phylogenetic framework.
@@
 Previous work has shown that SI species in Solanaceae are associated with higher diversification rates compared to their SC counterparts \citep{goldberg_2010}. 
-When we inferred only the role of breeding system in the diversification process, we were able to replicate the slower net diversification for SC (Figure 2C).  %B: fix fig ref
-Incorporating hidden states again resulted in SI associated with greater net diversification rates than SC in both hidden states (Figure 2D). 
+When we only considered the role of breeding system in the diversification process, we were able to replicate the slower net diversification for SC  (\cref{figure:netdivall}C).
+%B: we only considered <- we inferred only
+Incorporating hidden states again resulted in SI associated with greater net diversification rates than SC in both hidden states  (\cref{figure:netdivall}D). 
 However, the overall picture is a bit more complex.
@@
 
-Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.
+%B: I will yield to %R here, but I still think there is overwhelming support for this statement, and none of the cited papers show convincingly that, for example, diploidization happens within five orders of magnitude compared with polyploidization.
+%Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.
+Flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}. 
+Following polyploidization, it is possible for genome reorganization, downsizing, and loss \citep{dodsworth_2015, zenil_2016, mandakova_2018}.
+%B: commas, etc.: importantly rm: "has been observed" this is _only_ true on very, very, short time scales
 As a consequence, it is possible that nearly all extant species classified as ``diploid" in our analyses, based on cytogenetic data, are possibly secondary diploids, having undergone both polyploidization and re-diploidization.
+
 Unduly ignoring secondary diploidization would clearly underestimate the rate of transitions from polyploids to diploids, and would then likely cause inflated net diversification rates for diploids, because species considered in the diploid state, may have been ancestrally polyploid, instead.
-%B: This should be to mutual liking.
-IMO, the support is STRONG in the apparently meager analyses, one with only D/P), which yields highly unreasonable ancestral states.
-But it is decidedly not strong, i.e.
-"positive," in the apparently more informative joint trait analyses.
-We find strong support for including the diploidization rate parameter ($\delta$) in analyses that include only ploidy \cref{table:bayesfactors}. %B <- %Rosana, inspect this please. ->
+%B: added before conflict/edits
+We find strong support for including the diploidization rate parameter ($\delta$) in analyses that include only ploidy \cref{table:bayesfactors}.
+%B <- %Rosana, inspect this please.
+It's a merge of conflicted lines.
 However, we find only modest support for diploidization in the joint model, and conclusions regarding the relative effects of ploidy and breeding system on diversification are robust to its inclusion.
@@
 
-%B: "Genomic evidence" reads as if the genomes support this.
-They do not--the data used in studies, with requisite flaws and problems, are derived from genomes.
-But I don't think that it's okay to call it "genomic evidence".
+%B: pre-conflict: "Genomic evidence" reads as if the genomes support this.
+But genomes also do not support the statement (e.g., synteny).
+They are the data used in studies, not the evidence (it comes from modes of analysis).
 It would be a bit like referring to evidence in this paper as "evolutionary evidence".
-A few lines of evidence indicates that a paleo-polyploidization may have occurred prior to the origin of Solanaceae, rendering all current `diploids' secondarily derived. 
-First, \citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \textit{Solanum lycopersicum}, may have experienced one or more rounds of WGDs.
+A few lines of evidence indicate that polyploidy may have occurred prior to the origin of Solanaceae, rendering all extant `diploids' secondarily derived.
+%B: "polyploid event" or "polyploidy" works, but not "polyploidy event"; see above.
+First, \citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \textit{Solanum lycopersicum}, may have experienced one or more of WGDs.
 A subsequent analysis of synteny between grape and \textit{Solanum} genomes, as well as the distribution between inferred paralogs within genomes of \textit{Solanum} (tomato and potato) each suggested that this lineage experienced a likely round of ancient genome duplication or triplication \citep{tomato2012}.
@@
 
-%B: incorporate ancestral states in absence of I/C consideration to rag on \delta
+%B: incorporate ancestral states in absence of I/C consideration to "rag" on \delta? Rag on S-alleles for Ks, or just throw up arms?
 %Natural and induced tetraploids from a variety of families, including Fabaceae, Onagraceae, Rosaceae, and Solanaceae, almost invariably lose gametophytic SI (reviewed in \citealt{ramsey_1998,stone_2002,mcclure_2009}).

### 2019-03-13 manuscript/discussion.tex
- backup of discussion up to summary, with addition of Paradox paragraph

@@
 Although species are composed of vast assemblages of variable traits, many traits are heritable, and they may affect the propensity of species to perish and multiply \citep{lewontin_1970}, examining such macroevolutionary effects of many traits is rare, and remains challenging.
-We investigated the influence of two traits, ploidy and breeding system states, in diversification of Solanaceae.
+We investigated the influence of two traits, ploidy and breeding system, in diversification of Solanaceae.
 Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are overshadowed by hidden traits, or breeding system when added to models.
@@
 polyploid either noun or adjective).
-
 Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests of intricate hypotheses regarding a heterogenous diversification process.
 Below, we discuss key findings and highlight the most interesting and paradoxical implications.
-We additionally provide an outline for broader application of phylogenetic comparative methods in the context of interactions among traits and diversification process.
+Below, we provide an outline for broader application of phylogenetic comparative methods in the context of interactions among traits and diversification process.
 
-\subsection*{Disentangling trait interactions in a diversification context} %B: Emma may not like "effects,"?
-% R- changed the section name to be more accurate
+\subsection*{Disentangling trait interactions in the context of diversification} %RB: merged two titles; good?
 
@@
 
-Previous work has shown that SI species in Solanaceae are associated with higher diversification rates compared to their SC counterparts \citep{goldberg_2010}. 
+Earlier work has shown that SI species in Solanaceae are associated with higher diversification rates compared to their SC counterparts \citep{goldberg_2010}. 
 When we only considered the role of breeding system in the diversification process, we were able to replicate the slower net diversification for SC  (\cref{figure:netdivall}C).
@@
 
-%B: TODO: add below to future goals.
 %R: and I: polyploid lineage that has re-diploidized may enhance diversification.
@@
 %Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.
+
+Our exploratory analyses recover some support for inclusion of diploidization in the history of the family, but without significant effects on inferences of diversification dynamics.
+The implications of the results of our diploidization analyses, as well as much of the other recent work on WGDs, are interesting because they conflict with other lines of evidence, especially those concerning the observed simple synteny of genomes in Solanaceae, and patterns of evolution as self-incompatibility loci.
+
 Flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}. 
@@
 As a consequence, it is possible that nearly all extant species classified as ``diploid" in our analyses, based on cytogenetic data, are possibly secondary diploids, having undergone both polyploidization and re-diploidization.
-
-Unduly ignoring secondary diploidization would clearly underestimate the rate of transitions from polyploids to diploids, and would then likely cause inflated net diversification rates for diploids, because species considered in the diploid state, may have been ancestrally polyploid, instead.
-%B: added before conflict/edits
+Unduly ignoring secondary diploidization would necessarily underestimate the rate of transitions from polyploids to diploids.
+It would then likely cause inflated net diversification rates for diploids, because species considered diploid may have been ancestrally polyploid, instead. %B: added before conflict/edits
 We find strong support for including the diploidization rate parameter ($\delta$) in analyses that include only ploidy \cref{table:bayesfactors}.
@@
 However, we find only modest support for diploidization in the joint model, and conclusions regarding the relative effects of ploidy and breeding system on diversification are robust to its inclusion.
-Also, estimating the rate at which it occurs, based on ploidy level classifications, is highly uncertain (see parameter $\delta$ in \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}), and it can dramatically change ancestral reconstructions (\cref{suppfigure:DPnodipasr,suppfigure:DPasr,suppfigure:DPnodipABasr,suppfigure:DPABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPasr,suppfigure:IDCDCPnodipABasr,suppfigure:IDCDCPABasr}).
+Furthermore, estimating the rate of diploidization, based on our ploidy level classifications, is highly uncertain (see parameter $\delta$ in \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}), and yields dramatic changes in ancestral reconstructions (\cref{suppfigure:DPnodipasr,suppfigure:DPasr,suppfigure:DPnodipABasr,suppfigure:DPABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPasr,suppfigure:IDCDCPnodipABasr,suppfigure:IDCDCPABasr}).
 
-% ***
-%B: FIXME/TODO Then try to square the oval of history of DP and IC.
-Specifically, it is difficult to know what to make of the delta models.
-Their inclusion is 'positively' supported, but yields results that are hard to reconcile with other data.
-This is one of the more interesting tensions of the paper, in the sense that it binds all of the discussion point and something here has to come to a head.
-%Either what we think we know about SI/SC and reversals is wrong, or something about diploidization inference is wrong.
-Why?
-It is hard to square that we don't see IP state, almost ever!, but it's supposed to be the ancestral state, and not just in Solanaceae, across eudicots!
-Doesn't demand a full treatment here, but it's worth pointing out that if the SI/SC inference is wrong, then we must include reversals, qCI, here and elsewhere and it will change all kinds of inferences.
-Then, we'll have to explain maintenance of trans-generic polymorphism, ha, ha!
-Otherwise, something is off with the inference delta >> rho in some cases.
-If so, we'll have to explain paleo-WGD inference, etc.
-Either way, we fundamentally misunderstand the evolution of one of these two traits.
-% ***
-
-%B: pre-conflict: "Genomic evidence" reads as if the genomes support this.
-But genomes also do not support the statement (e.g., synteny).
-They are the data used in studies, not the evidence (it comes from modes of analysis).
-It would be a bit like referring to evidence in this paper as "evolutionary evidence".
-A few lines of evidence indicate that polyploidy may have occurred prior to the origin of Solanaceae, rendering all extant `diploids' secondarily derived.
+Some lines of evidence indicate that polyploidy may have occurred prior to the origin of Solanaceae, rendering all extant `diploids' secondarily derived.
 %B: "polyploid event" or "polyploidy" works, but not "polyploidy event"; see above.
@@
 A subsequent analysis of synteny between grape and \textit{Solanum} genomes, as well as the distribution between inferred paralogs within genomes of \textit{Solanum} (tomato and potato) each suggested that this lineage experienced a likely round of ancient genome duplication or triplication \citep{tomato2012}.
-The age of the peak of paralog Ks distances, is approximately 71 million years \citep{tomato2012}. 
-If this is the case, then all of the genomes may have been subsequently re-diploidized, yielding the common extant base chromosome numbers in this and related families, n=11-12, presently considered to be diploid \citep{robertson_2011}.
-%Not unlikely, because Arabidopsis n=5.
+The age of the peak of paralog Ks distances, is approximately 71$\pm$19 My \citep{tomato2012}. 
+If this is the case, then all of the genomes may have been subsequently re-diploidized, yielding the widespread and common chromosome numbers in this and related families, n=11-12, presently considered to be diploid \citep{robertson_2011}.
 
-%B: incorporate ancestral states in absence of I/C consideration to "rag" on \delta? Rag on S-alleles for Ks, or just throw up arms?
-%Natural and induced tetraploids from a variety of families, including Fabaceae, Onagraceae, Rosaceae, and Solanaceae, almost invariably lose gametophytic SI (reviewed in \citealt{ramsey_1998,stone_2002,mcclure_2009}).
-On the other hand, there is little evidence for the occurrence of re-diploidization in the time frame we considered, within the nightshades.
-First, the Ks-inferred duplication long pre-dates the origin of the family \citep{sarkinen_2013}. 
-The best-supported, most-recent WGD, which may represent a reversion from polyploidy to diploidy, is older than the lineage under consideration in this study.
-At best, the species we consider are ancestrally diploid, having re-diploidized after a suspected WGD.
-% E: But then how is there SI?! %B: Agreed, we'll cover it.
+On the other hand, there is little evidence for the occurrence of re-diploidization within Solanaceae, since the origin of the family---the time frame we considered.
+First, the Ks-inferred duplication likely pre-dates the origin of the family (49 My, HPD 46-53 My \citealt{sarkinen_2013}). 
+The best-supported, most-recent WGD \citet{tomato2012} is older than the lineage under consideration in this study.
+At best then, the species we consider are ancestrally diploid, having re-diploidized after a suspected WGD, but likely before the origin of the family.
 Second, studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
 Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species, including all of those in the relatively cytogenetically conserved `x=12' group, which includes tomato, potato, eggplant, pepper, and tobacco.
-%B: Once this is re-arranged, and made to fit with the rest of the discussion, add what this has to do with re-diploidization and how to interpret $\delta$.
-It is perhaps not surprising that models diploidization in our study recovered wide credible intervals for the diploidization parameter, $\delta$. 
-Despite the limitations of assignment of ploidy levels, we found some support for models that include pararmeter $\delta$, meaning that further study may be warranted, especially in other groups.
-%Moreover, rare state changes may not be easily recovered. Emma, you had a clever way to summarize this. 
-%     meaning of diploidization parameter
+Despite the limitations of assignment of ploidy levels, we found some support for models that include parameter $\delta$, meaning that further study may be warranted, especially in other groups.
 
-% generality of approach
-%
-% issues
-%     irreversibility SI -> SC assumption
+Regardless of whether genome polypoidization, followed by widespread diploidization, is a dominant mechanism of genome evolution in Solanaceae, it is clear that our present understanding of the evolution of ploidy and breeding systems is incomplete.
+An increasingly forceful weight of evidence seems to support WGDs in the ancestry of many angiosperms. %cite
+It is then easily inferred that many species have undergone diploidization, alongside genome size and chromosome number reductions. %cite
+Inference of such a common genome upheaval in eudicot history seems to clash with the data indicating that a homologous mechanism of SI, which breaks down nearly invariably in natural and induced tetraploids \citep{stone_2002,mcclure_2009}, has been present continually in many families, including Solanaceae \citep{igic_2006}.
+Most problematically in this context, it is unclear how to explain the maintenance of trans-generic polymorphism at the orthologous S-loci in this and other families.
+As well, chromosome numbers and genome comparisons within the family (esp. `x=12' clade, containing \textit{Solanum} and \textit{Nicotiana}) reveal strong conservation.
+\citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca.
+10 inversions and 11 translocations (likely underestimated).
+Recovery of such simple relationships would require outstanding convergent loss of duplicated segments.
+In either case, all of the approaches make ample assumptions, and it seems that at least some of them will necessitate deep revisions.
+Our study, for example, made numerous assumptions in coding trait states (e.g.
+what constitutes a polyploid or diploid, SI or SC species), in models used for analyses (e.g.
+irreversibility of SI loss underlain by trans-specific orthologous S-loci), and many more, which could each mislead inference.
+%B: -> %R: Rosana, take a look at this.
+% no discussion of magnitude of \delta vs \rho!
+% problems: this system has a big set of advantages - carefully curated (co-)occurrences of two traits, genetic basis, etc. and everything is still contradictory and opaque.
 
@@
 
-% B: STOPPED HERE
-
-\subsection*{Conclusions}
-% problems: this system has a big set of advantages - carefully curated (co-)occurrences of two traits, genetic basis, etc. and everything is still contradictory and opaque.
-
-% R-Here I copied the limitations of the models idea but needs to be reorganized to be cohesive % Fix this a lot more please
-% This is the idea on how thorough analyses should be done.
-Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as disscussed in  \citet{rabosky_2010, beaulieu_2015}),but also, a problem of inadequate  ``*SSE'' model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focus trait first used to model diversification.
-We presented a roadmap with a series of analyses that can help scientists to identify if an unknown or unobservable trait is worth pursuing.
+Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as disscussed in  \citet{rabosky_2010, beaulieu_2015}), but also, a problem of inadequate  ``*SSE'' model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focus trait first used to model diversification.
+We presented a roadmap with a series of analyses that can identify whether unknown or unobservable traits are worth pursuing. 
 In this roadmap, we have proposed fitting HiSSE models that have effectively shown that some binary states associated with diversification under BiSSE analyses are actually not different in net diversification terms, and instead a hidden state is linked to differences in the diversification rate \citep{beaulieu_2016}.
@@
  %%%%%%%%%%%
+
+%B: pre-conflict comment: "Genomic evidence" reads as if the "genomes" support this.
+But "genomic evidence" also does not support the statement (e.g., widespread synteny).
+Genomes are the data used in studies, not the evidence (it comes from different modes of genomic analysis).
 
@@
 %
-% - pathways
-% 
-%- diploidization
-% - rate, effect
-% - history in solanaceae
-%
-% - summary
-%
-% B: NEW REORG COMMENTS:
 
@@
 %
-% including delta makes for some nonsense: ancestral states are all P with rampant recent transitions to D (e.g.
-Fig.
-S4/S8.
-Not backed by any data, including synteny, S-allele trans-polymorphism, or anecdotally bearing experimental data about rates of spontaneous polyploidization and diploidization \citep{husband_2012}.
 

### 2019-03-13 manuscript/abstract.tex
- Karolis' comments and edits, mostly found typos and similar

@@
 %Two traits that have been repeatedly suggested as drivers of differential diversification are whether a lineage is diploid or polyploid, and whether it is self-incompatible or self-compatible.
-Ploidy (diploid vs. polyploid states) and breeding system (self-incompatible vs self-compatible states) have been repeatedly suggested as possibly drivers of differential diversification
+Ploidy (diploid vs. polyploid states) and breeding system (self-incompatible vs self-compatible states) have been repeatedly suggested as possibly drivers of differential diversification.
 We investigate the role of these traits, including their interaction, on speciation and extinction rates in Solanaceae.

### 2019-03-13 manuscript/methods.tex
- Karolis' comments and edits, mostly found typos and similar

@@
 For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus or based on SI/SC classification.
-For example, \textit{Solanum betaceum} did not have information about ploidy level but it has 24 chromosomes, so since $x=12$ is the base chromosome number of the \textit{Solanum} genus \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid.
+For example, \textit{Solanum betaceum} did not have information about ploidy level but it has 24 chromosomes, so since $x=12$ is the base chromosome number of the genus \textit{Solanum} \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid.
 Additionally, because of the absence of SI polyploids (discussed further below), species known to be SI could be scored as diploid.
@@
 Many species could unambiguously be coded as either I or C \citep{raduski_2012}.
-Following previous work, we coded as I any species with a functional SI system, even if SC or dioecy was also reported.
+Following previous work, we coded any species with a functional SI system as I, even if SC or dioecy was also reported.
 Dioecious species without functional SI were coded as C.
@@
 Hybrids and cultivars were excluded because ploidy and breeding system can be affected by artificial selection during domestication.
-Following the reasoning outlined in \citet{robertson_2011}, we examined closely the few species for which the merged ploidy and breeding system data indicated the presence of self-incompatible polyploids.
-Although SI populations frequently contain some SC individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find a convincing case of a naturally occurring SI and polyploid population.
+Following the reasoning outlined in \citet{robertson_2011}, we closely examined the few species for which the merged ploidy and breeding system data indicated the presence of self-incompatible polyploids.
+Although SI populations frequently contain some SC individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find convincing data for a naturally occurring SI polyploid population.
 The single instance of an SI polyploid individual appears to be an allopentaploid hybrid of \textit{Solanum oplocense} Hawkes x \textit{Solanum gourlayii} Hawkes, reported by \citet{camadro_1981}.
@@
 We retained all of these species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
-A total of 405 taxa without any information about breeding system or polyploidy were excluded.
+A total of 405 taxa without any information about breeding system or ploidy were excluded.
 %Tips without trait data might not improve point estimation of diversification parameters while increasing uncertainty in the estimates of trait linked diversification
@@
 
-Self-incompatibility is homologous in all species in which S-alleles have been cloned and controlled crosses performed.
+Self-incompatibility is homologous in all Solanaceae species in which S-alleles have been cloned and controlled crosses performed.
 All species sampled to date possess a non-self recognition, RNase-based gametophytic self-incompatibility \citep[shared even with other euasterid families;][]{ramanauskas_2017}.
@@
 
-Polyploidy and breeding system might influence lineage diversification individually, but these two traits also have complex interactions in Solanaceae species.
+Ploidy and breeding system might influence lineage diversification individually, but these two traits also have complex interactions in Solanaceae species.
 Therefore, we considered a multi-state model that investigates the contribution of both traits and the possible transitions between them.
@@
 
-The ID/CD/CP model could also suffer from a high type I error, just like a two state BiSSE model. 
+The ID/CD/CP model could also suffer from a high type I error, like the two-state BiSSE model. 
 We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018,  huang_2018}.
@@
 In the two-step pathway, the CD state is an intermediate: SI is first lost, and later the SC diploid undergoes polyploidization.
-We quantify the relative contribution of these pathways to polyploidy in two ways, each using the maximum a posteriori (MAP) estimates of rates from the IC/CD/CP model.
+We quantify the relative contribution of these pathways to polyploidy in two ways, each using the maximum \textit{a posteriori} (MAP) estimates of rates from the IC/CD/CP model.
 
@@
 We insert an artificial division in the CP state, so that one substate contains the CP species that arrived via the one-step pathway and the other substate contains the CP species that arrived via the two-step pathway.
-We consider only unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and not strong support for diploidization (see below).
-First, we consider only the rates of transitions between these states, placing them in the propogation matrix \myvec{Q}.
+We consider only unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and no strong support for diploidization (see below).
+First, we consider only the rates of transitions between these states, placing them in the propagation matrix \myvec{Q}.
 The matrix $\myvec{P} = \exp(\myvec{Q} t)$ then provides the probabilities of changing from one state to any other state after time $t$.
@@
 % "one-step predictions for P_I and P_C over time"
-% E & R todo later: for the case with net div, write out the propogation matrix and equations for pathway contributions
+% E & R todo later: for the case with net div, write out the propagation matrix and equations for pathway contributions
+%B: I corrected all instances of "propogation" in text and comments
 
@@
 As discussed by \citet{servedio_2014}, the presence or absence of a process can have an exploratory goal.
-In our case the diploidization parameter (or its absence, $\delta=0$) in our models is an opportunity to explore an assumption that might be important but that is not the most essential process to understand the interactions among polyploidy, breeding system, and diversification.
+In our models, the diploidization parameter (or its absence, $\delta=0$), represents an opportunity to explore a possibly significant assumption that may not decisively affect interactions among ploidy, breeding system, and diversification.
 
@@
 Additionally, ancestral states at each node in the phylogeny were sampled jointly during the MCMC analyses.
-Ancestral state reconstructions for all models can be found in the Supplementary Information and show the maximum a posteriori of the marginal probability distributions for each of the 594 internal nodes for each of the 10 models.
+Ancestral state reconstructions for all models can be found in the Supplementary Information and show the maximum posterior estimates of the marginal probability distributions for each of the 594 internal nodes for each of the 10 models.
 % TODO: add fig refs (or remove the mention of them)
+%B: there was an error in the above sentence: no estimates or values after a posteriori (also "posterior" is ok?)
 

### 2019-03-13 manuscript/results.tex
- Karolis' comments and edits, mostly found typos and similar

@@
 This result holds with (\cref{figure:netdivall}A) or without (\cref{figure:netdivnodip}A) the diploidization parameter, although including diploidization shifts the net diversification rate of polyploids to be non-negative.
-Incorporating a hidden state in this model, however, removes the clear separation in diversification between diploids and polyploids (D/P+A/B model; \cref{figure:netdivall}B, \cref{figure:netdivnodip}B).
-Thus, differences in net diversification are better explained by an unknown factor than by ploidy.
+Incorporating a hidden state in this model, however, removes the clear separation in diversification rate estimates between diploids and polyploids (D/P+A/B model; \cref{figure:netdivall}B, \cref{figure:netdivnodip}B).
+Thus, differences in net diversification rates are better explained by an unknown factor than by ploidy.
 Statistical model comparisons show very strong support for including the hidden state and strong support for including diploidization (\cref{table:bayesfactors}).
@@
 When considering breeding system alone (I/C model, \cref{figure:netdivall}C), we found a larger net diversification rate for SI than for SC species, in agreement with \citet{goldberg_2010}.
-When a hidden state is included (I/C+A/B model), the large net diversification difference persists for one of the hidden states but is removed for the other (\cref{figure:netdivall}D).
-Thus, differences in net diversification are best explained by both breeding system and an unknown factor \citep[as in Onagraceae;][]{freyman_2018}.
+When a hidden state is included (I/C+A/B model), the large net diversification rate difference persists for one of the hidden states but is removed for the other (\cref{figure:netdivall}D).
+Thus, differences in net diversification rates are best explained by both breeding system and an unknown factor \citep[as in Onagraceae;][]{freyman_2018}.
 % The transition rate from SI to SC is $q_{IC}=0.3$ for both the I/C and the I/C+A/B models.
@@
 When a hidden state is included (IC/CD/CP+A/B model), the same general pattern remains when diploidization is prevented (\cref{figure:netdivnodip}D), although the higher net diversification rate of ID is less clear within one of the hidden states.
-With diploidization, the net diversification rate of ID is still greater than CD within each hidden state, but diversification for P is highly uncertain and perhaps bimodal.
+With diploidization, the net diversification rate of ID is still greater than CD within each hidden state, but diversification estimate for P is highly uncertain and perhaps bimodal.
 Statistical model comparisons show very strong support for including the hidden state and at most positive support for including diploidization (\cref{table:bayesfactors}).

### 2019-03-13 manuscript/discussion.tex
- "Final" minor changes and typo/error corrections from Boris - on to Rosana for completion

@@
 Below, we discuss key findings and highlight the most interesting and paradoxical implications.
-Below, we provide an outline for broader application of phylogenetic comparative methods in the context of interactions among traits and diversification process.
+We provide an outline for broader application of phylogenetic comparative methods in the context of interactions among traits and diversification process.
 
@@
 
-% examine SC-only families
-% no causal correlation in SSI systems
-% paths to dioecy
-% path-dependent net-diversification
-
-% qualifying paragraph for pathways
-We find that the SI diploid lineages are much more likely to take a
-one-step  ID$\rightarrow$CP pathway to polyploidy rather than a two-step ID$\rightarrow$CD$\rightarrow$CP pathway, in Solanaceae.
+We find that SI diploid lineages are much more likely to take a one-step  ID$\rightarrow$CP pathway to polyploidy rather than a two-step ID$\rightarrow$CD$\rightarrow$CP pathway, in Solanaceae.
 This result is significant because it has a number of implications for the inference of history of breeding and sexual systems in flowering plants.
@@
 A more direct and definitive test of this hypothesis would additionally examine the propensity of polyploids generated through either pathway to become dioecious \citep{robertson_2011}.
-% -- mutational opportunity vs lineage transition rate reveals something about selection 
 Second, statistical phylogenetic approaches measure lineage transition rates.
@@
 TODO
+%
+% examine SC-only families
+% no causal correlation in SSI systems
+% paths to dioecy
+% path-dependent net-diversification
+%
 %B: also missing deeper discussion of with/without SDD/SID estimates and comparison with Robertson et al. 2011.
-
+%
 %B: And, are diversification rates of polyploid SC lineages different if one or other path was taken? % E: This is a great question.  We don't answer it with this round of analyses.  But discuss?
@@
 
-%B: I will yield to %R here, but I still think there is overwhelming support for this statement, and none of the cited papers show convincingly that, for example, diploidization happens within five orders of magnitude compared with polyploidization.
-%Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.
-
+%B: yielding to %R here.
+I still think there is overwhelming support for this statement, and few of the cited papers show convincingly that, for example, diploidization happens within five orders of magnitude compared with polyploidization, rendering "relatively rare" obsolete or invalid.
+(P.S.
+Rare not the same as unimportant!)
+%B: was: Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.
 Our exploratory analyses recover some support for inclusion of diploidization in the history of the family, but without significant effects on inferences of diversification dynamics.
@@
 Since MuSSE might also suffer from hight type I errors considering more heterogeneity using a hidden state model on top of the multistate SSE is necessary to adequately infer the effect of the complex interactions in the speciation and extinction process.
-
- %%%%%%%%%%%
- %% Notes to Self %%
- %%%%%%%%%%%
-
-%B: pre-conflict comment: "Genomic evidence" reads as if the "genomes" support this.
-But "genomic evidence" also does not support the statement (e.g., widespread synteny).
-Genomes are the data used in studies, not the evidence (it comes from different modes of genomic analysis).
-
-% outline
-% - disentangling effects
-% - general value of approach
-%  		another kind of confounding processes, not only transition and diversification, but "other diversifications" and "other transitions", esp. correlated
-%
-% - our study key findings and details
-% - effects of Ploidy (as %R)
-% - trumped by BS
-% - no difference on background of SC 
-% - discuss data/models and their long/shortcomings: 
-%	ClaSSE, will's idea for stochastic mapping
-%	irreversibility SI -> SC assumption
-% - how to improve / what next
-%  (may be useful to study in SC families)
-%
-
-% First, include all straightforward Results and related observations
-%
-
-% PATHWAYS -- mutational opportunity vs lineage transition rate reveals something about selection 
-%
-% Starting with SI-D, selection favoring establishment of SC-D seems more common than that of SC-P.
-% This is because the mutation rates are likely higher to SC-P, but rate 3x lower. "
-%Husband2012: The prevalence of polyploidy within plants has led to debate about the influence of genome duplication on net species diversification (Otto and Whitton 2000).
-Researchers have argued, on one hand, that polyploidy may simply accumulate as a unidirectional mutation with little influence on the rate or trajectory of evolutionary divergence (Stebbins 1971; Meyers and Levin 2006).
-This idea is based on the premise that ploidy is more likely to increase than decrease in time.
-Although reversals do occur, the resulting haploids are relaively rare and often of reduced fitness and high sterility (Stebbins 1980; Ramsey and Schemske 2002).
-Alternatively, polyploidy itself may enhance evolutionary diversification leading to increased divergence and species richness (Otto and Whitton 2000).
-Elevated diversification may occur if polyploidy accelerates adaptive divergence and evolution of reproductive isolation (Otto and Whitton 2000) or reduces species extinction.
-%The presence of intraspecific ploidy diversity has three main causes.
-First, genome duplication is an ongoing genomic mutation within many natural populations and ploidy polymorphism is therefore expected when poly- ploid formation is common.
-In fact, Ramsey and Schemske (1998) estimated that polyploids form at the rate of 10 5 per individual per generation, well above the genic mutation rate, and in some cases it may be even higher than that (Husband 2004).
-Second, depending on the frequency and pathways of polyploid formation, diploids and polyploids may be maintained as genetically homogeneous populations, which are justifiably treated as single taxonomic species.
-Finally, ploidy polymorphism may arise as an artefact of our imperfect taxonomic approach.
-Since morphological divergence is often used as a criterion for species delinea- tion, polyploid lineages may not be recognized at the species level, especially in autopolyploids, which often have a close resemblance to their diploid progenitors (Soltis et al.
-2007).

### 2019-03-13 manuscript/results.tex
- "Final" minor changes and typo/error corrections from Boris - on to Rosana for completion

@@
 
-There are two pathways by which a SI diploid lineages will eventually---given enough time---become a SC polyploid.
+There are two pathways by which SI diploid lineages eventually---given enough time---become a SC polyploid.
 In the one-step pathway, polyploidization directly disables SI.
@@
 
-We considered models both with or without diploidization in order to explore its effects on the estimates of state-dependent diversification.
+We considered models both with and without diploidization in order to explore its effects on the estimates of state-dependent diversification.
 For the two models that only include diploid and polyploid states (D/P and D/P+A/B), the net diversification rate of polyploid lineages is likely negative when diploidization (parameter $\delta$) is excluded but positive when $\delta$ is included (\cref{figure:netdivall}A versus \cref{figure:netdivnodip}A).
-For the two models that also included breeding system (ID/CD/CP and ID/CD/CP+A/B), the main effect of including diploidization is greatly increasing the uncertainty of the estimate of polyploid net diversification rate (\cref{figure:netdivall}E,F versus \cref{figure:netdivnodip}C,D).
+For the two models that also included breeding system (ID/CD/CP and ID/CD/CP+A/B), the main effect of including diploidization is greatly increasing the uncertainty of the estimated polyploid net diversification rate (\cref{figure:netdivall}E,F versus \cref{figure:netdivnodip}C,D).
 For all models, there is greater uncertainty in the estimate of diploidization rate than polyploidization rate, as judged by the width of the credibility intervals (see \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}).

### 2019-03-14 manuscript/discussion.tex
Small changes in the discussion, two final lines added

@@
 
-\subsection*{Disentangling trait interactions in the context of diversification} %RB: merged two titles; good?
+\subsection*{Disentangling trait interactions in the context of diversification} %RB: merged two titles; good? % R-Good
 
@@
 %B: was: Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.
-Our exploratory analyses recover some support for inclusion of diploidization in the history of the family, but without significant effects on inferences of diversification dynamics.
+%Our exploratory analyses recover some support for inclusion of diploidization in the history of the family, but without significant effects on inferences of diversification dynamics.
+%R- The line above is included in the bottom so no more need for that here
 The implications of the results of our diploidization analyses, as well as much of the other recent work on WGDs, are interesting because they conflict with other lines of evidence, especially those concerning the observed simple synteny of genomes in Solanaceae, and patterns of evolution as self-incompatibility loci.
@@
 Flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}. 
-Following polyploidization, it is possible for genome reorganization, downsizing, and loss \citep{dodsworth_2015, zenil_2016, mandakova_2018}.
+Following polyploidization, it is possible for genome reorganization, downsizing, and loss to occur \citep{dodsworth_2015, zenil_2016, mandakova_2018}.
 %B: commas, etc.: importantly rm: "has been observed" this is _only_ true on very, very, short time scales
-As a consequence, it is possible that nearly all extant species classified as ``diploid" in our analyses, based on cytogenetic data, are possibly secondary diploids, having undergone both polyploidization and re-diploidization.
+As a consequence, nearly all extant species classified as ``diploid" in our analyses, based on cytogenetic data, are possibly secondary diploids, having undergone both polyploidization and re-diploidization.
 Unduly ignoring secondary diploidization would necessarily underestimate the rate of transitions from polyploids to diploids.
 It would then likely cause inflated net diversification rates for diploids, because species considered diploid may have been ancestrally polyploid, instead. %B: added before conflict/edits
-We find strong support for including the diploidization rate parameter ($\delta$) in analyses that include only ploidy \cref{table:bayesfactors}.
-%B <- %Rosana, inspect this please.
-It's a merge of conflicted lines.
-However, we find only modest support for diploidization in the joint model, and conclusions regarding the relative effects of ploidy and breeding system on diversification are robust to its inclusion.
+%We find strong support for including the diploidization rate parameter ($\delta$) in analyses that include only ploidy . %B <- %Rosana, inspect this please. It's a merge of conflicted lines.
+% R- I put it up so the line above is not necessary anymore
+However, we find only modest support for diploidization in the joint model  (\cref{table:bayesfactors}), and conclusions regarding the relative effects of ploidy and breeding system on diversification are robust to its inclusion.
 Furthermore, estimating the rate of diploidization, based on our ploidy level classifications, is highly uncertain (see parameter $\delta$ in \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}), and yields dramatic changes in ancestral reconstructions (\cref{suppfigure:DPnodipasr,suppfigure:DPasr,suppfigure:DPnodipABasr,suppfigure:DPABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPasr,suppfigure:IDCDCPnodipABasr,suppfigure:IDCDCPABasr}).
@@
 % no discussion of magnitude of \delta vs \rho!
+% R- magnitude is not easy to discuss about, what we have are estimates and credible intervals which are completely sample size dependent. With a different sample "magnitude" can change.
 % problems: this system has a big set of advantages - carefully curated (co-)occurrences of two traits, genetic basis, etc. and everything is still contradictory and opaque.
@@
 When considering  a second candidate trait, one can model the complex interactions via a MuSSE model, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
-Since MuSSE might also suffer from hight type I errors considering more heterogeneity using a hidden state model on top of the multistate SSE is necessary to adequately infer the effect of the complex interactions in the speciation and extinction process.
+Since MuSSE might also suffer from hight type I errors considering more heterogeneity using a hidden state model on top of the multistate diversification model is necessary to adequately infer the effect of the complex interactions in the speciation and extinction process.
+% R- new lines
+Investigating the process of diversification linked to traits can be done via thorough statistical inferences that carefully evaluate multiple evolutionary histories. 
+Background heterogeneity in lineage of diversification is the rule rather than the exception.
+Inferences that question whether other complex trait interactions are hidden in this background heterogeneity are necessary because they critically contribute to the reconstruction and understanding of diversification.

### 2019-03-14 manuscript/discussion.tex
- Submission #1 final changes - added cover letter and manuscript proofs

@@
 Furthermore, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against the diploid genomic background. 
-%R: furthermore <- in the same vein
 Motivated by the widespread recent findings of polyploidy in the ancestry of angiosperms, we also examined and found modest support for diploidization in the history of the family, but without significant effects on inferences of diversification dynamics.
@@
 
-\subsection*{Disentangling trait interactions in the context of diversification} %RB: merged two titles; good? % R-Good
+\subsection*{Disentangling trait interactions in the context of diversification}
 
@@
 The addition of a hidden trait results in blurring of differences in diversification rates between diploids and polyploids.
-Simultaneous consideration of breeding system better explains the variation of the inferred net diversification rates among ploidy states. %RB: better <- further <- long way.
+Simultaneous consideration of breeding system better explains the variation of the inferred net diversification rates among ploidy states. 
 SI, which is itself associated with a faster diversification rate, is the breeding system context in which many diploids occur.
 Among SC species, diploids and polyploids are associated with similar diversification rates (\cref{figure:netdivall}E).
-These results indicate that some of the effects of a hidden trait can be explained by each of the two traits under consideration. %RB: explained <- dictated
+These results indicate that some of the effects of a hidden trait can be explained by each of the two traits under consideration. 
 The finding is unique, and implications may be of wide interest, because it evaluates the relative influence of these two exceptionally interesting traits in a comparative phylogenetic framework.
@@
 The implications of the results of our diploidization analyses, as well as much of the other recent work on WGDs, are interesting because they conflict with other lines of evidence, especially those concerning the observed simple synteny of genomes in Solanaceae, and patterns of evolution as self-incompatibility loci.
-
 Flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}. 
 Following polyploidization, it is possible for genome reorganization, downsizing, and loss to occur \citep{dodsworth_2015, zenil_2016, mandakova_2018}.
-%B: commas, etc.: importantly rm: "has been observed" this is _only_ true on very, very, short time scales
+%B: commas, etc.: importantly rm: "has been observed" this is _only_ true on very short time scales
 As a consequence, nearly all extant species classified as ``diploid" in our analyses, based on cytogenetic data, are possibly secondary diploids, having undergone both polyploidization and re-diploidization.
@@
 
-Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as disscussed in  \citet{rabosky_2010, beaulieu_2015}), but also, a problem of inadequate  ``*SSE'' model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focus trait first used to model diversification.
+%B: Rosana, read this carefully, please! I think it may need review:
+%B: I changed nothing here, except: (1) a citealt<-citet to avoid double parentheses.
+(2) sp.
+disscussed -> discussed, (3) "If that is the case ON a system" -> "...
+IN a system"?
+(4) "hight type I errors" -> high, type I errors [comma], (5) line break in penultimate sentence.
+Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as discussed in  \citealt{rabosky_2010, beaulieu_2015}), but also, a problem of inadequate  ``*SSE'' model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focus trait first used to model diversification.
 We presented a roadmap with a series of analyses that can identify whether unknown or unobservable traits are worth pursuing. 
 In this roadmap, we have proposed fitting HiSSE models that have effectively shown that some binary states associated with diversification under BiSSE analyses are actually not different in net diversification terms, and instead a hidden state is linked to differences in the diversification rate \citep{beaulieu_2016}.
-If that is the case on a system, raising the question of the identity of the hidden state should be considered as the next step in the modeling process. 
+If that is the case in a study system, raising the question of the identity of the hidden state should be considered as the next step in the modeling process. 
 When considering a second candidate trait, one can model the complex interactions via a MuSSE model, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
-Since MuSSE might also suffer from hight type I errors considering more heterogeneity using a hidden state model on top of the multistate diversification model is necessary to adequately infer the effect of the complex interactions in the speciation and extinction process.
+Since MuSSE might also suffer from high type I errors, considering more heterogeneity using a hidden state model on top of the multi-state diversification model is necessary to adequately infer the effect of the complex interactions in the speciation and extinction process.
 % R- new lines
 Investigating the process of diversification linked to traits can be done via thorough statistical inferences that carefully evaluate multiple evolutionary histories. 
-Background heterogeneity in lineage of diversification is the rule rather than the exception.
+Background heterogeneity in lineage of diversification is the rule, rather than the exception. 
 Inferences that question whether other complex trait interactions are hidden in this background heterogeneity are necessary because they critically contribute to the reconstruction and understanding of diversification.

### 2019-03-15 rev_analyses/hisse250K/hissenodipDPv1.sh
First push of Rev code

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=500mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o hisseoutput.log
+#PBS -e hisseerrorscluster.log
+#PBS -l walltime=96:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/hissenodip/hissenodipDP.Rev

### 2019-04-19 manuscript/response.tex
skeleton of response letter

\documentclass[11pt]{article}
\usepackage{evolution}
\setstretch{1}
\parindent 0pt
\parskip 8pt

\renewenvironment{quote}{\bigskip\noindent\itshape\ignorespaces}{\smallskip}

\begin{document}

\hfill \today

Dear editor and reviewers:

\medskip
Thank you very much for your close reading of our manuscript and supportive comments.
We appreciate your constructive suggestions, each of which we address below.

\medskip
Sincerely,

Rosana Zenil-Ferguson, on behalf of all authors
% E: Pardon my speaking for you, Rosana. :)  Obviously change whatever wording below---starting to draft responses just helped me think through the requests.

%--------------------------------------------------
\section{Editor}
%--------------------------------------------------
\vspace{-11pt}

\begin{quote}
Thanks for submitting your work to New Phytologist.
We have back the referee comments now, and in general all three referees found the work to be novel and interesting.
Congratulations on accomplishing such a nice study.

Of course, the referees also had a number of critical comments and suggestions for improving the manuscript and so I am suggesting that your paper be accepted with `major revision' due to the number and nature of the comments, some of which I agree with, and others perhaps not.
I will be interested to hear your reasoning for either accepting or not accepting these suggestions when you re-submit the revised manuscript.

Perhaps foremost among the issues that should be addressed in the revision is that the referees felt that readers are not likely to be familiar with the SSE-type models that incorporate hidden states.
More background information should be provided to make it clearer how the incorporation of hidden states alters the approach and especially the interpretation of results.
This becomes especially important for understanding the multistate models, and for later on, for following your reasoning and justification in reducing the number of parameter states in said models.
Related to the issue of providing useful background information, Referee \#2 suggests that readers who are less familiar with mating system function will benefit by having a clearer exposition of the causal links between polyploidization and the breakdown of SI, something else I too think would help the paper.

Both Referees \#2 and \#3 highlight the issue of symmetry in transition rates, and from what little I know about parameter estimation in these types of models, I am partial to the suggestion of this referee that such a constraint could influence the estimation of diversification rates.

Please also go though the other referee comments which when acted upon should improve the readability of the paper.

Best wishes,

Dan Schoen
\end{quote}

Thank you for assembling these substantive reviews.
We have provided more background information on interpreting the hidden state models, and on the relationship between polyploidization and breakdown of GSI.
These and other changes are detailed below.

%--------------------------------------------------
\section{Referee 1}
%--------------------------------------------------
\vspace{-11pt}

\begin{quote}
I think this is an excellent paper, both in its analytical approach, and in its biological insights.
In particular, I think the general approach---gradually adding complexity to the models, and comparing their inferences with those from simpler models and the literature---is very effective.
In addition, the use of hidden state and multistate SSE models provides a useful framework for other macroevolutionary studies. 
Most of my comments are specific things (grammar, etc.), although I think the paper would benefit from a slightly pulled-back concluding bit---what from these results do the authors think is broadly applicable, and what is likely specific to Solanaceae and S-locus incompatibility systems?
\end{quote}

Thank you for the kind words.
We have substantially revised the Discussion, including providing more perspective on the general versus specific findings. % TODO

\begin{quote}
Could the last two sentences of the abstract be made more concrete?
The phrases ``generates interesting outcomes'' and ``affects conclusions about \ldots different evolutionary pathways'' are both pretty vague (the authors do an impressive job of summarizing complex ideas in this abstract, and maybe it's as specific as possible given word number constraints etc).
\end{quote}

We improved this wording. % TODO

\begin{quote}
The introduction is broad, and generally well supported, but a little disjointed, and additional literature on the macroevolutionary consequences of each of polyploidy and breeding system could be included.
As well as their correlation with each other---this is a well-acknowledged problem.
Ln65 says this nicely, but this point could be introduced (with the appropriate literature cited) earlier.
Also, the introduction is explicitly limited to angiosperms, which doesn't seem necessary---other landplant lineages (and non-plants, for that matter) also have ploidy and breeding system variation.
\end{quote}

okay % TODO

\begin{quote}
--35 Is the Soltis et al.\ 2014 citation appropriate here?
That paper doesn't present much by way of new data
\end{quote}

check % TODO

\begin{quote}
--55 change ``effect disruption'' to ``disrupt''
\end{quote}

Fixed. % TODO

\begin{quote}
--57 vague pronoun -- ``this'' (as a general rule, don't use ``this'', ``that'' etc by themselves -- follow them with the noun that they're referring to)
\end{quote}

Fixed. % TODO

\begin{quote}
use commas before ``which''; if no comma, use ``that''
\end{quote}

check % TODO

\begin{quote}
--83 Nitpicky, but is it accurate to refer to this as a false-positive problem?
``False positive'' implies to me the erroneous rejection of the null, but in this case the situation is more of the null being mis-specified with respect to the question the researcher is after.
The null---no rate variation---is correctly rejected.
This section probably should include a reference to Rabosky and Goldberg.
\end{quote}

ugh, consider; cite B\&O % TODO

\begin{quote}
--115 I'm uncomfortable with this way of handling multiple cytotypes.
The most commonly reported cytotype could easily not be the most common cytotype in the wild (I assume that most taxa are represented by relatively few counts), and regardless, there is biology here.
Specifically, this methodology would artificially reduce the number of ploidy changes (if a taxon is polymorphic for ploidy, there's at least one event there), and the omitted changes would be disproportionately those towards the tips of the phylogeny, so not only would rates be depressed, but the bias is of a sort that the estimation of diversification parameters is particularly sensitive to (the authors will understand why this is so).
\end{quote}

Rosana? % TODO

\begin{quote}
--130 I don't follow this discussion.
There's a self-incompatible pentaploid?
How is it even sexual?
Regardless, the discussion of this pentaploid should be more clearly integrated back into the central argument here, about the absense of SI polyploids (I think that's the central argument?).
\end{quote}

Maybe skip more quickly over this weird case, if it is more of a distraction than an explanation?
Boris? % TODO

\begin{quote}
--133 change ``SI and polyploidy populations'' to ``polyploid SI populations'' (there are lots of polyploid populations, and lots of SI ones, too)
\end{quote}

Fixed. % TODO

\begin{quote}
--140 I lost track a little of how the data were coded -- maybe add an explicit statement as to how many tips had both forms of data, how many had only breeding system info, how many had only ploidy info (I see now, after the fact, that those data are included in the figure), and how missing data were handled
\end{quote}

Fixed. % TODO

\begin{quote}
--It's not intuitive to me what the effects of forcing the hidden state change rates to be symmetrical are.
Did the authors investigate this?
Or has it been explored in the literature?
Similarly, it seems to me that fixing the transition rates of the observed characters to be independent of the hidden state could do weird things---the point of these SSE models is that state change rates and state-dependent diversification rates need to be co-estimated
\end{quote}

Rosana?  Maybe one analysis with all rates different, to say there's so much uncertainty it doesn't add anything? % TODO

\begin{quote}
--221  Propagation
\end{quote}

Fixed. % TODO

\begin{quote}
--I would be more comfortable if each model was run at least twice, independently, to better assess whether the MCMC has indeed converged
\end{quote}

Ouch.  Rosana? % TODO

\begin{quote}
--the use of run time as a description of the analysis seems odd to me.
As a reader, I'm not interested in the power of a given computer -- I want to know how many generations the MCMC was run for, what the ESSs are, etc.
\end{quote}

Seems reasonable.  Rosana? % TODO

\begin{quote}
-325  These analyses (by accounting for diversification) may take care of this issue, but is there a problem with these results being driven by a non-stationary process being modeled as a stationary one?
\end{quote}

No? Rosana? % TODO

\begin{quote}
--340  Sentence fragment
\end{quote}

Fixed. % TODO

\begin{quote}
--beginning of the Discussion is perhaps unnecessary -- it's more like Introduction material
\end{quote}

consider % TODO

\begin{quote}
--are the results from the analyses that include diploidization coherent?
I'm not sure how to phrase this -- I guess I'm asking if there's any suggestion that that parameterization is indeed capturing the biological phenomenon of diploidization, or is it just soaking up noise?
\end{quote}

Interesting. Rosana? % TODO

\begin{quote}
--I think this is an excellent point -- ``SI, which is itself associated with a faster diversification rate, is the breeding system context in which many diploids occur.''
Perhaps an discussion of the extent to which these results are extendable to other plant groups would be appropriate, versus what is likely to be specific to this system
\end{quote}

Seems reasonable. % TODO

\begin{quote}
--366  I would use a more compelling descriptor than ``interesting'' to describe these traits
\end{quote}

Fixed. % TODO

\begin{quote}
--390  I don't fully follow.
Could this comparison be expanded upon?
And maybe avoid the parameter notation, and just describe the rates being compared.
\end{quote}

consider % TODO

\begin{quote}
--400  Didn't Mayrose et al.\ 2011 and Magnuson-Ford \& Otto 2012 look at this (cladogenic change)?
\end{quote}

yes % TODO

\begin{quote}
--408  In what way are the present results in conflict with other reports from the literature?
(It's not clear, at least at first, where this conflict lies)
\end{quote}

consider % TODO

\begin{quote}
--410  ``as self-incompatibility loci''?
\end{quote}

check % TODO

\begin{quote}
--417  Why would the rates for diploids be inflated?
So the diploids were ancestrally polyploidy -- why would that alter the rate inferences?
\end{quote}

consider % TODO

\begin{quote}
--430  remove the comma
\end{quote}

Fixed. % TODO

\begin{quote}
--436  use ndashes to indicate ranges (rather than hyphens)
\end{quote}

Fixed. % TODO

\begin{quote}
--437-8 I don't understand this sentence
\end{quote}

Rephrased. % TODO

\begin{quote}
--455  I don't quite follow where this section is leading -- I think these loose ends should be explicitly tied together.
There are reports of paleopolyploidy in Solanaceae and relatives, but there are reasons to be skeptical of them?
\end{quote}

consider % TODO

\begin{quote}
--457  differ by six..
\end{quote}

Fixed. % TODO

\begin{quote}
--467  focal trait
\end{quote}

Fixed. % TODO

%--------------------------------------------------
\section{Referee 2}
%--------------------------------------------------
\vspace{-11pt}

\begin{quote}
I really enjoyed reading the manuscript.
The subject is pertinent and interesting and the methodological approaches used are amongst the best available to answer such questions.
Polyploidy is well known to be evolutionarily correlated with self-incompatibility breakdown and their respective impact on diversification rates are best studied simultaneously.
The results presented show exactly this, but in addition show that some unknown hidden states could also affect diversification rates, suggesting that results must be interpreted with care.
These are very interesting results.
\end{quote}

We're pleased you found the results interesting!

\begin{quote}
I have two ``major'' comments.
The first is relatively easy to fix and is linked to the writing/structure of the paper.
I often had the impression that too much knowledge is assumed from the reader throughout the manuscript.
For instance, more information should be given in the intro regarding why self-incompatibility breakdowns with polyploidy (it is briefly explained, but only later in the manuscript).
Moreover, some methodological approaches could be better explained.
In particular, HISSE models are relatively recent and many readers may not know them well.
I thus think that it would be important to describe them better and also better explain the results (using biological terms) obtained.
I mention some examples below in the detailed notes where I think explanations could be added.
\end{quote}

We have rewritten parts of the manuscript, especially in the Introduction, to try to explain better the relationship between polyploidy and self-incompatibility, and the interpretations of hidden state models. % TODO

\begin{quote}
My second comment has to do with the parameterization of the models.
Overall, I think that the authors have done a great job in justifying their models and explaining them.
I think it is important to mention.
In addition, their implementation of the models in revBayes is noteworthy and although not stated precisely in the manuscript, I believe this represents a great scientific contribution. 
Nevertheless, I question the decision to restrict the transition rates to and from the hidden states to be all equal.
Indeed, we have no idea that could make us think that this would be the case.
The authors justify this by ``Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fit a simplified version with 16 parameters.''
I am not completely convinced by this argument.
First, constraining transition rates can have an impact on the estimation of diversification rates.
Second, specifically because the authors are not interested in estimating the rates themselves, they should not be too worried about increasing the number of parameters.
This is especially true in a Bayesian framework where it is sometimes better to have overparameterized models than under-parameterized models (see Lemmon and Moriarty. 2004. Systematic Biology 53:265–277).
I think it would have been important to at least present the results of an unconstrained model to show that this decision to constrain the rates do not affect the conclusions.
\end{quote}

Rosana? % TODO

\begin{quote}
Finally, I have a suggestion for the authors.
One thing that I would have like to see is a discussion regarding the distribution of the hidden states at the tip of the tree (e.g., Fig S6, S8).
The two peaks distributions obtained for the hidden and non-hidden states made me wonder if the high and low rate peaks could be restricted to specific clades on the phylogeny.
After inspecting the supplementary material, I see that it is not the case; hidden states are rather spread throughout the phylogeny.
I think this is interesting and that it could make an interesting discussion.
\end{quote}

Thanks for this suggestion, which we have incorporated in the Discussion. % TODO

% Minor points:

\begin{quote}
Line 60: Did the author mean in all polyploid examined to date?
\end{quote}

check % TODO

\begin{quote}
Line 70: Sentence not clear. Should explain better the results of Robertson.
\end{quote}

Fixed. % TODO

\begin{quote}
Line 87: Specify other sources. Cryptic
\end{quote}

Fixed. % TODO

\begin{quote}
Line 88: To which traits the ``these traits'' refer to?
\end{quote}

Fixed. % TODO

\begin{quote}
Line 133: Replace ``and polyploid populations'' by ``in polyploid populations''?
\end{quote}

Fixed. % TODO

\begin{quote}
Line 133-135: Should probably move this information in the intro.
\end{quote}

consider % TODO

\begin{quote}
Line 152: It would be useful to refer to the figures that illustrate the models in the methods.
It would help to understand them.
\end{quote}

Fixed. % TODO

\begin{quote}
Line 160: ``Simultaneous consideration of breeding system'' and what?
Polyploidy, hidden state?
The sentence is not clear
\end{quote}

Fixed. % TODO

\begin{quote}
Line 163: ``These results indicate that some of the effects of a hidden trait can be explained by each of the two traits under consideration''.
Not sure I understand this sentence. Is it possible to clarify?
\end{quote}

Fixed. % TODO

\begin{quote}
Line 369: The authors should explain in more details the link between the breakdown of SI and the evolution of gender dimorphism.
\end{quote}

Fixed. % TODO

\begin{quote}
Line 392: Remove one ``explained''.
\end{quote}

Fixed. % TODO

\begin{quote}
Line 392: ``our findings suggest that the effect of ploidy''.
Add ``on diversification rates'' after these words?
\end{quote}

Fixed. % TODO

\begin{quote}
Line 416: ``Unduly ignoring secondary diploidization would necessarily underestimate the rate of transitions from polyploids to diploids.''
Actually, if I understand well, models without diploidization simply fix this rate of transition from polyploidy to diploids at 0.
Am I right?
\end{quote}

consider % TODO

%--------------------------------------------------
\section{Referee 3}
%--------------------------------------------------
\vspace{-11pt}

\begin{quote}
Zenil-Ferguson et al.\ apply sophisticated modeling techniques to test for the combined impact that breeding system and ploidy have on diversification rates within Solanaceae.
These models used here include ``hidden'' states to account for the many additional unobserved trait that also impacts diversification elsewhere in the tree.
The authors find a strong correlation between ploidy and breeding system and that certain character combinations have an impact diversification.
Specifically, the authors find that as long as a lineage is diploid, being either self-compatible or incompatible result in higher diversification rates than if these breeding systems were associated with polyploids.
The main finding of diploids generally having higher diversification rates than polyploids (including a slight negative diversification rate for polyploids) is consistent with previous studies that have explored patterns of diversification in relation to ploidy in other lineages of angiosperms.
Overall, this is a very nice study, the manuscript very well written, and the results are interesting.
There is no question that this work will be of great interest to the readership at New Phytologist.

I have a few comments that I hope the authors will find useful.
\end{quote}

Thank you for this positive summary.

\begin{quote}
The first relates to the use of the hidden states in the models.
I agree completely with the rationale and meaning of including hidden states.
However, through my reading of the manuscript, I interpret the authors as using them simply to account for variation within a particular character state combination.
In other words, the authors are working under an assumption of trait-dependent diversification, with the hidden states simply providing a more nuanced view of the process.
But the authors are not considering a completely alternative view, which is that none of these trait combinations are strong drivers of diversification and that the shifts being inferred are simply coincidental.
Such an interpretation could be ruled out by explicitly fitting a model of trait-independent diversification.
This is achieved by linking the diversification rate parameters among the different character state combinations -- i.e., $ID_A=CD_A=CP_A$, and $ID_B=CD_B=CP_B$.
In other words, because the diversification rate differences are completely linked among the possible states, but unlinked by the hidden states, this forces an explicitly trait-independent model.
The fit of such a model could be included in the set of models in Table 1 and compared using Bayes Factors.
Of course, admittedly, I am not familiar with revBayes and how the model space is set up, so it's possible this issue is already being addressed.
But, if so, some clarity in the Methods/Results is warranted in how trait-independent models were evaluated (or even ruled out) in the analyses presented.
\end{quote}

Interesting. Rosana? % TODO

\begin{quote}
Along these same lines, how does revBayes deal with the symmetry problem with regards to the hidden states?
For example, imagine for one sample in the chain the speciation rate for $CP_A=0.1$ and $CP_B=0.05$, all else being equal.
Now, imagine in another sample the speciation rate parameters are exactly the same, but this time $CP_A=0.05$ and $CP_B=0.1$. 
Strangely, this would produce an identical likelihood as the previous sample (again, all else being equal), because A and B are really interchangeable.
This would, however, cause issues with how the posterior is summarized.
I should point out that nothing in the plots suggests this might be an issue, but this was something that I wondered about with the particular models used here.
\end{quote}

Need a clear explanation. % TODO

\begin{quote}
Finally, while I agree that the inclusion of ``hidden'' states provides a means for testing intuitions about a particular character, they can also ``soak'' up the variation in diversification rates due to various errors in the dataset (i.e., error in topology, branch lengths, etc.).
We just don't know.
In any case, it might be good to provide readers with this perspective when discussing the meaning of the hidden states in the context of the results (e.g., beginning on line 465).
I think the point about determining whether it is worth the effort to understand the meaning of the hidden states is a good one, but I also it worth keeping in mind the caveats that go with it.
\end{quote}

consider % TODO

% Other minor comments:

\begin{quote}
Line 392: Repeat of ``explained''
\end{quote}

Fixed. % TODO

\begin{quote}
Line 403-404: Would it be appropriate to cite Freyman and Hoehna (2017; Systematic Biology), and possibly Caetano et al (2018; Evolution)?
[Note, I am aware that Freyman is an author and would know if it is appropriate to cite his own paper here, but still\ldots]. 
Though I imagine the authors are referring to a hidden cladogenetic model that allows more than just a binary trait.
But, both of these papers, as far as I can tell, extend the basic SSE model to include both hidden states and cladogenetic events.
\end{quote}

check % TODO

\begin{quote}
Line 465-469: This sentence was a bit difficult to grasp during my first reading of it.
I get the point, but I think the authors might consider revising it for clarity.
\end{quote}

Rephrased. % TODO

\begin{quote}
Line 476-480: The type I error issues reported by Rabosky and Goldberg (2015) are, in my view, fairly universal.
So, I think it is safe to say that MuSSE ``does suffer'' issues of model misspecification/type I error.
Also, it is worth noting that there are several implementations that already include hidden states in a MuSSE model, most notably the HORRIBLY named SecSSE model, as well as the models used in a recent empirical paper on diatoms by Nakov et al (which I’ve seen on bioRxiv, though might be published by now).
\end{quote}

refs % TODO

\end{document}

### 2019-04-21 manuscript/response.tex
editing to respond to reviewers: Introduction
main changes are about why P breaks SI, and SSE models

@@
 
-We improved this wording. % TODO
+We improved this wording.
 
@@
 
-Fixed. % TODO
+Fixed.
 
@@
 --57 vague pronoun -- ``this'' (as a general rule, don't use ``this'', ``that'' etc by themselves -- follow them with the noun that they're referring to)
+
+--use commas before ``which''; if no comma, use ``that''
 \end{quote}
 
-Fixed. % TODO
-
-\begin{quote}
-use commas before ``which''; if no comma, use ``that''
-\end{quote}
-
-check % TODO
+We reworded this whole paragraph, as part of better explaining how polyploidization disables gametophytic SI.
 
@@
 
-ugh, consider; cite B\&O % TODO
+We reworded this whole paragraph, as part of better explaining the SSE family of models.
 
@@
 We have rewritten parts of the manuscript, especially in the Introduction, to try to explain better the relationship between polyploidy and self-incompatibility, and the interpretations of hidden state models.
-% TODO
+% E: I labeled the two paragraphs that I modified heavily.
 
@@
 
-check % TODO
+We reworded this whole paragraph, as part of better explaining how polyploidization disables gametophytic SI.
 

### 2019-04-21 manuscript/introduction.tex
editing to respond to reviewers: Introduction
main changes are about why P breaks SI, and SSE models

@@
 Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and crucially affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
-In particular, self-incompatibility (SI) causes plants to reject their own pollen, and loss of self-incompatibility, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974,igic_2008}.
+In particular, self-incompatibility (SI) causes plants to reject their own pollen, and loss of self-incompatibility, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974, igic_2008}.
 Previous analyses reported higher rates of diversification for SI than for SC lineages in Solanaceae \citep{goldberg_2010}. 
@@
 
+% E: re-worked, to explain better for R2
 Ploidy, for instance, is widely thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
 Whole genome duplication may help facilitate the loss of SI and the transition to selfing \citep{barringer2007, barrett2008, husband2008}, or it may directly cause the loss of SI \citep{stout1942, lewis1947}.
-Doubled number of alleles in pollen is thought to effect disruption of the genetic mechanisms in gametophytic SI systems, which prevent self-fertilization \citep{entani1999, tsukamoto2005, kubo2010}. 
-With exceedingly few exceptions \citep{hauck_2002,nunes_2006}, this generally precludes the existence of SI polyploids in the most widespread system of SI, which is regarded as ancestral in eudicots \citep{igic_2001,steinbachs_2002}.
-This system, RNase-based SI, is expressed in all SI species of Solanaceae examined to date.
+In gametophytic SI systems, self pollen fails because a pollen grain with one S-allele is unable to detoxify the RNase produced by the same allele in the style. % ref
+The pollen of polyploids, however, has multiple S-alleles which act complementarily, detoxifying the RNase produced by any maternal genotype \citep{entani1999, tsukamoto2005, kubo2010}.
+Polyploids are consequently capable of self-fertilization, with exceedingly few exceptions \citep{hauck_2002, nunes_2006}.
+This system of RNase-based SI is regarded as ancestral in eudicots \citep{igic_2001,steinbachs_2002}, and it is expressed in all SI species of Solanaceae examined to date.
 There is consequently a strong correlation between polyploidy and SC in this family \citep{robertson_2011}.
 
-%B: removed First, ... Second, ... because we had three consecutive counting paragraphs.
 Given that changes in ploidy and breeding systems are causally related, and may have profound effects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
@@
 
-%B: removed First, ... Second, ... because we had three consecutive counting paragraphs.
-Two recent methodological developments enable disentangling the complex interactions between two or more traits and diversification.
-\citet{fitzjohn_2012} extended the state-dependent diversification ``*SSE'' modeling framework \citep{maddison_2007} to allow estimating the effect of an arbitrary number of states (enabling multi-state, MuSSE, instead of binary state, BiSSE, speciation and extinction models).
-This allows multiple states to represent trait combinations, particularly useful if one is interested in explicitly testing whether combinations of traits non-additively affect speciation or extinction.
-Soon thereafter, \citet{beaulieu_2016} proposed an extension of the *SSE framework by introducing a model with unspecified, or `hidden' traits (HiSSE model), which can affect the diversification process.
-Because variation in speciation and extinction rates may not be solely due to the focal trait, HiSSE simultaneously enables hidden states to explain such variation, and it reduces the elevated false-positive error rates.
-A joint application of these advances allows simultaneous consideration of the effects of multiple known traits on the diversification process, along with the presence of other unobserved states that can affect speciation and extinction rates.
+% E: re-worked, to explain better for R2
+The possible associations of multiple traits with lineage diversification can be tested with the SSE model framework.
+This style of modeling is named after the binary state speciation and extinction model \citep[BiSSE; ][]{maddison_2007}, which simultaneously models transitions between the two states of a binary trait and different rates of speciation and extinction associated with each of those states.
+Although very powerful, this approach is prone to reporting an association between a trait and lineage diversification when the data contain additional, unmodeled diversification rate heterogeneity \citep{rabosky_2015}.
+We use two techniques to address this concern, both employing the model extension to an arbitrary number of states \citep[MuSSE][]{fitzjohn_2012}.
+First, by considering both ploidy and breeding system simultaneously, we reduce the chance that an effect of one trait on diversification would be mistaken for the effect of the other.
+Second, by incorporating a hidden state, we allow that an unknown factor could additionally be influencing diversification \citep[HiSSE][]{beaulieu_2016}.
+Our analyses jointly apply different combinations of these techniques in order to separate the associations of lineage diversification rates with breeding system alone, ploidy alone, the interaction between breeding system and ploidy, and additional unobserved factors.
 

### 2019-04-25 manuscript/introduction.tex
ref update

@@
 Previous analyses reported higher rates of diversification for SI than for SC lineages in Solanaceae \citep{goldberg_2010}. 
-Similarly, heterostylous SI lineages in Primulaceae seem to diversify faster \citep{devos2014}, as do outcrossing lineages in Onagraceae \citep{freyman_2018}.
+Similarly, heterostylous SI lineages in Primulaceae seem to diversify faster \citep{devos2014}, as do outcrossing lineages in Onagraceae \citep{freyman_2019}.
 Despite consistent inferences of their macroevolutionary role, breeding systems are clearly not the sole trait determining diversification rates, and may exhibit complex interactions with other traits.

### 2019-04-25 manuscript/results.tex
ref update

@@
 When a hidden state is included (I/C+A/B model), the large net diversification rate difference persists for one of the hidden states but is removed for the other (\cref{figure:netdivall}D).
-Thus, differences in net diversification rates are best explained by both breeding system and an unknown factor \citep[as in Onagraceae;][]{freyman_2018}.
+Thus, differences in net diversification rates are best explained by both breeding system and an unknown factor \citep[as in Onagraceae;][]{freyman_2019}.
 % The transition rate from SI to SC is $q_{IC}=0.3$ for both the I/C and the I/C+A/B models.

### 2019-04-25 manuscript/response.tex
more reviewer responses
mostly Methods and Results
including revised text to explain the pathway findings

@@
 Maybe skip more quickly over this weird case, if it is more of a distraction than an explanation?
+We could also maybe revise the following sentences once we have more explanation of no-SI-P in the Intro.
 Boris? % TODO
@@
 
-No? Rosana? % TODO
+This was a point worth considering, but our model does not assume stationarity.
+In Fig.~4, the contributions of each pathway are functions of time.
+We can see there both the short-term and long-term behavior.
+% Rosana, what do you think?
 

### 2019-04-25 manuscript/introduction.tex
more reviewer responses
mostly Methods and Results
including revised text to explain the pathway findings

@@
 Alternatively, the path from SI diploids to SC polyploids may involve a transition to SC, without an increase in ploidy, and then a subsequent polyploidization. 
-\citet{robertson_2011} found that the pathway from SI diploids to SC polyploids is dominated by the two-step pathway over long timescales, but generally proceeds in one step via polyploidization of SI species over short timescales.
-It remains to be seen whether these results hold true when allowing for the influence of these traits on lineage diversification.
+\citet{robertson_2011} compared the contributions of these two paths, finding that evolution from SI diploid to SC polyploid is more likely to proceed via the one-step pathway over short timescales, but via the two-step pathway over long timescales.
+Their analysis considered only transitions among the states, however, so it remains to be seen whether these results hold true when allowing for the influence of these traits on lineage diversification.
 

### 2019-04-25 manuscript/results.tex
more reviewer responses
mostly Methods and Results
including revised text to explain the pathway findings

@@
 
-There are two pathways by which SI diploid lineages eventually---given enough time---become a SC polyploid.
+There are two pathways by which SI diploid lineages eventually---given enough time---become SC polyploids.
 In the one-step pathway, polyploidization directly disables SI.
@@
 Determining the relative contribution of these pathways based on our estimated transition rates from the ID/CD/CP model, we find that the one-step pathway is more likely on short timescales and the two-step pathway is more likely on long timescales (\cref{figure:pathways}, left panels).
-Beginning with a single SI diploid (ID) lineage, when not much time has elapsed, the one-step pathway is more likely because it only necessitates a single event to effect transition to SC polyploid state.
+Beginning with a single SI diploid lineage, when not much time has elapsed, the one-step pathway is more likely because it only necessitates a single event to effect transition to SC polyploid state.
 When more time has elapsed, the two-step pathway is more likely because the rate of loss of SI within diploids, $q_{IC}$, is greater than the rate of polyploidization for SI species, $\rho_I$ (\cref{suppfigure:IDCDCPnodip}).
-These conclusions, however, ignore the changes in numbers of lineages in a particular state from speciation and extinction events.
-By analogy, frequent sinking of a middle stepping stone in a path would lower the value of such a path.
-We therefore examined the effects of diversification rate on relative importance of each pathway.
-%B: revisit this possibly terrible analogy.
 
-Indeed, incorporation of lineage diversification rates changes conclusions regarding the relative contribution of each pathway.
-Beginning with a single ID lineage, assessment of relative pathway importance necessitates consideration of proliferation in the ID state, as well as the CD state for the two-step pathway, and to track the transitions to the CP state.
-In this situation, even over long timescales, the two-step pathway contributes less to generation of CP (\cref{figure:pathways}, right panels).
-The lower rate of net diversification in the CD state, relative to ID, means that relatively fewer lineages are available to complete the second step of the two-step pathway, ending in CP.
+The preceding conclusions, however, ignore the changes in numbers of lineages in each state due to speciation and extinction.
+By analogy, envisioning the states as stepping stones, the extent to which each stone grows or shrinks over time affects the utility of each possible path.
+Allowing for the different net diversification rates for each state, we find a qualitative difference in the relative pathway contributions.
+In particular, the lower rate of net diversification in the CD state, relative to ID, means that relatively fewer lineages are available to complete the second step of the two-step pathway, ending in CP.
+Consequently, even over long timescales, the two-step pathway contributes less to generation of CP (\cref{figure:pathways}, right panels).
 

### 2019-05-09 manuscript/results.tex
small note

@@
 
-When analyzed separately, ploidy and breeding system each showed significant associations with net diversification differences in Solanaceae.
-When analyzed together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well.
+% NOTE: will need to update figure/panel refs if making no-delta models primary
+% R: I am changing no delta as just D/P and prevous D/P as +delta
+
+% E: I brutally shortened the text here and hence put all the models into one section for the "diversification" question.  But I think I retained all the ideas that were written out before.
+% Much of what I removed was descriptions about which distributions are/aren't overlapping with one another.
+I think that's easier for the reader to see in the figures than to read in the text.
+But now the text says things like if some estimates are "different", so perhaps in Methods we should explain different = non-overlapping?
+% For the former model comparison section, I moved the key conclusions into the paragraphs here and the mechanical details (bayes factor definition) into Methods.
+
+Individual trait-dependent diversification inference, of ploidy and breeding system considered separately, each showed significant associations with net diversification differences. 
+When considered together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well.
 
 When considering ploidy alone (D/P model), we found a greater net diversification rate for diploids than for polyploids, in agreement with \citep{mayrose_2011, mayrose_2015}.
-This result holds with (\cref{figure:netdivall}A) or without (\cref{figure:netdivnodip}A) the diploidization parameter, although including diploidization shifts the net diversification rate of polyploids to be non-negative.
-Incorporating a hidden state in this model, however, removes the clear separation in diversification rate estimates between diploids and polyploids (D/P+A/B model; \cref{figure:netdivall}B, \cref{figure:netdivnodip}B).
-Thus, differences in net diversification rates are better explained by an unknown factor than by ploidy.
+This result holds with (\cref{figure:netdivall}A) or without (\cref{figure:netdivnodip}A) the diploidization parameter, though including diploidization shifts the net diversification rate of polyploids to be non-negative.
+Incorporating a hidden state, however, removes the clear separation in diversification between diploids and polyploids (D/P+A/B model; \cref{figure:netdivall}B, \cref{figure:netdivnodip}B).
+Thus, differences in net diversification are better explained by an unknown factor than by ploidy.
 Statistical model comparisons show very strong support for including the hidden state and strong support for including diploidization (\cref{table:bayesfactors}).
+% E: I removed the description of relative extinction results.
+We don't discuss it for other models or show it in the main text, and I never really know what to make of that compound parameter.
+But it's good that it's in the supp info.
 
 When considering breeding system alone (I/C model, \cref{figure:netdivall}C), we found a larger net diversification rate for SI than for SC species, in agreement with \citet{goldberg_2010}.
-When a hidden state is included (I/C+A/B model), the large net diversification rate difference persists for one of the hidden states but is removed for the other (\cref{figure:netdivall}D).
-Thus, differences in net diversification rates are best explained by both breeding system and an unknown factor \citep[as in Onagraceae;][]{freyman_2019}.
+%B: moved \cref{} earlier; avoids seeming related to \citep{}
+When a hidden state is included (I/C+A/B model), the large net diversification difference persists for one of the hidden states but is removed for the other (\cref{figure:netdivall}D).
+Thus, differences in net diversification are best explained by both breeding system and an unknown factor, similar to results found by \citet{freyman_2018}.
 % The transition rate from SI to SC is $q_{IC}=0.3$ for both the I/C and the I/C+A/B models.
@@
 When considering ploidy and breeding system together (ID/CD/CP model), the net diversification rate for SI diploids was greater than for either SC diploids or SC polyploids, with or without diploidization (\cref{figure:netdivall}E, \cref{figure:netdivnodip}C).
-The difference in net diversification with breeding system persists when ploidy is included in the model, but not the reverse.
-That is, the association of ploidy with net diversification in the D/P model (\cref{figure:netdivall}A, \cref{figure:netdivnodip}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids are similar.
+It thus appears that the difference in net diversification with breeding system persists when ploidy is included in the model, but not the reverse.
+The association of ploidy with net diversification in the D/P model (\cref{figure:netdivall}A, \cref{figure:netdivnodip}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids are similar.
 %
 When a hidden state is included (IC/CD/CP+A/B model), the same general pattern remains when diploidization is prevented (\cref{figure:netdivnodip}D), although the higher net diversification rate of ID is less clear within one of the hidden states.
-With diploidization, the net diversification rate of ID is still greater than CD within each hidden state, but diversification estimate for P is highly uncertain and perhaps bimodal.
+With diploidization, the net diversification rate of ID is still greater than CD within each hidden state, but diversification for P is highly uncertain and perhaps bimodal.
 Statistical model comparisons show very strong support for including the hidden state and at most positive support for including diploidization (\cref{table:bayesfactors}).
@@
 
-There are two pathways by which SI diploid lineages eventually---given enough time---become SC polyploids.
-In the one-step pathway, polyploidization directly disables SI.
-In the two-step pathway, SI is first lost within the diploid state, followed by polyploidization.
+Evolution from self-incompatible diploids to self-compatible polyploids can proceed through two different pathways.
 Determining the relative contribution of these pathways based on our estimated transition rates from the ID/CD/CP model, we find that the one-step pathway is more likely on short timescales and the two-step pathway is more likely on long timescales (\cref{figure:pathways}, left panels).
-Beginning with a single SI diploid lineage, when not much time has elapsed, the one-step pathway is more likely because it only necessitates a single event to effect transition to SC polyploid state.
+When not much time has elapsed, the one-step pathway is more likely because only one event need happen.
 When more time has elapsed, the two-step pathway is more likely because the rate of loss of SI within diploids, $q_{IC}$, is greater than the rate of polyploidization for SI species, $\rho_I$ (\cref{suppfigure:IDCDCPnodip}).
+These conclusions are the same as those of \citet{robertson_2011} under one of their branch length approximations.
 
-The preceding conclusions, however, ignore the changes in numbers of lineages in each state due to speciation and extinction.
-By analogy, envisioning the states as stepping stones, the extent to which each stone grows or shrinks over time affects the utility of each possible path.
-Allowing for the different net diversification rates for each state, we find a qualitative difference in the relative pathway contributions.
-In particular, the lower rate of net diversification in the CD state, relative to ID, means that relatively fewer lineages are available to complete the second step of the two-step pathway, ending in CP.
-Consequently, even over long timescales, the two-step pathway contributes less to generation of CP (\cref{figure:pathways}, right panels).
+When rates of lineage diversification are considered, however, the conclusions change.
+Even over long timescales, the two-step pathway contributes less (\cref{figure:pathways}, right panels).
+The lower rate of net diversification in the CD state, relative to ID, means that relatively fewer lineages are available to complete the second step of the two-step pathway.
 
@@
 
-We considered models both with and without diploidization in order to explore its effects on the estimates of state-dependent diversification.
+We considered models both with or without diploidization in order to explore the effects of this process on the estimates of state-dependent diversification.
 For the two models that only include diploid and polyploid states (D/P and D/P+A/B), the net diversification rate of polyploid lineages is likely negative when diploidization (parameter $\delta$) is excluded but positive when $\delta$ is included (\cref{figure:netdivall}A versus \cref{figure:netdivnodip}A).
-For the two models that also included breeding system (ID/CD/CP and ID/CD/CP+A/B), the main effect of including diploidization is greatly increasing the uncertainty of the estimated polyploid net diversification rate (\cref{figure:netdivall}E,F versus \cref{figure:netdivnodip}C,D).
-For all models, there is greater uncertainty in the estimate of diploidization rate than polyploidization rate, as judged by the width of the credibility intervals (see \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}).
+For the two models that also included breeding system (ID/CD/CP and ID/CD/CP+A/B), the main effect of including diploidization is increasing the uncertainty of the estimate of polyploid net diversification rate (\cref{figure:netdivall}EF versus \cref{figure:netdivnodip}CD).
+For all models, there is greater uncertainty in the estimate of diploidization rate than polyploidization rate, as judged by the width of the credibility intervals (see supplementary information figures).
+% todo: precise figure refs here
+% E: Rosana, would this also be a good place to say something about what we learned regarding the meaningfulness of diploidization in these data?
+%B: maybe meaning of the data could go to a dedicated Discussion subsection?
+The idea would be to reconcile variation in $\delta$ across (the best-fitting) models considered here, and how the implied values do (or do not) agree with chromosome number changes and WGD inference from tree reconciliation and Ks approaches.
+(I can help here, if needed.)
+
 % R will add a bit about param confidence
@@
 
-    \begin{tabular}{lcccccc}
-                       & D/P   & D/P+A/B      & I/C    & I/C+A/B      & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
+    \begin{tabular}{lcccc}
+                       & D/P   & D/P+A/B      & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
         %                                                              
-        $r_D$          & 0.260 & 0.193, 0.658 & ---    & ---          & ---      & ---          \\
-        $r_P$          &-0.056 & 0.030, 0.187 & ---    & ---          & ---      & ---          \\
-        $r_I$          & ---   & ---          & 0.550  & 0.386, 0.877 & ---      & ---          \\
-        $r_C$          & ---   & ---          &-0.001  &-0.059, 0.606 & ---      & ---          \\
-        $r_{ID}$       & ---   & ---          & ---    & ---          & 0.455    & 0.309, 0.797 \\
-        $r_{CD}$       & ---   & ---          & ---    & ---          & 0.065    &-0.006, 0.587 \\
-        $r_{CP}$       & ---   & ---          & ---    & ---          &-0.088    &-0.074, 0.403 \\
+        $r_D$          & 0.260 & 0.193, 0.658 & ---      & ---          \\
+        $r_P$          &-0.056 & 0.030, 0.187 & ---      & ---          \\
+        $r_I$          & ---   & ---          & ---      & ---          \\
+        $r_C$          & ---   & ---          & ---      & ---          \\
+        $r_{ID}$       & ---   & ---          & 0.455    & 0.309, 0.797 \\
+        $r_{CD}$       & ---   & ---          & 0.065    &-0.006, 0.587 \\
+        $r_{CP}$       & ---   & ---          &-0.088    &-0.074, 0.403 \\
         %                                                              
-        $\rho$         & 0.047 & 0.047        & ---    & ---          & ---      & ---          \\
-        $\rho_I$       & ---   & ---          & ---    & ---          & 0.067    & 0.053        \\
-        $\rho_C$       & ---   & ---          & ---    & ---          & 0.033    & 0.032        \\
-        $q_{IC}$       & ---   & ---          & 0.364  & 0.261        & 0.198    & 0.145        
+        $\rho$         & 0.047 & 0.047        & ---      & ---          \\
+        $\rho_I$       & ---   & ---          & 0.067    & 0.053        \\
+        $\rho_C$       & ---   & ---          & 0.033    & 0.032        \\
+        $q_{IC}$       & ---   & ---          & 0.198    & 0.145        
     \end{tabular}
@@
 
-    \begin{tabular}{lcccc}
-                       & D/P   & D/P+A/B      & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
+    \begin{tabular}{lcccccc}
+                       & D/P   & D/P+A/B      & I/C    & I/C+A/B      & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
         %                                                                                      
-        $r_D$          & 0.382 & 0.698, 0.100 & ---      & ---          \\
-        $r_P$          & 0.109 & 0.587, 0.182 & ---      & ---          \\
-        $r_I$          & ---   & ---          & ---      & ---          \\
-        $r_C$          & ---   & ---          & ---      & ---          \\
-        $r_{ID}$       & ---   & ---          & 0.449    & 0.318, 0.789 \\
-        $r_{CD}$       & ---   & ---          & 0.050    &-0.248, 0.494 \\
-        $r_{CP}$       & ---   & ---          &-0.027    & 0.110, 0.634 \\
+        $r_D$          & 0.382 & 0.698, 0.100 & ---    & ---          & ---      & ---          \\
+        $r_P$          & 0.109 & 0.587, 0.182 & ---    & ---          & ---      & ---          \\
+        $r_I$          & ---   & ---          & 0.550  & 0.386, 0.877 & ---      & ---          \\
+        $r_C$          & ---   & ---          &-0.001  &-0.059, 0.606 & ---      & ---          \\
+        $r_{ID}$       & ---   & ---          & ---    & ---          & 0.449    & 0.318, 0.789 \\
+        $r_{CD}$       & ---   & ---          & ---    & ---          & 0.050    &-0.248, 0.494 \\
+        $r_{CP}$       & ---   & ---          & ---    & ---          &-0.027    & 0.110, 0.634 \\
         %                                                                                      
-        $\rho$         & 0.033 & 0.026        & ---      & ---          \\
-        $\rho_I$       & ---   & ---          & 0.063    & 0.047        \\
-        $\rho_C$       & ---   & ---          & 0.024    & 0.011        \\
-        $\delta$       & 0.050 & 0.162        & 0.022    & 0.107        \\
-        $q_{IC}$       & ---   & ---          & 0.194    & 0.164        
+        $\rho$         & 0.033 & 0.026        & ---    & ---          & ---      & ---          \\
+        $\rho_I$       & ---   & ---          & ---    & ---          & 0.063    & 0.047        \\
+        $\rho_C$       & ---   & ---          & ---    & ---          & 0.024    & 0.011        \\
+        $\delta$       & 0.050 & 0.162        & ---    & ---          & 0.022    & 0.107        \\
+        $q_{IC}$       & ---   & ---          & 0.364  &              & 0.194    & 0.164        
     \end{tabular}
@@
         Two comma-separated numbers refer to the $A$ and $B$ hidden states, and --- means the parameter was not present in the model.
+    %B: perhaps include this in the table so lazies don't have to go seeking subscripts?
     Net diversification rates ($r$) are subscripted with trait state initials (Diploid, Polyploid, Incompatible, Compatible).
@@
 
+%B: FIXME:
+%B: Table S2 (above) The I/C and I/C+A/B columns should maybe appear in the top panel, no?
+%B: I did not flip the panels, because I realized just now that we'll first discuss no-delta, then delta. Is that okay?
+%
 % check values of delta and rho D/P+A/B with(out) delta
@@
 
-%B: heads up: I added two clarifying descriptive sentences in the two figures below. Maybe something about noting x-axis scale differences would be useful.
+% TODO: swap panels between the figures so that the first is -delta and the second is +delta
+
 \begin{figure}
-\includegraphics[width=\textwidth]{allnetdivmodelswithoutdelta.pdf}
-\caption{Net diversification rates for all models without diploidization rate estimate. 
-Each panel contains a graphical summary of estimated model parameters and displays posterior distributions for net diversification estimates.}
+\includegraphics[width=\textwidth]{allnetdivmodelswithoutdelta.pdf} % FIXME: The panel labels should use "+A/B" instead of "-A/B".
+Rosana, if you add the .svg file for this figure, I can make small changes like this.
+And need a horizontal axis label.
+\caption{Net diversification rates for all models.}  
 \label{figure:netdivall}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{allnetdivmodelsdiploidization.pdf}
-\caption{Net diversification rates for all models that include diploidization ($\delta$). 
-Each panel contains a graphical summary of estimated model parameters and displays posterior distributions for net diversification estimates.}  
+\includegraphics[width=\textwidth]{allnetdivmodelsdiploidization.pdf} % FIXME: same note as above about "+A/B"
+\caption{Net diversification rates for all models that include diploidization.}  
 \label{figure:netdivnodip}
@@
 \includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P ploidy model} % XXX
+\caption{Posterior distribution for each of the parameters in the D/P, polyploidy model} % XXX
 \label{suppfigure:DPnodip}
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDP.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node of the D/P ploidy model} % XXX
-\label{suppfigure:DPnodipasr}
-\end{suppfigure}
-
-\begin{suppfigure}
 \includegraphics[width=\textwidth]{bisseDPposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P+ $\delta$ ploidy model} % XXX
+\caption{Posterior distribution for each of the parameters in the D/P+ $\delta$,polyploidy model} % XXX
 \label{suppfigure:DP}
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDPdelta.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the D/P+ $\delta$, ploidy model} % XXX
-\label{suppfigure:DPasr}
-\end{suppfigure}
-
-\begin{suppfigure}
 \includegraphics[width=\textwidth]{hisseDPnodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P+A/B ploidy model} % XXX
+\caption{Posterior distribution for each of the parameters in the D/P+A/B, polyploidy model} % XXX
 \label{suppfigure:DPnodipAB}
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDPAB.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the D/P D/P+A/B ploidy model} % XXX
-\label{suppfigure:DPnodipABasr}
-\end{suppfigure}
-
-\begin{suppfigure}
 \includegraphics[width=\textwidth]{hisseDPposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P+$\delta$+A/B ploidy model.  The axis is offset in one location so that the two overlapping distributions can be seen.} % XXX
+\caption{Posterior distribution for each of the parameters in the D/P+$\delta$+A/B, polyploidy model.  The axis is offset in one location so that the two overlapping distributions can be seen.} % XXX
 \label{suppfigure:DPAB}
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDPABdelta.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the D/P+$\delta$+A/B ploidy model.} % XXX
-\label{suppfigure:DPABasr}
-\end{suppfigure}
-
-\begin{suppfigure}
 \includegraphics[width=\textwidth]{bisseSIposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the I/C breeding system model} % XXX
+\caption{Posterior distribution for each of the parameters in the I/C, breeding system model} % XXX
 \label{suppfigure:IC}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIC.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the I/C breeding system model} % XXX
-\label{suppfigure:ICasr}
 \end{suppfigure}
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrICAB.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the I/C+A/B breeding system model} % XXX
-\label{suppfigure:ICABasr}
-\end{suppfigure}
-
-\begin{suppfigure}
 \includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the ID/CD/CP ploidy and breeding system model} % XXX
+\caption{Posterior distribution for each of the parameters in the ID/CD/CP, polyploidy and breeding system model} % XXX
 \label{suppfigure:IDCDCPnodip}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCP.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP ploidy and breeding system model} % XXX
-\label{suppfigure:IDCDCPnodipasr}
 \end{suppfigure}
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCPdelta.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP+$\delta$ ploidy and breeding system model} % XXX
-\label{suppfigure:IDCDCPasr}
-\end{suppfigure}
-
-\begin{suppfigure}
 \includegraphics[width=\textwidth]{muhisseDPSInodipposteriordist.pdf}
@@
 \label{suppfigure:IDCDCPnodipAB}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCPAB.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP+A/B polyploidy and breeding system model} % XXX
-\label{suppfigure:IDCDCPnodipABasr}
 \end{suppfigure}
@@
 \end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCPdeltaAB.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP+$\delta$+A/B, polyploidy and breeding system model} % XXX
-\label{suppfigure:IDCDCPABasr}
-\end{suppfigure}

### 2019-05-09 manuscript/abstract.tex
- Submission #1 final changes - added cover letter and manuscript proofs

--- 2019-02-28 manuscript/abstract.tex
@@
 
-% E: Here's the previous abstract.  The new one below is aligned more with the current Introduction and Results.
-%
-% The effect of polyploidy in diversification remains a contentious issue.
-On the one hand,  recent studies  that found that polyploids have slower speciation rates and higher extinction rates than diploids left scientist wondering if polyploidy is truly an evolutionary dead-end.
-On the other hand, botanist have found strong molecular support of multiple polyploidy events at the root of highly diverse clades which challenges the evolutionary dead-end conclusions reached by modeling approaches.
-We re-investigate the role of polyploidy in speciation and extinction from a new modeling perspective considering that patterns found in diversification models can be misleading and incorrectly attributed to polyploidy when other observed and unobserved plant traits are responsible of shaping diversification.
-Using  statistically robust comparative phylogenetic approaches, we show that it is possible to  detect whether  the contribution of polyploidy to speciation and extinction is significant  under the presence of  other potential traits also affect diversification.
-We use the phylogeny, polyploidy, and breeding system data of 595 Solanaceae species to understand the contribution of polyploidy to diversification.
-We ask if Solanaceae polyploids are evolutionary dead-ends, and whether breeding system or some other unobserved traits are responsible of the patterns of diversification observed in the phylogeny.
-
-% E:  Help making this less bland?  Except for the first sentence, which may be too grandiose?
 If particular traits consistently affect rates of speciation and extinction, broad macroevolutionary patterns can be understood as consequences of selection at high levels of the biological hierarchy.
-Identifying traits associated with diversification rate differences is tricky, though, because of the many traits available to consider and the statistical challenge of testing for associations from phylogenetic data.
-Two traits that have been repeatedly suggested as drivers of differential diversification are whether a lineage is diploid or polyploid, and whether it is self-incompatible or self-compatible.
-We investigate the role of each of these traits, and their interaction, on speciation and extinction rates in Solanaceae.
+Identifying traits associated with diversification rate differences is tricky, however, because of the wide variety of traits to consider and the statistical challenges of testing for associations from comparative phylogenetic data.
+Ploidy (diploid vs.\ polyploid states) and breeding system (self-incompatible vs.\ self-compatible states) have been repeatedly suggested as possibly drivers of differential diversification.
+We investigate the roles of these traits, including their interaction, on speciation and extinction rates in Solanaceae.
 We find that the effect of ploidy can largely be explained by its correlation with breeding system, and that additional unknown factors work with breeding system to determine diversification rates.
-These results are largely robust to assumptions about whether diploidization occurs.
-Finally, we show that allowing for state-dependent diversification affects conclusions about the relative contribution of different evolutionary pathways to self-compatible polyploids.
+These results are largely robust to allowing for diploidization.
+Finally, we show that the lower rate of diversification for self-compatible diploids makes it less likely for that state to lie along the evolutionary pathway from self-incompatible diploids to polyploids.

### 2019-05-09 manuscript/discussion.tex

@@
 
-Previous analyses of the effects of ploidy on diversification showed that diploids had greater net diversification rates than polyploids when exploring across multiple clades of the angiosperm phylogeny \citet{mayrose_2011, mayrose_2015}.
-We obtain a broadly similar outcome, when considering only the effects of ploidy, but the results change significantly with joint consideration of hidden states, and breeding systems (\cref{figure:netdivall}A and B).
-The addition of a hidden trait results in blurring of differences in diversification rates between diploids and polyploids.
+Previous analyses of the effects of ploidy on diversification showed that diploids had greater net diversification rates than polyploids across multiple clades of the angiosperm phylogeny \citep{mayrose_2011, mayrose_2015}.
+We obtain a broadly similar outcome when considering only the effects of ploidy, but the results change significantly with joint consideration of hidden states and breeding systems (\cref{figure:netdivall}A and B).
+The addition of a hidden trait blurs differences in diversification rates between diploids and polyploids.
 Simultaneous consideration of breeding system better explains the variation of the inferred net diversification rates among ploidy states. 
@@
 % qualifying paragraph for pathways
-We implicitly modeled trait changes in both of these pathways as anagenetic, transitions occurring within lineages, without associated speciation events.
+We modeled trait changes in both of these pathways as anagenetic, transitions occurring within lineages, without associated speciation events.
 However, losses of SI (I$\rightarrow$C transitions) are often associated with cladogenesis in Solanaceae \citep{goldberg_2012}, and polyploidization (D$\rightarrow$P transition) also seem to co-occur with speciation events in a number of other families \citep{zhan_2016,freyman_2017}.
@@
 First, the Ks-inferred duplication likely pre-dates the origin of the family (49 My, HPD 46-53 My \citealt{sarkinen_2013}). 
-The best-supported, most-recent WGD \citet{tomato2012} is older than the lineage under consideration in this study.
+The best-supported, most-recent WGD \citep{tomato2012} is older than the lineage under consideration in this study.
 At best then, the species we consider are ancestrally diploid, having re-diploidized after a suspected WGD, but likely before the origin of the family.
@@
 As well, chromosome numbers and genome comparisons within the family (esp. `x=12' clade, containing \textit{Solanum} and \textit{Nicotiana}) reveal strong conservation.
-\citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca.
-10 inversions and 11 translocations (likely underestimated).
+\citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca.\ 10 inversions and 11 translocations (likely underestimated).
 Recovery of such simple relationships would require outstanding convergent loss of duplicated segments.
 In either case, all of the approaches make ample assumptions, and it seems that at least some of them will necessitate deep revisions.
-Our study, for example, made numerous assumptions in coding trait states (e.g.
-what constitutes a polyploid or diploid, SI or SC species), in models used for analyses (e.g.
-irreversibility of SI loss underlain by trans-specific orthologous S-loci), and many more, which could each mislead inference.
+Our study, for example, made numerous assumptions in coding trait states (\eg what constitutes a polyploid or diploid, SI or SC species), in models used for analyses (\eg irreversibility of SI loss underlain by trans-specific orthologous S-loci), and many more, which could each mislead inference.
 %B: -> %R: Rosana, take a look at this.
@@
 (4) "hight type I errors" -> high, type I errors [comma], (5) line break in penultimate sentence.
-Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as discussed in  \citealt{rabosky_2010, beaulieu_2015}), but also, a problem of inadequate  ``*SSE'' model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focus trait first used to model diversification.
+Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as discussed in  \citealt{rabosky_2010, beaulieu_2015}), but also, a problem of inadequate  ``*SSE'' model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focal trait first used to model diversification.
 We presented a roadmap with a series of analyses that can identify whether unknown or unobservable traits are worth pursuing. 
@@
 If that is the case in a study system, raising the question of the identity of the hidden state should be considered as the next step in the modeling process.
+% E: I would say not necessarily.
+There could be other non-trait weirdnesses.
 When considering a second candidate trait, one can model the complex interactions via a MuSSE model, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.

### 2019-05-09 manuscript/methods.tex

@@
 For example, \textit{Solanum betaceum} did not have information about ploidy level but it has 24 chromosomes, so since $x=12$ is the base chromosome number of the genus \textit{Solanum} \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid.
-Additionally, because of the absence of SI polyploids (discussed further below), species known to be SI could be scored as diploid.
+Additionally, because of the absence of SI polyploids (explained above and below), species known to be SI could be scored as diploid.
 Species with more than one ploidy level were assigned the most frequent ploidy level recorded or the smallest ploidy in case of frequency ties.
@@
 Breeding system was scored as self-incompatible (I) or self-compatible (C) based on results curated from the literature and  original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
-Many species could unambiguously be coded as either I or C \citep{raduski_2012}.
+Most species could unambiguously be coded as either I or C \citep{raduski_2012}.
 Following previous work, we coded any species with a functional SI system as I, even if SC or dioecy was also reported.
@@
 Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain SI, so long as they express a single allele at the S-locus \citep[discussed in][]{robertson_2011}.
-Because of the resulting absence of SI and polyploid populations, as well as the linked functional explanation for disabling of gametophytic self-incompatibility systems with non-self recognition, following whole genome duplication \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).
+Because of the resulting absence of polyploid SI populations, as well as the linked functional explanation for disabling of gametophytic self-incompatibility systems with non-self recognition, following whole genome duplication \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).
 %B: I added the non-self recognition bit, because in Prunus, heteroallelic pollen may not cause SC.
@@
 Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 595 species with ploidy and/or breeding system information on the tree.
-Binary or three-state classification of ploidy and breeding system for the 595 taxa is summarized in \cref{figure:stateclassifications}.
-We retained all of these species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
+Of these, 290 had information for both states.
+The number of species in each combination of states is summarized in \cref{figure:stateclassifications}.
+We retained all 595 species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
 A total of 405 taxa without any information about breeding system or ploidy were excluded.
@@
 In particular, if the net diversification rate is small for CD, the two-step pathway will contribute relatively less.
-We therefore include the difference between speciation and extinction along the diagonal elements of the propogation matrix.
+We therefore include the difference between speciation and extinction along the diagonal elements of the propagation matrix.
 As before, matrix exponentiation provides the relative chance of changing from one state to any other state after time $t$.
@@
 % E & R todo later: for the case with net div, write out the propagation matrix and equations for pathway contributions
-%B: I corrected all instances of "propogation" in text and comments
 

### 2019-05-13 basicdata/check.R
comparing old data with current codings

@@ -1,0 +1,65 @@
+### Compare breeding system data from 2014 compilation with the current file.
+### (Don't worry about uncertain/polymorphic species for now.)
+
+library(ape)
+phy <- read.nexus("sarkinen1.nex")  # just the first of the 100 trees
+
+#--------------------------------------------------
+# Data files from 2014
+#--------------------------------------------------
+
+dat1 <- read.csv("statesICDP.csv", as.is=T)
+dat1 <- subset(dat1, Species %in% phy$tip.label)
+
+dat1 <- subset(dat1, !(ID==CD & CD==CP)) # remove 0.33 0.33 0.33
+dat1$SI <- dat1$ID
+dat1$SC <- dat1$CD + dat1$CP
+dat1 <- subset(dat1, SI != SC) # remove 0.5 0.5
+dat1[which(!(dat1$SI %in% c(0, 1))),] # Solanum_brevicaule Solanum_candolleanum Solanum_chacoense
+dat1 <- dat1[-which(!(dat1$SI %in% c(0, 1))),] # drop further uncertainty
+table(dat1$SI)
+table(dat1$SC)
+# 331 SC, 143 SI
+subset(dat1, Species %in% c("Capsicum_schottianum", "Capsicum_scolnikianum", "Lycium_californicum")) # for comparison with dat2, below
+
+# fully cleaned
+dat1 <- data.frame(name=dat1$Species, bs1=c("SC", "SI")[dat1$SI+1])
+
+#--------------------------------------------------
+# Current data file
+#--------------------------------------------------
+
+dat2 <- read.csv("solanaceaerecords.csv", as.is=T)
+dat2 <- subset(dat2, fullname %in% phy$tip.label)
+
+dat2 <- unique(dat2[,c("fullname", "selfincomp")])
+dat2 <- subset(dat2, selfincomp != "")
+dat2[which(duplicated(dat2$fullname)),]
+subset(dat2, fullname %in% c("Capsicum_schottianum", "Lycium_californicum")) # the Capsicum was coded as SC only in dat1; the Lycium is truly polymorphic
+
+# fully cleaned
+dat2 <- subset(dat2, !(fullname %in% c("Capsicum_schottianum", "Capsicum_scolnikianum", "Lycium_californicum")))
+names(dat2) <- c("name", "bs2")
+
+#--------------------------------------------------
+# Compare
+#--------------------------------------------------
+
+dat <- merge(dat1, dat2, all=T)
+
[…]

### 2019-05-14 manuscript/methods.tex
- addressed pentaploid exception in response.tex and methods.tex

@@
 Following the reasoning outlined in \citet{robertson_2011}, we closely examined the few species for which the merged ploidy and breeding system data indicated the presence of self-incompatible polyploids.
-Although SI populations frequently contain some SC individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find convincing data for a naturally occurring SI polyploid population.
-The single instance of an SI polyploid individual appears to be an allopentaploid hybrid of \textit{Solanum oplocense} Hawkes x \textit{Solanum gourlayii} Hawkes, reported by \citet{camadro_1981}.
-Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain SI, so long as they express a single allele at the S-locus \citep[discussed in][]{robertson_2011}.
+%B: Modified this section for post-review response:
+Although SI populations frequently contain some SC individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find convincing data for a naturally occurring SI polyploid population  \citep[discussed in][]{robertson_2011}.
+%B rm: The single instance of an SI polyploid individual appears to be an allopentaploid hybrid of \textit{Solanum oplocense} Hawkes x \textit{Solanum gourlayii} Hawkes, reported by \citet{camadro_1981}.
+% Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain SI, so long as they express a single allele at the S-locus \citep[discussed in][]{robertson_2011}.
 Because of the resulting absence of polyploid SI populations, as well as the linked functional explanation for disabling of gametophytic self-incompatibility systems with non-self recognition, following whole genome duplication \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).
-%B: I added the non-self recognition bit, because in Prunus, heteroallelic pollen may not cause SC.
 

### 2019-05-16 manuscript/response.tex

@@
 
-Maybe skip more quickly over this weird case, if it is more of a distraction than an explanation?
-We could also maybe revise the following sentences once we have more explanation of no-SI-P in the Intro.
-Boris? % TODO
+%B: I edited the manuscript (methods.tex) as explained below:
+We removed the discussion and simply reference a more detailed treatment of this single individual.
+We agree that this observation by \citet{camadro_1981} may be distracting, and now instead  point the reader to our previously published paper \citep{robertson_2011}, where the possible existence of pentaploid that fails to set seed following self-pollination is discussed at some length.
+
+%Although SI populations frequently contain some SC individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find convincing data for a naturally occurring SI polyploid population  \citep[discussed in][]{robertson_2011}.
+%B rm: The single instance of an SI polyploid individual appears to be an allopentaploid hybrid of \textit{Solanum oplocense} Hawkes x \textit{Solanum gourlayii} Hawkes, reported by \citet{camadro_1981}.
+% Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain SI, so long as they express a single allele at the S-locus \citep[discussed in][]{robertson_2011}.
 

### 2019-05-16 manuscript/discussion.tex
fixed up more small points from reviewer 1

@@
 These results indicate that some of the effects of a hidden trait can be explained by each of the two traits under consideration. 
-The finding is unique, and implications may be of wide interest, because it evaluates the relative influence of these two exceptionally interesting traits in a comparative phylogenetic framework.
+The finding thus demonstrates how the relative influence of two traits on diversification can be disentangled in a comparative phylogenetic framework.
 
@@
 We modeled trait changes in both of these pathways as anagenetic, transitions occurring within lineages, without associated speciation events.
-However, losses of SI (I$\rightarrow$C transitions) are often associated with cladogenesis in Solanaceae \citep{goldberg_2012}, and polyploidization (D$\rightarrow$P transition) also seem to co-occur with speciation events in a number of other families \citep{zhan_2016,freyman_2017}.
-Models that fail to consider such cladogenetic changes may make misleading estimates of anagenetic transition rates.
-In the only available data bearing on this problem, \citet{goldberg_2012} found that extending the *SSE framework to allow cladogenetic changes did not substantially affect inference of net diversification rates.
+However, losses of SI (I$\rightarrow$C transitions) are often associated with cladogenesis in Solanaceae \citep{goldberg_2012}, and polyploidization (D$\rightarrow$P transition) also seem to co-occur with speciation events in a number of other families \citep{zhan_2016, freyman_2017}.
+Methods do exist to separate anagenetic and cladogenetic changes \citep{mayrose_2011, goldberg_2012, magnuson-ford_2012}, though \citet{goldberg_2012} found that extending the *SSE framework to allow cladogenetic changes did not substantially affect inference of net diversification rates.
 Nevertheless, a distinct possibility remains that allowing for cladogenetic transitions could tip the scales further in favor of the one-step pathway to SC polyploidy (CP), but this remains to be tested in future work.
@@
 %R- The line above is included in the bottom so no more need for that here
-The implications of the results of our diploidization analyses, as well as much of the other recent work on WGDs, are interesting because they conflict with other lines of evidence, especially those concerning the observed simple synteny of genomes in Solanaceae, and patterns of evolution as self-incompatibility loci.
+The implications of the results of our diploidization analyses, as well as much of the other recent work on WGDs, are interesting because they conflict with other lines of evidence, especially those concerning the observed simple synteny of genomes in Solanaceae, and patterns of evolution at self-incompatibility loci.
 Flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}. 
@@
 A subsequent analysis of synteny between grape and \textit{Solanum} genomes, as well as the distribution between inferred paralogs within genomes of \textit{Solanum} (tomato and potato) each suggested that this lineage experienced a likely round of ancient genome duplication or triplication \citep{tomato2012}.
-The age of the peak of paralog Ks distances, is approximately 71$\pm$19 My \citep{tomato2012}. 
-If this is the case, then all of the genomes may have been subsequently re-diploidized, yielding the widespread and common chromosome numbers in this and related families, n=11-12, presently considered to be diploid \citep{robertson_2011}.
+The age of the peak of paralog Ks distances is approximately 71$\pm$19 My \citep{tomato2012}. 
+If this is the case, then all of the genomes may have been subsequently re-diploidized, yielding the widespread and common chromosome numbers in this and related families, n=11--12, presently considered to be diploid \citep{robertson_2011}.
 
 On the other hand, there is little evidence for the occurrence of re-diploidization within Solanaceae, since the origin of the family---the time frame we considered.
-First, the Ks-inferred duplication likely pre-dates the origin of the family (49 My, HPD 46-53 My \citealt{sarkinen_2013}). 
-The best-supported, most-recent WGD \citep{tomato2012} is older than the lineage under consideration in this study.
+First, the Ks-inferred duplication likely pre-dates the origin of the family (49 My, HPD 46--53 My \citealt{sarkinen_2013}). 
+The best-supported, most-recent WGD \citep{tomato2012} is older than the clade under consideration in this study.
 At best then, the species we consider are ancestrally diploid, having re-diploidized after a suspected WGD, but likely before the origin of the family.
@@
 Most problematically in this context, it is unclear how to explain the maintenance of trans-generic polymorphism at the orthologous S-loci in this and other families.
-As well, chromosome numbers and genome comparisons within the family (esp. `x=12' clade, containing \textit{Solanum} and \textit{Nicotiana}) reveal strong conservation.
-\citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca.\ 10 inversions and 11 translocations (likely underestimated).
+As well, chromosome numbers and genome comparisons within the family (esp.\ `x=12' clade, containing \textit{Solanum} and \textit{Nicotiana}) reveal strong conservation.
+\citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ by six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca.\ 10 inversions and 11 translocations (likely underestimated).
 Recovery of such simple relationships would require outstanding convergent loss of duplicated segments.
@@
 (4) "hight type I errors" -> high, type I errors [comma], (5) line break in penultimate sentence.
-Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as discussed in  \citealt{rabosky_2010, beaulieu_2015}), but also, a problem of inadequate  ``*SSE'' model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focal trait first used to model diversification.
+Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as discussed in \citealt{rabosky_2010, beaulieu_2015}), but also a problem of inadequate ``*SSE'' model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focal trait first used to model diversification.
 We presented a roadmap with a series of analyses that can identify whether unknown or unobservable traits are worth pursuing. 

### 2019-05-16 manuscript/response.tex
citations in response letter now work
and no bibliography printed at the end
fixed up more small points from reviewer 1
fixed up small points from reviewer 2

@@
 \usepackage{evolution}
+\usepackage{bibentry}
 \setstretch{1}
@@
 
-yes % TODO
+Yes.
+We adjusted the text accordingly.
 
@@
 
-check % TODO
+Fixed.
 
@@
 
-Rephrased. % TODO
+Fixed.
 
@@
 
-Fixed. % TODO
+Reworded.
 
@@
 
-consider % TODO
+We now explain the effect of polyploidization on self-incompatibility more thoroughly in the Introduction.
 
@@
 \begin{quote}
-Line 160: ``Simultaneous consideration of breeding system'' and what?
+Line 360: ``Simultaneous consideration of breeding system'' and what?
 Polyploidy, hidden state?
@@
 
-Fixed. % TODO
+Fixed.
 
 \begin{quote}
-Line 163: ``These results indicate that some of the effects of a hidden trait can be explained by each of the two traits under consideration''.
+Line 363: ``These results indicate that some of the effects of a hidden trait can be explained by each of the two traits under consideration''.
 Not sure I understand this sentence. Is it possible to clarify?
@@
 
-Fixed. % TODO
+Reworded.
 
 \begin{quote}
-Line 369: The authors should explain in more details the link between the breakdown of SI and the evolution of gender dimorphism.
+Line 380: The authors should explain in more details the link between the breakdown of SI and the evolution of gender dimorphism.
 \end{quote}
 
-Fixed. % TODO
+We added a brief explanation.
 
@@
 
+\nobibliography{refs}
+\bibliographystyle{evolution}
+
 \end{document}

### 2019-05-16 manuscript/discussion.tex
fixed up small points from reviewer 2

@@
 The addition of a hidden trait blurs differences in diversification rates between diploids and polyploids.
-Simultaneous consideration of breeding system better explains the variation of the inferred net diversification rates among ploidy states. 
+The model with breeding system also included better explains the variation of the inferred net diversification rates among ploidy states. 
 SI, which is itself associated with a faster diversification rate, is the breeding system context in which many diploids occur.
 Among SC species, diploids and polyploids are associated with similar diversification rates (\cref{figure:netdivall}E).
-These results indicate that some of the effects of a hidden trait can be explained by each of the two traits under consideration. 
+These results indicate that the hidden trait in the analysis of one focal trait can in part be explained by the other focal trait.
 The finding thus demonstrates how the relative influence of two traits on diversification can be disentangled in a comparative phylogenetic framework.
@@
 This result is significant because it has a number of implications for the inference of history of breeding and sexual systems in flowering plants.
-First, support for the direct pathway lends credence to the idea that breakdown of SI by WGD---by generation of diploid `heteroalleleic' pollen---may often trigger the evolution of gender dimorphism across angiosperms \citep{miller_2000}.
+First, support for the direct pathway lends credence to the idea that breakdown of SI by WGD---by generation of diploid `heteroalleleic' pollen---may often trigger the evolution of gender dimorphism across angiosperms, as a different mechanism of inbreeding avoidance \citep{miller_2000}.
 A more direct and definitive test of this hypothesis would additionally examine the propensity of polyploids generated through either pathway to become dioecious \citep{robertson_2011}.
@@
 We nevertheless recover transition rates from ID to CD ($q_{IC}$) threefold greater than $\rho_{I}$, indicating that selection restricts the fixation rate of new polyploids compared with new SC mutants \citep{robertson_2011}.
-Overall, our findings suggest that the effect of ploidy is largely explained explained by its correlation with breeding system, and that other unknown factors work with breeding system to determine diversification rates.
+Overall, our findings suggest that the effect of ploidy on lineage diversification rate is largely explained by its correlation with breeding system, and that other unknown factors work with breeding system to determine diversification rates.
 

### 2019-05-16 manuscript/introduction.tex
fixed up small points from reviewer 2

@@
 Here, we examine the roles of ploidy and breeding system, as well as their interaction, on the diversification process in Solanaceae.
-Using extensive trait data, and considering other possible sources of heterogeneity in the diversification process, we assess the influence of these traits on speciation and extinction rates.
+Importantly, our analyses aim to parse out the associations of these two traits on speciation and extinction rates while simultaneously allowing that other processes could influence diversification as well.
 First, we present models in which the effect of ploidy and breeding system are considered separately, and we compare the inferences to previously published results. 

### 2019-05-16 manuscript/response.tex
fixed up small points from reviewer 3

@@
 So, I think it is safe to say that MuSSE ``does suffer'' issues of model misspecification/type I error.
-Also, it is worth noting that there are several implementations that already include hidden states in a MuSSE model, most notably the HORRIBLY named SecSSE model, as well as the models used in a recent empirical paper on diatoms by Nakov et al (which I’ve seen on bioRxiv, though might be published by now).
+Also, it is worth noting that there are several implementations that already include hidden states in a MuSSE model, most notably the HORRIBLY named SecSSE model, as well as the models used in a recent empirical paper on diatoms by Nakov et al (which I've seen on bioRxiv, though might be published by now).
 \end{quote}
 
-refs % TODO
+Yes, and indeed this has already been shown.
+We fixed the text and added a reference.
 

### 2019-05-16 manuscript/discussion.tex
fixed up small points from reviewer 3

@@
 When considering a second candidate trait, one can model the complex interactions via a MuSSE model, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
-Since MuSSE might also suffer from high type I errors, considering more heterogeneity using a hidden state model on top of the multi-state diversification model is necessary to adequately infer the effect of the complex interactions in the speciation and extinction process.
+Since MuSSE can also falsely associate traits with diversification \citep{fitzjohn_2012}, considering more heterogeneity using a hidden state model on top of the multi-state diversification model is necessary to adequately infer the effect of the complex interactions in the speciation and extinction process.
 % R- new lines

### 2019-05-22 manuscript/introduction.tex
editing throughout the Introduction
Mostly trying to sign-post better.  Reads well overall!

--- 2019-04-25 manuscript/introduction.tex
@@
 Species accumulate across the tree of life at different rates. 
-A possible explanation for this phenomenon is that species possess various traits or character states that differentially affect rates of diversification. 
-Dramatic increases in the available phylogenetic and phenotypic data, as well as methodological advances, have greatly accelerated the search for traits that influence diversification traits.
-Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a difficult problem \citep[\eg][]{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}.
-Problems arise at least in part because speciation and extinction may not depend on a single trait, and the context in which traits occur can lead to complex interactions resulting in a heterogenous speciation and extinction rates \citep{beaulieu_2016, caetano_2018, herrera_2018}.
+One possible explanation for this phenomenon is that the various traits or character states of species differentially affect rates of diversification. 
+Dramatic increases in phylogenetic and phenotypic data and methodological advances have greatly accelerated the search for traits that influence diversification traits.
+Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a challenge \citep[\eg][]{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}.
+One difficulty is that speciation and extinction likely do not depend on a single trait, so the biological and environmental contexts in which traits occur can lead to complex interactions that affect lineage diversification \citep{beaulieu_2016, caetano_2018, herrera_2018}.
 Consequently, examining the association of only one trait with diversification patterns can be misleading. 
+Here, we embrace this challenge by jointly investigating two traits thought to influence speciation and extinction rates---ploidy level and breeding system---while allowing for their interactions and other confounding factors.
+% E: We need a "Here, we..." here.  Otherwise, we've just told the reader "not one trait!" and then launch into one trait.
 
@@
 The widespread prevalence of variation in ploidy has long been considered a salient feature of flowering plant lineages \citep{stebbins1938}. 
-Increase in ploidy can alter many traits and impact a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
+Increase in ploidy can alter many traits and affect a variety of evolutionary and ecological processes \citep{ramsey_2002, sessa_2019}.
 %
-Polyploids are thought to have lower diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
-Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications (WGDs) have played an important macroevolutionary role driving innovation and diversification in plants.
-Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies pervasiveness of diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
-This could affect the interpretation of different net diversification rates for diploids and polyploids, especially if diploids have a relatively higher net diversification rate if they were ancestrally polyploid than if they had no polyploid ancestry.
+Polyploids have been documented to have an overall lower diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
+Other studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications have played an important macroevolutionary role driving innovation and diversification in plants.
+Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies pervasive diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
+Our analyses re-examine the association between ploidy and lineage diversification by extending the approach of \citep{mayrose_2011, mayrose_2015} to include these additional processes.
+We allow for transitions both toward and away from polyploidy, and for heterogeneous diversification within each ploidy level, for example if diploids have a relatively higher net diversification rate if they were ancestrally polyploid than if they had no polyploid ancestry.
 
 Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and crucially affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
-In particular, self-incompatibility (SI) causes plants to reject their own pollen, and loss of self-incompatibility, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974, igic_2008}.
+In particular, genetic self-incompatibility (SI) systems cause plants to reject their own pollen, and loss of self-incompatibility, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974, igic_2008}.
 Previous analyses reported higher rates of diversification for SI than for SC lineages in Solanaceae \citep{goldberg_2010}. 
 Similarly, heterostylous SI lineages in Primulaceae seem to diversify faster \citep{devos2014}, as do outcrossing lineages in Onagraceae \citep{freyman_2019}.
-Despite consistent inferences of their macroevolutionary role, breeding systems are clearly not the sole trait determining diversification rates, and may exhibit complex interactions with other traits.
+Although these findings suggest a consistent macroevolutionary role of breeding system, it is unlikely to be the sole trait determining lineage diversification.
+We thus investigate the relationship of breeding system to speciation and extinction rates in the context of ploidy and other background factors.
 
-% E: re-worked, to explain better for R2
-Ploidy, for instance, is widely thought to affect the propensity for self-fertilization \citep{stebbins1950}. 
-Whole genome duplication may help facilitate the loss of SI and the transition to selfing \citep{barringer2007, barrett2008, husband2008}, or it may directly cause the loss of SI \citep{stout1942, lewis1947}.
-In gametophytic SI systems, self pollen fails because a pollen grain with one S-allele is unable to detoxify the RNase produced by the same allele in the style. % ref
-The pollen of polyploids, however, has multiple S-alleles which act complementarily, detoxifying the RNase produced by any maternal genotype \citep{entani1999, tsukamoto2005, kubo2010}.
+Polyploidy and self-fertilization are widely observed to be associated \citep{stebbins1950}.
+There are several possible general explanations: whole genome duplication may facilitate the transition to selfing by masking inbreeding depression, self-fertilization may facilitate polyploidy establishment by avoiding the triploid bridge, or both selfing and polyploidy may be correlates of weedy or peripheral populations \citep{ramsey_1998, barringer2007, barrett2008, husband2008}.
+% Boris, can you please check?  We need to be totally convincing about no SI-P.
+Additionally, specifically in RNase-based gametophytic SI systems, polyploidization may directly cause the loss of SI \citep{stout1942, lewis1947}.
+In these systems, self pollen fails because a pollen grain with one S-allele at the locus controlling the SI response is unable to detoxify the RNase produced by the same S-allele in the style. % ref?
+The pollen of polyploids, however, has multiple S-alleles which act complementarily, detoxifying the RNases produced by any maternal genotype \citep{entani1999, tsukamoto2005, kubo2010}.
 Polyploids are consequently capable of self-fertilization, with exceedingly few exceptions \citep{hauck_2002, nunes_2006}.
 This system of RNase-based SI is regarded as ancestral in eudicots \citep{igic_2001,steinbachs_2002}, and it is expressed in all SI species of Solanaceae examined to date.
-There is consequently a strong correlation between polyploidy and SC in this family \citep{robertson_2011}.
+The consequent absence of SI polyploids in this family yields a strong correlation between ploidy and breeding system \citep{robertson_2011}.
 
-Given that changes in ploidy and breeding systems are causally related, and may have profound effects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
-If the effects of ploidy and breeding system on lineage diversification are not additive, studying each trait separately may not reveal the effects of their combinations.
-As a result, it may be of interest to establish their joint influence on speciation and extinction rates.
-The two traits may also affect each other's state transitions.
-For example, losses of SI in diploids could be caused by polyploidization, resulting in a one-step pathway to polyploid SC species. 
-Alternatively, the path from SI diploids to SC polyploids may involve a transition to SC, without an increase in ploidy, and then a subsequent polyploidization. 
+We address two macroevolutionary questions about the correlated evolution of ploidy and breeding system.
+First, we investigate their joint influence on rates of speciation and extinction.
+Each trait alone has been shown to associate with differential lineage diversification, but if their effects are not additive, studying each trait separately may not reveal the effects of their combinations.
+Second, we examine the order of transitions in the two traits.
+Evolution is predominantly from diploid to polyploid, and from SI to SC, but there are two paths by which diploid SI lineages can eventually become SC polyploids.
+Loss of SI in diploids could be directly caused by polyploidization (as explained above, for gametophytic SI systems), resulting in a one-step pathway to polyploid SC species. 
+Alternatively, SI diploids could first transition to SC without an increase in ploidy, and then undergo polyploidization, resulting in a two-step pathway.
 \citet{robertson_2011} compared the contributions of these two paths, finding that evolution from SI diploid to SC polyploid is more likely to proceed via the one-step pathway over short timescales, but via the two-step pathway over long timescales.
-Their analysis considered only transitions among the states, however, so it remains to be seen whether these results hold true when allowing for the influence of these traits on lineage diversification.
+Their analysis considered only transitions among the states, however, so we investigate whether these results hold true when allowing for the influence of these traits on lineage diversification.
 
-% E: re-worked, to explain better for R2
-The possible associations of multiple traits with lineage diversification can be tested with the SSE model framework.
-This style of modeling is named after the binary state speciation and extinction model \citep[BiSSE; ][]{maddison_2007}, which simultaneously models transitions between the two states of a binary trait and different rates of speciation and extinction associated with each of those states.
-Although very powerful, this approach is prone to reporting an association between a trait and lineage diversification when the data contain additional, unmodeled diversification rate heterogeneity \citep{rabosky_2015}.
-We use two techniques to address this concern, both employing the model extension to an arbitrary number of states \citep[MuSSE][]{fitzjohn_2012}.
+Undertaking these questions requires powerful phylogenetic comparative methods.
+We employ the SSE model framework, which simultaneously models transitions between the discrete states of a trait and different rates of speciation and extinction associated with each of those states \citep{maddison_2007, fitzjohn_2012}.
+Although potentially very powerful, this approach is prone to reporting a false association between a trait and lineage diversification when the data contain additional, unmodeled diversification rate heterogeneity \citep{maddison_2015, rabosky_2015}.
+We address this concern in two ways.
 First, by considering both ploidy and breeding system simultaneously, we reduce the chance that an effect of one trait on diversification would be mistaken for the effect of the other.
-Second, by incorporating a hidden state, we allow that an unknown factor could additionally be influencing diversification \citep[HiSSE][]{beaulieu_2016}.
-Our analyses jointly apply different combinations of these techniques in order to separate the associations of lineage diversification rates with breeding system alone, ploidy alone, the interaction between breeding system and ploidy, and additional unobserved factors.
+Second, by incorporating a hidden state, we allow that an unknown factor could additionally be influencing diversification \citep{beaulieu_2016}.
 
-Here, we examine the roles of ploidy and breeding system, as well as their interaction, on the diversification process in Solanaceae.
-Using extensive trait data, and considering other possible sources of heterogeneity in the diversification process, we assess the influence of these traits on speciation and extinction rates.
-First, we present models in which the effect of ploidy and breeding system are considered separately, and we compare the inferences to previously published results. 
-Second, we add a hidden trait to each of those models and investigate whether the focal trait still explains differences in diversification.
-Third, we model ploidy and breeding system jointly to assess their combined relationship to diversification, with or without an additional hidden trait.
+Our macroevolutionary analyses in Solanaceae apply different combinations of these techniques in order to parse out the associations of ploidy and breeding system on speciation and extinction rates, while simultaneously allowing that other processes could influence diversification as well.
+% separate the associations of lineage diversification rates with breeding system alone, ploidy alone, the interaction between breeding system and ploidy, and additional unobserved factors.
+% Here, we examine the roles of ploidy and breeding system, as well as their interaction, on the diversification process in Solanaceae.
+% Importantly, our analyses aim to parse out the associations of these two traits on speciation and extinction rates while simultaneously allowing that other processes could influence diversification as well.
+We first present models in which the effect of ploidy and breeding system are considered separately, and we compare the inferences to previously published results. 
+Then we add a hidden trait to each of those models and investigate whether the focal trait still explains differences in diversification.
+Finally, we model ploidy and breeding system jointly to assess their combined relationship to diversification, with or without an additional hidden trait.
 With this joint trait model, we also quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
 For all of the models involving ploidy, we investigate the potential effects of including diploidization.
-Our results highlight the importance of considering non-additive effects of traits on net diversification rates, especially when there are strong biologically driven correlations among them.
+Our results highlight the importance of considering non-additive effects of traits on net diversification rates, especially when there are strong biologically-driven correlations among them.

### 2019-05-23 manuscript/methods.tex
fixed up small points from reviewer 2
fig note
edits to Data

@@
 
-\subsection{Ploidy and breeding system data}
+\subsection{Data}
 
 %B: IMO: The data subsection would be better organized with parts, (1) ploidy, (2) breeding system, (3) tree, (4) models, (5) model selection and statistical inference.
+% E: I changed the subsectioning, but to a different arrangement: (1) Data, (2) Models, (3) Statistical inference, with subsubsections as necessary.
 
@@
 CCDB contains records from original sources that have multiple complex symbol patterns denoting multivalence, or irregularites of chromosome counts.
-After a first full round of curation via CCDBcurator, we contrasted results by hand and corrected chromosomes where CCDBcurator's output was not able to adequately correct records.
+After a first round of automatic cleaning, we examined results by hand and corrected records as necessary.
 Our hand-curated records were also contrasted against the ploidy dataset from \citet{robertson_2011} and against ploidy data in the C-value DNA dataset from \citet{bennett_2005}.
-%B: perhaps "cleaned" should be explained, unless it's hard to identically repeat as-is. R- I added more (it only took 4months to hand-curate 14,000 entries
-By comparing three different sources of information, we were able to code taxa as diploid (D) or polyploid (P).
+By comparing three different sources of information, we were able to code taxa as diploid, $D$, or polyploid, $P$.
 For the majority of species, ploidy was assigned according to information from the original publications included in the  C-value DNA dataset \citep{bennett_2005}.
@@
 
-Breeding system was scored as self-incompatible (I) or self-compatible (C) based on results curated from the literature and  original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
-Most species could unambiguously be coded as either I or C \citep{raduski_2012}.
-Following previous work, we coded any species with a functional SI system as I, even if SC or dioecy was also reported.
-Dioecious species without functional SI were coded as C.
+Breeding system was scored as self-incompatible, $I$, or self-compatible, $C$, based on results curated from the literature and original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
+Most species could unambiguously be coded as either $I$ or $C$ \citep{raduski_2012}.
+Following previous work, we coded any species with a functional SI system as $I$, even if SC or dioecy was also reported.
+Dioecious species without functional SI were coded as $C$.
 % E: I am trying to stick with I/C when talking about the state values, but using SI/SC when talking about the meaning of the trait (because this what people use normally). 
@@
 The Supplementary Information contains citations for the numerous original data sources. % todo: say how many refs?
-Resolution of taxonomic synonymy followed the conventions provided in Solanaceae Source \citep{solsource}. 
+Resolution of taxonomic synonymy followed Solanaceae Source \citep{solsource}. 
 Hybrids and cultivars were excluded because ploidy and breeding system can be affected by artificial selection during domestication.
@@
 % Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain SI, so long as they express a single allele at the S-locus \citep[discussed in][]{robertson_2011}.
-Because of the resulting absence of polyploid SI populations, as well as the linked functional explanation for disabling of gametophytic self-incompatibility systems with non-self recognition, following whole genome duplication \citep[reviewed in][]{ramsey_1998,stone_2002}, we consider only three observed character states: self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids which are always self-compatible (CP).
+Because of the resulting absence of polyploid SI populations, as well as the functional explanation for polyploidy disabling gametophytic SI systems with non-self recognition (see the Introduction), we consider only three observed character states: self-incompatible diploids $(ID)$, self-compatible diploids $(CD)$, and self-compatible polyploids $(CP)$.
 
@@
 
-% D/P ploidy model
-% D/P+A/B ploidy hidden state model
-% I/C breeding system model
-% I/C+A/B breeding system and hidden state model
-% ID/CD/CP ploidy and breeding system model
-% ID/CD/CP+A/B ploidy, breeding system, and hidden state model
+\subsection{Models}
 
-\subsection{Models for ploidy and diversification}
+% TODO: Refer to the model diagrams for each model? (request from referee 2)
+
+\subsubsection{Ploidy and diversification}
 
@@
 
-\subsection{Models for breeding system and diversification}
+\subsubsection{Breeding system and diversification}
 
@@
 
-\subsection{Models for ploidy, breeding system, and diversification}
+\subsubsection{Ploidy, breeding system, and diversification}
 
@@
 
-\subsection{Pathways to polyploidy}
+\subsubsection{Pathways to polyploidy}
 
@@
 
-\subsection{Diploidization as an exploratory hypothesis}
+\subsubsection{Diploidization as an exploratory hypothesis}
 
@@
 
-\subsection{Statistical inference under the models}
+\subsection{Statistical inference}
+
+\subsubsection{Model fitting}
 
@@
 
-\subsection{Model selection}
+\subsubsection{Model selection}
 

### 2019-05-23 manuscript/methods.tex
edits to Models

@@
 
-The Supplementary Information contains citations for the numerous original data sources. % todo: say how many refs?
 Resolution of taxonomic synonymy followed Solanaceae Source \citep{solsource}. 
@@
 
+The Supplementary Information contains citations for the numerous original data sources. % FIXME (and say here how many refs?)
+The Dryad archive contains the data and tree files used for analyses. % FIXME
+% \url{https://github.com/roszenil/solploidy}  E: we should use dryad instead, because github repos are not guaranteed to last
+
 \subsection{Models}
 
-% TODO: Refer to the model diagrams for each model? (request from referee 2)
+We employed a suite of SSE-style models \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016} to test associations between ploidy, breeding system, and lineage diversification.
+The full set of models is listed in \cref{table:marginallike}.
 
@@
 
-In order to investigate the association between ploidy level and diversification, we first defined a binary state speciation and extinction model (BiSSE, \citealt{maddison_2007}) in which taxa were classified as diploid (D) or polyploid (P).
-We call this the D/P ploidy model.
-The model parameters are speciation rates ($\lambda_D$, $\lambda_P$) and extinction rates ($\mu_D$, $\mu_P$) for each state, and a rate of transition from D to P ($\rho$).
-This is the same model fitted by \citet{mayrose_2011} but applied to the larger Solanaceae dataset.
+In order to investigate the association between ploidy level and diversification, we first defined a binary state speciation and extinction model \citep[BiSSE,][]{maddison_2007} in which taxa were classified as diploid $(D)$ or polyploid $(P)$.
+In this $D/P$ ploidy model, the model parameters are speciation rates ($\lambda_D$, $\lambda_P$) and extinction rates ($\mu_D$, $\mu_P$) for each state, and a rate of transition from D to P ($\rho$).
+This is the same model \citet{mayrose_2011} used for this question across an array of plant genera.
 % R- Itay wanted to hightlight that is the same model. Definitely a BiSSE. He did the ClaSSE later in the 2014 response.
@@
 
-Our second model addressed diversification due to ploidy differences while also parsing out the heterogeneity of diversification rates due a possible unobserved trait.
-Diversification rate differences explained by some trait other than ploidy were accommodated by adding a hidden state (HiSSE model; \citealt{beaulieu_2016}).
+Our second model addressed diversification due to ploidy differences while also parsing out heterogeneity of diversification rates due another factor, accommodated by adding a hidden state \citep[HiSSE,][]{beaulieu_2016}.
 %B: inserted "(trait)", because that seems to be the goal - a biological explanation?
-% R- Now this is mentioned in the intro so that  makes it easier to write here
+E: I would say hisse captures some factor that evolves like a binary trait, not necessarily a real biological trait
 In this model, each of the observed diploid and polyploid states were subdivided by a binary hidden trait with states $A$ and $B$.
-This model was called the D/P+A/B ploidy and hidden state model. 
-Model parameters include speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$) and extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$) for each subdivided state.
-The polyploidization rate is again represented by $\rho$, and the rates of transition between hidden states are symmetrical with parameter $\alpha$.
+In this $D/P+A/B$ ploidy and hidden state model, parameters include speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$) and extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$) for each subdivided state.
+The polyploidization rate is again represented by $\rho$, and the rates of transition between hidden states are symmetrical with parameter $\alpha$.  % FIXME ?
+
+Diagrams of these two ploidy models are shown in \cref{figure:netdivall}AB.  % request from referee 2
 
@@
 
-To assess the effects of breeding system in the diversification process, we first fit a model in which the states are self-incompatible (I) or self-compatible (C).
-This is the same as the analysis of \citet{goldberg_2010}, but with an updated phylogeny \citep{sarkinen_2013}.
-We call this BiSSE model the I/C breeding system model. 
-It has parameters for state-specific speciation ($\lambda_I,\ \lambda_C$) and extinction ($\mu_I,\ \mu_C$), and for transitions from I to C ($q_{IC}$).
+To assess the effects of breeding system in the diversification process, we first fit a model with self-incompatible $(I)$ or self-compatible $(C)$ states.
+This is the same as the analysis of \citet{goldberg_2010} but with an updated phylogeny \citep{sarkinen_2013}.
+This $I/C$ breeding system model has parameters for state-specific speciation ($\lambda_I,\ \lambda_C$) and extinction ($\mu_I,\ \mu_C$), and for transitions from $I$ to $C$ ($q_{IC}$).
 
-To parse out the effect of breeding system on diversification, while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
-We call this HiSSE model the I/C+A/B breeding system and hidden state model.
-It also has parameters for state-specific speciation ($\lambda_{I_A},\ \lambda_{I_B},\ \lambda_{C_A},\ \lambda_{C_B}$) and extinction ($\mu_{I_A},\ \mu_{I_B},\ \mu_{C_A},\ \mu_{C_B}$), for loss of SI ($q_{IC}$), and for transitions between hidden states ($\alpha$).
-An I/C+A/B model was used by \citet{freyman_2017} used to study diversification in Onagraceae, but here we restricted the hidden state transitions to be symmetrical.
+To parse out the effect of breeding system on diversification while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of the breeding system states into hidden states $A$ and $B$.
+This $I/C+A/B$ breeding system and hidden state model has parameters for state-specific speciation ($\lambda_{I_A},\ \lambda_{I_B},\ \lambda_{C_A},\ \lambda_{C_B}$) and extinction ($\mu_{I_A},\ \mu_{I_B},\ \mu_{C_A},\ \mu_{C_B}$), for loss of SI ($q_{IC}$), and for transitions between hidden states ($\alpha$).
+An similar model was used by \citet{freyman_2017} used to study diversification in Onagraceae, but here we restricted the hidden state transitions to be symmetrical. % FIXME still true?
 
@@
 
+Diagrams of these two breeding system models are shown in \cref{figure:netdivall}CD.  % request from referee 2
+
 \subsubsection{Ploidy, breeding system, and diversification}
 
-Ploidy and breeding system might influence lineage diversification individually, but these two traits also have complex interactions in Solanaceae species.
-Therefore, we considered a multi-state model that investigates the contribution of both traits and the possible transitions between them.
-Self-incompatible diploids (ID), self-compatible diploids (CD), and polyploids, which are always self-compatible (CP) were used in a MuSSE model \citep{fitzjohn_2012}.
-We did not include a state for self-incompatible polyploids because they are not observed in the data, and  that trait combination state is mechanistically predicted not to occur.
-We call this the ID/CD/CP ploidy and breeding system model.
-It has six parameters for diversification in each state ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation, $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction), and three for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
+Ploidy and breeding system might influence lineage diversification individually, but these two traits also have an intricate association (see the Introduction).
+Therefore, we considered a multi-state model that investigates the contribution of both traits and the allowable transitions between them.
+This $ID/CD/CP$ ploidy and breeding system model has states for self-incompatible diploids $(ID)$, self-compatible diploids $(CD)$, and self-compatible polyploids $(CP)$; self-incompatible polyploids do not occur, as explained in the Introduction.
+There are six parameters for state-specific diversification rates ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation; $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and three for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
 The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
-% Diploidization from $CP$ to $ID$ is not allowed because it would represent a simultaneous regain of SI. % E: moved into the diploidization section
 
-The ID/CD/CP model could also suffer from a high type I error, like the two-state BiSSE model. 
+Although the $ID/CD/CP$ model allows for two traits to potentially influence diversification, other unknown factors could as well.
 We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018, huang_2018}.
-We refer to this as the ID/CD/CP+A/B model.
-A fully parameterized version of this model would have 26 rate parameters. 
+A fully parameterized version of this $ID/CD/CP+A/B$ model would have 26 rate parameters. 
 Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fit a simplified version with 16 parameters.
-% You were missing counting the alpha
-The reduction in parameter space was achieved by fixing the rates for transitions among hidden states to be equal with rate $\alpha$, and fixing the transition rates between observed states to be independent of the hidden state (rates $\rho_I,\rho_C,\ q_{IC}$ as defined for the ID/CD/CP model).
+% FIXME still true?
+The reduction in parameter space was achieved by fixing the rates for transitions among hidden states to be equal with rate $\alpha$, and fixing the transition rates between observed states to be independent of the hidden state (rates $\rho_I,\rho_C,\ q_{IC}$ as defined for the $ID/CD/CP$ model).
 There are additionally twelve diversification rate parameters ($\lambda_{ID_A},  \lambda_{ID_B},  \lambda_{CD_A}, \lambda_{CD_B}, \lambda_{CP_A}, \lambda_{CP_B}, \mu_{ID_A}, \mu_{ID_B}, \mu_{CD_A}, \mu_{CD_B},  \mu_{CP_A}, \mu_{CP_B}$).
@@
 
+Diagrams of these two ploidy and breeding system models are shown in \cref{figure:netdivall}EF.  % request from referee 2
+
 \subsubsection{Pathways to polyploidy}
@@
 In the two-step pathway, the CD state is an intermediate: SI is first lost, and later the SC diploid undergoes polyploidization.
-We quantify the relative contribution of these pathways to polyploidy in two ways, each using the maximum \textit{a posteriori} (MAP) estimates of rates from the IC/CD/CP model.
+We quantify the relative contribution of these pathways to polyploidy in two ways, each using the maximum \textit{a posteriori} (MAP) estimates of rates from the $IC/CD/CP$ model.
 
-Both of our methods are based on a propagation matrix that describes flow from ID to CP, as in \citet{robertson_2011}.
-We insert an artificial division in the CP state, so that one substate contains the CP species that arrived via the one-step pathway and the other substate contains the CP species that arrived via the two-step pathway.
+Both of our methods are based on a propagation matrix that describes flow from $ID$ to $CP$, as in \citet{robertson_2011}.
+We insert an artificial division in the $CP$ state, so that one substate contains the $CP$ species that arrived via the one-step pathway and the other substate contains the $CP$ species that arrived via the two-step pathway.
 We consider only unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and no strong support for diploidization (see below).
+
 First, we consider only the rates of transitions between these states, placing them in the propagation matrix \myvec{Q}.
@@
 Our results will differ from those of \citet{robertson_2011} because our transition rate estimates come from a dated phylogeny and a model that allows for state-dependent diversification.
-%
+
 Second, we consider not only transitions between states but also diversification within each state.
 State-dependent diversification can change the relative contributions of the two pathways.
-In particular, if the net diversification rate is small for CD, the two-step pathway will contribute relatively less.
+For example, if the net diversification rate is small for $CD$, the two-step pathway will contribute relatively less.
 We therefore include the difference between speciation and extinction along the diagonal elements of the propagation matrix.
@@
 In this case, however, these are not probabilities because diversification changes the number of lineages as time passes.
-We can still use their ratios to consider the relative contribution of each pathway, though, analogous to the normalized age structure in a growing population. % R- here a reference would be nice
+We can still use their ratios to consider the relative contribution of each pathway, though, analogous to the normalized age structure in a growing population.
+% R- here a reference would be nice  E: Does Jake have a favorite ref?
+Otherwise, I guess a basic ecology textbook.
 
-% "one-step predictions for P_I and P_C over time"
-% E & R todo later: for the case with net div, write out the propagation matrix and equations for pathway contributions
+% E & R todo ?: for the case with net div, write out the propagation matrix and equations for pathway contributions
 
@@
 
-For the models with ploidy changes, we allowed diploidization by adding a parameter $\delta$ for transitions from P back to D (in the D/P and D/P+A/B models) or from CP back to CD (in the I/C and I/C+A/B models).
-We did not allow transitions from CP back to ID because that would include regain of SI.
-The models are denoted with $+\delta$ in \cref{table:marginallike}.
-Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010, glick2014, mayrose_2015, freyman_2017}.
-These types of classifications do not allow for a ploidy reversion.
-Where indicated, the classification of ploidy for the data used in our models was based on chromosome multiplicity at the genus level (see \url{https://github.com/roszenil/solploidy} for archived dataset).
-% (data available at https://www.github.com/roszenil/solploidy/datasets).
-However, the majority of the ploidy classifications were adopted from original studies with alternative sources of information (\eg geographic distribution, genus ploidy distribution) where ploidy was defined by authors that found evidence for it.
-Since it is not clear whether diploidization can be detected under alternative ploidy classifications or even classifications based on chromosome number multiplicity at the genus level, we fit models with and without diploidization in order to test  whether the conclusions about diversification are sensitive to including diploidization.
-As discussed by \citet{servedio_2014}, the presence or absence of a process can have an exploratory goal.
+The models described above allow polyploidization but not its reverse.
+We explored the effects of allowing diploidization by adding a parameter $\delta$ for transitions from $P$ back to $D$ (in the $D/P$ and $D/P+A/B$ models) or from $CP$ back to $CD$ (in the $I/C$ and $I/C+A/B$ models); we did not allow transitions from $CP$ back to $ID$ because that would include regain of SI.
+Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010, glick2014, mayrose_2015, freyman_2017}, which assume only aneuploid change and polyploidization.
+Our ploidy classifications, however, when possible also incorporated information on chromosome multiplicity at the genus level and author conclusions within original studies with alternative sources of information (\eg geographic distribution, genus ploidy distribution).
+Since it is not known whether diploidization can be detected under these additional ploidy classifications, we fit models with and without diploidization in order to test whether the conclusions about diversification are sensitive to including diploidization.
+% As discussed by \citet{servedio_2014}, the presence or absence of a process can have an exploratory goal.
 In our models, the diploidization parameter (or its absence, $\delta=0$), represents an opportunity to explore a possibly significant assumption that may not decisively affect interactions among ploidy, breeding system, and diversification.

### 2019-05-23 manuscript/methods.tex
edits to Stat Inference

@@
 
-Parameters for each of the ten models described were estimated using RevBayes \citep{hoehna_2016}.
-Code for analyses and key results are available at \url{https://github.com/roszenil/solploidy}.
-We used the method of \citet{fitzjohn_2009} to account for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $595/3000$ since the Solanaceae family has approximately 3,000 species as estimated by the Solanaceae Source project \citep{solsource}.
+Parameters for each of the ten models were estimated using RevBayes \citep{hoehna_2016}.
+Code for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.
+We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $595/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
 For all models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
@@
 For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run for 96 hours in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the 10 models.
-%B: the number of models changes here?
-R-my mistake!
-fixed it
+% FIXME see referee comment about reporting convergence, not runtime
 Convergence and mixing of each MCMC chain was determined by ensuring the effective sample size of each parameter was over 200.
@@
 Additionally, ancestral states at each node in the phylogeny were sampled jointly during the MCMC analyses.
-Ancestral state reconstructions for all models can be found in the Supplementary Information and show the maximum posterior estimates of the marginal probability distributions for each of the 594 internal nodes for each of the 10 models.
-% TODO: add fig refs (or remove the mention of them)
-%B: there was an error in the above sentence: no estimates or values after a posteriori (also "posterior" is ok?)
+Ancestral state reconstructions for all models show the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 594 internal nodes for each of the 10 models (\cref{suppfigure:DPnodipasr,suppfigure:DPasr,suppfigure:DPnodipABasr,suppfigure:DPABasr,suppfigure:ICasr,suppfigure:ICABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPasr,suppfigure:IDCDCPnodipABasr,suppfigure:IDCDCPABasr}).
+% FIXME: I'm confused about if these are joint or marginal?
+A bit more explanation would help.
 
@@
 
-We calculated the marginal likelihood for each of the ten models in RevBayes \citep{hoehna_2016}.
-Marginal likelihoods were calculated using 50 stepping stone steps under the methodology of \citet{xie_2010}.
+We calculated the marginal likelihood for each of the ten models using fifty stepping stone steps under the methodology of \citet{xie_2010}, implemented in RevBayes \citep{hoehna_2016}.
 Each stepping stone step was found by calculating 500 generations of burn-in followed by a total of 1,000 MCMC steps (\cref{table:marginallike}).
-The calculation of each marginal likelihood ran for 24 hours on a high-performance computational cluster.
+The calculation of each marginal likelihood ran for 24 hours on a high-performance computational cluster. % FIXME: remove?
 
 Using the marginal likelihood values, we calculated thirteen different Bayes factors.
-Six compared the models of ploidy against one other (D/P and D/P+A/B, each with or without diploidization), one compared the breeding system models (I/C and I/C+A/B), and six compared the models with both traits (ID/CD/CP and ID/CD/CP+A/B, each with or without diploidization) (\cref{table:bayesfactors}).
+Six compared the models of ploidy against one other ($D/P$ and $D/P+A/B$, each with or without diploidization), one compared the breeding system models ($I/C$ and $I/C+A/B$), and six compared the models with both traits ($ID/CD/CP$ and $ID/CD/CP+A/B$, each with or without diploidization) (\cref{table:bayesfactors}).
 %
 Other comparisons between these models are not valid because the input data are different under the different state space codings (\cref{figure:stateclassifications}).
-In mathematical terms, the D/P, I/C, and ID/CD/CP state spaces are not `lumpable' with respect to one another \citep{tarasov_2019}.
-% E: I think this short explanation is clear enough, especially with the figure.  I'm not sure whether the hierarchical concept is exactly applicable here.
-%R: Boris in the introduction mentions this possible causal relationship would that be equivalent to say hierarchical?
+In mathematical terms, the $D/P$, $I/C$, and $ID/CD/CP$ state spaces are not `lumpable' with respect to one another \citep{tarasov_2019}.
+% E: I'm honestly still confused about whether this is true.
+From ID/CD/CP, we can lump ID/CD to get D/P, and we can lump CD/CP to get I/C.
+FIXME
 Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $BF(M_0,\ M_1) = \ln[ P(\mathbf{X} | M_0) - P(\mathbf{X} | M_1)]$.

### 2019-05-23 manuscript/results.tex
edits to Models
edits to Stat Inference
edits to Results

--- 2019-04-25 manuscript/results.tex
@@
 
-When considering ploidy alone (D/P model), we found a greater net diversification rate for diploids than for polyploids, in agreement with \citep{mayrose_2011, mayrose_2015}.
+When considering ploidy alone ($D/P$ model), we found a greater net diversification rate for diploids than for polyploids, in agreement with \citet{mayrose_2011, mayrose_2015}.
 This result holds with (\cref{figure:netdivall}A) or without (\cref{figure:netdivnodip}A) the diploidization parameter, although including diploidization shifts the net diversification rate of polyploids to be non-negative.
-Incorporating a hidden state in this model, however, removes the clear separation in diversification rate estimates between diploids and polyploids (D/P+A/B model; \cref{figure:netdivall}B, \cref{figure:netdivnodip}B).
-Thus, differences in net diversification rates are better explained by an unknown factor than by ploidy.
+Incorporating a hidden state in this model, however, reduces the clear separation in diversification rate estimates between diploids and polyploids ($D/P+A/B$ model; \cref{figure:netdivall}B, \cref{figure:netdivnodip}B).
+Thus, differences in net diversification rates are explained by an unknown factor as well as by ploidy.
+% E: I edited the interpretation of D/P+A/B.  There still is some separation between D/P within each A/B.  The A/B separation isn't any clearer, at least.
 Statistical model comparisons show very strong support for including the hidden state and strong support for including diploidization (\cref{table:bayesfactors}).
 
-When considering breeding system alone (I/C model, \cref{figure:netdivall}C), we found a larger net diversification rate for SI than for SC species, in agreement with \citet{goldberg_2010}.
-When a hidden state is included (I/C+A/B model), the large net diversification rate difference persists for one of the hidden states but is removed for the other (\cref{figure:netdivall}D).
+When considering breeding system alone ($I/C$ model, \cref{figure:netdivall}C), we found a larger net diversification rate for SI than for SC species, in agreement with \citet{goldberg_2010}.
+When a hidden state is included ($I/C+A/B$ model), the large net diversification rate difference persists for one of the hidden states but is mostly removed for the other (\cref{figure:netdivall}D).
 Thus, differences in net diversification rates are best explained by both breeding system and an unknown factor \citep[as in Onagraceae;][]{freyman_2019}.
@@
 
-When considering ploidy and breeding system together (ID/CD/CP model), the net diversification rate for SI diploids was greater than for either SC diploids or SC polyploids, with or without diploidization (\cref{figure:netdivall}E, \cref{figure:netdivnodip}C).
-The difference in net diversification with breeding system persists when ploidy is included in the model, but not the reverse.
-That is, the association of ploidy with net diversification in the D/P model (\cref{figure:netdivall}A, \cref{figure:netdivnodip}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids are similar.
+When considering ploidy and breeding system together ($ID/CD/CP$ model), the net diversification rate for SI diploids was greater than for either SC diploids or SC polyploids, with or without diploidization (\cref{figure:netdivall}E, \cref{figure:netdivnodip}C).
+Thus, the difference in net diversification associated with breeding system persists even when ploidy is included in the model.
+The reverse is not true, though: the association of ploidy with net diversification in the $D/P$ model (\cref{figure:netdivall}A, \cref{figure:netdivnodip}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids not as distinct.
 %
-When a hidden state is included (IC/CD/CP+A/B model), the same general pattern remains when diploidization is prevented (\cref{figure:netdivnodip}D), although the higher net diversification rate of ID is less clear within one of the hidden states.
-With diploidization, the net diversification rate of ID is still greater than CD within each hidden state, but diversification estimate for P is highly uncertain and perhaps bimodal.
+When a hidden state is included ($IC/CD/CP+A/B$ model), the same general pattern remains when diploidization is prevented (\cref{figure:netdivnodip}D), although the higher net diversification rate of $ID$ is less clear within one of the hidden states.
+With diploidization, the net diversification rate of $ID$ is still greater than $CD$ within each hidden state, but diversification estimate for $CP$ is highly uncertain and perhaps bimodal.
 Statistical model comparisons show very strong support for including the hidden state and at most positive support for including diploidization (\cref{table:bayesfactors}).
@@
 In the two-step pathway, SI is first lost within the diploid state, followed by polyploidization.
-Determining the relative contribution of these pathways based on our estimated transition rates from the ID/CD/CP model, we find that the one-step pathway is more likely on short timescales and the two-step pathway is more likely on long timescales (\cref{figure:pathways}, left panels).
-Beginning with a single SI diploid lineage, when not much time has elapsed, the one-step pathway is more likely because it only necessitates a single event to effect transition to SC polyploid state.
+Determining the relative contribution of these pathways based on our estimated transition rates from the $ID/CD/CP$ model, we find that the one-step pathway is more likely on short timescales and the two-step pathway is more likely on long timescales (\cref{figure:pathways}, left panels).
+Beginning with a single SI diploid lineage, when not much time has elapsed, the one-step pathway is more likely because it only necessitates a single event to effect transition to the SC polyploid state.
 When more time has elapsed, the two-step pathway is more likely because the rate of loss of SI within diploids, $q_{IC}$, is greater than the rate of polyploidization for SI species, $\rho_I$ (\cref{suppfigure:IDCDCPnodip}).
+That is, an $ID$ lineage is more likely to begin its path to polyploidy with a transition to $CD$, but completing this path to $CP$ takes longer.
 
@@
 Allowing for the different net diversification rates for each state, we find a qualitative difference in the relative pathway contributions.
-In particular, the lower rate of net diversification in the CD state, relative to ID, means that relatively fewer lineages are available to complete the second step of the two-step pathway, ending in CP.
-Consequently, even over long timescales, the two-step pathway contributes less to generation of CP (\cref{figure:pathways}, right panels).
+In particular, the lower rate of net diversification in the $CD$ state, relative to $ID$, means that relatively fewer lineages are available to complete the second step of the two-step pathway, ending in $CP$.
+Consequently, even over long timescales, the two-step pathway contributes less to the formation of polyploids (\cref{figure:pathways}, right panels).
 
@@
 We considered models both with and without diploidization in order to explore its effects on the estimates of state-dependent diversification.
-For the two models that only include diploid and polyploid states (D/P and D/P+A/B), the net diversification rate of polyploid lineages is likely negative when diploidization (parameter $\delta$) is excluded but positive when $\delta$ is included (\cref{figure:netdivall}A versus \cref{figure:netdivnodip}A).
-For the two models that also included breeding system (ID/CD/CP and ID/CD/CP+A/B), the main effect of including diploidization is greatly increasing the uncertainty of the estimated polyploid net diversification rate (\cref{figure:netdivall}E,F versus \cref{figure:netdivnodip}C,D).
+For the two models that only include diploid and polyploid states ($D/P$ and $D/P+A/B$), the net diversification rate of polyploid lineages is likely negative when diploidization (parameter $\delta$) is excluded but positive when it is included (\cref{figure:netdivall}A versus \cref{figure:netdivnodip}A).
+For the two models that also included breeding system ($ID/CD/CP$ and $ID/CD/CP+A/B$), the main effect of including diploidization is greatly increasing the uncertainty of the estimated polyploid net diversification rate (\cref{figure:netdivall}EF versus \cref{figure:netdivnodip}CD).
 For all models, there is greater uncertainty in the estimate of diploidization rate than polyploidization rate, as judged by the width of the credibility intervals (see \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}).
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{allnetdivmodelswithoutdelta.pdf}
+\includegraphics[width=\textwidth]{allnetdivmodelswithoutdelta.pdf} % FIXME: need horizontal axis label
 \caption{Net diversification rates for all models without diploidization rate estimate. 
@@
 
-% E todo: could make this look a little less klunky
+% E todo: include a diagram of the transitions (panels a/b at the top?); re-run with new rate estimates; add panel letters
 \begin{figure}

### 2019-05-23 manuscript/methods.tex
edited states figure
included Boris' graphics changes and a few more
adjusted the caption

@@
 Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 595 species with ploidy and/or breeding system information on the tree.
+% FIXME: update numbers around here
 Of these, 290 had information for both states.
-The number of species in each combination of states is summarized in \cref{figure:stateclassifications}.
+The number of species in each combination of states is summarized in \cref{figure:stateclassifications}A.
 We retained all 595 species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
@@
 %
-Other comparisons between these models are not valid because the input data are different under the different state space codings (\cref{figure:stateclassifications}).
+Other comparisons between these models are not valid because the input data are different under the different state space codings (\cref{figure:stateclassifications}B).
 In mathematical terms, the $D/P$, $I/C$, and $ID/CD/CP$ state spaces are not `lumpable' with respect to one another \citep{tarasov_2019}.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{states.pdf}
+\includegraphics[width=0.5\textwidth]{states} % FIXME: update numbers
 \caption{
@@
 Each species retained on the tree belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known.
-The number of species in each is shown under the corresponding circles in the top row.
-These categories were then grouped in a manner appropriate to the states of each model.
-For example, there are 34 species that are self-compatible ($C$) with unknown ploidy, and 40 species known to be polyploid($P$), but without breeding system data; the latter can be inferred, because all polyploid species in the dataset are self-compatible (see \citealt{robertson_2011}, and text).
-%B: I added stuff here
-In the D/P models, these are coded as either $D$ or $P$ (uncertain/consistent with either state); in the I/C  models, they are coded as $C$; and in the ID/CD/CP models, they are coded as either $CD$ or $CP$.
-In the hidden state models, species were coded as either $A$ or $B$.
-The model-specific coding is shown in different rows. 
-%B: I think that these two sentences are more appropriate for the model comparison tables and Methods text. If we are admonished to cut text, they could go.
-They each use state spaces that are not directly comparable with one another.
-For example, we cannot test whether the D/P model fits better than the I/C model because they use states that are not the same, and are therefore not `lumpable' \citep{tarasov_2019}.
-%B: [This] will be very difficult as a source of clarity for someone reading the paper the first time.
-%E: Yes, all I species are D and all P species are C (no SI polyploids).
-I don't think there are polymorphic species because species with some SI and some SC pops are called I, and the most common ploidy level was used.
+(A) The five categories and the number of species in each.
+Abbreviations are $I$ for self-incompatible, $C$ for self-compatible, $D$ for diploid, $P$ for polyploid, $?$ for unknown.
+Because polyploidization breaks this form of self-incompatibility, SI species with unobserved ploidy ($I?$) are taken to be diploid ($ID$), and polyploid species with unobserved breeding system ($?P$) are taken to be SC.
+(B) Category groupings into states for each model type.
+For example, there are 34 species that are SC with unknown ploidy. % FIXME: update number if necessary
+In the $D/P$ models these are coded as either $D$ or $P$ (uncertain/consistent with either state), in the $I/C$ models they are coded as $C$, and in the ID/CD/CP models they are coded as either $CD$ or $CP$.
+In the hidden state models, all species were coded as consistent with both $A$ and $B$.
 }
@@
     Bayes factors for model comparisons.
-    Each of the three boxes contains models that can be compared with one another, based on the character states they include (see \cref{figure:stateclassifications}).
+    Each of the three boxes contains models that can be compared with one another, based on the character states they include (see \cref{figure:stateclassifications}B).
     Models are numbered in as \cref{table:marginallike}. % FIXME: they aren't, but it's a pain to change; change in prev table instead?

### 2019-05-24 manuscript/discussion.tex

@@
 There could be other non-trait weirdnesses.
+TODO: include a more comprehensive discussion of how to interpret hisse/hidden state
 When considering a second candidate trait, one can model the complex interactions via a MuSSE model, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.

### 2019-05-24 manuscript/introduction.tex

@@
 There are several possible general explanations: whole genome duplication may facilitate the transition to selfing by masking inbreeding depression, self-fertilization may facilitate polyploidy establishment by avoiding the triploid bridge, or both selfing and polyploidy may be correlates of weedy or peripheral populations \citep{ramsey_1998, barringer2007, barrett2008, husband2008}.
-% Boris, can you please check?  We need to be totally convincing about no SI-P.
-Additionally, specifically in RNase-based gametophytic SI systems, polyploidization may directly cause the loss of SI \citep{stout1942, lewis1947}.
-In these systems, self pollen fails because a pollen grain with one S-allele at the locus controlling the SI response is unable to detoxify the RNase produced by the same S-allele in the style. % ref?
-The pollen of polyploids, however, has multiple S-alleles which act complementarily, detoxifying the RNases produced by any maternal genotype \citep{entani1999, tsukamoto2005, kubo2010}.
-Polyploids are consequently capable of self-fertilization, with exceedingly few exceptions \citep{hauck_2002, nunes_2006}.
+Additionally, in RNase-based gametophytic SI systems, polyploidization alone directly causes the loss of SI \citep{stout1942, lewis1947}.
+%B: disclaimer: ceteris paribus statement omitted; polyploidization ordinarily causes SC because of the formation of heterodiploid pollen.
+On the background of deletion of S-locus, or with heterologous chromosome homozygosity at the S-locus may not do so.
+These, however, appear to be sufficiently rare that they are recovered perhaps twice in dicots: once (or twice) naturally, in sour cherry and it's close relative (Prunus), and once in a experimentally induced and selected synthetic pentaploid Solanum (which has not been replicated).
+Strictly speaking, this is a feature of non-self recognition systems of the kind found in RNase-based GSI.
+In these systems, self-incompatibility is thought to occur because haploid self-pollen grains, with one S-allele at the locus controlling the SI response, are unable to detoxify the S-RNase produced by the same S-allele in the style \kubo{2010}.
+The unreduced pollen of tetraploids, however, can contain two S-alleles expressed in pollen, which jointly gain the ability to detoxify the S-RNases produced by any maternal genotype \citep{entani1999, tsukamoto2005, kubo2010}.
+Polyploid individuals are consequently initially capable of self-fertilization, with exceedingly few exceptions \citep{hauck_2002, nunes_2006}.
 This system of RNase-based SI is regarded as ancestral in eudicots \citep{igic_2001,steinbachs_2002}, and it is expressed in all SI species of Solanaceae examined to date.

### 2019-05-24 manuscript/introduction.tex
latex fix

@@
 Strictly speaking, this is a feature of non-self recognition systems of the kind found in RNase-based GSI.
-In these systems, self-incompatibility is thought to occur because haploid self-pollen grains, with one S-allele at the locus controlling the SI response, are unable to detoxify the S-RNase produced by the same S-allele in the style \kubo{2010}.
+In these systems, self-incompatibility is thought to occur because haploid self-pollen grains, with one S-allele at the locus controlling the SI response, are unable to detoxify the S-RNase produced by the same S-allele in the style \citep{kubo2010}.
 The unreduced pollen of tetraploids, however, can contain two S-alleles expressed in pollen, which jointly gain the ability to detoxify the S-RNases produced by any maternal genotype \citep{entani1999, tsukamoto2005, kubo2010}.

### 2019-05-27 manuscript/lumpability.tex
Lumpability tex file

\documentclass{article}
\usepackage{amsmath}
\begin{document}
Three state matrix
$$
\begin{pmatrix} 
-q_{12}-q_{13} & q_{12} & q_{13} \\
q_{21} & -q_{21}-q_{23} & q_{23}\\
q_{31} & q_{32} & -q_{31}-q_{32} 
\end{pmatrix}
$$

We want to lump states $2$ and $3$ together. According to Tarasov (2018, Figure 3), symmetries in the rate matrix are necessary for strong lumpability. This is call the row-wise sum rule.\\
To aggregate our three state model into a two state model we first define the new partition $\{f1=\{1\},f2=\{2,3\}\}$. The new rates for the two state model are going to be $\hat{q}_{ij}$. The aggregation is lumpable if and only if two conditions happen
$$ \hat{q}_{12}=q_{12}+q_{13} \qquad \textrm{and} \qquad \hat{q}_{21}=q_{21}=q_{31}$$

So the two state matrix for the lumped states will be
$$
\begin{pmatrix} 
-\hat{q}_{12} &\hat{q}_{12} \\
\hat{q}_{21} & -\hat{q}_{21}
\end{pmatrix}
$$

Now, let's assume we want to lump the model  for breeding system, therefore $\{f1=\{1=ID\}, f2=\{2=CD,3=CP\}\}$. We would have that

\begin{eqnarray}
 \hat{q}_{12} & = &q_{12}+q_{13}\\
\hat{q}_{IC}& = & q_{IC}+\rho_I \label{firstcondition}
\end{eqnarray}

and

\begin{eqnarray}\label{secondcondition}
\hat{q}_{21}&=q_{21}=q_{31}\\
0 & =0=0 \label{firstcondition}
\end{eqnarray}

The second condition (4) is met. This is what follows in the S3a Appendix result in Tarasov (2018). However, the first condition (2) is difficult to meet since the two state model would have to be defined via two parameters that are non-identifiable.\\

One way to overcome this issue could be defining in the three state model $q_{IC}=\hat{q}_{IC}*\gamma$ and $\rho_I=\hat{q}_{IC}(1-\gamma)$ with $\gamma$ is a r.v. between $(0,1)$. This assumption is okay if we are assuming that the evolution from SI to SC is sometimes straight forward in diploids and sometimes it happens due to polyploid. The statistical complication of this assumption is that $\gamma$ and $q_{IC}$ will be non-identifiable (they always appear in a functional relationship).  
\end{document}

### 2019-05-28 manuscript/introduction.tex
- fixed a sentence with unreduced pollen in Intro

@@
 In these systems, self-incompatibility is thought to occur because haploid self-pollen grains, with one S-allele at the locus controlling the SI response, are unable to detoxify the S-RNase produced by the same S-allele in the style \citep{kubo2010}.
-The unreduced pollen of tetraploids, however, can contain two S-alleles expressed in pollen, which jointly gain the ability to detoxify the S-RNases produced by any maternal genotype \citep{entani1999, tsukamoto2005, kubo2010}.
-Polyploid individuals are consequently initially capable of self-fertilization, with exceedingly few exceptions \citep{hauck_2002, nunes_2006}.
+The unreduced pollen of diploids, however, can contain two S-alleles expressed in pollen, which jointly gain the ability to detoxify the S-RNases produced by any maternal genotype \citep{entani1999, tsukamoto2005, kubo2010}.
+Initial mutant individuals with pollen containing doubled haploid genomes are consequently capable of self-fertilization, with exceedingly few exceptions \citep{hauck_2002, nunes_2006}.
+%B: this was rephrased.
 This system of RNase-based SI is regarded as ancestral in eudicots \citep{igic_2001,steinbachs_2002}, and it is expressed in all SI species of Solanaceae examined to date.

### 2019-05-28 summarystats/solanaceaesummarystats.R
New data files

@@ -10,11 +10,15 @@ solanaceae.tree<-read.tree("~/Dropbox/solploidy/basicdata/solcorrecttip.tre")
 #idnumber, source,fullname, genus, species,cultivarhybrid,csomenumber,lifehistoryclean, ploidy selfincomp, ploidybyZF, namechangedfrom
 solanaceae.allrecords<-read.csv("~/Dropbox/solploidypersonal/rawdata/solanaceaerecords.csv",header=TRUE,sep=",", stringsAsFactors=FALSE, na.strings="")
 
+cultivars<-which(solanaceae.allrecords$cultivarhybrid==1)
 #How many species, how many genera
 species<-unique(solanaceae.allrecords$fullname)
 total.species<-length(species)# 2268 total species
 genera<-unique(solanaceae.allrecords$genus)
 total.genera<-length(genera)#97 genera
+
+
+solanaceae.allrecords<-solanaceae.allrecords[-cultivars,]
 
 #Calculating the mode(s)
 Mode <- function(x) {
@@ -61,7 +65,7 @@ nor.ploidysi<-which(is.na(simplified.solanaceae[,5])==TRUE & is.na(simplified.so
 norploidysi.solanaceae<-simplified.solanaceae[-c(nor.ploidysi),]
 
 ###########################################
-### Matching tree with matching dataset if they have either ploidy or self-incomp information that is 550 species
+### Matching tree with matching dataset if they have either ploidy or self-incomp information that is 
 
 
 ##Just checking here that all data has either ploidy or breeding system info
@@ -73,12 +77,12 @@ write.nexus(matched$phy,file="~/Dropbox/solploidy/basicdata/fullmatchtree.nex")
 
 diploidsi<-which(matched$dat[,4]==2 & matched$dat[,5]=="SI")# 97
 diploidsc<-which(matched$dat[,4]==2 & matched$dat[,5]=="SC")#152
-diploidsna<-which(matched$dat[,4]==2 & is.na(matched$dat[,5])==TRUE) #223
-polyploids<-which(matched$dat[,4]>2) #86
-polyploidsc<-which(matched$dat[,4]>2 & matched$dat[,5]=="SC")#43
-polyploidsna<-which(matched$dat[,4]>2 & is.na(matched$dat[,5])==TRUE)#43
[…]

### 2019-05-28 summarystats/solanaceaesummarystats.R
New datasets with 635 species

@@ -8,14 +8,14 @@ solanaceae.tree<-read.tree("~/Dropbox/solploidy/basicdata/solcorrecttip.tre")
 
 # Dataset created by RZF from CCDB and independent records by EEG
 #idnumber, source,fullname, genus, species,cultivarhybrid,csomenumber,lifehistoryclean, ploidy selfincomp, ploidybyZF, namechangedfrom
-solanaceae.allrecords<-read.csv("~/Dropbox/solploidypersonal/rawdata/solanaceaerecords.csv",header=TRUE,sep=",", stringsAsFactors=FALSE, na.strings="")
+solanaceae.allrecords<-read.csv("~/Dropbox/solploidy/basicdata/solanaceaerecords.csv",header=TRUE,sep=",", stringsAsFactors=FALSE, na.strings="")
 
 cultivars<-which(solanaceae.allrecords$cultivarhybrid==1)
 #How many species, how many genera
 species<-unique(solanaceae.allrecords$fullname)
-total.species<-length(species)# 2268 total species
+total.species<-length(species)# 2270 total species
 genera<-unique(solanaceae.allrecords$genus)
-total.genera<-length(genera)#97 genera
+total.genera<-length(genera)#99 genera
 
 
 solanaceae.allrecords<-solanaceae.allrecords[-cultivars,]
@@ -71,18 +71,18 @@ norploidysi.solanaceae<-simplified.solanaceae[-c(nor.ploidysi),]
 ##Just checking here that all data has either ploidy or breeding system info
 ##nor.ploidysi2<-which(is.na(norploidysi.solanaceae[,5])==TRUE & is.na(norploidysi.solanaceae[,6])==TRUE)
 
-matched<-make.treedata(solanaceae.tree,norploidysi.solanaceae) #595
+matched<-make.treedata(solanaceae.tree,norploidysi.solanaceae) #635 species!!!!
 write.nexus(matched$phy,file="~/Dropbox/solploidy/basicdata/fullmatchtree.nex")
 
 
-diploidsi<-which(matched$dat[,4]==2 & matched$dat[,5]=="SI")# 97
-diploidsc<-which(matched$dat[,4]==2 & matched$dat[,5]=="SC")#152
-diploidsna<-which(matched$dat[,4]==2 & is.na(matched$dat[,5])==TRUE) #219
[…]

### 2019-07-09 basicdata/matrixdataprocessing.R
all diversification figures

@@ -4,28 +4,19 @@ library("treeplyr")
 
 ## Solanaceae time-tree with corrected names Sarkinen et al 2013
 solanaceae.tree<-read.tree("~/Dropbox/solploidy/basicdata/solcorrecttip.tre")
-
-cultivars<-which(solanaceae.allrecords$cultivarhybrid==1)
-#How many species, how many genera
-species<-unique(solanaceae.allrecords$fullname)
-total.species<-length(species)# 2270 total species
-genera<-unique(solanaceae.allrecords$genus)
-total.genera<-length(genera)#99 genera
-
-# Dataset created by RZF from CCDB and independent records by EEG
-#idnumber, source,fullname, genus, species,cultivarhybrid,csomenumber,lifehistoryclean, ploidy selfincomp, ploidybyZF, namechangedfrom
 solanaceae.allrecords<-read.csv("~/Dropbox/solploidy/basicdata/solanaceaerecords.csv",header=TRUE,sep=",", stringsAsFactors=FALSE, na.strings="")
-
 cultivars<-which(solanaceae.allrecords$cultivarhybrid==1)
 solanaceae.allrecords<-solanaceae.allrecords[-cultivars,]
 
-uncertain.species<-c("Nierembergia_rigida","Anisodus_luridus","Lycium_europaeum","Lycium_chilense","Lycium_ciliatum","Chamaesaracha_coronopus","Chamaesaracha_sordida","Physalis_hederifolia","Solanum_immite","Solanum_acroscopicum","Solanum_multiinterruptum","Solanum_brevicaule","Solanum_bahamense","Solanum_juvenale","Solanum_hieronymi","Solanum_campylacanthum","Solanum_discolor","Mandragora_officinarum")
+#How many species, how many genera
+species<-unique(solanaceae.allrecords$fullname)
+total.species<-length(species)# 2258 total species
+genera<-unique(solanaceae.allrecords$genus)
+total.genera<-length(genera)#100 genera
 
-long1<-length(uncertain.species)
-for (i in 1:long1){
-	aux1<-which(solanaceae.allrecords$fullname==uncertain.species[i])
-	solanaceae.allrecords<-solanaceae.allrecords[-aux1,]
-}
+# Dataset created by RZF from CCDB and independent records by EEG
[…]

### 2019-07-09 basicdata/revbayesdataprocessing.R

@@ -1,0 +1,58 @@
+# Processing for Revbayes files
+library("ape")
+library("phytools")
+library("treeplyr")
+solanaceae.tree<-read.tree("~/Dropbox/solploidy/basicdata/solcorrecttip.tre")
+threestate.matrix <- read.csv("~/Dropbox/solploidy/basicdata/newstatesICDP.csv", as.is=T)
+threestate.matrix<-threestate.matrix[,-1]
+
+matched<-make.treedata(solanaceae.tree,threestate.matrix) #638 species!!!!
+write.nexus(matched$phy,file="~/Dropbox/solploidy/basicdata/fullmatchtree.nex")
+
+diploidsi<-which((matched$dat[,1]==1 & matched$dat[,2]==0 & matched$dat[,3]==0)==TRUE)# 144
+diploidsc<-which((matched$dat[,1]==0 & matched$dat[,2]==1 & matched$dat[,3]==0)==TRUE)#157
+polyploids<-which((matched$dat[,1]==0 & matched$dat[,2]==0 & matched$dat[,3]==1)==TRUE) #108
+
+diploidsna<-which((matched$dat[,1]==1 & matched$dat[,2]==1 & matched$dat[,3]==0)==TRUE) #170
+sc<-which((matched$dat[,1]==0 & matched$dat[,2]==1 & matched$dat[,3]==1)==TRUE)#70
+idcp<-which(matched$dat[,1]==1 & matched$dat[,2]==0 & matched$dat[,3]==1)#2
+
+## Data for threestate analyses
+threestate<- rep(0,dim(matched$dat)[1])
+threestate[diploidsc]="0"
+threestate[polyploids]="1"
+threestate[diploidsi]="2"
+threestate[sc]="(0 1)"
+threestate[diploidsna]<-"(0 2)"
+threestate[idcp]<-"(1 2)"
+threestate<-as.data.frame(threestate)
+row.names(threestate)<-matched$phy$tip.label
+
+write.table(threestate,file="~/Dropbox/solploidy/basicdata/threestate2.tsv",sep="\t")
+
+##binary ploidy
+binary.ploidy<-rep(0,dim(matched$dat)[1])
+binary.ploidy[diploidsc]<-"0"
+binary.ploidy[polyploids]<- "1"
+binary.ploidy[diploidsi]<-"0"
+binary.ploidy[sc]<-"?"
+binary.ploidy[diploidsna]<-"0"
[…]

### 2019-07-09 manuscript/introduction.tex
updated some text for the additional models

@@
 Dramatic increases in phylogenetic and phenotypic data and methodological advances have greatly accelerated the search for traits that influence diversification traits.
-Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a challenge \citep[\eg][]{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}.
+Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a challenge \citep[\eg][]{maddison_2015, rabosky_2015, beaulieu_2016, rabosky_2017}. 
 One difficulty is that speciation and extinction likely do not depend on a single trait, so the biological and environmental contexts in which traits occur can lead to complex interactions that affect lineage diversification \citep{beaulieu_2016, caetano_2018, herrera_2018}.
@@
 % E: We need a "Here, we..." here.  Otherwise, we've just told the reader "not one trait!" and then launch into one trait.
+% Could also state that we newly test whether adding one more trait is worthwhile.
 
@@
 Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies pervasive diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
-Our analyses re-examine the association between ploidy and lineage diversification by extending the approach of \citep{mayrose_2011, mayrose_2015} to include these additional processes.
-We allow for transitions both toward and away from polyploidy, and for heterogeneous diversification within each ploidy level, for example if diploids have a relatively higher net diversification rate if they were ancestrally polyploid than if they had no polyploid ancestry.
+Our analyses re-examine the association between ploidy and lineage diversification by extending the approach of \citet{mayrose_2011, mayrose_2015} to include these additional processes.
+We allow for transitions not only toward polyploidy, but also away from it.
+We also allow for heterogeneous diversification within each ploidy level, which could capture, for example, if diploids with a past history of polyploidy diversify more rapidly than diploids with no polyploid ancestry.
 
@@
 Polyploidy and self-fertilization are widely observed to be associated \citep{stebbins1950}.
-There are several possible general explanations: whole genome duplication may facilitate the transition to selfing by masking inbreeding depression, self-fertilization may facilitate polyploidy establishment by avoiding the triploid bridge, or both selfing and polyploidy may be correlates of weedy or peripheral populations \citep{ramsey_1998, barringer2007, barrett2008, husband2008}.
+There are several possible general explanations: whole genome duplication may facilitate the transition to selfing by masking inbreeding depression, self-fertilization may facilitate polyploidy establishment by avoiding the triploid bridge, or both selfing and polyploidy may tend to be found in weedy or peripheral populations \citep{ramsey_1998, barringer2007, barrett2008, husband2008}.
 Additionally, in RNase-based gametophytic SI systems, polyploidization alone directly causes the loss of SI \citep{stout1942, lewis1947}.
@@
 On the background of deletion of S-locus, or with heterologous chromosome homozygosity at the S-locus may not do so.
-These, however, appear to be sufficiently rare that they are recovered perhaps twice in dicots: once (or twice) naturally, in sour cherry and it's close relative (Prunus), and once in a experimentally induced and selected synthetic pentaploid Solanum (which has not been replicated).
+These, however, appear to be sufficiently rare that they are recovered perhaps twice in dicots: once (or twice) naturally, in sour cherry and its close relative (Prunus), and once in a experimentally induced and selected synthetic pentaploid Solanum (which has not been replicated).
 Strictly speaking, this is a feature of non-self recognition systems of the kind found in RNase-based GSI.
-In these systems, self-incompatibility is thought to occur because haploid self-pollen grains, with one S-allele at the locus controlling the SI response, are unable to detoxify the S-RNase produced by the same S-allele in the style \citep{kubo2010}.
+In these systems, self-incompatibility occurs because haploid self-pollen grains, with one S-allele at the locus controlling the SI response, are unable to detoxify the S-RNase produced by the same S-allele in the style \citep{kubo2010}.
 The unreduced pollen of diploids, however, can contain two S-alleles expressed in pollen, which jointly gain the ability to detoxify the S-RNases produced by any maternal genotype \citep{entani1999, tsukamoto2005, kubo2010}.
 Initial mutant individuals with pollen containing doubled haploid genomes are consequently capable of self-fertilization, with exceedingly few exceptions \citep{hauck_2002, nunes_2006}.
-%B: this was rephrased.
 This system of RNase-based SI is regarded as ancestral in eudicots \citep{igic_2001,steinbachs_2002}, and it is expressed in all SI species of Solanaceae examined to date.
@@
 Our macroevolutionary analyses in Solanaceae apply different combinations of these techniques in order to parse out the associations of ploidy and breeding system on speciation and extinction rates, while simultaneously allowing that other processes could influence diversification as well.
-% separate the associations of lineage diversification rates with breeding system alone, ploidy alone, the interaction between breeding system and ploidy, and additional unobserved factors.
-% Here, we examine the roles of ploidy and breeding system, as well as their interaction, on the diversification process in Solanaceae.
-% Importantly, our analyses aim to parse out the associations of these two traits on speciation and extinction rates while simultaneously allowing that other processes could influence diversification as well.
 We first present models in which the effect of ploidy and breeding system are considered separately, and we compare the inferences to previously published results. 
@@
 Finally, we model ploidy and breeding system jointly to assess their combined relationship to diversification, with or without an additional hidden trait.
+% E: say something here about testing whether it's worth adding in another real trait?
 With this joint trait model, we also quantify the relative contributions of the two pathways from SI diploids to SC polyploids.

### 2019-07-09 manuscript/methods.tex
updated some text for the additional models

@@
 %B: IMO: The data subsection would be better organized with parts, (1) ploidy, (2) breeding system, (3) tree, (4) models, (5) model selection and statistical inference.
-% E: I changed the subsectioning, but to a different arrangement: (1) Data, (2) Models, (3) Statistical inference, with subsubsections as necessary.
+% E: I changed the subsectioning, but to a different arrangement: (1) Data, (2) Models, (3) Statistical inference.  With subsubsections as necessary.
 
@@
 Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 595 species with ploidy and/or breeding system information on the tree.
-% FIXME: update numbers around here
+% FIXME: Rosana will update numbers around here
 Of these, 290 had information for both states.
@@
 
-Our second model addressed diversification due to ploidy differences while also parsing out heterogeneity of diversification rates due another factor, accommodated by adding a hidden state \citep[HiSSE,][]{beaulieu_2016}.
+To address diversification due to ploidy differences while also parsing out heterogeneity of diversification rates due another factor, we used models that include a hidden state \citep[HiSSE,][]{beaulieu_2016}.
 %B: inserted "(trait)", because that seems to be the goal - a biological explanation?
 E: I would say hisse captures some factor that evolves like a binary trait, not necessarily a real biological trait
-In this model, each of the observed diploid and polyploid states were subdivided by a binary hidden trait with states $A$ and $B$.
-In this $D/P+A/B$ ploidy and hidden state model, parameters include speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$) and extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$) for each subdivided state.
-The polyploidization rate is again represented by $\rho$, and the rates of transition between hidden states are symmetrical with parameter $\alpha$.  % FIXME ?
+In these models, the observed diploid and polyploid states were each subdivided by a binary hidden trait with states $A$ and $B$.
+In these $D/P+A/B$ ploidy and hidden state models, parameters include speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$) and extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$) for each subdivided state.
+In the simplest case, the polyploidization rate, $\rho$, is the same within each hidden state, and the rates of transition between hidden states are symmetrical with parameter $\alpha$.
+We also considered the more complex cases with different polyploidization rates in each hidden state ($\rho_A$ and $\rho_B$) and asymmetric transitions between hidden states ($\alpha$ and $\beta$).
 
-Diagrams of these two ploidy models are shown in \cref{figure:netdivall}AB.  % request from referee 2
+% E: Rosana, please check this paragraph.  Did I get it correct?
+To test if diversification rate heterogeneity is due only to a hidden trait and not the ploidy trait, we used character-independent diversification models \citep[CID,][]{beaulieu_2016}.
+Parameters in this case are speciation and extinction rates for the hidden states ($\lambda_A,\ \lambda_B,\ \mu_A,\ \mu_B$), along with polyploidization ($\rho$) and transitions between hidden states ($\alpha$).
+
+Diagrams of these ploidy models are shown in \cref{figure:netdivall}AB.  % request from referee 2
+% FIXME: ref to additional D/P models
 
@@
 This $I/C+A/B$ breeding system and hidden state model has parameters for state-specific speciation ($\lambda_{I_A},\ \lambda_{I_B},\ \lambda_{C_A},\ \lambda_{C_B}$) and extinction ($\mu_{I_A},\ \mu_{I_B},\ \mu_{C_A},\ \mu_{C_B}$), for loss of SI ($q_{IC}$), and for transitions between hidden states ($\alpha$).
-An similar model was used by \citet{freyman_2017} used to study diversification in Onagraceae, but here we restricted the hidden state transitions to be symmetrical. % FIXME still true?
+We also considered more complex models with different transition rates in each hidden state (???) and asymmetric transitions between hidden states ($\alpha$ and $\beta$). % FIXME
+Similar models were used by \citet{freyman_2017} used to study diversification in Onagraceae.
+
+% E: Rosana, please check this paragraph.  Did I get it correct?
+We further used character-independent diversification models to test if diversification rate heterogeneity is due only to a hidden trait and not the breeding system trait.
+Parameters in this case are speciation and extinction rates for the hidden states ($\lambda_A,\ \lambda_B,\ \mu_A,\ \mu_B$), along with loss of SI ($q_{IC}$) and transitions between hidden states ($\alpha$).
 
@@
 
-Diagrams of these two breeding system models are shown in \cref{figure:netdivall}CD.  % request from referee 2
+Diagrams of these breeding system models are shown in \cref{figure:netdivall}CD.  % request from referee 2
+% FIXME: ref to additional I/C models
 
@@
 The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
+% E: Can we call it something other than q_IC?  It's a bit confusing to have the same parameter mean different things in different models.
+
+% E: This paragraph is new.  Rosana, please check.
+We would like to test whether including an additional trait significantly improves the model fit.
+For example, does the model with both breeding system and ploidy provide a significantly better fit than the model with only ploidy?
+To answer this, it is not valid to compare the $IC/CD/CP$ and $D/P$ model fits, as defined above, because they use different data.
+Similarly, the $IC/CD/CP$ and $I/C$ models cannot be compared statistically.
+Instead, we define lumped models \citep{tarasov_2019} that use the state space of the $IC/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
+% should ref a figure
+
+% E: We now have so many models to describe, I'm feeling like all this text may not be the best way.  I'm wondering if we can rely on figures/tables instead, and only write out the overall strategy.
 
@@
 Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fit a simplified version with 16 parameters.
-% FIXME still true?
+% FIXME Rosana, please update about the parameters
 The reduction in parameter space was achieved by fixing the rates for transitions among hidden states to be equal with rate $\alpha$, and fixing the transition rates between observed states to be independent of the hidden state (rates $\rho_I,\rho_C,\ q_{IC}$ as defined for the $ID/CD/CP$ model).
@@
 
-Diagrams of these two ploidy and breeding system models are shown in \cref{figure:netdivall}EF.  % request from referee 2
+Diagrams of these ploidy and breeding system models are shown in \cref{figure:netdivall}EF.  % request from referee 2
+% FIXME: ref to additional models?
 
@@
 For example, there are 34 species that are SC with unknown ploidy. % FIXME: update number if necessary
-In the $D/P$ models these are coded as either $D$ or $P$ (uncertain/consistent with either state), in the $I/C$ models they are coded as $C$, and in the ID/CD/CP models they are coded as either $CD$ or $CP$.
+In the $D/P$ models these are coded as either $D$ or $P$ (uncertain/consistent with either state), in the $I/C$ models they are coded as $C$, and in the $ID/CD/CP$ models they are coded as either $CD$ or $CP$.
 In the hidden state models, all species were coded as consistent with both $A$ and $B$.

### 2019-07-10 manuscript/response.tex
Response to reviewers 07/09/19

@@
 
-Dear editor and reviewers:
+Dear Dr. Schoen and reviewers,
 
@@
 We appreciate your constructive suggestions, each of which we address below.
+You will find a big difference on the number of  models  that we have fitted for the revision based on your suggestions (originally 12 now 26).
+The two main reasons to increase the number of models were to first address one key question by Referees \#2 and \# 3 about the symmetry of the hidden states, and second to address a key question about more complex model.
+For the symmetry of the hidden states we found that is relevant to have asymmetric hidden state rates and in our case, it strengthened the differences amongst net diversification rates.
+The second question,  comes from a vague idea that we had while submitting the first version of this article: by increasing from two to three states the models, are we explaining better the history of diversification?
+This second question could not be addressed directly by model comparison statistics because the two-state and three-state models receive different input (i.e.
+(0,1) versus (0,1,2)), thus different data.
+However, the species and the tree are the same and should be delivering the same evidence despite any particular coding.
+During this revision, we figured a way to recode the three-state models to make them equivalent to the two-state models (the lumped models).
+Thanks to this equivalency, two-state and three-state models can now be compared to argue for the importance of adding key trait information and states in diversification analyses.
+
+We hope the changes and the new models are satisfactory and exciting to you and the reviewers.
 
@@
 % E: Pardon my speaking for you, Rosana. :)  Obviously change whatever wording below---starting to draft responses just helped me think through the requests.
+% R: I'm super excited now as you can see
 
@@
 \end{quote}
-
-Rosana? % TODO
+We acknowledge this is potentially a problem.
+It is true that chromosome number multiplicity can bias towards more (or less) transitions because it heavily depends on the most common cytotype.
+For the majority of species (~85\%)  the ploidy was not directly dictated by us but by the original sources of the data.
+In the originally sources ploidy has been classified not only most common cytotype but geographical distribution, multivalence, and/or  genome sizes.
+The remaining 15\% is what we assigned based on genus or family cytotype multiplicity.
+We cross-validated all of the assignments with the data from \citet{robertson_2011} where ploidy was assigned via a chromosome number change model using ChromEvol \citep{mayrose_2010}.
+We differ in exactly 35 species from \citet{robertson_2011} were we found a new reference or a new chromosome count.
+Classifying using a chromosome number change model only could also suffer from inflation of polyploidy events due to the lack of diploidization parameters.
 
@@
 \end{quote}
-
-Rosana?  Maybe one analysis with all rates different, to say there's so much uncertainty it doesn't add anything? % TODO
+We appreciate this suggestion and we followed.
+The main result after fitting models with symmetrical vs.
+asymmetrical rates is that the direction of the results does not change but the amount of difference amongst net diversification rates does.
+We found using bayes factors that a model including asymmetrical hidden rates is preferred over a model with equal hidden rates.
+Also,  the asymmetry in the trait rates did not add much to the differences according to the bayes factors.
+These results can be found on table and figures %TODO put number.
 
@@
 
-Ouch.  Rosana? % TODO
+Ouch.  Rosana? % TODO: Still need to calculate some stats here
 
@@
 \end{quote}
-
-Seems reasonable.  Rosana? % TODO
+We ran each model for 200,000 generations.
+For the main models with asymmetrical hidden rates all of the parameters have ESS of 300.
+For the most difficult models were all rates were different  the minimum is 125.
+Unfortunately, we don't have access to more of 96 hours of computational time to increase those but we have calculated the convergence as indicated above.
 
@@
 We can see there both the short-term and long-term behavior.
-% Rosana, what do you think?
+% Rosana, what do you think?- I think this is a good answer
 
@@
 
-Interesting. Rosana? % TODO
+The soaking noise comes from the hidden state so in the presence of hidden states finding that diploidization matters is key.
+All the bayes factors indicate strong support for models that include diploidization.
+However, the diploidzation parameter per se is difficult to estimate from ploidy data because of the quality of the classification.
+For those two reasons we are conservative about overstating the importance of diploidization.
 
@@
 
-Yes.
-We adjusted the text accordingly.
+Yes. We adjusted the text accordingly.
 
@@
 \end{quote}
-
-Rosana? % TODO
+We have now fitted models with asymmetrical hidden rates and asymmetrical trait rates.
+We found that the former are important and strengthened the differences.
+We found no evidence that having all asymmetrical rates is preferable over just asymmetrical hidden for our specific example.
+(Figs and table)
 
@@
 
-Interesting. Rosana? % TODO
+We have fitted the trait independent models (character-independent CID).
+We appreciated this suggestion because  we hadn't considered it and it allowed us to parse out the effect of the traits in the diversification process.
+By comparing bayes factors, models with traits are strongly preferred over CID (table )
+%TODO: write table number
 
@@
 
-Need a clear explanation. % TODO
+We have added the asymmetry on the rates. % TODO
 
@@
 \end{quote}
-
-check % TODO
+We have cited them accordingly. % TODO
 
@@
 Yes, and indeed this has already been shown.
-We fixed the text and added a reference.
+We fixed the text and added a reference. Nakov paper has not been accepted yet for publication or it has not been announced (July 9th, 2019).
 

### 2019-07-10 manuscript/methods.tex
Editing methods and adding the SSE models fig

@@
 After a first round of automatic cleaning, we examined results by hand and corrected records as necessary.
-Our hand-curated records were also contrasted against the ploidy dataset from \citet{robertson_2011} and against ploidy data in the C-value DNA dataset from \citet{bennett_2005}.
+Our hand-curated records were also contrasted against the ploidy dataset from \citet{robertson_2011}, original references therein, and against ploidy data in the C-value DNA dataset from \citet{bennett_2005}.
 By comparing three different sources of information, we were able to code taxa as diploid, $D$, or polyploid, $P$.
 For the majority of species, ploidy was assigned according to information from the original publications included in the  C-value DNA dataset \citep{bennett_2005}.
-For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus or based on SI/SC classification.
+For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus/family, or based on SI/SC classification.
 For example, \textit{Solanum betaceum} did not have information about ploidy level but it has 24 chromosomes, so since $x=12$ is the base chromosome number of the genus \textit{Solanum} \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid.
@@
 
-Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 595 species with ploidy and/or breeding system information on the tree.
-% FIXME: Rosana will update numbers around here
+Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 651 species with ploidy and/or breeding system information on the tree.
+% FIXME: Rosana will update numbers around here TODO
 Of these, 290 had information for both states.
 The number of species in each combination of states is summarized in \cref{figure:stateclassifications}A.
-We retained all 595 species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
+We retained all 651 species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
 A total of 405 taxa without any information about breeding system or ploidy were excluded.
@@
 
-% E: Rosana, please check this paragraph.  Did I get it correct?
+% E: Rosana, please check this paragraph.  Did I get it correct?- R: Yes that is correct
 To test if diversification rate heterogeneity is due only to a hidden trait and not the ploidy trait, we used character-independent diversification models \citep[CID,][]{beaulieu_2016}.
-Parameters in this case are speciation and extinction rates for the hidden states ($\lambda_A,\ \lambda_B,\ \mu_A,\ \mu_B$), along with polyploidization ($\rho$) and transitions between hidden states ($\alpha$).
+Parameters in this case are speciation and extinction rates for the hidden states ($\lambda_A,\ \lambda_B,\ \mu_A,\ \mu_B$), along with polyploidization ($\rho$) and transitions between hidden states ($(\alpha,\ \beta$).
 
@@
 This $I/C+A/B$ breeding system and hidden state model has parameters for state-specific speciation ($\lambda_{I_A},\ \lambda_{I_B},\ \lambda_{C_A},\ \lambda_{C_B}$) and extinction ($\mu_{I_A},\ \mu_{I_B},\ \mu_{C_A},\ \mu_{C_B}$), for loss of SI ($q_{IC}$), and for transitions between hidden states ($\alpha$).
-We also considered more complex models with different transition rates in each hidden state (???) and asymmetric transitions between hidden states ($\alpha$ and $\beta$). % FIXME
+We also considered more complex models with different transition rates in each hidden state and asymmetric transitions between hidden states ($\alpha$ and $\beta$). % FIXME
 Similar models were used by \citet{freyman_2017} used to study diversification in Onagraceae.
@@
 
-Ploidy and breeding system might influence lineage diversification individually, but these two traits also have an intricate association (see the Introduction).
+Ploidy and breeding system might influence lineage diversification individually, but these two traits also have an intricate association as discussed in the Introduction.
 Therefore, we considered a multi-state model that investigates the contribution of both traits and the allowable transitions between them.
-This $ID/CD/CP$ ploidy and breeding system model has states for self-incompatible diploids $(ID)$, self-compatible diploids $(CD)$, and self-compatible polyploids $(CP)$; self-incompatible polyploids do not occur, as explained in the Introduction.
+This $ID/CD/CP$ ploidy and breeding system model has states for self-incompatible diploids $(ID)$, self-compatible diploids $(CD)$, and self-compatible polyploids $(CP)$; self-incompatible polyploids do not occur.
 There are six parameters for state-specific diversification rates ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation; $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and three for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
@@
 % E: This paragraph is new.  Rosana, please check.
-We would like to test whether including an additional trait significantly improves the model fit.
-For example, does the model with both breeding system and ploidy provide a significantly better fit than the model with only ploidy?
-To answer this, it is not valid to compare the $IC/CD/CP$ and $D/P$ model fits, as defined above, because they use different data.
+When including a new trait the question on whether 
+the model with both breeding system and ploidy could provide a significantly better fit than the model with only ploidy was key.
+Under all the statistical model comparison approaches, it is not valid to compare the $IC/CD/CP$ and $D/P$ model fits, as defined above, because they use different coding (i.e. (0,1) versus (0,1,2)). 
 Similarly, the $IC/CD/CP$ and $I/C$ models cannot be compared statistically.
-Instead, we define lumped models \citep{tarasov_2019} that use the state space of the $IC/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
+However, the evidence in the species and the tree are the same for both two-state and three-state models.
+Instead, we define  and extend lumped models, as described first by \citet{tarasov_2019} for ontologies.
+The lumped models proposed use the state space of the $IC/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
 % should ref a figure
 
-% E: We now have so many models to describe, I'm feeling like all this text may not be the best way.  I'm wondering if we can rely on figures/tables instead, and only write out the overall strategy.
+% E: We now have so many models to describe, I'm feeling like all this text may not be the best way.
+I'm wondering if we can rely on figures/tables instead, and only write out the overall strategy.
+R- I feel the same but it might be very difficult to change at this point
 
@@
 We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018, huang_2018}.
-A fully parameterized version of this $ID/CD/CP+A/B$ model would have 26 rate parameters. 
-Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fit a simplified version with 16 parameters.
-% FIXME Rosana, please update about the parameters
-The reduction in parameter space was achieved by fixing the rates for transitions among hidden states to be equal with rate $\alpha$, and fixing the transition rates between observed states to be independent of the hidden state (rates $\rho_I,\rho_C,\ q_{IC}$ as defined for the $ID/CD/CP$ model).
-There are additionally twelve diversification rate parameters ($\lambda_{ID_A},  \lambda_{ID_B},  \lambda_{CD_A}, \lambda_{CD_B}, \lambda_{CP_A}, \lambda_{CP_B}, \mu_{ID_A}, \mu_{ID_B}, \mu_{CD_A}, \mu_{CD_B},  \mu_{CP_A}, \mu_{CP_B}$).
-%B: I can't tell, but it seems to me that this 26->16 parameter reduction makes it possible to underestimate the effect of these hidden states.Perhaps this complication is covered in the Discussion section.
-%R- It is possible, but adding all the parameters can also result in an underestimation of the hidden states contribution because it is not enough sample to accurately estimate transitions.
-In my simulations for diversification-free models, accurately estimating 10 parameters can only be achieved with trees with more than 500 tips.
-Because of that experience I think 16 parameters is already pushing really the inference machine hard.
-This happens because the phylogeny is not an independent sample so your ESS is much much smaller and you might end up over-parameterizing.
+We fitted then $ID/CD/CP+A/B$ models, one with equal hidden rate $\alpha$ and equal transition rates between observed states to be independent of the hidden state (rates $\rho_I,\rho_C,\ q_{IC}$ as defined for the $ID/CD/CP$ model); second model were the hidden states are asymmetrical, and a third model were all parameters were different.
+For each of these models, there are additionally twelve diversification rate parameters ($\lambda_{ID_A},  \lambda_{ID_B},  \lambda_{CD_A}, \lambda_{CD_B}, \lambda_{CP_A}, \lambda_{CP_B}, \mu_{ID_A}, \mu_{ID_B}, \mu_{CD_A}, \mu_{CD_B},  \mu_{CP_A}, \mu_{CP_B}$).
 
@@
 Code for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.
-We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $595/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
+We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $651/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
 For all models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
 Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
-For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run for 96 hours in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the 10 models.
+For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the  models.
 % FIXME see referee comment about reporting convergence, not runtime
@@
 
-We calculated the marginal likelihood for each of the ten models using fifty stepping stone steps under the methodology of \citet{xie_2010}, implemented in RevBayes \citep{hoehna_2016}.
-Each stepping stone step was found by calculating 500 generations of burn-in followed by a total of 1,000 MCMC steps (\cref{table:marginallike}).
-The calculation of each marginal likelihood ran for 24 hours on a high-performance computational cluster. % FIXME: remove?
+We calculated the marginal likelihood for each of the models using fifty stepping stone steps under the methodology of \citet{xie_2010}, implemented in RevBayes \citep{hoehna_2016}.
+Each stepping stone step was found by calculating at least 500 generations of burn-in followed by a total of 1,000 MCMC steps (\cref{table:marginallike}).
 
-Using the marginal likelihood values, we calculated thirteen different Bayes factors.
-Six compared the models of ploidy against one other ($D/P$ and $D/P+A/B$, each with or without diploidization), one compared the breeding system models ($I/C$ and $I/C+A/B$), and six compared the models with both traits ($ID/CD/CP$ and $ID/CD/CP+A/B$, each with or without diploidization) (\cref{table:bayesfactors}).
-%
-Other comparisons between these models are not valid because the input data are different under the different state space codings (\cref{figure:stateclassifications}B).
-In mathematical terms, the $D/P$, $I/C$, and $ID/CD/CP$ state spaces are not `lumpable' with respect to one another \citep{tarasov_2019}.
-% E: I'm honestly still confused about whether this is true.
-From ID/CD/CP, we can lump ID/CD to get D/P, and we can lump CD/CP to get I/C.
-FIXME
-Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $BF(M_0,\ M_1) = \ln[ P(\mathbf{X} | M_0) - P(\mathbf{X} | M_1)]$.
-There is `positive' support for $M_0$ when this value is more than 2, `strong' support when it is more than 6, and `very strong' support when it is more than 10 \citep{kass1995}.
+Using the marginal likelihood values, we calculated Bayes factors to argue five key biological and methodological questions:
+\begin{enumerate}
+\item Are hidden state necessary to add in this context?- Comparison of simple models against hidden state models
+\item Is a second trait adding information to the diversification process?- Comparison of lumped models against IC/CD/CP models
+\item Are diversification patterns only determined by hidden states and not the traits of interest?- Comparison of character independent models against hidden state 
+\item Is there evidence for diploidization?- Comparison amongst models with and without diploidization
+\item Do asymmetrical rates in the model affect the pattern of diversification? - Comparison amongst hidden states models with equal hidden rates and asymmetrical rates.
+\end{enumerate}
+
+Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $BF(M_1,\ M_0) = \ln[ P(\mathbf{X} | M_1) - P(\mathbf{X} | M_0)]$.
+There is `strong evidence'  for $M_1$ when this value is more than 2, `very strong' support when it is more than 6, and `extreme' support when it is more than 10 \citep{kass1995}.
 
@@
 %\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
-Model                     & Ploidy & Diploidization & Breeding  & Hidden & Num        & Marginal       \\
-                          &        &                & System    & State  & Parameters & Log-Likelihood \\ \midrule
-1. D/P                    & Yes    & No             & No        & No     & 5          & -1193.66\\
-2. D/P +$\delta$          & Yes    & Yes            & No        & No     & 6          & -1182.93 \\
-3. D/P+A/B                & Yes    & No             & No        & Yes    & 10         & -1150.99\\
-4. D/P+A/B+$\delta$       & Yes    & Yes            & No        & Yes    & 11         & \textbf{-1145.69}\\
-5. I/C                    & No     & No             & Yes       & No     & 5          & -1194.80 \\
-6. I/C+A/B                & No     & No             & Yes       & Yes    & 10         & \textbf{-1155.37}\\
-7. ID/CD/CP               & Yes    & No             & Yes       & No     & 9          & -1345.87\\
-8. ID/CD/CP+$\delta$      & Yes    & Yes            & Yes       & No     & 10         & -1344.50\\
-9. ID/CD/CP+A/B           & Yes    & No             & Yes       & Yes    & 15         & -1303.55 \\ 
-10. ID/CD/CP+A/B+$\delta$ & Yes    & Yes            & Yes       & Yes    & 16         & \textbf{-1300.35} \\
-\bottomrule
-\end{tabular}
-\caption{
-    The ten models and their marginal likelihoods.
-    Values in bold are for the best models within each class that are comparable (see \cref{table:bayesfactors}).
-    Abbreviations are D: diploid, P: polyploid, I: self-incompatible, C: self-compatible, A: one state of hidden trait, B: other state of hidden trait, $\delta$: diploidization.}
-\label{table:marginallike}
-\end{table}
+%Model                     & Ploidy & Diploidization & Breeding  & Hidden & Num        & Marginal       \\
+%                          &        &                & System    & State  & Parameters & Log-Likelihood \\ \midrule
+%1. D/P                    & Yes    & No             & No        & No     & 5          & -1193.66\\
+%2. D/P +$\delta$          & Yes    & Yes            & No        & No     & 6          & -1182.93 \\
+%3. D/P+A/B                & Yes    & No             & No        & Yes    & 10         & -1150.99\\
+%4. D/P+A/B+$\delta$       & Yes    & Yes            & No        & Yes    & 11         & \textbf{-1145.69}\\
+%5. I/C                    & No     & No             & Yes       & No     & 5          & -1194.80 \\
+%6. I/C+A/B                & No     & No             & Yes       & Yes    & 10         & \textbf{-1155.37}\\
+%7. ID/CD/CP               & Yes    & No             & Yes       & No     & 9          & -1345.87\\
+%8. ID/CD/CP+$\delta$      & Yes    & Yes            & Yes       & No     & 10         & -1344.50\\
+%9. ID/CD/CP+A/B           & Yes    & No             & Yes       & Yes    & 15         & -1303.55 \\ 
+%10. ID/CD/CP+A/B+$\delta$ & Yes    & Yes            & Yes       & Yes    & 16         & \textbf{-1300.35} \\
+%\bottomrule
+%\end{tabular}
+%\caption{
+   % The ten models and their marginal likelihoods.
+    %Values in bold are for the best models within each class that are comparable (see \cref{table:bayesfactors}).
+    %Abbreviations are D: diploid, P: polyploid, I: self-incompatible, C: self-compatible, A: one state of hidden trait, B: other state of hidden trait, $\delta$: diploidization.}
+%\label{table:marginallike}
+%\end{table}
 
@@
 \begin{tabular}{|c|c|c|}
-\toprule
-Ploidy Models & Breeding System Models & Ploidy and Breeding System Models \\ \midrule
-{\begin{tabular}{lcccc}
-            & 1 & 2 & 3 & 4 \\
-1. D/P$+\delta$ & $\cdot$     &10.72 &    -37.24    &-31.94\\
-2. D/P  &$\cdot$&$\cdot$ &    -47.97    &-42.66\\
-\textbf{3. D/P+A/B$+\delta$}  &$\cdot$  & $\cdot$&    $\cdot$    & 5.30 \\
-4. D/P+A/B  &$\cdot$& $\cdot$ & $\cdot$&$\cdot$ \\
-\end{tabular}
-}  & 
-{\begin{tabular}{lcc}
- & 5 & 6\\
-5. I/C &$\cdot$ & -39.43 \\
-\textbf{6. I/C+A/B} &$\cdot$& $\cdot$ \\
-& & \\
-& & \\
-\end{tabular}
-} & 
-{\begin{tabular}{lcccc}
-& 7 & 8 & 9 & 10\\
-7. ID/CD/CP$+\delta$ & $\cdot$    &1.36 & -44.15&-40.95 \\
-8. ID/CD/CP & $\cdot$ & $\cdot$ & -45.51 &-42.31 \\
-\textbf{9. ID/CD/CP+A/B+$\delta$}& $\cdot$ & $\cdot$ &$\cdot$    & 3.2 \\ 
-10. ID/P/CD+A/B & $\cdot$& $\cdot$&$\cdot$ & $\cdot$\\
-\end{tabular}
-}\\
 \bottomrule

### 2019-07-11 manuscript/tables.tex
File tables.tex added. Bayes Factors results complete

\documentclass{article}
\usepackage{amsmath}
\begin{document}

\begin{table}
\addtolength{\tabcolsep}{-5pt}
\begin{tabular}{|l|c|c|c|l|}
\hline
Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
M3. D/P+A/B                                 & -1234.52                     &    M3 vs. M4                 & 11.239                  & M4 (Strong)   \\
\textbf{M4. D/P+ A/B asym}            & -1223.28                     &   M4 vs. M5              & -1.658                  & M4 (Moderate)   \\
M5. D/P+A/B all asym                 & -1224.93                     &                     &                         &                         \\ \hline
M13. I/C+ A/B                                & -1270.47                     &   M13 vs. M14                  & 22.75                   & M14 (Strong)   \\
\textbf{M14. I/C+ A/B asym}            & \textbf{-1247.72}            & M14 vs. M15         & {0.05}           & No evidence             \\
M15. I/C+ A/B all asym               & -1247.67                     &                     &                         &                         \\ \hline
M18. IC/CD/CP+A/B                            & -1414.00                     &     M18 vs. M19                & 20.79                   & M19 (Strong)    \\
\textbf{M19. IC/CD/CP+ A/B asym}       & \textbf{-1393.21}            & M19 vs. M20          & {-1.919}         & M19 (Moderate)   \\
M20. IC/CD/CP+ A/B all asym           & -1395.129                    &                     &                         &                         \\ \hline
M8. D/P +A/B+$\delta$                          & -1225.05                     &  M8 vs. M9                   & 10.58                   & M9 (Strong)    \\
\textbf{M9. D/P+ A/B+ $\delta$ asym}      & \textbf{-1214.46}                     &    M9 vs. M10                 & -0.52                   & M10 (Moderate) \\
M10. D/P+A/B+$\delta$ all asym        & -1214.98                     &                    &                         &                         \\ \hline
M23. IC/CD/DP+A/B+$\delta$                 & -1408.65                     &     M23 vs M24                & 21.91                   & M24(Strong)     \\
\textbf{M24. IC/CD/DP+A/B+$\delta$ asym} & \textbf{-1386.74}                     &   M24 vs M25              & -2.78                   & M24 (Moderate) \\
M25. IC/CD/DP+A/B+ $\delta$ all asym     & -1389.52                     &                     &                         &                         \\ \hline
\end{tabular}
\caption{Test for symmetry of the hidden and trait rates via Bayes factors. For all models asymmetric hidden transition rates are preferred over models with equal hidden rates. Adding more complexity by assuming asymmetry in the trait rate within hidden states does not improve models.}
\label{table:testsymmetry}
\end{table}

\begin{table}
\addtolength{\tabcolsep}{-5pt}
\begin{tabular}{|l|c|c|c|c|}
\hline
Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
M1. D/P                                  & -1238.76                                                          & M1 vs. M4   & 60.47           & M4 (Strong)                                                          \\
\textbf{M4. D/P+A/B asym}                & \textbf{-1223.28}                                                &             &                 &                                                                      \\ \hline
M11. I/C                                 & -1309.07                                                          & M11 vs. M14 & 61.35           & M14 (Strong)                                                         \\
\textbf{M14. I/C+A/B asym}               & \textbf{-1247.72}                                                 &             &                 &                                                                      \\ \hline
M16. ID/CD/CP                            & -1459.11                                                          & M16 vs. M19 & 65.90           & M19 (Strong)                                                         \\
\textbf{M19. ID/CD/CP+A/B asym}          & \textbf{-1393.20}                                                &             &                 &                                                                      \\ \hline
M6. D/P+$\delta$                         & -1283.76                                                          & M6 vs. M9   & 69.3            & M9 (Strong)                                                          \\
\textbf{M9. D/P+A/B+$\delta$ asym}       & \textbf{-1214.46}                                                 &             &                 &                                                                      \\ \hline
M21. IC/CD/CP+$\delta$                   & -1454.68                                                          & M21 vs. M24 & 68.48           & M24 (Strong)                                                         \\
\textbf{M24. IC/CD/CP+A/B+$\delta$ asym} & \textbf{-1386.20}                                                 &             &                 &                                                                      \\ \hline
\end{tabular}
\caption{Test for addition of hidden states in models via Bayes factors. Models with hidden states are strongly preferred over simpler models}
\label{table:testaddhidden}
\end{table}

\begin{table}
\addtolength{\tabcolsep}{-5pt}
\begin{tabular}{|l|c|c|c|c|}
\hline
Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
M1. D/P                                  & -1238.76                                                          & M1 vs. M6   & 65.92           & M6 (Strong)                                                          \\
\textbf{M6. D/P+$\delta$}                & \textbf{-1267.84}                                                 &             &                 &                                                                      \\ \hline
M4. D/P+A/B asym                         & -1223.28                                                          & M4 vs. M9   & 8.81            & M9 (Moderate)                                                        \\
\textbf{M9. D/P+A/B+$\delta$ asym}       & \textbf{-1214.46}                                                 &             &                 &                                                                      \\ \hline
M16. ID/CD/CP                            & -1459.11                                                          & M16 vs. M21 & 4.41            & M21 (Moderate)                                                       \\
\textbf{M21. ID/CD/CP+$\delta$}          & \textbf{-1454.68}                                                 &             &                 &                                                                      \\ \hline
M19. IC/CD/CP asym                       & -1393.20                                                          & M19 vs. M24 & 6.46            & M24 (Moderate)                                                       \\
\textbf{M24. IC/CD/CP+A/B+$\delta$ asym} & \textbf{-1386.20}                                                 &             &                 &                                                                      \\ \hline
\end{tabular}
\caption{Test for inclusion of a diploidization rate via Bayes factors. Models with diploidzation are moderately preferred over models that do not include a diploidization rate}
\label{table:testdiploidization}
\end{table}

\begin{table}[]
\addtolength{\tabcolsep}{-5pt}
\begin{tabular}{|l|c|c|c|c|}
\hline
Model                       & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
M26. Lumped D/P             & -1463.22                                                          & M26 vs. M16 & 4.10            & M16 (Moderate)                                                       \\
\textbf{M16. ID/CD/CP}      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
M27. Lumped D/P+A/B         & -1417.67                                                          & M27 vs. M23 & 3.678           & M23 (Moderate)                                                       \\
\textbf{M23. ID/CD/CP +A/B} & \textbf{-1414.00}                                                 &             &                 &                                                                      \\ \hline
M28. Lumped I/C             & -1458.41                                                          & M28 vs. M16 & -0.69           & No evidence                                                          \\
\textbf{M16. ID/CD/CP}      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
M29. Lumped I/C+ A/B        & -1416.60                                                          & M29 vs. M23 & 2.60            & M23 (Moderate)                                                       \\
\textbf{M23. IC/CD/CP+A/B}  & \textbf{-1414.00}                                                 &             &                 &                                                                      \\ \hline
\end{tabular}
\caption{Test for adding a focal trait to a binary state model via Bayes factors. Three-state modes are moderately preferred over two-state models. The evidence is stronger towards the inclusion of breeding system in ploidy models (M26, M27) than the inclusion of ploidy in breeding system models (M29, M29)}
\label{table:testextrastate}
\end{table}

\begin{table}[]
\addtolength{\tabcolsep}{-5pt}
\begin{tabular}{|l|c|c|c|c|}
\hline
Model                                     & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
M2.  CID D/P                              & -1223.86                                                          & M2 vs. M4   & 0.58            & No evidence                                                          \\
\textbf{M4. D/P +A/B asym}                & \textbf{-1223.28}                                                 &             &                 &                                                                      \\ \hline
\textbf{M7. CID D/P+$\delta$}             & \textbf{-1212.42}                                                 & M7 vs. M9   & -2.04           & M7(Moderate)                                                         \\
M9. D/P+A/B+$\delta$ asym                 & -1214.47                                                          &             &                 &                                                                      \\ \hline
M12. CID I/C                              & -1267.93                                                          & M12 vs. M14 & 20.22           & M14 (Strong)                                                         \\
\textbf{M14. I/C+ A/B asym}               & \textbf{-1247.72}                                                 &             &                 &                                                                      \\ \hline
M22. CID ID/CD/CP+$\delta$                & -1399.20                                                          & M22 vs. M24 & 12.45           & M24 (Strong)                                                         \\
\textbf{M24. ID/CD/CP +A/B+$\delta$ asym} & \textbf{-1386.74}                                                 &             &                 &                                                                      \\ \hline
\end{tabular}
\caption{Test for the inclusion of focal traits in diversification models via Bayes factors. Models that include ploidy (M4, M9) are moderately or non-preferred over models where the diversification is independent of the ploidy trait (M2, M4). Models that include breeding system are strongly preferred over models where the diversification process is independent of the breeding system trait. A comparison between model M17 versus M19 is missing because of challenges calculating the log-likelihood marginal for M17.}
\label{table:cidtests}
\end{table}

\end{document}

### 2019-07-11 basicdata/matrixdataprocessing (Autosaved).R
Number of taxa in states updated

@@ -360,7 +360,7 @@ b_tmp %in% a_tmp 
 dat2[639:651,]
 #651 species!!!!
 #diploidsi 123+ 4 =127
-# 160 +3=163
+# diploidsc 160 +3=163
 # diploidsna 178
 # polyploidsc<-which(matched$dat[,4]>2 & matched$dat[,5]=="SC")#92
 # polyploidsna 17+5=22

### 2019-07-11 manuscript/methods.tex
updated the figure of all models
Changed the layout and re-numbered the models accordingly.

The figure is now Fig S1.  It won't fit on a normal-size page, so it'll
have to be a separate supp file.  For now, the Makefile attaches the
giant figure as the final page of main.pdf.
edits to Models section of Methods
Relying now on the giant figure of all models.  The text isn't much
shorter, but I think its logic is clearer.

@@
 
-We employed a suite of SSE-style models \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016} to test associations between ploidy, breeding system, and lineage diversification.
-The full set of models is listed in \cref{table:marginallike}.
+To test our hypotheses about lineage diversification and trait macroevolution, we fit a large suite of phylogenetic models.
+All models are from the SSE framework \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016}.
+This style of model has parameters that describe per-lineage rates of speciation and extinction, specific to each character state (denoted $\lambda$ and $\mu$, respectively, with subscripts to indicate the state), along with rates of transitions between states (denoted $\rho$ for polyploidization, $\delta$ for diploidization, and $q_{IC}$ for loss of self-incompatibility).
+
+The full set of models and all their rate parameters are detailed in \cref{fig:allmodels}.
+Here we summarize how each type of model allows us to assess whether diversification is best explained by ploidy, breeding system, their combination, or some unknown factor.
 
@@
 
-In order to investigate the association between ploidy level and diversification, we first defined a binary state speciation and extinction model \citep[BiSSE,][]{maddison_2007} in which taxa were classified as diploid $(D)$ or polyploid $(P)$.
-In this $D/P$ ploidy model, the model parameters are speciation rates ($\lambda_D$, $\lambda_P$) and extinction rates ($\mu_D$, $\mu_P$) for each state, and a rate of transition from D to P ($\rho$).
+In order to investigate the association between ploidy level and diversification, we first defined a model in which taxa were classified as diploid $(D)$ or polyploid $(P)$ (M1).
 This is the same model \citet{mayrose_2011} used for this question across an array of plant genera.
 % R- Itay wanted to hightlight that is the same model. Definitely a BiSSE. He did the ClaSSE later in the 2014 response.
+Although this model can be a powerful means to distinguish state-specific speciation and extinction from trait changes, it is prone to incorrectly reporting that a trait is associated with diversification differences \citep{maddison_2015, rabosky_2015}.
+We therefore define several models that incorporate additional forms of diversification rate heterogeneity.
 
-% E: I moved r and nu to later, under statistical inference.  It made it seem like there were too many parameters here.
-% I also removed italics around the model names because we don't actually use those names later on (more just the state codes).
+Our second ploidy model includes a binary hidden trait that subdivides each observed state (M2).
+In this CID model \citep{beaulieu_2016}, the hidden trait can affect diversification but the observed trait does not.
+Comparing M1 and M2 allows us to test whether diversification rate heterogeneity is better explained by ploidy or by some unknown factor.
 
-To address diversification due to ploidy differences while also parsing out heterogeneity of diversification rates due another factor, we used models that include a hidden state \citep[HiSSE,][]{beaulieu_2016}.
-%B: inserted "(trait)", because that seems to be the goal - a biological explanation?
-E: I would say hisse captures some factor that evolves like a binary trait, not necessarily a real biological trait
-In these models, the observed diploid and polyploid states were each subdivided by a binary hidden trait with states $A$ and $B$.
-In these $D/P+A/B$ ploidy and hidden state models, parameters include speciation rates ($\lambda_{D_A},\ \lambda_{D_B},\ \lambda_{P_A},\ \lambda_{P_B}$) and extinction rates ($\mu_{D_A},\ \mu_{D_B},\ \mu_{P_A},\ \mu_{P_B}$) for each subdivided state.
-In the simplest case, the polyploidization rate, $\rho$, is the same within each hidden state, and the rates of transition between hidden states are symmetrical with parameter $\alpha$.
-We also considered the more complex cases with different polyploidization rates in each hidden state ($\rho_A$ and $\rho_B$) and asymmetric transitions between hidden states ($\alpha$ and $\beta$).
+We additionally fit three models in which both ploidy and a hidden trait could influence diversification (M3--M5).
+These models differ in whether transitions between the hidden states are symmetric or asymmetric, and whether the polyploidization rate depends on the hidden state.
+Comparing M1--M5 allows us to test whether ploidy is associated with diversification differences on top of differences explained by an unknown factor.
 
-% E: Rosana, please check this paragraph.  Did I get it correct?- R: Yes that is correct
-To test if diversification rate heterogeneity is due only to a hidden trait and not the ploidy trait, we used character-independent diversification models \citep[CID,][]{beaulieu_2016}.
-Parameters in this case are speciation and extinction rates for the hidden states ($\lambda_A,\ \lambda_B,\ \mu_A,\ \mu_B$), along with polyploidization ($\rho$) and transitions between hidden states ($(\alpha,\ \beta$).
-
-Diagrams of these ploidy models are shown in \cref{figure:netdivall}AB.  % request from referee 2
-% FIXME: ref to additional D/P models
+We further fit the analogues of these five models but including a rate parameter for transitions from polyploid to diploid (M6--M10).
+This allows us to assess whether our conclusions about ploidy and diversification are robust to the possibility of diploidization.
 
@@
 
-To assess the effects of breeding system in the diversification process, we first fit a model with self-incompatible $(I)$ or self-compatible $(C)$ states.
+Our first five breeding system models follow the same logic as the first ploidy models.
+We begin with each species classified as self-incompatible $(I)$ or self-compatible $(C)$ (M11).
 This is the same as the analysis of \citet{goldberg_2010} but with an updated phylogeny \citep{sarkinen_2013}.
-This $I/C$ breeding system model has parameters for state-specific speciation ($\lambda_I,\ \lambda_C$) and extinction ($\mu_I,\ \mu_C$), and for transitions from $I$ to $C$ ($q_{IC}$).
 
-To parse out the effect of breeding system on diversification while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of the breeding system states into hidden states $A$ and $B$.
-This $I/C+A/B$ breeding system and hidden state model has parameters for state-specific speciation ($\lambda_{I_A},\ \lambda_{I_B},\ \lambda_{C_A},\ \lambda_{C_B}$) and extinction ($\mu_{I_A},\ \mu_{I_B},\ \mu_{C_A},\ \mu_{C_B}$), for loss of SI ($q_{IC}$), and for transitions between hidden states ($\alpha$).
-We also considered more complex models with different transition rates in each hidden state and asymmetric transitions between hidden states ($\alpha$ and $\beta$). % FIXME
+We then add models to allow diversification to be influenced by only a hidden trait (M12), or by both breeding system and a hidden trait (M13--M15, with varying degrees of complexity in the hidden trait transitions).
 Similar models were used by \citet{freyman_2017} used to study diversification in Onagraceae.
-
-% E: Rosana, please check this paragraph.  Did I get it correct?
-We further used character-independent diversification models to test if diversification rate heterogeneity is due only to a hidden trait and not the breeding system trait.
-Parameters in this case are speciation and extinction rates for the hidden states ($\lambda_A,\ \lambda_B,\ \mu_A,\ \mu_B$), along with loss of SI ($q_{IC}$) and transitions between hidden states ($\alpha$).
 
@@
 
-Diagrams of these breeding system models are shown in \cref{figure:netdivall}CD.  % request from referee 2
-% FIXME: ref to additional I/C models
-
 \subsubsection{Ploidy, breeding system, and diversification}
 
-Ploidy and breeding system might influence lineage diversification individually, but these two traits also have an intricate association as discussed in the Introduction.
-Therefore, we considered a multi-state model that investigates the contribution of both traits and the allowable transitions between them.
-This $ID/CD/CP$ ploidy and breeding system model has states for self-incompatible diploids $(ID)$, self-compatible diploids $(CD)$, and self-compatible polyploids $(CP)$; self-incompatible polyploids do not occur.
-There are six parameters for state-specific diversification rates ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation; $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and three for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
-The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
+Ploidy and breeding system might influence lineage diversification individually, but these two traits also have an intricate association (discussed in the Introduction).
+Therefore, we considered several multi-state models that investigate the contribution of both traits and the allowable transitions between them.
+
+The simplest classifies each species as either self-incompatible diploid $(ID)$, self-compatible diploid $(CD)$, or self-compatible polyploid $(CP)$; recall that self-incompatible polyploids do not occur (M16).
+Each of these states may again be associated with different rates of speciation and extinction, and the allowable transitions are loss of SI within the diploid state (from $ID$ to $CD$), loss of SI via polyploidization (from $ID$ to $CP$), and polyploidization while SC (from $CD$ to $CP$).
+% The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
 % E: Can we call it something other than q_IC?  It's a bit confusing to have the same parameter mean different things in different models.
 
-% E: This paragraph is new.  Rosana, please check.
-When including a new trait the question on whether 
-the model with both breeding system and ploidy could provide a significantly better fit than the model with only ploidy was key.
-Under all the statistical model comparison approaches, it is not valid to compare the $IC/CD/CP$ and $D/P$ model fits, as defined above, because they use different coding (i.e. (0,1) versus (0,1,2)). 
-Similarly, the $IC/CD/CP$ and $I/C$ models cannot be compared statistically.
-However, the evidence in the species and the tree are the same for both two-state and three-state models.
-Instead, we define  and extend lumped models, as described first by \citet{tarasov_2019} for ontologies.
-The lumped models proposed use the state space of the $IC/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
-% should ref a figure
+As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by both ploidy and breeding system as well as a hidden trait (M18--M20, with varying degrees of complexity in the hidden trait transitions).
+We add this hidden trait layer on top of our three-state model \citep[similar to][]{caetano_2018, huang_2018}.
+We also fit the analogous models but allowing for diploidization (M21--M25).
 
-% E: We now have so many models to describe, I'm feeling like all this text may not be the best way.
-I'm wondering if we can rely on figures/tables instead, and only write out the overall strategy.
-R- I feel the same but it might be very difficult to change at this point
+\subsubsection{Lumped models}
 
-Although the $ID/CD/CP$ model allows for two traits to potentially influence diversification, other unknown factors could as well.
-We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018, huang_2018}.
-We fitted then $ID/CD/CP+A/B$ models, one with equal hidden rate $\alpha$ and equal transition rates between observed states to be independent of the hidden state (rates $\rho_I,\rho_C,\ q_{IC}$ as defined for the $ID/CD/CP$ model); second model were the hidden states are asymmetrical, and a third model were all parameters were different.
-For each of these models, there are additionally twelve diversification rate parameters ($\lambda_{ID_A},  \lambda_{ID_B},  \lambda_{CD_A}, \lambda_{CD_B}, \lambda_{CP_A}, \lambda_{CP_B}, \mu_{ID_A}, \mu_{ID_B}, \mu_{CD_A}, \mu_{CD_B},  \mu_{CP_A}, \mu_{CP_B}$).
+The models above allow us to assess the contributions of our two focal traits---ploidy and breeding system---to lineage diversification.
+They do not, however, reveal whether it was ``worth it'' to include both traits in the analysis.
+That is, they do not allow us to perform statistical model comparisons to test whether diversification heterogeneity is better explain by both traits than by either one of them.
+This is because the models in the three sections above use different state spaces: $D/P$ versus $I/C$ versus $IC/CD/CP$.
+Model comparisons work with the likelihood of the data given the model, but they are not valid when the data used differ from model to model.
 
-Diagrams of these ploidy and breeding system models are shown in \cref{figure:netdivall}EF.  % request from referee 2
-% FIXME: ref to additional models?
+In order to compare model fits for ploidy only versus breeding system only versus both traits, we use the technique of \citet{tarasov_2019} for lumping states together.
+We use the state space of the $IC/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
+First we lump together $ID$ and $CD$ to form the diploid state, mimicking the $D/P$ model (M26).
+Then we lump together $CD$ and $CP$ to form the self-compatible state, mimicking the $I/C$ model (M28).
+We further add a hidden trait to each of these models (M27 and M29).
+
+We do not include additional models with diploidization because this reverse transition renders the models non-lumpable \citep{tarasov_2019}.  % E: further explanation needed?
 
@@
 Considering ploidy and breeding system together, there are two evolutionary pathways from SI diploid to SC polyploid \citep{brunet2001, robertson_2011}.
-In the one-step pathway, the CP state is produced directly from the ID state when whole genome duplication disables SI.
-In the two-step pathway, the CD state is an intermediate: SI is first lost, and later the SC diploid undergoes polyploidization.
-We quantify the relative contribution of these pathways to polyploidy in two ways, each using the maximum \textit{a posteriori} (MAP) estimates of rates from the $IC/CD/CP$ model.
+In the one-step pathway, the $CP$ state is produced directly from the $ID$ state when whole genome duplication disables SI.
+In the two-step pathway, the $CD$ state is an intermediate: SI is first lost, and later the SC diploid undergoes polyploidization.
+We quantify the relative contribution of these pathways to polyploidy in two ways, each using the maximum \textit{a posteriori} (MAP) estimates of rates from the simplest model that includes both traits (M16).
 
@@
 
-\subsubsection{Diploidization as an exploratory hypothesis}
-
-The models described above allow polyploidization but not its reverse.
-We explored the effects of allowing diploidization by adding a parameter $\delta$ for transitions from $P$ back to $D$ (in the $D/P$ and $D/P+A/B$ models) or from $CP$ back to $CD$ (in the $I/C$ and $I/C+A/B$ models); we did not allow transitions from $CP$ back to $ID$ because that would include regain of SI.
-Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010, glick2014, mayrose_2015, freyman_2017}, which assume only aneuploid change and polyploidization.
-Our ploidy classifications, however, when possible also incorporated information on chromosome multiplicity at the genus level and author conclusions within original studies with alternative sources of information (\eg geographic distribution, genus ploidy distribution).
-Since it is not known whether diploidization can be detected under these additional ploidy classifications, we fit models with and without diploidization in order to test whether the conclusions about diversification are sensitive to including diploidization.
-% As discussed by \citet{servedio_2014}, the presence or absence of a process can have an exploratory goal.
-In our models, the diploidization parameter (or its absence, $\delta=0$), represents an opportunity to explore a possibly significant assumption that may not decisively affect interactions among ploidy, breeding system, and diversification.
+% E: I don't think we need this section now that diploidization is in each of the sections above.  But I'm saving this text in case we want to put some of it in Results.
+%--------------------------------------------------
+% \subsubsection{Diploidization as an exploratory hypothesis}
+%
+% The models described above allow polyploidization but not its reverse.
+% We explored the effects of allowing diploidization by adding a parameter $\delta$ for transitions from $P$ back to $D$ (in the $D/P$ and $D/P+A/B$ models) or from $CP$ back to $CD$ (in the $I/C$ and $I/C+A/B$ models); we did not allow transitions from $CP$ back to $ID$ because that would include regain of SI.
+% Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010, glick2014, mayrose_2015, freyman_2017}, which assume only aneuploid change and polyploidization.
+% Our ploidy classifications, however, when possible also incorporated information on chromosome multiplicity at the genus level and author conclusions within original studies with alternative sources of information (\eg geographic distribution, genus ploidy distribution).
+% Since it is not known whether diploidization can be detected under these additional ploidy classifications, we fit models with and without diploidization in order to test whether the conclusions about diversification are sensitive to including diploidization.
+% In our models, the diploidization parameter (or its absence, $\delta=0$), represents an opportunity to explore a possibly significant assumption that may not decisively affect interactions among ploidy, breeding system, and diversification.
+%--------------------------------------------------
 
@@
 
-\begin{table}
-\begin{tabular}{lcccccc} \toprule
+\begin{suppfigure}
+    \caption{ All models.  [This figure is provided as a separate, large-format page.] }
+    \label{fig:allmodels}
+\end{suppfigure}
+
+%%% Below here is old tables.  I think the new versions should instead go in results.tex or a separate file. %%%
+
+% \begin{table}
+% \begin{tabular}{lcccccc} \toprule
 % \begin{tabular}{@{}lcccccc@{}} \toprule

### 2019-07-11 manuscript/response.tex
Editing more response to reviewers

--- 2019-05-16 manuscript/response.tex
@@
 We appreciate your constructive suggestions, each of which we address below.
+You will find a big difference on the number of  models  that we have fitted for the revision based on your suggestions (originally 12 now 29).
+The two main reasons to increase the number of models were to first address one key question by made by all referees about the symmetry of the hidden states, and second to address a remaining question about adding complexity to two-state models.
+For the symmetry of the hidden states  question we found that is relevant to have asymmetric hidden state rates and in our case, it strengthened the differences amongst net diversification rates.
+The second question,  comes from a vague idea that we had while submitting the first version of this article: by increasing from two to three states the models, are we explaining better the history of diversification?
+This second question could not be addressed directly by model comparison statistics because the two-state and three-state models need different input (i.e.
+(0,1) versus (0,1,2)).
+However, the species and the tree are the same and our intuition indicated that they should carry the same biological evidence despite any particular coding needed as input.
+During this revision, we figured a way to recode the three-state models to make them equivalent to the two-state models (the new lumped models).
+Thanks to this equivalency, two-state and three-state models can now be compared to argue for the importance of adding key trait information and states in diversification analyses.
+We also revised the original dataset thanks to the suggestion of Reviewer 1, and we were able to correct discrepancies with previous dataset as well as add more sample (651 taxa).
+
+We hope the changes and the new models are satisfactory and exciting to you and the reviewers.
 
@@
 \end{quote}
-
-Rosana? % TODO
+We acknowledge this is a potential problem.
+It is true that chromosome number multiplicity can bias coding towards more (or less) transitions when it heavily depends on the most common cytotype.
+For the majority of species (~85\%)  the ploidy was not directly dictated by us but by the original sources of the data.
+In the originally sources ploidy has been classified not only most common cytotype but geographical distribution, multivalence, and/or  genome sizes.
+The remaining 15\% of the sample is what we assigned based on genus or family cytotype multiplicity.
+We cross-validated all of the assignments with the data from \citet{robertson_2011} where ploidy was assigned via a chromosome number change model using ChromEvol \citep{mayrose_2010}.
+We differ in exactly 35 species from \citet{robertson_2011} were we found a new reference or a new chromosome count.
+Classifying using a chromosome number change models only could also suffer from inflation of polyploidy events due to the lack of diploidization parameters, we found some specific species in the  \citet{robertson_2011} sample that were clearly aneuploids (one chromosome difference) that were classified as polyploids using chromosome number models, we manually corrected for those hoping to minimize obvious mistakes.
 
@@
 
-Rosana?  Maybe one analysis with all rates different, to say there's so much uncertainty it doesn't add anything? % TODO
+We appreciate this suggestion and we followed it.
+The main result after fitting models with symmetric vs.
+asymmetric rates is that the direction  of the results does not change but the amount of difference amongst net diversification posterior distributions does.
+We found using Bayes factors that a model including asymmetric hidden rates is preferred over a model with equal hidden rates.
+Also,  the asymmetry in the trait rates have moderate to not support of being better than a model with equal trait transition rates.
+%TODO put number.
 
@@
 \end{quote}
+Unfortunately, we weren't able to run multiple chains for all models due to some limitations to our computer resources.
+However, we were able to test for the convergence of  three key models using Gelman's $\hat{R}$ test statistic.
+We tested convergence with two runs of 130,000 generations (after 5,000 generations of burn-in) for models D/P (BiSSE), D/P+A/B asym (HiSSE) and ID/CD/CP+A/B asym (MuHiSSE).
+The statistics as well as the Gelman plots in  \ref{figure:convergencestats} show that with that number of generations the models were converging.
+We increased the number of generations to be safe for all the models that we did not test and visually assessed the mixing of  models on top of demanding ESS larger than 200 for parameters.
 
-Ouch.  Rosana? % TODO
+\begin{figure}
+    \centering 
+      \begin{minipage}[b]{0.33\textwidth}
+    \includegraphics[width=\textwidth]{Gelmanbissenodip.pdf} 
+    \end{minipage}
+    \begin{minipage}[b]{0.33\textwidth}
+        \includegraphics[width=\textwidth]{Gelmanhissenodipasym.pdf}
+        \end{minipage}
+    \begin{minipage}[b]{0.33\textwidth}
+             \includegraphics[width=\textwidth]{Gelmanmuhissenodipasym.pdf}
+                 \end{minipage}
+    \caption{ Convergence statistics and Gelman plots for D/P (BiSSE), D/P+A/B asym (HiSSE) at the top and ID/CD/CP+A/B asym (MuHiSSE,bottom).
+Upper limits of Gelman's test statistics should be close to 1 for an MCMC to be considered convergent.}
+    \label{figure:convergencestats}
+\end{figure}
 
@@
 \end{quote}
-
-Seems reasonable.  Rosana? % TODO
+We ran each model for 200,000 generations. Unfortunately, we don't have access to more of 96 hours of computational time to increase those but we have calculated the convergence as indicated above.
 
@@
 We can see there both the short-term and long-term behavior.
-% Rosana, what do you think?
+% Rosana, what do you think?- I think this is a good answer- I think it is good
 
@@
 
-Interesting. Rosana? % TODO
+The soaking noise comes from the hidden state so in the presence of hidden states finding that diploidization matters is key.
+All the Bayes factors indicated strong support for models that include diploidization.
+However, the diploidzation parameter per se is difficult to estimate (wide credible intervals) mainly because diploidization is a genome size process that can be partially or minimally recover from ploidy data.
+This is the reason why we do not overstate this finding,  the importance of diploidization needs to be studied using genomes.
 
@@
 \end{quote}
-
-Rosana? % TODO
+We have now fitted models with asymmetric hidden rates and asymmetric trait rates.
+We found that the former are important and strengthened the differences.
+We found no evidence that having all asymmetric rates is preferable over just asymmetric hidden for our specific example.
+%TODO (Figs and table)
 
@@
 
-Fixed. % TODO
+Fixed. % Done: Emma did the new figure
 
@@
 
-Interesting. Rosana? % TODO
+We have fitted the trait independent models (character-independent CID).
+We appreciated this suggestion because it allowed us to parse out the effect of the traits in the diversification process.
+By comparing Bayes factors, models with traits are strongly preferred over CID when the breeding system is the focal trait, and there is little to no evidence that models that include ploidy are better than CID.
+These results strongly suggest that breeding system is much more informative than ploidy in the diversification dynamics of Solanaceae.
+%TODO: write table number
 
@@
 
-Need a clear explanation. % TODO
+When we presented the first submission we were pushing the limits on implementations of diversification models in RevBayes.
+During the revision we found a work-around to be able to implement all the extra models we are presenting.
+New models with asymmetric traits are preferred and strengthen the results.
 

### 2019-07-11 manuscript/methods.tex
Editing more methodss

@@
 Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 651 species with ploidy and/or breeding system information on the tree.
-% FIXME: Rosana will update numbers around here TODO
+% FIXME: Rosana will update numbers around here TODO:STILL NOT CORRECT!!!
 Of these, 290 had information for both states.
@@
 
-To test our hypotheses about lineage diversification and trait macroevolution, we fit a large suite of phylogenetic models.
-All models are from the SSE framework \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016}.
-This style of model has parameters that describe per-lineage rates of speciation and extinction, specific to each character state (denoted $\lambda$ and $\mu$, respectively, with subscripts to indicate the state), along with rates of transitions between states (denoted $\rho$ for polyploidization, $\delta$ for diploidization, and $q_{IC}$ for loss of self-incompatibility).
+In order to test our hypotheses about lineage diversification and trait macroevolution, we fitted 29 state-dependent speciation and extinction models (SSE) based on previous work from \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016}.
+SSE models parameters that describe per-lineage rates of speciation and extinction, specific to each character state (denoted $\lambda$ and $\mu$, respectively, with subscripts to indicate the state), along with rates of transitions between states (denoted $\rho$ for polyploidization, $\delta$ for diploidization, and $q_{IC}$ for loss of self-incompatibility).
 
@@
 In order to investigate the association between ploidy level and diversification, we first defined a model in which taxa were classified as diploid $(D)$ or polyploid $(P)$ (M1).
-This is the same model \citet{mayrose_2011} used for this question across an array of plant genera.
+This is the same as one of the models presented in  \citet{mayrose_2011} that was used for the exact same question across an array of plant genera.
 % R- Itay wanted to hightlight that is the same model. Definitely a BiSSE. He did the ClaSSE later in the 2014 response.
@@
 Our second ploidy model includes a binary hidden trait that subdivides each observed state (M2).
-In this CID model \citep{beaulieu_2016}, the hidden trait can affect diversification but the observed trait does not.
+In this character independent model known as CID \citep{beaulieu_2016}, the hidden trait can affect diversification but the observed trait does not.
 Comparing M1 and M2 allows us to test whether diversification rate heterogeneity is better explained by ploidy or by some unknown factor.
@@
 We begin with each species classified as self-incompatible $(I)$ or self-compatible $(C)$ (M11).
-This is the same as the analysis of \citet{goldberg_2010} but with an updated phylogeny \citep{sarkinen_2013}.
+This is the same as the analysis of \citet{goldberg_2010} but with an updated phylogeny \citep{sarkinen_2013} and larger aggregated dataset.
 
@@
 
-We do not include additional models with diploidization because this reverse transition renders the models non-lumpable \citep{tarasov_2019}.  % E: further explanation needed?
+We do not include additional models with diploidization because this reverse transition renders the models non-lumpable \citep{tarasov_2019}.
+The lumpability property requires that the rates going from the lumped state to the singular state are equal.
+Under the diploidization models CP goes back to CD with rate $\delta$ but the rate from CP back to ID is 0, and $\delta \neq 0$.
+Therefore, two-state models with diplodization cannot be re-coded to be compared to the tree-state models.
+% E: further explanation needed? R- I added a simple math explanation. Let me know if it is okay
 
@@
 
-Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $BF(M_1,\ M_0) = \ln[ P(\mathbf{X} | M_1) - P(\mathbf{X} | M_0)]$.
-There is `strong evidence'  for $M_1$ when this value is more than 2, `very strong' support when it is more than 6, and `extreme' support when it is more than 10 \citep{kass1995}.
+Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $K=ln(BF(M_1,\ M_0)) = \ln[ P(\mathbf{X} | M_1) - P(\mathbf{X} | M_0)]$.
+There is `strong evidence'  for $M_1$ when this value is more than 10, moderate support if the value is more than 1 and no evidence if the value is between -1 and 1.
+If the vaue of  $K$ is negative the evidence goes towards $M_0$.
+\citep{kass1995}.
 

### 2019-07-11 manuscript/tables.tex
more tables

@@
 \begin{document}
+
+\begin{table}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|r|c|c|c|c|l|}
+\hline
+Model                & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M2    & M3     & M4    & M5     & Evidence                                                                                     \\ \hline
+M1.
+D/P              & -1283.76                                                          & 59.90 & 49.23  & 60.48 & 58.82  & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M1\end{tabular}            \\ \hline
+\textbf{M2.
+CID D/P} & \textbf{-1223.86}                                                 &       & -10.66 & 0.579 & -1.079 & \begin{tabular}[c]{@{}l@{}}Moderate to strong \\ preference for M2\end{tabular}              \\ \hline
+M3.
+D/P+A/B          & -1234.52                                                          &       &        & 11.24 & 9.58   & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}   \\ \hline
+M4.
+D/P+A/B asym     & -1223.28                                                          &       &        &       & -1.658 & \begin{tabular}[c]{@{}l@{}}Moderate evidence for\\ only asymmetric hidden rates\end{tabular} \\\hline
+M5.
+D/P+A/B all asym & -1224.93                                                          &       &        &       &        &          \\ \hline
+\end{tabular}
+\caption{Bayes factors of  ploidy only models without diplodization.
+Results indicate that a character independent model (M2) is strongly preferred over every other model.
+The following best model is M4 a HiSSE model with asymetric hidden rates.}
+\label{table:M1M5}
+\end{table}
+
+\begin{table}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|r|c|c|c|c|l|}
+\hline
+Model                      & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M12   & M13   & M14   & M15   & Evidence                                                                                       \\ \hline
+M11.
+I/C                   & -1309.07                                                          & 41.13 & 38.59 & 61.39 & 61.35 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M111\end{tabular}            \\ \hline
+M12.
+CID I/C               & -1267.93                                                          &       & -2.53 & 20.37 & 20.21 & \begin{tabular}[c]{@{}l@{}}Models with asymmetric rates\\ are preferred over  M12\end{tabular} \\ \hline
+M13.
+I/C+A/B               & -1270.47                                                          &       &       & 22.80 & 22.75 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}   \\ \hline
+\textbf{M14.
+I/C+A/B asym} & \textbf{-1247.66}                                                 &       &       &       & -0.05 & No evidence                                                                                    \\ \hline
+M15.
+I/C+A/B all asym      & -1247.72                                                          &       &       &       &       &       \\ \hline
+\end{tabular}
+\caption{Bayes factors of breeding system only models.
+Results indicate that the HiSSE model with asymmetric hidden rates (M14) is strongly preferred over every other model very closely followed by M15 the model that adds asymmetry to the transition rates within hidden states.}
+\label{table:M11M15}
+\end{table}
 

### 2019-07-11 rev_analyses/mussenodip250K/mussedpsinodipv1.sh
adding mussenodip code

@@ -1,0 +1,12 @@
+#!/bin/bash -l
+#PBS -l pmem=500mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o mussenodip.log
+#PBS -e mussenodip.log
+#PBS -l walltime=464:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/mussenodip/MuSSEDPSInodip.Rev
+

### 2019-07-11 manuscript/discussion.tex
discussion notes-to-self
substantial edits in Discussion
diversification section
pathways section

haven't tackled diploidization section yet

@@
 %
-Although species are composed of vast assemblages of variable traits, many traits are heritable, and they may affect the propensity of species to perish and multiply \citep{lewontin_1970}, examining such macroevolutionary effects of many traits is rare, and remains challenging.
-We investigated the influence of two traits, ploidy and breeding system, in diversification of Solanaceae.
-Most significantly, we found that the effects of ploidy, inferred as significant when it is the only trait under consideration, are overshadowed by hidden traits, or breeding system when added to models.
-Interestingly, the difference in diversification rates between the diploids and polyploids is negligible on the background of self-compatibility.
-Furthermore, we recovered a smaller difference in diversification rates between self-incompatible and self-compatible lineages against the diploid genomic background. 
-Motivated by the widespread recent findings of polyploidy in the ancestry of angiosperms, we also examined and found modest support for diploidization in the history of the family, but without significant effects on inferences of diversification dynamics.
-%B: how about "modest"?
-%R- I think modest is a good way to express it because it is the best we can do with chromosome numbers.
-%B: changed "polyploidy events" (polyploidy is a noun, only.
-polyploid either noun or adjective).
-Disentangling the patterns of diversification linked to multiple traits is difficult, but our strategy allows tests of intricate hypotheses regarding a heterogenous diversification process.
-Below, we discuss key findings and highlight the most interesting and paradoxical implications.
-We provide an outline for broader application of phylogenetic comparative methods in the context of interactions among traits and diversification process.
+Species are composed of vast assemblages of variable traits, many traits are heritable, and many traits could affect the propensity of species to perish or multiply \citep{lewontin_1970}.
+Examining the effects of complex trait combinations on lineage diversification, however, remains challenging.
+%
+Focusing first on ploidy and then on breeding system, we found that considering each trait in isolation provides an incomplete story.
+Considering them together, and in conjunction with another hidden factor, provides a fuller picture of macroevolutionary dynamics within Solanaceae.
+We hope our work serves as an example of how phylogenetic comparative methods can be used to disentangle the contributions of interacting traits to heterogeneous lineage diversification.
 
-\subsection*{Disentangling trait interactions in the context of diversification}
+\subsection{Interacting traits and lineage diversification}
 
-Previous analyses of the effects of ploidy on diversification showed that diploids had greater net diversification rates than polyploids across multiple clades of the angiosperm phylogeny \citep{mayrose_2011, mayrose_2015}.
-We obtain a broadly similar outcome when considering only the effects of ploidy, but the results change significantly with joint consideration of hidden states and breeding systems (\cref{figure:netdivall}A and B).
-The addition of a hidden trait blurs differences in diversification rates between diploids and polyploids.
-The model with breeding system also included better explains the variation of the inferred net diversification rates among ploidy states. 
-SI, which is itself associated with a faster diversification rate, is the breeding system context in which many diploids occur.
-Among SC species, diploids and polyploids are associated with similar diversification rates (\cref{figure:netdivall}E).
-These results indicate that the hidden trait in the analysis of one focal trait can in part be explained by the other focal trait.
-The finding thus demonstrates how the relative influence of two traits on diversification can be disentangled in a comparative phylogenetic framework.
+% E: should this results summary be much shorter?
+Previous analyses of the effects of ploidy on diversification showed that diploids had greater net diversification rates than polyploids across many angiosperm clades \citep{mayrose_2011, mayrose_2015}.
+We obtain a similar outcome when examining ploidy alone in Solanaceae (\cref{figure:netdivall}A), but a consistent effect of ploidy on diversification is not clear when we also allow that a hidden factor could affect diversification as well (\cref{figure:netdivall}B).
+%
+Previous analyses of breeding system in this family showed that SI species are associated with higher diversification rates than their SC counterparts (\citealt{goldberg_2010}; \cref{figure:netdivall}C).
+Our new analyses with a hidden trait show that this pattern persists on different backgrounds.
+One on background, however, SC species experience positive net diversification (\ie not a dead end; cf.\ \citealt{igic_2013}) that exceeds the diversification rate of SI species on the other background (\cref{figure:netdivall}D).
+%
+Furthermore, our models of the joint macroevolution of ploidy and breeding system provided better statistical fits and revealed how the trait combinations associate with diversification.
+We find that the higher overall high net diversification rate of diploids is dominated by SI diploids, while SC diploids have a lower diversification rate that is comparable to polyploids, which are also SC (\cref{figure:netdivall}E).
+Thus, breeding system seems to provide the dominant explanation of diversification differences, with ploidy providing a secondary effect within SC species.
 
-Earlier work has shown that SI species in Solanaceae are associated with higher diversification rates compared to their SC counterparts \citep{goldberg_2010}. 
-When we only considered the role of breeding system in the diversification process, we were able to replicate the slower net diversification for SC  (\cref{figure:netdivall}C).
-%B: we only considered <- we inferred only
-Incorporating hidden states again resulted in SI associated with greater net diversification rates than SC in both hidden states  (\cref{figure:netdivall}D). 
-However, the overall picture is a bit more complex.
-Depending on the background hidden state, it is possible that a class of SC species may experience higher diversification rates than SI species.
-Moreover, these SC lineages may not constitute a dead end (with negative diversification rates).
-
+% These results indicate that the hidden trait in the analysis of one focal trait can in part be explained by the other focal trait.
 % more generally, not okay to approximate musse with bisse when states are correlated (cf Pyron)
 
+Throughout our numerous model comparisons, models that included a hidden trait emerged as significantly better than those that did not. % supptable ref
+This is entirely consistent with the expectation that many processes can affect speciation and extinction, far beyond the one or two traits of particular interest to us.
+A natural follow-up question is, what is this hidden trait?
+Our results show that to some extent, breeding system functions as a hidden trait in the ploidy-centered analysis, and vice versa.
+But there is also strong statistical support for a hidden trait beyond ploidy and breeding system.
+One temptation is to propose specific possible traits to add into the analysis.
+For example, the rapidly-diversifying clade of mostly SC species within \emph{Solanum} suggests that geography may play a role. % Boris, please check!  Australian?  fixme
 % alpha is always low
 %   new supp fig showing how A & B are inferred on the tree?  at least for ID/CD/CP+A/B model
-%     maybe we can hypothesize factors that underlie the hidden state (geography?)
+Alternatively, we could view the hidden trait as merely a statistical trick, providing an easy way to model extra heterogeneity without directly representing a real trait.
+
+Although we fit an extensive set of models in order to relax a variety of their assumptions, we did not explore the process of trait change in conjunction with speciation.
+That is, our models all assume anagenetic but not cladogenetic trait evolution.
+Phylogenetic methods do exist to separate anagenetic and cladogenetic changes \citep{mayrose_2011, goldberg_2012, magnuson-ford_2012}.
+These have been applied to show that losses of SI are often associated with cladogenesis in Solanaceae \citep{goldberg_2012}, and that polyploidization co-occurs with speciation events in other families \citep{zhan_2016, freyman_2017}.
+Exploring cladogenetic trait change was outside the scope of our modeling complexity.
+Although \citet{goldberg_2012} found that allowing cladogenetic changes did not substantially affect inference of net diversification rates associated with breeding system, future work could test whether this process affects diversification rate estimates with the more complex state spaces of our other models.
+
+\subsection{Pathways to polyploidy}
+
+With evolution predominantly in the direction from diploid to polyploid, and from SI to SC, surviving lineages will tend eventually to become SC polyploids.
+We found that in Solanaceae, the pathway to this state is more likely to consist of a single step ($ID \rightarrow CP$) than two steps ($ID \rightarrow CD \rightarrow CP$).
+We also showed that although this question focuses on the process of state transitions, its answer is affected by the process of lineage diversification.
+We used a simple mathematical approach to investigate the contributions of the two pathways, but future work could use stochastic character mapping to more directly estimate the numbers of each type of transition within the clade.
+
+Macroevolutionary transition rates represent a combination of time spent waiting for individuals to arise with a new trait value, and for that new trait value to become widespread within the species.
+For our traits, this consists of mutations that break SI or generate polyploid individuals, and selective pressures that cause loss or fixation of these mutants.
+Estimates of mutation rates are highly uncertain, but the chance of breakdown of SI within diploids is perhaps $10^{-5}$ per pollen grain; this includes breakdown by tetraploidization \citep{lewis1979}, which is itself estimated to occur at about the same rate \citep{ramsey_1998}.
+% unit check?
+% Seemingly, then, the simple genic mutation rate that leads to loss of SI is at best equal to tetraploidization mutation rate, and possibly far lower.
+In contrast, we estimate a macroevolutionary transition rate from $ID$ to $CD$ that is three-fold greater than the rate from $ID$ to $CP$, indicating that selection restricts the fixation of new polyploids more than of new SC mutants \citep{robertson_2011}.
+% todo: check if new rates are still 3x
+
+Our findings prompt several further questions about the macroevolutionary pathways of ploidy and breeding system.
+%
+First, our support for the direct pathway is consistent with the idea that breakdown of SI by whole genome duplication---via diploid `heteroalleleic' pollen---may often trigger the evolution of gender dimorphism as a different mechanism of inbreeding avoidance \citep{miller_2000}.
+A further test of this hypothesis would additionally examine the propensity of polyploids generated through either pathway to become dioecious \citep{robertson_2011}.
+%
+Second, we might wonder whether the propensity for a polyploid species to diversify depends on whether it arose via the one-step or two-step pathway.
+This could be tested with a different form of a hidden state model, in which the polyploid state is subdivided based on the pathway taken into that state.
+This would be similar to what we did for our pathways analysis, but it would also include the possibility of different diversification rates in those $CP$ substates.
+%
+Third, the generality of our findings in other families remains to be assessed.
+An identical procedure could be used in other families with gametophytic SI.
+In clades with sporophytic self-incompatibility systems, however, SI is not disabled by whole genome duplication, so there is no one-step pathway. % ref?
+The correlation between breeding system and ploidy may therefore be different, and it is currently unknown whether one of the two-step pathways might dominate.
 
@@
 
-\subsection*{Pathways to polyploidy}
+\subsection*{Diploidization}
 
-We find that SI diploid lineages are much more likely to take a one-step  ID$\rightarrow$CP pathway to polyploidy rather than a two-step ID$\rightarrow$CD$\rightarrow$CP pathway, in Solanaceae.
-This result is significant because it has a number of implications for the inference of history of breeding and sexual systems in flowering plants.
-First, support for the direct pathway lends credence to the idea that breakdown of SI by WGD---by generation of diploid `heteroalleleic' pollen---may often trigger the evolution of gender dimorphism across angiosperms, as a different mechanism of inbreeding avoidance \citep{miller_2000}.
-A more direct and definitive test of this hypothesis would additionally examine the propensity of polyploids generated through either pathway to become dioecious \citep{robertson_2011}.
-Second, statistical phylogenetic approaches measure lineage transition rates.
-Transitions depend on a combination of availability of individual mutation rates, which generate trait variation, and selection, which may cause disproportionate loss or fixation of these mutants.
-Estimates of mutation rates are highly uncertain, but the total rate of breakdown of SI within diploids ranges is estimated to be on the order of  $\mu_{IC} = 10^{-5}$ per pollen grain, and includes breakdown by tetraploidization \citep{lewis1979}, which is itself estimated to occur on the order of $\mu_{DP} = 10^{-5}$ \citep{ramsey_1998}.
-Seemingly, then, the simple genic mutation rate that leads to loss of SI is at best equal to tetraploidization mutation rate, and possibly far lower.
-We nevertheless recover transition rates from ID to CD ($q_{IC}$) threefold greater than $\rho_{I}$, indicating that selection restricts the fixation rate of new polyploids compared with new SC mutants \citep{robertson_2011}.
-Overall, our findings suggest that the effect of ploidy on lineage diversification rate is largely explained by its correlation with breeding system, and that other unknown factors work with breeding system to determine diversification rates.
-
-% qualifying paragraph for pathways
-We modeled trait changes in both of these pathways as anagenetic, transitions occurring within lineages, without associated speciation events.
-However, losses of SI (I$\rightarrow$C transitions) are often associated with cladogenesis in Solanaceae \citep{goldberg_2012}, and polyploidization (D$\rightarrow$P transition) also seem to co-occur with speciation events in a number of other families \citep{zhan_2016, freyman_2017}.
-Methods do exist to separate anagenetic and cladogenetic changes \citep{mayrose_2011, goldberg_2012, magnuson-ford_2012}, though \citet{goldberg_2012} found that extending the *SSE framework to allow cladogenetic changes did not substantially affect inference of net diversification rates.
-Nevertheless, a distinct possibility remains that allowing for cladogenetic transitions could tip the scales further in favor of the one-step pathway to SC polyploidy (CP), but this remains to be tested in future work.
-%B Will's comment alluded to, but not directly.
-Specifically, the multi/hidden state SSE-based framework we employed here could be extended to incorporate cladogenetic transitions. %B: We could also list %W+%R idea above, for stochastic mapping?
-Another useful line of inquiry would examine the path-dependent diversification rates, asking whether diversification rates of polyploid SC lineages differ depending on whether they were generated via a one- or two-step pathway.
-%B: could elaborate if you see necessary.
-TODO
-%
-% examine SC-only families
-% no causal correlation in SSI systems
-% paths to dioecy
-% path-dependent net-diversification
-%
-%B: also missing deeper discussion of with/without SDD/SID estimates and comparison with Robertson et al. 2011.
-%
-%B: And, are diversification rates of polyploid SC lineages different if one or other path was taken? % E: This is a great question.  We don't answer it with this round of analyses.  But discuss?
-% E: To elaborate: In the IC/CD/CP model, the CP state would be subdivided into two hidden values, corresponding to the influx from the two pathways.
-Those two states, CP_X and CP_Y, could have different lambda and mu.
-%
-%W: 3) I'll bet allowing cladogenetic transitions would make the one-step ID/CP even more dominant.
-%    I added the paragraph below regarding cladogenetic transitions, but it doesn't really fit with
-%    the rest of the discussion, so modify or discard as needed:
-% E: A small defense of excluding clado is that it didn't affect the diversification results in goldberg2010 vs goldberg2012
-
-\subsection*{Diploidization}
+% Motivated by the widespread recent findings of polyploidy in the ancestry of angiosperms, we also examined and found modest support for diploidization in the history of the family, but without significant effects on inferences of diversification dynamics.
+%B: how about "modest"?
+%R- I think modest is a good way to express it because it is the best we can do with chromosome numbers.
+%B: changed "polyploidy events" (polyploidy is a noun, only.
+polyploid either noun or adjective).
 
@@
 Inferences that question whether other complex trait interactions are hidden in this background heterogeneity are necessary because they critically contribute to the reconstruction and understanding of diversification.
+
+%%%%%%%%%%
+% E notes:
+%
+% summarize div findings
+%     breeding system takes from ploidy
+%     but not simple SI/SC pattern
+% what is the hidden state?
+%     could be a real trait, see tip ASR, geography?
+%     but don't be circular, test in different clade
+%     could be a stats trick
+%
+% pathways
+%     likely way to lose SI
+%     likely way to become P
+%     (compare those relative rates with robertson2011)
+%     different consequences for eventual D?
+%     div effects on pathway commonness
+%     path-dep div in CP?
+%     clado
+%
+% diploidization
+%     some model support
+%     but still SC
+% paleopolyploidy
+%     if before root, no SI?

### 2019-07-12 manuscript/methods.tex
Adding the correct figures and tables to results and first part of supplement

@@
 
-In order to test our hypotheses about lineage diversification and trait macroevolution, we fitted 29 state-dependent speciation and extinction models (SSE) based on previous work from \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016}.
+In order to test our hypotheses about lineage diversification and trait macroevolution, we fitted 29 state-dependent speciation and extinction models (SSE) to previous work on diversification \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016}.
 SSE models parameters that describe per-lineage rates of speciation and extinction, specific to each character state (denoted $\lambda$ and $\mu$, respectively, with subscripts to indicate the state), along with rates of transitions between states (denoted $\rho$ for polyploidization, $\delta$ for diploidization, and $q_{IC}$ for loss of self-incompatibility).
@@
 These models differ in whether transitions between the hidden states are symmetric or asymmetric, and whether the polyploidization rate depends on the hidden state.
-Comparing M1--M5 allows us to test whether ploidy is associated with diversification differences on top of differences explained by an unknown factor.
+Comparing M1--M5 allows us to test whether ploidy is associated with diversification differences on top of the differences potentially explained by an unknown factor\cref{table:bayesfactors}.
 
-We further fit the analogues of these five models but including a rate parameter for transitions from polyploid to diploid (M6--M10).
+We further fit the analogues of these five models but including a rate parameter for transitions from polyploid to diploid (M6--M10) \cref{supptable:M6M10}.
 This allows us to assess whether our conclusions about ploidy and diversification are robust to the possibility of diploidization.
@@
 
-We then add models to allow diversification to be influenced by only a hidden trait (M12), or by both breeding system and a hidden trait (M13--M15, with varying degrees of complexity in the hidden trait transitions).
+We then add models to allow diversification to be influenced by only a hidden trait (M12), or by both breeding system and a hidden trait (M13--M15, with varying degrees of complexity in the hidden trait transitions \cref{table:M11M15}).
 Similar models were used by \citet{freyman_2017} used to study diversification in Onagraceae.
@@
 
-As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by both ploidy and breeding system as well as a hidden trait (M18--M20, with varying degrees of complexity in the hidden trait transitions).
+As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by both ploidy and breeding system as well as a hidden trait (M18--M20, with varying degrees of complexity in the hidden trait transitions \cref{supptable:M16M20})
 We add this hidden trait layer on top of our three-state model \citep[similar to][]{caetano_2018, huang_2018}.
-We also fit the analogous models but allowing for diploidization (M21--M25).
+We also fit the analogous models but allowing for diploidization (M21--M25, \cref{supptable:M21M25}).
 
@@
 We further add a hidden trait to each of these models (M27 and M29).
+We compared those models in table \cref{table:lumped}
 
@@
 
-Parameters for each of the ten models were estimated using RevBayes \citep{hoehna_2016}.
+Parameters for each of the 29 models were estimated using RevBayes \citep{hoehna_2016}.
 Code for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.
@@
 
-We calculated the marginal likelihood for each of the models using fifty stepping stone steps under the methodology of \citet{xie_2010}, implemented in RevBayes \citep{hoehna_2016}.
+We calculated the marginal log-likelihood for each of the models using fifty stepping stone steps under the methodology of \citet{xie_2010}, implemented in RevBayes \citep{hoehna_2016}.
 Each stepping stone step was found by calculating at least 500 generations of burn-in followed by a total of 1,000 MCMC steps (\cref{table:marginallike}).
@@
    % The ten models and their marginal likelihoods.
-    %Values in bold are for the best models within each class that are comparable (see \cref{table:bayesfactors}).
+    %Values in bold are for the best models within each class that are comparable (see \cref{table:}).
     %Abbreviations are D: diploid, P: polyploid, I: self-incompatible, C: self-compatible, A: one state of hidden trait, B: other state of hidden trait, $\delta$: diploidization.}
@@
 \addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|c|c|c|}
-\bottomrule
+\begin{tabular}{|l|r|c|c|c|c|l|}
+\hline
+Model                & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M2    & M3     & M4    & M5     & Evidence                                                                                     \\ \hline
+M1.
+D/P              & -1283.76                                                          & 59.90 & 49.23  & 60.48 & 58.82  & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M1\end{tabular}            \\ \hline
+\textbf{M2.
+CID D/P} & \textbf{-1223.86}                                                 &       & -10.66 & 0.579 & -1.079 & \begin{tabular}[c]{@{}l@{}}Strong preference for \\ M2 over M1\end{tabular}              \\ \hline
+M3.
+D/P+A/B          & -1234.52                                                          &       &        & 11.24 & 9.58   & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}   \\ \hline
+\textbf{M4.
+D/P+A/B asym}     & \textbf{-1223.28}                                                          &       &        &       & -1.658 & \begin{tabular}[c]{@{}l@{}}Moderate evidence for\\ only asymmetric hidden rates\end{tabular} \\\hline
+M5.
+D/P+A/B all asym & -1224.93                                                          &       &        &       &        &          \\ \hline
 \end{tabular}
-% E: We will probably have to change the layout here because it's not technically a table.  But good enough for now.
-\caption{
-    Bayes factors for model comparisons.
-    Each of the three boxes contains models that can be compared with one another, based on the character states they include (see \cref{figure:stateclassifications}B).
-    Models are numbered in as \cref{table:marginallike}. % FIXME: they aren't, but it's a pain to change; change in prev table instead?
-    Bayes factors are reported on the natural log scale, so numbers greater than $+2$ mean that the model in the row label has `positive' support relative to the model in the column label; numbers less than $-2$ mean that model in the column label is the preferred one.
-    Conventional thresholds for `strong' and `very strong' support are 6 and 10, respectively.
-    The best model in each set is written in bold.
-    In each case, it is the most complex model of the set.
+\caption{Bayes factors of  ploidy only models without diplodization.
+Numbers greater that 1 mean moderate support relative to the model in the row label.
+Conventional threshold for `strong' support is 10.
+The best models are written in bold.
 }
-\label{table:bayesfactors}
+\label{table: bayesfactors}
+\end{table}
+
+\begin{table}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|r|c|c|c|c|l|}
+\hline
+Model                      & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M12   & M13   & M14   & M15   & Evidence                                                                                       \\ \hline
+M11.
+I/C                   & -1309.07                                                          & 41.13 & 38.59 & 61.34 & 61.40 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M111\end{tabular}            \\ \hline
+M12.
+CID I/C               & -1267.93                                                          &       & -2.53 & 20.21 & 20.37 & \begin{tabular}[c]{@{}l@{}}Models with asymmetric rates\\ are preferred over  M12\end{tabular} \\ \hline
+M13.
+I/C+A/B               & -1270.47                                                          &       &       & 22.75 & 22.80 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}   \\ \hline
+\textbf{M14.
+I/C+A/B asym} & \textbf{-1247.72}                                                 &       &       &       & 0.05 & No evidence                                                                                    \\ \hline
+\textbf{M15.
+I/C+A/B all asym }     & \textbf{ -1247.66}                                                          &       &       &       &       &       \\ \hline
+\end{tabular}
+\caption{Bayes factors of breeding system only models. Results indicate that the HiSSE models with asymmetric rates (M14, M15) are strongly preferred over every other model.}
+\label{table:M11M15}
 \end{table}
@@
 % E: Could we include a figure of the traits on the tree?  At least as supp info. FIXME
+\begin{table}[]
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+Model                       & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=ln(BF(M1,M2)) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
+M26.
+Lumped D/P             & -1463.22                                                          & M26 vs.
+M16 & 4.10            & M16 (Moderate)                                                       \\
+\textbf{M16.
+ID/CD/CP}      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
+M27.
+Lumped D/P+A/B         & -1417.67                                                          & M27 vs.
+M23 & 3.678           & M23 (Moderate)                                                       \\
+\textbf{M23.
+ID/CD/CP +A/B} & \textbf{-1414.00}                                                 &             &                 &                                                                      \\ \hline
+M28.
+Lumped I/C             & -1458.41                                                          & M28 vs.
+M16 & -0.69           & No evidence                                                          \\
+\textbf{M16.
+ID/CD/CP}      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
+M29.
+Lumped I/C+ A/B        & -1416.60                                                          & M29 vs.
+M23 & 2.60            & M23 (Moderate)                                                       \\
+\textbf{M23.
+IC/CD/CP+A/B}  & \textbf{-1414.00}                                                 &             &                 &                                                                      \\ \hline
+\end{tabular}
+\caption{Test for adding a focal trait to a binary state model via Bayes factors.
+Three-state models are moderately preferred over two-state models.
+The evidence is stronger towards the inclusion of breeding systems in ploidy models (M26, M27) than the inclusion of ploidy in breeding system models (M29, M29)}
+\label{table:lumped}
+\end{table}

### 2019-07-12 manuscript/results.tex
fixed up pathways figure
added a diagram of the two pathways
nicer fonts and spacing in the plots
discussion notes-to-self
- fixed %Boris mark in introduction - added modified pathstates fig to separate and match line type
Adding the correct figures and tables to results and first part of supplement

@@
 
+% see Will's notes in discussion.tex, FIXME
+%   also use transition rates from transition-only (no diversification) analysis
+%   also use stochatic maps, (a) for CP tips, what paths were taken? (b) for steps away from ID, which path was started?
+
 The preceding conclusions, however, ignore the changes in numbers of lineages in each state due to speciation and extinction.
@@
 \begin{supptable}
-    \begin{center}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|r|c|c|c|c|l|}
+\hline
+Model                           & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M7    & M8     & M9    & M10   & Evidence                                                                                     \\ \hline
+M6.
+D/P+$\delta$                & -1268.83                                                          & 55.41 & 42.78  & 53.37 & 52.85 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M6\end{tabular}            \\ \hline
+\textbf{M7.
+CID D/P+$\delta$}   & \textbf{-1212.42}                                                 &       & -12.62 & -2.04 & -2.04 & \begin{tabular}[c]{@{}l@{}}Model M7 moderately\\ preferred over M9 and M10\end{tabular}      \\ \hline
+M8.
+D/P+A/B $\delta$            & -1214.46                                                          &       &        & 10.58 & 10.07 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular} \\ \hline
+M9.
+D/P+A/B +$\delta$ asym      & -1214.46                                                          &       &        &       & -0.51 & No evidence                                                                                  \\ \hline
+M10.
+D/P+A/B +$\delta$ all asym & -1214.97                                                          &       &        &       &       &         \\ \hline
+\end{tabular}
+\caption{Bayes factors of  ploidy only models with diplodization.
+Results indicate that a character independent model (M6) is strongly preferred over BiSSE (M1)  model.
+M6 is moderately preferred over any of the HiSSE model swith asymmetric hidden rates (M9, M10).}
+\label{supptable:M6M10}
+\end{supptable}
 
-    \begin{tabular}{lcccccc}
-                       & D/P   & D/P+A/B      & I/C    & I/C+A/B      & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
-        %                                                                                      
-        $r_D$          & 0.260 & 0.193, 0.658 & ---    & ---          & ---      & ---          \\
-        $r_P$          &-0.056 & 0.030, 0.187 & ---    & ---          & ---      & ---          \\
-        $r_I$          & ---   & ---          & 0.550  & 0.386, 0.877 & ---      & ---          \\
-        $r_C$          & ---   & ---          &-0.001  &-0.059, 0.606 & ---      & ---          \\
-        $r_{ID}$       & ---   & ---          & ---    & ---          & 0.455    & 0.309, 0.797 \\
-        $r_{CD}$       & ---   & ---          & ---    & ---          & 0.065    &-0.006, 0.587 \\
-        $r_{CP}$       & ---   & ---          & ---    & ---          &-0.088    &-0.074, 0.403 \\
-        %                                                                                      
-        $\rho$         & 0.047 & 0.047        & ---    & ---          & ---      & ---          \\
-        $\rho_I$       & ---   & ---          & ---    & ---          & 0.067    & 0.053        \\
-        $\rho_C$       & ---   & ---          & ---    & ---          & 0.033    & 0.032        \\
-        $q_{IC}$       & ---   & ---          & 0.364  & 0.261        & 0.198    & 0.145        
+\begin{supptable}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|r|c|c|c|c|l|}
+\hline
+Model                            & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M18   & M19   & M20   & Evidence                                                                                                   \\ \hline
+M16.
+ID/CP/CD                    & -1459.11                                                          & 45.11 & 65.91 & 63.98 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M16\end{tabular}                         \\ \hline
+M17.
+CID ID/CP/CD                & *                                                                 &       &       &       &                                                                                                            \\
+M18.
+ID/CP/CD+A/B                & -1414                                                             &       & 20.79 & 18.87 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}               \\ \hline
+\textbf{M19.
+ID/CP/CD +A/B asym} & \textbf{-1393.20}                                                 &       &       & -1.92 & \begin{tabular}[c]{@{}l@{}}Asymmetric hidden rates moderately\\ preferred over all asymmetric\end{tabular} \\ \hline
+M20.
+ID/CP/CD +A/B all asym      & -1393.12                                                          &       &       &       &   \\ \hline
 \end{tabular}
+\caption{Bayes factors of ploidy and breeding system withou diploidization.
+Results indicate that the MuHiSSE model with asymmetric hidden rates (17) is  strongly preferred over M16 and M18 and moderately preferred over the MuHiSSe with all rates asymetric (M20).
+*Marginal log-likelihood for M17 could not be calculated in our limited amount of computational time.}
+\label{supptable:M16M20}
+\end{supptable}
 
-    \vspace{12pt}
+\begin{supptable}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|r|c|c|c|c|l|}
+\hline
+Model                                 & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M22   & M23    & M24   & M25   & Evidence                                                                                         \\ \hline
+M21.
+ID/CP/CD+$\delta$                & -1454.68                                                          & 55.48 & 46.031 & 67.94 & 65.15 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M21\end{tabular}               \\ \hline
+M22.
+CID ID/CP/CD+$\delta$            & -1399.201                                                         &       & -9.452 & 12.45 & 9.675 & \begin{tabular}[c]{@{}l@{}}Model M24 strongly preferred \\ over M22\end{tabular}                 \\ \hline
+M23.
+ID/CP/CD+A/B $\delta$            & -1408.65                                                          &       &        & 21.91 & 19.12 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}     \\ \hline
+\textbf{M24.
+ID/CP/CD +$\delta$ asym} & \textbf{-1386.74}                                                 &       &        &       & -2.78 & \begin{tabular}[c]{@{}l@{}}Asymmetric hidden rates \\ preferred over all asymmetric\end{tabular} \\ \hline
+M25.
+ID/CP/CD +$\delta$ all asym      & -1389.52                                                          &       &        &       &       &                      \\ \hline
+\end{tabular}
+\caption{Bayes factors of ploidy and breeding system with diploidization.
+Results indicate that the MuHiSSE model with asymmetric hidden rates (M24) is  strongly preferred over M21-M23 and moderately preferred over the MuHiSSe with all rates asymetric (M25).}
+\label{supptable:M21M25}
+\end{supptable}
 
-    \begin{tabular}{lcccc}
-                       & D/P   & D/P+A/B      & ID/CD/CP & ID/CD/CP+A/B \\ \midrule
-        %                                                              
-        $r_D$          & 0.382 & 0.698, 0.100 & ---      & ---          \\
-        $r_P$          & 0.109 & 0.587, 0.182 & ---      & ---          \\
-        $r_I$          & ---   & ---          & ---      & ---          \\
-        $r_C$          & ---   & ---          & ---      & ---          \\
-        $r_{ID}$       & ---   & ---          & 0.449    & 0.318, 0.789 \\
-        $r_{CD}$       & ---   & ---          & 0.050    &-0.248, 0.494 \\
-        $r_{CP}$       & ---   & ---          &-0.027    & 0.110, 0.634 \\
-        %                                                              
-        $\rho$         & 0.033 & 0.026        & ---      & ---          \\
-        $\rho_I$       & ---   & ---          & 0.063    & 0.047        \\
-        $\rho_C$       & ---   & ---          & 0.024    & 0.011        \\
-        $\delta$       & 0.050 & 0.162        & 0.022    & 0.107        \\
-        $q_{IC}$       & ---   & ---          & 0.194    & 0.164        
+\begin{supptable}
+\addtolength{\tabcolsep}{-5pt}
+\begin{tabular}{|l|c|c|c|l|}
+\hline
+Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
+M3. D/P+A/B                                 & -1234.52                     &    M3 vs. M4                 & 11.239                  & M4 (Strong)   \\
+\textbf{M4. D/P+ A/B asym}            & -1223.28                     &   M4 vs. M5              & -1.658                  & M4 (Moderate)   \\
+M5. D/P+A/B all asym                 & -1224.93                     &                     &                         &                         \\ \hline
+M13. I/C+ A/B                                & -1270.47                     &   M13 vs. M14                  & 22.75                   & M14 (Strong)   \\
+\textbf{M14. I/C+ A/B asym}            & \textbf{-1247.72}            & M14 vs. M15         & {0.05}           & No evidence             \\
+M15. I/C+ A/B all asym               & -1247.67                     &                     &                         &                         \\ \hline
+M18. IC/CD/CP+A/B                            & -1414.00                     &     M18 vs. M19                & 20.79                   & M19 (Strong)    \\
+\textbf{M19. IC/CD/CP+ A/B asym}       & \textbf{-1393.21}            & M19 vs. M20          & {-1.919}         & M19 (Moderate)   \\
+M20. IC/CD/CP+ A/B all asym           & -1395.129                    &                     &                         &                         \\ \hline
+M8. D/P +A/B+$\delta$                          & -1225.05                     &  M8 vs. M9                   & 10.58                   & M9 (Strong)    \\
+\textbf{M9. D/P+ A/B+ $\delta$ asym}      & \textbf{-1214.46}                     &    M9 vs. M10                 & -0.52                   & M10 (Moderate) \\
+M10. D/P+A/B+$\delta$ all asym        & -1214.98                     &                    &                         &                         \\ \hline
+M23. IC/CD/DP+A/B+$\delta$                 & -1408.65                     &     M23 vs M24                & 21.91                   & M24(Strong)     \\
+\textbf{M24. IC/CD/DP+A/B+$\delta$ asym} & \textbf{-1386.74}                     &   M24 vs M25              & -2.78                   & M24 (Moderate) \\
+M25. IC/CD/DP+A/B+ $\delta$ all asym     & -1389.52                     &                     &                         &                         \\ \hline
 \end{tabular}
-
-    \end{center}
-    \caption{
-        Median rate estimates for all fitted models.
-        Units are per million years.
-        Two comma-separated numbers refer to the $A$ and $B$ hidden states, and --- means the parameter was not present in the model.
-    Net diversification rates ($r$) are subscripted with trait state initials (Diploid, Polyploid, Incompatible, Compatible).
-Transition rates are $\rho$ (polyploidization), subscripted with background breeding system state; $\delta$ (diploidization); and  $q_{IC}$ (loss of self-incompatibility).
-    The upper section is for models without diploidization, and the lower section is for models with diploidization.
-    The supplemental figures show the corresponding distributions of parameter estimates.
-    }
-    \label{table:estimates}
+\caption{Test for symmetry of the hidden and trait rates via Bayes factors.
+For all models asymmetric hidden transition rates are preferred over models with equal hidden rates.
+Adding more complexity by assuming asymmetry in the trait rate within hidden states does not improve models.}
+\label{supptable:asymmetry}
 \end{supptable}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{allnetdivmodelswithoutdelta.pdf} % FIXME: need horizontal axis label
+\includegraphics[width=\textwidth]{alldiversificationasymhidden.pdf} 
 \caption{Net diversification rates for all models without diploidization rate estimate. 
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{allnetdivmodelsdiploidization.pdf}
-\caption{Net diversification rates for all models that include diploidization ($\delta$). 
-Each panel contains a graphical summary of estimated model parameters and displays posterior distributions for net diversification estimates.}  
-\label{figure:netdivnodip}
+\includegraphics[width=\textwidth]{lumped.pdf} %lumped.pdf
+\caption{Comparing two and three state models is possible via lumpable models (M26, M28).
+These models used the same classifications as the three-state model M16 which allow for comparisons via Bayes Factors as shown in \cref{table:lumped}.}
+\label{figure:lumped}
 \end{figure}
 
-% E todo: include a diagram of the transitions (panels a/b at the top?); re-run with new rate estimates; add panel letters
+% E todo: re-run with new rate estimates; add panel letters?
 \begin{figure}
-    \centering \includegraphics[width=0.9\textwidth]{pathways}
+    \centering 
+    \includegraphics[width=0.5\textwidth]{pathstates2} \\ [40pt] 
+        \includegraphics[width=0.9\textwidth]{pathways}
     \caption{
@@
 \begin{suppfigure}
+\includegraphics[width=\textwidth]{alldiversificationasymhiddendip.pdf}
+\caption{Posterior distribution for all the best models with diploidization} % XXX
+\label{suppfigure:alldip}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{comparablesDP.pdf}
+\caption{Lumped ploidy models to assess the effect of adding breeding system to create a three-state model.
+Moderate evidence exists that adding a breeding system is necessary \cref{table:lumped} } % XXX
+\label{suppfigure:lumpedDP}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{comparablesIC.pdf}
+\caption{Lumped breeding system models to assess the effect of adding ploidy to create a three-state model.
+Moderate to insignificant evidence exists that adding ploidy is necessary \cref{table:lumped} } % XXX
+\label{suppfigure:lumpedIC}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{effectofasymmetry.pdf}
+\caption{Effect of asymmetric rates in hidden models. Models with asymmetric rates are preferred over models with equal rates \cref{supptable:asymmetry}} % XXX
+\label{suppfigure:asymmetric}
+\end{suppfigure}
+
+\begin{suppfigure}
 \includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
@@
 
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDP.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node of the D/P ploidy model} % XXX
-\label{suppfigure:DPnodipasr}
-\end{suppfigure}
+%\begin{suppfigure}
+%\includegraphics[width=\textwidth]{asrDP.pdf}
+%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node of the D/P ploidy model} % XXX
+%\label{suppfigure:DPnodipasr}
+%\end{suppfigure}
 
@@
 
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDPdelta.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the D/P+ $\delta$, ploidy model} % XXX
-\label{suppfigure:DPasr}
-\end{suppfigure}
+%\begin{suppfigure}
+%\includegraphics[width=\textwidth]{asrDPdelta.pdf}
+%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the D/P+ $\delta$, ploidy model} % XXX
+%\label{suppfigure:DPasr}
+%\end{suppfigure}
 
@@
 
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDPAB.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the D/P D/P+A/B ploidy model} % XXX
-\label{suppfigure:DPnodipABasr}
-\end{suppfigure}
+%\begin{suppfigure}
+%\includegraphics[width=\textwidth]{asrDPAB.pdf}
+%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the D/P D/P+A/B ploidy model} % XXX
+%\label{suppfigure:DPnodipABasr}
+%\end{suppfigure}
 
@@
 
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDPABdelta.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the D/P+$\delta$+A/B ploidy model.} % XXX
-\label{suppfigure:DPABasr}
-\end{suppfigure}
+%\begin{suppfigure}
+%\includegraphics[width=\textwidth]{asrDPABdelta.pdf}
+%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the D/P+$\delta$+A/B ploidy model.} % XXX
+%\label{suppfigure:DPABasr}
+%\end{suppfigure}
 
@@
 
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIC.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the I/C breeding system model} % XXX
-\label{suppfigure:ICasr}
-\end{suppfigure}
+%\begin{suppfigure}
+%\includegraphics[width=\textwidth]{asrIC.pdf}
+%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the I/C breeding system model} % XXX
+%\label{suppfigure:ICasr}
+%\end{suppfigure}
 
@@
 
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrICAB.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the I/C+A/B breeding system model} % XXX
-\label{suppfigure:ICABasr}
-\end{suppfigure}
+%\begin{suppfigure}
+%\includegraphics[width=\textwidth]{asrICAB.pdf}
+%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the I/C+A/B breeding system model} % XXX
+%\label{suppfigure:ICABasr}
+%\end{suppfigure}
 
@@
 
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCP.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP ploidy and breeding system model} % XXX
-\label{suppfigure:IDCDCPnodipasr}
-\end{suppfigure}
+%\begin{suppfigure}
+%\includegraphics[width=\textwidth]{asrIDCDCP.pdf}
+%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP ploidy and breeding system model} % XXX
+%\label{suppfigure:IDCDCPnodipasr}
+%\end{suppfigure}
 
@@
 
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCPdelta.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP+$\delta$ ploidy and breeding system model} % XXX
-\label{suppfigure:IDCDCPasr}
-\end{suppfigure}
+%\begin{suppfigure}
+%\includegraphics[width=\textwidth]{asrIDCDCPdelta.pdf}
+%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP+$\delta$ ploidy and breeding system model} % XXX
+%\label{suppfigure:IDCDCPasr}
+%\end{suppfigure}
 
@@
 
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCPAB.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP+A/B polyploidy and breeding system model} % XXX
-\label{suppfigure:IDCDCPnodipABasr}
-\end{suppfigure}
+%\begin{suppfigure}
+%\includegraphics[width=\textwidth]{asrIDCDCPAB.pdf}
+%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP+A/B polyploidy and breeding system model} % XXX
+%\label{suppfigure:IDCDCPnodipABasr}
+%\end{suppfigure}
 
@@
 
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCPdeltaAB.pdf}
-\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP+$\delta$+A/B, polyploidy and breeding system model} % XXX
-\label{suppfigure:IDCDCPABasr}
-\end{suppfigure}
+%\begin{suppfigure}
+%\includegraphics[width=\textwidth]{asrIDCDCPdeltaAB.pdf}
+%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP+$\delta$+A/B, polyploidy and breeding system model} % XXX
+%\label{suppfigure:IDCDCPABasr}
+%\end{suppfigure}

### 2019-07-12 manuscript/tables.tex
Adding the correct figures and tables to results and first part of supplement

--- 2019-07-11 manuscript/tables.tex
@@
 \begin{document}
+
+\begin{table}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|r|c|c|c|c|l|}
+\hline
+Model                & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M2    & M3     & M4    & M5     & Evidence                                                                                     \\ \hline
+M1.
+D/P              & -1283.76                                                          & 59.90 & 49.23  & 60.48 & 58.82  & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M1\end{tabular}            \\ \hline
+\textbf{M2.
+CID D/P} & \textbf{-1223.86}                                                 &       & -10.66 & 0.579 & -1.079 & \begin{tabular}[c]{@{}l@{}}Strong preference for \\ M2 over M1\end{tabular}              \\ \hline
+M3.
+D/P+A/B          & -1234.52                                                          &       &        & 11.24 & 9.58   & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}   \\ \hline
+\textbf{M4.
+D/P+A/B asym}     & \textbf{-1223.28}                                                          &       &        &       & -1.658 & \begin{tabular}[c]{@{}l@{}}Moderate evidence for\\ only asymmetric hidden rates\end{tabular} \\\hline
+M5.
+D/P+A/B all asym & -1224.93                                                          &       &        &       &        &          \\ \hline
+\end{tabular}
+\caption{Bayes factors of  ploidy only models without diplodization.
+Results indicate that a character independent model (M2) is strongly preferred over BiSSE (M1)  model.
+There is no evidence that character independent model is different from M4, the HiSSE model with asymmetric hidden rates.
+M4 is moderately preferred over the more complicated M5, a HiSSE model that has asymmetry for all transition rates.}
+\label{table:M1M5}
+\end{table}
+
+\begin{table}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|r|c|c|c|c|l|}
+\hline
+Model                           & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M7    & M8     & M9    & M10   & Evidence                                                                                     \\ \hline
+M6.
+D/P+$\delta$                & -1268.83                                                          & 55.41 & 42.78  & 53.37 & 52.85 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M6\end{tabular}            \\ \hline
+\textbf{M7.
+CID D/P+$\delta$}   & \textbf{-1212.42}                                                 &       & -12.62 & -2.04 & -2.04 & \begin{tabular}[c]{@{}l@{}}Model M7 moderately\\ preferred over M9 and M10\end{tabular}      \\ \hline
+M8.
+D/P+A/B $\delta$            & -1214.46                                                          &       &        & 10.58 & 10.07 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular} \\ \hline
+M9.
+D/P+A/B +$\delta$ asym      & -1214.46                                                          &       &        &       & -0.51 & No evidence                                                                                  \\ \hline
+M10.
+D/P+A/B +$\delta$ all asym & -1214.97                                                          &       &        &       &       &         \\ \hline
+\end{tabular}
+\caption{Bayes factors of  ploidy only models with diplodization.
+Results indicate that a character independent model (M6) is strongly preferred over BiSSE (M1)  model.
+M6 is moderately preferred over any of the HiSSE model swith asymmetric hidden rates (M9, M10).}
+\label{table:M6M10}
+\end{table}
+
+\begin{table}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|r|c|c|c|c|l|}
+\hline
+Model                      & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M12   & M13   & M14   & M15   & Evidence                                                                                       \\ \hline
+M11.
+I/C                   & -1309.07                                                          & 41.13 & 38.59 & 61.34 & 61.40 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M111\end{tabular}            \\ \hline
+M12.
+CID I/C               & -1267.93                                                          &       & -2.53 & 20.21 & 20.37 & \begin{tabular}[c]{@{}l@{}}Models with asymmetric rates\\ are preferred over  M12\end{tabular} \\ \hline
+M13.
+I/C+A/B               & -1270.47                                                          &       &       & 22.75 & 22.80 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}   \\ \hline
+\textbf{M14.
+I/C+A/B asym} & \textbf{-1247.72}                                                 &       &       &       & 0.05 & No evidence                                                                                    \\ \hline
+\textbf{M15.
+I/C+A/B all asym }     & \textbf{ -1247.66}                                                          &       &       &       &       &       \\ \hline
+\end{tabular}
+\caption{Bayes factors of breeding system only models. Results indicate that the HiSSE models with asymmetric rates (M14, M15) are strongly preferred over every other model.}
+\label{table:M11M15}
+\end{table}
+
+\begin{table}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|r|c|c|c|c|l|}
+\hline
+Model                            & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M18   & M19   & M20   & Evidence                                                                                                   \\ \hline
+M16.
+ID/CP/CD                    & -1459.11                                                          & 45.11 & 65.91 & 63.98 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M16\end{tabular}                         \\ \hline
+M17.
+CID ID/CP/CD                & *                                                                 &       &       &       &                                                                                                            \\
+M18.
+ID/CP/CD+A/B                & -1414                                                             &       & 20.79 & 18.87 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}               \\ \hline
+\textbf{M19.
+ID/CP/CD +A/B asym} & \textbf{-1393.20}                                                 &       &       & -1.92 & \begin{tabular}[c]{@{}l@{}}Asymmetric hidden rates moderately\\ preferred over all asymmetric\end{tabular} \\ \hline
+M20.
+ID/CP/CD +A/B all asym      & -1393.12                                                          &       &       &       &   \\ \hline
+\end{tabular}
+\caption{Bayes factors of ploidy and breeding system withou diploidization.
+Results indicate that the MuHiSSE model with asymmetric hidden rates (17) is  strongly preferred over M16 and M18 and moderately preferred over the MuHiSSe with all rates asymetric (M20).
+*Marginal log-likelihood for M17 could not be calculated in our limited amount of computational time.}
+\end{table}
+
+\begin{table}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|r|c|c|c|c|l|}
+\hline
+Model                                 & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M22   & M23    & M24   & M25   & Evidence                                                                                         \\ \hline
+M21.
+ID/CP/CD+$\delta$                & -1454.68                                                          & 55.48 & 46.031 & 67.94 & 65.15 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M21\end{tabular}               \\ \hline
+M22.
+CID ID/CP/CD+$\delta$            & -1399.201                                                         &       & -9.452 & 12.45 & 9.675 & \begin{tabular}[c]{@{}l@{}}Model M24 strongly preferred \\ over M22\end{tabular}                 \\ \hline
+M23.
+ID/CP/CD+A/B $\delta$            & -1408.65                                                          &       &        & 21.91 & 19.12 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}     \\ \hline
+\textbf{M24.
+ID/CP/CD +$\delta$ asym} & \textbf{-1386.74}                                                 &       &        &       & -2.78 & \begin{tabular}[c]{@{}l@{}}Asymmetric hidden rates \\ preferred over all asymmetric\end{tabular} \\ \hline
+M25.
+ID/CP/CD +$\delta$ all asym      & -1389.52                                                          &       &        &       &       &                      \\ \hline
+\end{tabular}
+\caption{Bayes factors of ploidy and breeding system with diploidization.
+Results indicate that the MuHiSSE model with asymmetric hidden rates (M24) is  strongly preferred over M21-M23 and moderately preferred over the MuHiSSe with all rates asymetric (M25).}
+\label{table:M21M25}
+\end{table}
 
@@
 \begin{table}[]
-\addtolength{\tabcolsep}{-5pt}
+\addtolength{\tabcolsep}{-3pt}
 \begin{tabular}{|l|c|c|c|c|}
 \hline
-Model                       & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
+Model                       & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=ln(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
 M26.

### 2019-07-12 manuscript/response.tex
- extensive changes to response.tex letter - minor changes to others, mainly table legends (%B: comments)

@@
 
-Dear Dr. Schoen and reviewers,
+Dear Dr. Schoen and Reviewers,
 
 \medskip
+
 Thank you very much for your close reading of our manuscript and supportive comments.
 We appreciate your constructive suggestions, each of which we address below.
-You will find a big difference on the number of  models  that we have fitted for the revision based on your suggestions (originally 12 now 29).
-The two main reasons to increase the number of models were to first address one key question by made by all referees about the symmetry of the hidden states, and second to address a remaining question about adding complexity to two-state models.
-For the symmetry of the hidden states  question we found that is relevant to have asymmetric hidden state rates and in our case, it strengthened the differences amongst net diversification rates.
-The second question,  comes from a vague idea that we had while submitting the first version of this article: by increasing from two to three states the models, are we explaining better the history of diversification?
-This second question could not be addressed directly by model comparison statistics because the two-state and three-state models need different input (i.e.
-(0,1) versus (0,1,2)).
-However, the species and the tree are the same and our intuition indicated that they should carry the same biological evidence despite any particular coding needed as input.
-During this revision, we figured a way to recode the three-state models to make them equivalent to the two-state models (the new lumped models).
-Thanks to this equivalency, two-state and three-state models can now be compared to argue for the importance of adding key trait information and states in diversification analyses.
-We also revised the original dataset thanks to the suggestion of Reviewer 1, and we were able to correct discrepancies with previous dataset as well as add more sample (651 taxa).
+You will find a great difference in the number of models that we have fitted for the revision based on reviewer suggestions (originally 12 now 29).
+There are two main reasons for this sizable increase.
 
-We hope the changes and the new models are satisfactory and exciting to you and the reviewers.
+The first reason is additional work to address a key concern from all referees, about the transition symmetry among the hidden states. 
+Briefly, we found that is relevant to have asymmetric hidden state rates and, in our case, it increased the differences among net diversification rates. 
+We added and altered the text to reflect these additional analyses.
 
+The second reason is related to a question we mulled over after submitting the first version of this article.
+Specifically, we wondered whether increasing the state space from two- to three-states in our models of the diversification process explains the data better (and how they ought to be best compared).
+This  trait coding question could not be addressed directly by model comparison statistics because the two-state and three-state models need different input (i.e. (0,1) versus (0,1,2)). 
+The species and the tree are identical, however, and our intuition indicated that they should carry the same biological evidence despite a particular coding input. 
+During this revision, we figured a way to recode the three-state models to make them equivalent to the two-state models (the new ``lumped" models). 
+This equivalency enables comparison of two-state and three-state models, and assessment of their relative explanatory value after adding key trait information and states in diversification analyses. 
+We also revised the original dataset (thanks to the suggestion of Reviewer 1) and we were able to correct discrepancies with a previous dataset, as well as add more data (now 651 taxa).
+
+We hope the changes and the new models are satisfactory to the editorial staff and the reviewers.
 \medskip
@@
 Rosana Zenil-Ferguson, on behalf of all authors
-% E: Pardon my speaking for you, Rosana. :)  Obviously change whatever wording below---starting to draft responses just helped me think through the requests.
-% R: I'm super excited now as you can see
 
@@
 Yes, and indeed this has already been shown.
-We fixed the text and added a reference. Nakov paper has not been accepted yet for publication or it has not been announced (July 9th, 2019).
+We fixed the text and added a reference. Nakov paper has not been accepted yet for publication or it has not been announced as such (as of July 9th, 2019).
 

### 2019-07-12 manuscript/methods.tex
- extensive changes to response.tex letter - minor changes to others, mainly table legends (%B: comments)

@@
 
-In order to compare model fits for ploidy only versus breeding system only versus both traits, we use the technique of \citet{tarasov_2019} for lumping states together.
+In order to compare model fits for ploidy-only versus breeding system-only versus both traits, we use the technique of \citet{tarasov_2019} for lumping states together.
 We use the state space of the $IC/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
@@
 We insert an artificial division in the $CP$ state, so that one substate contains the $CP$ species that arrived via the one-step pathway and the other substate contains the $CP$ species that arrived via the two-step pathway.
-We consider only unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and no strong support for diploidization (see below).
+We consider unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and no strong support for diploidization (see below).
 
@@
 \end{tabular}
-\caption{Bayes factors of  ploidy only models without diplodization.
+\caption{Bayes factors of  ploidy-only models without diplodization.
 Numbers greater that 1 mean moderate support relative to the model in the row label.
@@
 
+% B: M111->M11
 \begin{table}
@@
 M11.
-I/C                   & -1309.07                                                          & 41.13 & 38.59 & 61.34 & 61.40 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M111\end{tabular}            \\ \hline
+I/C                   & -1309.07                                                          & 41.13 & 38.59 & 61.34 & 61.40 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M11\end{tabular}            \\ \hline
 M12.
@@
 \end{tabular}
-\caption{Bayes factors of breeding system only models. Results indicate that the HiSSE models with asymmetric rates (M14, M15) are strongly preferred over every other model.}
+\caption{Bayes factors of breeding system-only models. Results indicate that the HiSSE models with asymmetric rates (M14, M15) are strongly preferred over every other model.}
 \label{table:M11M15}
@@
 % E: Could we include a figure of the traits on the tree?  At least as supp info. FIXME
+%B: "K=ln(BF(M1,M2))" -> "K=ln(BF($M_{i}$,$M_{ii}$))" because we use M1 and M2 for model names.
 \begin{table}[]
@@
 \hline
-Model                       & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=ln(BF(M1,M2)) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
+Model                       & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=ln(BF($M_{i}$,$M_{ii}$)) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
 M26.

### 2019-07-12 manuscript/results.tex
- extensive changes to response.tex letter - minor changes to others, mainly table legends (%B: comments)

@@
 \end{tabular}
-\caption{Bayes factors of  ploidy only models with diplodization.
+\caption{Bayes factors of  ploidy-only models with diplodization.
 Results indicate that a character independent model (M6) is strongly preferred over BiSSE (M1)  model.
-M6 is moderately preferred over any of the HiSSE model swith asymmetric hidden rates (M9, M10).}
+M6 is moderately preferred over any of the HiSSE models with asymmetric hidden rates (M9, M10).}
 \label{supptable:M6M10}

### 2019-07-12 manuscript/tables.tex
- extensive changes to response.tex letter - minor changes to others, mainly table legends (%B: comments)

@@
 Results indicate that a character independent model (M6) is strongly preferred over BiSSE (M1)  model.
-M6 is moderately preferred over any of the HiSSE model swith asymmetric hidden rates (M9, M10).}
+M6 is moderately preferred over any of the HiSSE models with asymmetric hidden rates (M9, M10).}
 \label{table:M6M10}
@@
 
+% M111 -> M11
 \begin{table}
@@
 
+%K=log(BF(M1,M2) -> K=ln(BF($M_{i}$,$M_{ii}$))
 \begin{table}
@@
 \hline
-Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
+Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=ln(BF($M_{i}$,$M_{ii}$))& \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
 M3. D/P+A/B                                 & -1234.52                     &    M3 vs. M4                 & 11.239                  & M4 (Strong)   \\

### 2019-07-12 manuscript/results.tex
Figures and tables in the supplement fixed

@@
 \begin{suppfigure}
+\includegraphics[width=\textwidth]{asrDP.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of theM1. D/P ploidy model} % XXX
+\label{suppfigure:DPnodipasr}
+\end{suppfigure}
+
+\begin{suppfigure}
 \includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P ploidy model} % XXX
+\caption{Posterior distribution for each of the parameters in the M1. D/P model} % XXX
 \label{suppfigure:DPnodip}
@@
 
-%\begin{suppfigure}
-%\includegraphics[width=\textwidth]{asrDP.pdf}
-%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node of the D/P ploidy model} % XXX
-%\label{suppfigure:DPnodipasr}
-%\end{suppfigure}
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrDPAB.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the M4. D/P+A/B asym model} % XXX
+\label{suppfigure:DPnodipABasr}
+\end{suppfigure}
 
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{bisseDPposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P+ $\delta$ ploidy model} % XXX
-\label{suppfigure:DP}
-\end{suppfigure}
-
-%\begin{suppfigure}
-%\includegraphics[width=\textwidth]{asrDPdelta.pdf}
-%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the D/P+ $\delta$, ploidy model} % XXX
-%\label{suppfigure:DPasr}
-%\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{hisseDPnodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P+A/B ploidy model} % XXX
+\includegraphics[width=\textwidth]{hisseDPnodipasymposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the M4. D/P+A/B asym model} % XXX
 \label{suppfigure:DPnodipAB}
@@
 
-%\begin{suppfigure}
-%\includegraphics[width=\textwidth]{asrDPAB.pdf}
-%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the D/P D/P+A/B ploidy model} % XXX
-%\label{suppfigure:DPnodipABasr}
-%\end{suppfigure}
-
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{hisseDPposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the D/P+$\delta$+A/B ploidy model.  The axis is offset in one location so that the two overlapping distributions can be seen.} % XXX
-\label{suppfigure:DPAB}
+\includegraphics[width=\textwidth]{asrIC.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the M11.I/C model} % XXX
+\label{suppfigure:ICasr}
 \end{suppfigure}
-
-%\begin{suppfigure}
-%\includegraphics[width=\textwidth]{asrDPABdelta.pdf}
-%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the D/P+$\delta$+A/B ploidy model.} % XXX
-%\label{suppfigure:DPABasr}
-%\end{suppfigure}
 
@@
 \includegraphics[width=\textwidth]{bisseSIposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the I/C breeding system model} % XXX
+\caption{Posterior distribution for each of the parameters in the M11. I/C model} % XXX
 \label{suppfigure:IC}
@@
 
-%\begin{suppfigure}
-%\includegraphics[width=\textwidth]{asrIC.pdf}
-%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the I/C breeding system model} % XXX
-%\label{suppfigure:ICasr}
-%\end{suppfigure}
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrICAB.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the M14. I/C+A/B asym model} % XXX
+\label{suppfigure:ICABasr}
+\end{suppfigure}
 
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{hisseSInoretposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the I/C+A/B, breeding system model} % XXX
+\includegraphics[width=\textwidth]{hisseSIasymposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the M14. I/C+A/B asym model} % XXX
 \label{suppfigure:ICAB}
@@
 
-%\begin{suppfigure}
-%\includegraphics[width=\textwidth]{asrICAB.pdf}
-%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the I/C+A/B breeding system model} % XXX
-%\label{suppfigure:ICABasr}
-%\end{suppfigure}
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIDCDCP.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M16. ID/CD/CP model} % XXX
+\label{suppfigure:IDCDCPnodipasr}
+\end{suppfigure}
 
@@
 \includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the ID/CD/CP ploidy and breeding system model} % XXX
+\caption{Posterior distribution for each of the parameters in the  M16. ID/CD/CP model} % XXX
 \label{suppfigure:IDCDCPnodip}
@@
 
-%\begin{suppfigure}
-%\includegraphics[width=\textwidth]{asrIDCDCP.pdf}
-%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP ploidy and breeding system model} % XXX
-%\label{suppfigure:IDCDCPnodipasr}
-%\end{suppfigure}
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIDCDCPAB.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M19. ID/CD/CP+A/B asym model} % XXX
+\label{suppfigure:IDCDCPnodipABasr}
+\end{suppfigure}
 
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{musseDPSIposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the ID/CD/CP+$\delta$, polyploidy and breeding system model} % XXX
-\label{suppfigure:IDCDCP}
-\end{suppfigure}
-
-%\begin{suppfigure}
-%\includegraphics[width=\textwidth]{asrIDCDCPdelta.pdf}
-%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP+$\delta$ ploidy and breeding system model} % XXX
-%\label{suppfigure:IDCDCPasr}
-%\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{muhisseDPSInodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the ID/CD/CP+A/B polyploidy and breeding system model} % XXX
+\includegraphics[width=\textwidth]{muhissenodipasymposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the M19. ID/CD/CP+A/B asym model} % XXX
 \label{suppfigure:IDCDCPnodipAB}
 \end{suppfigure}
-
-%\begin{suppfigure}
-%\includegraphics[width=\textwidth]{asrIDCDCPAB.pdf}
-%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP+A/B polyploidy and breeding system model} % XXX
-%\label{suppfigure:IDCDCPnodipABasr}
-%\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{muhisseDPSIposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the ID/CD/CP+$\delta$+A/B, polyploidy and breeding system model} % XXX
-\label{suppfigure:IDCDCPAB}
-\end{suppfigure}
-
-%\begin{suppfigure}
-%\includegraphics[width=\textwidth]{asrIDCDCPdeltaAB.pdf}
-%\caption{Ancestral state reconstruction showing the maximum a posteriori for each node in the ID/CD/CP+$\delta$+A/B, polyploidy and breeding system model} % XXX
-%\label{suppfigure:IDCDCPABasr}
-%\end{suppfigure}

### 2019-07-12 manuscript/results.tex
updated pathways figure with new rate values

@@
 In the two-step pathway, SI is first lost within the diploid state, followed by polyploidization.
-Determining the relative contribution of these pathways based on our estimated transition rates from the $ID/CD/CP$ model, we find that the one-step pathway is more likely on short timescales and the two-step pathway is more likely on long timescales (\cref{figure:pathways}, left panels).
+Determining the relative contribution of these pathways based on our estimated transition rates from the $ID/CD/CP$ model (median rate values from M16), we find that the one-step pathway is more likely on short timescales and the two-step pathway is more likely on long timescales (\cref{figure:pathways}, left panels).
 Beginning with a single SI diploid lineage, when not much time has elapsed, the one-step pathway is more likely because it only necessitates a single event to effect transition to the SC polyploid state.
@@
 
-% see Will's notes in discussion.tex, FIXME
-%   also use transition rates from transition-only (no diversification) analysis
-%   also use stochatic maps, (a) for CP tips, what paths were taken? (b) for steps away from ID, which path was started?
+% also use transition rates from transition-only (no diversification) analysis? todo
 
@@
 By analogy, envisioning the states as stepping stones, the extent to which each stone grows or shrinks over time affects the utility of each possible path.
-Allowing for the different net diversification rates for each state, we find a qualitative difference in the relative pathway contributions.
+Allowing for the different net diversification rates for each state (again using median rate estimates from M16), we find a qualitative difference in the relative pathway contributions.
 In particular, the lower rate of net diversification in the $CD$ state, relative to $ID$, means that relatively fewer lineages are available to complete the second step of the two-step pathway, ending in $CP$.

### 2019-07-12 summarystats/pathways.R

@@ -51,11 +51,13 @@ path.probs2 <- function(t, q13, q14, q34)
 # Wa = 1 -> 4
 # Wb = 1 -> 3 -> 4
 
-# MAP values from EstimatesIDCDPnodelta.csv
-q14 <- 0.0681          # ID -> CP, rho_I
-q13 <- 0.2066          # ID -> CD, q_IC
-q34 <- 0.0370          # CD -> CP, rho_C
+# median values from MuSSE_ploidysinodip250K.log (via tracer)
+q14 <- 0.0779         # ID -> CP, rho_I
+q13 <- 0.1895         # ID -> CD, q_IC
+q34 <- 0.0188         # CD -> CP, rho_C
 q41 <- q31 <- q43 <- 0 # irreversible
+
+# also use CID estimates?
 
 # Q is only needed for method 1
 Q <- matrix(c(
@@ -86,9 +88,9 @@ abline(h=0.5, lty=3)
 
 ### Try with diversification ###
 
-r1 <-  0.4607
-r3 <-  0.0684
-r4 <- -0.0877
+r1 <-  0.4618
+r3 <-  0.0605
+r4 <- -0.0121
 
 Qd <- matrix(c(
        -q13 - q14 + r1,  q13,             q14,       0,

### 2019-07-12 manuscript/response.tex
- reworded a few responses and spaced the query/answer formatting
- Emma/Rosana CHECK: paragraph re: why not include polymorphisms
  without transitions

@@
 
-okay % TODO
+[Considered.] % TODO
 
@@
 
-check % TODO
+[Considered.] % TODO
 
@@
 
-Fixed.
+Done.
 
@@
 
-We reworded this whole paragraph, as part of better explaining how polyploidization disables gametophytic SI.
+We reworded this whole paragraph, with the aim of providing a better explanation for how polyploidization disables gametophytic SI.
 
@@
 
-We reworded this whole paragraph, as part of better explaining the SSE family of models.
+We reworded this whole paragraph and aimed to provide a better explanation of the SSE family of models.
 
@@
 \end{quote}
-We acknowledge this is a potential problem.
-It is true that chromosome number multiplicity can bias coding towards more (or less) transitions when it heavily depends on the most common cytotype.
-For the majority of species (~85\%)  the ploidy was not directly dictated by us but by the original sources of the data.
+
+This is a part of an important broader point that we hope to clarify. 
+First, it is true that chromosome number multiplicity can bias coding towards more (or less) transitions when it heavily depends on the most common cytotype. 
+For the majority of species (~85\%), we did not make the ploidy call, which we instead got from the original sources of the data.
 In the originally sources ploidy has been classified not only most common cytotype but geographical distribution, multivalence, and/or  genome sizes.
@@
 We differ in exactly 35 species from \citet{robertson_2011} were we found a new reference or a new chromosome count. 
-Classifying using a chromosome number change models only could also suffer from inflation of polyploidy events due to the lack of diploidization parameters, we found some specific species in the  \citet{robertson_2011} sample that were clearly aneuploids (one chromosome difference) that were classified as polyploids using chromosome number models, we manually corrected for those hoping to minimize obvious mistakes.
+%
+%
+%B: I don't think this is quite right.
+It's not the "lack of diploidization parameter" but the level of organization that we are measuring that is the key objection to coding each individual mutation as a transition.
+The key point is that *SSE models+tip states are definitely a poor estimator of mutation rate.
+(Irony about its lack of accuracy at the "species" level noted.)
+%rm: Classifying using a chromosome number change models only could also suffer from inflation of polyploidy events due to the lack of diploidization parameters, we found some specific species in the  \citet{robertson_2011} sample that were clearly aneuploids (one chromosome difference) that were classified as polyploids using chromosome number models, we manually corrected for those hoping to minimize obvious mistakes.
+
+% EMMA see below:
+%B: I think this should be added to clarify; as well, it is something someone should write up as a paper, because it's been a topic of review conversations for at least a decade.
+Second, we are interested estimating lineage transition rates and not `events', the individual mutations to polyploid state.
+Following a mutation, the outcome of an interplay between multiple evolutionary forces may or may not cause a transition. 
+Coding each polyploid tissue separately (indeed, it is possible that a polyploid tissue and not individual is being scored in a study) would then confound  lineage transition and mutation on multiple levels (somatic tissue, individual, and species transition).
+Our macroevolutionary parameters would then include estimate of mutation rate as well as the lineage transition rate, with the result representing a drastic and perhaps ill-advised overestimate of the lineage transition rate.
 
@@
 How is it even sexual?
-Regardless, the discussion of this pentaploid should be more clearly integrated back into the central argument here, about the absense of SI polyploids (I think that's the central argument?).
+Regardless, the discussion of this pentaploid should be more clearly integrated back into the central argument here, about the absence of SI polyploids (I think that's the central argument?).
 \end{quote}
@@
 %B: I edited the manuscript (methods.tex) as explained below:
-We removed the discussion and simply reference a more detailed treatment of this single individual.
-We agree that this observation by \citet{camadro_1981} may be distracting, and now instead  point the reader to our previously published paper \citep{robertson_2011}, where the possible existence of pentaploid that fails to set seed following self-pollination is discussed at some length.
+We removed the discussion and simply reference a more detailed treatment of this single individual potato plant.
+This observation \citet{camadro_1981} and the manifold convoluted cases by which pentaploids may be sterile are distracting, and were previously covered in \citep{robertson_2011}.
+Instead, we point the reader to \citet{robertson_2011}, where the possible existence of a sterile or incompatible pentaploid, which fails to set seed following self-pollination, is discussed at some length.
 
@@
 
-Fixed.
+Added.
 
@@
 We found using Bayes factors that a model including asymmetric hidden rates is preferred over a model with equal hidden rates. 
-Also,  the asymmetry in the trait rates have moderate to not support of being better than a model with equal trait transition rates.
-%TODO put number.
+Also,  the asymmetry in the trait rates have moderate to not support of being better than a model with equal trait transition rates.  %TODO put number.
 
@@
 \end{quote}
-Unfortunately, we weren't able to run multiple chains for all models due to some limitations to our computer resources.
-However, we were able to test for the convergence of  three key models using Gelman's $\hat{R}$ test statistic.
+
+Unfortunately, we were not able to run multiple chains for all models due to limited computer resources. 
+We were, however, able to test for the convergence of  three key models using Gelman's $\hat{R}$ test statistic. 
 We tested convergence with two runs of 130,000 generations (after 5,000 generations of burn-in) for models D/P (BiSSE), D/P+A/B asym (HiSSE) and ID/CD/CP+A/B asym (MuHiSSE). 
-The statistics as well as the Gelman plots in  \ref{figure:convergencestats} show that with that number of generations the models were converging.
-We increased the number of generations to be safe for all the models that we did not test and visually assessed the mixing of  models on top of demanding ESS larger than 200 for parameters.
+The statistics and the Gelman plots in  \ref{figure:convergencestats} show that with that number of generations the models were converging. 
+We increased the number of generations for all the models that we did not test, and visually assessed the mixing of models on top of demanding ESS larger than 200 for parameters.
 
@@
                  \end{minipage}
-    \caption{ Convergence statistics and Gelman plots for D/P (BiSSE), D/P+A/B asym (HiSSE) at the top and ID/CD/CP+A/B asym (MuHiSSE,bottom).
-Upper limits of Gelman's test statistics should be close to 1 for an MCMC to be considered convergent.}
+    \caption{ Convergence statistics and Gelman plots for D/P (BiSSE, left), D/P+A/B asym (HiSSE, right) at the top and ID/CD/CP+A/B asym (MuHiSSE, bottom).
+Upper limits of Gelman's test statistics should be close to 1.0 for MCMC to be considered convergent.}
     \label{figure:convergencestats}
@@
 \end{quote}
-We ran each model for 200,000 generations. Unfortunately, we don't have access to more of 96 hours of computational time to increase those but we have calculated the convergence as indicated above.
+
+We ran each model for 200,000 generations. Unfortunately, we don't have access to more than 96 hours of computational time to increase those, but we have calculated the convergence as indicated above.
 
@@
 In Fig.~4, the contributions of each pathway are functions of time.
-We can see there both the short-term and long-term behavior.
-% Rosana, what do you think?- I think this is a good answer- I think it is good
+There, we can see both the short-term and long-term behavior.
 
@@
 
-consider % TODO
+[Considered.] % TODO %We can defend this on account of repeating some of the points because of numerous and complex  questions; aimed to prevent flipping pages back
 
@@
 
+%B: Rosana, you know this best, but it seems to me that the current phrasing (below) is unclear.
 The soaking noise comes from the hidden state so in the presence of hidden states finding that diploidization matters is key. 
 All the Bayes factors indicated strong support for models that include diploidization. 
-However, the diploidzation parameter per se is difficult to estimate (wide credible intervals) mainly because diploidization is a genome size process that can be partially or minimally recover from ploidy data.
+However, the diploidization parameter is difficult to estimate (wide credible intervals) mainly because diploidization is a genome size process that can be partially or minimally recover from ploidy data.
 This is the reason why we do not overstate this finding, the importance of diploidization needs to be studied using genomes.
@@
 
-Seems reasonable. % TODO
+[Seems reasonable.] % TODO
 
@@
 
-consider % TODO
+[Considered.] % TODO
 
@@
 
-Fixed.
+Clarified.
 
@@
 
-We're pleased you found the results interesting!
+We're pleased R2 found the results interesting!
 
@@
 \end{quote}
+
 We have now fitted models with asymmetric hidden rates and asymmetric trait rates.
@@
 
-Thanks for this suggestion, which we have incorporated in the Discussion. % TODO
+Thank you for this suggestion, which we have incorporated in the Discussion. % TODO
 
@@
 
-Fixed. % Done: Emma did the new figure
+Done. % Done: Emma did the new figure
 
@@
 
-Fixed.
+Modified.
 
@@
 
-Thank you for this positive summary.
+We appreciate this positive summary.
 
@@
 \end{quote}
+
 We have cited them accordingly. % TODO
@@
 
-Yes, and indeed this has already been shown.
-We fixed the text and added a reference. Nakov paper has not been accepted yet for publication or it has not been announced as such (as of July 9th, 2019).
+Yes, this has already been shown.
+We fixed the text and added a reference. 
+Nakov paper has not been accepted yet for publication or it has not been announced as such (as of July 9th, 2019).
 

### 2019-07-12 summarystats/pathways.R
CID pathway results
qualitatively the same as trans-only under SDD model

@@ -58,6 +58,10 @@ q34 <- 0.0188         # CD -> CP, rho_C
 q41 <- q31 <- q43 <- 0 # irreversible
 
 # also use CID estimates?
+q14 <- 0.0228         # ID -> CP, rho_I
+q13 <- 0.0703         # ID -> CD, q_IC
+q34 <- 0.0334         # CD -> CP, rho_C
+q41 <- q31 <- q43 <- 0 # irreversible
 
 # Q is only needed for method 1
 Q <- matrix(c(

### 2019-07-12 manuscript/discussion.tex
edits/reorg of Diploidization section of Discussion

@@
 For example, the rapidly-diversifying clade of mostly SC species within \emph{Solanum} suggests that geography may play a role. % Boris, please check!  Australian?  fixme
-% alpha is always low
-%   new supp fig showing how A & B are inferred on the tree?  at least for ID/CD/CP+A/B model
+%     but don't be circular, test in different clade
+% alpha is always low: new supp fig showing how A & B are inferred on the tree?  at least for ID/CD/CP+A/B model
 Alternatively, we could view the hidden trait as merely a statistical trick, providing an easy way to model extra heterogeneity without directly representing a real trait.
@@
 %     \rho_I > \rho_C in all!
-
-% Will: I really like the pathways analyses and have a couple comments
-%
-% 1) Were the "without diversification" and "with diversification" analyses computed
-%    using the same MAP transition rate estimates? Transition rates among states will
-%    be very different if estimated using a model that is diversification-independent 
-%    compared to when estimated with a trait-dependent diversification model. Since the point of 
-%    showing these side-by-side seems to be emphasizing the importance of considering trait-dependent
-%    diversification, I wonder how different the "without diversification" contributions would be
-%    if calculated from transition rates estimated using a diversification-independent model.
-%  R- Good point. Emma will clarify
-%  E: Yes, good point.  Rosana, do you have results for state-independent diversification?  We could compare with using those rates.
-%
-% 2) I think stochastic mapping would be another useful perspective on the pathways question
-%    (and yeah, its a bit late in the game to propose analyses, so maybe next time!). With stochastic 
-%    mapping we could use the entire posterior estimate of the transition rates rather than a point
-%    estimate. But more importantly we'd get the relative contributions of the pathways estimated 
-%    over the actual Solanceae tree (which has a lot of short interdependent branches rather than 
-%    hypothetical long branches). 
-%R- It will, it was actually on my code I think for the ID/CD/CP I have the stochastic mapping trees.
-We can do it for the revision.
-I think the only model where it could not finish was the ID/CD/CP+A/B (was taking forever and we had the restriction of time from MSI)
+%     likely way to lose SI
+%     likely way to become P
+%     (compare those relative rates with robertson2011)
 
@@
 
-% Motivated by the widespread recent findings of polyploidy in the ancestry of angiosperms, we also examined and found modest support for diploidization in the history of the family, but without significant effects on inferences of diversification dynamics.
-%B: how about "modest"?
-%R- I think modest is a good way to express it because it is the best we can do with chromosome numbers.
-%B: changed "polyploidy events" (polyploidy is a noun, only.
-polyploid either noun or adjective).
-
-%R: and I: polyploid lineage that has re-diploidized may enhance diversification.
+Polyploidization is known to be common in plants, but the commonness of the reverse process---diploidization---is under active investigation.
+%
+In our analyses, ignoring diploidization if it is common could cause us to underestimate the polyploidization rate, overestimate of the diploid diversification rate, or otherwise confound our conclusions.
+We therefore included a slate of models with a diploidization parameter, and they showed that our main conclusions are robust to this process.
+These models also suggested modest statistical support for diploidization occuring within Solanaceae \cref{supptableFIXME}, although our estimates of the diploidization rate were highly uncertain.
+Furthermore, additional lines of evidence for classifying species as diploid or polyploid (beyond the genus chromosome multiplicity that we primarily relied on) would be needed for more reliable conclusions from an analysis such as ours.
+%In the cases where ploidy level was assigned by us based chromosome multiplicity at the genus level, estimating diploidization might be a potential issue due to ploidy misclassification.  
 %R: By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with comparative phylogenetic models.
-%In the cases where ploidy level was assigned by us based chromosome multiplicity at the genus level, estimating diploidization  might be a potential issue due to ploidy misclassification.
-
 %B: yielding to %R here.
@@
 %B: was: Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.
-%Our exploratory analyses recover some support for inclusion of diploidization in the history of the family, but without significant effects on inferences of diversification dynamics.
-%R- The line above is included in the bottom so no more need for that here
-The implications of the results of our diploidization analyses, as well as much of the other recent work on WGDs, are interesting because they conflict with other lines of evidence, especially those concerning the observed simple synteny of genomes in Solanaceae, and patterns of evolution at self-incompatibility loci.
-Flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}. 
-Following polyploidization, it is possible for genome reorganization, downsizing, and loss to occur \citep{dodsworth_2015, zenil_2016, mandakova_2018}.
-%B: commas, etc.: importantly rm: "has been observed" this is _only_ true on very short time scales
-As a consequence, nearly all extant species classified as ``diploid" in our analyses, based on cytogenetic data, are possibly secondary diploids, having undergone both polyploidization and re-diploidization.
-Unduly ignoring secondary diploidization would necessarily underestimate the rate of transitions from polyploids to diploids.
-It would then likely cause inflated net diversification rates for diploids, because species considered diploid may have been ancestrally polyploid, instead. %B: added before conflict/edits
-%We find strong support for including the diploidization rate parameter ($\delta$) in analyses that include only ploidy . %B <- %Rosana, inspect this please. It's a merge of conflicted lines.
-% R- I put it up so the line above is not necessary anymore
-However, we find only modest support for diploidization in the joint model  (\cref{table:bayesfactors}), and conclusions regarding the relative effects of ploidy and breeding system on diversification are robust to its inclusion.
-Furthermore, estimating the rate of diploidization, based on our ploidy level classifications, is highly uncertain (see parameter $\delta$ in \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}), and yields dramatic changes in ancestral reconstructions (\cref{suppfigure:DPnodipasr,suppfigure:DPasr,suppfigure:DPnodipABasr,suppfigure:DPABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPasr,suppfigure:IDCDCPnodipABasr,suppfigure:IDCDCPABasr}).
 
-Some lines of evidence indicate that polyploidy may have occurred prior to the origin of Solanaceae, rendering all extant `diploids' secondarily derived.
-%B: "polyploid event" or "polyploidy" works, but not "polyploidy event"; see above.
-First, \citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \textit{Solanum lycopersicum}, may have experienced one or more of WGDs.
-A subsequent analysis of synteny between grape and \textit{Solanum} genomes, as well as the distribution between inferred paralogs within genomes of \textit{Solanum} (tomato and potato) each suggested that this lineage experienced a likely round of ancient genome duplication or triplication \citep{tomato2012}.
-The age of the peak of paralog Ks distances is approximately 71$\pm$19 My \citep{tomato2012}. 
-If this is the case, then all of the genomes may have been subsequently re-diploidized, yielding the widespread and common chromosome numbers in this and related families, n=11--12, presently considered to be diploid \citep{robertson_2011}.
+% E: Boris, could you please revisit the following two paragraphs?  TODO
+% Two questions:
+%    Did my reorganization of this section mess anything up?
+%    Can this text be shorter?  It's seems disproportionately long, and I'm hesitant to revise it heavily myself because the logical connections aren't entirely clear to me.
 
-On the other hand, there is little evidence for the occurrence of re-diploidization within Solanaceae, since the origin of the family---the time frame we considered.
-First, the Ks-inferred duplication likely pre-dates the origin of the family (49 My, HPD 46--53 My \citealt{sarkinen_2013}). 
-The best-supported, most-recent WGD \citep{tomato2012} is older than the clade under consideration in this study.
-At best then, the species we consider are ancestrally diploid, having re-diploidized after a suspected WGD, but likely before the origin of the family.
+Other lines of evidence about the prevalence of diploidization within Solanaceae or its ancestors are mixed or even conflicting.
+% recent work on WGDs conflict with other lines of evidence, especially those concerning the observed simple synteny of genomes in Solanaceae, and patterns of evolution at self-incompatibility loci
+%
+% Part I: evidence for WGD in/near family
+On the one hand, polyploidy may have occurred prior to the origin of Solanaceae, rendering all extant `diploids' secondarily derived.
+\Citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \emph{Solanum lycopersicum}, may have experienced one or more whole genome duplications.
+A subsequent analysis of synteny between grape and \emph{Solanum} genomes, as well as the distribution between inferred paralogs within genomes of \emph{Solanum} (tomato and potato) each suggested that this lineage experienced a likely round of ancient genome duplication or triplication \citep{tomato2012}.
+The age of the peak of paralog Ks distances is approximately $71 \pm 19$ My \citep{tomato2012}. 
+If this is the case, then all of the genomes may have been subsequently diploidized, yielding the widespread and common chromosome numbers in this and related families, n=11--12, presently considered to be diploid \citep{robertson_2011}.
+
+% Part II: evidence against WGD in/near family
+On the other hand, there is little evidence for the occurrence of diploidization within Solanaceae itself.
+First, the Ks-inferred duplication \citep{tomato2012} likely pre-dates the origin of the family (49 My, HPD 46--53 My; \citealt{sarkinen_2013}). 
+Thus, the species in our analyes could be ancestrally polyploid, but would have diploidized before the root of the family.
 Second, studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
-Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species, including all of those in the relatively cytogenetically conserved `x=12' group, which includes tomato, potato, eggplant, pepper, and tobacco.
-Despite the limitations of assignment of ploidy levels, we found some support for models that include parameter $\delta$, meaning that further study may be warranted, especially in other groups.
-
-Regardless of whether genome polypoidization, followed by widespread diploidization, is a dominant mechanism of genome evolution in Solanaceae, it is clear that our present understanding of the evolution of ploidy and breeding systems is incomplete.
-An increasingly forceful weight of evidence seems to support WGDs in the ancestry of many angiosperms. %cite
-It is then easily inferred that many species have undergone diploidization, alongside genome size and chromosome number reductions. %cite
-Inference of such a common genome upheaval in eudicot history seems to clash with the data indicating that a homologous mechanism of SI, which breaks down nearly invariably in natural and induced tetraploids \citep{stone_2002,mcclure_2009}, has been present continually in many families, including Solanaceae \citep{igic_2006}.
-Most problematically in this context, it is unclear how to explain the maintenance of trans-generic polymorphism at the orthologous S-loci in this and other families.
-As well, chromosome numbers and genome comparisons within the family (esp.\ `x=12' clade, containing \textit{Solanum} and \textit{Nicotiana}) reveal strong conservation.
+Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species.
+% including all of those in the relatively cytogenetically conserved x=12 group, which includes tomato, potato, eggplant, pepper, and tobacco.
+Specifically, chromosome numbers and genome comparisons within the family (esp.\ the x=12 clade containing \textit{Solanum} and \textit{Nicotiana}) reveal strong conservation.
 \citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ by six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca.\ 10 inversions and 11 translocations (likely underestimated).
 Recovery of such simple relationships would require outstanding convergent loss of duplicated segments.
-In either case, all of the approaches make ample assumptions, and it seems that at least some of them will necessitate deep revisions.
-Our study, for example, made numerous assumptions in coding trait states (\eg what constitutes a polyploid or diploid, SI or SC species), in models used for analyses (\eg irreversibility of SI loss underlain by trans-specific orthologous S-loci), and many more, which could each mislead inference.
-%B: -> %R: Rosana, take a look at this.
+%
+% Part IIb: evidence from SI [might need more explaining?]
+Furthermore, whole genome duplication in a eudicot lineage ancestral to Solanaceae would clash with the evidence that the homologous mechanism of SI that has been present continually in many families \citep{igic_2006} breaks down nearly invariably in natural and induced tetraploids \citep{stone_2002, mcclure_2009}.
+Most problematically in this context, it is unclear how to explain the maintenance of trans-generic polymorphism at the orthologous S-loci in Solanaceae and other families if SI was previously broken down by polyploidization.
+%
+Regardless of whether genome polypoidization, followed by widespread diploidization, is a dominant mechanism of genome evolution in Solanaceae, it is clear that more work is needed for a complete understanding of the joint evolution of ploidy and breeding systems.
+
+% Flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}. 
+% Following polyploidization, it is possible for genome reorganization, downsizing, and loss to occur \citep{dodsworth_2015, zenil_2016, mandakova_2018}.
+% As a consequence, nearly all extant species classified as ``diploid" in our analyses are possibly secondary diploids, having undergone both polyploidization and subsequent diploidization.
+
 % no discussion of magnitude of \delta vs \rho!
-% R- magnitude is not easy to discuss about, what we have are estimates and credible intervals which are completely sample size dependent. With a different sample "magnitude" can change.
+
 % problems: this system has a big set of advantages - carefully curated (co-)occurrences of two traits, genetic basis, etc. and everything is still contradictory and opaque.
@@
 (4) "hight type I errors" -> high, type I errors [comma], (5) line break in penultimate sentence.
-Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as discussed in \citealt{rabosky_2010, beaulieu_2015}), but also a problem of inadequate ``*SSE'' model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focal trait first used to model diversification.
+Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as discussed in \citealt{rabosky_2010, beaulieu_2015}), but also a problem of inadequate SSE model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focal trait first used to model diversification.
 We presented a roadmap with a series of analyses that can identify whether unknown or unobservable traits are worth pursuing. 
@@
 Inferences that question whether other complex trait interactions are hidden in this background heterogeneity are necessary because they critically contribute to the reconstruction and understanding of diversification.
-
-%%%%%%%%%%
-% E notes:
-%
-% summarize div findings
-%     breeding system takes from ploidy
-%     but not simple SI/SC pattern
-% what is the hidden state?
-%     could be a real trait, see tip ASR, geography?
-%     but don't be circular, test in different clade
-%     could be a stats trick
-%
-% pathways
-%     likely way to lose SI
-%     likely way to become P
-%     (compare those relative rates with robertson2011)
-%     different consequences for eventual D?
-%     div effects on pathway commonness
-%     path-dep div in CP?
-%     clado
-%
-% diploidization
-%     some model support
-%     but still SC
-% paleopolyploidy
-%     if before root, no SI?

### 2019-07-12 manuscript/discussion.tex
shortened the final section of Discussion

@@
 
-\bigskip
+\section{Conclusion}
 
-%B: Rosana, read this carefully, please! I think it may need review:
-%B: I changed nothing here, except: (1) a citealt<-citet to avoid double parentheses.
-(2) sp.
-disscussed -> discussed, (3) "If that is the case ON a system" -> "...
-IN a system"?
-(4) "hight type I errors" -> high, type I errors [comma], (5) line break in penultimate sentence.
-Estimating rates of trait linked diversification models is not only a problem of difficult parametric inference (as discussed in \citealt{rabosky_2010, beaulieu_2015}), but also a problem of inadequate SSE model specification that can result in misleading inferences when the presence of a second trait has a complex interaction with the focal trait first used to model diversification.
-We presented a roadmap with a series of analyses that can identify whether unknown or unobservable traits are worth pursuing. 
-In this roadmap, we have proposed fitting HiSSE models that have effectively shown that some binary states associated with diversification under BiSSE analyses are actually not different in net diversification terms, and instead a hidden state is linked to differences in the diversification rate \citep{beaulieu_2016}.
-If that is the case in a study system, raising the question of the identity of the hidden state should be considered as the next step in the modeling process.
-% E: I would say not necessarily.
-There could be other non-trait weirdnesses.
-TODO: include a more comprehensive discussion of how to interpret hisse/hidden state
-When considering a second candidate trait, one can model the complex interactions via a MuSSE model, especially in systems where multiple traits are suspected to influence diversification, and where interactions between those traits are well understood.
-Since MuSSE can also falsely associate traits with diversification \citep{fitzjohn_2012}, considering more heterogeneity using a hidden state model on top of the multi-state diversification model is necessary to adequately infer the effect of the complex interactions in the speciation and extinction process.
-% R- new lines
-Investigating the process of diversification linked to traits can be done via thorough statistical inferences that carefully evaluate multiple evolutionary histories. 
-Background heterogeneity in lineage of diversification is the rule, rather than the exception. 
-Inferences that question whether other complex trait interactions are hidden in this background heterogeneity are necessary because they critically contribute to the reconstruction and understanding of diversification.
+% E: The text previously here was great, but it read to me more like an abstract or methods/results.
+I tried to make it more like a big-picture summary.
+What other general points do we want to hit here?
+
+Heterogeneity in lineage diversification across time and clades is the rule, rather than the exception.
+This background heterogeneity makes it difficult to test for the association of any one, isolated trait with different rates of speciation or extinction.
+Our study provides an example of how diversification linked to a particular trait can be better assessed by a suite of more inclusive models that allow for alternative explanations---whether other traits or unknown factors.
+Additionally, our analysis of evolutionary pathways to polyploidy shows the importance of including diversification effects even when addressing questions that focus on trait evolution.

### 2019-07-12 manuscript/main.tex
- backups of all the changes -- mainly to discussion
New Phyt style for lit cited
misc small

@@
 
-% New comparative approaches for the study of breeding system, polyploidy, and lineage diversification % E: the tentative title given to New Phytologist
-%
-% Interacting effects of ploidy and breeding system on lineage diversification % E: but "effects" is too causal
+% Original title given to New Phytologist:
+% {New comparative approaches for the study of breeding system, polyploidy, and lineage diversification} 
 %
@@
 
+%B: added three! can have eight
 \noindent \textit{Keywords:} 
 Polyploidy,
-Breeding System,
-Diversification, SSE models
+Breeding Systems,
+Self-incompatibility,
+Diversification, 
+Ploidy,
+SSE models,
+Diploidization.
 
@@
 
-This work was supported by the National Science Foundation (NSF DEB-1655478 to EEG and NSF DEB-1655692 to BI).
+This work was supported by the National Science Foundation (NSF DEB-1655478 to EEG and NSF DEB-1655692 to BI) and the Binational Science Foundation (BSF-2016550 to IM).
 The computing resources were provided by the Minnesota Supercomputing Institute (MSI) at the University of Minnesota.
@@
 \bibliography{refs}
-\bibliographystyle{evolution} % E: remove URLs
+\bibliographystyle{newphytologist}
 \setstretch{\stretchby}

### 2019-07-12 manuscript/results.tex
CID pathway results
qualitatively the same as trans-only under SDD model
major editing of results

@@
 
-\subsection{Traits and diversification}
+\subsection{Models}
+\subsubsection{Ploidy and diversification}
+When considering ploidy alone (M1. $D/P$ model), we found a greater net diversification rate for diploids than for polyploids, in agreement with \citet{mayrose_2011, mayrose_2015}.
+This result holds with (\cref{figure:netdivall}A) or without ( M6 \cref{suppfigure:alldip}A) the diploidization parameter.
+Incorporating a hidden state in this model, however, reduces the clear separation in diversification rate estimates between diploids and polyploids (M4.
+$D/P+A/B$ asym model; \cref{figure:netdivall}B, and M9 \cref{figure:alldip}B).
+Thus, differences in net diversification rates are explained by an unknown factor as well as by ploidy.
+This specific result is also supported by the Bayes factors shown in table \cref{table:bayesfactors} where both model M4 and the character independent model M2 have equal evidence of  being preferred over every other model of ploidy.
+See table \cref{supptable:M6M7} showing analogous results when diploidzation is accounted for.
 
-When analyzed separately, ploidy and breeding system each showed significant associations with net diversification differences in Solanaceae.
-When analyzed together, however, the effect of breeding system dominated the effect of ploidy, although hidden factors played an important role as well.
+\subsubsection{Breeding system and diversification}
+When considering breeding system alone (M11. $I/C$ model, \cref{figure:netdivall}C), we found a larger net diversification rate for SI than for SC species, in agreement with \citet{goldberg_2010}.
+When a hidden state is included (M14.
+$I/C+A/B$ model), the large net diversification rate difference persists for hidden state $A$ but and it slightly overlaps in the  hidden state $B$ (\cref{figure:netdivall}D).
+When calculating Bayes factors for all breeding system models (M11-M15) we found that the best two supported models are the models with hidden states and asymmetric transition rates (M14 and M15 \cref{table:M11M15})
+Thus, differences in net diversification rates are best explained by both breeding system and an unknown factor \citep[as in Onagraceae;][]{freyman_2019} and not only by the focal trait nor an unknown factor.
+% The transition rate from SI to SC is $q_{IC}=0.3$ for both the I/C and the I/C+A/B models.
 
-When considering ploidy alone ($D/P$ model), we found a greater net diversification rate for diploids than for polyploids, in agreement with \citet{mayrose_2011, mayrose_2015}.
-This result holds with (\cref{figure:netdivall}A) or without (\cref{figure:netdivnodip}A) the diploidization parameter, although including diploidization shifts the net diversification rate of polyploids to be non-negative.
-Incorporating a hidden state in this model, however, reduces the clear separation in diversification rate estimates between diploids and polyploids ($D/P+A/B$ model; \cref{figure:netdivall}B, \cref{figure:netdivnodip}B).
-Thus, differences in net diversification rates are explained by an unknown factor as well as by ploidy.
-% E: I edited the interpretation of D/P+A/B.  There still is some separation between D/P within each A/B.  The A/B separation isn't any clearer, at least.
-Statistical model comparisons show very strong support for including the hidden state and strong support for including diploidization (\cref{table:bayesfactors}).
+\subsubsection{Ploidy, breeding system, and diversification}
+When considering ploidy and breeding system together (M16.
+$ID/CD/CP$ model), the net diversification rate for SI diploids was greater than for either SC diploids or SC polyploids, with or without diploidization (M16 \cref{figure:netdivall}E, M21\cref{suppfigure:dipall}E).
+Thus, the difference in net diversification associated with breeding system persists even when ploidy is included in the model.
+The reverse is not true, though: the association of ploidy with net diversification in the M1.
+$D/P$ model (\cref{figure:netdivall}A, \cref{suppfigure:dipall}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids not as distinct.
+%
+When a hidden state is included (M19.
+$IC/CD/CP+A/B$ model) the net diversification rate separation amongst states ID and (CD, CP)  is clear under hidden state A, but under hidden state B the separation exist between states ID and CP while CD overlaps with the former two (\cref{figure:netdivall}F).
+The same general pattern remains when diploidization is included (\cref{suppfigure:alldip}F), but the hidden states are reversed.
+Bayes factors for the three state models that include the character independent model show very strong support preferring asymmetric hidden state with the focal traits models shown in table \cref{supptable:M16M20} (with diploidization \cref{supptable:M21M25}).
 
-When considering breeding system alone ($I/C$ model, \cref{figure:netdivall}C), we found a larger net diversification rate for SI than for SC species, in agreement with \citet{goldberg_2010}.
-When a hidden state is included ($I/C+A/B$ model), the large net diversification rate difference persists for one of the hidden states but is mostly removed for the other (\cref{figure:netdivall}D).
-Thus, differences in net diversification rates are best explained by both breeding system and an unknown factor \citep[as in Onagraceae;][]{freyman_2019}.
-% The transition rate from SI to SC is $q_{IC}=0.3$ for both the I/C and the I/C+A/B models.
-The statistical model comparison shows very strong support for including the hidden state (\cref{table:bayesfactors}).
+\subsubsection{Lumped models}
+When comparing model M26.
+Lumped $D/P$ against M16.
+$ID/CD/CP$ (\cref{figure:lumped}B and C) Bayes factors reveal moderate support ($K=4.1$) preferring the three state model instead of the two state model(\cref{table:lumped}).
+A similar result was obtained when comparing the lumped model for ploidy but in the presence of hidden states (model M27 \textit{vs.} M23 \cref{table:lumped},\cref{suppfigure:lumpedDP}E and F).
 
-When considering ploidy and breeding system together ($ID/CD/CP$ model), the net diversification rate for SI diploids was greater than for either SC diploids or SC polyploids, with or without diploidization (\cref{figure:netdivall}E, \cref{figure:netdivnodip}C).
-Thus, the difference in net diversification associated with breeding system persists even when ploidy is included in the model.
-The reverse is not true, though: the association of ploidy with net diversification in the $D/P$ model (\cref{figure:netdivall}A, \cref{figure:netdivnodip}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids not as distinct.
-%
-When a hidden state is included ($IC/CD/CP+A/B$ model), the same general pattern remains when diploidization is prevented (\cref{figure:netdivnodip}D), although the higher net diversification rate of $ID$ is less clear within one of the hidden states.
-With diploidization, the net diversification rate of $ID$ is still greater than $CD$ within each hidden state, but diversification estimate for $CP$ is highly uncertain and perhaps bimodal.
-Statistical model comparisons show very strong support for including the hidden state and at most positive support for including diploidization (\cref{table:bayesfactors}).
+We also compared lumped models for breeding system to test for the inclusion of ploidy state (M28\cref{figure:lumped}E \textit{vs.} M16\cref{figure:lumped}C) and we found no evidence preferring the three-state over the two state-model ($K=-0.6$, \cref{table:lumped}).
+However, when accounting for hidden state,  the three-state model MM18.
+$I/C/CD/CP+A/B$ (\cref{suppigure:lumpedIC}F) was moderately preferred ($K=2.6$)over the lumped model M29.(\cref{suppfigure:lumpedIC}E).
 
 \subsection{Pathways to polyploidy}
-
 There are two pathways by which SI diploid lineages eventually---given enough time---become SC polyploids.
@@
 That is, an $ID$ lineage is more likely to begin its path to polyploidy with a transition to $CD$, but completing this path to $CP$ takes longer.
-
-% also use transition rates from transition-only (no diversification) analysis? todo
+(The results are qualitatively the same when using transition rate estimates from the model that does not allow diversification differences related to the observed states, M17).
 
@@
 
-\subsection{Diploidization as an exploratory hypothesis}
+\subsection{Model selection key questions}
+\subsubsection{Are diversification patterns only determined by hidden states?}
+Whether the determination of the diversification is dependent on the focal traits and not a hidden state only was done via the comparisons of character independent models against the focal traits  models, or against the focal traits with hidden states (results shown in tables \cref{table: bayesfactors},\cref{table:M11M15},\cref{supptable:M6M10},\cref{supptable:M16M20},\cref{supptable:M21M25}).
+The general pattern is to moderately to strongly prefer models with focal traits and asymetric hidden states over the rest of the models with the exception of the ploidy only model where the character independent model (M2) seems to be equally preferred($K=0.5$) as the model with ploidy and asymmetric hidden rates (M4, \cref{table:revbayes}).
 
+\subsubsection{Are hidden states necessary?}
+In the comparisons of simple diversification models of focal traits against models with hidden states we found that models with asymmetric hidden traits are strongly preferred over simpler models (summarized in table \cref{supptable:testaddhidden}).
+
+\subsubsection{Is a second trait adding information to the diversification process?}
+As summarized in table \cref{table:lumped} it is moderately preferred to add breeding system to ploidy only models (cref{figure:lumpedDP}).
+When adding ploidy information to breeding system models to create a three-state model, we found that there is no evidence that it is preferred (M28 vs. M16 \cref{figure:lumpedIC}B and C). 
+However, it becomes moderately preferable in the presence of hidden states (M29 vs. M23  \cref{figure:lumpedIC}E and C)
+
+\subsubsection{Is there evidence for diploidization?}
 We considered models both with and without diploidization in order to explore its effects on the estimates of state-dependent diversification.
-For the two models that only include diploid and polyploid states ($D/P$ and $D/P+A/B$), the net diversification rate of polyploid lineages is likely negative when diploidization (parameter $\delta$) is excluded but positive when it is included (\cref{figure:netdivall}A versus \cref{figure:netdivnodip}A).
-For the two models that also included breeding system ($ID/CD/CP$ and $ID/CD/CP+A/B$), the main effect of including diploidization is greatly increasing the uncertainty of the estimated polyploid net diversification rate (\cref{figure:netdivall}EF versus \cref{figure:netdivnodip}CD).
-For all models, there is greater uncertainty in the estimate of diploidization rate than polyploidization rate, as judged by the width of the credibility intervals (see \cref{suppfigure:DP,suppfigure:DPAB,suppfigure:IDCDCP,suppfigure:IDCDCPAB}).
-% R will add a bit about param confidence
+For the models that only include diploid and polyploid states (M6-M10) the results are similar in terms of the effect of diversification (\cref{figure:dipall}).
+When comparing models with diploidization against models without diploidization we found moderate evidence that models containing the parameter $\delta$ are preferred over models without diploidization (summarized in table \cref{supptable:testdiploidization})
+
+\subsubsection{Do asymmetric rates in hidden states models affect the pattern of diversification?}
+When comparing different types of asymmetry across hidden states models we found that the asymmetry did not change the direction of effects in net diversification results as shown in \cref{suppfigure:asymmetric}.
+However, models with asymmetric hidden rates are strongly preferred over models with equal rates between hidden states (summarized in \cref{supptable:asymmetry}).
 
@@
 \label{supptable:M21M25}
+\end{supptable}
+
+\begin{supptable}
+\addtolength{\tabcolsep}{-5pt}
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
+M1.
+D/P                                  & -1238.76                                                          & M1 vs.
+M4   & 60.47           & M4 (Strong)                                                          \\
+\textbf{M4.
+D/P+A/B asym}                & \textbf{-1223.28}                                                &             &                 &                                                                      \\ \hline
+M11.
+I/C                                 & -1309.07                                                          & M11 vs.
+M14 & 61.35           & M14 (Strong)                                                         \\
+\textbf{M14.
+I/C+A/B asym}               & \textbf{-1247.72}                                                 &             &                 &                                                                      \\ \hline
+M16.
+ID/CD/CP                            & -1459.11                                                          & M16 vs.
+M19 & 65.90           & M19 (Strong)                                                         \\
+\textbf{M19.
+ID/CD/CP+A/B asym}          & \textbf{-1393.20}                                                &             &                 &                                                                      \\ \hline
+M6.
+D/P+$\delta$                         & -1283.76                                                          & M6 vs.
+M9   & 69.3            & M9 (Strong)                                                          \\
+\textbf{M9.
+D/P+A/B+$\delta$ asym}       & \textbf{-1214.46}                                                 &             &                 &                                                                      \\ \hline
+M21.
+IC/CD/CP+$\delta$                   & -1454.68                                                          & M21 vs.
+M24 & 68.48           & M24 (Strong)                                                         \\
+\textbf{M24.
+IC/CD/CP+A/B+$\delta$ asym} & \textbf{-1386.20}                                                 &             &                 &                                                                      \\ \hline
+\end{tabular}
+\caption{Test for addition of hidden states in models via Bayes factors. Models with hidden states are strongly preferred over simpler models}
+\label{table:testaddhidden}
+\end{supptable}
+
+\begin{supptable}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
+M1.
+D/P                                  & -1238.76                                                          & M1 vs.
+M6   & 65.92           & M6 (Strong)                                                          \\
+\textbf{M6.
+D/P+$\delta$}                & \textbf{-1267.84}                                                 &             &                 &                                                                      \\ \hline
+M4.
+D/P+A/B asym                         & -1223.28                                                          & M4 vs.
+M9   & 8.81            & M9 (Moderate)                                                        \\
+\textbf{M9.
+D/P+A/B+$\delta$ asym}       & \textbf{-1214.46}                                                 &             &                 &                                                                      \\ \hline
+M16.
+ID/CD/CP                            & -1459.11                                                          & M16 vs.
+M21 & 4.41            & M21 (Moderate)                                                       \\
+\textbf{M21.
+ID/CD/CP+$\delta$}          & \textbf{-1454.68}                                                 &             &                 &                                                                      \\ \hline
+M19.
+IC/CD/CP asym                       & -1393.20                                                          & M19 vs.
+M24 & 6.46            & M24 (Moderate)                                                       \\
+\textbf{M24.
+IC/CD/CP+A/B+$\delta$ asym} & \textbf{-1386.20}                                                 &             &                 &                                                                      \\ \hline
+\end{tabular}
+\caption{Test for inclusion of a diploidization rate via Bayes factors. Models with diploidzation are moderately preferred over models that do not include a diploidization rate}
+\label{supptable:testdiploidization}
 \end{supptable}

### 2019-07-12 manuscript/results.tex
minor editing to the resuls (referrences fixed)

@@
 \subsubsection{Ploidy and diversification}
-When considering ploidy alone (M1. $D/P$ model), we found a greater net diversification rate for diploids than for polyploids, in agreement with \citet{mayrose_2011, mayrose_2015}.
-This result holds with (\cref{figure:netdivall}A) or without ( M6 \cref{suppfigure:alldip}A) the diploidization parameter.
+When considering ploidy alone (M1. $D/P$ model), we found a larger net diversification rate for diploids than for polyploids, in agreement with \citet{mayrose_2011, mayrose_2015}.
+This result holds with (\cref{figure:netdivall}A) or with the diploidization parameter (model M6 \cref{suppfigure:alldip}A) .
 Incorporating a hidden state in this model, however, reduces the clear separation in diversification rate estimates between diploids and polyploids (M4.
-$D/P+A/B$ asym model; \cref{figure:netdivall}B, and M9 \cref{figure:alldip}B).
-Thus, differences in net diversification rates are explained by an unknown factor as well as by ploidy.
-This specific result is also supported by the Bayes factors shown in table \cref{table:bayesfactors} where both model M4 and the character independent model M2 have equal evidence of  being preferred over every other model of ploidy.
-See table \cref{supptable:M6M7} showing analogous results when diploidzation is accounted for.
+$D/P+A/B$ asym model; \cref{figure:netdivall}B, and M9 \cref{suppfigure:alldip}B).
+Thus, differences in net diversification rates are explained by an unknown factor.
+The lack of differences between diploid and polyploid net diversification rates is also supported by the Bayes factors shown in table \cref{table:bayesfactors} where both model M4 and the character independent model M2 have equal evidence of  being preferred over every other model of ploidy.
+See table \cref{supptable:M6M10} showing analogous results when diploidzation is accounted for.
 
@@
 $I/C+A/B$ model), the large net diversification rate difference persists for hidden state $A$ but and it slightly overlaps in the  hidden state $B$ (\cref{figure:netdivall}D).
-When calculating Bayes factors for all breeding system models (M11-M15) we found that the best two supported models are the models with hidden states and asymmetric transition rates (M14 and M15 \cref{table:M11M15})
+When calculating Bayes factors for all breeding system models (M11-M15) we found that the best two supported models are the models with at least asymmetric hidden states (M14 and M15 \cref{table:M11M15})
 Thus, differences in net diversification rates are best explained by both breeding system and an unknown factor \citep[as in Onagraceae;][]{freyman_2019} and not only by the focal trait nor an unknown factor.
@@
 When considering ploidy and breeding system together (M16.
-$ID/CD/CP$ model), the net diversification rate for SI diploids was greater than for either SC diploids or SC polyploids, with or without diploidization (M16 \cref{figure:netdivall}E, M21\cref{suppfigure:dipall}E).
+$ID/CD/CP$ model), the net diversification rate for SI diploids was greater than for either SC diploids or SC polyploids, with or without diploidization (M16 \cref{figure:netdivall}E, M21\cref{suppfigure:alldip}E).
 Thus, the difference in net diversification associated with breeding system persists even when ploidy is included in the model.
-The reverse is not true, though: the association of ploidy with net diversification in the M1.
-$D/P$ model (\cref{figure:netdivall}A, \cref{suppfigure:dipall}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids not as distinct.
-%
+The reverse is not true, the association of ploidy with net diversification in the M1.
+$D/P$ model (\cref{figure:netdivall}A, \cref{suppfigure:alldip}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids is not distinct.
+
 When a hidden state is included (M19.
-$IC/CD/CP+A/B$ model) the net diversification rate separation amongst states ID and (CD, CP)  is clear under hidden state A, but under hidden state B the separation exist between states ID and CP while CD overlaps with the former two (\cref{figure:netdivall}F).
-The same general pattern remains when diploidization is included (\cref{suppfigure:alldip}F), but the hidden states are reversed.
-Bayes factors for the three state models that include the character independent model show very strong support preferring asymmetric hidden state with the focal traits models shown in table \cref{supptable:M16M20} (with diploidization \cref{supptable:M21M25}).
+$IC/CD/CP+A/B$ model) the net diversification rate separation amongst states ID and (CD, CP)  is clear under hidden state A, but under hidden state B the separation exist between states ID and CP, while CD overlaps with the other two states (\cref{figure:netdivall}F).
+The same general pattern remains when diploidization is included (\cref{suppfigure:alldip}F), but the hidden states are reversed in the x-axis.
+Bayes factors for the three state models that include the character independent model show very strong support preferring asymmetric hidden state with the focal traits models shown in  \cref{supptable:M16M20} (with diploidization \cref{supptable:M21M25}).
 
@@
 Lumped $D/P$ against M16.
-$ID/CD/CP$ (\cref{figure:lumped}B and C) Bayes factors reveal moderate support ($K=4.1$) preferring the three state model instead of the two state model(\cref{table:lumped}).
-A similar result was obtained when comparing the lumped model for ploidy but in the presence of hidden states (model M27 \textit{vs.} M23 \cref{table:lumped},\cref{suppfigure:lumpedDP}E and F).
+$ID/CD/CP$ (\cref{figure:lumped}B and C) Bayes factors reveal moderate support ($K=4.1$) preferring the three state model instead of the two state model (\cref{table:lumped}).
+A similar result was obtained when comparing the lumped model for ploidy but in the presence of hidden states (model M27 \textit{vs.} M23 \cref{table:lumped}, \cref{suppfigure:lumpedDP}E and F).
 
-We also compared lumped models for breeding system to test for the inclusion of ploidy state (M28\cref{figure:lumped}E \textit{vs.} M16\cref{figure:lumped}C) and we found no evidence preferring the three-state over the two state-model ($K=-0.6$, \cref{table:lumped}).
-However, when accounting for hidden state,  the three-state model MM18.
-$I/C/CD/CP+A/B$ (\cref{suppigure:lumpedIC}F) was moderately preferred ($K=2.6$)over the lumped model M29.(\cref{suppfigure:lumpedIC}E).
+We also compared lumped models for breeding system to test for the inclusion of ploidy state (M28 \cref{figure:lumped}E \textit{vs.} M16 \cref{figure:lumped}C) and we found no evidence preferring the three-state over the two state-model ($K=-0.6$, \cref{table:lumped}).
+However, when accounting for hidden state,  the three-state model M18.
+$I/C/CD/CP+A/B$ (\cref{suppfigure:lumpedIC}F) was moderately preferred ($K=2.6$) over the lumped model M29.(\cref{suppfigure:lumpedIC}E).
 
@@
 \subsubsection{Are diversification patterns only determined by hidden states?}
-Whether the determination of the diversification is dependent on the focal traits and not a hidden state only was done via the comparisons of character independent models against the focal traits  models, or against the focal traits with hidden states (results shown in tables \cref{table: bayesfactors},\cref{table:M11M15},\cref{supptable:M6M10},\cref{supptable:M16M20},\cref{supptable:M21M25}).
-The general pattern is to moderately to strongly prefer models with focal traits and asymetric hidden states over the rest of the models with the exception of the ploidy only model where the character independent model (M2) seems to be equally preferred($K=0.5$) as the model with ploidy and asymmetric hidden rates (M4, \cref{table:revbayes}).
+Whether the determination of the diversification is dependent on the focal traits and not only on the hidden state was done via the comparisons of character independent models against the focal traits w/o hidden states models (results shown in  \cref{table:bayesfactors},\cref{table:M11M15},\cref{supptable:M6M10},\cref{supptable:M16M20},\cref{supptable:M21M25}).
+The general pattern is to moderate- to strongly prefer models with focal traits and asymmetric hidden states over the rest of the models with the exception of the ploidy only model where the character independent model (M2) seems to be equally preferred($K=0.5$) as the model with ploidy and asymmetric hidden rates (M4, \cref{table:bayesfactors}).
 
@@
 \subsubsection{Is a second trait adding information to the diversification process?}
-As summarized in table \cref{table:lumped} it is moderately preferred to add breeding system to ploidy only models (cref{figure:lumpedDP}).
-When adding ploidy information to breeding system models to create a three-state model, we found that there is no evidence that it is preferred (M28 vs. M16 \cref{figure:lumpedIC}B and C). 
-However, it becomes moderately preferable in the presence of hidden states (M29 vs. M23  \cref{figure:lumpedIC}E and C)
+As summarized in  \cref{table:lumped}, it is moderately preferred to add breeding system to ploidy only models (\cref{suppfigure:lumpedDP}).
+When adding ploidy information to breeding system models to create a three-state model, we found that there is no evidence that it is preferred (M28 vs. M16 \cref{suppfigure:lumpedIC}B and C). 
+However, it becomes moderately preferable in the presence of hidden states (M29 vs. M23  \cref{suppfigure:lumpedIC}E and C).
 
@@
 We considered models both with and without diploidization in order to explore its effects on the estimates of state-dependent diversification.
-For the models that only include diploid and polyploid states (M6-M10) the results are similar in terms of the effect of diversification (\cref{figure:dipall}).
-When comparing models with diploidization against models without diploidization we found moderate evidence that models containing the parameter $\delta$ are preferred over models without diploidization (summarized in table \cref{supptable:testdiploidization})
+For the models that only include diploid and polyploid states (M6-M10) the results are similar in terms of the effect of diversification (\cref{figure:alldip}).
+When comparing models with diploidization against models without diploidization we found moderate evidence that models containing the parameter $\delta$ are preferred over models without diploidization (summarized in \cref{supptable:testdiploidization}).
 
 \subsubsection{Do asymmetric rates in hidden states models affect the pattern of diversification?}
-When comparing different types of asymmetry across hidden states models we found that the asymmetry did not change the direction of effects in net diversification results as shown in \cref{suppfigure:asymmetric}.
-However, models with asymmetric hidden rates are strongly preferred over models with equal rates between hidden states (summarized in \cref{supptable:asymmetry}).
+When comparing different types of asymmetry across hidden states models we found that the asymmetry did not change the direction of effects in net diversification results .
+However, models with asymmetric hidden rates reduced the overlap amongst net diversification posterior distributions as shown in \cref{suppfigure:asymmetric}.
+Furthermore, hidden state models with asymmetric traits are strongly preferred over models with equal rates between hidden states (summarized in \cref{supptable:asymmetry}).
 
@@
 \caption{Test for addition of hidden states in models via Bayes factors. Models with hidden states are strongly preferred over simpler models}
-\label{table:testaddhidden}
+\label{supptable:testaddhidden}
 \end{supptable}

### 2019-07-13 manuscript/methods.tex
misc small
minor editing to the resuls (referrences fixed)
all edits from RZF in Methods and Results

@@
 
-In order to investigate the association between ploidy level and diversification, we first defined a model in which taxa were classified as diploid $(D)$ or polyploid $(P)$ (M1).
-This is the same as one of the models presented in  \citet{mayrose_2011} that was used for the exact same question across an array of plant genera.
+In order to investigate the association between ploidy level and diversification, we first defined a model in which taxa were classified as diploid $(D)$ or polyploid $(P)$ (model M1).
+This is the same as one of the models presented in  \citet{mayrose_2011}, which was used for the exact same question across an array of plant genera.
 % R- Itay wanted to hightlight that is the same model. Definitely a BiSSE. He did the ClaSSE later in the 2014 response.
@@
 These models differ in whether transitions between the hidden states are symmetric or asymmetric, and whether the polyploidization rate depends on the hidden state.
-Comparing M1--M5 allows us to test whether ploidy is associated with diversification differences on top of the differences potentially explained by an unknown factor\cref{table:bayesfactors}.
+Comparing M1--M5 allows us to test whether ploidy is associated with diversification differences on top of the differences potentially explained by an unknown factor (\cref{table:bayesfactors}).
 
-We further fit the analogues of these five models but including a rate parameter for transitions from polyploid to diploid (M6--M10) \cref{supptable:M6M10}.
-This allows us to assess whether our conclusions about ploidy and diversification are robust to the possibility of diploidization.
+We further fit the analogues of these five models but including a rate parameter $\delta$ for transitions from polyploid to diploid (M6--M10,  \cref{supptable:M6M10}).
+These comparisons allow us to assess whether our conclusions about ploidy and diversification are robust to the possibility of diploidization.
 
@@
 
-Our first five breeding system models follow the same logic as the first ploidy models.
-We begin with each species classified as self-incompatible $(I)$ or self-compatible $(C)$ (M11).
-This is the same as the analysis of \citet{goldberg_2010} but with an updated phylogeny \citep{sarkinen_2013} and larger aggregated dataset.
+We propose five breeding system models following the same logic as the first ploidy models.
+Under the simplest breeding system and diversification model,  species are classified as self-incompatible $(I)$ or self-compatible $(C)$ (M11).
+This is the same model as the analysis presented in \citet{goldberg_2010} but with an updated phylogeny \citep{sarkinen_2013} and larger aggregated dataset.
 
@@
 Furthermore, species that are distantly related within this family carry closely-related alleles, with deep trans-specific polymorphism at the locus that controls the SI response \citep{ioerger_1990, igic_2006}.
-This represents very strong evidence that the $I$ state is ancestral in Solanaceae, and that the SI mechanism was not regained within this family.
+The breeding system in Solanaceae has strong evidence that the $I$ state is ancestral in the family, and that the SI mechanism was not regained.
 Therefore, for all breeding system models, we estimated a transition rate from $I$ to $C$  but not the reverse ($q_{CI}=0$).
@@
 
-The simplest classifies each species as either self-incompatible diploid $(ID)$, self-compatible diploid $(CD)$, or self-compatible polyploid $(CP)$; recall that self-incompatible polyploids do not occur (M16).
+The simplest model classifies each species as either self-incompatible diploid $(ID)$, self-compatible diploid $(CD)$, or self-compatible polyploid $(CP)$; recall that self-incompatible polyploids do not occur (M16).
 Each of these states may again be associated with different rates of speciation and extinction, and the allowable transitions are loss of SI within the diploid state (from $ID$ to $CD$), loss of SI via polyploidization (from $ID$ to $CP$), and polyploidization while SC (from $CD$ to $CP$).
@@
 
-The models above allow us to assess the contributions of our two focal traits---ploidy and breeding system---to lineage diversification.
-They do not, however, reveal whether it was ``worth it'' to include both traits in the analysis.
-That is, they do not allow us to perform statistical model comparisons to test whether diversification heterogeneity is better explain by both traits than by either one of them.
-This is because the models in the three sections above use different state spaces: $D/P$ versus $I/C$ versus $IC/CD/CP$.
-Model comparisons work with the likelihood of the data given the model, but they are not valid when the data used differ from model to model.
+The models described so far allow us to assess the contributions of our two focal traits---ploidy and breeding system---to lineage diversification.
+They do not, however, reveal whether it is `worth it' to include both traits in the analysis.
+Model comparisons to test whether diversification heterogeneity is better explain by both traits than by either one require the calculation of the likelihood function of the data given an specific model.
+However, the likelihood of the data given a model differs in the two-state versus the three-state models because the proposed models use different state spaces: $D/P$ versus $I/C$ versus $IC/CD/CP$.
+Therefore, different data results in incomparable models.
 
-In order to compare model fits for ploidy-only versus breeding system-only versus both traits, we use the technique of \citet{tarasov_2019} for lumping states together.
+In order to compare model fits for ploidy-only versus breeding system-only versus both traits, we extend a new technique proposed in ontological work by \citet{tarasov_2019} that allows for state `lumping' .
 We use the state space of the $IC/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
@@
 We do not include additional models with diploidization because this reverse transition renders the models non-lumpable \citep{tarasov_2019}.
-The lumpability property requires that the rates going from the lumped state to the singular state are equal.
-Under the diploidization models CP goes back to CD with rate $\delta$ but the rate from CP back to ID is 0, and $\delta \neq 0$.
-Therefore, two-state models with diplodization cannot be re-coded to be compared to the tree-state models.
+This means, that to lump two states a necessary and sufficient condition is that the rates going from the lumped state to the singular state are equal. 
+When lumping compatible state $CD$ and $CP$ the the rate at which they back to state $ID$ is zero for both, thus, compatible states are lumpable. 
+Then, we define a single $q_0$ parameter as the new rate from incompatible to compatible going into the lumped state $(CD,CP)$.
+By doing so, we can define the lumped model M28 (\cref{figure:lumped}E) that is equivalent to M1 but comparable to M16.
+We repeat this process to create the lumped model M26for $D/P$ by restricting the polyploidy rate to be $rho_0$ out of the lumped state $(ID,CD)$
+For the diploidization models, $CP$ goes back to $CD$ with rate $\delta$ but the rate from $CP$ back to $ID$ is 0.
+Since $\delta \neq 0$, the models with diplodization cannot be lumped to be compared to the tree-state models.
 % E: further explanation needed? R- I added a simple math explanation. Let me know if it is okay
@@
 
-We report posterior distributions for the model parameters, and also for the compound parameters of net diversification ($\lambda - \mu$) and extinction fraction ($\mu / \lambda$) for each state.
-Additionally, ancestral states at each node in the phylogeny were sampled jointly during the MCMC analyses.
-Ancestral state reconstructions for all models show the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 594 internal nodes for each of the 10 models (\cref{suppfigure:DPnodipasr,suppfigure:DPasr,suppfigure:DPnodipABasr,suppfigure:DPABasr,suppfigure:ICasr,suppfigure:ICABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPasr,suppfigure:IDCDCPnodipABasr,suppfigure:IDCDCPABasr}).
+We report posterior distributions for the model parameters in \cref{figure:netdivall}, and also for the compound parameters of net diversification ($\lambda - \mu$) and extinction fraction ($\mu / \lambda$) for each state.
+Additionally, ancestral states at each node in the phylogeny were sampled jointly during the MCMC analyses every 100 generations.
+Ancestral state estimations for all models show the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes for each of the models in \cref{figure:netdivall}.
+(\cref{suppfigure:DPnodipasr,suppfigure:DPnodipABasr,suppfigure:ICasr,suppfigure:ICABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPnodipABasr}).
 % FIXME: I'm confused about if these are joint or marginal?
 A bit more explanation would help.
+R: Marginal explanation added
 
-\subsubsection{Model selection}
+\subsubsection{Model selection key questions}
 
@@
 \begin{enumerate}
-\item Are hidden state necessary to add in this context?- Comparison of simple models against hidden state models
-\item Is a second trait adding information to the diversification process?- Comparison of lumped models against IC/CD/CP models
-\item Are diversification patterns only determined by hidden states and not the traits of interest?- Comparison of character independent models against hidden state 
-\item Is there evidence for diploidization?- Comparison amongst models with and without diploidization
-\item Do asymmetrical rates in the model affect the pattern of diversification? - Comparison amongst hidden states models with equal hidden rates and asymmetrical rates.
+\item Are diversification patterns only determined by hidden states and not the traits of interest?- Comparison of character independent models against hidden state (\cref{table:bayesfactors,table:M11M15,supptable:M6M10,supptable:M16M20,supptable:M21M25}).
+\item Are hidden state necessary to add in this context?- Comparison of simple models against hidden state models (\cref{supptable:testaddhidden}).
+\item Is a second trait adding information to the diversification process?- Comparison of lumped models against IC/CD/CP models (\cref{figure:lumped,table:lumped}).
+\item Is there evidence for diploidization?- Comparison amongst models with and without diploidization (\cref{suppfigure:alldip,supptable:testdiploidization}).
+\item Do asymmetrical rates in the model affect the pattern of diversification?
+- Comparison amongst hidden states models with equal hidden rates and asymmetrical rates (\cref{suppfigure:asymmetric,supptable:asymmetry}).
 \end{enumerate}
@@
 }
-\label{table: bayesfactors}
+\label{table:bayesfactors}
 \end{table}
@@
 \end{table}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{alldiversificationasymhidden.pdf} 
+\caption{Net diversification rates for all models without diploidization rate estimate. 
+Each panel contains a graphical summary of estimated model parameters and displays posterior distributions for net diversification estimates.}
+\label{figure:netdivall}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{lumped.pdf} %lumped.pdf
+\caption{Comparing two and three state models is possible via lumpable models (M26, M28).
+These models used the same classifications as the three-state model M16 which allow for comparisons via Bayes Factors as shown in \cref{table:lumped}.}
+\label{figure:lumped}
+\end{figure}
+
+% E todo: re-run with new rate estimates; add panel letters?
+\begin{figure}
+    \centering 
+    \includegraphics[width=0.5\textwidth]{pathstates2} \\ [40pt] 
+        \includegraphics[width=0.9\textwidth]{pathways}
+    \caption{
+        Contributions of the two pathways to polyploidy.
+        The one-step pathway is direct ID$\rightarrow$CP transitions.
+        The two-step pathway consists of ID$\rightarrow$CD$\rightarrow$CP transitions.
+        When considering only rates of transitions among the states (ignoring the diversification rate parameters), the one-step pathway dominates on short timescales and the two-step on long timescales (left panels).
+        When also considering diversification within each state, the one-step pathway, in which polyploidization breaks down SI, dominates over any timescale (right panels).
+        The top panels show the separate contributions of each pathway.
+        The bottom panels show the proportional contribution of the one-step pathway (\ie one-step / [one-step + two-step]).
+    }
+    \label{figure:pathways}
+\end{figure}
+
+\begin{supptable}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|r|c|c|c|c|l|}
+\hline
+Model                           & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M7    & M8     & M9    & M10   & Evidence                                                                                     \\ \hline
+M6.
+D/P+$\delta$                & -1268.83                                                          & 55.41 & 42.78  & 53.37 & 52.85 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M6\end{tabular}            \\ \hline
+\textbf{M7.
+CID D/P+$\delta$}   & \textbf{-1212.42}                                                 &       & -12.62 & -2.04 & -2.04 & \begin{tabular}[c]{@{}l@{}}Model M7 moderately\\ preferred over M9 and M10\end{tabular}      \\ \hline
+M8.
+D/P+A/B $\delta$            & -1214.46                                                          &       &        & 10.58 & 10.07 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular} \\ \hline
+M9.
+D/P+A/B +$\delta$ asym      & -1214.46                                                          &       &        &       & -0.51 & No evidence                                                                                  \\ \hline
+M10.
+D/P+A/B +$\delta$ all asym & -1214.97                                                          &       &        &       &       &         \\ \hline
+\end{tabular}
+\caption{Bayes factors of  ploidy-only models with diplodization.
+Results indicate that a character independent model (M6) is strongly preferred over BiSSE (M1)  model.
+M6 is moderately preferred over any of the HiSSE models with asymmetric hidden rates (M9, M10).}
+\label{supptable:M6M10}
+\end{supptable}
+
+\begin{supptable}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|r|c|c|c|c|l|}
+\hline
+Model                            & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M18   & M19   & M20   & Evidence                                                                                                   \\ \hline
+M16.
+ID/CP/CD                    & -1459.11                                                          & 45.11 & 65.91 & 63.98 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M16\end{tabular}                         \\ \hline
+M17.
+CID ID/CP/CD                & *                                                                 &       &       &       &                                                                                                            \\
+M18.
+ID/CP/CD+A/B                & -1414                                                             &       & 20.79 & 18.87 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}               \\ \hline
+\textbf{M19.
+ID/CP/CD +A/B asym} & \textbf{-1393.20}                                                 &       &       & -1.92 & \begin{tabular}[c]{@{}l@{}}Asymmetric hidden rates moderately\\ preferred over all asymmetric\end{tabular} \\ \hline
+M20.
+ID/CP/CD +A/B all asym      & -1393.12                                                          &       &       &       &   \\ \hline
+\end{tabular}
+\caption{Bayes factors of ploidy and breeding system withou diploidization.
+Results indicate that the MuHiSSE model with asymmetric hidden rates (17) is  strongly preferred over M16 and M18 and moderately preferred over the MuHiSSe with all rates asymetric (M20).
+*Marginal log-likelihood for M17 could not be calculated in our limited amount of computational time.}
+\label{supptable:M16M20}
+\end{supptable}
+
+\begin{supptable}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|r|c|c|c|c|l|}
+\hline
+Model                                 & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M22   & M23    & M24   & M25   & Evidence                                                                                         \\ \hline
+M21.
+ID/CP/CD+$\delta$                & -1454.68                                                          & 55.48 & 46.031 & 67.94 & 65.15 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M21\end{tabular}               \\ \hline
+M22.
+CID ID/CP/CD+$\delta$            & -1399.201                                                         &       & -9.452 & 12.45 & 9.675 & \begin{tabular}[c]{@{}l@{}}Model M24 strongly preferred \\ over M22\end{tabular}                 \\ \hline
+M23.
+ID/CP/CD+A/B $\delta$            & -1408.65                                                          &       &        & 21.91 & 19.12 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}     \\ \hline
+\textbf{M24.
+ID/CP/CD +$\delta$ asym} & \textbf{-1386.74}                                                 &       &        &       & -2.78 & \begin{tabular}[c]{@{}l@{}}Asymmetric hidden rates \\ preferred over all asymmetric\end{tabular} \\ \hline
+M25.
+ID/CP/CD +$\delta$ all asym      & -1389.52                                                          &       &        &       &       &                      \\ \hline
+\end{tabular}
+\caption{Bayes factors of ploidy and breeding system with diploidization.
+Results indicate that the MuHiSSE model with asymmetric hidden rates (M24) is  strongly preferred over M21-M23 and moderately preferred over the MuHiSSe with all rates asymetric (M25).}
+\label{supptable:M21M25}
+\end{supptable}
+
+\begin{supptable}
+\addtolength{\tabcolsep}{-5pt}
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
+M1.
+D/P                                  & -1238.76                                                          & M1 vs.
+M4   & 60.47           & M4 (Strong)                                                          \\
+\textbf{M4.
+D/P+A/B asym}                & \textbf{-1223.28}                                                &             &                 &                                                                      \\ \hline
+M11.
+I/C                                 & -1309.07                                                          & M11 vs.
+M14 & 61.35           & M14 (Strong)                                                         \\
+\textbf{M14.
+I/C+A/B asym}               & \textbf{-1247.72}                                                 &             &                 &                                                                      \\ \hline
+M16.
+ID/CD/CP                            & -1459.11                                                          & M16 vs.
+M19 & 65.90           & M19 (Strong)                                                         \\
+\textbf{M19.
+ID/CD/CP+A/B asym}          & \textbf{-1393.20}                                                &             &                 &                                                                      \\ \hline
+M6.
+D/P+$\delta$                         & -1283.76                                                          & M6 vs.
+M9   & 69.3            & M9 (Strong)                                                          \\
+\textbf{M9.
+D/P+A/B+$\delta$ asym}       & \textbf{-1214.46}                                                 &             &                 &                                                                      \\ \hline
+M21.
+IC/CD/CP+$\delta$                   & -1454.68                                                          & M21 vs.
+M24 & 68.48           & M24 (Strong)                                                         \\
+\textbf{M24.
+IC/CD/CP+A/B+$\delta$ asym} & \textbf{-1386.20}                                                 &             &                 &                                                                      \\ \hline
+\end{tabular}
+\caption{Test for addition of hidden states in models via Bayes factors. Models with hidden states are strongly preferred over simpler models}
+\label{supptable:testaddhidden}
+\end{supptable}
+
+\begin{supptable}
+\addtolength{\tabcolsep}{-3pt}
+\begin{tabular}{|l|c|c|c|c|}
+\hline
+Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
+M1.
+D/P                                  & -1238.76                                                          & M1 vs.
+M6   & 65.92           & M6 (Strong)                                                          \\
+\textbf{M6.
+D/P+$\delta$}                & \textbf{-1267.84}                                                 &             &                 &                                                                      \\ \hline
+M4.
+D/P+A/B asym                         & -1223.28                                                          & M4 vs.
+M9   & 8.81            & M9 (Moderate)                                                        \\
+\textbf{M9.
+D/P+A/B+$\delta$ asym}       & \textbf{-1214.46}                                                 &             &                 &                                                                      \\ \hline
+M16.
+ID/CD/CP                            & -1459.11                                                          & M16 vs.
+M21 & 4.41            & M21 (Moderate)                                                       \\
+\textbf{M21.
+ID/CD/CP+$\delta$}          & \textbf{-1454.68}                                                 &             &                 &                                                                      \\ \hline
+M19.
+IC/CD/CP asym                       & -1393.20                                                          & M19 vs.
+M24 & 6.46            & M24 (Moderate)                                                       \\
+\textbf{M24.
+IC/CD/CP+A/B+$\delta$ asym} & \textbf{-1386.20}                                                 &             &                 &                                                                      \\ \hline
+\end{tabular}
+\caption{Test for inclusion of a diploidization rate via Bayes factors. Models with diploidzation are moderately preferred over models that do not include a diploidization rate}
+\label{supptable:testdiploidization}
+\end{supptable}
+
+\begin{supptable}
+\addtolength{\tabcolsep}{-5pt}
+\begin{tabular}{|l|c|c|c|l|}
+\hline
+Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
+M3. D/P+A/B                                 & -1234.52                     &    M3 vs. M4                 & 11.239                  & M4 (Strong)   \\
+\textbf{M4. D/P+ A/B asym}            & -1223.28                     &   M4 vs. M5              & -1.658                  & M4 (Moderate)   \\
+M5. D/P+A/B all asym                 & -1224.93                     &                     &                         &                         \\ \hline
+M13. I/C+ A/B                                & -1270.47                     &   M13 vs. M14                  & 22.75                   & M14 (Strong)   \\
+\textbf{M14. I/C+ A/B asym}            & \textbf{-1247.72}            & M14 vs. M15         & {0.05}           & No evidence             \\
+M15. I/C+ A/B all asym               & -1247.67                     &                     &                         &                         \\ \hline
+M18. IC/CD/CP+A/B                            & -1414.00                     &     M18 vs. M19                & 20.79                   & M19 (Strong)    \\
+\textbf{M19. IC/CD/CP+ A/B asym}       & \textbf{-1393.21}            & M19 vs. M20          & {-1.919}         & M19 (Moderate)   \\
+M20. IC/CD/CP+ A/B all asym           & -1395.129                    &                     &                         &                         \\ \hline
+M8. D/P +A/B+$\delta$                          & -1225.05                     &  M8 vs. M9                   & 10.58                   & M9 (Strong)    \\
+\textbf{M9. D/P+ A/B+ $\delta$ asym}      & \textbf{-1214.46}                     &    M9 vs. M10                 & -0.52                   & M10 (Moderate) \\
+M10. D/P+A/B+$\delta$ all asym        & -1214.98                     &                    &                         &                         \\ \hline
+M23. IC/CD/DP+A/B+$\delta$                 & -1408.65                     &     M23 vs M24                & 21.91                   & M24(Strong)     \\
+\textbf{M24. IC/CD/DP+A/B+$\delta$ asym} & \textbf{-1386.74}                     &   M24 vs M25              & -2.78                   & M24 (Moderate) \\
+M25. IC/CD/DP+A/B+ $\delta$ all asym     & -1389.52                     &                     &                         &                         \\ \hline
+\end{tabular}
+\caption{Test for symmetry of the hidden and trait rates via Bayes factors.
+For all models asymmetric hidden transition rates are preferred over models with equal hidden rates.
+Adding more complexity by assuming asymmetry in the trait rate within hidden states does not improve models.}
+\label{supptable:asymmetry}
+\end{supptable}
+
+% check values of delta and rho D/P+A/B with(out) delta
+% i.e. parameter values in figure S1 & S4 vs Table S2
+
+%B: heads up: I added two clarifying descriptive sentences in the two figures below. Maybe something about noting x-axis scale differences would be useful.
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{alldiversificationasymhiddendip.pdf}
+\caption{Posterior distribution for all the best models with diploidization} % XXX
+\label{suppfigure:alldip}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{comparablesDP.pdf}
+\caption{Lumped ploidy models to assess the effect of adding breeding system to create a three-state model.
+Moderate evidence exists that adding a breeding system is necessary \cref{table:lumped} } % XXX
+\label{suppfigure:lumpedDP}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{comparablesIC.pdf}
+\caption{Lumped breeding system models to assess the effect of adding ploidy to create a three-state model.
+Moderate to insignificant evidence exists that adding ploidy is necessary \cref{table:lumped} } % XXX
+\label{suppfigure:lumpedIC}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{effectofasymmetry.pdf}
+\caption{Effect of asymmetric rates in hidden models. Models with asymmetric rates are preferred over models with equal rates \cref{supptable:asymmetry}} % XXX
+\label{suppfigure:asymmetric}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrDP.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of theM1. D/P ploidy model} % XXX
+\label{suppfigure:DPnodipasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the M1. D/P model} % XXX
+\label{suppfigure:DPnodip}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrDPAB.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the M4. D/P+A/B asym model} % XXX
+\label{suppfigure:DPnodipABasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{hisseDPnodipasymposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the M4. D/P+A/B asym model} % XXX
+\label{suppfigure:DPnodipAB}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIC.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the M11.I/C model} % XXX
+\label{suppfigure:ICasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{bisseSIposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the M11. I/C model} % XXX
+\label{suppfigure:IC}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrICAB.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the M14. I/C+A/B asym model} % XXX
+\label{suppfigure:ICABasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{hisseSIasymposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the M14. I/C+A/B asym model} % XXX
+\label{suppfigure:ICAB}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIDCDCP.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M16. ID/CD/CP model} % XXX
+\label{suppfigure:IDCDCPnodipasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the  M16. ID/CD/CP model} % XXX
+\label{suppfigure:IDCDCPnodip}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIDCDCPAB.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M19. ID/CD/CP+A/B asym model} % XXX
+\label{suppfigure:IDCDCPnodipABasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{muhissenodipasymposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the M19. ID/CD/CP+A/B asym model} % XXX
+\label{suppfigure:IDCDCPnodipAB}
+\end{suppfigure}

### 2019-07-13 manuscript/results.tex
all edits from RZF in Methods and Results

@@
 \subsubsection{Are diversification patterns only determined by hidden states?}
-Whether the determination of the diversification is dependent on the focal traits and not only on the hidden state was done via the comparisons of character independent models against the focal traits w/o hidden states models (results shown in  \cref{table:bayesfactors},\cref{table:M11M15},\cref{supptable:M6M10},\cref{supptable:M16M20},\cref{supptable:M21M25}).
+Whether the determination of the diversification is dependent on the focal traits and not only on the hidden state was done via the comparisons of character independent models against the focal traits w/o hidden states models (results shown in  \cref{table:bayesfactors,table:M11M15,supptable:M6M10,supptable:M16M20,supptable:M21M25}).
 The general pattern is to moderate- to strongly prefer models with focal traits and asymmetric hidden states over the rest of the models with the exception of the ploidy only model where the character independent model (M2) seems to be equally preferred($K=0.5$) as the model with ploidy and asymmetric hidden rates (M4, \cref{table:bayesfactors}).
@@
 We considered models both with and without diploidization in order to explore its effects on the estimates of state-dependent diversification.
-For the models that only include diploid and polyploid states (M6-M10) the results are similar in terms of the effect of diversification (\cref{figure:alldip}).
+For the models that only include diploid and polyploid states (M6-M10) the results are similar in terms of the effect of diversification (\cref{suppfigure:alldip}).
 When comparing models with diploidization against models without diploidization we found moderate evidence that models containing the parameter $\delta$ are preferred over models without diploidization (summarized in \cref{supptable:testdiploidization}).
@@
 Furthermore, hidden state models with asymmetric traits are strongly preferred over models with equal rates between hidden states (summarized in \cref{supptable:asymmetry}).
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|r|c|c|c|c|l|}
-\hline
-Model                           & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M7    & M8     & M9    & M10   & Evidence                                                                                     \\ \hline
-M6.
-D/P+$\delta$                & -1268.83                                                          & 55.41 & 42.78  & 53.37 & 52.85 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M6\end{tabular}            \\ \hline
-\textbf{M7.
-CID D/P+$\delta$}   & \textbf{-1212.42}                                                 &       & -12.62 & -2.04 & -2.04 & \begin{tabular}[c]{@{}l@{}}Model M7 moderately\\ preferred over M9 and M10\end{tabular}      \\ \hline
-M8.
-D/P+A/B $\delta$            & -1214.46                                                          &       &        & 10.58 & 10.07 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular} \\ \hline
-M9.
-D/P+A/B +$\delta$ asym      & -1214.46                                                          &       &        &       & -0.51 & No evidence                                                                                  \\ \hline
-M10.
-D/P+A/B +$\delta$ all asym & -1214.97                                                          &       &        &       &       &         \\ \hline
-\end{tabular}
-\caption{Bayes factors of  ploidy-only models with diplodization.
-Results indicate that a character independent model (M6) is strongly preferred over BiSSE (M1)  model.
-M6 is moderately preferred over any of the HiSSE models with asymmetric hidden rates (M9, M10).}
-\label{supptable:M6M10}
-\end{supptable}
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|r|c|c|c|c|l|}
-\hline
-Model                            & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M18   & M19   & M20   & Evidence                                                                                                   \\ \hline
-M16.
-ID/CP/CD                    & -1459.11                                                          & 45.11 & 65.91 & 63.98 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M16\end{tabular}                         \\ \hline
-M17.
-CID ID/CP/CD                & *                                                                 &       &       &       &                                                                                                            \\
-M18.
-ID/CP/CD+A/B                & -1414                                                             &       & 20.79 & 18.87 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}               \\ \hline
-\textbf{M19.
-ID/CP/CD +A/B asym} & \textbf{-1393.20}                                                 &       &       & -1.92 & \begin{tabular}[c]{@{}l@{}}Asymmetric hidden rates moderately\\ preferred over all asymmetric\end{tabular} \\ \hline
-M20.
-ID/CP/CD +A/B all asym      & -1393.12                                                          &       &       &       &   \\ \hline
-\end{tabular}
-\caption{Bayes factors of ploidy and breeding system withou diploidization.
-Results indicate that the MuHiSSE model with asymmetric hidden rates (17) is  strongly preferred over M16 and M18 and moderately preferred over the MuHiSSe with all rates asymetric (M20).
-*Marginal log-likelihood for M17 could not be calculated in our limited amount of computational time.}
-\label{supptable:M16M20}
-\end{supptable}
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|r|c|c|c|c|l|}
-\hline
-Model                                 & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M22   & M23    & M24   & M25   & Evidence                                                                                         \\ \hline
-M21.
-ID/CP/CD+$\delta$                & -1454.68                                                          & 55.48 & 46.031 & 67.94 & 65.15 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M21\end{tabular}               \\ \hline
-M22.
-CID ID/CP/CD+$\delta$            & -1399.201                                                         &       & -9.452 & 12.45 & 9.675 & \begin{tabular}[c]{@{}l@{}}Model M24 strongly preferred \\ over M22\end{tabular}                 \\ \hline
-M23.
-ID/CP/CD+A/B $\delta$            & -1408.65                                                          &       &        & 21.91 & 19.12 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}     \\ \hline
-\textbf{M24.
-ID/CP/CD +$\delta$ asym} & \textbf{-1386.74}                                                 &       &        &       & -2.78 & \begin{tabular}[c]{@{}l@{}}Asymmetric hidden rates \\ preferred over all asymmetric\end{tabular} \\ \hline
-M25.
-ID/CP/CD +$\delta$ all asym      & -1389.52                                                          &       &        &       &       &                      \\ \hline
-\end{tabular}
-\caption{Bayes factors of ploidy and breeding system with diploidization.
-Results indicate that the MuHiSSE model with asymmetric hidden rates (M24) is  strongly preferred over M21-M23 and moderately preferred over the MuHiSSe with all rates asymetric (M25).}
-\label{supptable:M21M25}
-\end{supptable}
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-5pt}
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
-M1.
-D/P                                  & -1238.76                                                          & M1 vs.
-M4   & 60.47           & M4 (Strong)                                                          \\
-\textbf{M4.
-D/P+A/B asym}                & \textbf{-1223.28}                                                &             &                 &                                                                      \\ \hline
-M11.
-I/C                                 & -1309.07                                                          & M11 vs.
-M14 & 61.35           & M14 (Strong)                                                         \\
-\textbf{M14.
-I/C+A/B asym}               & \textbf{-1247.72}                                                 &             &                 &                                                                      \\ \hline
-M16.
-ID/CD/CP                            & -1459.11                                                          & M16 vs.
-M19 & 65.90           & M19 (Strong)                                                         \\
-\textbf{M19.
-ID/CD/CP+A/B asym}          & \textbf{-1393.20}                                                &             &                 &                                                                      \\ \hline
-M6.
-D/P+$\delta$                         & -1283.76                                                          & M6 vs.
-M9   & 69.3            & M9 (Strong)                                                          \\
-\textbf{M9.
-D/P+A/B+$\delta$ asym}       & \textbf{-1214.46}                                                 &             &                 &                                                                      \\ \hline
-M21.
-IC/CD/CP+$\delta$                   & -1454.68                                                          & M21 vs.
-M24 & 68.48           & M24 (Strong)                                                         \\
-\textbf{M24.
-IC/CD/CP+A/B+$\delta$ asym} & \textbf{-1386.20}                                                 &             &                 &                                                                      \\ \hline
-\end{tabular}
-\caption{Test for addition of hidden states in models via Bayes factors. Models with hidden states are strongly preferred over simpler models}
-\label{supptable:testaddhidden}
-\end{supptable}
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
-M1.
-D/P                                  & -1238.76                                                          & M1 vs.
-M6   & 65.92           & M6 (Strong)                                                          \\
-\textbf{M6.
-D/P+$\delta$}                & \textbf{-1267.84}                                                 &             &                 &                                                                      \\ \hline
-M4.
-D/P+A/B asym                         & -1223.28                                                          & M4 vs.
-M9   & 8.81            & M9 (Moderate)                                                        \\
-\textbf{M9.
-D/P+A/B+$\delta$ asym}       & \textbf{-1214.46}                                                 &             &                 &                                                                      \\ \hline
-M16.
-ID/CD/CP                            & -1459.11                                                          & M16 vs.
-M21 & 4.41            & M21 (Moderate)                                                       \\
-\textbf{M21.
-ID/CD/CP+$\delta$}          & \textbf{-1454.68}                                                 &             &                 &                                                                      \\ \hline
-M19.
-IC/CD/CP asym                       & -1393.20                                                          & M19 vs.
-M24 & 6.46            & M24 (Moderate)                                                       \\
-\textbf{M24.
-IC/CD/CP+A/B+$\delta$ asym} & \textbf{-1386.20}                                                 &             &                 &                                                                      \\ \hline
-\end{tabular}
-\caption{Test for inclusion of a diploidization rate via Bayes factors. Models with diploidzation are moderately preferred over models that do not include a diploidization rate}
-\label{supptable:testdiploidization}
-\end{supptable}
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-5pt}
-\begin{tabular}{|l|c|c|c|l|}
-\hline
-Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
-M3. D/P+A/B                                 & -1234.52                     &    M3 vs. M4                 & 11.239                  & M4 (Strong)   \\
-\textbf{M4. D/P+ A/B asym}            & -1223.28                     &   M4 vs. M5              & -1.658                  & M4 (Moderate)   \\
-M5. D/P+A/B all asym                 & -1224.93                     &                     &                         &                         \\ \hline
-M13. I/C+ A/B                                & -1270.47                     &   M13 vs. M14                  & 22.75                   & M14 (Strong)   \\
-\textbf{M14. I/C+ A/B asym}            & \textbf{-1247.72}            & M14 vs. M15         & {0.05}           & No evidence             \\
-M15. I/C+ A/B all asym               & -1247.67                     &                     &                         &                         \\ \hline
-M18. IC/CD/CP+A/B                            & -1414.00                     &     M18 vs. M19                & 20.79                   & M19 (Strong)    \\
-\textbf{M19. IC/CD/CP+ A/B asym}       & \textbf{-1393.21}            & M19 vs. M20          & {-1.919}         & M19 (Moderate)   \\
-M20. IC/CD/CP+ A/B all asym           & -1395.129                    &                     &                         &                         \\ \hline
-M8. D/P +A/B+$\delta$                          & -1225.05                     &  M8 vs. M9                   & 10.58                   & M9 (Strong)    \\
-\textbf{M9. D/P+ A/B+ $\delta$ asym}      & \textbf{-1214.46}                     &    M9 vs. M10                 & -0.52                   & M10 (Moderate) \\
-M10. D/P+A/B+$\delta$ all asym        & -1214.98                     &                    &                         &                         \\ \hline
-M23. IC/CD/DP+A/B+$\delta$                 & -1408.65                     &     M23 vs M24                & 21.91                   & M24(Strong)     \\
-\textbf{M24. IC/CD/DP+A/B+$\delta$ asym} & \textbf{-1386.74}                     &   M24 vs M25              & -2.78                   & M24 (Moderate) \\
-M25. IC/CD/DP+A/B+ $\delta$ all asym     & -1389.52                     &                     &                         &                         \\ \hline
-\end{tabular}
-\caption{Test for symmetry of the hidden and trait rates via Bayes factors.
-For all models asymmetric hidden transition rates are preferred over models with equal hidden rates.
-Adding more complexity by assuming asymmetry in the trait rate within hidden states does not improve models.}
-\label{supptable:asymmetry}
-\end{supptable}
-
-% check values of delta and rho D/P+A/B with(out) delta
-% i.e. parameter values in figure S1 & S4 vs Table S2
-
-%B: heads up: I added two clarifying descriptive sentences in the two figures below. Maybe something about noting x-axis scale differences would be useful.
-\begin{figure}
-\includegraphics[width=\textwidth]{alldiversificationasymhidden.pdf} 
-\caption{Net diversification rates for all models without diploidization rate estimate. 
-Each panel contains a graphical summary of estimated model parameters and displays posterior distributions for net diversification estimates.}
-\label{figure:netdivall}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{lumped.pdf} %lumped.pdf
-\caption{Comparing two and three state models is possible via lumpable models (M26, M28).
-These models used the same classifications as the three-state model M16 which allow for comparisons via Bayes Factors as shown in \cref{table:lumped}.}
-\label{figure:lumped}
-\end{figure}
-
-% E todo: re-run with new rate estimates; add panel letters?
-\begin{figure}
-    \centering 
-    \includegraphics[width=0.5\textwidth]{pathstates2} \\ [40pt] 
-        \includegraphics[width=0.9\textwidth]{pathways}
-    \caption{
-        Contributions of the two pathways to polyploidy.
-        The one-step pathway is direct ID$\rightarrow$CP transitions.
-        The two-step pathway consists of ID$\rightarrow$CD$\rightarrow$CP transitions.
-        When considering only rates of transitions among the states (ignoring the diversification rate parameters), the one-step pathway dominates on short timescales and the two-step on long timescales (left panels).
-        When also considering diversification within each state, the one-step pathway, in which polyploidization breaks down SI, dominates over any timescale (right panels).
-        The top panels show the separate contributions of each pathway.
-        The bottom panels show the proportional contribution of the one-step pathway (\ie one-step / [one-step + two-step]).
-    }
-    \label{figure:pathways}
-\end{figure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{alldiversificationasymhiddendip.pdf}
-\caption{Posterior distribution for all the best models with diploidization} % XXX
-\label{suppfigure:alldip}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{comparablesDP.pdf}
-\caption{Lumped ploidy models to assess the effect of adding breeding system to create a three-state model.
-Moderate evidence exists that adding a breeding system is necessary \cref{table:lumped} } % XXX
-\label{suppfigure:lumpedDP}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{comparablesIC.pdf}
-\caption{Lumped breeding system models to assess the effect of adding ploidy to create a three-state model.
-Moderate to insignificant evidence exists that adding ploidy is necessary \cref{table:lumped} } % XXX
-\label{suppfigure:lumpedIC}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{effectofasymmetry.pdf}
-\caption{Effect of asymmetric rates in hidden models. Models with asymmetric rates are preferred over models with equal rates \cref{supptable:asymmetry}} % XXX
-\label{suppfigure:asymmetric}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDP.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of theM1. D/P ploidy model} % XXX
-\label{suppfigure:DPnodipasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M1. D/P model} % XXX
-\label{suppfigure:DPnodip}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDPAB.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the M4. D/P+A/B asym model} % XXX
-\label{suppfigure:DPnodipABasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{hisseDPnodipasymposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M4. D/P+A/B asym model} % XXX
-\label{suppfigure:DPnodipAB}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIC.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the M11.I/C model} % XXX
-\label{suppfigure:ICasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{bisseSIposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M11. I/C model} % XXX
-\label{suppfigure:IC}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrICAB.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the M14. I/C+A/B asym model} % XXX
-\label{suppfigure:ICABasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{hisseSIasymposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M14. I/C+A/B asym model} % XXX
-\label{suppfigure:ICAB}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCP.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M16. ID/CD/CP model} % XXX
-\label{suppfigure:IDCDCPnodipasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the  M16. ID/CD/CP model} % XXX
-\label{suppfigure:IDCDCPnodip}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCPAB.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M19. ID/CD/CP+A/B asym model} % XXX
-\label{suppfigure:IDCDCPnodipABasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{muhissenodipasymposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M19. ID/CD/CP+A/B asym model} % XXX
-\label{suppfigure:IDCDCPnodipAB}
-\end{suppfigure}

### 2019-07-15 manuscript/abstract.tex
updated some text for the additional models
- Edits through end of Introduction

@@
 If particular traits consistently affect rates of speciation and extinction, broad macroevolutionary patterns can be understood as consequences of selection at high levels of the biological hierarchy.
-Identifying traits associated with diversification rate differences is tricky, however, because of the wide variety of traits to consider and the statistical challenges of testing for associations from comparative phylogenetic data.
+Identifying traits associated with diversification rate differences is complicated by the wide variety of characters under consideration and the statistical challenges of testing for associations from comparative phylogenetic data.
+%B: complicated by <- tricky, however, because of ...
+to consider; was %Identifying traits associated with diversification rate differences is tricky, however, because of the wide variety of traits to consider and the statistical challenges of testing for associations from comparative phylogenetic data.
 Ploidy (diploid vs.\ polyploid states) and breeding system (self-incompatible vs.\ self-compatible states) have been repeatedly suggested as possibly drivers of differential diversification.
-We investigate the roles of these traits, including their interaction, on speciation and extinction rates in Solanaceae.
-We find that the effect of ploidy can largely be explained by its correlation with breeding system, and that additional unknown factors work with breeding system to determine diversification rates.
-These results are largely robust to allowing for diploidization.
-Finally, we show that the lower rate of diversification for self-compatible diploids makes it less likely for that state to lie along the evolutionary pathway from self-incompatible diploids to polyploids.
+We investigate the connections of these traits, including their interaction, to speciation and extinction rates in Solanaceae.
+We show that the effect of ploidy on diversification can be largely explained by its correlation with breeding system and that additional unknown factors, alongside breeding system, determine diversification rates.
+%B: was: We find that the effect of ploidy can largely be explained by its correlation with breeding system, and that additional unknown factors work with breeding system to determine diversification rates.
+These results are largely robust to allowing for diploidization. %B: left along, but I prefer: "These results are largely robust to inclusion of diploidization."
+Finally, we find that the direct pathway, in which breakdowns of self-incompatibility arise as byproducts of polyploidization, rather than a step-wise pathway with breakdowns followed by polyploidization, is disproportionately important in the evolutionary history of the nightshade family.
+%B: was: "from self-incompatible diploids lower rate of diversification associated with self-compatible diploids makes it less likely for that state to lie along the evolutionary pathway from self-incompatible diploids to polyploids."

### 2019-07-15 manuscript/introduction.tex
- Edits through end of Introduction

@@
 Species accumulate across the tree of life at different rates. 
-One possible explanation for this phenomenon is that the various traits or character states of species differentially affect rates of diversification. 
-Dramatic increases in phylogenetic and phenotypic data and methodological advances have greatly accelerated the search for traits that influence diversification traits.
+One possible explanation for this phenomenon is that the various traits differentially affect rates of diversification.
+%B: note: traits and character states are synonymous.
+I've edited the rest of the manuscript so that "trait" is retained where ambiguity does not matter, but where directly contrasted (vs.: as here), I changed the wording.
+Dramatic increases in phylogenetic and phenotypic data along with methodological advances have greatly accelerated the search for traits that influence diversification traits. 
+%B: was: "phenotypic data and methodological advances"
 Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a challenge \citep[\eg][]{maddison_2015, rabosky_2015, beaulieu_2016, rabosky_2017}. 
-One difficulty is that speciation and extinction likely do not depend on a single trait, so the biological and environmental contexts in which traits occur can lead to complex interactions that affect lineage diversification \citep{beaulieu_2016, caetano_2018, herrera_2018}.
-Consequently, examining the association of only one trait with diversification patterns can be misleading. 
-Here, we embrace this challenge by jointly investigating two traits thought to influence speciation and extinction rates---ploidy level and breeding system---while allowing for their interactions and other confounding factors.
+One difficulty is that speciation and extinction likely do not depend on a single character, so the biological and environmental contexts in which traits occur can lead to complex interactions that affect lineage diversification \citep{beaulieu_2016, caetano_2018, herrera_2018}.
+Consequently, examining the association of only one character with diversification patterns can be misleading. 
+Here, we embrace this challenge by jointly investigating two characters thought to influence speciation and extinction rates---ploidy level and breeding system---while allowing for their interactions and other confounding factors.
 % E: We need a "Here, we..." here.  Otherwise, we've just told the reader "not one trait!" and then launch into one trait.
 % Could also state that we newly test whether adding one more trait is worthwhile.
+%B: I think you are correct, but it introduces a new concern: we now say N=1 is too few, so not why try N>2 characters?
+Why is N=2 the Goldilocks number?
+I didn't change anything but it's clear fodder for Discussion.
 
@@
 Increase in ploidy can alter many traits and affect a variety of evolutionary and ecological processes \citep{ramsey_2002, sessa_2019}.
-%
-Polyploids have been documented to have an overall lower diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
-Other studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications have played an important macroevolutionary role driving innovation and diversification in plants.
+%B: The contrast below was between: "Polyploids have been documented..." vs.
+"Other studies find...".
+I modified it to: "Polyploids were generally found..." vs.
+"more recent studies ..."; If you feel that "Many studies" or "Recent studies" is better, please alter as appropriate.
+%B: Separately, "common" and "numerous" are synonymous as used here (I removed "common" because that term also has the "unrefined" meaning).
+Polyploids were generally found to have an overall lower diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
+Many recent genomic studies, however, find numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications may have played an important macroevolutionary role driving innovation and diversification in plants.
 Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies pervasive diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
-Our analyses re-examine the association between ploidy and lineage diversification by extending the approach of \citet{mayrose_2011, mayrose_2015} to include these additional processes.
-We allow for transitions not only toward polyploidy, but also away from it.
-We also allow for heterogeneous diversification within each ploidy level, which could capture, for example, if diploids with a past history of polyploidy diversify more rapidly than diploids with no polyploid ancestry.
+Our analyses re-examine the association between ploidy and lineage diversification by extending the approach of \citet{mayrose_2011, mayrose_2015} to include transitions between diploid and polyploid states.
+We also allow for heterogeneous diversification rates within each ploidy level, which could examine whether, for example, diploids with ancestral polyploidy diversify more rapidly than diploids without it.
 
-Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and crucially affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
-In particular, genetic self-incompatibility (SI) systems cause plants to reject their own pollen, and loss of self-incompatibility, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974, igic_2008}.
+Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are also remarkably common and crucially affect the distribution and amount of genetic variation in populations \citep{barrett2013}.
+%B: note: we use "remarkably" to describe the frequency of both characters; hence "also."
+In particular, a variety of genetic self-incompatibility (SI) systems cause plants to reject their own pollen, and loss of such mechanisms, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974, igic_2008}.
 Previous analyses reported higher rates of diversification for SI than for SC lineages in Solanaceae \citep{goldberg_2010}. 
 Similarly, heterostylous SI lineages in Primulaceae seem to diversify faster \citep{devos2014}, as do outcrossing lineages in Onagraceae \citep{freyman_2019}.
-Although these findings suggest a consistent macroevolutionary role of breeding system, it is unlikely to be the sole trait determining lineage diversification.
+Although these findings suggest a consistent macroevolutionary role of breeding system, it is unlikely to be the sole character determining lineage diversification.
 We thus investigate the relationship of breeding system to speciation and extinction rates in the context of ploidy and other background factors.
 
-Polyploidy and self-fertilization are widely observed to be associated \citep{stebbins1950}.
+Polyploidy and self-fertilization are widely thought to be associated \citep{stebbins1950}.
 There are several possible general explanations: whole genome duplication may facilitate the transition to selfing by masking inbreeding depression, self-fertilization may facilitate polyploidy establishment by avoiding the triploid bridge, or both selfing and polyploidy may tend to be found in weedy or peripheral populations \citep{ramsey_1998, barringer2007, barrett2008, husband2008}.
-Additionally, in RNase-based gametophytic SI systems, polyploidization alone directly causes the loss of SI \citep{stout1942, lewis1947}.
+Additionally, in RNase-based gametophytic SI systems, polyploidization directly causes the loss of SI \citep{stout1942, lewis1947}.
+%B: rm "alone" in  "...polyploidization alone directly causes..." because it incorrectly implies that nothing other than polyploidization causes SC.
 %B: disclaimer: ceteris paribus statement omitted; polyploidization ordinarily causes SC because of the formation of heterodiploid pollen.
@@
 Strictly speaking, this is a feature of non-self recognition systems of the kind found in RNase-based GSI.
-In these systems, self-incompatibility occurs because haploid self-pollen grains, with one S-allele at the locus controlling the SI response, are unable to detoxify the S-RNase produced by the same S-allele in the style \citep{kubo2010}.
+In these systems, SI occurs because haploid self-pollen grains, with one S-allele at the locus controlling the SI response, are unable to detoxify the S-RNase produced by the same S-allele in the style \citep{kubo2010}.
 The unreduced pollen of diploids, however, can contain two S-alleles expressed in pollen, which jointly gain the ability to detoxify the S-RNases produced by any maternal genotype \citep{entani1999, tsukamoto2005, kubo2010}.
 Initial mutant individuals with pollen containing doubled haploid genomes are consequently capable of self-fertilization, with exceedingly few exceptions \citep{hauck_2002, nunes_2006}.
-This system of RNase-based SI is regarded as ancestral in eudicots \citep{igic_2001,steinbachs_2002}, and it is expressed in all SI species of Solanaceae examined to date.
+RNase-based SI is regarded as ancestral in eudicots \citep{igic_2001,steinbachs_2002}, and it is expressed in all SI species of Solanaceae examined to date.
 The consequent absence of SI polyploids in this family yields a strong correlation between ploidy and breeding system \citep{robertson_2011}.
@@
 First, we investigate their joint influence on rates of speciation and extinction.
-Each trait alone has been shown to associate with differential lineage diversification, but if their effects are not additive, studying each trait separately may not reveal the effects of their combinations.
-Second, we examine the order of transitions in the two traits.
-Evolution is predominantly from diploid to polyploid, and from SI to SC, but there are two paths by which diploid SI lineages can eventually become SC polyploids.
-Loss of SI in diploids could be directly caused by polyploidization (as explained above, for gametophytic SI systems), resulting in a one-step pathway to polyploid SC species. 
-Alternatively, SI diploids could first transition to SC without an increase in ploidy, and then undergo polyploidization, resulting in a two-step pathway.
-\citet{robertson_2011} compared the contributions of these two paths, finding that evolution from SI diploid to SC polyploid is more likely to proceed via the one-step pathway over short timescales, but via the two-step pathway over long timescales.
-Their analysis considered only transitions among the states, however, so we investigate whether these results hold true when allowing for the influence of these traits on lineage diversification.
+Each character alone has been shown to associate with differential lineage diversification, but if their effects are not additive, studying them separately may not reveal the effects of their combinations.
+Second, we examine the order of transitions in the two characters.
+Even though evolution may commonly proceed from diploid to polyploid, and from SI to SC states, there are two paths by which diploid SI lineages can eventually become SC polyploids.
+Loss of SI in diploids could be directly caused by polyploidization (as explained above, for RNase-based SI systems), resulting in a one-step pathway to SC polyploids. 
+Alternatively, SI diploids could first transition to SC without an increase in ploidy, and subsequently undergo polyploidization, resulting in a two-step pathway to SC polyploids.
+\citet{robertson_2011} compared the contributions of these two paths, finding that evolution from SI diploids to SC polyploids is more likely to proceed via the one-step pathway, over short timescales, but via the two-step pathway over long timescales.
+They considered only transitions among the states, however, so we investigate whether these results hold true when allowing for differences in lineage diversification.
 
-Undertaking these questions requires powerful phylogenetic comparative methods.
-We employ the SSE model framework, which simultaneously models transitions between the discrete states of a trait and different rates of speciation and extinction associated with each of those states \citep{maddison_2007, fitzjohn_2012}.
-Although potentially very powerful, this approach is prone to reporting a false association between a trait and lineage diversification when the data contain additional, unmodeled diversification rate heterogeneity \citep{maddison_2015, rabosky_2015}.
+%B: redundant with "very powerful" below rm: Examining these questions requires powerful phylogenetic comparative methods.
+To address these questions, we employ an extended SSE model framework, which simultaneously models transitions between the discrete states of a trait and different rates of speciation and extinction associated with each of those states \citep{maddison_2007, fitzjohn_2012}.
+Although potentially very powerful, this approach can be prone to reporting a false association between a trait and lineage diversification when the data contain additional, unmodeled diversification rate heterogeneity \citep{maddison_2015, rabosky_2015}.
 We address this concern in two ways.
-First, by considering both ploidy and breeding system simultaneously, we reduce the chance that an effect of one trait on diversification would be mistaken for the effect of the other.
-Second, by incorporating a hidden state, we allow that an unknown factor could additionally be influencing diversification \citep{beaulieu_2016}.
+First, we consider both ploidy and breeding system simultaneously, which reduces the chance that an effect of one trait on diversification would be mistaken for the effect of the other.
+Second, we employ a hidden state model, which allows for an unknown factor to additionally influence diversification \citep{beaulieu_2016}.
 
-Our macroevolutionary analyses in Solanaceae apply different combinations of these techniques in order to parse out the associations of ploidy and breeding system on speciation and extinction rates, while simultaneously allowing that other processes could influence diversification as well.
-We first present models in which the effect of ploidy and breeding system are considered separately, and we compare the inferences to previously published results. 
-Then we add a hidden trait to each of those models and investigate whether the focal trait still explains differences in diversification.
-Finally, we model ploidy and breeding system jointly to assess their combined relationship to diversification, with or without an additional hidden trait.
+%B: I altered this line, but it is still a bit redundant given previous sentence.
+Our analyses evaluate combinations of macroevolutionary models to parse out the associations of ploidy and breeding system on speciation and extinction rates in Solanaceae, while simultaneously accounting for other processes that influence diversification.
+First, we present models in which the effects of ploidy and breeding system are considered separately and compare the inferences to previously published results. 
+Second, we add a hidden character to each model and investigate if state-dependent diversification in focal character retains its explanatory power, or whether the added hidden character obscures it.
+Next, we model ploidy and breeding system jointly to assess their combined influence on diversification, with or without an additional hidden character.
 % E: say something here about testing whether it's worth adding in another real trait?
 With this joint trait model, we also quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
-For all of the models involving ploidy, we investigate the potential effects of including diploidization.
+Finally, for all models involving ploidy, we investigate the potential effects of including diploidization.
 Our results highlight the importance of considering non-additive effects of traits on net diversification rates, especially when there are strong biologically-driven correlations among them.

### 2019-07-15 manuscript/main.tex
acknowledging awesome people
- Edits through end of Introduction

@@
 \begin{center}
-    \textsc{Interaction between ploidy, breeding system, and lineage diversification}
+    \textsc{Interaction between Ploidy, Breeding System, and Lineage Diversification}
 
@@
 %--------------------------------------------------
-
-This work was supported by the National Science Foundation (NSF DEB-1655478 to EEG and NSF DEB-1655692 to BI) and the Binational Science Foundation (BSF-2016550 to IM).
+Special thanks to Bob Thomson for donating computational resources that made possible testing the convergence of the models. 
+We also thank Sergei Tarasov for clarifying lumpability concepts, and Carrie Tribble for troubleshooting some ancestral estimate figures. 
 The computing resources were provided by the Minnesota Supercomputing Institute (MSI) at the University of Minnesota.
-%BI would like to thank SCHB for inspiration, sparring, and encouragement; maybe EEG, as well?
+This work was supported by the National Science Foundation grant DEB-1655478 (to EEG), DEB-1655692 and NESCent sabbatical scholar award EF-0905606 (to BI), and the Binational Science Foundation BSF-2016550 (to IM).
+%B: new:
+We thank Dan Schoen and three anonymous reviewers for providing uncommonly thoughtful and helpful reviews, as well as Spencer Barrett for continually inspiring our work.
 

### 2019-07-15 manuscript/results.tex
started editing Results
Got through the "traits and diversification" section.
A lot of small wording edits, but no changes to the general presentation.

@@
 
-\subsection{Models}
-\subsubsection{Ploidy and diversification}
-When considering ploidy alone (M1. $D/P$ model), we found a larger net diversification rate for diploids than for polyploids, in agreement with \citet{mayrose_2011, mayrose_2015}.
-This result holds with (\cref{figure:netdivall}A) or with the diploidization parameter (model M6 \cref{suppfigure:alldip}A) .
-Incorporating a hidden state in this model, however, reduces the clear separation in diversification rate estimates between diploids and polyploids (M4.
-$D/P+A/B$ asym model; \cref{figure:netdivall}B, and M9 \cref{suppfigure:alldip}B).
-Thus, differences in net diversification rates are explained by an unknown factor.
-The lack of differences between diploid and polyploid net diversification rates is also supported by the Bayes factors shown in table \cref{table:bayesfactors} where both model M4 and the character independent model M2 have equal evidence of  being preferred over every other model of ploidy.
-See table \cref{supptable:M6M10} showing analogous results when diploidzation is accounted for.
+% E: I removed the model names (like, "I/C+A/B") because we don't use them in the Methods anymore.
+I think it works find just to say in real words what each model includes, as it comes up, and the model numbers are good for shorthand.
 
-\subsubsection{Breeding system and diversification}
-When considering breeding system alone (M11. $I/C$ model, \cref{figure:netdivall}C), we found a larger net diversification rate for SI than for SC species, in agreement with \citet{goldberg_2010}.
-When a hidden state is included (M14.
-$I/C+A/B$ model), the large net diversification rate difference persists for hidden state $A$ but and it slightly overlaps in the  hidden state $B$ (\cref{figure:netdivall}D).
-When calculating Bayes factors for all breeding system models (M11-M15) we found that the best two supported models are the models with at least asymmetric hidden states (M14 and M15 \cref{table:M11M15})
-Thus, differences in net diversification rates are best explained by both breeding system and an unknown factor \citep[as in Onagraceae;][]{freyman_2019} and not only by the focal trait nor an unknown factor.
-% The transition rate from SI to SC is $q_{IC}=0.3$ for both the I/C and the I/C+A/B models.
+\subsection{Trait-dependent diversification}
 
-\subsubsection{Ploidy, breeding system, and diversification}
-When considering ploidy and breeding system together (M16.
-$ID/CD/CP$ model), the net diversification rate for SI diploids was greater than for either SC diploids or SC polyploids, with or without diploidization (M16 \cref{figure:netdivall}E, M21\cref{suppfigure:alldip}E).
+\subsubsection{Ploidy only}
+
+When considering ploidy alone, we found a larger net diversification rate for diploids than for polyploids, in agreement with \citet{mayrose_2011, mayrose_2015}.
+This result holds with (model M1, \cref{figure:netdivall}A) or without the diploidization parameter (M6, \cref{suppfigure:alldip}A).
+Incorporating a hidden state in this model, however, reduces the clear separation in diversification rate estimates between diploids and polyploids (M4, \cref{figure:netdivall}B; M9, \cref{suppfigure:alldip}B).
+Statistical model comparisons show a clear preference for models in which only a hidden state affects diversification or a hidden state as well as ploidy (M2 and M4; \cref{table:bayesfactors}).
+Results are similar when diploidization is included (\cref{supptable:M6M10}).
+Thus, the role of ploidy in net diversification is unclear, with rate differences perhaps better explained by another factor.
+
+\subsubsection{Breeding system only}
+
+When considering breeding system alone (M11, \cref{figure:netdivall}C), we found a larger net diversification rate for SI than for SC species, in agreement with \citet{goldberg_2010}.
+When a hidden state is included, the large net diversification rate difference persists for one hidden state but is diminished for the other (M14, \cref{figure:netdivall}D).
+In the statistical model comparisons, the best two supported models include diversification differences due both to breeding system and to a hidden trait (M14 and M15, \cref{table:M11M15}).
+Thus, breeding system seems to play a role in diversification differences, though a hidden factor does as well.
+
+\subsubsection{Ploidy and breeding system together}
+
+When considering ploidy and breeding system together, the net diversification rate for SI diploids was greater than for either SC diploids or SC polyploids, with or without diploidization (M16, \cref{figure:netdivall}E; M21, \cref{suppfigure:alldip}E).
 Thus, the difference in net diversification associated with breeding system persists even when ploidy is included in the model.
-The reverse is not true, the association of ploidy with net diversification in the M1.
-$D/P$ model (\cref{figure:netdivall}A, \cref{suppfigure:alldip}A) appears to be driven by the subset of diploids that are SI; among SC species, net diversification rates for diploids and polyploids is not distinct.
+The reverse is not true: the association of ploidy with net diversification in the simplest ploidy-only model (M1, \cref{figure:netdivall}A, \cref{suppfigure:alldip}A) appears to be driven by the subset of diploids that are SI, while among SC species, net diversification rates for diploids and polyploids are similar.
 
-When a hidden state is included (M19.
-$IC/CD/CP+A/B$ model) the net diversification rate separation amongst states ID and (CD, CP)  is clear under hidden state A, but under hidden state B the separation exist between states ID and CP, while CD overlaps with the other two states (\cref{figure:netdivall}F).
-The same general pattern remains when diploidization is included (\cref{suppfigure:alldip}F), but the hidden states are reversed in the x-axis.
-Bayes factors for the three state models that include the character independent model show very strong support preferring asymmetric hidden state with the focal traits models shown in  \cref{supptable:M16M20} (with diploidization \cref{supptable:M21M25}).
+When a hidden state is included, the separation in net diversification rate of ID \vs CD and CP persists within one hidden state but is reduced in the other (M19, \cref{figure:netdivall}F).
+The same general pattern remains when diploidization is included (\cref{suppfigure:alldip}F).
+Model comparisons clearly favor models that include ploidy, breeding system, and the hidden trait, against the character-independent model in which the focal traits do not influence diversification (\cref{supptable:M16M20}; \cref{supptable:M21M25} with diploidization).
+% add model numbers
 
-\subsubsection{Lumped models}
-When comparing model M26.
+Using the lumped models, we find moderate support for obtaining a significantly better fit by adding breeding system to the ploidy-only model (M26 \vs M16, \cref{table:lumped}, \cref{figure:lumped}BC).
+This is also true when a hidden trait is included (M27 \vs M23, \cref{table:lumped}, \cref{suppfigure:lumpedDP}EF).
+% When comparing model M26.
 Lumped $D/P$ against M16.
 $ID/CD/CP$ (\cref{figure:lumped}B and C) Bayes factors reveal moderate support ($K=4.1$) preferring the three state model instead of the two state model (\cref{table:lumped}).
-A similar result was obtained when comparing the lumped model for ploidy but in the presence of hidden states (model M27 \textit{vs.} M23 \cref{table:lumped}, \cref{suppfigure:lumpedDP}E and F).
+% A similar result was obtained when comparing the lumped model for ploidy but in the presence of hidden states (model M27 \textit{vs.} M23 \cref{table:lumped}, \cref{suppfigure:lumpedDP}E and F).
+A similar comparison in which ploidy is added to the breeding system-only model shows no preference for the model that also includes ploidy (M28 \vs M16, \cref{table:lumped}, \cref{figure:lumped}CE).
+When including a hidden state, however, the model with both focal traits is moderately preferred over the model with only breeding system (M29 \vs M18, \cref{table:lumped}, \cref{suppfigure:lumpedIC}EF).
+% We also compared lumped models for breeding system to test for the inclusion of ploidy state (M28 \cref{figure:lumped}E \textit{vs.} M16 \cref{figure:lumped}C) and we found no evidence preferring the three-state over the two state-model ($K=-0.6$, \cref{table:lumped}).
+% However, when accounting for hidden state,  the three-state model M18.
+$I/C/CD/CP+A/B$ (\cref{suppfigure:lumpedIC}F) was moderately preferred ($K=2.6$) over the lumped model M29.(\cref{suppfigure:lumpedIC}E).
 
-We also compared lumped models for breeding system to test for the inclusion of ploidy state (M28 \cref{figure:lumped}E \textit{vs.} M16 \cref{figure:lumped}C) and we found no evidence preferring the three-state over the two state-model ($K=-0.6$, \cref{table:lumped}).
-However, when accounting for hidden state,  the three-state model M18.
-$I/C/CD/CP+A/B$ (\cref{suppfigure:lumpedIC}F) was moderately preferred ($K=2.6$) over the lumped model M29.(\cref{suppfigure:lumpedIC}E).
+From all of these types of statistical evidence, we conclude that breeding system (and a hidden factor) are strongly associated with diversification differences, and that ploidy plays a smaller role.
 

### 2019-07-15 manuscript/methods.tex
fixing the number of states (figure and methods)

@@
 Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 651 species with ploidy and/or breeding system information on the tree.
-% FIXME: Rosana will update numbers around here TODO:STILL NOT CORRECT!!!
-Of these, 290 had information for both states.
+% Fixed!
+Of these, 368 had information for both states.
 The number of species in each combination of states is summarized in \cref{figure:stateclassifications}A.
 We retained all 651 species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
-A total of 405 taxa without any information about breeding system or ploidy were excluded.
+A total of 372 taxa without any information about breeding system or ploidy were excluded.
 %Tips without trait data might not improve point estimation of diversification parameters while increasing uncertainty in the estimates of trait linked diversification
@@
 \centering
-\includegraphics[width=0.5\textwidth]{states} % FIXME: update numbers
+\includegraphics[width=0.5\textwidth]{states.pdf} % FIXME: update numbers
 \caption{
@@
 (B) Category groupings into states for each model type.
-For example, there are 34 species that are SC with unknown ploidy. % FIXME: update number if necessary
+For example, there are 70 species that are SC with unknown ploidy. % Fixed
 In the $D/P$ models these are coded as either $D$ or $P$ (uncertain/consistent with either state), in the $I/C$ models they are coded as $C$, and in the $ID/CD/CP$ models they are coded as either $CD$ or $CP$.
 In the hidden state models, all species were coded as consistent with both $A$ and $B$.
+Two species \textit{Lycium californicum} and \textit{Solanum bulbocastanum} are simultaneously ID and CP and by adding them the sample adds to the total of 651 taxa used for analyses.
 }

### 2019-07-16 manuscript/introduction.tex
Changed the end of the intro

@@
 %B: redundant with "very powerful" below rm: Examining these questions requires powerful phylogenetic comparative methods.
-To address these questions, we employ an extended SSE model framework, which simultaneously models transitions between the discrete states of a trait and different rates of speciation and extinction associated with each of those states \citep{maddison_2007, fitzjohn_2012}.
-Although potentially very powerful, this approach can be prone to reporting a false association between a trait and lineage diversification when the data contain additional, unmodeled diversification rate heterogeneity \citep{maddison_2015, rabosky_2015}.
-We address this concern in two ways.
-First, we consider both ploidy and breeding system simultaneously, which reduces the chance that an effect of one trait on diversification would be mistaken for the effect of the other.
-Second, we employ a hidden state model, which allows for an unknown factor to additionally influence diversification \citep{beaulieu_2016}.
+In the present study, we employ an extended framework of state speciation and extincion models (SSE), which simultaneously models transitions between the discrete states of a trait and different rates of speciation and extinction associated with each of those states \citep{maddison_2007, fitzjohn_2012}.
+All the models proposed for this study are coded as graphical models and Bayesian inference was performed over the proposed models using RevBayes software \citep{hoehna_2016}.
+%Although potentially very powerful, this approach can be prone to reporting a false association between a trait and lineage diversification when the data contain additional, unmodeled diversification rate heterogeneity \citep{maddison_2015, rabosky_2015}.
+We start by fitting binary state speciation and extinction models to each ploidy and breeding system independently \citep{maddison_2007}. We follow by 
+fitting models that incorporate hidden states, which reduce the chance that an effect of one trait on diversification would be mistaken for the effect of an unobserved factor that is changing the diversification pattern \citep{beaulieu_2016}.
+We compare the proposed models against their character independent counterparts \citep{beaulieu_2016} to investigate whether ploidy or breeding system are actually linked to the diversification process.
+Next, we model ploidy and breeding system jointly to assess their combined influence on diversification, with or without an additional hidden character.
+With the joint trait model without hidden state, we also quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
+Later, we propose new models that aggregate states in the ploidy and breeding system models to make them equivalent to the single trait models.
+The new `lumped' models allow us to compare whether increasing model complexity from one trait to two traits significantly improves our understanding of the diversification process.
+%B: I altered this line, but it is still a bit redundant given previous sentence.
+%Our analyses evaluate combinations of macroevolutionary models to parse out the associations of ploidy and breeding system on speciation and extinction rates in Solanaceae, while simultaneously accounting for other processes that influence diversification.
+%First, we present models in which the effects of ploidy and breeding system are considered separately and compare the inferences to previously published results. 
+%Second, we add a hidden character to each model and investigate if state-dependent diversification in focal character retains its explanatory power, or whether the added hidden character obscures it.
+%
+% E: say something here about testing whether it's worth adding in another real trait?
+Then, we extend all the models involving ploidy, to investigate the potential effects of including diploidization.
+Finally, we explore models of diversification using asymmetric rates between and within hidden states to investigate the sensitivity of the statistical inference to model assumptions.
 
-%B: I altered this line, but it is still a bit redundant given previous sentence.
-Our analyses evaluate combinations of macroevolutionary models to parse out the associations of ploidy and breeding system on speciation and extinction rates in Solanaceae, while simultaneously accounting for other processes that influence diversification.
-First, we present models in which the effects of ploidy and breeding system are considered separately and compare the inferences to previously published results. 
-Second, we add a hidden character to each model and investigate if state-dependent diversification in focal character retains its explanatory power, or whether the added hidden character obscures it.
-Next, we model ploidy and breeding system jointly to assess their combined influence on diversification, with or without an additional hidden character.
-% E: say something here about testing whether it's worth adding in another real trait?
-With this joint trait model, we also quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
-Finally, for all models involving ploidy, we investigate the potential effects of including diploidization.
 Our results highlight the importance of considering non-additive effects of traits on net diversification rates, especially when there are strong biologically-driven correlations among them.

### 2019-07-16 manuscript/results.tex
small edits to Pathways in Results

@@
 \subsection{Pathways to polyploidy}
+
 There are two pathways by which SI diploid lineages eventually---given enough time---become SC polyploids.
@@
 In the two-step pathway, SI is first lost within the diploid state, followed by polyploidization.
-Determining the relative contribution of these pathways based on our estimated transition rates from the $ID/CD/CP$ model (median rate values from M16), we find that the one-step pathway is more likely on short timescales and the two-step pathway is more likely on long timescales (\cref{figure:pathways}, left panels).
-Beginning with a single SI diploid lineage, when not much time has elapsed, the one-step pathway is more likely because it only necessitates a single event to effect transition to the SC polyploid state.
+Determining the relative contribution of these pathways based on our transition rate estimates (median transition rate values from M16), we find that the one-step pathway is more likely on short timescales and the two-step pathway is more likely on long timescales (\cref{figure:pathways}, left panels).
+Beginning with a single SI diploid lineage, when not much time has elapsed, the one-step pathway is more likely because it only necessitates a single event to reach the SC polyploid state.
 When more time has elapsed, the two-step pathway is more likely because the rate of loss of SI within diploids, $q_{IC}$, is greater than the rate of polyploidization for SI species, $\rho_I$ (\cref{suppfigure:IDCDCPnodip}).
 That is, an $ID$ lineage is more likely to begin its path to polyploidy with a transition to $CD$, but completing this path to $CP$ takes longer.
-(The results are qualitatively the same when using transition rate estimates from the model that does not allow diversification differences related to the observed states, M17).
+\Citet{robertson_2011} reached the same conclusion.
+Our result is qualitatively unchanged when using transition rate estimates from the model that does not allow diversification differences related to the observed states (M17).
 
@@
 Allowing for the different net diversification rates for each state (again using median rate estimates from M16), we find a qualitative difference in the relative pathway contributions.
-In particular, the lower rate of net diversification in the $CD$ state, relative to $ID$, means that relatively fewer lineages are available to complete the second step of the two-step pathway, ending in $CP$.
-Consequently, even over long timescales, the two-step pathway contributes less to the formation of polyploids (\cref{figure:pathways}, right panels).
+The lower rate of net diversification in the $CD$ state, relative to $ID$, means that relatively fewer lineages are available to complete the second step of the two-step pathway, ending in $CP$.
+Consequently, even over long timescales, we find that the two-step pathway contributes less to the formation of polyploids (\cref{figure:pathways}, right panels) when considering diversification as well as transitions.
 
 \subsection{Model selection key questions}
+
 \subsubsection{Are diversification patterns only determined by hidden states?}

### 2019-07-16 manuscript/methods.tex
edits to "model comparison key questions"
In Results, moved the Q/A section to before Pathways, because the
questions are mostly about diversification.

@@
 \begin{enumerate}
-\item Are diversification patterns only determined by hidden states and not the traits of interest?- Comparison of character independent models against hidden state (\cref{table:bayesfactors,table:M11M15,supptable:M6M10,supptable:M16M20,supptable:M21M25}).
-\item Are hidden state necessary to add in this context?- Comparison of simple models against hidden state models (\cref{supptable:testaddhidden}).
-\item Is a second trait adding information to the diversification process?- Comparison of lumped models against IC/CD/CP models (\cref{figure:lumped,table:lumped}).
-\item Is there evidence for diploidization?- Comparison amongst models with and without diploidization (\cref{suppfigure:alldip,supptable:testdiploidization}).
-\item Do asymmetrical rates in the model affect the pattern of diversification?
-- Comparison amongst hidden states models with equal hidden rates and asymmetrical rates (\cref{suppfigure:asymmetric,supptable:asymmetry}).
+\item Are diversification patterns only determined by hidden states and not the traits of interest?---Comparison of character independent models against hidden state (\cref{table:bayesfactors,table:M11M15,supptable:M6M10,supptable:M16M20,supptable:M21M25}).
+\item Are hidden states necessary to explain diversification rate heterogeneity?---Comparison of simple models against hidden state models (\cref{supptable:testaddhidden}).
+\item Does a second focal trait add information about the diversification process?---Comparison of lumped models against IC/CD/CP models (\cref{figure:lumped,table:lumped}).
+\item Are conclusions robust to assumptions about hidden state transitions?---Comparison amongst hidden states models with equal hidden rates and asymmetrical rates (\cref{suppfigure:asymmetric,supptable:asymmetry}).
+\item Is there evidence for diploidization?---Comparison amongst models with and without diploidization (\cref{suppfigure:alldip,supptable:testdiploidization}).
 \end{enumerate}
@@
 Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $K=ln(BF(M_1,\ M_0)) = \ln[ P(\mathbf{X} | M_1) - P(\mathbf{X} | M_0)]$.
-There is `strong evidence'  for $M_1$ when this value is more than 10, moderate support if the value is more than 1 and no evidence if the value is between -1 and 1.
-If the vaue of  $K$ is negative the evidence goes towards $M_0$.
-\citep{kass1995}.
+There is `strong evidence' for $M_1$ when this value is more than 10, moderate support if the value is more than 1, and no evidence if the value is between -1 and 1.
+If the value of $K$ is negative the evidence goes towards $M_0$ \citep{kass1995}.
 

### 2019-07-16 manuscript/results.tex
edits to "model comparison key questions"
In Results, moved the Q/A section to before Pathways, because the
questions are mostly about diversification.

@@
 
+\subsection{Key questions about diversification and transitions}
+
+The above results include several of our statistical model comparison findings.
+Here we return to the five specific questions we targeted with our model comparisons.
+
+\begin{enumerate}
+
+    \item Are diversification patterns only determined by hidden states and not the traits of interest?
+    No, our focal traits are supported as having associations with diversification differences.
+    In most cases, we find moderate to strong preference for models with the focal traits as well as hidden states, over models with only hidden states \cref{table:bayesfactors,table:M11M15,supptable:M6M10,supptable:M16M20,supptable:M21M25}).
+
+    \item Are hidden states necessary to explain diversification rate heterogeneity?
+    Yes, models with hidden states that influence diversification are strongly preferred over models containing only the focal traits (\cref{supptable:testaddhidden}).
+    This means that there are potentially many sources of diversification shifts within the family.
+
+    \item Does a second focal trait add information about the diversification process?
+    Yes, in most cases models with both ploidy and breeding system are preferred over models with only one of the focal traits (\cref{table:lumped,suppfigure:lumpedDP,suppfigure:lumpedIC}).
+
+    \item Are conclusions robust to assumptions about hidden state transitions?
+    Yes, we found that allowing different types of asymmetry in transitions within and between hidden states did not change our conclusions about net diversification differences.
+    Hidden state models with asymmetric rates are, however, strongly preferred over models with equal rates between hidden states (\cref{supptable:asymmetry}), and they show stronger differences between some net diversification rates \cref{suppfigure:asymmetric}.
+
+    \item Is there evidence for diploidization?
+    Perhaps: when comparing models with diploidization against models without it, we found moderate evidence that models containing diploidization are preferred (\cref{supptable:testdiploidization}).
+    We discuss later some further challenges in identifying diploidization.
+    We further found that our main conclusions about net diversification differences are not dependent on whether diploidization in included (\cref{suppfigure:alldip}).
+
+\end{enumerate}
+
 \subsection{Pathways to polyploidy}
@@
 
-\subsection{Model selection key questions}
-
-\subsubsection{Are diversification patterns only determined by hidden states?}
-Whether the determination of the diversification is dependent on the focal traits and not only on the hidden state was done via the comparisons of character independent models against the focal traits w/o hidden states models (results shown in  \cref{table:bayesfactors,table:M11M15,supptable:M6M10,supptable:M16M20,supptable:M21M25}).
-The general pattern is to moderate- to strongly prefer models with focal traits and asymmetric hidden states over the rest of the models with the exception of the ploidy only model where the character independent model (M2) seems to be equally preferred($K=0.5$) as the model with ploidy and asymmetric hidden rates (M4, \cref{table:bayesfactors}).
-
-\subsubsection{Are hidden states necessary?}
-In the comparisons of simple diversification models of focal traits against models with hidden states we found that models with asymmetric hidden traits are strongly preferred over simpler models (summarized in table \cref{supptable:testaddhidden}).
-
-\subsubsection{Is a second trait adding information to the diversification process?}
-As summarized in  \cref{table:lumped}, it is moderately preferred to add breeding system to ploidy only models (\cref{suppfigure:lumpedDP}).
-When adding ploidy information to breeding system models to create a three-state model, we found that there is no evidence that it is preferred (M28 vs. M16 \cref{suppfigure:lumpedIC}B and C). 
-However, it becomes moderately preferable in the presence of hidden states (M29 vs. M23  \cref{suppfigure:lumpedIC}E and C).
-
-\subsubsection{Is there evidence for diploidization?}
-We considered models both with and without diploidization in order to explore its effects on the estimates of state-dependent diversification.
-For the models that only include diploid and polyploid states (M6-M10) the results are similar in terms of the effect of diversification (\cref{suppfigure:alldip}).
-When comparing models with diploidization against models without diploidization we found moderate evidence that models containing the parameter $\delta$ are preferred over models without diploidization (summarized in \cref{supptable:testdiploidization}).
-
-\subsubsection{Do asymmetric rates in hidden states models affect the pattern of diversification?}
-When comparing different types of asymmetry across hidden states models we found that the asymmetry did not change the direction of effects in net diversification results .
+%--------------------------------------------------
+% \subsection{Model selection key questions}
+% 
+% \subsubsection{Are diversification patterns only determined by hidden states?}
+% Whether the determination of the diversification is dependent on the focal traits and not only on the hidden state was done via the comparisons of character independent models against the focal traits w/o hidden states models (results shown in  \cref{table:bayesfactors,table:M11M15,supptable:M6M10,supptable:M16M20,supptable:M21M25}).
+% The general pattern is to moderate- to strongly prefer models with focal traits and asymmetric hidden states over the rest of the models with the exception of the ploidy only model where the character independent model (M2) seems to be equally preferred($K=0.5$) as the model with ploidy and asymmetric hidden rates (M4, \cref{table:bayesfactors}).
+% 
+% \subsubsection{Are hidden states necessary?}
+% In the comparisons of simple diversification models of focal traits against models with hidden states we found that models with asymmetric hidden traits are strongly preferred over simpler models (summarized in table \cref{supptable:testaddhidden}).
+% 
+% \subsubsection{Is a second trait adding information to the diversification process?}
+% As summarized in  \cref{table:lumped}, it is moderately preferred to add breeding system to ploidy only models (\cref{suppfigure:lumpedDP}).
+% When adding ploidy information to breeding system models to create a three-state model, we found that there is no evidence that it is preferred (M28 vs. M16 \cref{suppfigure:lumpedIC}B and C). 
+% However, it becomes moderately preferable in the presence of hidden states (M29 vs. M23  \cref{suppfigure:lumpedIC}E and C).
+% 
+% \subsubsection{Is there evidence for diploidization?}
+% We considered models both with and without diploidization in order to explore its effects on the estimates of state-dependent diversification.
+% For the models that only include diploid and polyploid states (M6-M10) the results are similar in terms of the effect of diversification (\cref{suppfigure:alldip}).
+% When comparing models with diploidization against models without diploidization we found moderate evidence that models containing the parameter $\delta$ are preferred over models without diploidization (summarized in \cref{supptable:testdiploidization}).
+% 
+% \subsubsection{Do asymmetric rates in hidden states models affect the pattern of diversification?}
+% When comparing different types of asymmetry across hidden states models we found that the asymmetry did not change the direction of effects in net diversification results .
 However, models with asymmetric hidden rates reduced the overlap amongst net diversification posterior distributions as shown in \cref{suppfigure:asymmetric}.
 Furthermore, hidden state models with asymmetric traits are strongly preferred over models with equal rates between hidden states (summarized in \cref{supptable:asymmetry}).
+%--------------------------------------------------

### 2019-07-16 manuscript/discussion.tex
Rosana edited the whole thing

@@
 Considering them together, and in conjunction with another hidden factor, provides a fuller picture of macroevolutionary dynamics within Solanaceae.
-We hope our work serves as an example of how phylogenetic comparative methods can be used to disentangle the contributions of interacting traits to heterogeneous lineage diversification.
+We hope our work serves as an example of how phylogenetic comparative methods can be used to disentangle the contributions of interacting traits to heterogeneous lineage diversification, and how to statistically argue for increasing complexity in diversification modeling.
 
@@
 Previous analyses of the effects of ploidy on diversification showed that diploids had greater net diversification rates than polyploids across many angiosperm clades \citep{mayrose_2011, mayrose_2015}.
-We obtain a similar outcome when examining ploidy alone in Solanaceae (\cref{figure:netdivall}A), but a consistent effect of ploidy on diversification is not clear when we also allow that a hidden factor could affect diversification as well (\cref{figure:netdivall}B).
+We obtained a similar outcome when examining ploidy alone in Solanaceae (\cref{figure:netdivall}A), but a consistent effect of ploidy on diversification was not clear when we added a  hidden factor linked to diversification(\cref{figure:netdivall}B).
 %
-Previous analyses of breeding system in this family showed that SI species are associated with higher diversification rates than their SC counterparts (\citealt{goldberg_2010}; \cref{figure:netdivall}C).
-Our new analyses with a hidden trait show that this pattern persists on different backgrounds.
-One on background, however, SC species experience positive net diversification (\ie not a dead end; cf.\ \citealt{igic_2013}) that exceeds the diversification rate of SI species on the other background (\cref{figure:netdivall}D).
+Previous analyses of breeding system in this family showed that SI species were associated with higher diversification rates than their SC counterparts (\citealt{goldberg_2010}; \cref{figure:netdivall}C).
+Our new analyses with a hidden trait showed that this pattern persists even under the presences of an unobserved trait linked to diversification.
+Under one of the hidden states, however, SC species experienced positive net diversification (\ie not a dead end; cf.\ \citealt{igic_2013}) that exceeded the diversification rate of SI species under the second hidden state (\cref{figure:netdivall}D).
 %
-Furthermore, our models of the joint macroevolution of ploidy and breeding system provided better statistical fits and revealed how the trait combinations associate with diversification.
-We find that the higher overall high net diversification rate of diploids is dominated by SI diploids, while SC diploids have a lower diversification rate that is comparable to polyploids, which are also SC (\cref{figure:netdivall}E).
+Furthermore, our models of the joint macroevolution of ploidy and breeding system that were statistically preferred revealed how the trait combinations associate with diversification.
+We found that the highest overall net diversification rate of diploids was dominated by SI diploids, while SC diploids had a lower diversification rate that overlapped with the net diversification of polyploids, which are also SC (\cref{figure:netdivall}E).
 Thus, breeding system seems to provide the dominant explanation of diversification differences, with ploidy providing a secondary effect within SC species.
@@
 
-Throughout our numerous model comparisons, models that included a hidden trait emerged as significantly better than those that did not. % supptable ref
+Throughout our numerous model comparisons, models that included a hidden trait emerged as significantly better than those that did not (\cref{supptable:testaddhidden}). 
 This is entirely consistent with the expectation that many processes can affect speciation and extinction, far beyond the one or two traits of particular interest to us.
 A natural follow-up question is, what is this hidden trait?
-Our results show that to some extent, breeding system functions as a hidden trait in the ploidy-centered analysis, and vice versa.
-But there is also strong statistical support for a hidden trait beyond ploidy and breeding system.
+Our results show that to some extent, breeding system functions as a hidden trait in the ploidy-centered analysis, and vice versa (\cref{suppfigure:lumpedDP,suppfigure:lumpedIC}).
+But there is also strong statistical support for a hidden trait beyond ploidy and breeding system (\cref{supptable:testaddhidden}).
 One temptation is to propose specific possible traits to add into the analysis.
@@
 % alpha is always low: new supp fig showing how A & B are inferred on the tree?  at least for ID/CD/CP+A/B model
-Alternatively, we could view the hidden trait as merely a statistical trick, providing an easy way to model extra heterogeneity without directly representing a real trait.
+%Here Rosana adds the value of adding traits with the lumped models
+When analyzing the value of adding more traits to a complex model, our suggestion is to opt for the use of lumped models (\cref{figure:lumped,table:lumped}), and verify that adding a trait is statistically preferred.
+In the absence of additional information, the hidden states can be view as merely a statistical trick, providing an easy way to model extra heterogeneity without directly representing a specific trait.
+In our analyses, the transition rates between and within hidden traits did not affect the direction of the net diversification difference.
+However, we found that it is important to allow for the hidden state transition rates to be asymmetrical since those models have a better fit over models that fix the hidden state transition rates to be equal (\cref{suppfigure:asymmetric,supptable:asymmetry}).
 
@@
 
-Polyploidization is known to be common in plants, but the commonness of the reverse process---diploidization---is under active investigation.
+Polyploidization is known to be common in plants, but the commonness of the reverse process---diploidization---is under active investigation \citep{dodsworth_2015,mandakova_2018}.
 %
-In our analyses, ignoring diploidization if it is common could cause us to underestimate the polyploidization rate, overestimate of the diploid diversification rate, or otherwise confound our conclusions.
-We therefore included a slate of models with a diploidization parameter, and they showed that our main conclusions are robust to this process.
-These models also suggested modest statistical support for diploidization occuring within Solanaceae \cref{supptableFIXME}, although our estimates of the diploidization rate were highly uncertain.
+Ignoring diploidization if it is common, could cause us  for models that ignore it to underestimate the polyploidization rate and overestimate of the diploid diversification rate leading to incorrect conclusions.
+Therefore, we included a slate of models with a diploidization parameter, and they showed that our main conclusions are robust to this process (\cref{suppfigure:alldip}).
+These models also suggested modest statistical support for diploidization occuring within Solanaceae (\cref{supptable:testdiploidization}), although our estimates of the diploidization rate were highly uncertain.
 Furthermore, additional lines of evidence for classifying species as diploid or polyploid (beyond the genus chromosome multiplicity that we primarily relied on) would be needed for more reliable conclusions from an analysis such as ours.

### 2019-07-16 manuscript/introduction.tex
Rosana edited the whole thing

@@
 I've edited the rest of the manuscript so that "trait" is retained where ambiguity does not matter, but where directly contrasted (vs.: as here), I changed the wording.
-Dramatic increases in phylogenetic and phenotypic data along with methodological advances have greatly accelerated the search for traits that influence diversification traits. 
+Dramatic increases in phylogenetic and phenotypic data along with methodological advances have greatly accelerated the search for traits that influence diversification. 
 %B: was: "phenotypic data and methodological advances"
@@
 Consequently, examining the association of only one character with diversification patterns can be misleading. 
-Here, we embrace this challenge by jointly investigating two characters thought to influence speciation and extinction rates---ploidy level and breeding system---while allowing for their interactions and other confounding factors.
+We embrace this challenge by jointly investigating two characters thought to influence speciation and extinction rates---ploidy level and breeding system---while allowing for their interactions and other confounding factors to change diversification.
 % E: We need a "Here, we..." here.  Otherwise, we've just told the reader "not one trait!" and then launch into one trait.
@@
 %B: Separately, "common" and "numerous" are synonymous as used here (I removed "common" because that term also has the "unrefined" meaning).
-Polyploids were generally found to have an overall lower diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
-Many recent genomic studies, however, find numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications may have played an important macroevolutionary role driving innovation and diversification in plants.
+Polyploids were  found to have an overall lower net diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
+However, many recent genomic studies,  have found numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications may have played an important macroevolutionary role driving innovation and diversification in plants.
 Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies pervasive diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
-Our analyses re-examine the association between ploidy and lineage diversification by extending the approach of \citet{mayrose_2011, mayrose_2015} to include transitions between diploid and polyploid states.
-We also allow for heterogeneous diversification rates within each ploidy level, which could examine whether, for example, diploids with ancestral polyploidy diversify more rapidly than diploids without it.
+Our analyses re-examine the association between ploidy and lineage diversification by extending the approach of \citet{mayrose_2011, mayrose_2015} by including transitions from polyploid to diploid states and potential unobserved factors affecting diversification patterns.
+%R- No we don't do the below thing and it was confusing that's what Jacob Landis did in 2018
+%We also allow for heterogeneous diversification rates within each ploidy level, which could examine whether, for example, diploids with ancestral polyploidy diversify more rapidly than diploids without it.
 
@@
 Although these findings suggest a consistent macroevolutionary role of breeding system, it is unlikely to be the sole character determining lineage diversification.
-We thus investigate the relationship of breeding system to speciation and extinction rates in the context of ploidy and other background factors.
+We investigate the relationship of breeding system to speciation and extinction rates in the context of ploidy and other unobserved factors.
 
@@
 We start by fitting binary state speciation and extinction models to each ploidy and breeding system independently \citep{maddison_2007}. We follow by 
-fitting models that incorporate hidden states, which reduce the chance that an effect of one trait on diversification would be mistaken for the effect of an unobserved factor that is changing the diversification pattern \citep{beaulieu_2016}.
-We compare the proposed models against their character independent counterparts \citep{beaulieu_2016} to investigate whether ploidy or breeding system are actually linked to the diversification process.
+fitting models that incorporate hidden states, which reduce the chance that an effect of one trait on diversification would be mistaken for the effect of an unobserved factor \citep{beaulieu_2016}.
+We compare the proposed models against their character independent counterparts \citep{beaulieu_2016} to investigate whether ploidy or breeding system are actually informative of the diversification process.
 Next, we model ploidy and breeding system jointly to assess their combined influence on diversification, with or without an additional hidden character.
-With the joint trait model without hidden state, we also quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
-Later, we propose new models that aggregate states in the ploidy and breeding system models to make them equivalent to the single trait models.
+Using thw ploidy and breeding system model  without hidden state, we quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
+Later, we propose new models that aggregate states in the ploidy and breeding system models to make them equivalent to the binary state models.
 The new `lumped' models allow us to compare whether increasing model complexity from one trait to two traits significantly improves our understanding of the diversification process.
@@
 % E: say something here about testing whether it's worth adding in another real trait?
-Then, we extend all the models involving ploidy, to investigate the potential effects of including diploidization.
-Finally, we explore models of diversification using asymmetric rates between and within hidden states to investigate the sensitivity of the statistical inference to model assumptions.
+Furthermore, we extend all the models involving ploidy, to investigate the potential effects of including diploidization.
+And finally, we explore models of diversification using asymmetric rates between and within hidden states to investigate the sensitivity of the statistical inference to model assumptions.
 
-Our results highlight the importance of considering non-additive effects of traits on net diversification rates, especially when there are strong biologically-driven correlations among them.
+Our results highlight the importance of considering non-additive effects of traits on net diversification rates  under the presence of unobserved factors, in order to detect strong biologically-driven processes dictating the diversification patterns.

### 2019-07-16 manuscript/abstract.tex
minor edits in Abstract and Introduction

@@
 Identifying traits associated with diversification rate differences is complicated by the wide variety of characters under consideration and the statistical challenges of testing for associations from comparative phylogenetic data.
-%B: complicated by <- tricky, however, because of ...
-to consider; was %Identifying traits associated with diversification rate differences is tricky, however, because of the wide variety of traits to consider and the statistical challenges of testing for associations from comparative phylogenetic data.
-Ploidy (diploid vs.\ polyploid states) and breeding system (self-incompatible vs.\ self-compatible states) have been repeatedly suggested as possibly drivers of differential diversification.
+Ploidy (diploid \vs polyploid states) and breeding system (self-incompatible \vs self-compatible states) have been repeatedly suggested as possible drivers of differential diversification.
 We investigate the connections of these traits, including their interaction, to speciation and extinction rates in Solanaceae.
 We show that the effect of ploidy on diversification can be largely explained by its correlation with breeding system and that additional unknown factors, alongside breeding system, determine diversification rates.
-%B: was: We find that the effect of ploidy can largely be explained by its correlation with breeding system, and that additional unknown factors work with breeding system to determine diversification rates.
-These results are largely robust to allowing for diploidization. %B: left along, but I prefer: "These results are largely robust to inclusion of diploidization."
-Finally, we find that the direct pathway, in which breakdowns of self-incompatibility arise as byproducts of polyploidization, rather than a step-wise pathway with breakdowns followed by polyploidization, is disproportionately important in the evolutionary history of the nightshade family.
-%B: was: "from self-incompatible diploids lower rate of diversification associated with self-compatible diploids makes it less likely for that state to lie along the evolutionary pathway from self-incompatible diploids to polyploids."
+These results are largely robust to allowing for diploidization.
+Finally, we find that the evolutionary pathway to polyploidy in Solanaceae is most often via the direct breakdown of self-incompatibility by whole genome duplication, rather than indirectly via breakdown followed by polyploidization.

### 2019-07-16 manuscript/introduction.tex
minor edits in Abstract and Introduction

@@
 Species accumulate across the tree of life at different rates. 
-One possible explanation for this phenomenon is that the various traits differentially affect rates of diversification.
+One possible explanation for this phenomenon is that various traits differentially affect rates of diversification.
 %B: note: traits and character states are synonymous.
@@
 %B: Separately, "common" and "numerous" are synonymous as used here (I removed "common" because that term also has the "unrefined" meaning).
-Polyploids were  found to have an overall lower net diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
-However, many recent genomic studies,  have found numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications may have played an important macroevolutionary role driving innovation and diversification in plants.
+Polyploids were found to have an overall lower net diversification rate than diploids across many vascular plant clades \citep{mayrose_2011, mayrose_2015}. 
+However, recent genomic studies have found numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications may have played an important macroevolutionary role driving innovation and diversification in plants.
 Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies pervasive diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
-Our analyses re-examine the association between ploidy and lineage diversification by extending the approach of \citet{mayrose_2011, mayrose_2015} by including transitions from polyploid to diploid states and potential unobserved factors affecting diversification patterns.
+Our analyses re-examine the association between ploidy and lineage diversification by extending the approach of \citet{mayrose_2011, mayrose_2015} to include transitions from polyploid to diploid states and potential unobserved factors affecting diversification patterns.
 %R- No we don't do the below thing and it was confusing that's what Jacob Landis did in 2018
@@
 Second, we examine the order of transitions in the two characters.
-Even though evolution may commonly proceed from diploid to polyploid, and from SI to SC states, there are two paths by which diploid SI lineages can eventually become SC polyploids.
+Evolution commonly proceeds from diploid to polyploid, and from SI to SC states, but there are two paths by which diploid SI lineages can eventually become SC polyploids.
 Loss of SI in diploids could be directly caused by polyploidization (as explained above, for RNase-based SI systems), resulting in a one-step pathway to SC polyploids. 
 Alternatively, SI diploids could first transition to SC without an increase in ploidy, and subsequently undergo polyploidization, resulting in a two-step pathway to SC polyploids.
-\citet{robertson_2011} compared the contributions of these two paths, finding that evolution from SI diploids to SC polyploids is more likely to proceed via the one-step pathway, over short timescales, but via the two-step pathway over long timescales.
+\citet{robertson_2011} compared the contributions of these two paths, finding that evolution from SI diploids to SC polyploids is more likely to proceed via the one-step pathway over short timescales, but via the two-step pathway over long timescales.
 They considered only transitions among the states, however, so we investigate whether these results hold true when allowing for differences in lineage diversification.
@@
 %B: redundant with "very powerful" below rm: Examining these questions requires powerful phylogenetic comparative methods.
-In the present study, we employ an extended framework of state speciation and extincion models (SSE), which simultaneously models transitions between the discrete states of a trait and different rates of speciation and extinction associated with each of those states \citep{maddison_2007, fitzjohn_2012}.
-All the models proposed for this study are coded as graphical models and Bayesian inference was performed over the proposed models using RevBayes software \citep{hoehna_2016}.
+In the present study, we employ an extended framework of state speciation and extinction models, which simultaneously model transitions between the discrete states of a trait and different rates of speciation and extinction associated with each of those states \citep[`SSE' models;][]{maddison_2007, fitzjohn_2012}.
+% All the models proposed for this study are coded as graphical models and Bayesian inference was performed over the proposed models using RevBayes software \citep{hoehna_2016}.
+% E: The models aren't drawn graphically in any of the figures (not in the revbayes sense, at least).
+And we don't need methods in the intro.
 %Although potentially very powerful, this approach can be prone to reporting a false association between a trait and lineage diversification when the data contain additional, unmodeled diversification rate heterogeneity \citep{maddison_2015, rabosky_2015}.
-We start by fitting binary state speciation and extinction models to each ploidy and breeding system independently \citep{maddison_2007}. We follow by 
-fitting models that incorporate hidden states, which reduce the chance that an effect of one trait on diversification would be mistaken for the effect of an unobserved factor \citep{beaulieu_2016}.
+We start by fitting binary state speciation and extinction models to each ploidy and breeding system independently \citep{maddison_2007}.
+We follow by fitting models that incorporate hidden states, which reduce the chance that an effect of one trait on diversification would be mistaken for the effect of an unobserved factor \citep{beaulieu_2016}.
 We compare the proposed models against their character independent counterparts \citep{beaulieu_2016} to investigate whether ploidy or breeding system are actually informative of the diversification process.
 Next, we model ploidy and breeding system jointly to assess their combined influence on diversification, with or without an additional hidden character.
-Using thw ploidy and breeding system model  without hidden state, we quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
-Later, we propose new models that aggregate states in the ploidy and breeding system models to make them equivalent to the binary state models.
-The new `lumped' models allow us to compare whether increasing model complexity from one trait to two traits significantly improves our understanding of the diversification process.
+Using the ploidy and breeding system model without hidden states, we quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
+We also lump together states within these joint models of ploidy and breeding system in order to test whether increasing complexity from one trait to two traits significantly improves our understanding of the diversification process.
 %B: I altered this line, but it is still a bit redundant given previous sentence.
@@
 %Second, we add a hidden character to each model and investigate if state-dependent diversification in focal character retains its explanatory power, or whether the added hidden character obscures it.
-%
-% E: say something here about testing whether it's worth adding in another real trait?
-Furthermore, we extend all the models involving ploidy, to investigate the potential effects of including diploidization.
-And finally, we explore models of diversification using asymmetric rates between and within hidden states to investigate the sensitivity of the statistical inference to model assumptions.
-
+Furthermore, we extend all the models involving ploidy to investigate the potential effects of including diploidization.
+% And finally, we explore models of diversification using asymmetric rates between and within hidden states to investigate the sensitivity of the statistical inference to model assumptions.
 Our results highlight the importance of considering non-additive effects of traits on net diversification rates  under the presence of unobserved factors, in order to detect strong biologically-driven processes dictating the diversification patterns.

### 2019-07-16 manuscript/methods.tex

@@
 
-Chromosome number data were obtained for all Solanaceae taxa in the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were curated semi-automatically using the CCDBcurator R package \citep{rivero_2019}.
+Chromosome number data were obtained for all Solanaceae taxa in the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were curated semi-automatically using the \mbox{CCDBcurator} R package \citep{rivero_2019}.
 CCDB contains records from original sources that have multiple complex symbol patterns denoting multivalence, or irregularites of chromosome counts.
@@
 
-Breeding system was scored as self-incompatible, $I$, or self-compatible, $C$, based on results curated from the literature and original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
+Breeding system states were scored as self-incompatible, $I$, or self-compatible, $C$, based on results curated from the literature and original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
 Most species could unambiguously be coded as either $I$ or $C$ \citep{raduski_2012}.
@@
 Following the reasoning outlined in \citet{robertson_2011}, we closely examined the few species for which the merged ploidy and breeding system data indicated the presence of self-incompatible polyploids.
-%B: Modified this section for post-review response:
 Although SI populations frequently contain some SC individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find convincing data for a naturally occurring SI polyploid population  \citep[discussed in][]{robertson_2011}.
@@
 
-The Supplementary Information contains citations for the numerous original data sources. % FIXME (and say here how many refs?)
+The Supplementary Information contains citations for the numerous original data sources. % FIXME (and say here how many refs?) %B: no need for an additional counting and updating headache?
 The Dryad archive contains the data and tree files used for analyses. % FIXME
@@
 SSE models parameters that describe per-lineage rates of speciation and extinction, specific to each character state (denoted $\lambda$ and $\mu$, respectively, with subscripts to indicate the state), along with rates of transitions between states (denoted $\rho$ for polyploidization, $\delta$ for diploidization, and $q_{IC}$ for loss of self-incompatibility).
-
 The full set of models and all their rate parameters are detailed in \cref{fig:allmodels}.
-Here we summarize how each type of model allows us to assess whether diversification is best explained by ploidy, breeding system, their combination, or some unknown factor.
+Here, we summarize how each model allows us to assess whether diversification is best explained by variation in ploidy, breeding system, their combination, or some unknown factor.
 
@@
 
-In order to investigate the association between ploidy level and diversification, we first defined a model in which taxa were classified as diploid $(D)$ or polyploid $(P)$ (model M1).
-This is the same as one of the models presented in  \citet{mayrose_2011}, which was used for the exact same question across an array of plant genera.
-% R- Itay wanted to hightlight that is the same model. Definitely a BiSSE. He did the ClaSSE later in the 2014 response.
-Although this model can be a powerful means to distinguish state-specific speciation and extinction from trait changes, it is prone to incorrectly reporting that a trait is associated with diversification differences \citep{maddison_2015, rabosky_2015}.
+In order to investigate the association between ploidy level and diversification, we first employed a model  (M1), previously used by \citet{mayrose_2011}, with each taxon classified as diploid $(D)$ or polyploid $(P)$.
+Although this model can be powerful in studies of trait evolution, it is prone to incorrectly reporting that a trait is associated with diversification differences \citep{maddison_2015, rabosky_2015}.
 We therefore define several models that incorporate additional forms of diversification rate heterogeneity.
 
-Our second ploidy model includes a binary hidden trait that subdivides each observed state (M2).
-In this character independent model known as CID \citep{beaulieu_2016}, the hidden trait can affect diversification but the observed trait does not.
+The second ploidy model (M2) includes a binary hidden trait that subdivides each observed state.
+In this trait-independent model known as CID \citep{beaulieu_2016}, hidden traits can affect diversification but the observed traits do not.
 Comparing M1 and M2 allows us to test whether diversification rate heterogeneity is better explained by ploidy or by some unknown factor.
 
-We additionally fit three models in which both ploidy and a hidden trait could influence diversification (M3--M5).
+We additionally fit three models in which both ploidy and a hidden states could influence diversification (M3--M5).
 These models differ in whether transitions between the hidden states are symmetric or asymmetric, and whether the polyploidization rate depends on the hidden state.
@@
 We propose five breeding system models following the same logic as the first ploidy models.
-Under the simplest breeding system and diversification model,  species are classified as self-incompatible $(I)$ or self-compatible $(C)$ (M11).
-This is the same model as the analysis presented in \citet{goldberg_2010} but with an updated phylogeny \citep{sarkinen_2013} and larger aggregated dataset.
+Under the simplest breeding system and diversification model (M11), species are classified as self-incompatible $(I)$ or self-compatible $(C)$.
+This is the same model as in the analysis presented in \citet{goldberg_2010} but with an updated phylogeny \citep{sarkinen_2013} and a larger aggregated dataset.
 
@@
 
-The simplest model classifies each species as either self-incompatible diploid $(ID)$, self-compatible diploid $(CD)$, or self-compatible polyploid $(CP)$; recall that self-incompatible polyploids do not occur (M16).
+The simplest model classifies each species as either SI diploid $(ID)$, SC diploid $(CD)$, or SC polyploid $(CP)$; recall that SI polyploids do not occur (M16).
 Each of these states may again be associated with different rates of speciation and extinction, and the allowable transitions are loss of SI within the diploid state (from $ID$ to $CD$), loss of SI via polyploidization (from $ID$ to $CP$), and polyploidization while SC (from $CD$ to $CP$).
@@
 
-As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by both ploidy and breeding system as well as a hidden trait (M18--M20, with varying degrees of complexity in the hidden trait transitions \cref{supptable:M16M20})
+As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by both ploidy, breeding system, as well as a hidden trait (M18--M20) with varying degrees of complexity in the hidden trait transitions (\cref{supptable:M16M20}).
 We add this hidden trait layer on top of our three-state model \citep[similar to][]{caetano_2018, huang_2018}.
@@
 
-The models described so far allow us to assess the contributions of our two focal traits---ploidy and breeding system---to lineage diversification.
-They do not, however, reveal whether it is `worth it' to include both traits in the analysis.
-Model comparisons to test whether diversification heterogeneity is better explain by both traits than by either one require the calculation of the likelihood function of the data given an specific model.
-However, the likelihood of the data given a model differs in the two-state versus the three-state models because the proposed models use different state spaces: $D/P$ versus $I/C$ versus $IC/CD/CP$.
-Therefore, different data results in incomparable models.
+The models described so far allow us to assess the contributions of our two focal characters---ploidy and breeding system---to lineage diversification, but they do not reveal whether it is valuable to include both characters in the analysis.
+Assessing whether diversification heterogeneity is better explained by simultaneous consideration of both characters, rather than either one alone, requires the calculation of the likelihood function of the data given a specific model.
+However, the likelihood of the data given a model differs---in our case, two-state versus three-state models---because the proposed models use different state spaces: $D/P$ versus $I/C$ versus $IC/CD/CP$.
+Therefore, the use of different data results in incomparable models.
 
-In order to compare model fits for ploidy-only versus breeding system-only versus both traits, we extend a new technique proposed in ontological work by \citet{tarasov_2019} that allows for state `lumping' .
+In order to compare fits of ploidy-only vs. breeding system-only vs. combined trait models, we extend a  technique proposed in the work by \citet{tarasov_2019} for state `lumping.'
 We use the state space of the $IC/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
@@
 Then we lump together $CD$ and $CP$ to form the self-compatible state, mimicking the $I/C$ model (M28).
-We further add a hidden trait to each of these models (M27 and M29).
-We compared those models in table \cref{table:lumped}
+We further add a hidden character to each of these models (M27 and M29), and the compare this group of models in table \cref{table:lumped}
 
-We do not include additional models with diploidization because this reverse transition renders the models non-lumpable \citep{tarasov_2019}.
-This means, that to lump two states a necessary and sufficient condition is that the rates going from the lumped state to the singular state are equal. 
+We do not include additional models with diploidization because this reverse ploidy transition renders the models non-lumpable \citep{tarasov_2019}.
+To lump two states, it is a necessary and sufficient condition that the transition rates from the lumped state to the singular state are equal  \citet{tarasov_2019}.
 When lumping compatible state $CD$ and $CP$ the the rate at which they back to state $ID$ is zero for both, thus, compatible states are lumpable. 
@@
 Both of our methods are based on a propagation matrix that describes flow from $ID$ to $CP$, as in \citet{robertson_2011}.
-We insert an artificial division in the $CP$ state, so that one substate contains the $CP$ species that arrived via the one-step pathway and the other substate contains the $CP$ species that arrived via the two-step pathway.
+We insert an artificial division in the $CP$ state, so that one sub-state contains the $CP$ species that arrived via the one-step pathway and the other substate contains the $CP$ species that arrived via the two-step pathway.
 We consider unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and no strong support for diploidization (see below).
@@
 In this case, however, these are not probabilities because diversification changes the number of lineages as time passes.
-We can still use their ratios to consider the relative contribution of each pathway, though, analogous to the normalized age structure in a growing population.
+%B: FIXME, we'll hear complaints again about "this," "these," and ...
+"their." (below)
+We can still use their ratios to consider the relative contribution of each pathway, analogous to the normalized age structure in a growing population.
 % R- here a reference would be nice  E: Does Jake have a favorite ref?
 Otherwise, I guess a basic ecology textbook.
-
+%
 % E & R todo ?: for the case with net div, write out the propagation matrix and equations for pathway contributions
-
+%
 % E: I don't think we need this section now that diploidization is in each of the sections above.  But I'm saving this text in case we want to put some of it in Results.
@@
 Parameters for each of the 29 models were estimated using RevBayes \citep{hoehna_2016}.
-Code for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.
+Scripts for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.
 We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $651/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
@@
 Each species retained on the tree belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known.
-(A) The five categories and the number of species in each.
-Abbreviations are $I$ for self-incompatible, $C$ for self-compatible, $D$ for diploid, $P$ for polyploid, $?$ for unknown.
-Because polyploidization breaks this form of self-incompatibility, SI species with unobserved ploidy ($I?$) are taken to be diploid ($ID$), and polyploid species with unobserved breeding system ($?P$) are taken to be SC.
+Character state abbreviations: $I$ for self-incompatible, $C$ for self-compatible, $D$ for diploid, $P$ for polyploid, $?$ for unknown.
+(A) The five categories and the number of species in our data for each category.
+Because polyploidization breaks this form of self-incompatibility, self-incompatible species with unobserved ploidy ($I?$) are taken to be diploid ($ID$), and polyploid species with unobserved breeding system ($?P$) are taken to be SC.
 (B) Category groupings into states for each model type.
@@
 Results indicate that the MuHiSSE model with asymmetric hidden rates (17) is  strongly preferred over M16 and M18 and moderately preferred over the MuHiSSe with all rates asymetric (M20).
-*Marginal log-likelihood for M17 could not be calculated in our limited amount of computational time.}
+*Marginal log-likelihood for M17 could not be calculated within allotted computer time.}
 \label{supptable:M16M20}
@@
 \caption{Bayes factors of ploidy and breeding system with diploidization.
-Results indicate that the MuHiSSE model with asymmetric hidden rates (M24) is  strongly preferred over M21-M23 and moderately preferred over the MuHiSSe with all rates asymetric (M25).}
+Results indicate that the MuHiSSE model with asymmetric hidden rates (M24) is  strongly preferred over M21-M23 and moderately preferred over the MuHiSSe with all rates asymmetric (M25).}
 \label{supptable:M21M25}
@@
 \end{tabular}
-\caption{Test for inclusion of a diploidization rate via Bayes factors. Models with diploidzation are moderately preferred over models that do not include a diploidization rate}
+\caption{Test for inclusion of a diploidization rate via Bayes factors. Models with diploidization are moderately preferred over models that do not include a diploidization rate}
 \label{supptable:testdiploidization}

### 2019-07-17 manuscript/methods.tex
small edits in Methods

@@
 
-In order to test our hypotheses about lineage diversification and trait macroevolution, we fitted 29 state-dependent speciation and extinction models (SSE) to previous work on diversification \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016}.
-SSE models parameters that describe per-lineage rates of speciation and extinction, specific to each character state (denoted $\lambda$ and $\mu$, respectively, with subscripts to indicate the state), along with rates of transitions between states (denoted $\rho$ for polyploidization, $\delta$ for diploidization, and $q_{IC}$ for loss of self-incompatibility).
+In order to test our hypotheses about lineage diversification and trait macroevolution, we fit 29 state-dependent speciation and extinction models \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016}.
+SSE models contain parameters that describe per-lineage rates of speciation and extinction, specific to each character state (denoted $\lambda$ and $\mu$, respectively, with subscripts to indicate the state), along with rates of transitions between states (denoted $\rho$ for polyploidization, $\delta$ for diploidization, and $q_{IC}$ for loss of self-incompatibility).
 The full set of models and all their rate parameters are detailed in \cref{fig:allmodels}.
@@
 
-In order to investigate the association between ploidy level and diversification, we first employed a model  (M1), previously used by \citet{mayrose_2011}, with each taxon classified as diploid $(D)$ or polyploid $(P)$.
+In order to investigate the association between ploidy level and diversification, we first employed a model (labeled M1), previously used by \citet{mayrose_2011}, with each species classified as diploid $(D)$ or polyploid $(P)$.
 Although this model can be powerful in studies of trait evolution, it is prone to incorrectly reporting that a trait is associated with diversification differences \citep{maddison_2015, rabosky_2015}.
@@
 
-We additionally fit three models in which both ploidy and a hidden states could influence diversification (M3--M5).
+We additionally fit three models in which both ploidy and a hidden trait could influence diversification (M3--M5).
 These models differ in whether transitions between the hidden states are symmetric or asymmetric, and whether the polyploidization rate depends on the hidden state.
@@
 
-We propose five breeding system models following the same logic as the first ploidy models.
+We propose five breeding system models following the same logic as the ploidy models above.
 Under the simplest breeding system and diversification model (M11), species are classified as self-incompatible $(I)$ or self-compatible $(C)$.
@@
 Furthermore, species that are distantly related within this family carry closely-related alleles, with deep trans-specific polymorphism at the locus that controls the SI response \citep{ioerger_1990, igic_2006}.
-The breeding system in Solanaceae has strong evidence that the $I$ state is ancestral in the family, and that the SI mechanism was not regained.
+Thus, there is strong evidence in Solanaceae that the $I$ state is ancestral in the family, and that the SI mechanism was not regained.
 Therefore, for all breeding system models, we estimated a transition rate from $I$ to $C$  but not the reverse ($q_{CI}=0$).
@@
 
-The simplest model classifies each species as either SI diploid $(ID)$, SC diploid $(CD)$, or SC polyploid $(CP)$; recall that SI polyploids do not occur (M16).
+The simplest model (M16) classifies each species as either SI diploid $(ID)$, SC diploid $(CD)$, or SC polyploid $(CP)$; recall that SI polyploids do not occur.
 Each of these states may again be associated with different rates of speciation and extinction, and the allowable transitions are loss of SI within the diploid state (from $ID$ to $CD$), loss of SI via polyploidization (from $ID$ to $CP$), and polyploidization while SC (from $CD$ to $CP$).
@@
 
-As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by both ploidy, breeding system, as well as a hidden trait (M18--M20) with varying degrees of complexity in the hidden trait transitions (\cref{supptable:M16M20}).
+As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by ploidy, breeding system, and a hidden trait (M18--M20) with varying degrees of complexity in the hidden trait transitions (\cref{supptable:M16M20}).
 We add this hidden trait layer on top of our three-state model \citep[similar to][]{caetano_2018, huang_2018}.
@@
 The models described so far allow us to assess the contributions of our two focal characters---ploidy and breeding system---to lineage diversification, but they do not reveal whether it is valuable to include both characters in the analysis.
-Assessing whether diversification heterogeneity is better explained by simultaneous consideration of both characters, rather than either one alone, requires the calculation of the likelihood function of the data given a specific model.
-However, the likelihood of the data given a model differs---in our case, two-state versus three-state models---because the proposed models use different state spaces: $D/P$ versus $I/C$ versus $IC/CD/CP$.
+To answer this question with statistical model comparisons requires comparing the likelihood of the data given each model.
+This is impossible for the ploidy and breeding system models presented so far, however, because the data are different for the different models: they use either the $D/P$ or the $I/C$ or the $IC/CD/CP$ state spaces.
 Therefore, the use of different data results in incomparable models.
 
-In order to compare fits of ploidy-only vs. breeding system-only vs. combined trait models, we extend a  technique proposed in the work by \citet{tarasov_2019} for state `lumping.'
+In order to compare fits of ploidy-only \vs breeding system-only \vs combined trait models, we use the technique of `lumping' states together \citep{tarasov_2019}.
 We use the state space of the $IC/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
@@
 Then we lump together $CD$ and $CP$ to form the self-compatible state, mimicking the $I/C$ model (M28).
-We further add a hidden character to each of these models (M27 and M29), and the compare this group of models in table \cref{table:lumped}
+We further add a hidden character to each of these models (M27 and M29), and then compare this group of models (\cref{table:lumped}).
 
-We do not include additional models with diploidization because this reverse ploidy transition renders the models non-lumpable \citep{tarasov_2019}.
-To lump two states, it is a necessary and sufficient condition that the transition rates from the lumped state to the singular state are equal  \citet{tarasov_2019}.
-When lumping compatible state $CD$ and $CP$ the the rate at which they back to state $ID$ is zero for both, thus, compatible states are lumpable. 
-Then, we define a single $q_0$ parameter as the new rate from incompatible to compatible going into the lumped state $(CD,CP)$.
-By doing so, we can define the lumped model M28 (\cref{figure:lumped}E) that is equivalent to M1 but comparable to M16.
-We repeat this process to create the lumped model M26for $D/P$ by restricting the polyploidy rate to be $rho_0$ out of the lumped state $(ID,CD)$
-For the diploidization models, $CP$ goes back to $CD$ with rate $\delta$ but the rate from $CP$ back to $ID$ is 0.
-Since $\delta \neq 0$, the models with diplodization cannot be lumped to be compared to the tree-state models.
-% E: further explanation needed? R- I added a simple math explanation. Let me know if it is okay
+We do not include additional models with diploidization because this reverse ploidy transition renders the models non-lumpable.
+To lump states requires that the transition rates from the lumped state to the singular state be equal \citep{tarasov_2019}.
+When including diploidization, transitions from $CP$ to $CD$ are at rate $\delta$ but transitions from $CP$ to $ID$ do not occur.
+Because $ID$ and $CD$ would be lumped to mimic the $D/P$ model, this model is non-lumpable when $\delta \ne 0$.
+When lumping $CD$ and $CP$, however, the transition rate to $ID$ is zero for both, so the SC states are lumpable. 
+Thus, we can compare models to test whether it is advantageous to include both traits, but only when ignoring diploidization.
+% E: Below was all good, but it seemed like more detail than necessary.
+% Then we define a single parameter $q_0$ going into the lumped state $(CD,CP)$ as the new rate from SI to SC.
+% This defines the lumped model M28 (\cref{figure:lumped}E) that is equivalent to M1 but comparable to M16.
+% We repeat this process to create the lumped model M26 for $D/P$ by restricting the polyploidy rate to be $rho_0$ out of the lumped state $(ID,CD)$.
+% When including diploidization, however, transitions from $CP$ to $CD$ are at rate $\delta$ but transitions from $CP$ to $ID$ do not occur.
+% Since $\delta \neq 0$, the models with diplodization cannot be lumped to be compared to the three-state models.
 
@@
 In the two-step pathway, the $CD$ state is an intermediate: SI is first lost, and later the SC diploid undergoes polyploidization.
-We quantify the relative contribution of these pathways to polyploidy in two ways, each using the maximum \textit{a posteriori} (MAP) estimates of rates from the simplest model that includes both traits (M16).
+We quantify the relative contribution of these pathways to polyploidy in two ways, each using the median estimates of rates from the simplest model that includes both traits (M16).
 
@@
 "their." (below)
-We can still use their ratios to consider the relative contribution of each pathway, analogous to the normalized age structure in a growing population.
+We can still use ratios, however, to consider the relative contribution of each pathway, analogous to the normalized age structure in a growing population.
 % R- here a reference would be nice  E: Does Jake have a favorite ref?
@@
 For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the models.
-% FIXME see referee comment about reporting convergence, not runtime
 Convergence and mixing of each MCMC chain was determined by ensuring the effective sample size of each parameter was over 200.
@@
 (\cref{suppfigure:DPnodipasr,suppfigure:DPnodipABasr,suppfigure:ICasr,suppfigure:ICABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPnodipABasr}).
-% FIXME: I'm confused about if these are joint or marginal?
-A bit more explanation would help.
-R: Marginal explanation added
 
@@
 We calculated the marginal log-likelihood for each of the models using fifty stepping stone steps under the methodology of \citet{xie_2010}, implemented in RevBayes \citep{hoehna_2016}.
-Each stepping stone step was found by calculating at least 500 generations of burn-in followed by a total of 1,000 MCMC steps (\cref{table:marginallike}).
+Each stepping stone step was found by calculating at least 500 generations of burn-in followed by a total of 1,000 MCMC steps. % (\cref{table:marginallike}) updated ref?
 
-Using the marginal likelihood values, we calculated Bayes factors to argue five key biological and methodological questions:
+Using the marginal likelihood values, we calculated Bayes factors to answer five key biological and methodological questions:
 \begin{enumerate}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{states.pdf} % FIXME: update numbers
+\includegraphics[width=0.5\textwidth]{states.pdf}
 \caption{
@@
 (A) The five categories and the number of species in our data for each category.
-Because polyploidization breaks this form of self-incompatibility, self-incompatible species with unobserved ploidy ($I?$) are taken to be diploid ($ID$), and polyploid species with unobserved breeding system ($?P$) are taken to be SC.
+Because polyploidization breaks this form of self-incompatibility, self-incompatible species with unobserved ploidy ($I?$) are taken to be diploid ($ID$), and polyploid species with unobserved breeding system ($?P$) are taken to be SC ($CP$).
 (B) Category groupings into states for each model type.
-For example, there are 70 species that are SC with unknown ploidy. % Fixed
-In the $D/P$ models these are coded as either $D$ or $P$ (uncertain/consistent with either state), in the $I/C$ models they are coded as $C$, and in the $ID/CD/CP$ models they are coded as either $CD$ or $CP$.
+For example, there are 70 species that are SC with unknown ploidy.
+In the $D/P$ models these are coded as $D$ and $P$ (uncertain/consistent with either state), in the $I/C$ models they are coded as $C$, and in the $ID/CD/CP$ models they are coded as $CD$ and $CP$.
 In the hidden state models, all species were coded as consistent with both $A$ and $B$.
-Two species \textit{Lycium californicum} and \textit{Solanum bulbocastanum} are simultaneously ID and CP and by adding them the sample adds to the total of 651 taxa used for analyses.
+Two species, \emph{Lycium californicum} and \emph{Solanum bulbocastanum}, are simultaneously $ID$ and $CP$, and by adding them the sample adds to the total of 651 taxa used for analyses.
 }

### 2019-07-17 manuscript/discussion.tex
Discussion edits down to Diploidization section (including related changes in other sections)

--- 2019-07-12 manuscript/discussion.tex
@@
 %
-Species are composed of vast assemblages of variable traits, many traits are heritable, and many traits could affect the propensity of species to perish or multiply \citep{lewontin_1970}.
+Species are composed of vast assemblages of variable traits.
+Many traits are both heritable and possibly affect the propensity of species to perish or multiply \citep{lewontin_1970}.
+%B: I subtly altered the previously second sentence and split into two, to link heritable traits with those that affect diversification.
+The implication was not clear---"many are heritable and many affect fitness" is not necessarily that some traits fall under both categories.
+%was: Species are composed of vast assemblages of variable traits, many traits are heritable, and many traits could affect the propensity of species to perish or multiply \citep{lewontin_1970}.
 Examining the effects of complex trait combinations on lineage diversification, however, remains challenging.
@@
 % E: should this results summary be much shorter?
-Previous analyses of the effects of ploidy on diversification showed that diploids had greater net diversification rates than polyploids across many angiosperm clades \citep{mayrose_2011, mayrose_2015}.
-We obtain a similar outcome when examining ploidy alone in Solanaceae (\cref{figure:netdivall}A), but a consistent effect of ploidy on diversification is not clear when we also allow that a hidden factor could affect diversification as well (\cref{figure:netdivall}B).
+% B: I think it's fine, but I switched it to present tense (not strictly, but almost) for both reasons of style (aesthetic preference, perhaps), to separate previous work from current work, and to avoid awkward implications about the process not being ongoing.
+Moreover, the use of "show" or "demonstrate" is very loose in the context of our weak inference resulting from--what we all hopefully agree--are potentially flawed models.
+Previous analyses of the effects of ploidy on diversification found that diploids are associated with greater net diversification rates than polyploids across many angiosperm clades \citep{mayrose_2011, mayrose_2015}.
+We obtain a similar outcome when examining ploidy alone in Solanaceae (\cref{figure:netdivall}A), but a consistent effect of ploidy on diversification is not clear when we incorporate a hidden factor linked to diversification (\cref{figure:netdivall}B).
 %
-Previous analyses of breeding system in this family showed that SI species are associated with higher diversification rates than their SC counterparts (\citealt{goldberg_2010}; \cref{figure:netdivall}C).
-Our new analyses with a hidden trait show that this pattern persists on different backgrounds.
-One on background, however, SC species experience positive net diversification (\ie not a dead end; cf.\ \citealt{igic_2013}) that exceeds the diversification rate of SI species on the other background (\cref{figure:netdivall}D).
+Previous analyses of breeding system in this family found that SI may cause higher diversification rates, compared with SC (\citealt{goldberg_2010}; \cref{figure:netdivall}C).
+Our analyses that include a hidden trait recover the same pattern, with one important difference.
+On the background of one hidden state, we recover a net diversification rate for SC species that is positive and greater than the diversification rate of SI species on the background of the other hidden state (\cref{figure:netdivall}D).%B: I rewrote this to separate the point of the side clause.
+Therefore, SC may not be a `dead end' (cf.\ \citealt{igic_2013}) when coupled with some unknown trait combinations (or processes not modeled by these two traits). 
+%Separately, I should note here that nothing about this analysis seems to me to strongly imply that SC is not a dead end in combination with a trait (as we discuss immediately below!).
+Much of the explanatory power derived for the A/B model may come from (for instance) accommodating tree scaling problems or rate heterogeneity in molecular evolution.
+It just fills up all of the mystery space without pointing to a cause.
+Side notes: (1) The I/C + A/B model ancestral states basically have the root of the tree as being in one hidden state and then tips in another.
+(2) Is there any place where we can see how much of the tree is in hidden state A or state B under some model?
 %
-Furthermore, our models of the joint macroevolution of ploidy and breeding system provided better statistical fits and revealed how the trait combinations associate with diversification.
-We find that the higher overall high net diversification rate of diploids is dominated by SI diploids, while SC diploids have a lower diversification rate that is comparable to polyploids, which are also SC (\cref{figure:netdivall}E).
-Thus, breeding system seems to provide the dominant explanation of diversification differences, with ploidy providing a secondary effect within SC species.
+Our analyses also reveal that models of joint evolution of ploidy and breeding system are statistically preferred, and hint at how various trait combinations may be linked with diversification.
+We find that the highest net diversification rate is associated with SI diploids, while SC diploids have a lower diversification rate that overlaps with the net diversification of SC polyploids (\cref{figure:netdivall}E).
+Thus, breeding system appears to provide a relatively better explanation of diversification rate differences, with ploidy providing a secondary effect within SC species.
 
@@
 
-Throughout our numerous model comparisons, models that included a hidden trait emerged as significantly better than those that did not. % supptable ref
-This is entirely consistent with the expectation that many processes can affect speciation and extinction, far beyond the one or two traits of particular interest to us.
-A natural follow-up question is, what is this hidden trait?
-Our results show that to some extent, breeding system functions as a hidden trait in the ploidy-centered analysis, and vice versa.
-But there is also strong statistical support for a hidden trait beyond ploidy and breeding system.
-One temptation is to propose specific possible traits to add into the analysis.
-For example, the rapidly-diversifying clade of mostly SC species within \emph{Solanum} suggests that geography may play a role. % Boris, please check!  Australian?  fixme
+Throughout our numerous model comparisons, we find that inclusion of hidden traits provides a considerably better fit (\cref{supptable:testaddhidden}). 
+This is consistent with the expectation that many processes, beyond those associated with the focal traits, can affect inference of speciation and extinction.
+%removed "one or two" because it's true for N traits.
+%A natural follow-up question is, what is this hidden trait?
+It is, however, unclear exactly which processes are captured by the hidden traits.
+For example, our results show that, to a varying extent, breeding system functions as a hidden trait in the ploidy-centered analysis, and vice versa (\cref{suppfigure:lumpedDP,suppfigure:lumpedIC}).
+But the strong statistical support for processes not well explained by ploidy and breeding system (\cref{supptable:testaddhidden}) tempts one to interpret the remaining variation as the effect of other measurable traits.
+For example, our data appear to show a rapidly-diversifying Australasian clade of mostly SC species within \emph{Solanum}, which suggests that geography may play a role.
+% Boris, please check!
+Australian?
+fixme
+Nevertheless, it is also possible that the addition of hidden traits instead explains variation stemming from any of a number of unrelated processes or methodological artifacts, such as variation in rates of molecular evolution or misspecified models of sequence evolution.
+%Rosana and Emma check me.
 %     but don't be circular, test in different clade
 % alpha is always low: new supp fig showing how A & B are inferred on the tree?  at least for ID/CD/CP+A/B model
-Alternatively, we could view the hidden trait as merely a statistical trick, providing an easy way to model extra heterogeneity without directly representing a real trait.
+%B: this transition from the meaning of hidden states to best practices of lumping is a bit hard.
+%Here Rosana adds the value of adding traits with the lumped models
+More generally, when analyzing the value of additional traits to a complex model, it may be best to opt for the use of lumped models (\cref{figure:lumped,table:lumped}), and verify that adding a trait is statistically preferred.
+In the absence of additional information, the hidden states can be view as a statistical trick, providing an easy way to model extra heterogeneity without directly representing a specific trait.
+In our analyses, the transition rates between and within hidden traits did not affect the direction of the net diversification difference. 
+We find that it is important to allow for the hidden state transition rates to be asymmetrical, because those models tend to provide a better fit than models that fix the hidden state transition rates to be equal (\cref{suppfigure:asymmetric,supptable:asymmetry}).
 
-Although we fit an extensive set of models in order to relax a variety of their assumptions, we did not explore the process of trait change in conjunction with speciation.
-That is, our models all assume anagenetic but not cladogenetic trait evolution.
-Phylogenetic methods do exist to separate anagenetic and cladogenetic changes \citep{mayrose_2011, goldberg_2012, magnuson-ford_2012}.
-These have been applied to show that losses of SI are often associated with cladogenesis in Solanaceae \citep{goldberg_2012}, and that polyploidization co-occurs with speciation events in other families \citep{zhan_2016, freyman_2017}.
-Exploring cladogenetic trait change was outside the scope of our modeling complexity.
+Although we fit an extensive set of models in order to relax a variety of assumptions, we did not explore the process of trait change in conjunction with speciation.
+That is, our models all assume anagenetic trait evolution and ignore cladogenetic shifts.
+Anagenetic and cladogenetic changes can be separated with phylogenetic models \citep{mayrose_2011, goldberg_2012, magnuson-ford_2012}. 
+These have been previously applied to estimate the relative contribution of anagenetic and cladogenetic shifts in breeding system \citep{goldberg_2012} and polyploidization \citep{zhan_2016, freyman_2017}.
+We did not explore cladogenetic trait change because of the vast increase in state space and model number, outside the scope of our already complex and extensive modeling framework. 
 Although \citet{goldberg_2012} found that allowing cladogenetic changes did not substantially affect inference of net diversification rates associated with breeding system, future work could test whether this process affects diversification rate estimates with the more complex state spaces of our other models.
@@
 
-With evolution predominantly in the direction from diploid to polyploid, and from SI to SC, surviving lineages will tend eventually to become SC polyploids.
-We found that in Solanaceae, the pathway to this state is more likely to consist of a single step ($ID \rightarrow CP$) than two steps ($ID \rightarrow CD \rightarrow CP$).
-We also showed that although this question focuses on the process of state transitions, its answer is affected by the process of lineage diversification.
-We used a simple mathematical approach to investigate the contributions of the two pathways, but future work could use stochastic character mapping to more directly estimate the numbers of each type of transition within the clade.
+With evolution predominantly in the direction from diploid to polyploid, and from SI to SC, surviving lineages will tend to become SC polyploids.
+We find that in Solanaceae, the pathway to this state is more likely to consist of a single step ($ID \rightarrow CP$) than two steps ($ID \rightarrow CD \rightarrow CP$; \cref{figure:pathways}).
+Although this question focuses on the process of state transitions, we also show that its answer is affected by the process of lineage diversification.
+We used a simple mathematical approach to investigate the contributions of the two pathways, but future work could instead rely on stochastic character mapping to more directly estimate the numbers of each type of transition.
 
-Macroevolutionary transition rates represent a combination of time spent waiting for individuals to arise with a new trait value, and for that new trait value to become widespread within the species.
-For our traits, this consists of mutations that break SI or generate polyploid individuals, and selective pressures that cause loss or fixation of these mutants.
-Estimates of mutation rates are highly uncertain, but the chance of breakdown of SI within diploids is perhaps $10^{-5}$ per pollen grain; this includes breakdown by tetraploidization \citep{lewis1979}, which is itself estimated to occur at about the same rate \citep{ramsey_1998}.
-% unit check?
+Macroevolutionary transition rates represent a combination of time spent waiting for individuals with a new character state, and for that new state to become widespread within the species.
+For our traits, this consists of mutations that break SI or generate polyploid individuals, and selective pressures that cause fixation (or loss) of these mutants.
+Estimates of mutation rates are highly uncertain, but the chance of breakdown of SI within diploids is perhaps $10^{-5}$ per pollen grain; this includes breakdown by autopolyploidization \citep{lewis1979}, which is by itself estimated to occur approximately within the same order of magnitude \citep{ramsey_1998}.
 % Seemingly, then, the simple genic mutation rate that leads to loss of SI is at best equal to tetraploidization mutation rate, and possibly far lower.
-In contrast, we estimate a macroevolutionary transition rate from $ID$ to $CD$ that is three-fold greater than the rate from $ID$ to $CP$, indicating that selection restricts the fixation of new polyploids more than of new SC mutants \citep{robertson_2011}.
-% todo: check if new rates are still 3x
+In contrast, we infer a macroevolutionary transition rate from $ID$ to $CD$ that is three-fold greater than the rate from $ID$ to $CP$, indicating that selection restricts the fixation of new polyploids more than of new SC mutants  (\cref{figure:pathways}, \citealt{robertson_2011}).
+% todo: check if new rates are still 3x %FIXME
 
@@
 %
-First, our support for the direct pathway is consistent with the idea that breakdown of SI by whole genome duplication---via diploid `heteroalleleic' pollen---may often trigger the evolution of gender dimorphism as a different mechanism of inbreeding avoidance \citep{miller_2000}.
+First, our support for the direct pathway is consistent with the idea that breakdown of SI by whole genome duplication---via diploid `heteroallelic' pollen---may often trigger the evolution of gender dimorphism as a different mechanism of inbreeding avoidance \citep{miller_2000}.
 A further test of this hypothesis would additionally examine the propensity of polyploids generated through either pathway to become dioecious \citep{robertson_2011}.
@@
 Second, we might wonder whether the propensity for a polyploid species to diversify depends on whether it arose via the one-step or two-step pathway.
-This could be tested with a different form of a hidden state model, in which the polyploid state is subdivided based on the pathway taken into that state.
-This would be similar to what we did for our pathways analysis, but it would also include the possibility of different diversification rates in those $CP$ substates.
+This could be tested with a different form of a hidden state model, in which the polyploid state is subdivided, based on the path taken into that state.
+This would be similar our analysis, but it would also include the possibility of different diversification rates in those $CP$ substates.
 %
@@
 An identical procedure could be used in other families with gametophytic SI.
-In clades with sporophytic self-incompatibility systems, however, SI is not disabled by whole genome duplication, so there is no one-step pathway. % ref?
-The correlation between breeding system and ploidy may therefore be different, and it is currently unknown whether one of the two-step pathways might dominate.
+In clades with sporophytic SI systems, however, SI is not disabled by whole genome duplication, so there is no one-step pathway \citep{miller_2000,mable_2004}.
+The correlation between breeding system and ploidy may therefore be different in sporophytic systems, and it is unclear whether one of the two-step pathways might predominate.
 
@@
 
-Polyploidization is known to be common in plants, but the commonness of the reverse process---diploidization---is under active investigation.
+Polyploidization is known to be common in plants, but the relative frequency of the reverse process---diploidization---remains unclear, and it is under active investigation \citep{dodsworth_2015,mandakova_2018}.
 %
-In our analyses, ignoring diploidization if it is common could cause us to underestimate the polyploidization rate, overestimate of the diploid diversification rate, or otherwise confound our conclusions.
-We therefore included a slate of models with a diploidization parameter, and they showed that our main conclusions are robust to this process.
-These models also suggested modest statistical support for diploidization occuring within Solanaceae \cref{supptableFIXME}, although our estimates of the diploidization rate were highly uncertain.
-Furthermore, additional lines of evidence for classifying species as diploid or polyploid (beyond the genus chromosome multiplicity that we primarily relied on) would be needed for more reliable conclusions from an analysis such as ours.
+Ignoring diploidization, if it is common, could cause us to both underestimate the polyploidization rate and overestimate of the diploid diversification rate, leading to incorrect conclusions.
+Therefore, we included a slate of models with a diploidization parameter, and show that our main conclusions are robust to this process (\cref{suppfigure:alldip}).
+These models also suggest modest statistical support for diploidization occurring within Solanaceae (\cref{supptable:testdiploidization}), although our estimates of the diploidization rate were highly uncertain.
+Furthermore, additional lines of evidence for classifying species as diploid or polyploid (beyond the genus-level chromosome multiplicity that we primarily relied on) are needed for more reliable conclusions.
 %In the cases where ploidy level was assigned by us based chromosome multiplicity at the genus level, estimating diploidization might be a potential issue due to ploidy misclassification.  
@@
 %
-Regardless of whether genome polypoidization, followed by widespread diploidization, is a dominant mechanism of genome evolution in Solanaceae, it is clear that more work is needed for a complete understanding of the joint evolution of ploidy and breeding systems.
+Regardless of whether genome polyploidization, followed by widespread diploidization, is a dominant mechanism of genome evolution in Solanaceae, it is clear that more work is needed for a complete understanding of the joint evolution of ploidy and breeding systems.
 

### 2019-07-17 manuscript/introduction.tex
Discussion edits down to Diploidization section (including related changes in other sections)

@@
 One possible explanation for this phenomenon is that various traits differentially affect rates of diversification. 
-%B: note: traits and character states are synonymous.
-I've edited the rest of the manuscript so that "trait" is retained where ambiguity does not matter, but where directly contrasted (vs.: as here), I changed the wording.
 Dramatic increases in phylogenetic and phenotypic data along with methodological advances have greatly accelerated the search for traits that influence diversification. 
-%B: was: "phenotypic data and methodological advances"
 Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a challenge \citep[\eg][]{maddison_2015, rabosky_2015, beaulieu_2016, rabosky_2017}. 
@@
 Increase in ploidy can alter many traits and affect a variety of evolutionary and ecological processes \citep{ramsey_2002, sessa_2019}.
-%B: The contrast below was between: "Polyploids have been documented..." vs.
-"Other studies find...".
-I modified it to: "Polyploids were generally found..." vs.
-"more recent studies ..."; If you feel that "Many studies" or "Recent studies" is better, please alter as appropriate.
-%B: Separately, "common" and "numerous" are synonymous as used here (I removed "common" because that term also has the "unrefined" meaning).
 Polyploids were found to have an overall lower net diversification rate than diploids across many vascular plant clades \citep{mayrose_2011, mayrose_2015}. 
-However, recent genomic studies have found numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications may have played an important macroevolutionary role driving innovation and diversification in plants.
+However, recent genomic studies have inferred numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications may have played an important macroevolutionary role driving innovation and diversification in plants.
 Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies pervasive diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
 Our analyses re-examine the association between ploidy and lineage diversification by extending the approach of \citet{mayrose_2011, mayrose_2015} to include transitions from polyploid to diploid states and potential unobserved factors affecting diversification patterns.
-%R- No we don't do the below thing and it was confusing that's what Jacob Landis did in 2018
-%We also allow for heterogeneous diversification rates within each ploidy level, which could examine whether, for example, diploids with ancestral polyploidy diversify more rapidly than diploids without it.
 
@@
 Additionally, in RNase-based gametophytic SI systems, polyploidization directly causes the loss of SI \citep{stout1942, lewis1947}.
-%B: rm "alone" in  "...polyploidization alone directly causes..." because it incorrectly implies that nothing other than polyploidization causes SC.
 %B: disclaimer: ceteris paribus statement omitted; polyploidization ordinarily causes SC because of the formation of heterodiploid pollen.
@@
 \citet{robertson_2011} compared the contributions of these two paths, finding that evolution from SI diploids to SC polyploids is more likely to proceed via the one-step pathway over short timescales, but via the two-step pathway over long timescales.
-They considered only transitions among the states, however, so we investigate whether these results hold true when allowing for differences in lineage diversification.
+They considered only transitions among the states, however, and we investigate whether these results hold true when allowing for differences in lineage diversification.
 
-%B: redundant with "very powerful" below rm: Examining these questions requires powerful phylogenetic comparative methods.
 In the present study, we employ an extended framework of state speciation and extinction models, which simultaneously model transitions between the discrete states of a trait and different rates of speciation and extinction associated with each of those states \citep[`SSE' models;][]{maddison_2007, fitzjohn_2012}.

### 2019-07-17 manuscript/discussion.tex
I think I'm done with everything (through bottom of Discussion). Let me know if I missed anything or left something unclear!

@@
 More generally, when analyzing the value of additional traits to a complex model, it may be best to opt for the use of lumped models (\cref{figure:lumped,table:lumped}), and verify that adding a trait is statistically preferred.
-In the absence of additional information, the hidden states can be view as a statistical trick, providing an easy way to model extra heterogeneity without directly representing a specific trait.
+In the absence of additional information, the hidden states can be viewed as a statistical trick, providing an easy way to model extra heterogeneity without directly representing a specific trait.
 In our analyses, the transition rates between and within hidden traits did not affect the direction of the net diversification difference. 
@@
 Therefore, we included a slate of models with a diploidization parameter, and show that our main conclusions are robust to this process (\cref{suppfigure:alldip}).
-These models also suggest modest statistical support for diploidization occurring within Solanaceae (\cref{supptable:testdiploidization}), although our estimates of the diploidization rate were highly uncertain.
+These models also suggest modest statistical support for diploidization (\cref{supptable:testdiploidization}), although our estimates of its rate were highly uncertain. 
 Furthermore, additional lines of evidence for classifying species as diploid or polyploid (beyond the genus-level chromosome multiplicity that we primarily relied on) are needed for more reliable conclusions.
@@
 Other lines of evidence about the prevalence of diploidization within Solanaceae or its ancestors are mixed or even conflicting.
-% recent work on WGDs conflict with other lines of evidence, especially those concerning the observed simple synteny of genomes in Solanaceae, and patterns of evolution at self-incompatibility loci
-%
 % Part I: evidence for WGD in/near family
@@
 The age of the peak of paralog Ks distances is approximately $71 \pm 19$ My \citep{tomato2012}. 
-If this is the case, then all of the genomes may have been subsequently diploidized, yielding the widespread and common chromosome numbers in this and related families, n=11--12, presently considered to be diploid \citep{robertson_2011}.
+If this is the case, then all of the genomes may have been subsequently diploidized, yielding the widespread and frequent chromosome numbers in this and closely related families, n=11--12, all presently considered to be diploid \citep{robertson_2011}.
 
@@
 First, the Ks-inferred duplication \citep{tomato2012} likely pre-dates the origin of the family (49 My, HPD 46--53 My; \citealt{sarkinen_2013}). 
-Thus, the species in our analyes could be ancestrally polyploid, but would have diploidized before the root of the family.
+Thus, the species in our analyses could be ancestrally polyploid, but would have diploidized before the root of the family.
 Second, studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
@@
 Specifically, chromosome numbers and genome comparisons within the family (esp.\ the x=12 clade containing \textit{Solanum} and \textit{Nicotiana}) reveal strong conservation.
-\citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ by six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca.\ 10 inversions and 11 translocations (likely underestimated).
+\citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ by as few as six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca.\ 10 inversions and 11 translocations.
 Recovery of such simple relationships would require outstanding convergent loss of duplicated segments.
@@
 % Part IIb: evidence from SI [might need more explaining?]
-Furthermore, whole genome duplication in a eudicot lineage ancestral to Solanaceae would clash with the evidence that the homologous mechanism of SI that has been present continually in many families \citep{igic_2006} breaks down nearly invariably in natural and induced tetraploids \citep{stone_2002, mcclure_2009}.
+Furthermore, whole genome duplication in a eudicot lineage ancestral to Solanaceae would clash with the evidence that the homologous mechanism of SI, which has been present continually in many families \citep{igic_2006}, breaks down nearly invariably in natural and induced tetraploids \citep{stone_2002, mcclure_2009}.
 Most problematically in this context, it is unclear how to explain the maintenance of trans-generic polymorphism at the orthologous S-loci in Solanaceae and other families if SI was previously broken down by polyploidization.

### 2019-07-17 manuscript/discussion.tex
Fixed all todos and revised discussion issues

@@
 
-Throughout our numerous model comparisons, we find that inclusion of hidden traits provides a considerably better fit (\cref{supptable:testaddhidden}). 
+Throughout our numerous model comparisons, we find that inclusion of hidden states provides a considerably better fit (\cref{supptable:testaddhidden}). 
 This is consistent with the expectation that many processes, beyond those associated with the focal traits, can affect inference of speciation and extinction.
@@
 %A natural follow-up question is, what is this hidden trait?
-It is, however, unclear exactly which processes are captured by the hidden traits.
-For example, our results show that, to a varying extent, breeding system functions as a hidden trait in the ploidy-centered analysis, and vice versa (\cref{suppfigure:lumpedDP,suppfigure:lumpedIC}).
+It is, however, unclear exactly which processes are captured by the hidden states.
+For example, our results show that, to a varying extent, breeding system functions as a hidden state in the ploidy-centered analysis, and vice versa (\cref{suppfigure:lumpedDP,suppfigure:lumpedIC}).
 But the strong statistical support for processes not well explained by ploidy and breeding system (\cref{supptable:testaddhidden}) tempts one to interpret the remaining variation as the effect of other measurable traits.
@@
 fixme
-Nevertheless, it is also possible that the addition of hidden traits instead explains variation stemming from any of a number of unrelated processes or methodological artifacts, such as variation in rates of molecular evolution or misspecified models of sequence evolution.
-%Rosana and Emma check me.
+Nevertheless, it is also possible that the addition of hidden states instead explains variation stemming from any of a number of unrelated processes or methodological artifacts, as previously discussed by \citet{beaulieu_2016}.
+In the absence of additional information, the hidden states can be viewed as a statistical trick, providing an easy way to model extra heterogeneity without directly representing a specific trait.
+
+%such as variation in rates of molecular evolution or misspecified models of sequence evolution. %Rosana and Emma check me.
+%R- let's not get into the weeds of what is a hidden state (I remove the use of trait  for state because a state is what the model receives) .
+Jeremy does a good job discussing this in his paper so no need.
+
 %     but don't be circular, test in different clade
-% alpha is always low: new supp fig showing how A & B are inferred on the tree?  at least for ID/CD/CP+A/B model
-%B: this transition from the meaning of hidden states to best practices of lumping is a bit hard.
+% alpha is always low: new supp fig showing how A & B are inferred on the tree?
+at least for ID/CD/CP+A/B model - it is low because not enough transitions and mostly a different process towards the tips-who knows what is happening there?
+%B: this transition from the meaning of hidden states to best practices of lumping is a bit hard.- Fixed below
 %Here Rosana adds the value of adding traits with the lumped models
-More generally, when analyzing the value of additional traits to a complex model, it may be best to opt for the use of lumped models (\cref{figure:lumped,table:lumped}), and verify that adding a trait is statistically preferred.
-In the absence of additional information, the hidden states can be viewed as a statistical trick, providing an easy way to model extra heterogeneity without directly representing a specific trait.
-In our analyses, the transition rates between and within hidden traits did not affect the direction of the net diversification difference. 
-We find that it is important to allow for the hidden state transition rates to be asymmetrical, because those models tend to provide a better fit than models that fix the hidden state transition rates to be equal (\cref{suppfigure:asymmetric,supptable:asymmetry}).
+As more trait information becomes available for macroevolutionary studies, it is important not only recognizing the role of hidden states as a part of a general modeling approach, but also questioning whether adding more traits to diversification studies is relevant.
+More generally, when analyzing the value of adding additional traits to an already complex model with or without hidden states, we find that using lumped models  is a key tool to verify that adding a trait improves our understanding of the diversification patterns (\cref{figure:lumped,table:lumped}).
+
+%R-removed this, it is really not important to discuss
+%In our analyses, we find that the transition rates between and within hidden traits did not affect the direction of the net diversification differences. 
+%We find that it is important to allow for the hidden state transition rates to be asymmetrical, because those models tend to provide a better fit than models that fix the hidden state transition rates to be equal (\cref{suppfigure:asymmetric,supptable:asymmetry}).
 
@@
 % Seemingly, then, the simple genic mutation rate that leads to loss of SI is at best equal to tetraploidization mutation rate, and possibly far lower.
-In contrast, we infer a macroevolutionary transition rate from $ID$ to $CD$ that is three-fold greater than the rate from $ID$ to $CP$, indicating that selection restricts the fixation of new polyploids more than of new SC mutants  (\cref{figure:pathways}, \citealt{robertson_2011}).
-% todo: check if new rates are still 3x %FIXME
+In contrast, we infer a macroevolutionary transition rate from $ID$ to $CP$ that is 2.
+5 times greater than the rate from $ID$ to $CD$, indicating that selection restricts the fixation of new polyploids more than of new SC mutants  (\cref{figure:pathways}, \citealt{robertson_2011}).
+% todo: check if new rates are still 3x 
+%R- fixed, it is two and a half more and it was written upside down ID to CP 0.194 ID to CD 0.073 which means more evolution to CP than to CD.
 
@@
 \Citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \emph{Solanum lycopersicum}, may have experienced one or more whole genome duplications.
-A subsequent analysis of synteny between grape and \emph{Solanum} genomes, as well as the distribution between inferred paralogs within genomes of \emph{Solanum} (tomato and potato) each suggested that this lineage experienced a likely round of ancient genome duplication or triplication \citep{tomato2012}.
-The age of the peak of paralog Ks distances is approximately $71 \pm 19$ My \citep{tomato2012}. 
-If this is the case, then all of the genomes may have been subsequently diploidized, yielding the widespread and frequent chromosome numbers in this and closely related families, n=11--12, all presently considered to be diploid \citep{robertson_2011}.
+A subsequent analysis of synteny between grape and \emph{Solanum} genomes, as well as the distribution between inferred paralogs within genomes of \emph{Solanum} (tomato and potato) each suggested that this lineage experienced a likely round of ancient genome duplication or triplication ( $71 \pm 19$ My, \citet{tomato2012})  likely pre-dating the origin of the family ($49 \pm 3$ My; \citealt{sarkinen_2013}).
+
+%The age of the peak of paralog Ks distances is approximately $71 \pm 19$ My \citep{tomato2012}. 
+% R- not entering into the weeds of Ks plots
+%f this is the case, then all of the genomes may have been subsequently diploidized, yielding the widespread and frequent chromosome numbers in this and closely related families, n=11--12, all presently considered to be diploid \citep{robertson_2011}.
 
 % Part II: evidence against WGD in/near family
-On the other hand, there is little evidence for the occurrence of diploidization within Solanaceae itself.
-First, the Ks-inferred duplication \citep{tomato2012} likely pre-dates the origin of the family (49 My, HPD 46--53 My; \citealt{sarkinen_2013}). 
-Thus, the species in our analyses could be ancestrally polyploid, but would have diploidized before the root of the family.
-Second, studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
+On the other hand, there is little evidence for the occurrence of diploidization after the origin of Solanaceae itself.
+% R- This is redundant now we said it above
+%First, the Ks-inferred duplication \citep{tomato2012} likely pre-dates the origin of the family (49 My, HPD 46--53 My; \citealt{sarkinen_2013}). 
+%Thus, the species in our analyses could be ancestrally polyploid, but would have diploidized before the root of the family.
+Studies comparing map-based genome synteny within the family find no evidence for recent diploidization (extensively discussed in \citet{wu_2010a}).
 Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species.
@@
 Specifically, chromosome numbers and genome comparisons within the family (esp.\ the x=12 clade containing \textit{Solanum} and \textit{Nicotiana}) reveal strong conservation.
-\citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ by as few as six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca.\ 10 inversions and 11 translocations.
-Recovery of such simple relationships would require outstanding convergent loss of duplicated segments.
+
+% R- removed the  inversion discussion I think the point was made above about the map base genome is sufficient. We don't need to describe more specifics
+
+%\citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ by as few as six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca.\ 10 inversions and 11 translocations.
+%Recovery of such simple relationships would require outstanding convergent loss of duplicated segments.
 %

### 2019-07-17 manuscript/introduction.tex
Fixed all todos and revised discussion issues

@@
 Consequently, examining the association of only one character with diversification patterns can be misleading. 
-We embrace this challenge by jointly investigating two characters thought to influence speciation and extinction rates---ploidy level and breeding system---while allowing for their interactions and other confounding factors to change diversification.
+Here, we embrace this challenge by jointly investigating two characters thought to influence speciation and extinction rates---ploidy level and breeding system---while allowing for their interactions and other confounding factors to change diversification.
+We also test whether adding one more trait and increasing model complexity is worthwhile.
 % E: We need a "Here, we..." here.  Otherwise, we've just told the reader "not one trait!" and then launch into one trait.
@@
 I didn't change anything but it's clear fodder for Discussion.
+% R. added Emma's suggetion
 
@@
 And we don't need methods in the intro.
+% R- The graphical model coding is not about the figures but about how you code the nodes in the code. But I agree that it might be confusing especiallly for most botanists.
 %Although potentially very powerful, this approach can be prone to reporting a false association between a trait and lineage diversification when the data contain additional, unmodeled diversification rate heterogeneity \citep{maddison_2015, rabosky_2015}.
@@
 Using the ploidy and breeding system model without hidden states, we quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
-We also lump together states within these joint models of ploidy and breeding system in order to test whether increasing complexity from one trait to two traits significantly improves our understanding of the diversification process.
-%B: I altered this line, but it is still a bit redundant given previous sentence.
+We also aggregate  states within these joint models of ploidy and breeding system in order to test whether increasing complexity from one trait to two traits significantly improves our understanding of the diversification process.
+%B: I altered this line, but it is still a bit redundant given previous sentence.R-fixed
 %Our analyses evaluate combinations of macroevolutionary models to parse out the associations of ploidy and breeding system on speciation and extinction rates in Solanaceae, while simultaneously accounting for other processes that influence diversification.

### 2019-07-17 manuscript/methods.tex
Fixed all todos and revised discussion issues

@@
 The Supplementary Information contains citations for the numerous original data sources. % FIXME (and say here how many refs?) %B: no need for an additional counting and updating headache?
+% R- There is few data that I added so I can find those references and send them to you, so it would be simply the same references that you previously had plus another 20-25 I think
 The Dryad archive contains the data and tree files used for analyses. % FIXME
 % \url{https://github.com/roszenil/solploidy}  E: we should use dryad instead, because github repos are not guaranteed to last
+%This fixme happens after publications
 
@@
 % E: Can we call it something other than q_IC?  It's a bit confusing to have the same parameter mean different things in different models.
+% R: In the lumped models I call them something different q_0 and rho_0 because they are different parameters but I don't think we should explain.
 
@@
 As before, matrix exponentiation provides the relative chance of changing from one state to any other state after time $t$.
-In this case, however, these are not probabilities because diversification changes the number of lineages as time passes.
-%B: FIXME, we'll hear complaints again about "this," "these," and ...
-"their." (below)
+The calculations of the propagation matrix are not probabilities because diversification changes the number of lineages as time passes. 
+%B: FIXME, we'll hear complaints again about "this," "these," and ... "their." (below)
+% R- fixed
 We can still use ratios, however, to consider the relative contribution of each pathway, analogous to the normalized age structure in a growing population.
@@
 % E & R todo ?: for the case with net div, write out the propagation matrix and equations for pathway contributions
+% R- couldn't solve to get the exact solutions turns out it is really hard.
 %
@@
 
-Parameters for each of the 29 models were estimated using RevBayes \citep{hoehna_2016}.
-Scripts for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.
+Parameters for each of the 29 models were coded as graphical models and Bayesian statistical inference was performed RevBayes \citep{hoehna_2016}.
+Scripts for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.- Happens after publication
 We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $651/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
@@
 We calculated the marginal log-likelihood for each of the models using fifty stepping stone steps under the methodology of \citet{xie_2010}, implemented in RevBayes \citep{hoehna_2016}.
-Each stepping stone step was found by calculating at least 500 generations of burn-in followed by a total of 1,000 MCMC steps. % (\cref{table:marginallike}) updated ref?
+Each stepping stone step was found by calculating at least 500 generations of burn-in followed by a total of 1,000 MCMC steps (\cref{table:bayesfactors}).
 
@@
 For example, there are 70 species that are SC with unknown ploidy.
-In the $D/P$ models these are coded as $D$ and $P$ (uncertain/consistent with either state), in the $I/C$ models they are coded as $C$, and in the $ID/CD/CP$ models they are coded as $CD$ and $CP$.
+In the ploidy only models (M1- M10) these are coded as $D$ and $P$ (uncertain/consistent with either state), in the breeding system only models (M11-M15) they are coded as $C$, and in the ploidy and breeding system models (M16-M29) they are coded as $CD$ and $CP$.
 In the hidden state models, all species were coded as consistent with both $A$ and $B$.
@@
 \includegraphics[width=\textwidth]{alldiversificationasymhidden.pdf} 
-\caption{Net diversification rates for all models without diploidization rate estimate. 
+\caption{Net diversification rates for SSE models of focal traits with or without hidden state. 
 Each panel contains a graphical summary of estimated model parameters and displays posterior distributions for net diversification estimates.}
@@
 \end{tabular}
-\caption{Bayes factors of ploidy and breeding system withou diploidization.
+\caption{Bayes factors of ploidy and breeding system without diploidization.
 Results indicate that the MuHiSSE model with asymmetric hidden rates (17) is  strongly preferred over M16 and M18 and moderately preferred over the MuHiSSe with all rates asymetric (M20).
@@
 M19.
-IC/CD/CP asym                       & -1393.20                                                          & M19 vs.
+IC/CD/CP +A/B asym                       & -1393.20                                                          & M19 vs.
 M24 & 6.46            & M24 (Moderate)                                                       \\

### 2019-07-17 manuscript/response.tex
Fixed all response to reviews

@@
 
-[Considered.] % TODO
+Introduction was edited.
 
@@
 
-[Considered.] % TODO
+It is appropriate based on suggestions for re-analyses.
 
@@
 asymmetric rates is that the direction of the results does not change but the amount of difference amongst net diversification posterior distributions does.
-We found using Bayes factors that a model including asymmetric hidden rates is preferred over a model with equal hidden rates. 
-Also,  the asymmetry in the trait rates have moderate to not support of being better than a model with equal trait transition rates.  %TODO put number.
+We found using Bayes factors that a model including asymmetric hidden rates is preferred over a model with equal hidden rates (Table S6, Figure S5). 
+Also,  the asymmetry in the trait rates have moderate to not support of being better than a model with equal trait transition rates.
 
@@
 
-[Considered.] % TODO %We can defend this on account of repeating some of the points because of numerous and complex  questions; aimed to prevent flipping pages back
+The discussion was heavily edited.
+However, we consider that it is important to refresh some of the introductory key points due to the numerous and complex questions that are being addressed with SSE models.
+ % TODO %We can defend this on account of repeating some of the points because of numerous and complex  questions; aimed to prevent flipping pages back
+% Fixed
 
@@
 %B: Rosana, you know this best, but it seems to me that the current phrasing (below) is unclear.
-The soaking noise comes from the hidden state so in the presence of hidden states finding that diploidization matters is key. 
-All the Bayes factors indicated strong support for models that include diploidization. 
-However, the diploidization parameter is difficult to estimate (wide credible intervals) mainly because diploidization is a genome size process that can be partially or minimally recover from ploidy data.
-This is the reason why we do not overstate this finding, the importance of diploidization needs to be studied using genomes.
+%R- rewritten
+The soaking noise comes from adding  hidden states to models.
+Therefore in the presence of hidden states finding that diploidization  parameter is different from zero truly represents potential evidence of the reversibility of polyploidy.
+Furthermore, all the Bayes factors indicated strong preference for models that include diploidization. 
+However, diploidization is a genomic process that can be partially or minimally recover from ploidy data. 
+And as discussed at the end of the article, genomic evidence for diploidization in Solanaceae  is still debatable.
 
@@
 
-[Seems reasonable.] % TODO
+Considered.
 
@@
 
-[Considered.] % TODO
+Clarified.
 
@@
 
-[Considered.] % TODO
+Fixed.
 
@@
 
-[Considered.] % TODO
+We edited the discussion.
+We found not evidence of conclusions about diversification changing in the presence of diploidization.
+The  the estimates for rate of polyploidization in the absence of diploidization also remained the same (\cref{figure:hissedip}.
+However, ignoring diploidization, if it is common, could cause to underestimate the polyploidization rate or create larger credible intervals.
 
-\begin{quote}
+We have corrected the phrase accordingly.
+Thank you for the suggestion
+\begin{figure}
+    \centering 
+      \begin{minipage}[b]{0.5\textwidth}
+    \includegraphics[width=\textwidth]{hisseDPnodipasymposteriordist.pdf} 
+    \end{minipage}\\
+    \vspace{1cm}
+    \begin{minipage}[b]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{hisseDPasymposteriordist.pdf}
+        \end{minipage}
+    \caption{ Top 6 panels: Ploidy model with hidden states without diploidization.
+Bottom 6 panels: Ploidy model with hidden states with diploidization.
+Rate of polyploidy has a wider credible interval in the absence of diploidization}
+    \label{figure:hissedip}
+\end{figure}
+
+\begin{quote}1
 --430  remove the comma
@@
 \end{quote}
-
-[Considered.] % TODO
+Fixed
 
@@
 We found that the former are important and strengthened the differences.
-We found no evidence that having all asymmetric rates is preferable over just asymmetric hidden for our specific example.
-%TODO (Figs and table)
+We found no evidence that having all asymmetric rates is preferable over just asymmetric hidden for our specific example (Figure S5, Table S6).
 
@@
 
-Thank you for this suggestion, which we have incorporated in the Discussion. % TODO
+Thank you for this suggestion.
+Fitting the new models the only trend we can notice is that for ploidy only models hidden states seem to transition mostly towards recent lineages but still spread throughout lineages (Figure S8).
+For breeding system models there are a two or three clades where transitions between hidden states happen sooner (Figure S12 and S16).
+However, we don't have a clear interpretation for this phenomenon.
 
@@
 
-[Considered.] % TODO
+Correct, and inference remains the same.
+\cref{figure:hissedip}.
+However, ignoring diploidization, if it is common, could cause to underestimate the polyploidization rate or create larger credible intervals.
+We have corrected the phrase accordingly.
 
@@
 We appreciated this suggestion because it allowed us to parse out the effect of the traits in the diversification process. 
-By comparing Bayes factors, models with traits are strongly preferred over CID when the breeding system is the focal trait, and there is little to no evidence that models that include ploidy are better than CID.
-These results strongly suggest that breeding system is much more informative than ploidy in the diversification dynamics of Solanaceae.
-%TODO: write table number
+By comparing Bayes factors, models with traits are strongly preferred over CID when the breeding system is the focal trait, and there is little to no evidence that models that include ploidy are better than CID (Tables 1-2 and S1-S3).
+These results strongly suggest that breeding system is much more informative than ploidy in the diversification dynamics of Solanaceae (Tables 1-2).
+%TODO: write table number- Fixed
 
@@
 During the revision we found a work-around to be able to implement all the extra models we are presenting.
-New models with asymmetric traits are preferred and strengthen the results.
+New models with asymmetric traits are preferred and strengthen the results (Figure S5, Table S6).
 
@@
 
-[Considered.] % TODO
+Discussion edited.
 
@@
 
-We have cited them accordingly. % TODO
+We have cited them accordingly.
 
@@
 
-Rephrased. % TODO
+Rephrased.
 

### 2019-07-17 manuscript/discussion.tex
Fixed all response to reviews

@@
 %
-Ignoring diploidization, if it is common, could cause us to both underestimate the polyploidization rate and overestimate of the diploid diversification rate, leading to incorrect conclusions.
+Ignoring diploidization, if it is common, could cause underestimation and increase of uncertainty in  the polyploidization rate. 
 Therefore, we included a slate of models with a diploidization parameter, and show that our main conclusions are robust to this process (\cref{suppfigure:alldip}).

### 2019-07-17 manuscript/abstract.tex
Small edits to space etc. in abstract and intro. Returned one sentence in each of two discussion paragraphs (the point of Ks timing is that the event is not within the family, which is critical for the section and inference evaluation.)

@@
 These results are largely robust to allowing for diploidization.
-Finally, we find that the evolutionary pathway to polyploidy in Solanaceae is most often via the direct breakdown of self-incompatibility by whole genome duplication, rather than indirectly via breakdown followed by polyploidization.
+Finally, we find that the most common evolutionary pathway to polyploidy in Solanaceae occurs via direct breakdown of self-incompatibility by whole genome duplication, rather than indirectly via breakdown followed by polyploidization.

### 2019-07-17 manuscript/discussion.tex
Small edits to space etc. in abstract and intro. Returned one sentence in each of two discussion paragraphs (the point of Ks timing is that the event is not within the family, which is critical for the section and inference evaluation.)

@@
 % E: should this results summary be much shorter?
-% B: I think it's fine, but I switched it to present tense (not strictly, but almost) for both reasons of style (aesthetic preference, perhaps), to separate previous work from current work, and to avoid awkward implications about the process not being ongoing.
-Moreover, the use of "show" or "demonstrate" is very loose in the context of our weak inference resulting from--what we all hopefully agree--are potentially flawed models.
 Previous analyses of the effects of ploidy on diversification found that diploids are associated with greater net diversification rates than polyploids across many angiosperm clades \citep{mayrose_2011, mayrose_2015}.
@@
 Furthermore, additional lines of evidence for classifying species as diploid or polyploid (beyond the genus-level chromosome multiplicity that we primarily relied on) are needed for more reliable conclusions.
-%In the cases where ploidy level was assigned by us based chromosome multiplicity at the genus level, estimating diploidization might be a potential issue due to ploidy misclassification.  
-%R: By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with comparative phylogenetic models.
-%B: yielding to %R here.
-I still think there is overwhelming support for this statement, and few of the cited papers show convincingly that, for example, diploidization happens within five orders of magnitude compared with polyploidization, rendering "relatively rare" obsolete or invalid.
-(P.S.
-Rare not the same as unimportant!)
-%B: was: Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.
-
-% E: Boris, could you please revisit the following two paragraphs?  TODO
-% Two questions:
-%    Did my reorganization of this section mess anything up?
-%    Can this text be shorter?  It's seems disproportionately long, and I'm hesitant to revise it heavily myself because the logical connections aren't entirely clear to me.
 
@@
 A subsequent analysis of synteny between grape and \emph{Solanum} genomes, as well as the distribution between inferred paralogs within genomes of \emph{Solanum} (tomato and potato) each suggested that this lineage experienced a likely round of ancient genome duplication or triplication ( $71 \pm 19$ My, \citet{tomato2012})  likely pre-dating the origin of the family ($49 \pm 3$ My; \citealt{sarkinen_2013}).
-
-%The age of the peak of paralog Ks distances is approximately $71 \pm 19$ My \citep{tomato2012}. 
-% R- not entering into the weeds of Ks plots
-%f this is the case, then all of the genomes may have been subsequently diploidized, yielding the widespread and frequent chromosome numbers in this and closely related families, n=11--12, all presently considered to be diploid \citep{robertson_2011}.
+The age of the peak of paralog Ks distances is approximately $71 \pm 19$ My \citep{tomato2012}. 
+%B: We cannot selectively delete the weak evidentiary basis for ku2000, blanc2004, and tomato2012.
+It's the only mention of their inference date and method, and it is a single sentence.
+%R- not entering into the weeds of Ks plots
 
@@
 On the other hand, there is little evidence for the occurrence of diploidization after the origin of Solanaceae itself.
-% R- This is redundant now we said it above
-%First, the Ks-inferred duplication \citep{tomato2012} likely pre-dates the origin of the family (49 My, HPD 46--53 My; \citealt{sarkinen_2013}). 
-%Thus, the species in our analyses could be ancestrally polyploid, but would have diploidized before the root of the family.
-Studies comparing map-based genome synteny within the family find no evidence for recent diploidization (extensively discussed in \citet{wu_2010a}).
-Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution between a number of species.
-% including all of those in the relatively cytogenetically conserved x=12 group, which includes tomato, potato, eggplant, pepper, and tobacco.
-Specifically, chromosome numbers and genome comparisons within the family (esp.\ the x=12 clade containing \textit{Solanum} and \textit{Nicotiana}) reveal strong conservation.
-
-% R- removed the  inversion discussion I think the point was made above about the map base genome is sufficient. We don't need to describe more specifics
-
-%\citet{wu_2010a} review the evidence from map-based genome comparisons and find that tomato and potato differ by as few as six inversions, tomato and eggplant by 24 inversions and five translocations, tomato and pepper by 19 inversions and six translocations, and tomato and tobacco by ca.\ 10 inversions and 11 translocations.
-%Recovery of such simple relationships would require outstanding convergent loss of duplicated segments.
-%
-% Part IIb: evidence from SI [might need more explaining?]
+First, the Ks-inferred duplication \citep{tomato2012} pre-dates the origin of the family (49 My, HPD 46--53 My; \citealt{sarkinen_2013}). 
+Studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
+Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution within species in the cytogenetically conserved `x=12' group, which includes tomato, potato, eggplant, pepper, and tobacco.
+Recovery of comparatively few rearrangements would require outstanding convergent loss of duplicated segments.
 Furthermore, whole genome duplication in a eudicot lineage ancestral to Solanaceae would clash with the evidence that the homologous mechanism of SI, which has been present continually in many families \citep{igic_2006}, breaks down nearly invariably in natural and induced tetraploids \citep{stone_2002, mcclure_2009}.
@@
 
-% Flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}. 
-% Following polyploidization, it is possible for genome reorganization, downsizing, and loss to occur \citep{dodsworth_2015, zenil_2016, mandakova_2018}.
-% As a consequence, nearly all extant species classified as ``diploid" in our analyses are possibly secondary diploids, having undergone both polyploidization and subsequent diploidization.
-
-% no discussion of magnitude of \delta vs \rho!
-
-% problems: this system has a big set of advantages - carefully curated (co-)occurrences of two traits, genetic basis, etc. and everything is still contradictory and opaque.
-
 \section{Conclusion}
-
-% E: The text previously here was great, but it read to me more like an abstract or methods/results.
-I tried to make it more like a big-picture summary.
-What other general points do we want to hit here?
 

### 2019-07-17 manuscript/introduction.tex
Small edits to space etc. in abstract and intro. Returned one sentence in each of two discussion paragraphs (the point of Ks timing is that the event is not within the family, which is critical for the section and inference evaluation.)

@@
 One possible explanation for this phenomenon is that various traits differentially affect rates of diversification. 
-Dramatic increases in phylogenetic and phenotypic data along with methodological advances have greatly accelerated the search for traits that influence diversification. 
+Dramatic increases in phylogenetic and phenotypic data, along with methodological advances, have greatly accelerated the search for traits that influence diversification.
 Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a challenge \citep[\eg][]{maddison_2015, rabosky_2015, beaulieu_2016, rabosky_2017}. 
@@
 Consequently, examining the association of only one character with diversification patterns can be misleading. 
-Here, we embrace this challenge by jointly investigating two characters thought to influence speciation and extinction rates---ploidy level and breeding system---while allowing for their interactions and other confounding factors to change diversification.
+Here, we embrace this challenge by jointly investigating two characters thought to influence speciation and extinction rates---ploidy level and breeding system---while allowing for their interactions, and other confounding factors, to change diversification.
 We also test whether adding one more trait and increasing model complexity is worthwhile.
-% E: We need a "Here, we..." here.  Otherwise, we've just told the reader "not one trait!" and then launch into one trait.
-% Could also state that we newly test whether adding one more trait is worthwhile.
-%B: I think you are correct, but it introduces a new concern: we now say N=1 is too few, so not why try N>2 characters?
-Why is N=2 the Goldilocks number?
-I didn't change anything but it's clear fodder for Discussion.
-% R. added Emma's suggetion
 
@@
 Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are also remarkably common and crucially affect the distribution and amount of genetic variation in populations \citep{barrett2013}.
-%B: note: we use "remarkably" to describe the frequency of both characters; hence "also."
 In particular, a variety of genetic self-incompatibility (SI) systems cause plants to reject their own pollen, and loss of such mechanisms, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974, igic_2008}.
@@
 In the present study, we employ an extended framework of state speciation and extinction models, which simultaneously model transitions between the discrete states of a trait and different rates of speciation and extinction associated with each of those states \citep[`SSE' models;][]{maddison_2007, fitzjohn_2012}.
-% All the models proposed for this study are coded as graphical models and Bayesian inference was performed over the proposed models using RevBayes software \citep{hoehna_2016}.
-% E: The models aren't drawn graphically in any of the figures (not in the revbayes sense, at least).
-And we don't need methods in the intro.
-% R- The graphical model coding is not about the figures but about how you code the nodes in the code. But I agree that it might be confusing especiallly for most botanists.
-%Although potentially very powerful, this approach can be prone to reporting a false association between a trait and lineage diversification when the data contain additional, unmodeled diversification rate heterogeneity \citep{maddison_2015, rabosky_2015}.
 We start by fitting binary state speciation and extinction models to each ploidy and breeding system independently \citep{maddison_2007}.
@@
 We also aggregate states within these joint models of ploidy and breeding system in order to test whether increasing complexity from one trait to two traits significantly improves our understanding of the diversification process.
-%B: I altered this line, but it is still a bit redundant given previous sentence.R-fixed
-%Our analyses evaluate combinations of macroevolutionary models to parse out the associations of ploidy and breeding system on speciation and extinction rates in Solanaceae, while simultaneously accounting for other processes that influence diversification.
-%First, we present models in which the effects of ploidy and breeding system are considered separately and compare the inferences to previously published results. 
-%Second, we add a hidden character to each model and investigate if state-dependent diversification in focal character retains its explanatory power, or whether the added hidden character obscures it.
 Furthermore, we extend all the models involving ploidy to investigate the potential effects of including diploidization.

### 2019-07-18 manuscript/discussion.tex
I edited (slightly) the Diploidization section. It's a bit shorter now, and very similar to Rosana's last version (in content).

@@
 On the one hand, polyploidy may have occurred prior to the origin of Solanaceae, rendering all extant `diploids' secondarily derived.
-\Citet{ku2000} and \citet{blanc2004} posited that the lineage leading to tomato, \emph{Solanum lycopersicum}, may have experienced one or more whole genome duplications.
-A subsequent analysis of synteny between grape and \emph{Solanum} genomes, as well as the distribution between inferred paralogs within genomes of \emph{Solanum} (tomato and potato) each suggested that this lineage experienced a likely round of ancient genome duplication or triplication ( $71 \pm 19$ My, \citet{tomato2012})  likely pre-dating the origin of the family ($49 \pm 3$ My; \citealt{sarkinen_2013}).
-The age of the peak of paralog Ks distances is approximately $71 \pm 19$ My \citep{tomato2012}. 
-%B: We cannot selectively delete the weak evidentiary basis for ku2000, blanc2004, and tomato2012.
-It's the only mention of their inference date and method, and it is a single sentence.
-%R- not entering into the weeds of Ks plots
+\Citet{ku2000} and \citet{blanc2004} posited that the lineage leading to cultivated tomato, \emph{Solanum lycopersicum}, may have experienced one or more whole genome duplications.
+A subsequent analysis of synteny between grape and \emph{Solanum} genomes, as well as genetic distances (Ks) between inferred paralogs within genomes of \emph{Solanum} (tomato and potato), each suggested that this lineage experienced a likely round of ancient genome duplication or triplication ( $71 \pm 19$ My, \citet{tomato2012})  likely pre-dating the origin of the family ($49 \pm 3$ My; \citealt{sarkinen_2013}).
 
@@
 On the other hand, there is little evidence for the occurrence of diploidization after the origin of Solanaceae itself.
-First, the Ks-inferred duplication \citep{tomato2012} pre-dates the origin of the family (49 My, HPD 46--53 My; \citealt{sarkinen_2013}). 
-Studies comparing map-based genome synteny within the family find no evidence for recent diploidization \citep{wu_2010a}.
+Studies comparing genetic map-based genome synteny within a number of species in this family find no evidence for diploidization \citep{wu_2010a}.
 Instead, simple genome re-arrangements appear sufficient to explain chromosomal evolution within species in the cytogenetically conserved `x=12' group, which includes tomato, potato, eggplant, pepper, and tobacco.

### 2019-07-18 manuscript/discussion.tex
- fixing bad parentheses/spacing

@@
 \Citet{ku2000} and \citet{blanc2004} posited that the lineage leading to cultivated tomato, \emph{Solanum lycopersicum}, may have experienced one or more whole genome duplications.
-A subsequent analysis of synteny between grape and \emph{Solanum} genomes, as well as genetic distances (Ks) between inferred paralogs within genomes of \emph{Solanum} (tomato and potato), each suggested that this lineage experienced a likely round of ancient genome duplication or triplication ( $71 \pm 19$ My, \citet{tomato2012})  likely pre-dating the origin of the family ($49 \pm 3$ My; \citealt{sarkinen_2013}).
+A subsequent analysis of synteny between grape and \emph{Solanum} genomes, as well as genetic distances (Ks) between inferred paralogs within genomes of \emph{Solanum} (tomato and potato), each suggested that this lineage experienced a likely round of ancient genome duplication or triplication ($71 \pm 19$ My; \citealt{tomato2012})  likely pre-dating the origin of the family ($49 \pm 3$ My; \citealt{sarkinen_2013}).
 

### 2019-07-18 manuscript/introduction.tex
Oops! Fixing the opening quote and adding its reference! - adding final sentence emphsizing the same statement from honoree Barrett

@@
 \begin{quote}
-\em{``Among life history traits, reproductive characters that determine mating system patterns are perhaps the most influential in governing macroevolution." 
+\em{``Among life history traits, reproductive characters that determine mating patterns are perhaps the most influential in governing macroevolution." 
 }

### 2019-07-18 manuscript/response.tex
- polishing some minor stuff in the response - diploidization support: strong -> moderate

@@
 \medskip
-
 Thank you very much for your close reading of our manuscript and supportive comments.
 We appreciate your constructive suggestions, each of which we address below.
-You will find a great difference in the number of models that we have fitted for the revision based on reviewer suggestions (originally 12 now 29).
+You will find a great difference in the number of models that we have fitted for the revision based on reviewer suggestions (originally 12, now 29, with a special supplementary figure to enable easy visualization).
 There are two main reasons for this sizable increase.
@@
 We hope the changes and the new models are satisfactory to the editorial staff and the reviewers.
+
 \medskip
@@
 
-Rosana Zenil-Ferguson, on behalf of all authors
+The Authors \\
+(Rosana Zenil-Ferguson et al.)
 
@@
 %R- rewritten
-The soaking noise comes from adding  hidden states to models.
-Therefore in the presence of hidden states finding that diploidization  parameter is different from zero truly represents potential evidence of the reversibility of polyploidy.
-Furthermore, all the Bayes factors indicated strong preference for models that include diploidization. 
-However, diploidization is a genomic process that can be partially or minimally recover from ploidy data. 
-And as discussed at the end of the article, genomic evidence for diploidization in Solanaceae  is still debatable.
+Adding hidden states to models more likely `soaks up' noise.
+Therefore, in the presence of hidden states, finding that the diploidization parameter is different from zero truly represents potential evidence of the reversibility of polyploidy. 
+Furthermore, all the Bayes factors indicated a moderate preference for models that include diploidization.
+%CHECK: I changed this because the rest of the manuscript states "moderate" and our tables show that the hidden state model shows "moderate" preference.
+However, diploidization is a genomic process that may not be fully recovered from ploidy data. 
+As discussed at the end of the article, genomic evidence in favor or against diploidization in Solanaceae is far from final.
 
@@
 
-Considered.
+We include a necessarily limited discussion of other systems.
 
@@
 We edited the discussion. 
-We found not evidence of conclusions about diversification changing in the presence of diploidization.
-The  the estimates for rate of polyploidization in the absence of diploidization also remained the same (\cref{figure:hissedip}.
-However, ignoring diploidization, if it is common, could cause to underestimate the polyploidization rate or create larger credible intervals.
+We found no conclusive evidence about diversification changing in the presence of diploidization. 
+The estimates for rate of polyploidization in the absence of diploidization also remained the similar (\cref{figure:hissedip}. 
+However, ignoring diploidization, if it is common, could cause an underestimate of the polyploidization rate or create larger credible intervals.
 
 We have corrected the phrase accordingly.
-Thank you for the suggestion
+Thank you for the suggestion.
 \begin{figure}
@@
         \end{minipage}
-    \caption{ Top 6 panels: Ploidy model with hidden states without diploidization.
-Bottom 6 panels: Ploidy model with hidden states with diploidization.
-Rate of polyploidy has a wider credible interval in the absence of diploidization}
+    \caption{Top six panels: ploidy model with hidden states without diploidization.
+Bottom six panels: ploidy model with hidden states with diploidization.
+Rate of polyploidy has a wider credible interval in the absence of diploidization.}
     \label{figure:hissedip}
@@
 2004.
-Systematic Biology 53:265–277).
+Systematic Biology 53:265--277).
 I think it would have been important to at least present the results of an unconstrained model to show that this decision to constrain the rates do not affect the conclusions.
@@
 Thank you for this suggestion. 
-Fitting the new models the only trend we can notice is that for ploidy only models hidden states seem to transition mostly towards recent lineages but still spread throughout lineages (Figure S8).
-For breeding system models there are a two or three clades where transitions between hidden states happen sooner (Figure S12 and S16).
-However, we don't have a clear interpretation for this phenomenon.
+After fitting the new models, the only trend we can notice is that for ploidy-only models, hidden states seem to transition mostly towards recent lineages but are still spread across the tree (Figure S8).
+For breeding system-only models, there are a two or three clades in which transitions between hidden states happen sooner (Figure S12 and S16). 
+However, we do not have a clear interpretation for these phenomena.
 
@@
 
-Done. % Done: Emma did the new figure
+Done.
 
@@
 
-Correct, and inference remains the same.
-\cref{figure:hissedip}.
+Correct, and inference remains the same. \cref{figure:hissedip}. 
 However, ignoring diploidization, if it is common, could cause to underestimate the polyploidization rate or create larger credible intervals.

### 2019-07-18 manuscript/discussion.tex
Oops! Fixing the opening quote and adding its reference! - adding final sentence emphsizing the same statement from honoree Barrett
Itay's comments related to discussion.tex & main.tex (#28-#32)

@@
 Our analyses that include a hidden trait recover the same pattern, with one important difference.
-On the background of one hidden state, we recover a net diversification rate for SC species that is positive and greater than the diversification rate of SI species on the background of the other hidden state (\cref{figure:netdivall}D).%B: I rewrote this to separate the point of the side clause.
-Therefore, SC may not be a `dead end' (cf.\ \citealt{igic_2013}) when coupled with some unknown trait combinations (or processes not modeled by these two traits). 
-%Separately, I should note here that nothing about this analysis seems to me to strongly imply that SC is not a dead end in combination with a trait (as we discuss immediately below!).
+On the background of one hidden state, we recover a net diversification rate for SC species that is positive and greater than the diversification rate of SI species on the background of the other hidden state (\cref{figure:netdivall}D).
+Therefore, SC may not be a `dead end' when coupled with some unknown trait combinations or processes not modeled by these two traits. 
+%B: nothing about this analysis seems to me to strongly imply that SC is not a dead end in combination with a trait (as we discuss immediately below!).
 Much of the explanatory power derived for the A/B model may come from (for instance) accommodating tree scaling problems or rate heterogeneity in molecular evolution.
@@
 This is consistent with the expectation that many processes, beyond those associated with the focal traits, can affect inference of speciation and extinction. 
-%removed "one or two" because it's true for N traits.
-%A natural follow-up question is, what is this hidden trait?
 It is, however, unclear exactly which processes are captured by the hidden states. 
@@
 For example, our data appear to show a rapidly-diversifying Australasian clade of mostly SC species within \emph{Solanum}, which suggests that geography may play a role. 
-% Boris, please check!
-Australian?
-fixme
 Nevertheless, it is also possible that the addition of hidden states instead explains variation stemming from any of a number of unrelated processes or methodological artifacts, as previously discussed by \citet{beaulieu_2016}.
@@
 
-%such as variation in rates of molecular evolution or misspecified models of sequence evolution. %Rosana and Emma check me.
-%R- let's not get into the weeds of what is a hidden state (I remove the use of trait  for state because a state is what the model receives) .
+%R- let's not get into the weeds of what is a hidden state (I remove the use of trait  for state because a state is what the model receives).
 Jeremy does a good job discussing this in his paper so no need.
-
 %     but don't be circular, test in different clade
@@
 at least for ID/CD/CP+A/B model - it is low because not enough transitions and mostly a different process towards the tips-who knows what is happening there?
-%B: this transition from the meaning of hidden states to best practices of lumping is a bit hard.- Fixed below
 %Here Rosana adds the value of adding traits with the lumped models
@@
 More generally, when analyzing the value of adding additional traits to an already complex model with or without hidden states, we find that using lumped models  is a key tool to verify that adding a trait improves our understanding of the diversification patterns (\cref{figure:lumped,table:lumped}).
-
-%R-removed this, it is really not important to discuss
-%In our analyses, we find that the transition rates between and within hidden traits did not affect the direction of the net diversification differences. 
+%R-removed this: %In our analyses, we find that the transition rates between and within hidden traits did not affect the direction of the net diversification differences. 
 %We find that it is important to allow for the hidden state transition rates to be asymmetrical, because those models tend to provide a better fit than models that fix the hidden state transition rates to be equal (\cref{suppfigure:asymmetric,supptable:asymmetry}).
@@
 These have been previously applied to estimate the relative contribution of anagenetic and cladogenetic shifts in breeding system \citep{goldberg_2012} and polyploidization \citep{zhan_2016, freyman_2017}.
-We did not explore cladogenetic trait change because of the vast increase in state space and model number, outside the scope of our already complex and extensive modeling framework. 
-Although \citet{goldberg_2012} found that allowing cladogenetic changes did not substantially affect inference of net diversification rates associated with breeding system, future work could test whether this process affects diversification rate estimates with the more complex state spaces of our other models.
+We did not explore cladogenetic trait change because of the vast increase in parameter space and model number, outside the scope of our already complex and extensive modeling framework. %IM: #29
+Although \citet{goldberg_2012} found that allowing cladogenetic changes did not substantially affect inference of net diversification rates associated with breeding system, future work could test whether this process affects diversification rate estimates with the more complex state and parameter spaces of our other models.
 
@@
 A further test of this hypothesis would additionally examine the propensity of polyploids generated through either pathway to become dioecious \citep{robertson_2011}.
-%
 Second, we might wonder whether the propensity for a polyploid species to diversify depends on whether it arose via the one-step or two-step pathway.
-This could be tested with a different form of a hidden state model, in which the polyploid state is subdivided, based on the path taken into that state.
-This would be similar our analysis, but it would also include the possibility of different diversification rates in those $CP$ substates.
-%
+This could be tested with a different form of a hidden state model, in which the polyploid state is subdivided into parts, reflecting the arriving path taken to that state. %IM: #30
+Such an estimate of path-dependence could also include the possibility of different diversification rates for the subdivided $CP$ substates.
 Third, the generality of our findings in other families remains to be assessed.
@@
 The correlation between breeding system and ploidy may therefore be different in sporophytic systems, and it is unclear whether one of the two-step pathways might predominate.
-
 % pathways: compare/contrast with robertson2011
@@
 %     likely way to become P
-%     (compare those relative rates with robertson2011)
 
@@
 
-Polyploidization is known to be common in plants, but the relative frequency of the reverse process---diploidization---remains unclear, and it is under active investigation \citep{dodsworth_2015,mandakova_2018}.
-%
+Polyploidization is known to be common in plants, but the pace and relative frequency of the reverse process---diploidization---remains unclear, and it is under active investigation \citep{dodsworth_2015,mandakova_2018}.
+%IM: #31
 Ignoring diploidization, if it is common, could cause underestimation and increase of uncertainty in  the polyploidization rate. 
@@
 Additionally, our analysis of evolutionary pathways to polyploidy shows the importance of including diversification effects even when addressing questions that focus on trait evolution.
+Finally, although a growing number of plant traits have been studied to date, breeding systems indeed seem to be among the most influential in governing macroevolution \citep{barrett1996}.

### 2019-07-18 manuscript/main.tex
Itay's comments related to discussion.tex & main.tex (#28-#32)

@@
 The computing resources were provided by the Minnesota Supercomputing Institute (MSI) at the University of Minnesota.
-This work was supported by the National Science Foundation grant DEB-1655478 (to EEG), DEB-1655692 and NESCent sabbatical scholar award EF-0905606 (to BI), and the Binational Science Foundation BSF-2016550 (to IM).
+This work was supported by the National Science Foundation grant DEB-1655478 (to EEG), DEB-1655692 and NESCent sabbatical scholar award EF-0905606 (to BI), and the Israel Science Foundation 961/17 (to IM).
+%IM: #32
 %B: new:

### 2019-07-18 manuscript/methods.tex
- Itay's comments #21-33, only. Not done: #1-20, #34-35.

@@
 \caption{
-Character states used for each of the models.
-Each species retained on the tree belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known.
-Character state abbreviations: $I$ for self-incompatible, $C$ for self-compatible, $D$ for diploid, $P$ for polyploid, $?$ for unknown.
-(A) The five categories and the number of species in our data for each category.
+Character states used in the models.
+(A) Each species retained on the phylogeny belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known. 
+Number of species in each category is indicated; for example, 70 species are self-compatible with unknown ploidy.
+Character state abbreviations are: $I$ for self-incompatible, $C$ for self-compatible, $D$ for diploid, $P$ for polyploid, $?$ for unknown.
 Because polyploidization breaks this form of self-incompatibility, self-incompatible species with unobserved ploidy ($I?$) are taken to be diploid ($ID$), and polyploid species with unobserved breeding system ($?P$) are taken to be SC ($CP$).
-(B) Category groupings into states for each model type.
-For example, there are 70 species that are SC with unknown ploidy.
-In the ploidy only models (M1- M10) these are coded as $D$ and $P$ (uncertain/consistent with either state), in the breeding system only models (M11-M15) they are coded as $C$, and in the ploidy and breeding system models (M16-M29) they are coded as $CD$ and $CP$.
-In the hidden state models, all species were coded as consistent with both $A$ and $B$.
+(B) Category groupings into states for each model class.
+In the ploidy-only models (M1- M10), states are coded as $D$ \& $P$ when uncertain/consistent with either state; in the breeding system-only models (M11-M15) such states are coded as $C$; in the ploidy and breeding system models (M16-M29), they are coded as $CD$ \& $CP$.
+In the hidden-trait models, all species could take on either of two `hidden' character states. %IM: #33 %B: NOT USING A/B because that confuses panel A and panel B designations unnecessarily. 
 Two species, \emph{Lycium californicum} and \emph{Solanum bulbocastanum}, are simultaneously $ID$ and $CP$, and by adding them the sample adds to the total of 651 taxa used for analyses.

### 2019-07-18 manuscript/results.tex
Fixed all todos and revised discussion issues
- Itay's comments #21-33, only. Not done: #1-20, #34-35.

@@
 Results are similar when diploidization is included (\cref{supptable:M6M10}).
-Thus, the role of ploidy in net diversification is unclear, with rate differences perhaps better explained by another factor.
+Thus, when other traits are ignored, the role of ploidy in net diversification is unclear, with marginal support for diploids having higher diversification rates, but rate differences perhaps better explained by another factor.
+%IM: #21
 
@@
 When a hidden state is included, the large net diversification rate difference persists for one hidden state but is diminished for the other (M14, \cref{figure:netdivall}D).
-In the statistical model comparisons, the best two supported models include diversification differences due both to breeding system and to a hidden trait (M14 and M15, \cref{table:M11M15}).
-Thus, breeding system seems to play a role in diversification differences, though a hidden factor does as well.
+In the statistical model comparisons, the best two supported models include diversification differences due to both breeding system and to a hidden trait (M14 and M15, \cref{table:M11M15}). %IM: #22
+Breeding system seems to play a role in diversification differences, though a hidden factor does as well.
 
@@
 Model comparisons clearly favor models that include ploidy, breeding system, and the hidden trait, against the character-independent model in which the focal traits do not influence diversification (\cref{supptable:M16M20}; \cref{supptable:M21M25} with diploidization).
-% add model numbers
 
@@
     No, our focal traits are supported as having associations with diversification differences.
-    In most cases, we find moderate to strong preference for models with the focal traits as well as hidden states, over models with only hidden states \cref{table:bayesfactors,table:M11M15,supptable:M6M10,supptable:M16M20,supptable:M21M25}).
+    In most cases, we find moderate to strong preference for models with the focal traits as well as hidden states, over models with only hidden states (\cref{table:bayesfactors,table:M11M15,supptable:M6M10,supptable:M16M20,supptable:M21M25}).
+%IM: #23
 
@@
     Yes, models with hidden states that influence diversification are strongly preferred over models containing only the focal traits (\cref{supptable:testaddhidden}).
-    This means that there are potentially many sources of diversification shifts within the family.
+    This means that there are potentially many factors underlying diversification shifts within the family. %IM: #24
 
@@
     Yes, we found that allowing different types of asymmetry in transitions within and between hidden states did not change our conclusions about net diversification differences.
-    Hidden state models with asymmetric rates are, however, strongly preferred over models with equal rates between hidden states (\cref{supptable:asymmetry}), and they show stronger differences between some net diversification rates \cref{suppfigure:asymmetric}.
+    Hidden state models with asymmetric rates are, however, strongly preferred over models with equal rates between hidden states (\cref{supptable:asymmetry}), and they show stronger differences between some net diversification rates (\cref{suppfigure:asymmetric}).
 
@@
     We discuss later some further challenges in identifying diploidization.
-    We further found that our main conclusions about net diversification differences are not dependent on whether diploidization in included (\cref{suppfigure:alldip}).
+    We further found that our main conclusions about net diversification differences are not dependent on whether diploidization is included (\cref{suppfigure:alldip}). %IM: #26
 
@@
 By analogy, envisioning the states as stepping stones, the extent to which each stone grows or shrinks over time affects the utility of each possible path.
-Allowing for the different net diversification rates for each state (again using median rate estimates from M16), we find a qualitative difference in the relative pathway contributions.
+Allowing different net diversification rates for each state (again using median rate estimates from M16), we find a qualitative difference in the relative pathway contributions. %IM: #27
 The lower rate of net diversification in the $CD$ state, relative to $ID$, means that relatively fewer lineages are available to complete the second step of the two-step pathway, ending in $CP$.
 Consequently, even over long timescales, we find that the two-step pathway contributes less to the formation of polyploids (\cref{figure:pathways}, right panels) when considering diversification as well as transitions.
-
-%--------------------------------------------------
-% \subsection{Model selection key questions}
-% 
-% \subsubsection{Are diversification patterns only determined by hidden states?}
-% Whether the determination of the diversification is dependent on the focal traits and not only on the hidden state was done via the comparisons of character independent models against the focal traits w/o hidden states models (results shown in  \cref{table:bayesfactors,table:M11M15,supptable:M6M10,supptable:M16M20,supptable:M21M25}).
-% The general pattern is to moderate- to strongly prefer models with focal traits and asymmetric hidden states over the rest of the models with the exception of the ploidy only model where the character independent model (M2) seems to be equally preferred($K=0.5$) as the model with ploidy and asymmetric hidden rates (M4, \cref{table:bayesfactors}).
-% 
-% \subsubsection{Are hidden states necessary?}
-% In the comparisons of simple diversification models of focal traits against models with hidden states we found that models with asymmetric hidden traits are strongly preferred over simpler models (summarized in table \cref{supptable:testaddhidden}).
-% 
-% \subsubsection{Is a second trait adding information to the diversification process?}
-% As summarized in  \cref{table:lumped}, it is moderately preferred to add breeding system to ploidy only models (\cref{suppfigure:lumpedDP}).
-% When adding ploidy information to breeding system models to create a three-state model, we found that there is no evidence that it is preferred (M28 vs. M16 \cref{suppfigure:lumpedIC}B and C). 
-% However, it becomes moderately preferable in the presence of hidden states (M29 vs. M23  \cref{suppfigure:lumpedIC}E and C).
-% 
-% \subsubsection{Is there evidence for diploidization?}
-% We considered models both with and without diploidization in order to explore its effects on the estimates of state-dependent diversification.
-% For the models that only include diploid and polyploid states (M6-M10) the results are similar in terms of the effect of diversification (\cref{suppfigure:alldip}).
-% When comparing models with diploidization against models without diploidization we found moderate evidence that models containing the parameter $\delta$ are preferred over models without diploidization (summarized in \cref{supptable:testdiploidization}).
-% 
-% \subsubsection{Do asymmetric rates in hidden states models affect the pattern of diversification?}
-% When comparing different types of asymmetry across hidden states models we found that the asymmetry did not change the direction of effects in net diversification results .
-However, models with asymmetric hidden rates reduced the overlap amongst net diversification posterior distributions as shown in \cref{suppfigure:asymmetric}.
-Furthermore, hidden state models with asymmetric traits are strongly preferred over models with equal rates between hidden states (summarized in \cref{supptable:asymmetry}).
-%--------------------------------------------------

### 2019-07-18 manuscript/tables.tex
- Itay's comments #21-33, only. Not done: #1-20, #34-35.

@@
 \end{tabular}
-\caption{Bayes factors of  ploidy only models without diplodization.
+\caption{Bayes factors for ploidy-only models without diplodization.
 Results indicate that a character independent model (M2) is strongly preferred over BiSSE (M1)  model.
@@
 \end{tabular}
-\caption{Bayes factors of  ploidy only models with diplodization.
-Results indicate that a character independent model (M6) is strongly preferred over BiSSE (M1)  model.
+\caption{Bayes factors for ploidy-only models with diplodization. 
+Results indicate that a character-independent model (M6) is strongly preferred over BiSSE (M1) model. 
 M6 is moderately preferred over any of the HiSSE models with asymmetric hidden rates (M9, M10).}
@@
 \end{tabular}
-\caption{Bayes factors of breeding system only models. Results indicate that the HiSSE models with asymmetric rates (M14, M15) are strongly preferred over every other model.}
+\caption{Bayes factors for breeding system-only models. 
+Results indicate that the HiSSE models with asymmetric rates (M14, M15) are strongly preferred over other models.}
 \label{table:M11M15}
@@
 \end{tabular}
-\caption{Test for addition of hidden states in models via Bayes factors. Models with hidden states are strongly preferred over simpler models}
+\caption{Test for addition of hidden states in models via Bayes factors. 
+Models with hidden states are strongly preferred over simpler models}
 \label{table:testaddhidden}
@@
 \end{tabular}
-\caption{Test for inclusion of a diploidization rate via Bayes factors. Models with diploidzation are moderately preferred over models that do not include a diploidization rate}
+\caption{Test for inclusion of a diploidization rate via Bayes factors. 
+Models with diploidzation are moderately preferred over models that do not include a diploidization rate}
 \label{table:testdiploidization}

### 2019-07-18 manuscript/abstract.tex
All co-authors input added, figures corrected

@@
 We investigate the connections of these traits, including their interaction, to speciation and extinction rates in Solanaceae.
-We show that the effect of ploidy on diversification can be largely explained by its correlation with breeding system and that additional unknown factors, alongside breeding system, determine diversification rates.
+We show that the effect of ploidy on diversification can be largely explained by its correlation with breeding system and that additional unknown factors, alongside breeding system, influence diversification rates.
 These results are largely robust to allowing for diploidization.

### 2019-07-18 manuscript/discussion.tex
All co-authors input added, figures corrected

@@
 Focusing first on ploidy and then on breeding system, we found that considering each trait in isolation provides an incomplete story.
-Considering them together, and in conjunction with another hidden factor, provides a fuller picture of macroevolutionary dynamics within Solanaceae.
+Considering them together, and in conjunction with another hidden factor, provides a more complete picture of macroevolutionary dynamics within Solanaceae.
 We hope our work serves as an example of how phylogenetic comparative methods can be used to disentangle the contributions of interacting traits to heterogeneous lineage diversification, and how to statistically argue for increasing complexity in diversification modeling.
@@
 %Here Rosana adds the value of adding traits with the lumped models
-As more trait information becomes available for macroevolutionary studies, it is important not only recognizing the role of hidden states as a part of a general modeling approach, but also questioning whether adding more traits to diversification studies is relevant.
-More generally, when analyzing the value of adding additional traits to an already complex model with or without hidden states, we find that using lumped models  is a key tool to verify that adding a trait improves our understanding of the diversification patterns (\cref{figure:lumped,table:lumped}).
+As more trait information becomes available for macroevolutionary studies, it is not only important to recognize the role of hidden states as a part of a general modeling approach, but also to question whether adding more traits to diversification studies is justified.
+More generally, we find that lumped models are useful for assessing the value of adding additional traits to already complex diversification models(\cref{figure:lumped,table:lumped}).
 %R-removed this: %In our analyses, we find that the transition rates between and within hidden traits did not affect the direction of the net diversification differences. 
@@
 Anagenetic and cladogenetic changes can be separated with phylogenetic models \citep{mayrose_2011, goldberg_2012, magnuson-ford_2012}. 
-These have been previously applied to estimate the relative contribution of anagenetic and cladogenetic shifts in breeding system \citep{goldberg_2012} and polyploidization \citep{zhan_2016, freyman_2017}.
-We did not explore cladogenetic trait change because of the vast increase in parameter space and model number, outside the scope of our already complex and extensive modeling framework. %IM: #29
+These have been applied to estimate the relative contribution of anagenetic and cladogenetic shifts in breeding system \citep{goldberg_2012} and polyploidization \citep{zhan_2016, freyman_2019}.
+We did not explore cladogenetic trait change because they would greatly increase in state space and model number of our already complex and extensive modeling framework.
+%IM: #29 and GB suggested that chance
 Although \citet{goldberg_2012} found that allowing cladogenetic changes did not substantially affect inference of net diversification rates associated with breeding system, future work could test whether this process affects diversification rate estimates with the more complex state and parameter spaces of our other models.
@@
 Although this question focuses on the process of state transitions, we also show that its answer is affected by the process of lineage diversification.
-We used a simple mathematical approach to investigate the contributions of the two pathways, but future work could instead rely on stochastic character mapping to more directly estimate the numbers of each type of transition.
+We used a simple mathematical approach to investigate the contributions of the two pathways, but future work could instead rely on stochastic character mapping to estimate the numbers of each type of transition more directly.
 
@@
 %
-First, our support for the direct pathway is consistent with the idea that breakdown of SI by whole genome duplication---via diploid `heteroallelic' pollen---may often trigger the evolution of gender dimorphism as a different mechanism of inbreeding avoidance \citep{miller_2000}.
+First, our support for the direct pathway is consistent with the idea that breakdown of SI by whole genome duplication---via diploid `heteroallelic' pollen---may trigger the evolution of gender dimorphism as a different mechanism of inbreeding avoidance \citep{miller_2000}.
 A further test of this hypothesis would additionally examine the propensity of polyploids generated through either pathway to become dioecious \citep{robertson_2011}.
 Second, we might wonder whether the propensity for a polyploid species to diversify depends on whether it arose via the one-step or two-step pathway.
-This could be tested with a different form of a hidden state model, in which the polyploid state is subdivided into parts, reflecting the arriving path taken to that state. %IM: #30
+This could be tested with a different form of a hidden state model, in which the polyploid state is subdivided into parts, reflecting the arriving path taken to that state. %IM: #30 and GB 
 Such an estimate of path-dependence could also include the possibility of different diversification rates for the subdivided $CP$ substates.

### 2019-07-18 manuscript/introduction.tex
All co-authors input added, figures corrected

@@
 Polyploidization is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
-The widespread prevalence of variation in ploidy has long been considered a salient feature of flowering plant lineages \citep{stebbins1938}. 
-Increase in ploidy can alter many traits and affect a variety of evolutionary and ecological processes \citep{ramsey_2002, sessa_2019}.
-Polyploids were found to have an overall lower net diversification rate than diploids across many vascular plant clades \citep{mayrose_2011, mayrose_2015}. 
-However, recent genomic studies have inferred numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications may have played an important macroevolutionary role driving innovation and diversification in plants.
+The widespread variation in ploidy has long been considered a salient feature of flowering plant lineages \citep{stebbins1938}. 
+An increase in ploidy can alter many traits and affect a variety of evolutionary and ecological processes \citep{ramsey_2002, sessa_2019}.
+At shallow evolutionary time scale, polyploids were found to have an overall lower net diversification rate than diploids across many vascular plant clades \citep{mayrose_2011, mayrose_2015}.
+%IM suggestion
+However, recent genomic studies have inferred numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications may have played an important  role driving innovation and diversification in plants.
 Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies pervasive diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
@@
 Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are also remarkably common and crucially affect the distribution and amount of genetic variation in populations \citep{barrett2013}.
-In particular, a variety of genetic self-incompatibility (SI) systems cause plants to reject their own pollen, and loss of such mechanisms, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974, igic_2008}.
+In particular, a variety of genetic self-incompatibility (SI) systems cause plants to reject their own pollen, and the loss of such mechanisms, yielding self-compatibility (SC), is a commonly observed transition in flowering plant evolution \citep{stebbins1974, igic_2008}.
 Previous analyses reported higher rates of diversification for SI than for SC lineages in Solanaceae \citep{goldberg_2010}. 
-Similarly, heterostylous SI lineages in Primulaceae seem to diversify faster \citep{devos2014}, as do outcrossing lineages in Onagraceae \citep{freyman_2019}.
+Similarly, heterostylous SI lineages in Primulaceae seem to diversify faster than SC lineages \citep{devos2014}, as do outcrossing lineages in Onagraceae \citep{freyman_2019}.
 Although these findings suggest a consistent macroevolutionary role of breeding system, it is unlikely to be the sole character determining lineage diversification.
@@
 Polyploidy and self-fertilization are widely thought to be associated \citep{stebbins1950}.
-There are several possible general explanations: whole genome duplication may facilitate the transition to selfing by masking inbreeding depression, self-fertilization may facilitate polyploidy establishment by avoiding the triploid bridge, or both selfing and polyploidy may tend to be found in weedy or peripheral populations \citep{ramsey_1998, barringer2007, barrett2008, husband2008}.
+Whole genome duplication may facilitate the transition to selfing by masking inbreeding depression, or self-fertilization may facilitate polyploidy establishment by by avoiding the lower fitness that characterize early-forming polyploids \citep{levin_1975,ramsey_1998, barringer2007, barrett2008, husband2008}.%IM suggestion and Gordon suggestion
 Additionally, in RNase-based gametophytic SI systems, polyploidization directly causes the loss of SI \citep{stout1942, lewis1947}.
@@
 RNase-based SI is regarded as ancestral in eudicots \citep{igic_2001,steinbachs_2002}, and it is expressed in all SI species of Solanaceae examined to date.
-The consequent absence of SI polyploids in this family yields a strong correlation between ploidy and breeding system \citep{robertson_2011}.
+The absence of SI polyploids in this family yields a strong correlation between ploidy and breeding system \citep{robertson_2011}.
 
@@
 First, we investigate their joint influence on rates of speciation and extinction.
-Each character alone has been shown to associate with differential lineage diversification, but if their effects are not additive, studying them separately may not reveal the effects of their combinations.
+Each character alone has been shown to be associated with differential lineage diversification, but if their effects are not additive, studying them separately may not reveal their combined effect.
 Second, we examine the order of transitions in the two characters.
@@
 In the present study, we employ an extended framework of state speciation and extinction models, which simultaneously model transitions between the discrete states of a trait and different rates of speciation and extinction associated with each of those states \citep[`SSE' models;][]{maddison_2007, fitzjohn_2012}.
-We start by fitting binary state speciation and extinction models to each ploidy and breeding system independently \citep{maddison_2007}.
-We follow by fitting models that incorporate hidden states, which reduce the chance that an effect of one trait on diversification would be mistaken for the effect of an unobserved factor \citep{beaulieu_2016}.
-We compare the proposed models against their character independent counterparts \citep{beaulieu_2016} to investigate whether ploidy or breeding system are actually informative of the diversification process.
-Next, we model ploidy and breeding system jointly to assess their combined influence on diversification, with or without an additional hidden character.
+We start by fitting binary state speciation and extinction models to ploidy and breeding system independently \citep{maddison_2007}.
+We follow by fitting models that incorporate hidden states, which reduce the chance that the effect of the focal trait (ploidy or breeding system) on diversification is found to be significant when in reality, it is simply background heterogeneity in the diversification process what produces the diversification patterns in the phylogeny \citep{beaulieu_2016}.
+%IM: added a little more explanation
+We compare the proposed models against their character independent counterparts \citep{beaulieu_2016} to investigate whether ploidy or breeding system are actually affect the diversification process.
+Next, we model ploidy and breeding system jointly to assess their combined influence on diversification, with or without an additional hidden state.
 Using the ploidy and breeding system model without hidden states, we quantify the relative contributions of the two pathways from SI diploids to SC polyploids.

### 2019-07-18 manuscript/main.tex
All co-authors input added, figures corrected

@@
 The computing resources were provided by the Minnesota Supercomputing Institute (MSI) at the University of Minnesota.
-This work was supported by the National Science Foundation grant DEB-1655478 (to EEG), DEB-1655692 and NESCent sabbatical scholar award EF-0905606 (to BI), and the Israel Science Foundation 961/17 (to IM).
+This work was supported by the National Science Foundation grant DEB-1655478 (to EEG and IM), DEB-1655692 and NESCent sabbatical scholar award EF-0905606 (to BI), and the Israel Science Foundation 961/17 (to IM).
 %IM: #32

### 2019-07-18 manuscript/methods.tex
All co-authors input added, figures corrected

@@
 For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus/family, or based on SI/SC classification.
-For example, \textit{Solanum betaceum} did not have information about ploidy level but it has 24 chromosomes, so since $x=12$ is the base chromosome number of the genus \textit{Solanum} \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid.
+For example, \textit{Solanum betaceum} did not have information about ploidy level but it has 2n=24 chromosomes, so since $x=12$ is the base chromosome number of the genus \textit{Solanum} \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid.
 Additionally, because of the absence of SI polyploids (explained above and below), species known to be SI could be scored as diploid.
@@
 
-Breeding system states were scored as self-incompatible, $I$, or self-compatible, $C$, based on results curated from the literature and original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
+Breeding system states were scored as self-incompatible, $I$, or self-compatible, $C$, based on results curated from the literature \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
+%IM suggestion
 Most species could unambiguously be coded as either $I$ or $C$ \citep{raduski_2012}.
@@
 We additionally fit three models in which both ploidy and a hidden trait could influence diversification (M3--M5).
-These models differ in whether transitions between the hidden states are symmetric or asymmetric, and whether the polyploidization rate depends on the hidden state.
-Comparing M1--M5 allows us to test whether ploidy is associated with diversification differences on top of the differences potentially explained by an unknown factor (\cref{table:bayesfactors}).
+These models differ in whether transitions between the hidden states are symmetric (M3) or asymmetric(M4), and whether the polyploidization rate depends on the hidden state(M5). %WF suggestion
+Comparing M1--M5 allows us to test whether ploidy is associated with diversification differences on top of the differences potentially explained by an unknown factor.
 
-We further fit the analogues of these five models but including a rate parameter $\delta$ for transitions from polyploid to diploid (M6--M10,  \cref{supptable:M6M10}).
+We further fit the analogues of these five models but including a rate parameter $\delta$ for transitions from polyploid to diploid (M6--M10).
 These comparisons allow us to assess whether our conclusions about ploidy and diversification are robust to the possibility of diploidization.
@@
 We then add models to allow diversification to be influenced by only a hidden trait (M12), or by both breeding system and a hidden trait (M13--M15, with varying degrees of complexity in the hidden trait transitions \cref{table:M11M15}).
-Similar models were used by \citet{freyman_2017} used to study diversification in Onagraceae.
+Similar models were used by \citet{freyman_2019} to study diversification in Onagraceae.
 
@@
 Thus, there is strong evidence in Solanaceae that the $I$ state is ancestral in the family, and that the SI mechanism was not regained.
-Therefore, for all breeding system models, we estimated a transition rate from $I$ to $C$  but not the reverse ($q_{CI}=0$).
+For all breeding system models, we estimated a transition rate from $I$ to $C$  but not the reverse ($q_{CI}=0$).
 % This is an important assumption because ancestral state reconstruction in the models might differ when polyploidy and breeding system are analyzed as independent in the Solanaceae tree.
@@
 
-As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by ploidy, breeding system, and a hidden trait (M18--M20) with varying degrees of complexity in the hidden trait transitions (\cref{supptable:M16M20}).
-We add this hidden trait layer on top of our three-state model \citep[similar to][]{caetano_2018, huang_2018}.
-We also fit the analogous models but allowing for diploidization (M21--M25, \cref{supptable:M21M25}).
+As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by ploidy, breeding system, and a hidden trait (M18--M20) with varying degrees of complexity in the hidden trait transitions
+\citep[similar to][]{caetano_2018, huang_2018}. %IM suggestion
+We also fit the analogous models but allowing for diploidization (M21--M25).
 
@@
 To answer this question with statistical model comparisons requires comparing the likelihood of the data given each model.
-This is impossible for the ploidy and breeding system models presented so far, however, because the data are different for the different models: they use either the $D/P$ or the $I/C$ or the $IC/CD/CP$ state spaces.
+This is impossible for the ploidy and breeding system models presented so far, however, because the data are different for the different models: they use either the $D/P$ or the $I/C$ or the $ID/CD/CP$ state spaces.
 Therefore, the use of different data results in incomparable models.
@@
 In order to compare fits of ploidy-only \vs breeding system-only \vs combined trait models, we use the technique of `lumping' states together \citep{tarasov_2019}.
-We use the state space of the $IC/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
+We use the state space of the $ID/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
+To lump states requires that the transition rates from the lumped state to the singular state be equal \citep{tarasov_2019}.
 First we lump together $ID$ and $CD$ to form the diploid state, mimicking the $D/P$ model (M26). 
-Then we lump together $CD$ and $CP$ to form the self-compatible state, mimicking the $I/C$ model (M28).
+Proposing a lumped ploidy model by aggregating $ID$ and $CD$ requires forcing the rate of polyploidization from $ID$ and $CD$ to $CP$ to be equal (i.e. $\rho_0=\rho_I=\rho_c$), but also requires
+assuming that the rates of speciation and extinciton for $ID$ and $CD$ to be equal.
+Therefore, we define the new parameters $\lambda_D$ and $\mu_D$  that are the same for each of the two diploid states $ID$ and $CD$.%IM suggestion, describe more the lumping
+We used the same procedure to lump together $CD$ and $CP$ to form the self-compatible state, mimicking the $I/C$ model (M28).
+In this particular case, the rate from $CD$ to $CP$ back to $ID$ is zero and equal for both, so the model is lumpable. 
+However, to fully mimic the breeding system model, we assume that the rate of selfing is equal (i.e. $q_0=q_{IC}=rho_I$) and the rates
+of speciation and extinction for both $CD$ and $CP$ are the same (new parameters $\lambda_C$ and $\mu_C$).
 We further add a hidden character to each of these models (M27 and M29), and then compare this group of models (\cref{table:lumped}).
@@
 We do not include additional models with diploidization because this reverse ploidy transition renders the models non-lumpable.
-To lump states requires that the transition rates from the lumped state to the singular state be equal \citep{tarasov_2019}.
 When including diploidization, transitions from $CP$ to $CD$ are at rate $\delta$ but transitions from $CP$ to $ID$ do not occur.
 Because $ID$ and $CD$ would be lumped to mimic the $D/P$ model, this model is non-lumpable when $\delta \ne 0$.
-When lumping $CD$ and $CP$, however, the transition rate to $ID$ is zero for both, so the SC states are lumpable. 
 Thus, we can compare models to test whether it is advantageous to include both traits, but only when ignoring diploidization.
@@
 We quantify the relative contribution of these pathways to polyploidy in two ways, each using the median estimates of rates from the simplest model that includes both traits (M16).
+Our results will differ from those of \citet{robertson_2011} because our transition rate estimates come from a dated phylogeny and a model that allows for state-dependent diversification.
+%IM suggestion
 
@@
 Closed-form solutions for the two pathway probabilities are provided in \citet{robertson_2011}.
-Our results will differ from those of \citet{robertson_2011} because our transition rate estimates come from a dated phylogeny and a model that allows for state-dependent diversification.
 
@@
 % R- fixed
-We can still use ratios, however, to consider the relative contribution of each pathway, analogous to the normalized age structure in a growing population.
-% R- here a reference would be nice  E: Does Jake have a favorite ref?
-Otherwise, I guess a basic ecology textbook.
+We can still use ratios, however, to consider the relative contribution of each pathway, analogous to the normalized age structure in a growing population \citep{leslie_1945}.
+ % R- here a reference would be nice  E: Does Jake have a favorite ref? Otherwise, I guess a basic ecology textbook.-DONE
 %
@@
 
-Parameters for each of the 29 models were coded as graphical models and Bayesian statistical inference was performed RevBayes \citep{hoehna_2016}.
+Parameters for each of the 29 models were coded as graphical models and Bayesian statistical inference was performed with RevBayes \citep{hoehna_2016}.
 Scripts for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.- Happens after publication
@@
 Additionally, ancestral states at each node in the phylogeny were sampled jointly during the MCMC analyses every 100 generations.
+% Itay says remove but some reviewers were interested in the distribution of the states. I say leave we don't need to discuss. I'll mention briefly.
 Ancestral state estimations for all models show the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes for each of the models in \cref{figure:netdivall}.

### 2019-07-18 manuscript/results.tex
All co-authors input added, figures corrected

@@
     Hidden state models with asymmetric rates are, however, strongly preferred over models with equal rates between hidden states (\cref{supptable:asymmetry}), and they show stronger differences between some net diversification rates (\cref{suppfigure:asymmetric}).
+    %Rosana added the following lines to alleviate the comment from Itay about hidden states
+    The effect of the asymmetry of the hidden rate transitions  is better observed in the ancestral state estimations (\cref{suppfigure:DPnodipABasr,suppfigure:ICABasr,suppfigure:IDCDCPnodipABasr}), which show that moving out of state A (dark colors) happens quickly with rate $\alpha$,  whereas evolving out of hidden state B (light colors) is slow with rate $\beta$.
 

### 2019-07-18 manuscript/response.tex
slight change in the signature of response

@@
 
-We hope the changes and the new models are satisfactory to the editorial staff and the reviewers.
+W hope the changes and the new models are satisfactory to the editorial staff and the reviewers.
 
@@
 
-The Authors \\
-(Rosana Zenil-Ferguson et al.)
+Rosana Zenil-Ferguson, Gordon Burleigh, Boris Igi\'c, Itay Mayrose, and Emma Goldberg.
 

### 2019-07-19 manuscript/discussion.tex
- Final ms changes pass-through, mainly spaces, commas, spelling, etc. - Leaving FIXME marks for galley proofs. - Will update archive once more with submission version.

@@
 As more trait information becomes available for macroevolutionary studies, it is not only important to recognize the role of hidden states as a part of a general modeling approach, but also to question whether adding more traits to diversification studies is justified.
-More generally, we find that lumped models are useful for assessing the value of adding additional traits to already complex diversification models(\cref{figure:lumped,table:lumped}).
+More generally, we find that lumped models are useful for assessing the value of adding additional traits to already complex diversification models (\cref{figure:lumped,table:lumped}).
 %R-removed this: %In our analyses, we find that the transition rates between and within hidden traits did not affect the direction of the net diversification differences. 
@@
 These have been applied to estimate the relative contribution of anagenetic and cladogenetic shifts in breeding system \citep{goldberg_2012} and polyploidization \citep{zhan_2016, freyman_2019}.
-We did not explore cladogenetic trait change because they would greatly increase in state space and model number of our already complex and extensive modeling framework.
-%IM: #29 and GB suggested that chance
+We did not explore cladogenetic trait change because of the requisite increase in parameter space and model number, outside the scope of our already complex and extensive modeling framework.
+%B: restored IM:#29 and removed "they" pointing to "trait change."
 Although \citet{goldberg_2012} found that allowing cladogenetic changes did not substantially affect inference of net diversification rates associated with breeding system, future work could test whether this process affects diversification rate estimates with the more complex state and parameter spaces of our other models.

### 2019-07-19 manuscript/introduction.tex
- Final ms changes pass-through, mainly spaces, commas, spelling, etc. - Leaving FIXME marks for galley proofs. - Will update archive once more with submission version.

@@
 Polyploidy and self-fertilization are widely thought to be associated \citep{stebbins1950}.
-Whole genome duplication may facilitate the transition to selfing by masking inbreeding depression, or self-fertilization may facilitate polyploidy establishment by by avoiding the lower fitness that characterize early-forming polyploids \citep{levin_1975,ramsey_1998, barringer2007, barrett2008, husband2008}.%IM suggestion and Gordon suggestion
+Whole genome duplication may facilitate the transition to selfing by masking inbreeding depression, or self-fertilization may facilitate establishment of early-forming polyploids by avoiding the lower fitness \citep{levin_1975,ramsey_1998, barringer2007, barrett2008, husband2008}.%IM suggestion and Gordon suggestion %B: reworded because polyploidy is a process, so its establishment is not the issue.
 Additionally, in RNase-based gametophytic SI systems, polyploidization directly causes the loss of SI \citep{stout1942, lewis1947}.
@@
 First, we investigate their joint influence on rates of speciation and extinction.
-Each character alone has been shown to be associated with differential lineage diversification, but if their effects are not additive, studying them separately may not reveal their combined effect.
+Each character alone is thought to be associated with differential lineage diversification but, if their effects are not additive, studying them separately may not reveal their combined effect.
 Second, we examine the order of transitions in the two characters.
@@
 We start by fitting binary state speciation and extinction models to ploidy and breeding system independently \citep{maddison_2007}.
-We follow by fitting models that incorporate hidden states, which reduce the chance that the effect of the focal trait (ploidy or breeding system) on diversification is found to be significant when in reality, it is simply background heterogeneity in the diversification process what produces the diversification patterns in the phylogeny \citep{beaulieu_2016}.
+We follow by fitting models that incorporate hidden states, which reduce the chance that the effect of the focal trait (ploidy or breeding system) on diversification is found to be significant when in reality, it may simply represent background heterogeneity in the diversification process that produces the phylogenetic patterns \citep{beaulieu_2016}.
 %IM: added a little more explanation
-We compare the proposed models against their character independent counterparts \citep{beaulieu_2016} to investigate whether ploidy or breeding system are actually affect the diversification process.
-Next, we model ploidy and breeding system jointly to assess their combined influence on diversification, with or without an additional hidden state.
+We compare the proposed models against their character-independent counterparts \citep{beaulieu_2016} to investigate whether ploidy or breeding system affect the diversification process.
+Next, we model ploidy and breeding system jointly to assess their combined influence on diversification, with or without an additional hidden character.
+%B: additional state implies that we would introduce another state, say 'C' to 'A' and 'B'.
 Using the ploidy and breeding system model without hidden states, we quantify the relative contributions of the two pathways from SI diploids to SC polyploids.

### 2019-07-19 manuscript/main.tex
- Final ms changes pass-through, mainly spaces, commas, spelling, etc. - Leaving FIXME marks for galley proofs. - Will update archive once more with submission version.

--- 2019-07-18 manuscript/main.tex
@@
 The computing resources were provided by the Minnesota Supercomputing Institute (MSI) at the University of Minnesota.
-This work was supported by the National Science Foundation grant DEB-1655478 (to EEG), DEB-1655692 and NESCent sabbatical scholar award EF-0905606 (to BI), and the Israel Science Foundation 961/17 (to IM).
+This work was supported by the National Science Foundation grants DEB-1655478 (to EEG and IM), NSF DEB-1655692 \& NESCent sabbatical scholar award EF-0905606 (to BI), and the Israel Science Foundation 961/17 (to IM).
 %IM: #32
-%B: new:
 We thank Dan Schoen and three anonymous reviewers for providing uncommonly thoughtful and helpful reviews, as well as Spencer Barrett for continually inspiring our work.

### 2019-07-19 manuscript/methods.tex
- Final ms changes pass-through, mainly spaces, commas, spelling, etc. - Leaving FIXME marks for galley proofs. - Will update archive once more with submission version.

@@
 For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus/family, or based on SI/SC classification.
-For example, \textit{Solanum betaceum} did not have information about ploidy level but it has 2n=24 chromosomes, so since $x=12$ is the base chromosome number of the genus \textit{Solanum} \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid.
+For example, \textit{Solanum betaceum} did not have information about ploidy level, but it has 2n=24 chromosomes, and  $x=12$ is the base chromosome number of the genus \textit{Solanum} \citep{olmstead_2007}, so we assigned \textit{S.~betaceum} as diploid.
 Additionally, because of the absence of SI polyploids (explained above and below), species known to be SI could be scored as diploid.
@@
 
-Breeding system states were scored as self-incompatible, $I$, or self-compatible, $C$, based on results curated from the literature \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
-%IM suggestion
+Breeding system states were scored as self-incompatible, $I$, or self-compatible, $C$, based on results curated from the literature  \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012} and original experimental crosses (B.I.
+and E.E.G., unpub.
+data).
+%IM suggestion %B:  I returned the aside about original data.
+I don't think Itay knows that we performed original crosses (many), and they are incorporated in the data, without being present in "the literature"
+%was originally: Breeding system states were scored as self-incompatible, $I$, or self-compatible, $C$, based on results curated from the literature and original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
 Most species could unambiguously be coded as either $I$ or $C$ \citep{raduski_2012}.
@@
 
-In order to investigate the association between ploidy level and diversification, we first employed a model (labeled M1), previously used by \citet{mayrose_2011}, with each species classified as diploid $(D)$ or polyploid $(P)$.
+We first employed a model (labeled M1), previously used by \citet{mayrose_2011}, with each species classified as diploid $(D)$ or polyploid $(P)$.
 Although this model can be powerful in studies of trait evolution, it is prone to incorrectly reporting that a trait is associated with diversification differences \citep{maddison_2015, rabosky_2015}.
 We therefore define several models that incorporate additional forms of diversification rate heterogeneity.
-
 The second ploidy model (M2) includes a binary hidden trait that subdivides each observed state.
@@
 
-We additionally fit three models in which both ploidy and a hidden trait could influence diversification (M3--M5).
+We fit three models in which both ploidy and a hidden trait could influence diversification (M3--M5).
 These models differ in whether transitions between the hidden states are symmetric (M3) or asymmetric(M4), and whether the polyploidization rate depends on the hidden state(M5). %WF suggestion
 Comparing M1--M5 allows us to test whether ploidy is associated with diversification differences on top of the differences potentially explained by an unknown factor.
-
 We further fit the analogues of these five models but including a rate parameter $\delta$ for transitions from polyploid to diploid (M6--M10).
@@
 This is the same model as in the analysis presented in \citet{goldberg_2010} but with an updated phylogeny \citep{sarkinen_2013} and a larger aggregated dataset.
-
 We then add models to allow diversification to be influenced by only a hidden trait (M12), or by both breeding system and a hidden trait (M13--M15, with varying degrees of complexity in the hidden trait transitions \cref{table:M11M15}).
@@
 % R: In the lumped models I call them something different q_0 and rho_0 because they are different parameters but I don't think we should explain.
-
 As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by ploidy, breeding system, and a hidden trait (M18--M20) with varying degrees of complexity in the hidden trait transitions
@@
 We use the state space of the $ID/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
-To lump states requires that the transition rates from the lumped state to the singular state be equal \citep{tarasov_2019}.
-First we lump together $ID$ and $CD$ to form the diploid state, mimicking the $D/P$ model (M26). 
+Lumping states requires that the transition rates from the lumped state to the singular state be equal \citep{tarasov_2019}.
+First, we lump together $ID$ and $CD$ to form the diploid state, mimicking the $D/P$ model (M26). 
 Proposing a lumped ploidy model by aggregating $ID$ and $CD$ requires forcing the rate of polyploidization from $ID$ and $CD$ to $CP$ to be equal (i.e. $\rho_0=\rho_I=\rho_c$), but also requires
 assuming that the rates of speciation and extinciton for $ID$ and $CD$ to be equal.
-Therefore, we define the new parameters $\lambda_D$ and $\mu_D$  that are the same for each of the two diploid states $ID$ and $CD$.%IM suggestion, describe more the lumping
+Therefore, we define the new parameters $\lambda_D$ and $\mu_D$  that are the same for each of the two diploid states $ID$ and $CD$. %IM suggestion, describe more the lumping
 We used the same procedure to lump together $CD$ and $CP$ to form the self-compatible state, mimicking the $I/C$ model (M28).
 In this particular case, the rate from $CD$ to $CP$ back to $ID$ is zero and equal for both, so the model is lumpable. 
-However, to fully mimic the breeding system model, we assume that the rate of selfing is equal (i.e. $q_0=q_{IC}=rho_I$) and the rates
-of speciation and extinction for both $CD$ and $CP$ are the same (new parameters $\lambda_C$ and $\mu_C$).
+However, to fully mimic the breeding system model, we assume that the rate of selfing is equal (i.e.
+$q_0=q_{IC}=rho_I$) and the rates of speciation and extinction for both $CD$ and $CP$ are the same (new parameters $\lambda_C$ and $\mu_C$).
 We further add a hidden character to each of these models (M27 and M29), and then compare this group of models (\cref{table:lumped}).
@@
 We quantify the relative contribution of these pathways to polyploidy in two ways, each using the median estimates of rates from the simplest model that includes both traits (M16). 
-Our results will differ from those of \citet{robertson_2011} because our transition rate estimates come from a dated phylogeny and a model that allows for state-dependent diversification.
-%IM suggestion
+Our results differ from those of \citet{robertson_2011} in part because our approach relies on a dated phylogeny and models that allow for state-dependent diversification. %IM suggestion
 
@@
 We consider unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and no strong support for diploidization (see below).
-
 First, we consider only the rates of transitions between these states, placing them in the propagation matrix \myvec{Q}.
@@
 Closed-form solutions for the two pathway probabilities are provided in \citet{robertson_2011}.
-
 Second, we consider not only transitions between states but also diversification within each state.
@@
 The calculations of the propagation matrix are not probabilities because diversification changes the number of lineages as time passes. 
-%B: FIXME, we'll hear complaints again about "this," "these," and ... "their." (below)
-% R- fixed
 We can still use ratios, however, to consider the relative contribution of each pathway, analogous to the normalized age structure in a growing population \citep{leslie_1945}.
- % R- here a reference would be nice  E: Does Jake have a favorite ref? Otherwise, I guess a basic ecology textbook.-DONE
-%
 % E & R todo ?: for the case with net div, write out the propagation matrix and equations for pathway contributions
 % R- couldn't solve to get the exact solutions turns out it is really hard.
-%
-% E: I don't think we need this section now that diploidization is in each of the sections above.  But I'm saving this text in case we want to put some of it in Results.
-%--------------------------------------------------
-% \subsubsection{Diploidization as an exploratory hypothesis}
-%
-% The models described above allow polyploidization but not its reverse.
-% We explored the effects of allowing diploidization by adding a parameter $\delta$ for transitions from $P$ back to $D$ (in the $D/P$ and $D/P+A/B$ models) or from $CP$ back to $CD$ (in the $I/C$ and $I/C+A/B$ models); we did not allow transitions from $CP$ back to $ID$ because that would include regain of SI.
-% Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010, glick2014, mayrose_2015, freyman_2017}, which assume only aneuploid change and polyploidization.
-% Our ploidy classifications, however, when possible also incorporated information on chromosome multiplicity at the genus level and author conclusions within original studies with alternative sources of information (\eg geographic distribution, genus ploidy distribution).
-% Since it is not known whether diploidization can be detected under these additional ploidy classifications, we fit models with and without diploidization in order to test whether the conclusions about diversification are sensitive to including diploidization.
-% In our models, the diploidization parameter (or its absence, $\delta=0$), represents an opportunity to explore a possibly significant assumption that may not decisively affect interactions among ploidy, breeding system, and diversification.
-%--------------------------------------------------
 
@@
 Parameters for each of the 29 models were coded as graphical models and Bayesian statistical inference was performed with RevBayes \citep{hoehna_2016}.
-Scripts for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.- Happens after publication
+Scripts for analyses and key results are available as Supplementary Information. 
 We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $651/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
@@
 Additionally, ancestral states at each node in the phylogeny were sampled jointly during the MCMC analyses every 100 generations. 
-% Itay says remove but some reviewers were interested in the distribution of the states. I say leave we don't need to discuss. I'll mention briefly.
+%R: Itay says remove but some reviewers were interested in the distribution of the states.
+I say leave we don't need to discuss.
+I'll mention briefly.
+%B: I am not sure what this comment refers to (up or down?), but I'm not touching.
 Ancestral state estimations for all models show the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes for each of the models in \cref{figure:netdivall}.
@@
 In the ploidy-only models (M1- M10), states are coded as $D$ \& $P$ when uncertain/consistent with either state; in the breeding system-only models (M11-M15) such states are coded as $C$; in the ploidy and breeding system models (M16-M29), they are coded as $CD$ \& $CP$.
-In the hidden-trait models, all species could take on either of two `hidden' character states. %IM: #33 %B: NOT USING A/B because that confuses panel A and panel B designations unnecessarily. 
+In the hidden-trait models, all species could take on either of two `hidden' character states. %IM: #33 
 Two species, \emph{Lycium californicum} and \emph{Solanum bulbocastanum}, are simultaneously $ID$ and $CP$, and by adding them the sample adds to the total of 651 taxa used for analyses.
@@
 
-% E: Could we include a figure of the traits on the tree?  At least as supp info. FIXME
-%B: "K=ln(BF(M1,M2))" -> "K=ln(BF($M_{i}$,$M_{ii}$))" because we use M1 and M2 for model names.
+% E: Could we include a figure of the traits on the tree?  At least as supp info. %FIXME
+
 \begin{table}[]
@@
 
-% E todo: re-run with new rate estimates; add panel letters?
+% E: %FIXME: re-run with new rate estimates; add panel letters?
 \begin{figure}
@@
 \end{tabular}
-\caption{Test for symmetry of the hidden and trait rates via Bayes factors.
-For all models asymmetric hidden transition rates are preferred over models with equal hidden rates.
-Adding more complexity by assuming asymmetry in the trait rate within hidden states does not improve models.}
+\caption{Test for symmetry of the hidden trait transition rates via Bayes factors.
+For all models, asymmetric transition rates between hidden trait states are preferred over models with equal rates.
+Adding more complexity by assuming asymmetry in all rates within both hidden states does not improve models.} %FIXME/CHECK
 \label{supptable:asymmetry}
@@
 
-%B: heads up: I added two clarifying descriptive sentences in the two figures below. Maybe something about noting x-axis scale differences would be useful.
-
 \begin{suppfigure}
 \includegraphics[width=\textwidth]{alldiversificationasymhiddendip.pdf}
-\caption{Posterior distribution for all the best models with diploidization} % XXX
+\caption{Posterior distribution for all the best models with diploidization.} % XXX
 \label{suppfigure:alldip}
@@
 \caption{Lumped ploidy models to assess the effect of adding breeding system to create a three-state model.
-Moderate evidence exists that adding a breeding system is necessary \cref{table:lumped} } % XXX
+Moderate evidence exists that adding a breeding system is necessary \cref{table:lumped}.
+} % XXX
 \label{suppfigure:lumpedDP}
@@
 \caption{Lumped breeding system models to assess the effect of adding ploidy to create a three-state model.
-Moderate to insignificant evidence exists that adding ploidy is necessary \cref{table:lumped} } % XXX
+Moderate to insignificant evidence exists that adding ploidy is necessary \cref{table:lumped}.
+} % XXX
 \label{suppfigure:lumpedIC}
@@
 \includegraphics[width=\textwidth]{effectofasymmetry.pdf}
-\caption{Effect of asymmetric rates in hidden models. Models with asymmetric rates are preferred over models with equal rates \cref{supptable:asymmetry}} % XXX
+\caption{Effect of asymmetric rates in hidden models. Models with asymmetric rates are preferred over models with equal rates \cref{supptable:asymmetry}. } % XXX
 \label{suppfigure:asymmetric}
@@
 \includegraphics[width=\textwidth]{asrDP.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of theM1. D/P ploidy model} % XXX
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the M1, D/P ploidy model. } % XXX
 \label{suppfigure:DPnodipasr}
@@
 \includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M1. D/P model} % XXX
+\caption{Posterior distribution for each of the parameters in the M1, D/P ploidy model.} % XXX
 \label{suppfigure:DPnodip}
@@
 \includegraphics[width=\textwidth]{asrDPAB.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the M4. D/P+A/B asym model} % XXX
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the M4, D/P+A/B asym model.} % XXX
 \label{suppfigure:DPnodipABasr}
@@
 \includegraphics[width=\textwidth]{hisseDPnodipasymposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M4. D/P+A/B asym model} % XXX
+\caption{Posterior distribution for each of the parameters in the M4, D/P+A/B asym model.} % XXX
 \label{suppfigure:DPnodipAB}
@@
 \includegraphics[width=\textwidth]{asrIC.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the M11.I/C model} % XXX
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the M1, I/C breeding system model.} % XXX
 \label{suppfigure:ICasr}
@@
 \includegraphics[width=\textwidth]{bisseSIposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M11. I/C model} % XXX
+\caption{Posterior distribution for each of the parameters in the M11, I/C breeding system model.} % XXX
 \label{suppfigure:IC}
@@
 \includegraphics[width=\textwidth]{asrICAB.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the M14. I/C+A/B asym model} % XXX
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the M14, I/C+A/B asym model.} % XXX
 \label{suppfigure:ICABasr}
@@
 \includegraphics[width=\textwidth]{hisseSIasymposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M14. I/C+A/B asym model} % XXX
+\caption{Posterior distribution for each of the parameters in the M14, I/C+A/B asym model.} % XXX
 \label{suppfigure:ICAB}
@@
 \includegraphics[width=\textwidth]{asrIDCDCP.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M16. ID/CD/CP model} % XXX
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M16, ID/CD/CP model.} % XXX
 \label{suppfigure:IDCDCPnodipasr}
@@
 \includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the  M16. ID/CD/CP model} % XXX
+\caption{Posterior distribution for each of the parameters in the  M16, ID/CD/CP model.} % XXX
 \label{suppfigure:IDCDCPnodip}
@@
 \includegraphics[width=\textwidth]{asrIDCDCPAB.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M19. ID/CD/CP+A/B asym model} % XXX
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M19, ID/CD/CP+A/B asym model.} % XXX
 \label{suppfigure:IDCDCPnodipABasr}
@@
 \includegraphics[width=\textwidth]{muhissenodipasymposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M19. ID/CD/CP+A/B asym model} % XXX
+\caption{Posterior distribution for each of the parameters in the M19, ID/CD/CP+A/B asym model.} % XXX
 \label{suppfigure:IDCDCPnodipAB}

### 2019-07-19 manuscript/results.tex
- Final ms changes pass-through, mainly spaces, commas, spelling, etc. - Leaving FIXME marks for galley proofs. - Will update archive once more with submission version.

--- 2019-07-18 manuscript/results.tex
@@
     Hidden state models with asymmetric rates are, however, strongly preferred over models with equal rates between hidden states (\cref{supptable:asymmetry}), and they show stronger differences between some net diversification rates (\cref{suppfigure:asymmetric}).
+   The effect of the asymmetry of the hidden state transitions is better observed in the ancestral state estimations (\cref{suppfigure:DPnodipABasr,suppfigure:ICABasr,suppfigure:IDCDCPnodipABasr}), which show that moving out of state A (dark colors) happens quickly, with rate $\alpha$,  whereas evolving out of hidden state B (light colors) is slow, with rate $\beta$.
 

### 2019-07-19 manuscript/response.tex
- Final ms changes pass-through, mainly spaces, commas, spelling, etc. - Leaving FIXME marks for galley proofs. - Will update archive once more with submission version.
- tidying up response text on top

--- 2019-07-18 manuscript/response.tex
@@
 
-\hfill \today
+\hfill \today.
 
@@
 
-The first reason is additional work to address a key concern from all referees, about the transition symmetry among the hidden states. 
+First, we performed additional work to address a key concern from all referees about the transition symmetry among the hidden states. 
 Briefly, we found that is relevant to have asymmetric hidden state rates and, in our case, it increased the differences among net diversification rates. 
@@
 
-The second reason is related to a question we mulled over after submitting the first version of this article.
-Specifically, we wondered whether increasing the state space from two- to three-states in our models of the diversification process explains the data better (and how they ought to be best compared).
-This  trait coding question could not be addressed directly by model comparison statistics because the two-state and three-state models need different input (i.e. (0,1) versus (0,1,2)). 
+Second the time since the first version allowed us to mull over an important issue. 
+Specifically, we wondered whether increasing the state space from two-states (single-trait) to three-states (two traits) explains the data better (and how they ought to be best compared). 
+This trait-coding question could not be addressed directly by model comparison, because the two-state and three-state models need different input (i.e. (0,1) versus (0,1,2)). 
 The species and the tree are identical, however, and our intuition indicated that they should carry the same biological evidence despite a particular coding input. 
-During this revision, we figured a way to recode the three-state models to make them equivalent to the two-state models (the new ``lumped" models). 
+During this revision, with the help of Sergei Tarasov's recent paper \citep{tarasov_2019}, we found a way to recode the three-state models to make them equivalent to the two-state models (these are called ``lumped models").
 This equivalency enables comparison of two-state and three-state models, and assessment of their relative explanatory value after adding key trait information and states in diversification analyses.
-We also revised the original dataset (thanks to the suggestion of Reviewer 1) and we were able to correct discrepancies with a previous dataset, as well as add more data (now 651 taxa).
+We also revised the original dataset (thanks to the suggestion of Reviewer 1), we were able to correct discrepancies with a previous dataset, as well as add more data (now 651 taxa).
 
-We hope the changes and the new models are satisfactory to the editorial staff and the reviewers.
+We hope the changes and the new models are satisfactory to all of you.
 
@@
 
-The Authors \\
-(Rosana Zenil-Ferguson et al.)
+Rosana Zenil-Ferguson, Gordon Burleigh, Boris Igi\'c, Itay Mayrose, and Emma Goldberg
 

### 2019-07-19 manuscript/main.tex
REVISION VERSION (JULY 19, 2019) NEW PHYTOLOGIST NPH-MS-2019-29554.R1
- Note: final submission will require all figures to be uploaded separately
- Figures are currently spread across two directories with some stale versions
- Author Contributions is commented out, but will be populated in the next commit
- There are FIXME marks remaining in a few places
- Author contributions copied from Rosana's email and adapted to narrative form

@@
 This work was supported by the National Science Foundation grants DEB-1655478 (to EEG and IM), NSF DEB-1655692 \& NESCent sabbatical scholar award EF-0905606 (to BI), and the Israel Science Foundation 961/17 (to IM).
-%IM: #32
+%IM: #32; ISF funding source not recognized by NP database.
 We thank Dan Schoen and three anonymous reviewers for providing uncommonly thoughtful and helpful reviews, as well as Spencer Barrett for continually inspiring our work.
+
+%--------------------------------------------------
+%\section{Author Contributions}
+%--------------------------------------------------
+
+%B: these are from July 19th, 2019, email from RZF:
+RZF, BI, and EEG designed the research; RZF and EEG performed the research; RZF conducted data analyses; RZF, BI, and EEG collected the data; RZF, BJG, WF, BI, EEG were involved in data interpretation; RZF, BJG, WF, BI, IM, and EEG wrote the manuscript.
 

### 2019-08-20 manuscript/abstract.tex
Summary 200 words fixed, Title page fixed

--- 2019-07-16 manuscript/abstract.tex
@@
-\section{Abstract}
+\section{Summary}
 
-If particular traits consistently affect rates of speciation and extinction, broad macroevolutionary patterns can be understood as consequences of selection at high levels of the biological hierarchy.
-Identifying traits associated with diversification rate differences is complicated by the wide variety of characters under consideration and the statistical challenges of testing for associations from comparative phylogenetic data.
-Ploidy (diploid \vs polyploid states) and breeding system (self-incompatible \vs self-compatible states) have been repeatedly suggested as possible drivers of differential diversification.
-We investigate the connections of these traits, including their interaction, to speciation and extinction rates in Solanaceae.
-We show that the effect of ploidy on diversification can be largely explained by its correlation with breeding system and that additional unknown factors, alongside breeding system, determine diversification rates.
-These results are largely robust to allowing for diploidization.
-Finally, we find that the evolutionary pathway to polyploidy in Solanaceae is most often via the direct breakdown of self-incompatibility by whole genome duplication, rather than indirectly via breakdown followed by polyploidization.
+\begin{itemize}
+\item If particular traits consistently affect rates of speciation and extinction, broad macroevolutionary patterns can be interpreted as consequences of selection at high levels of the biological hierarchy.
+Identifying traits associated with diversification rates is difficult because of the wide variety of characters under consideration and the statistical challenges of testing for associations from comparative phylogenetic data.
+Ploidy (diploid \vs polyploid states) and breeding system (self-incompatible \vs self-compatible states) have been repeatedly suggested as possible drivers of differential diversification in angiosperms.
+
+\item  We investigate the connections of ploidy and breeding system, including their interaction, to speciation and extinction rates in Solanaceae by fitting and comparing twenty-nine diversification models.
+
+\item We show that the effect of ploidy on diversification can be largely explained by its correlation with breeding system and additional unobserved factors changing diversification rates.
+We find that the most common evolutionary pathway to polyploidy in Solanaceae occurs via direct breakdown of self-incompatibility by whole-genome duplication, rather than indirectly via breakdown followed by polyploidization.
+
+\item Finally, we found that lineage diversification in Solanaceae is the result of a complex interaction, amongst ploidy, breeding system, and unobserved factors.
+We show that modeling complex trait interactions is key to our understanding of diversification.
+\end{itemize}

### 2019-08-21 manuscript/methods.tex
figure order fixed

@@
 
-In order to test our hypotheses about lineage diversification and trait macroevolution, we fit 29 state-dependent speciation and extinction models \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016}.
+In order to test our hypotheses about lineage diversification and trait macroevolution, we fit twenty-nine state-dependent speciation and extinction models \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016}.
 SSE models contain parameters that describe per-lineage rates of speciation and extinction, specific to each character state (denoted $\lambda$ and $\mu$, respectively, with subscripts to indicate the state), along with rates of transitions between states (denoted $\rho$ for polyploidization, $\delta$ for diploidization, and $q_{IC}$ for loss of self-incompatibility).
@@
 These models differ in whether transitions between the hidden states are symmetric (M3) or asymmetric(M4), and whether the polyploidization rate depends on the hidden state(M5). %WF suggestion
-Comparing M1--M5 allows us to test whether ploidy is associated with diversification differences on top of the differences potentially explained by an unknown factor.
+Comparing M1--M5 (as shown in \cref{table:bayesfactors}) allows us to test whether ploidy is associated with diversification differences on top of the differences potentially explained by an unknown factor.
 We further fit the analogues of these five models but including a rate parameter $\delta$ for transitions from polyploid to diploid (M6--M10).
@@
 To answer this question with statistical model comparisons requires comparing the likelihood of the data given each model.
-This is impossible for the ploidy and breeding system models presented so far, however, because the data are different for the different models: they use either the $D/P$ or the $I/C$ or the $ID/CD/CP$ state spaces.
+This is impossible for the ploidy and breeding system models presented so far, however, because the data are different for the different models: they use either the $D/P$ or the $I/C$ or the $ID/CD/CP$ state spaces (see figure \cref{figure:stateclassifications} for state classifications).
 Therefore, the use of different data results in incomparable models.
@@
 However, to fully mimic the breeding system model, we assume that the rate of selfing is equal (i.e.
-$q_0=q_{IC}=rho_I$) and the rates of speciation and extinction for both $CD$ and $CP$ are the same (new parameters $\lambda_C$ and $\mu_C$).
+$q_0=q_{IC}=\rho_I$) and the rates of speciation and extinction for both $CD$ and $CP$ are the same (new parameters $\lambda_C$ and $\mu_C$).
 We further add a hidden character to each of these models (M27 and M29), and then compare this group of models (\cref{table:lumped}).
@@
 Parameters for each of the 29 models were coded as graphical models and Bayesian statistical inference was performed with RevBayes \citep{hoehna_2016}.
-Scripts for analyses and key results are available as Supplementary Information. 
+Scripts for analyses and key results are available in Dryad. %TODO
 We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $651/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
@@
 
-We report posterior distributions for the model parameters in \cref{figure:netdivall}, and also for the compound parameters of net diversification ($\lambda - \mu$) and extinction fraction ($\mu / \lambda$) for each state.
+We report posterior distributions for the net diversification parameters ($\lambda - \mu$)  in \cref{figure:netdivall} and full results of all parameters of the diversification models in \cref{suppfigure:DPnodip,suppfigure:DPnodipAB,suppfigure:IC,suppfigure:ICAB,suppfigure:IDCDCPnodip,suppfigure:IDCDCPnodipAB}.
 Additionally, ancestral states at each node in the phylogeny were sampled jointly during the MCMC analyses every 100 generations. 
@@
 \item Are hidden states necessary to explain diversification rate heterogeneity?---Comparison of simple models against hidden state models (\cref{supptable:testaddhidden}).
-\item Does a second focal trait add information about the diversification process?---Comparison of lumped models against IC/CD/CP models (\cref{figure:lumped,table:lumped}).
-\item Are conclusions robust to assumptions about hidden state transitions?---Comparison amongst hidden states models with equal hidden rates and asymmetrical rates (\cref{suppfigure:asymmetric,supptable:asymmetry}).
-\item Is there evidence for diploidization?---Comparison amongst models with and without diploidization (\cref{suppfigure:alldip,supptable:testdiploidization}).
+\item Does a second focal trait add information about the diversification process?---Comparison of lumped models against IC/CD/CP models (\cref{figure:lumped,table:lumped,suppfigure:lumpedDP,suppfigure:lumpedIC}).
+\item Are conclusions robust to assumptions about hidden state transitions?---Comparison amongst hidden states models with equal hidden rates and asymmetrical rates (\cref{supptable:asymmetry,suppfigure:asymmetric}).
+\item Is there evidence for diploidization?---Comparison amongst models with and without diploidization (\cref{supptable:testdiploidization,suppfigure:alldip}).
 \end{enumerate}
@@
 \begin{supptable}
+\addtolength{\tabcolsep}{-5pt}
+\begin{tabular}{|l|c|c|c|l|}
+\hline
+Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
+M3. D/P+A/B                                 & -1234.52                     &    M3 vs. M4                 & 11.239                  & M4 (Strong)   \\
+\textbf{M4. D/P+ A/B asym}            & -1223.28                     &   M4 vs. M5              & -1.658                  & M4 (Moderate)   \\
+M5. D/P+A/B all asym                 & -1224.93                     &                     &                         &                         \\ \hline
+M13. I/C+ A/B                                & -1270.47                     &   M13 vs. M14                  & 22.75                   & M14 (Strong)   \\
+\textbf{M14. I/C+ A/B asym}            & \textbf{-1247.72}            & M14 vs. M15         & {0.05}           & No evidence             \\
+M15. I/C+ A/B all asym               & -1247.67                     &                     &                         &                         \\ \hline
+M18. IC/CD/CP+A/B                            & -1414.00                     &     M18 vs. M19                & 20.79                   & M19 (Strong)    \\
+\textbf{M19. IC/CD/CP+ A/B asym}       & \textbf{-1393.21}            & M19 vs. M20          & {-1.919}         & M19 (Moderate)   \\
+M20. IC/CD/CP+ A/B all asym           & -1395.129                    &                     &                         &                         \\ \hline
+M8. D/P +A/B+$\delta$                          & -1225.05                     &  M8 vs. M9                   & 10.58                   & M9 (Strong)    \\
+\textbf{M9. D/P+ A/B+ $\delta$ asym}      & \textbf{-1214.46}                     &    M9 vs. M10                 & -0.52                   & M10 (Moderate) \\
+M10. D/P+A/B+$\delta$ all asym        & -1214.98                     &                    &                         &                         \\ \hline
+M23. IC/CD/DP+A/B+$\delta$                 & -1408.65                     &     M23 vs M24                & 21.91                   & M24(Strong)     \\
+\textbf{M24. IC/CD/DP+A/B+$\delta$ asym} & \textbf{-1386.74}                     &   M24 vs M25              & -2.78                   & M24 (Moderate) \\
+M25. IC/CD/DP+A/B+ $\delta$ all asym     & -1389.52                     &                     &                         &                         \\ \hline
+\end{tabular}
+\caption{Test for symmetry of the hidden trait transition rates via Bayes factors.
+For all models, asymmetric transition rates between hidden trait states are preferred over models with equal rates.
+Adding more complexity by assuming asymmetry in all rates within both hidden states does not improve models.} %FIXME/CHECK
+\label{supptable:asymmetry}
+\end{supptable}
+
+\begin{supptable}
 \addtolength{\tabcolsep}{-3pt}
@@
 
-\begin{supptable}
-\addtolength{\tabcolsep}{-5pt}
-\begin{tabular}{|l|c|c|c|l|}
-\hline
-Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
-M3. D/P+A/B                                 & -1234.52                     &    M3 vs. M4                 & 11.239                  & M4 (Strong)   \\
-\textbf{M4. D/P+ A/B asym}            & -1223.28                     &   M4 vs. M5              & -1.658                  & M4 (Moderate)   \\
-M5. D/P+A/B all asym                 & -1224.93                     &                     &                         &                         \\ \hline
-M13. I/C+ A/B                                & -1270.47                     &   M13 vs. M14                  & 22.75                   & M14 (Strong)   \\
-\textbf{M14. I/C+ A/B asym}            & \textbf{-1247.72}            & M14 vs. M15         & {0.05}           & No evidence             \\
-M15. I/C+ A/B all asym               & -1247.67                     &                     &                         &                         \\ \hline
-M18. IC/CD/CP+A/B                            & -1414.00                     &     M18 vs. M19                & 20.79                   & M19 (Strong)    \\
-\textbf{M19. IC/CD/CP+ A/B asym}       & \textbf{-1393.21}            & M19 vs. M20          & {-1.919}         & M19 (Moderate)   \\
-M20. IC/CD/CP+ A/B all asym           & -1395.129                    &                     &                         &                         \\ \hline
-M8. D/P +A/B+$\delta$                          & -1225.05                     &  M8 vs. M9                   & 10.58                   & M9 (Strong)    \\
-\textbf{M9. D/P+ A/B+ $\delta$ asym}      & \textbf{-1214.46}                     &    M9 vs. M10                 & -0.52                   & M10 (Moderate) \\
-M10. D/P+A/B+$\delta$ all asym        & -1214.98                     &                    &                         &                         \\ \hline
-M23. IC/CD/DP+A/B+$\delta$                 & -1408.65                     &     M23 vs M24                & 21.91                   & M24(Strong)     \\
-\textbf{M24. IC/CD/DP+A/B+$\delta$ asym} & \textbf{-1386.74}                     &   M24 vs M25              & -2.78                   & M24 (Moderate) \\
-M25. IC/CD/DP+A/B+ $\delta$ all asym     & -1389.52                     &                     &                         &                         \\ \hline
-\end{tabular}
-\caption{Test for symmetry of the hidden trait transition rates via Bayes factors.
-For all models, asymmetric transition rates between hidden trait states are preferred over models with equal rates.
-Adding more complexity by assuming asymmetry in all rates within both hidden states does not improve models.} %FIXME/CHECK
-\label{supptable:asymmetry}
-\end{supptable}
-
 % check values of delta and rho D/P+A/B with(out) delta
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{alldiversificationasymhiddendip.pdf}
-\caption{Posterior distribution for all the best models with diploidization.} % XXX
-\label{suppfigure:alldip}
+\includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the M1, D/P ploidy model.} % XXX
+\label{suppfigure:DPnodip}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{hisseDPnodipasymposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the M4, D/P+A/B asym model.} % XXX
+\label{suppfigure:DPnodipAB}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{bisseSIposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the M11, I/C breeding system model.} % XXX
+\label{suppfigure:IC}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{hisseSIasymposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the M14, I/C+A/B asym model.} % XXX
+\label{suppfigure:ICAB}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the  M16, ID/CD/CP model.} % XXX
+\label{suppfigure:IDCDCPnodip}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{muhissenodipasymposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the M19, ID/CD/CP+A/B asym model.} % XXX
+\label{suppfigure:IDCDCPnodipAB}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrDP.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the M1, D/P ploidy model. } % XXX
+\label{suppfigure:DPnodipasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrDPAB.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the M4, D/P+A/B asym model.} % XXX
+\label{suppfigure:DPnodipABasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIC.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the M1, I/C breeding system model.} % XXX
+\label{suppfigure:ICasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrICAB.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the M14, I/C+A/B asym model.} % XXX
+\label{suppfigure:ICABasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIDCDCP.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M16, ID/CD/CP model.} % XXX
+\label{suppfigure:IDCDCPnodipasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIDCDCPAB.pdf}
+\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M19, ID/CD/CP+A/B asym model.} % XXX
+\label{suppfigure:IDCDCPnodipABasr}
 \end{suppfigure}
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDP.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the M1, D/P ploidy model. } % XXX
-\label{suppfigure:DPnodipasr}
+\includegraphics[width=\textwidth]{alldiversificationasymhiddendip.pdf}
+\caption{Posterior distribution for all the best models with diploidization.} % XXX
+\label{suppfigure:alldip}
 \end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M1, D/P ploidy model.} % XXX
-\label{suppfigure:DPnodip}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDPAB.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the M4, D/P+A/B asym model.} % XXX
-\label{suppfigure:DPnodipABasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{hisseDPnodipasymposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M4, D/P+A/B asym model.} % XXX
-\label{suppfigure:DPnodipAB}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIC.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the M1, I/C breeding system model.} % XXX
-\label{suppfigure:ICasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{bisseSIposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M11, I/C breeding system model.} % XXX
-\label{suppfigure:IC}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrICAB.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the M14, I/C+A/B asym model.} % XXX
-\label{suppfigure:ICABasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{hisseSIasymposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M14, I/C+A/B asym model.} % XXX
-\label{suppfigure:ICAB}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCP.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M16, ID/CD/CP model.} % XXX
-\label{suppfigure:IDCDCPnodipasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the  M16, ID/CD/CP model.} % XXX
-\label{suppfigure:IDCDCPnodip}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCPAB.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M19, ID/CD/CP+A/B asym model.} % XXX
-\label{suppfigure:IDCDCPnodipABasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{muhissenodipasymposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M19, ID/CD/CP+A/B asym model.} % XXX
-\label{suppfigure:IDCDCPnodipAB}
-\end{suppfigure}

### 2019-08-22 manuscript/abstract.tex
- edited and commented on Abstract

@@
 Identifying traits associated with diversification rates is difficult because of the wide variety of characters under consideration and the statistical challenges of testing for associations from comparative phylogenetic data.
-Ploidy (diploid \vs polyploid states) and breeding system (self-incompatible \vs self-compatible states) have been repeatedly suggested as possible drivers of differential diversification in angiosperms.
+Ploidy (diploid \vs polyploid states) and breeding system (self-incompatible \vs self-compatible states) are both thought to be drivers of differential diversification in angiosperms.
+%B: rm: have been repeatedly suggested as possible
 
-\item  We investigate the connections of ploidy and breeding system, including their interaction, to speciation and extinction rates in Solanaceae by fitting and comparing twenty-nine diversification models.
+\item  We investigate the effect of ploidy and breeding system, including their interaction, on speciation and extinction rates in Solanaceae by fitting and comparing twenty-nine diversification models.
+%B: rm: connection ...
+including interaction to speciation
 
 \item We show that the effect of ploidy on diversification can be largely explained by its correlation with breeding system and additional unobserved factors changing diversification rates.
-We find that the most common evolutionary pathway to polyploidy in Solanaceae occurs via direct breakdown of self-incompatibility by whole-genome duplication, rather than indirectly via breakdown followed by polyploidization.
+%B: Why not just directly state here that the effect seems to be primarily explained by breeding system?
+(unaltered)
+These results are largely robust to allowing for diploidization.  %B: this sentence was apparently accidentally dropped?
+We also find that the most common evolutionary pathway to polyploidy in Solanaceae occurs via direct breakdown of self-incompatibility by whole-genome duplication, rather than indirectly via breakdown followed by polyploidization.
 
-\item Finally, we found that lineage diversification in Solanaceae is the result of a complex interaction, amongst ploidy, breeding system, and unobserved factors.
+\item  Lineage diversification in Solanaceae appears to be the result of a complex interaction, among ploidy, breeding system, and unobserved factors. 
 We show that modeling complex trait interactions is key to our understanding of diversification.
+%B was: Our models of trait interactions alongside unobserved factors may enhance our understanding of diversification.
 \end{itemize}

### 2019-08-22 manuscript/methods.tex
references Rosana to be merged with boris and partial figure editing

@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{states.pdf}
-\caption{
-Character states used in the models.
-(A) Each species retained on the phylogeny belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known. 
-Number of species in each category is indicated; for example, 70 species are self-compatible with unknown ploidy.
-Character state abbreviations are: $I$ for self-incompatible, $C$ for self-compatible, $D$ for diploid, $P$ for polyploid, $?$ for unknown.
-Because polyploidization breaks this form of self-incompatibility, self-incompatible species with unobserved ploidy ($I?$) are taken to be diploid ($ID$), and polyploid species with unobserved breeding system ($?P$) are taken to be SC ($CP$).
-(B) Category groupings into states for each model class.
-In the ploidy-only models (M1- M10), states are coded as $D$ \& $P$ when uncertain/consistent with either state; in the breeding system-only models (M11-M15) such states are coded as $C$; in the ploidy and breeding system models (M16-M29), they are coded as $CD$ \& $CP$.
-In the hidden-trait models, all species could take on either of two `hidden' character states. %IM: #33 
-Two species, \emph{Lycium californicum} and \emph{Solanum bulbocastanum}, are simultaneously $ID$ and $CP$, and by adding them the sample adds to the total of 651 taxa used for analyses.
-}
-\label{figure:stateclassifications}
-\end{figure}
-
-\begin{suppfigure}
-    \caption{ All models.  [This figure is provided as a separate, large-format page.] }
-    \label{fig:allmodels}
-\end{suppfigure}
-
 %%% Below here is old tables.  I think the new versions should instead go in results.tex or a separate file. %%%
@@
 \end{tabular}
-\caption{Bayes factors of  ploidy-only models without diplodization.
+\caption{Bayes factors for ploidy only models without diplodization.
 Numbers greater that 1 mean moderate support relative to the model in the row label.
 Conventional threshold for `strong' support is 10.
-The best models are written in bold.
+The preferred models are M2 and M4 (bold).
 }
@@
 \end{tabular}
-\caption{Bayes factors of breeding system-only models. Results indicate that the HiSSE models with asymmetric rates (M14, M15) are strongly preferred over every other model.}
+\caption{Bayes factors for breeding system only models. Results indicate that the HiSSE models (bold) with asymmetric rates (M14, M15) are strongly preferred over every other model.}
 \label{table:M11M15}
@@
 M16 & -0.69           & No evidence                                                          \\
-\textbf{M16.
-ID/CD/CP}      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
+M16. ID/CD/CP      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
 M29.
@@
 \end{tabular}
-\caption{Test for adding a focal trait to a binary state model via Bayes factors.
+\caption{Testing the addition of a focal trait to a binary state model via Bayes factors.
 Three-state models are moderately preferred over two-state models.
-The evidence is stronger towards the inclusion of breeding systems in ploidy models (M26, M27) than the inclusion of ploidy in breeding system models (M29, M29)}
+There is moderate evidence towards the inclusion of breeding systems in ploidy models (M16, M23, bold).
+There is moderate or no evidence towards the  inclusion of ploidy in breeding system models (M23, bold).}
 \label{table:lumped}
@@
 \begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{states.pdf}
+\caption{
+Character states used in the models.
+(A) Each species retained on the phylogeny belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known. 
+Number of species in each category is indicated; for example, 70 species are self-compatible with unknown ploidy.
+Character state abbreviations are: $I$ for self-incompatible, $C$ for self-compatible, $D$ for diploid, $P$ for polyploid, $?$ for unknown.
+Because polyploidization breaks this form of self-incompatibility, self-incompatible species with unobserved ploidy ($I?$) are taken to be diploid ($ID$), and polyploid species with unobserved breeding system ($?P$) are taken to be SC ($CP$).
+(B) Category groupings into states for each model class.
+In the ploidy-only models (M1- M10), states are coded as $D$ \& $P$ when uncertain/consistent with either state; in the breeding system-only models (M11-M15) such states are coded as $C$; in the ploidy and breeding system models (M16-M29), they are coded as $CD$ \& $CP$.
+In the hidden-trait models, all species could take on either of two `hidden' character states. %IM: #33 
+Two species, \emph{Lycium californicum} and \emph{Solanum bulbocastanum}, are simultaneously $ID$ and $CP$, and by adding them the sample adds to the total of 651 taxa used for analyses.
+}\label{figure:stateclassifications}
+\end{figure}
+
+\begin{figure}
 \includegraphics[width=\textwidth]{alldiversificationasymhidden.pdf} 
 \caption{Net diversification rates for SSE models of focal traits with or without hidden state. 
-Each panel contains a graphical summary of estimated model parameters and displays posterior distributions for net diversification estimates.}
+(A) Ploidy only model (M1) showing higher net diversification linked to diploid state $D$ compared to polyploid state $P$.
+(B) Ploidy only model with hidden states  model (M4) showing that the net diversification is higher for hidden state $B$ compared to hidden state $A$ and both diploid and polyploid states within each hidden traits have overlapping net diversification rates.
+(C) Breeding system model only (M9) showing higher net diversification linked to self-incompatible $I$ state compared to self-compatible state $C$.
+(D) Breeding system with hidden states model (M11) showing differences in both hidden states in net diversification and little to no overlapping in between self-compatible \vs self-compatible states.
+(E) Ploidy and breeding system model (M13) showing higher net diversification linked to self-incompatible diploid state $ID$ compared to both self-compatible states despite ploidy level ($CD, CP$).
+(F) Ploidy, breeding system, and hidden states modl (M16) showing a similar pattern that panel (E) within each hidden state $A$ and $B$.
+For hidden state $B$ there is a larger overlap of net diversification between states $CD_B$ and $ID_B$.}
 \label{figure:netdivall}
@@
 \begin{suppfigure}
+    \caption{ All models.  [This figure is provided as a separate, large-format page.] }
+    \label{fig:allmodels}
+\end{suppfigure}
+
+\begin{suppfigure}
 \includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}

### 2019-08-22 manuscript/abstract.tex
- a few tiny changes to Abstract (errant comma and plural)

@@
 
+% Right now at exactly 200 words!
 \begin{itemize}
@@
 
-\item  We investigate the effect of ploidy and breeding system, including their interaction, on speciation and extinction rates in Solanaceae by fitting and comparing twenty-nine diversification models.
-%B: rm: connection ...
+\item  We investigate the effect of ploidy and breeding system, including their interaction, and unobserved traits, on speciation and extinction rates in Solanaceae by fitting and comparing twenty-nine diversification models.
+%B: add unobserved traits rm: connection ...
 including interaction to speciation
@@
 (unaltered)
-These results are largely robust to allowing for diploidization.  %B: this sentence was apparently accidentally dropped?
+%These results are largely robust to allowing for diploidization.  %B: I now dropped it again to squeeze under 200 words. %B: this sentence was apparently accidentally dropped?
 We also find that the most common evolutionary pathway to polyploidy in Solanaceae occurs via direct breakdown of self-incompatibility by whole-genome duplication, rather than indirectly via breakdown followed by polyploidization.
 
-\item  Lineage diversification in Solanaceae appears to be the result of a complex interaction, among ploidy, breeding system, and unobserved factors. 
-We show that modeling complex trait interactions is key to our understanding of diversification.
-%B was: Our models of trait interactions alongside unobserved factors may enhance our understanding of diversification.
+\item  Lineage diversification in Solanaceae appears to be the result of complex interactions among breeding system, ploidy, and unobserved factors. 
+Examination of models that include trait interactions alongside `hidden' traits may generally enhance our understanding of diversification.
+%B was: We show that modeling complex trait interactions is key to our understanding of diversification.
 \end{itemize}

### 2019-08-22 manuscript/abstract.tex
edited summary (Boris please check)

@@
 
-\item We show that the effect of ploidy on diversification can be largely explained by its correlation with breeding system and additional unobserved factors changing diversification rates.
+\item We show that the diversification patterns in Solanaceae can largely explained by changes in breeding system and additional unobserved factors linked to speciation and rates, even under the presence of ploidy information.
 %B: Why not just directly state here that the effect seems to be primarily explained by breeding system?
 (unaltered)
-%These results are largely robust to allowing for diploidization.  %B: I now dropped it again to squeeze under 200 words. %B: this sentence was apparently accidentally dropped?
+%These results are largely robust to allowing for diploidization.
+%B: I now dropped it again to squeeze under 200 words.
+%B: this sentence was apparently accidentally dropped?
+% I removed the diploidization because the story was complicated so I preferred not to add it in the abstract
+%R- Is the above sentence better now?
+
 We also find that the most common evolutionary pathway to polyploidy in Solanaceae occurs via direct breakdown of self-incompatibility by whole-genome duplication, rather than indirectly via breakdown followed by polyploidization.
 
-\item  Lineage diversification in Solanaceae appears to be the result of complex interactions among breeding system, ploidy, and unobserved factors. 
-Examination of models that include trait interactions alongside `hidden' traits may generally enhance our understanding of diversification.
+\item Comparing multiple stochastic diversification models that include complex trait interactions alongside hidden states enhances our understanding of the macroevolutionary patterns in plant phylogenies.
 %B was: We show that modeling complex trait interactions is key to our understanding of diversification.

### 2019-08-24 manuscript/abstract.tex
edits to Summary

@@
 
-% Right now at exactly 200 words!
+% max 200 words (188 now)
 \begin{itemize}
+
+% Rationale
 \item If particular traits consistently affect rates of speciation and extinction, broad macroevolutionary patterns can be interpreted as consequences of selection at high levels of the biological hierarchy.
@@
 Ploidy (diploid \vs polyploid states) and breeding system (self-incompatible \vs self-compatible states) are both thought to be drivers of differential diversification in angiosperms.
-%B: rm: have been repeatedly suggested as possible
 
-\item  We investigate the effect of ploidy and breeding system, including their interaction, and unobserved traits, on speciation and extinction rates in Solanaceae by fitting and comparing twenty-nine diversification models.
-%B: add unobserved traits rm: connection ...
-including interaction to speciation
+% Methods
+\item To extensive trait and phylogenetic data in Solanaceae, we fit twenty-nine diversification models to investigate how speciation and extinction rate differences are associated with ploidy, breeding system, and the interaction between these traits.
+% \item  We investigate the effect of ploidy and breeding system, including their interaction, and unobserved traits, on speciation and extinction rates in Solanaceae by fitting and comparing twenty-nine diversification models.
 
-\item We show that the diversification patterns in Solanaceae can largely explained by changes in breeding system and additional unobserved factors linked to speciation and rates, even under the presence of ploidy information.
-%B: Why not just directly state here that the effect seems to be primarily explained by breeding system?
-(unaltered)
+% Results
+\item We show that diversification patterns in Solanaceae are better explained by breeding system and an additional unobserved factor, rather than by ploidy.
+% E: changed because we don't show that breeding system and one hidden trait are sufficient
 %These results are largely robust to allowing for diploidization.
@@
 % I removed the diploidization because the story was complicated so I preferred not to add it in the abstract
-%R- Is the above sentence better now?
-
 We also find that the most common evolutionary pathway to polyploidy in Solanaceae occurs via direct breakdown of self-incompatibility by whole-genome duplication, rather than indirectly via breakdown followed by polyploidization.
 
+% Conclusions
 \item Comparing multiple stochastic diversification models that include complex trait interactions alongside hidden states enhances our understanding of the macroevolutionary patterns in plant phylogenies.
 %B was: We show that modeling complex trait interactions is key to our understanding of diversification.
+
 \end{itemize}

### 2019-08-24 manuscript/introduction.tex
what a nice Intro!

@@
 \begin{quote}
-\em{``Among life history traits, reproductive characters that determine mating patterns are perhaps the most influential in governing macroevolution." 
+\em{``Among life history traits, reproductive characters that determine mating patterns are perhaps the most influential in governing macroevolution.'' 
 }
 \\
-\hspace*{\fill}\rm{ Barrett et al. (1996)}
+\hspace*{\fill}\rm{Barrett et al.~(1996)}
 \end{quote}
@@
 An increase in ploidy can alter many traits and affect a variety of evolutionary and ecological processes \citep{ramsey_2002, sessa_2019}.
-At shallow evolutionary time scale, polyploids were found to have an overall lower net diversification rate than diploids across many vascular plant clades \citep{mayrose_2011, mayrose_2015}.
-%IM suggestion
+At shallow evolutionary time scales, polyploids were found to have an overall lower net diversification rate than diploids across many vascular plant clades \citep{mayrose_2011, mayrose_2015}.
 However, recent genomic studies have inferred numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications may have played an important  role driving innovation and diversification in plants.
@@
 Polyploidy and self-fertilization are widely thought to be associated \citep{stebbins1950}.
-Whole genome duplication may facilitate the transition to selfing by masking inbreeding depression, or self-fertilization may facilitate establishment of early-forming polyploids by avoiding the lower fitness \citep{levin_1975,ramsey_1998, barringer2007, barrett2008, husband2008}.%IM suggestion and Gordon suggestion %B: reworded because polyploidy is a process, so its establishment is not the issue.
+Whole genome duplication may facilitate the transition to selfing by masking inbreeding depression, or self-fertilization may facilitate establishment of polyploids by avoiding the lower fitness of triploids \citep{levin_1975,ramsey_1998, barringer2007, barrett2008, husband2008}.
+% or both selfing and polyploidy may tend to be found in weedy or peripheral populations \citep{ramsey_1998, barringer2007, barrett2008, husband2008}.
 Additionally, in RNase-based gametophytic SI systems, polyploidization directly causes the loss of SI \citep{stout1942, lewis1947}.
@@
 Strictly speaking, this is a feature of non-self recognition systems of the kind found in RNase-based GSI.
-In these systems, SI occurs because haploid self-pollen grains, with one S-allele at the locus controlling the SI response, are unable to detoxify the S-RNase produced by the same S-allele in the style \citep{kubo2010}.
-The unreduced pollen of diploids, however, can contain two S-alleles expressed in pollen, which jointly gain the ability to detoxify the S-RNases produced by any maternal genotype \citep{entani1999, tsukamoto2005, kubo2010}.
+In these systems, SI occurs because haploid self pollen grains, with one S-allele at the locus controlling the SI response, are unable to detoxify the S-RNase produced by the same S-allele in the style \citep{kubo2010}.
+The unreduced pollen of diploids, however, can contain two S-alleles expressed in pollen, which jointly provide the ability to detoxify the S-RNases produced by any maternal genotype \citep{entani1999, tsukamoto2005, kubo2010}.
 Initial mutant individuals with pollen containing doubled haploid genomes are consequently capable of self-fertilization, with exceedingly few exceptions \citep{hauck_2002, nunes_2006}.
@@
 We follow by fitting models that incorporate hidden states, which reduce the chance that the effect of the focal trait (ploidy or breeding system) on diversification is found to be significant when in reality, it may simply represent background heterogeneity in the diversification process that produces the phylogenetic patterns \citep{beaulieu_2016}.
-%IM: added a little more explanation
-We compare the proposed models against their character-independent counterparts \citep{beaulieu_2016} to investigate whether ploidy or breeding system affect the diversification process.
+We compare the proposed models against their character-independent counterparts \citep{beaulieu_2016} to investigate whether something other than ploidy or breeding system better explains the diversification process.
 Next, we model ploidy and breeding system jointly to assess their combined influence on diversification, with or without an additional hidden character.
-%B: additional state implies that we would introduce another state, say 'C' to 'A' and 'B'.
 Using the ploidy and breeding system model without hidden states, we quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
-We also aggregate states within these joint models of ploidy and breeding system in order to test whether increasing complexity from one trait to two traits significantly improves our understanding of the diversification process.
+We also aggregate states within these joint models of ploidy and breeding system in order to test whether increasing model complexity from one trait to two traits significantly improves our understanding of the diversification process.
 Furthermore, we extend all the models involving ploidy to investigate the potential effects of including diploidization.
 % And finally, we explore models of diversification using asymmetric rates between and within hidden states to investigate the sensitivity of the statistical inference to model assumptions.
-Our results highlight the importance of considering non-additive effects of traits on net diversification rates  under the presence of unobserved factors, in order to detect strong biologically-driven processes dictating the diversification patterns.
+Our results highlight the importance of considering non-additive effects of traits on net diversification rates  under the presence of unobserved factors, in order to detect strong biologically-driven processes dictating diversification patterns.

### 2019-08-24 manuscript/abstract.tex
- Removed old comments and rearranged one sentence of Summary

@@
 
-% max 200 words (188 now)
+% presently 187/200 words
 \begin{itemize}
@@
 % Methods
-\item To extensive trait and phylogenetic data in Solanaceae, we fit twenty-nine diversification models to investigate how speciation and extinction rate differences are associated with ploidy, breeding system, and the interaction between these traits.
-% \item  We investigate the effect of ploidy and breeding system, including their interaction, and unobserved traits, on speciation and extinction rates in Solanaceae by fitting and comparing twenty-nine diversification models.
+\item  We fit twenty-nine diversification models to extensive trait and phylogenetic data in Solanaceae and investigate how speciation and extinction rate differences are associated with ploidy, breeding system, and the interaction between these traits.%B: I re-ordered this and removed the second "to": To extensive trait and phylogenetic data in Solanaceae, we fit twenty-nine diversification models to investigate how speciation and extinction rate differences are associated with ploidy, breeding system, and the interaction between these traits.
 
@@
 \item We show that diversification patterns in Solanaceae are better explained by breeding system and an additional unobserved factor, rather than by ploidy. 
-% E: changed because we don't show that breeding system and one hidden trait are sufficient
-%These results are largely robust to allowing for diploidization.
-%B: I now dropped it again to squeeze under 200 words.
-%B: this sentence was apparently accidentally dropped?
-% I removed the diploidization because the story was complicated so I preferred not to add it in the abstract
 We also find that the most common evolutionary pathway to polyploidy in Solanaceae occurs via direct breakdown of self-incompatibility by whole-genome duplication, rather than indirectly via breakdown followed by polyploidization.
@@
 \item Comparing multiple stochastic diversification models that include complex trait interactions alongside hidden states enhances our understanding of the macroevolutionary patterns in plant phylogenies.
-%B was: We show that modeling complex trait interactions is key to our understanding of diversification.
 

### 2019-08-30 manuscript/methods.tex
All legends fixed

--- 2019-08-21 manuscript/methods.tex
@@
 \item Are conclusions robust to assumptions about hidden state transitions?---Comparison amongst hidden states models with equal hidden rates and asymmetrical rates (\cref{supptable:asymmetry,suppfigure:asymmetric}).
-\item Is there evidence for diploidization?---Comparison amongst models with and without diploidization (\cref{supptable:testdiploidization,suppfigure:alldip}).
+\item Is there evidence for diploidization?---Comparison amongst log-scalemodels with and without diploidization (\cref{supptable:testdiploidization,suppfigure:alldip}).
 \end{enumerate}
 
-Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $K=ln(BF(M_1,\ M_0)) = \ln[ P(\mathbf{X} | M_1) - P(\mathbf{X} | M_0)]$.
+Each model comparison is reported with a Bayes factor on the natural log-scale: the comparison between models $M_0$ and $M_1$ is $K=ln(BF(M_1,\ M_0)) = \ln[ P(\mathbf{X} | M_1) - P(\mathbf{X} | M_0)]$.
 There is `strong evidence' for $M_1$ when this value is more than 10, moderate support if the value is more than 1, and no evidence if the value is between -1 and 1.
 If the value of $K$ is negative the evidence goes towards $M_0$ \citep{kass1995}.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{states.pdf}
-\caption{
-Character states used in the models.
-(A) Each species retained on the phylogeny belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known. 
-Number of species in each category is indicated; for example, 70 species are self-compatible with unknown ploidy.
-Character state abbreviations are: $I$ for self-incompatible, $C$ for self-compatible, $D$ for diploid, $P$ for polyploid, $?$ for unknown.
-Because polyploidization breaks this form of self-incompatibility, self-incompatible species with unobserved ploidy ($I?$) are taken to be diploid ($ID$), and polyploid species with unobserved breeding system ($?P$) are taken to be SC ($CP$).
-(B) Category groupings into states for each model class.
-In the ploidy-only models (M1- M10), states are coded as $D$ \& $P$ when uncertain/consistent with either state; in the breeding system-only models (M11-M15) such states are coded as $C$; in the ploidy and breeding system models (M16-M29), they are coded as $CD$ \& $CP$.
-In the hidden-trait models, all species could take on either of two `hidden' character states. %IM: #33 
-Two species, \emph{Lycium californicum} and \emph{Solanum bulbocastanum}, are simultaneously $ID$ and $CP$, and by adding them the sample adds to the total of 651 taxa used for analyses.
-}
-\label{figure:stateclassifications}
-\end{figure}
-
-\begin{suppfigure}
-    \caption{ All models.  [This figure is provided as a separate, large-format page.] }
-    \label{fig:allmodels}
-\end{suppfigure}
 
@@
 \textbf{M2.
-CID D/P} & \textbf{-1223.86}                                                 &       & -10.66 & 0.579 & -1.079 & \begin{tabular}[c]{@{}l@{}}Strong preference for \\ M2 over M1\end{tabular}              \\ \hline
+CID D/P} & \textbf{-1223.86}                                                 &       & -10.66 & 0.579 & -1.079 & \begin{tabular}[c]{@{}l@{}}Strong preference for \\ M2 over M1 and M3\end{tabular}              \\ \hline
 M3.
@@
 \end{tabular}
-\caption{Bayes factors of  ploidy-only models without diplodization.
-Numbers greater that 1 mean moderate support relative to the model in the row label.
-Conventional threshold for `strong' support is 10.
-The best models are written in bold.
-}
+\caption{Bayes factors for ploidy only models without diplodization in log-scale.
+Numbers smaller than -1 indicate moderate preference for the model listed in the row, and values larger than 1 indicate moderate preference for the model listed in the respective column .
+Conventional threshold for `strong' preference is an absolute value larger than 10.
+The character independent model (M2, bold) is strongly preferred over the ploidy only model (M1) or a hidden state model with symmetric rates (M3).
+Moderate support exists for the model with ploidy and asymmetric hidden states (M4, bold) over the rest of the ploidy only models.}
 \label{table:bayesfactors}
@@
 \end{tabular}
-\caption{Bayes factors of breeding system-only models. Results indicate that the HiSSE models with asymmetric rates (M14, M15) are strongly preferred over every other model.}
+\caption{Bayes factors for breeding system only models in log-scale.
+Numbers smaller than -1 indicate moderate preference for the model listed on the row, and values larger than 1 indicate moderate preference for the model in the respective column.
+Conventional threshold for `strong' preference is an absolute value larger than 10.
+Moderate to strong preference exist for models M14 and M15 which include a hidden states with asymmetrical rates.}
 \label{table:M11M15}
@@
 \end{tabular}
-\caption{Test for adding a focal trait to a binary state model via Bayes factors.
-Three-state models are moderately preferred over two-state models.
-The evidence is stronger towards the inclusion of breeding systems in ploidy models (M26, M27) than the inclusion of ploidy in breeding system models (M29, M29)}
+\caption{Testing the addition of a focal trait to a binary state model via Bayes factors (log-scale).
+Numbers smaller than -1 indicate moderate preference for the model in the row, and values larger than 1 indicate moderate preference for the model in the respective column.
+Conventional threshold for `strong' preference is an absolute value larger than 10.
+Three-state models (M16, M23, bold) are moderately preferred over two-state models .
+Moderate evidence exists towards the inclusion of breeding systems in ploidy models (M16, M23, bold).
+Moderate to no evidence towards the  inclusion of ploidy in breeding system models is indicated by the comparisons of models M28 \vs M16 and M29 \vs M23 where the difference is the inclusion of ploidy information.}
 \label{table:lumped}
@@
 \begin{figure}
+\centering
+\includegraphics[width=0.5\textwidth]{states.pdf}
+\caption{
+Character states used in the models.
+(A) Each species retained on the phylogeny belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known. 
+Number of species in each category is indicated; for example, 70 species are self-compatible with unknown ploidy.
+Character state abbreviations are: $I$ for self-incompatible, $C$ for self-compatible, $D$ for diploid, $P$ for polyploid, $?$ for unknown.
+Because polyploidization breaks this form of self-incompatibility, self-incompatible species with unobserved ploidy ($I?$) are taken to be diploid ($ID$), and polyploid species with unobserved breeding system ($?P$) are taken to be SC ($CP$).
+(B) Category groupings into states for each model class.
+In the ploidy-only models (M1- M10), states are coded as $D$ \& $P$ when uncertain/consistent with either state; in the breeding system-only models (M11-M15) such states are coded as $C$; in the ploidy and breeding system models (M16-M29), they are coded as $CD$ \& $CP$.
+In the hidden-trait models, all species could take on either of two `hidden' character states. %IM: #33 
+Two species, \emph{Lycium californicum} and \emph{Solanum bulbocastanum}, are simultaneously $ID$ and $CP$, and by adding them the sample adds to the total of 651 taxa used for analyses.
+}\label{figure:stateclassifications}
+\end{figure}
+
+\begin{figure}
 \includegraphics[width=\textwidth]{alldiversificationasymhidden.pdf} 
 \caption{Net diversification rates for SSE models of focal traits with or without hidden state. 
-Each panel contains a graphical summary of estimated model parameters and displays posterior distributions for net diversification estimates.}
+(A) Ploidy only model (M1) showing higher net diversification linked to diploid state $D$ compared to polyploid state $P$.
+(B) Ploidy with hidden states model (M4) showing that the net diversification is higher for hidden state $B$ (lighter colors) compared to hidden state $A$ (darker colors) and both diploid and polyploid states within each hidden traits have overlapping net diversification rates.
+(C) Breeding system only model (M9) showing higher net diversification linked to self-incompatible $I$ state compared to self-compatible state $C$.
+(D) Breeding system with hidden states model (M11) showing diversification differences in both hidden states (light \vs.
+dark colors) and little to no overlapping in between self-compatible \vs self-compatible states.
+(E) Ploidy and breeding system model (M13) showing higher net diversification linked to self-incompatible diploid state $ID$ compared to both self-compatible states despite ploidy level ($CD, CP$).
+(F) Ploidy, breeding system, and hidden states model (M16) showing a similar pattern that panel (E) within each hidden state $A$ and $B$.
+For hidden state $B$ there is a larger overlap of net diversification between states $CD_B$ and $ID_B$.}
 \label{figure:netdivall}
@@
 \includegraphics[width=\textwidth]{lumped.pdf} %lumped.pdf
-\caption{Comparing two and three state models is possible via lumpable models (M26, M28).
-These models used the same classifications as the three-state model M16 which allow for comparisons via Bayes Factors as shown in \cref{table:lumped}.}
+\caption{Comparing two and three-state models using lumped models. (A) Ploidy model only (M1) where data enter as binary $D$ and $P$. 
+(B) Lumped model for ploidy (M26) where data are the three-state values ($ID,CP,CD$) but results are equivalent to model M1.  
+(C) Ploidy and breeding system model (M16) where  data enter as the three-state values. Models M26 and M16 are comparable whereas M1 and M16 are not.
+(D) Breeding system only model (M11) where data are entered as binary $I$ and $C$. 
+(E) Lumped model for breeding system (M28) where data are the three-state values ($ID,CP,CD$) but results are equivalent to model M11. Model M26 can be compared with model M16 from panel (C).
+Model comparisons are done via Bayes factors and results are shown in \cref{table:lumped}}  
 \label{figure:lumped}
@@
 
+% check values of delta and rho D/P+A/B with(out) delta
+% i.e. parameter values in figure S1 & S4 vs Table S2
+
+\begin{suppfigure}
+    \caption{ Twenty-nine models of diversification are proposed for the study of ploidy, breeding systems, and hidden states linked to the process of diversification.
+We divide the models by the type of focal trait studied (ploidy only, breeding system only, or ploidy and breeding system).
+The contributions of the focal trait to the diversification process can be measured by comparing the models in each of the columns.
+That is, the focal trait only models assume that speciation and extinction rates are only linked to the trait itself, the hidden trait only models assume that the diversification rates are  linked to unknown factors but not the trait of interest, and the focal trait with hidden trait models assume that both the focal trait and unknown factors are contributing to diversification.}
+    \label{fig:allmodels}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
+\caption{Posterior distributions for each of the parameters in the ploidy only model (M1).
+Red color represents diploid state $D$ and blue color represents polyploid state $P$.
+(A)Speciation rates.
+(B) Extinction rates.
+(C) Net diversification rates (speciation minus extinction from panels A and B).
+(D) Relative extinction rates (extinction divided by speciation from panels A and B).
+(E) Polyploidization rate ($\rho$).} % XXX
+\label{suppfigure:DPnodip}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{hisseDPnodipasymposteriordist.pdf}
+\caption{Posterior distributions for each of the parameters in the ploidy and hidden trait model (M4).
+Red color represents diploid state $D$ and blue color represents polyploid state $P$.
+Dark colors represent hidden state $A$ and light colors hidden state $B$.
+(A) Speciation rates.
+(B) Extinction rates.
+(C) Net diversification rates (speciation minus extinction from panels A and B).
+(D) Relative extinction rates (extinction divided by speciation from panels A and B).
+(E) Polyploidization rate ($\rho$).
+(F) Transition rates between hidden states ( $\alpha$ and $\beta$).} % XXX
+\label{suppfigure:DPnodipAB}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{bisseSIposteriordist.pdf}
+\caption{Posterior distributions for each of the parameters in the breeding system only model (M11).
+Green color represents self-incompatible state $I$ and purple color represents self-compatible state $C$.
+(A) Speciation rates.
+(B) Extinction rates.
+(C) Net diversification rates (speciation minus extinction from panels A and B).
+(D) Relative extinction rates (extinction divided by speciation from panels A and B).
+(E) Self-incompatible to self-compatible transition rate ($q_IC$).} % XXX
+\label{suppfigure:IC}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{hisseSIasymposteriordist.pdf}
+\caption{Posterior distributions for each of the parameters in the breeding system and hidden trait model (M14).
+Green color represents self-incompatible state $I$ and purple color represents self-compatible state $A$.
+Dark colors represent hidden state $A$ and light colors hidden state $B$.
+(A) Speciation rates.
+(B) Extinction rates.
+(C) Net diversification rates (speciation minus extinction from panels A and B).
+(D) Relative extinction rates (extinction divided by speciation from panels A and B).
+(E) Self-incompatible to self-compatible transition rate ($q_IC$).
+(F) Transition rates between hidden states ( $\alpha$ and $\beta$).} % XXX
+\label{suppfigure:ICAB}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the ploidy and breeding system model (M16).
+Green color represents self-incompatible diploid state $ID$, blue color is the self-compatible and diploid state $CD$ and pink represents self-compatible polyploid state $CP$.
+(A) Speciation rates.
+(B) Extinction rates.
+(C) Net diversification rates (speciation minus extinction from panels A and B).
+(D) Relative extinction rates (extinction divided by speciation from panels A and B).
+(E) Self-incompatible to self-compatible transition rate ($q_IC$, yellow), polyploidization rate from self-incompatible ($\rho_I,$ light pink), and polyploidization from self-compatible ($\rho_c$, orange).} % XXX
+\label{suppfigure:IDCDCPnodip}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{muhissenodipasymposteriordist.pdf}
+\caption{Posterior distribution for each of the parameters in the ploidy, breeding system, and hidden trait model (M19).
+Green color represents self-incompatible diploid state $ID$, blue color is the self-compatible and diploid state $CD$ and pink represents self-compatible polyploid state $CP$.
+Dark colors represent hidden state $A$ and light colors hidden state $B$.
+(A) Speciation rates.
+(B) Extinction rates.
+(C) Net diversification rates (speciation minus extinction from panels A and B).
+(D) Relative extinction rates (extinction divided by speciation from panels A and B).
+(E) Self-incompatible to self-compatible transition rate ($q_IC$, yellow), polyploidization rate from self-incompatible ($\rho_I,$ light pink), and polyploidization from self-compatible ($\rho_c$, orange).
+(F) Transition rates between hidden states ( $\alpha$ and $\beta$).}% XXX
+\label{suppfigure:IDCDCPnodipAB}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrDP.pdf}
+\caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under  the ploidy only model (M1).
+} % XXX
+\label{suppfigure:DPnodipasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrDPAB.pdf}
+\caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the ploidy and hidden states model (M4).} % XXX
+\label{suppfigure:DPnodipABasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIC.pdf}
+\caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the  breeding system only model (M11).} % XXX
+\label{suppfigure:ICasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrICAB.pdf}
+\caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the breeding system and hidden states model (M14.} % XXX
+\label{suppfigure:ICABasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIDCDCP.pdf}
+\caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the ploidy and breeding system model (M16).} % XXX
+\label{suppfigure:IDCDCPnodipasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{asrIDCDCPAB.pdf}
+\caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the  ploidy, breeding systems, and hidden states model (M19).} % XXX
+\label{suppfigure:IDCDCPnodipABasr}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{comparablesDP.pdf}
+\caption{Testing the addition of breeding system to ploidy models.
+(A) Ploidy only model (M1) where data enter as binary $D$ and $P$.
+(B) Lumped model for ploidy (M26) where data are the three-state values ($ID,CP,CD$) but results are equivalent to model M1.
+(C) Ploidy and breeding system model (M16) where  data enter as the three-state values.
+Models M26 and M16 are comparable whereas M1 and M16 are not.
+(D) Ploidy and hidden state model (M3) where data enter as binary $D$ and $P$.
+(E) Lumped model for ploidy and hidden state (M27) where data are the three-state values ($ID,CP,CD$) but results are equivalent to model M3.
+(F) Ploidy, breeding system, and hidden state model (M18) where  data enter as the three-state values.
+Models M27 and M18 are comparable whereas M3 and M18 are not.
+Bayes factors comparing the models are shown in \cref{table:lumped}.} % XXX
+\label{suppfigure:lumpedDP}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{comparablesIC.pdf}
+\caption{Testing the addition of ploidy to breeding system models.
+(A) Breeding system only model (M11) where data enter as binary $I$ and $C$.
+(B) Lumped model for breeding system (M28) where data are the three-state values ($ID,CP,CD$) but results are equivalent to model M11.
+(C) Ploidy and breeding system model (M16) where  data enter as the three-state values.
+Models M28 and M16 are comparable whereas M11 and M16 are not.
+(D) Breeding system and hidden state model (M13) where data enter as binary $I$ and $C$.
+(E) Lumped model for breeding system and hidden state (M29) where data are the three-state values ($ID,CP,CD$) but results are equivalent to model M13.
+(F) Ploidy, breeding system, and hidden state model (M18) where  data enter as the three-state values.
+Models M29 and M18 are comparable whereas M13 and M18 are not.
+Bayes factors comparing the models are shown in \cref{table:lumped}.} % XXX
+\label{suppfigure:lumpedIC}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{effectofasymmetry.pdf}
+\caption{Effect of asymmetric rates in hidden models.
+First column models M3, M13, and M18 assume that the rates between hidden states are equal.
+The models in the second column (M4, M14, M19) assume that the rates between hidden states are different .
+Column three models assumes that the rates between hidden state are asymmetric and that the transition rates within each hidden states are also different.
+Bayes factors in \cref{supptable:asymmetry} strongly preferred models with asymmetric rates between states (second column) over models with equal rates in hidden states (first column).
+Models in the second column are moderately or equally prefer to more complex models in column 3.
+} % XXX
+\label{suppfigure:asymmetric}
+\end{suppfigure}
+
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{alldiversificationasymhiddendip.pdf}
+\caption{Posterior distributions for the net diversification rates of the preferred models with diploidization.
+Red color represents diploid state $D$, blue color represents polyploid state $P$, green color represents self-incompatible $I$, purple color represents self-compatible $C$,  dark colors represent hidden state $A$ and light colors hidden state $B$.
+(A) Ploidy only model M6.
+(B) Ploidy and hidden states model M9 (C) Breeding systems only model M11.
+(D) Breeding systems and hidden state model M14.
+(E) Ploidy and breeding systems model M21.
+(F) Ploidy, breeding systems, and hidden states model M24.} % XXX
+\label{suppfigure:alldip}
+\end{suppfigure}
+
+% Supplementary TABLES
+
 \begin{supptable}
@@
 \end{tabular}
-\caption{Bayes factors of  ploidy-only models with diplodization.
-Results indicate that a character independent model (M6) is strongly preferred over BiSSE (M1)  model.
-M6 is moderately preferred over any of the HiSSE models with asymmetric hidden rates (M9, M10).}
+\caption{Bayes factors in log-scale of  ploidy only models with diplodization.
+Results indicate that a character independent model (M7) is strongly preferred over model M6 (BiSSE).
+Model M7 (bold) is also moderately preferred over any of the HiSSE models with asymmetric hidden rates (M9, M10).}
 \label{supptable:M6M10}
@@
 \end{tabular}
-\caption{Bayes factors of ploidy and breeding system without diploidization.
-Results indicate that the MuHiSSE model with asymmetric hidden rates (17) is  strongly preferred over M16 and M18 and moderately preferred over the MuHiSSe with all rates asymetric (M20).
+\caption{Bayes factors of ploidy and breeding system without diploidization in log-scale.
+Results indicate that the model with asymmetric hidden rates (M19, bold) is  strongly preferred over M16 and M18 and moderately preferred over the MuHiSSe with all rates asymetric (M20).
 *Marginal log-likelihood for M17 could not be calculated within allotted computer time.}
@@
 \end{tabular}
-\caption{Bayes factors of ploidy and breeding system with diploidization.
-Results indicate that the MuHiSSE model with asymmetric hidden rates (M24) is  strongly preferred over M21-M23 and moderately preferred over the MuHiSSe with all rates asymmetric (M25).}
+\caption{Bayes factors of ploidy and breeding system with diploidization in log-scale.
+Results indicate that the MuHiSSE model with asymmetric hidden rates (M24, bold) is  strongly preferred over M21-M23 and moderately preferred over the MuHiSSe with all rates asymmetric (M25).}
 \label{supptable:M21M25}
@@
 \end{tabular}
-\caption{Test for addition of hidden states in models via Bayes factors. Models with hidden states are strongly preferred over simpler models}
+\caption{Test for addition of hidden states in models via Bayes factors (in log-scale).
+Models with hidden states  (M4, M14, M19, M9, M24, bold) are strongly preferred over simpler models that do not include hidden s}
 \label{supptable:testaddhidden}
@@
 \end{tabular}
-\caption{Test for symmetry of the hidden trait transition rates via Bayes factors.
-For all models, asymmetric transition rates between hidden trait states are preferred over models with equal rates.
-Adding more complexity by assuming asymmetry in all rates within both hidden states does not improve models.} %FIXME/CHECK
+\caption{Test for asymmetry of the hidden trait transition rates via Bayes factors.
+For all models, asymmetric transition rates between hidden trait states are preferred over models with equal rates (bold).
+Adding more complexity by assuming asymmetry in all rates within both hidden states is not preferred over models with just asymmetry between hidden states.}
 \label{supptable:asymmetry}
@@
 \end{tabular}
-\caption{Test for inclusion of a diploidization rate via Bayes factors. Models with diploidization are moderately preferred over models that do not include a diploidization rate}
+\caption{Test for inclusion of a diploidization rate via Bayes factors. Models with diploidization are moderately preferred over models that do not include a diploidization rate (bold).}
 \label{supptable:testdiploidization}
 \end{supptable}
-
-% check values of delta and rho D/P+A/B with(out) delta
-% i.e. parameter values in figure S1 & S4 vs Table S2
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M1, D/P ploidy model.} % XXX
-\label{suppfigure:DPnodip}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{hisseDPnodipasymposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M4, D/P+A/B asym model.} % XXX
-\label{suppfigure:DPnodipAB}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{bisseSIposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M11, I/C breeding system model.} % XXX
-\label{suppfigure:IC}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{hisseSIasymposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M14, I/C+A/B asym model.} % XXX
-\label{suppfigure:ICAB}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the  M16, ID/CD/CP model.} % XXX
-\label{suppfigure:IDCDCPnodip}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{muhissenodipasymposteriordist.pdf}
-\caption{Posterior distribution for each of the parameters in the M19, ID/CD/CP+A/B asym model.} % XXX
-\label{suppfigure:IDCDCPnodipAB}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDP.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the M1, D/P ploidy model. } % XXX
-\label{suppfigure:DPnodipasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDPAB.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the M4, D/P+A/B asym model.} % XXX
-\label{suppfigure:DPnodipABasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIC.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the M1, I/C breeding system model.} % XXX
-\label{suppfigure:ICasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrICAB.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the M14, I/C+A/B asym model.} % XXX
-\label{suppfigure:ICABasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCP.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M16, ID/CD/CP model.} % XXX
-\label{suppfigure:IDCDCPnodipasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCPAB.pdf}
-\caption{Ancestral state estimation using the maximum a posteriori for each node of the  M19, ID/CD/CP+A/B asym model.} % XXX
-\label{suppfigure:IDCDCPnodipABasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{comparablesDP.pdf}
-\caption{Lumped ploidy models to assess the effect of adding breeding system to create a three-state model.
-Moderate evidence exists that adding a breeding system is necessary \cref{table:lumped}.
-} % XXX
-\label{suppfigure:lumpedDP}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{comparablesIC.pdf}
-\caption{Lumped breeding system models to assess the effect of adding ploidy to create a three-state model.
-Moderate to insignificant evidence exists that adding ploidy is necessary \cref{table:lumped}.
-} % XXX
-\label{suppfigure:lumpedIC}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{effectofasymmetry.pdf}
-\caption{Effect of asymmetric rates in hidden models. Models with asymmetric rates are preferred over models with equal rates \cref{supptable:asymmetry}. } % XXX
-\label{suppfigure:asymmetric}
-\end{suppfigure}
-
-\begin{suppfigure}
-\includegraphics[width=\textwidth]{alldiversificationasymhiddendip.pdf}
-\caption{Posterior distribution for all the best models with diploidization.} % XXX
-\label{suppfigure:alldip}
-\end{suppfigure}

### 2019-08-31 finalmanuscript/methods.tex
final manuscript for submission

--- 2019-08-30 manuscript/methods.tex
@@
 Of these, 368 had information for both states.
-The number of species in each combination of states is summarized in \cref{figure:stateclassifications}A.
+The number of species in each combination of states is summarized in \cref{figure:stateclassifications}A and in \cref{suppfigure:stateclassifications}.
 We retained all 651 species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
@@
 
-%%% Below here is old tables.  I think the new versions should instead go in results.tex or a separate file. %%%
-
-% \begin{table}
-% \begin{tabular}{lcccccc} \toprule
-% \begin{tabular}{@{}lcccccc@{}} \toprule
-%\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
-%Model                     & Ploidy & Diploidization & Breeding  & Hidden & Num        & Marginal       \\
-%                          &        &                & System    & State  & Parameters & Log-Likelihood \\ \midrule
-%1. D/P                    & Yes    & No             & No        & No     & 5          & -1193.66\\
-%2. D/P +$\delta$          & Yes    & Yes            & No        & No     & 6          & -1182.93 \\
-%3. D/P+A/B                & Yes    & No             & No        & Yes    & 10         & -1150.99\\
-%4. D/P+A/B+$\delta$       & Yes    & Yes            & No        & Yes    & 11         & \textbf{-1145.69}\\
-%5. I/C                    & No     & No             & Yes       & No     & 5          & -1194.80 \\
-%6. I/C+A/B                & No     & No             & Yes       & Yes    & 10         & \textbf{-1155.37}\\
-%7. ID/CD/CP               & Yes    & No             & Yes       & No     & 9          & -1345.87\\
-%8. ID/CD/CP+$\delta$      & Yes    & Yes            & Yes       & No     & 10         & -1344.50\\
-%9. ID/CD/CP+A/B           & Yes    & No             & Yes       & Yes    & 15         & -1303.55 \\ 
-%10. ID/CD/CP+A/B+$\delta$ & Yes    & Yes            & Yes       & Yes    & 16         & \textbf{-1300.35} \\
-%\bottomrule
-%\end{tabular}
-%\caption{
-   % The ten models and their marginal likelihoods.
-    %Values in bold are for the best models within each class that are comparable (see \cref{table:}).
-    %Abbreviations are D: diploid, P: polyploid, I: self-incompatible, C: self-compatible, A: one state of hidden trait, B: other state of hidden trait, $\delta$: diploidization.}
-%\label{table:marginallike}
-%\end{table}
-
 \begin{table}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{states.pdf}
+%\includegraphics[width=0.5\textwidth]{fig1.pdf}
 \caption{
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{alldiversificationasymhidden.pdf} 
+%\includegraphics[width=\textwidth]{fig2.pdf} 
 \caption{Net diversification rates for SSE models of focal traits with or without hidden state. 
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{lumped.pdf} %lumped.pdf
+%\includegraphics[width=\textwidth]{fig3.pdf} %lumped.pdf
 \caption{Comparing two and three-state models using lumped models. (A) Ploidy model only (M1) where data enter as binary $D$ and $P$. 
@@
 \begin{figure}
-    \centering 
-    \includegraphics[width=0.5\textwidth]{pathstates2} \\ [40pt] 
-        \includegraphics[width=0.9\textwidth]{pathways}
+   % \centering 
+    %\includegraphics[width=0.5\textwidth]{fig4a} \\ [40pt] 
+       % \includegraphics[width=0.9\textwidth]{fig4b}
     \caption{
@@
 % i.e. parameter values in figure S1 & S4 vs Table S2
+\begin{suppfigure}
+%\includegraphics[width=\textwidth]{figS1.pdf}
+    \caption{Ploidy and breeding system data according to three different classifications.
+For ploidy only models, classifications with states $D$ and $P$ were used (inner circle).
+For breeding system models classifications with states $I$ and $C$ were used (middle circle).
+For ploidy and breeding system models classifications using $ID, CD, CP$ were used (outer circle).
+Data with missing information in one of the traits was classified simultaneously as two possible states, for example, diploids without breeding system $?D$ were classified as $(CD, CP)$).}
+    \label{fig:allmodels}
+\end{suppfigure}
 
-\begin{suppfigure}
+\begin{suppfigure} %figS2 all models
     \caption{ Twenty-nine models of diversification are proposed for the study of ploidy, breeding systems, and hidden states linked to the process of diversification.
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
+%\includegraphics[width=\textwidth]{figS3.pdf}
 \caption{Posterior distributions for each of the parameters in the ploidy only model (M1).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{hisseDPnodipasymposteriordist.pdf}
+%\includegraphics[width=\textwidth]{figS4.pdf}
 \caption{Posterior distributions for each of the parameters in the ploidy and hidden trait model (M4).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{bisseSIposteriordist.pdf}
+%\includegraphics[width=\textwidth]{figS5.pdf}
 \caption{Posterior distributions for each of the parameters in the breeding system only model (M11).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{hisseSIasymposteriordist.pdf}
+%\includegraphics[width=\textwidth]{figS6.pdf}
 \caption{Posterior distributions for each of the parameters in the breeding system and hidden trait model (M14).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
+%\includegraphics[width=\textwidth]{figS7.pdf}
 \caption{Posterior distribution for each of the parameters in the ploidy and breeding system model (M16).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{muhissenodipasymposteriordist.pdf}
+%\includegraphics[width=\textwidth]{figS8.pdf}
 \caption{Posterior distribution for each of the parameters in the ploidy, breeding system, and hidden trait model (M19).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDP.pdf}
+%\includegraphics[width=\textwidth]{figS9.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under  the ploidy only model (M1).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDPAB.pdf}
+%\includegraphics[width=\textwidth]{figS10.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the ploidy and hidden states model (M4).} % XXX
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIC.pdf}
+%\includegraphics[width=\textwidth]{figS11.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the  breeding system only model (M11).} % XXX
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrICAB.pdf}
+%\includegraphics[width=\textwidth]{figS12.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the breeding system and hidden states model (M14.} % XXX
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCP.pdf}
+%\includegraphics[width=\textwidth]{figS13.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the ploidy and breeding system model (M16).} % XXX
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCPAB.pdf}
+%\includegraphics[width=\textwidth]{figS14.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the  ploidy, breeding systems, and hidden states model (M19).} % XXX
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{comparablesDP.pdf}
+%\includegraphics[width=\textwidth]{figS15.pdf}
 \caption{Testing the addition of breeding system to ploidy models.
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{comparablesIC.pdf}
+%\includegraphics[width=\textwidth]{figS16.pdf}
 \caption{Testing the addition of ploidy to breeding system models.
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{effectofasymmetry.pdf}
+%\includegraphics[width=\textwidth]{figS17.pdf}
 \caption{Effect of asymmetric rates in hidden models.
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{alldiversificationasymhiddendip.pdf}
+%\includegraphics[width=\textwidth]{figS18.pdf}
 \caption{Posterior distributions for the net diversification rates of the preferred models with diploidization.

### 2019-08-31 manuscript/onlyfigs.tex
final manuscript for submission

--- 2019-08-31 finalmanuscript/methods.tex
@@
-\section{Methods}
-
-\subsection{Data}
-
-%B: IMO: The data subsection would be better organized with parts, (1) ploidy, (2) breeding system, (3) tree, (4) models, (5) model selection and statistical inference.
-% E: I changed the subsectioning, but to a different arrangement: (1) Data, (2) Models, (3) Statistical inference.  With subsubsections as necessary.
-
-Chromosome number data were obtained for all Solanaceae taxa in the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were curated semi-automatically using the \mbox{CCDBcurator} R package \citep{rivero_2019}.
-CCDB contains records from original sources that have multiple complex symbol patterns denoting multivalence, or irregularites of chromosome counts.
-After a first round of automatic cleaning, we examined results by hand and corrected records as necessary.
-Our hand-curated records were also contrasted against the ploidy dataset from \citet{robertson_2011}, original references therein, and against ploidy data in the C-value DNA dataset from \citet{bennett_2005}.
-By comparing three different sources of information, we were able to code taxa as diploid, $D$, or polyploid, $P$.
-For the majority of species, ploidy was assigned according to information from the original publications included in the  C-value DNA dataset \citep{bennett_2005}.
-For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus/family, or based on SI/SC classification.
-For example, \textit{Solanum betaceum} did not have information about ploidy level, but it has 2n=24 chromosomes, and  $x=12$ is the base chromosome number of the genus \textit{Solanum} \citep{olmstead_2007}, so we assigned \textit{S.~betaceum} as diploid.
-Additionally, because of the absence of SI polyploids (explained above and below), species known to be SI could be scored as diploid.
-Species with more than one ploidy level were assigned the most frequent ploidy level recorded or the smallest ploidy in case of frequency ties.
-
-Breeding system states were scored as self-incompatible, $I$, or self-compatible, $C$, based on results curated from the literature  \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012} and original experimental crosses (B.I.
-and E.E.G., unpub.
-data).
-%IM suggestion %B:  I returned the aside about original data.
-I don't think Itay knows that we performed original crosses (many), and they are incorporated in the data, without being present in "the literature"
-%was originally: Breeding system states were scored as self-incompatible, $I$, or self-compatible, $C$, based on results curated from the literature and original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
-Most species could unambiguously be coded as either $I$ or $C$ \citep{raduski_2012}.
-Following previous work, we coded any species with a functional SI system as $I$, even if SC or dioecy was also reported.
-Dioecious species without functional SI were coded as $C$.
-% E: I am trying to stick with I/C when talking about the state values, but using SI/SC when talking about the meaning of the trait (because this what people use normally). 
-%R- Okay- I was getting confused as well thanks for clarifying
-
-Resolution of taxonomic synonymy followed Solanaceae Source \citep{solsource}. 
-Hybrids and cultivars were excluded because ploidy and breeding system can be affected by artificial selection during domestication.
-Following the reasoning outlined in \citet{robertson_2011}, we closely examined the few species for which the merged ploidy and breeding system data indicated the presence of self-incompatible polyploids.
-Although SI populations frequently contain some SC individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find convincing data for a naturally occurring SI polyploid population  \citep[discussed in][]{robertson_2011}.
-%B rm: The single instance of an SI polyploid individual appears to be an allopentaploid hybrid of \textit{Solanum oplocense} Hawkes x \textit{Solanum gourlayii} Hawkes, reported by \citet{camadro_1981}.
-% Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain SI, so long as they express a single allele at the S-locus \citep[discussed in][]{robertson_2011}.
-Because of the resulting absence of polyploid SI populations, as well as the functional explanation for polyploidy disabling gametophytic SI systems with non-self recognition (see the Introduction), we consider only three observed character states: self-incompatible diploids $(ID)$, self-compatible diploids $(CD)$, and self-compatible polyploids $(CP)$.
-
-Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 651 species with ploidy and/or breeding system information on the tree.
-% Fixed!
-Of these, 368 had information for both states.
-The number of species in each combination of states is summarized in \cref{figure:stateclassifications}A and in \cref{suppfigure:stateclassifications}.
-We retained all 651 species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
-A total of 372 taxa without any information about breeding system or ploidy were excluded.
-%Tips without trait data might not improve point estimation of diversification parameters while increasing uncertainty in the estimates of trait linked diversification
-% R-This is very important but I have removed it because I think it is a full on paper by its own
-%Including this many more species would have prohibitively slowed our analyses, especially those implementing the most complex models.
-
-The Supplementary Information contains citations for the numerous original data sources. % FIXME (and say here how many refs?) %B: no need for an additional counting and updating headache?
-% R- There is few data that I added so I can find those references and send them to you, so it would be simply the same references that you previously had plus another 20-25 I think
-The Dryad archive contains the data and tree files used for analyses. % FIXME
-% \url{https://github.com/roszenil/solploidy}  E: we should use dryad instead, because github repos are not guaranteed to last
-%This fixme happens after publications
-
-\subsection{Models}
-
-In order to test our hypotheses about lineage diversification and trait macroevolution, we fit twenty-nine state-dependent speciation and extinction models \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016}.
-SSE models contain parameters that describe per-lineage rates of speciation and extinction, specific to each character state (denoted $\lambda$ and $\mu$, respectively, with subscripts to indicate the state), along with rates of transitions between states (denoted $\rho$ for polyploidization, $\delta$ for diploidization, and $q_{IC}$ for loss of self-incompatibility).
-The full set of models and all their rate parameters are detailed in \cref{fig:allmodels}.
-Here, we summarize how each model allows us to assess whether diversification is best explained by variation in ploidy, breeding system, their combination, or some unknown factor.
-
-\subsubsection{Ploidy and diversification}
-
-We first employed a model (labeled M1), previously used by \citet{mayrose_2011}, with each species classified as diploid $(D)$ or polyploid $(P)$.
-Although this model can be powerful in studies of trait evolution, it is prone to incorrectly reporting that a trait is associated with diversification differences \citep{maddison_2015, rabosky_2015}.
-We therefore define several models that incorporate additional forms of diversification rate heterogeneity.
-The second ploidy model (M2) includes a binary hidden trait that subdivides each observed state.
-In this trait-independent model known as CID \citep{beaulieu_2016}, hidden traits can affect diversification but the observed traits do not.
-Comparing M1 and M2 allows us to test whether diversification rate heterogeneity is better explained by ploidy or by some unknown factor.
-
-We fit three models in which both ploidy and a hidden trait could influence diversification (M3--M5).
-These models differ in whether transitions between the hidden states are symmetric (M3) or asymmetric(M4), and whether the polyploidization rate depends on the hidden state(M5). %WF suggestion
-Comparing M1--M5 (as shown in \cref{table:bayesfactors}) allows us to test whether ploidy is associated with diversification differences on top of the differences potentially explained by an unknown factor.
-We further fit the analogues of these five models but including a rate parameter $\delta$ for transitions from polyploid to diploid (M6--M10).
-These comparisons allow us to assess whether our conclusions about ploidy and diversification are robust to the possibility of diploidization.
-
-\subsubsection{Breeding system and diversification}
-
-We propose five breeding system models following the same logic as the ploidy models above.
-Under the simplest breeding system and diversification model (M11), species are classified as self-incompatible $(I)$ or self-compatible $(C)$.
-This is the same model as in the analysis presented in \citet{goldberg_2010} but with an updated phylogeny \citep{sarkinen_2013} and a larger aggregated dataset.
-We then add models to allow diversification to be influenced by only a hidden trait (M12), or by both breeding system and a hidden trait (M13--M15, with varying degrees of complexity in the hidden trait transitions \cref{table:M11M15}).
-Similar models were used by \citet{freyman_2019} to study diversification in Onagraceae.
-
-Self-incompatibility is homologous in all Solanaceae species in which S-alleles have been cloned and controlled crosses performed.
-All species sampled to date possess a non-self recognition, RNase-based gametophytic self-incompatibility \citep[shared even with other euasterid families;][]{ramanauskas_2017}.
-Furthermore, species that are distantly related within this family carry closely-related alleles, with deep trans-specific polymorphism at the locus that controls the SI response \citep{ioerger_1990, igic_2006}.
-Thus, there is strong evidence in Solanaceae that the $I$ state is ancestral in the family, and that the SI mechanism was not regained.
-For all breeding system models, we estimated a transition rate from $I$ to $C$  but not the reverse ($q_{CI}=0$).
-% This is an important assumption because ancestral state reconstruction in the models might differ when polyploidy and breeding system are analyzed as independent in the Solanaceae tree.
-%
-%B: What happens if SC->SI is not fixed? Asking for a friend.  E: Surprisingly, Rosana reports, results are not turned on their head. 
-%B: Whew, I was worried.
-Er, "friend" says thanks.
-R- Your welcome, by mistake I did that model with reversibility from C to I.
-I didn't run it for as long as the rest but preliminary ancestral reconstructions seem almost identical when you add that parameter than when you don't.
-
-\subsubsection{Ploidy, breeding system, and diversification}
-
-Ploidy and breeding system might influence lineage diversification individually, but these two traits also have an intricate association (discussed in the Introduction).
-Therefore, we considered several multi-state models that investigate the contribution of both traits and the allowable transitions between them.
-
-The simplest model (M16) classifies each species as either SI diploid $(ID)$, SC diploid $(CD)$, or SC polyploid $(CP)$; recall that SI polyploids do not occur.
-Each of these states may again be associated with different rates of speciation and extinction, and the allowable transitions are loss of SI within the diploid state (from $ID$ to $CD$), loss of SI via polyploidization (from $ID$ to $CP$), and polyploidization while SC (from $CD$ to $CP$).
-% The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
-% E: Can we call it something other than q_IC?  It's a bit confusing to have the same parameter mean different things in different models.
-% R: In the lumped models I call them something different q_0 and rho_0 because they are different parameters but I don't think we should explain.
-As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by ploidy, breeding system, and a hidden trait (M18--M20) with varying degrees of complexity in the hidden trait transitions
-\citep[similar to][]{caetano_2018, huang_2018}. %IM suggestion
-We also fit the analogous models but allowing for diploidization (M21--M25).
-
-\subsubsection{Lumped models}
-
-The models described so far allow us to assess the contributions of our two focal characters---ploidy and breeding system---to lineage diversification, but they do not reveal whether it is valuable to include both characters in the analysis.
-To answer this question with statistical model comparisons requires comparing the likelihood of the data given each model.
-This is impossible for the ploidy and breeding system models presented so far, however, because the data are different for the different models: they use either the $D/P$ or the $I/C$ or the $ID/CD/CP$ state spaces (see figure \cref{figure:stateclassifications} for state classifications).
-Therefore, the use of different data results in incomparable models.
-
-In order to compare fits of ploidy-only \vs breeding system-only \vs combined trait models, we use the technique of `lumping' states together \citep{tarasov_2019}.
-We use the state space of the $ID/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
-Lumping states requires that the transition rates from the lumped state to the singular state be equal \citep{tarasov_2019}.
-First, we lump together $ID$ and $CD$ to form the diploid state, mimicking the $D/P$ model (M26). 
-Proposing a lumped ploidy model by aggregating $ID$ and $CD$ requires forcing the rate of polyploidization from $ID$ and $CD$ to $CP$ to be equal (i.e. $\rho_0=\rho_I=\rho_c$), but also requires
-assuming that the rates of speciation and extinciton for $ID$ and $CD$ to be equal.
-Therefore, we define the new parameters $\lambda_D$ and $\mu_D$  that are the same for each of the two diploid states $ID$ and $CD$. %IM suggestion, describe more the lumping
-We used the same procedure to lump together $CD$ and $CP$ to form the self-compatible state, mimicking the $I/C$ model (M28).
-In this particular case, the rate from $CD$ to $CP$ back to $ID$ is zero and equal for both, so the model is lumpable. 
-However, to fully mimic the breeding system model, we assume that the rate of selfing is equal (i.e.
-$q_0=q_{IC}=\rho_I$) and the rates of speciation and extinction for both $CD$ and $CP$ are the same (new parameters $\lambda_C$ and $\mu_C$).
-We further add a hidden character to each of these models (M27 and M29), and then compare this group of models (\cref{table:lumped}).
-
-We do not include additional models with diploidization because this reverse ploidy transition renders the models non-lumpable.
-When including diploidization, transitions from $CP$ to $CD$ are at rate $\delta$ but transitions from $CP$ to $ID$ do not occur.
-Because $ID$ and $CD$ would be lumped to mimic the $D/P$ model, this model is non-lumpable when $\delta \ne 0$.
-Thus, we can compare models to test whether it is advantageous to include both traits, but only when ignoring diploidization.
-% E: Below was all good, but it seemed like more detail than necessary.
-% Then we define a single parameter $q_0$ going into the lumped state $(CD,CP)$ as the new rate from SI to SC.
-% This defines the lumped model M28 (\cref{figure:lumped}E) that is equivalent to M1 but comparable to M16.
-% We repeat this process to create the lumped model M26 for $D/P$ by restricting the polyploidy rate to be $rho_0$ out of the lumped state $(ID,CD)$.
-% When including diploidization, however, transitions from $CP$ to $CD$ are at rate $\delta$ but transitions from $CP$ to $ID$ do not occur.
-% Since $\delta \neq 0$, the models with diplodization cannot be lumped to be compared to the three-state models.
-
-\subsubsection{Pathways to polyploidy}
-
-Considering ploidy and breeding system together, there are two evolutionary pathways from SI diploid to SC polyploid \citep{brunet2001, robertson_2011}.
-In the one-step pathway, the $CP$ state is produced directly from the $ID$ state when whole genome duplication disables SI.
-In the two-step pathway, the $CD$ state is an intermediate: SI is first lost, and later the SC diploid undergoes polyploidization.
-We quantify the relative contribution of these pathways to polyploidy in two ways, each using the median estimates of rates from the simplest model that includes both traits (M16). 
-Our results differ from those of \citet{robertson_2011} in part because our approach relies on a dated phylogeny and models that allow for state-dependent diversification. %IM suggestion
-
-Both of our methods are based on a propagation matrix that describes flow from $ID$ to $CP$, as in \citet{robertson_2011}.
-We insert an artificial division in the $CP$ state, so that one sub-state contains the $CP$ species that arrived via the one-step pathway and the other substate contains the $CP$ species that arrived via the two-step pathway.
-We consider unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and no strong support for diploidization (see below).
-First, we consider only the rates of transitions between these states, placing them in the propagation matrix \myvec{Q}.
-The matrix $\myvec{P} = \exp(\myvec{Q} t)$ then provides the probabilities of changing from one state to any other state after time $t$.
-Closed-form solutions for the two pathway probabilities are provided in \citet{robertson_2011}.
-Second, we consider not only transitions between states but also diversification within each state.
-State-dependent diversification can change the relative contributions of the two pathways.
-For example, if the net diversification rate is small for $CD$, the two-step pathway will contribute relatively less.
-We therefore include the difference between speciation and extinction along the diagonal elements of the propagation matrix.
-As before, matrix exponentiation provides the relative chance of changing from one state to any other state after time $t$.
-The calculations of the propagation matrix are not probabilities because diversification changes the number of lineages as time passes. 
-We can still use ratios, however, to consider the relative contribution of each pathway, analogous to the normalized age structure in a growing population \citep{leslie_1945}.
-% E & R todo ?: for the case with net div, write out the propagation matrix and equations for pathway contributions
-% R- couldn't solve to get the exact solutions turns out it is really hard.
-
-\subsection{Statistical inference}
-
-\subsubsection{Model fitting}
-
-Parameters for each of the 29 models were coded as graphical models and Bayesian statistical inference was performed with RevBayes \citep{hoehna_2016}.
-Scripts for analyses and key results are available in Dryad. %TODO
-We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $651/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
-For all models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
-Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
-For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the models.
-Convergence and mixing of each MCMC chain was determined by ensuring the effective sample size of each parameter was over 200.
-
-We report posterior distributions for the net diversification parameters ($\lambda - \mu$)  in \cref{figure:netdivall} and full results of all parameters of the diversification models in \cref{suppfigure:DPnodip,suppfigure:DPnodipAB,suppfigure:IC,suppfigure:ICAB,suppfigure:IDCDCPnodip,suppfigure:IDCDCPnodipAB}.
-Additionally, ancestral states at each node in the phylogeny were sampled jointly during the MCMC analyses every 100 generations. 
-%R: Itay says remove but some reviewers were interested in the distribution of the states.
-I say leave we don't need to discuss.
-I'll mention briefly.
-%B: I am not sure what this comment refers to (up or down?), but I'm not touching.
-Ancestral state estimations for all models show the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes for each of the models in \cref{figure:netdivall}.
-(\cref{suppfigure:DPnodipasr,suppfigure:DPnodipABasr,suppfigure:ICasr,suppfigure:ICABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPnodipABasr}).
-
-\subsubsection{Model selection key questions}
-
-We calculated the marginal log-likelihood for each of the models using fifty stepping stone steps under the methodology of \citet{xie_2010}, implemented in RevBayes \citep{hoehna_2016}.
-Each stepping stone step was found by calculating at least 500 generations of burn-in followed by a total of 1,000 MCMC steps (\cref{table:bayesfactors}).
-
-Using the marginal likelihood values, we calculated Bayes factors to answer five key biological and methodological questions:
-\begin{enumerate}
-\item Are diversification patterns only determined by hidden states and not the traits of interest?---Comparison of character independent models against hidden state (\cref{table:bayesfactors,table:M11M15,supptable:M6M10,supptable:M16M20,supptable:M21M25}).
-\item Are hidden states necessary to explain diversification rate heterogeneity?---Comparison of simple models against hidden state models (\cref{supptable:testaddhidden}).
-\item Does a second focal trait add information about the diversification process?---Comparison of lumped models against IC/CD/CP models (\cref{figure:lumped,table:lumped,suppfigure:lumpedDP,suppfigure:lumpedIC}).
-\item Are conclusions robust to assumptions about hidden state transitions?---Comparison amongst hidden states models with equal hidden rates and asymmetrical rates (\cref{supptable:asymmetry,suppfigure:asymmetric}).
-\item Is there evidence for diploidization?---Comparison amongst log-scalemodels with and without diploidization (\cref{supptable:testdiploidization,suppfigure:alldip}).
-\end{enumerate}
-
-Each model comparison is reported with a Bayes factor on the natural log-scale: the comparison between models $M_0$ and $M_1$ is $K=ln(BF(M_1,\ M_0)) = \ln[ P(\mathbf{X} | M_1) - P(\mathbf{X} | M_0)]$.
-There is `strong evidence' for $M_1$ when this value is more than 10, moderate support if the value is more than 1, and no evidence if the value is between -1 and 1.
-If the value of $K$ is negative the evidence goes towards $M_0$ \citep{kass1995}.
-
-\begin{table}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|r|c|c|c|c|l|}
-\hline
-Model                & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M2    & M3     & M4    & M5     & Evidence                                                                                     \\ \hline
-M1.
-D/P              & -1283.76                                                          & 59.90 & 49.23  & 60.48 & 58.82  & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M1\end{tabular}            \\ \hline
-\textbf{M2.
-CID D/P} & \textbf{-1223.86}                                                 &       & -10.66 & 0.579 & -1.079 & \begin{tabular}[c]{@{}l@{}}Strong preference for \\ M2 over M1 and M3\end{tabular}              \\ \hline
-M3.
-D/P+A/B          & -1234.52                                                          &       &        & 11.24 & 9.58   & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}   \\ \hline
-\textbf{M4.
-D/P+A/B asym}     & \textbf{-1223.28}                                                          &       &        &       & -1.658 & \begin{tabular}[c]{@{}l@{}}Moderate evidence for\\ only asymmetric hidden rates\end{tabular} \\\hline
-M5.
-D/P+A/B all asym & -1224.93                                                          &       &        &       &        &          \\ \hline
-\end{tabular}
-\caption{Bayes factors for ploidy only models without diplodization in log-scale.
-Numbers smaller than -1 indicate moderate preference for the model listed in the row, and values larger than 1 indicate moderate preference for the model listed in the respective column .
-Conventional threshold for `strong' preference is an absolute value larger than 10.
-The character independent model (M2, bold) is strongly preferred over the ploidy only model (M1) or a hidden state model with symmetric rates (M3).
-Moderate support exists for the model with ploidy and asymmetric hidden states (M4, bold) over the rest of the ploidy only models.}
-\label{table:bayesfactors}
-\end{table}
-
-% B: M111->M11
-\begin{table}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|r|c|c|c|c|l|}
-\hline
-Model                      & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M12   & M13   & M14   & M15   & Evidence                                                                                       \\ \hline
-M11.
-I/C                   & -1309.07                                                          & 41.13 & 38.59 & 61.34 & 61.40 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M11\end{tabular}            \\ \hline
-M12.
-CID I/C               & -1267.93                                                          &       & -2.53 & 20.21 & 20.37 & \begin{tabular}[c]{@{}l@{}}Models with asymmetric rates\\ are preferred over  M12\end{tabular} \\ \hline
-M13.
-I/C+A/B               & -1270.47                                                          &       &       & 22.75 & 22.80 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}   \\ \hline
-\textbf{M14.
-I/C+A/B asym} & \textbf{-1247.72}                                                 &       &       &       & 0.05 & No evidence                                                                                    \\ \hline
-\textbf{M15.
-I/C+A/B all asym }     & \textbf{ -1247.66}                                                          &       &       &       &       &       \\ \hline
-\end{tabular}
-\caption{Bayes factors for breeding system only models in log-scale.
-Numbers smaller than -1 indicate moderate preference for the model listed on the row, and values larger than 1 indicate moderate preference for the model in the respective column.
-Conventional threshold for `strong' preference is an absolute value larger than 10.
-Moderate to strong preference exist for models M14 and M15 which include a hidden states with asymmetrical rates.}
-\label{table:M11M15}
-\end{table}
-
-% E: Could we include a figure of the traits on the tree?  At least as supp info. %FIXME
-
-\begin{table}[]
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Model                       & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=ln(BF($M_{i}$,$M_{ii}$)) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
-M26.
-Lumped D/P             & -1463.22                                                          & M26 vs.
-M16 & 4.10            & M16 (Moderate)                                                       \\
-\textbf{M16.
-ID/CD/CP}      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
-M27.
-Lumped D/P+A/B         & -1417.67                                                          & M27 vs.
-M23 & 3.678           & M23 (Moderate)                                                       \\
-\textbf{M23.
-ID/CD/CP +A/B} & \textbf{-1414.00}                                                 &             &                 &                                                                      \\ \hline
-M28.
-Lumped I/C             & -1458.41                                                          & M28 vs.
-M16 & -0.69           & No evidence                                                          \\
-M16. ID/CD/CP      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
-M29.
-Lumped I/C+ A/B        & -1416.60                                                          & M29 vs.
-M23 & 2.60            & M23 (Moderate)                                                       \\
-\textbf{M23.
-IC/CD/CP+A/B}  & \textbf{-1414.00}                                                 &             &                 &                                                                      \\ \hline
-\end{tabular}
-\caption{Testing the addition of a focal trait to a binary state model via Bayes factors (log-scale).
-Numbers smaller than -1 indicate moderate preference for the model in the row, and values larger than 1 indicate moderate preference for the model in the respective column.
-Conventional threshold for `strong' preference is an absolute value larger than 10.
-Three-state models (M16, M23, bold) are moderately preferred over two-state models .
-Moderate evidence exists towards the inclusion of breeding systems in ploidy models (M16, M23, bold).
-Moderate to no evidence towards the  inclusion of ploidy in breeding system models is indicated by the comparisons of models M28 \vs M16 and M29 \vs M23 where the difference is the inclusion of ploidy information.}
-\label{table:lumped}
-\end{table}
-
 \begin{figure}
 \centering
-%\includegraphics[width=0.5\textwidth]{fig1.pdf}
+\includegraphics[width=0.5\textwidth]{fig1.pdf}
 \caption{
@@
 \begin{figure}
-%\includegraphics[width=\textwidth]{fig2.pdf} 
+\includegraphics[width=\textwidth]{fig2.pdf} 
 \caption{Net diversification rates for SSE models of focal traits with or without hidden state. 
@@
 \begin{figure}
-%\includegraphics[width=\textwidth]{fig3.pdf} %lumped.pdf
+\includegraphics[width=\textwidth]{fig3.pdf} %lumped.pdf
 \caption{Comparing two and three-state models using lumped models. (A) Ploidy model only (M1) where data enter as binary $D$ and $P$. 
@@
 \begin{figure}
-   % \centering 
-    %\includegraphics[width=0.5\textwidth]{fig4a} \\ [40pt] 
-       % \includegraphics[width=0.9\textwidth]{fig4b}
+    \centering 
+    \includegraphics[width=0.5\textwidth]{fig4a} \\ [40pt] 
+        \includegraphics[width=0.9\textwidth]{fig4b}
     \caption{
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS1.pdf}
+\includegraphics[width=\textwidth]{figS1.pdf}
     \caption{Ploidy and breeding system data according to three different classifications.
@@
 
-\begin{suppfigure} %figS2 all models
+\begin{suppfigure}% figS2 models
     \caption{ Twenty-nine models of diversification are proposed for the study of ploidy, breeding systems, and hidden states linked to the process of diversification.
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS3.pdf}
+\includegraphics[width=\textwidth]{figS3.pdf}
 \caption{Posterior distributions for each of the parameters in the ploidy only model (M1).
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS4.pdf}
+\includegraphics[width=\textwidth]{figS4.pdf}
 \caption{Posterior distributions for each of the parameters in the ploidy and hidden trait model (M4).
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS5.pdf}
+\includegraphics[width=\textwidth]{figS5.pdf}
 \caption{Posterior distributions for each of the parameters in the breeding system only model (M11).
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS6.pdf}
+\includegraphics[width=\textwidth]{figS6.pdf}
 \caption{Posterior distributions for each of the parameters in the breeding system and hidden trait model (M14).
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS7.pdf}
+\includegraphics[width=\textwidth]{figS7.pdf}
 \caption{Posterior distribution for each of the parameters in the ploidy and breeding system model (M16).
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS8.pdf}
+\includegraphics[width=\textwidth]{figS8.pdf}
 \caption{Posterior distribution for each of the parameters in the ploidy, breeding system, and hidden trait model (M19).
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS9.pdf}
+\includegraphics[width=\textwidth]{figS9.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under  the ploidy only model (M1).
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS10.pdf}
+\includegraphics[width=\textwidth]{figS12.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the ploidy and hidden states model (M4).} % XXX
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS11.pdf}
+\includegraphics[width=\textwidth]{figS11.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the  breeding system only model (M11).} % XXX
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS12.pdf}
+\includegraphics[width=\textwidth]{figS12.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the breeding system and hidden states model (M14.} % XXX
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS13.pdf}
+\includegraphics[width=\textwidth]{figS13.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the ploidy and breeding system model (M16).} % XXX
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS14.pdf}
+\includegraphics[width=\textwidth]{figS14.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the  ploidy, breeding systems, and hidden states model (M19).} % XXX
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS15.pdf}
+\includegraphics[width=\textwidth]{figS15.pdf}
 \caption{Testing the addition of breeding system to ploidy models.
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS16.pdf}
+\includegraphics[width=\textwidth]{figS16.pdf}
 \caption{Testing the addition of ploidy to breeding system models.
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS17.pdf}
+\includegraphics[width=\textwidth]{figS17.pdf}
 \caption{Effect of asymmetric rates in hidden models.
@@
 \begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS18.pdf}
+\includegraphics[width=\textwidth]{figS18.pdf}
 \caption{Posterior distributions for the net diversification rates of the preferred models with diploidization.
@@
 \end{suppfigure}
-
-% Supplementary TABLES
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|r|c|c|c|c|l|}
-\hline
-Model                           & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M7    & M8     & M9    & M10   & Evidence                                                                                     \\ \hline
-M6.
-D/P+$\delta$                & -1268.83                                                          & 55.41 & 42.78  & 53.37 & 52.85 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M6\end{tabular}            \\ \hline
-\textbf{M7.
-CID D/P+$\delta$}   & \textbf{-1212.42}                                                 &       & -12.62 & -2.04 & -2.04 & \begin{tabular}[c]{@{}l@{}}Model M7 moderately\\ preferred over M9 and M10\end{tabular}      \\ \hline
-M8.
-D/P+A/B $\delta$            & -1214.46                                                          &       &        & 10.58 & 10.07 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular} \\ \hline
-M9.
-D/P+A/B +$\delta$ asym      & -1214.46                                                          &       &        &       & -0.51 & No evidence                                                                                  \\ \hline
-M10.
-D/P+A/B +$\delta$ all asym & -1214.97                                                          &       &        &       &       &         \\ \hline
-\end{tabular}
-\caption{Bayes factors in log-scale of  ploidy only models with diplodization.
-Results indicate that a character independent model (M7) is strongly preferred over model M6 (BiSSE).
-Model M7 (bold) is also moderately preferred over any of the HiSSE models with asymmetric hidden rates (M9, M10).}
-\label{supptable:M6M10}
-\end{supptable}
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|r|c|c|c|c|l|}
-\hline
-Model                            & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M18   & M19   & M20   & Evidence                                                                                                   \\ \hline
-M16.
-ID/CP/CD                    & -1459.11                                                          & 45.11 & 65.91 & 63.98 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M16\end{tabular}                         \\ \hline
-M17.
-CID ID/CP/CD                & *                                                                 &       &       &       &                                                                                                            \\
-M18.
-ID/CP/CD+A/B                & -1414                                                             &       & 20.79 & 18.87 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}               \\ \hline
-\textbf{M19.
-ID/CP/CD +A/B asym} & \textbf{-1393.20}                                                 &       &       & -1.92 & \begin{tabular}[c]{@{}l@{}}Asymmetric hidden rates moderately\\ preferred over all asymmetric\end{tabular} \\ \hline
-M20.
-ID/CP/CD +A/B all asym      & -1393.12                                                          &       &       &       &   \\ \hline
-\end{tabular}
-\caption{Bayes factors of ploidy and breeding system without diploidization in log-scale.
-Results indicate that the model with asymmetric hidden rates (M19, bold) is  strongly preferred over M16 and M18 and moderately preferred over the MuHiSSe with all rates asymetric (M20).
-*Marginal log-likelihood for M17 could not be calculated within allotted computer time.}
-\label{supptable:M16M20}
-\end{supptable}
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|r|c|c|c|c|l|}
-\hline
-Model                                 & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M22   & M23    & M24   & M25   & Evidence                                                                                         \\ \hline
-M21.
-ID/CP/CD+$\delta$                & -1454.68                                                          & 55.48 & 46.031 & 67.94 & 65.15 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M21\end{tabular}               \\ \hline
-M22.
-CID ID/CP/CD+$\delta$            & -1399.201                                                         &       & -9.452 & 12.45 & 9.675 & \begin{tabular}[c]{@{}l@{}}Model M24 strongly preferred \\ over M22\end{tabular}                 \\ \hline
-M23.
-ID/CP/CD+A/B $\delta$            & -1408.65                                                          &       &        & 21.91 & 19.12 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}     \\ \hline
-\textbf{M24.
-ID/CP/CD +$\delta$ asym} & \textbf{-1386.74}                                                 &       &        &       & -2.78 & \begin{tabular}[c]{@{}l@{}}Asymmetric hidden rates \\ preferred over all asymmetric\end{tabular} \\ \hline
-M25.
-ID/CP/CD +$\delta$ all asym      & -1389.52                                                          &       &        &       &       &                      \\ \hline
-\end{tabular}
-\caption{Bayes factors of ploidy and breeding system with diploidization in log-scale.
-Results indicate that the MuHiSSE model with asymmetric hidden rates (M24, bold) is  strongly preferred over M21-M23 and moderately preferred over the MuHiSSe with all rates asymmetric (M25).}
-\label{supptable:M21M25}
-\end{supptable}
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-5pt}
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
-M1.
-D/P                                  & -1238.76                                                          & M1 vs.
-M4   & 60.47           & M4 (Strong)                                                          \\
-\textbf{M4.
-D/P+A/B asym}                & \textbf{-1223.28}                                                &             &                 &                                                                      \\ \hline
-M11.
-I/C                                 & -1309.07                                                          & M11 vs.
-M14 & 61.35           & M14 (Strong)                                                         \\
-\textbf{M14.
-I/C+A/B asym}               & \textbf{-1247.72}                                                 &             &                 &                                                                      \\ \hline
-M16.
-ID/CD/CP                            & -1459.11                                                          & M16 vs.
-M19 & 65.90           & M19 (Strong)                                                         \\
-\textbf{M19.
-ID/CD/CP+A/B asym}          & \textbf{-1393.20}                                                &             &                 &                                                                      \\ \hline
-M6.
-D/P+$\delta$                         & -1283.76                                                          & M6 vs.
-M9   & 69.3            & M9 (Strong)                                                          \\
-\textbf{M9.
-D/P+A/B+$\delta$ asym}       & \textbf{-1214.46}                                                 &             &                 &                                                                      \\ \hline
-M21.
-IC/CD/CP+$\delta$                   & -1454.68                                                          & M21 vs.
-M24 & 68.48           & M24 (Strong)                                                         \\
-\textbf{M24.
-IC/CD/CP+A/B+$\delta$ asym} & \textbf{-1386.20}                                                 &             &                 &                                                                      \\ \hline
-\end{tabular}
-\caption{Test for addition of hidden states in models via Bayes factors (in log-scale).
-Models with hidden states  (M4, M14, M19, M9, M24, bold) are strongly preferred over simpler models that do not include hidden s}
-\label{supptable:testaddhidden}
-\end{supptable}
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-5pt}
-\begin{tabular}{|l|c|c|c|l|}
-\hline
-Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
-M3. D/P+A/B                                 & -1234.52                     &    M3 vs. M4                 & 11.239                  & M4 (Strong)   \\
-\textbf{M4. D/P+ A/B asym}            & -1223.28                     &   M4 vs. M5              & -1.658                  & M4 (Moderate)   \\
-M5. D/P+A/B all asym                 & -1224.93                     &                     &                         &                         \\ \hline
-M13. I/C+ A/B                                & -1270.47                     &   M13 vs. M14                  & 22.75                   & M14 (Strong)   \\
-\textbf{M14. I/C+ A/B asym}            & \textbf{-1247.72}            & M14 vs. M15         & {0.05}           & No evidence             \\
-M15. I/C+ A/B all asym               & -1247.67                     &                     &                         &                         \\ \hline
-M18. IC/CD/CP+A/B                            & -1414.00                     &     M18 vs. M19                & 20.79                   & M19 (Strong)    \\
-\textbf{M19. IC/CD/CP+ A/B asym}       & \textbf{-1393.21}            & M19 vs. M20          & {-1.919}         & M19 (Moderate)   \\
-M20. IC/CD/CP+ A/B all asym           & -1395.129                    &                     &                         &                         \\ \hline
-M8. D/P +A/B+$\delta$                          & -1225.05                     &  M8 vs. M9                   & 10.58                   & M9 (Strong)    \\
-\textbf{M9. D/P+ A/B+ $\delta$ asym}      & \textbf{-1214.46}                     &    M9 vs. M10                 & -0.52                   & M10 (Moderate) \\
-M10. D/P+A/B+$\delta$ all asym        & -1214.98                     &                    &                         &                         \\ \hline
-M23. IC/CD/DP+A/B+$\delta$                 & -1408.65                     &     M23 vs M24                & 21.91                   & M24(Strong)     \\
-\textbf{M24. IC/CD/DP+A/B+$\delta$ asym} & \textbf{-1386.74}                     &   M24 vs M25              & -2.78                   & M24 (Moderate) \\
-M25. IC/CD/DP+A/B+ $\delta$ all asym     & -1389.52                     &                     &                         &                         \\ \hline
-\end{tabular}
-\caption{Test for asymmetry of the hidden trait transition rates via Bayes factors.
-For all models, asymmetric transition rates between hidden trait states are preferred over models with equal rates (bold).
-Adding more complexity by assuming asymmetry in all rates within both hidden states is not preferred over models with just asymmetry between hidden states.}
-\label{supptable:asymmetry}
-\end{supptable}
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
-M1.
-D/P                                  & -1238.76                                                          & M1 vs.
-M6   & 65.92           & M6 (Strong)                                                          \\
-\textbf{M6.
-D/P+$\delta$}                & \textbf{-1267.84}                                                 &             &                 &                                                                      \\ \hline
-M4.
-D/P+A/B asym                         & -1223.28                                                          & M4 vs.
-M9   & 8.81            & M9 (Moderate)                                                        \\
-\textbf{M9.
-D/P+A/B+$\delta$ asym}       & \textbf{-1214.46}                                                 &             &                 &                                                                      \\ \hline
-M16.
-ID/CD/CP                            & -1459.11                                                          & M16 vs.
-M21 & 4.41            & M21 (Moderate)                                                       \\
-\textbf{M21.
-ID/CD/CP+$\delta$}          & \textbf{-1454.68}                                                 &             &                 &                                                                      \\ \hline
-M19.
-IC/CD/CP +A/B asym                       & -1393.20                                                          & M19 vs.
-M24 & 6.46            & M24 (Moderate)                                                       \\
-\textbf{M24.
-IC/CD/CP+A/B+$\delta$ asym} & \textbf{-1386.20}                                                 &             &                 &                                                                      \\ \hline
-\end{tabular}
-\caption{Test for inclusion of a diploidization rate via Bayes factors. Models with diploidization are moderately preferred over models that do not include a diploidization rate (bold).}
-\label{supptable:testdiploidization}
-\end{supptable}

### 2019-08-31 basicdata/datawithphylo.R

@@ -1,0 +1,47 @@
+library("ggplot2")
+library("ggtree")
+library("treeio")
+library("ggnewscale")
+#library("RevGadgets")
+library("wesanderson")
+library("dyplr")
+
+cols<-c(wes_palette("Darjeeling1"),wes_palette("Darjeeling2"))
+cols2<- c("#7b3294","#c2a5cf","#a6dba0","#008837","#ffffbf")
+cols3<-c("#ece2f0", "#67a9cf","#e31a1c","#fd8d3c","#02818a","#014636")
+
+
+soltree<-read.nexus("~/Dropbox/solploidy/basicdata/fullmatchtree.nex")
+soldata<- read.csv("~/Dropbox/solploidy/basicdata/discretematrix.csv")
+dp<-data.frame(soldata[,2:4])
+rownames(dp)<-soltree$tip.label
+
+circ <- ggtree(soltree, layout = "circular")
+
+p1 <- gheatmap(circ, dp[, "Ploidy", drop=F], offset=.8, width=.1,
+               colnames_angle=90, colnames_offset_y = .25)
+p2 <- gheatmap(p1, dp[, "BS", drop=F], offset=5, width=.1,
+               colnames_angle=90, colnames_offset_y = .25)
+p3 <- gheatmap(p2, dp[, "Ploidybs", drop=F], offset=10, width=.1,
+               colnames_angle=90, colnames_offset_y = .25)
+
+col <- c(cols2[2],cols[7],cols2[1],cols[1],cols[9],cols[9],cols2[3],cols2[4],"yellow",cols2[3],cols[6],cols[6],"grey")
+names(col) = c("C","CD","Compatible","CP","D","Diploid","I","ID","IDCP","Incompatible","P","Polyploid","Unknown")
+#D = cols[9]
+#%P=cols[6]
+
+#CD cols[7]
+#CP cols[1]
+
+#Compatible=cols2[1],
+#C= cols2[2]
+#Incompatible= cols2[3]
+#I cols2[3])
+#ID= cols2[4]
+#IDCP= "yellow"
+#Diploid= cols[9]
+#Polyploid= cols[6]
+
+pp <- p3 + scale_fill_manual(values=col)
+require(phytools)
+plotTree(soltree,type="fan", cex=0.18,fsize=0.19,label.offset= 2)

### 2019-08-31 finalmanuscript/methods.tex
final manuscript for submission v2

@@
 \end{figure}
-
-% check values of delta and rho D/P+A/B with(out) delta
-% i.e. parameter values in figure S1 & S4 vs Table S2
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS1.pdf}
-    \caption{Ploidy and breeding system data according to three different classifications.
-For ploidy only models, classifications with states $D$ and $P$ were used (inner circle).
-For breeding system models classifications with states $I$ and $C$ were used (middle circle).
-For ploidy and breeding system models classifications using $ID, CD, CP$ were used (outer circle).
-Data with missing information in one of the traits was classified simultaneously as two possible states, for example, diploids without breeding system $?D$ were classified as $(CD, CP)$).}
-    \label{fig:allmodels}
-\end{suppfigure}
-
-\begin{suppfigure} %figS2 all models
-    \caption{ Twenty-nine models of diversification are proposed for the study of ploidy, breeding systems, and hidden states linked to the process of diversification.
-We divide the models by the type of focal trait studied (ploidy only, breeding system only, or ploidy and breeding system).
-The contributions of the focal trait to the diversification process can be measured by comparing the models in each of the columns.
-That is, the focal trait only models assume that speciation and extinction rates are only linked to the trait itself, the hidden trait only models assume that the diversification rates are  linked to unknown factors but not the trait of interest, and the focal trait with hidden trait models assume that both the focal trait and unknown factors are contributing to diversification.}
-    \label{fig:allmodels}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS3.pdf}
-\caption{Posterior distributions for each of the parameters in the ploidy only model (M1).
-Red color represents diploid state $D$ and blue color represents polyploid state $P$.
-(A)Speciation rates.
-(B) Extinction rates.
-(C) Net diversification rates (speciation minus extinction from panels A and B).
-(D) Relative extinction rates (extinction divided by speciation from panels A and B).
-(E) Polyploidization rate ($\rho$).} % XXX
-\label{suppfigure:DPnodip}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS4.pdf}
-\caption{Posterior distributions for each of the parameters in the ploidy and hidden trait model (M4).
-Red color represents diploid state $D$ and blue color represents polyploid state $P$.
-Dark colors represent hidden state $A$ and light colors hidden state $B$.
-(A) Speciation rates.
-(B) Extinction rates.
-(C) Net diversification rates (speciation minus extinction from panels A and B).
-(D) Relative extinction rates (extinction divided by speciation from panels A and B).
-(E) Polyploidization rate ($\rho$).
-(F) Transition rates between hidden states ( $\alpha$ and $\beta$).} % XXX
-\label{suppfigure:DPnodipAB}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS5.pdf}
-\caption{Posterior distributions for each of the parameters in the breeding system only model (M11).
-Green color represents self-incompatible state $I$ and purple color represents self-compatible state $C$.
-(A) Speciation rates.
-(B) Extinction rates.
-(C) Net diversification rates (speciation minus extinction from panels A and B).
-(D) Relative extinction rates (extinction divided by speciation from panels A and B).
-(E) Self-incompatible to self-compatible transition rate ($q_IC$).} % XXX
-\label{suppfigure:IC}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS6.pdf}
-\caption{Posterior distributions for each of the parameters in the breeding system and hidden trait model (M14).
-Green color represents self-incompatible state $I$ and purple color represents self-compatible state $A$.
-Dark colors represent hidden state $A$ and light colors hidden state $B$.
-(A) Speciation rates.
-(B) Extinction rates.
-(C) Net diversification rates (speciation minus extinction from panels A and B).
-(D) Relative extinction rates (extinction divided by speciation from panels A and B).
-(E) Self-incompatible to self-compatible transition rate ($q_IC$).
-(F) Transition rates between hidden states ( $\alpha$ and $\beta$).} % XXX
-\label{suppfigure:ICAB}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS7.pdf}
-\caption{Posterior distribution for each of the parameters in the ploidy and breeding system model (M16).
-Green color represents self-incompatible diploid state $ID$, blue color is the self-compatible and diploid state $CD$ and pink represents self-compatible polyploid state $CP$.
-(A) Speciation rates.
-(B) Extinction rates.
-(C) Net diversification rates (speciation minus extinction from panels A and B).
-(D) Relative extinction rates (extinction divided by speciation from panels A and B).
-(E) Self-incompatible to self-compatible transition rate ($q_IC$, yellow), polyploidization rate from self-incompatible ($\rho_I,$ light pink), and polyploidization from self-compatible ($\rho_c$, orange).} % XXX
-\label{suppfigure:IDCDCPnodip}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS8.pdf}
-\caption{Posterior distribution for each of the parameters in the ploidy, breeding system, and hidden trait model (M19).
-Green color represents self-incompatible diploid state $ID$, blue color is the self-compatible and diploid state $CD$ and pink represents self-compatible polyploid state $CP$.
-Dark colors represent hidden state $A$ and light colors hidden state $B$.
-(A) Speciation rates.
-(B) Extinction rates.
-(C) Net diversification rates (speciation minus extinction from panels A and B).
-(D) Relative extinction rates (extinction divided by speciation from panels A and B).
-(E) Self-incompatible to self-compatible transition rate ($q_IC$, yellow), polyploidization rate from self-incompatible ($\rho_I,$ light pink), and polyploidization from self-compatible ($\rho_c$, orange).
-(F) Transition rates between hidden states ( $\alpha$ and $\beta$).}% XXX
-\label{suppfigure:IDCDCPnodipAB}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS9.pdf}
-\caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under  the ploidy only model (M1).
-} % XXX
-\label{suppfigure:DPnodipasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS10.pdf}
-\caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the ploidy and hidden states model (M4).} % XXX
-\label{suppfigure:DPnodipABasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS11.pdf}
-\caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the  breeding system only model (M11).} % XXX
-\label{suppfigure:ICasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS12.pdf}
-\caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the breeding system and hidden states model (M14.} % XXX
-\label{suppfigure:ICABasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS13.pdf}
-\caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the ploidy and breeding system model (M16).} % XXX
-\label{suppfigure:IDCDCPnodipasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS14.pdf}
-\caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the  ploidy, breeding systems, and hidden states model (M19).} % XXX
-\label{suppfigure:IDCDCPnodipABasr}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS15.pdf}
-\caption{Testing the addition of breeding system to ploidy models.
-(A) Ploidy only model (M1) where data enter as binary $D$ and $P$.
-(B) Lumped model for ploidy (M26) where data are the three-state values ($ID,CP,CD$) but results are equivalent to model M1.
-(C) Ploidy and breeding system model (M16) where  data enter as the three-state values.
-Models M26 and M16 are comparable whereas M1 and M16 are not.
-(D) Ploidy and hidden state model (M3) where data enter as binary $D$ and $P$.
-(E) Lumped model for ploidy and hidden state (M27) where data are the three-state values ($ID,CP,CD$) but results are equivalent to model M3.
-(F) Ploidy, breeding system, and hidden state model (M18) where  data enter as the three-state values.
-Models M27 and M18 are comparable whereas M3 and M18 are not.
-Bayes factors comparing the models are shown in \cref{table:lumped}.} % XXX
-\label{suppfigure:lumpedDP}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS16.pdf}
-\caption{Testing the addition of ploidy to breeding system models.
-(A) Breeding system only model (M11) where data enter as binary $I$ and $C$.
-(B) Lumped model for breeding system (M28) where data are the three-state values ($ID,CP,CD$) but results are equivalent to model M11.
-(C) Ploidy and breeding system model (M16) where  data enter as the three-state values.
-Models M28 and M16 are comparable whereas M11 and M16 are not.
-(D) Breeding system and hidden state model (M13) where data enter as binary $I$ and $C$.
-(E) Lumped model for breeding system and hidden state (M29) where data are the three-state values ($ID,CP,CD$) but results are equivalent to model M13.
-(F) Ploidy, breeding system, and hidden state model (M18) where  data enter as the three-state values.
-Models M29 and M18 are comparable whereas M13 and M18 are not.
-Bayes factors comparing the models are shown in \cref{table:lumped}.} % XXX
-\label{suppfigure:lumpedIC}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS17.pdf}
-\caption{Effect of asymmetric rates in hidden models.
-First column models M3, M13, and M18 assume that the rates between hidden states are equal.
-The models in the second column (M4, M14, M19) assume that the rates between hidden states are different .
-Column three models assumes that the rates between hidden state are asymmetric and that the transition rates within each hidden states are also different.
-Bayes factors in \cref{supptable:asymmetry} strongly preferred models with asymmetric rates between states (second column) over models with equal rates in hidden states (first column).
-Models in the second column are moderately or equally prefer to more complex models in column 3.
-} % XXX
-\label{suppfigure:asymmetric}
-\end{suppfigure}
-
-\begin{suppfigure}
-%\includegraphics[width=\textwidth]{figS18.pdf}
-\caption{Posterior distributions for the net diversification rates of the preferred models with diploidization.
-Red color represents diploid state $D$, blue color represents polyploid state $P$, green color represents self-incompatible $I$, purple color represents self-compatible $C$,  dark colors represent hidden state $A$ and light colors hidden state $B$.
-(A) Ploidy only model M6.
-(B) Ploidy and hidden states model M9 (C) Breeding systems only model M11.
-(D) Breeding systems and hidden state model M14.
-(E) Ploidy and breeding systems model M21.
-(F) Ploidy, breeding systems, and hidden states model M24.} % XXX
-\label{suppfigure:alldip}
-\end{suppfigure}
-
-% Supplementary TABLES
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|r|c|c|c|c|l|}
-\hline
-Model                           & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M7    & M8     & M9    & M10   & Evidence                                                                                     \\ \hline
-M6.
-D/P+$\delta$                & -1268.83                                                          & 55.41 & 42.78  & 53.37 & 52.85 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M6\end{tabular}            \\ \hline
-\textbf{M7.
-CID D/P+$\delta$}   & \textbf{-1212.42}                                                 &       & -12.62 & -2.04 & -2.04 & \begin{tabular}[c]{@{}l@{}}Model M7 moderately\\ preferred over M9 and M10\end{tabular}      \\ \hline
-M8.
-D/P+A/B $\delta$            & -1214.46                                                          &       &        & 10.58 & 10.07 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular} \\ \hline
-M9.
-D/P+A/B +$\delta$ asym      & -1214.46                                                          &       &        &       & -0.51 & No evidence                                                                                  \\ \hline
-M10.
-D/P+A/B +$\delta$ all asym & -1214.97                                                          &       &        &       &       &         \\ \hline
-\end{tabular}
-\caption{Bayes factors in log-scale of  ploidy only models with diplodization.
-Results indicate that a character independent model (M7) is strongly preferred over model M6 (BiSSE).
-Model M7 (bold) is also moderately preferred over any of the HiSSE models with asymmetric hidden rates (M9, M10).}
-\label{supptable:M6M10}
-\end{supptable}
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|r|c|c|c|c|l|}
-\hline
-Model                            & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M18   & M19   & M20   & Evidence                                                                                                   \\ \hline
-M16.
-ID/CP/CD                    & -1459.11                                                          & 45.11 & 65.91 & 63.98 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M16\end{tabular}                         \\ \hline
-M17.
-CID ID/CP/CD                & *                                                                 &       &       &       &                                                                                                            \\
-M18.
-ID/CP/CD+A/B                & -1414                                                             &       & 20.79 & 18.87 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}               \\ \hline
-\textbf{M19.
-ID/CP/CD +A/B asym} & \textbf{-1393.20}                                                 &       &       & -1.92 & \begin{tabular}[c]{@{}l@{}}Asymmetric hidden rates moderately\\ preferred over all asymmetric\end{tabular} \\ \hline
-M20.
-ID/CP/CD +A/B all asym      & -1393.12                                                          &       &       &       &   \\ \hline
-\end{tabular}
-\caption{Bayes factors of ploidy and breeding system without diploidization in log-scale.
-Results indicate that the model with asymmetric hidden rates (M19, bold) is  strongly preferred over M16 and M18 and moderately preferred over the MuHiSSe with all rates asymetric (M20).
-*Marginal log-likelihood for M17 could not be calculated within allotted computer time.}
-\label{supptable:M16M20}
-\end{supptable}
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|r|c|c|c|c|l|}
-\hline
-Model                                 & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M22   & M23    & M24   & M25   & Evidence                                                                                         \\ \hline
-M21.
-ID/CP/CD+$\delta$                & -1454.68                                                          & 55.48 & 46.031 & 67.94 & 65.15 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M21\end{tabular}               \\ \hline
-M22.
-CID ID/CP/CD+$\delta$            & -1399.201                                                         &       & -9.452 & 12.45 & 9.675 & \begin{tabular}[c]{@{}l@{}}Model M24 strongly preferred \\ over M22\end{tabular}                 \\ \hline
-M23.
-ID/CP/CD+A/B $\delta$            & -1408.65                                                          &       &        & 21.91 & 19.12 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}     \\ \hline
-\textbf{M24.
-ID/CP/CD +$\delta$ asym} & \textbf{-1386.74}                                                 &       &        &       & -2.78 & \begin{tabular}[c]{@{}l@{}}Asymmetric hidden rates \\ preferred over all asymmetric\end{tabular} \\ \hline
-M25.
-ID/CP/CD +$\delta$ all asym      & -1389.52                                                          &       &        &       &       &                      \\ \hline
-\end{tabular}
-\caption{Bayes factors of ploidy and breeding system with diploidization in log-scale.
-Results indicate that the MuHiSSE model with asymmetric hidden rates (M24, bold) is  strongly preferred over M21-M23 and moderately preferred over the MuHiSSe with all rates asymmetric (M25).}
-\label{supptable:M21M25}
-\end{supptable}
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-5pt}
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
-M1.
-D/P                                  & -1238.76                                                          & M1 vs.
-M4   & 60.47           & M4 (Strong)                                                          \\
-\textbf{M4.
-D/P+A/B asym}                & \textbf{-1223.28}                                                &             &                 &                                                                      \\ \hline
-M11.
-I/C                                 & -1309.07                                                          & M11 vs.
-M14 & 61.35           & M14 (Strong)                                                         \\
-\textbf{M14.
-I/C+A/B asym}               & \textbf{-1247.72}                                                 &             &                 &                                                                      \\ \hline
-M16.
-ID/CD/CP                            & -1459.11                                                          & M16 vs.
-M19 & 65.90           & M19 (Strong)                                                         \\
-\textbf{M19.
-ID/CD/CP+A/B asym}          & \textbf{-1393.20}                                                &             &                 &                                                                      \\ \hline
-M6.
-D/P+$\delta$                         & -1283.76                                                          & M6 vs.
-M9   & 69.3            & M9 (Strong)                                                          \\
-\textbf{M9.
-D/P+A/B+$\delta$ asym}       & \textbf{-1214.46}                                                 &             &                 &                                                                      \\ \hline
-M21.
-IC/CD/CP+$\delta$                   & -1454.68                                                          & M21 vs.
-M24 & 68.48           & M24 (Strong)                                                         \\
-\textbf{M24.
-IC/CD/CP+A/B+$\delta$ asym} & \textbf{-1386.20}                                                 &             &                 &                                                                      \\ \hline
-\end{tabular}
-\caption{Test for addition of hidden states in models via Bayes factors (in log-scale).
-Models with hidden states  (M4, M14, M19, M9, M24, bold) are strongly preferred over simpler models that do not include hidden s}
-\label{supptable:testaddhidden}
-\end{supptable}
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-5pt}
-\begin{tabular}{|l|c|c|c|l|}
-\hline
-Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
-M3. D/P+A/B                                 & -1234.52                     &    M3 vs. M4                 & 11.239                  & M4 (Strong)   \\
-\textbf{M4. D/P+ A/B asym}            & -1223.28                     &   M4 vs. M5              & -1.658                  & M4 (Moderate)   \\
-M5. D/P+A/B all asym                 & -1224.93                     &                     &                         &                         \\ \hline
-M13. I/C+ A/B                                & -1270.47                     &   M13 vs. M14                  & 22.75                   & M14 (Strong)   \\
-\textbf{M14. I/C+ A/B asym}            & \textbf{-1247.72}            & M14 vs. M15         & {0.05}           & No evidence             \\
-M15. I/C+ A/B all asym               & -1247.67                     &                     &                         &                         \\ \hline
-M18. IC/CD/CP+A/B                            & -1414.00                     &     M18 vs. M19                & 20.79                   & M19 (Strong)    \\
-\textbf{M19. IC/CD/CP+ A/B asym}       & \textbf{-1393.21}            & M19 vs. M20          & {-1.919}         & M19 (Moderate)   \\
-M20. IC/CD/CP+ A/B all asym           & -1395.129                    &                     &                         &                         \\ \hline
-M8. D/P +A/B+$\delta$                          & -1225.05                     &  M8 vs. M9                   & 10.58                   & M9 (Strong)    \\
-\textbf{M9. D/P+ A/B+ $\delta$ asym}      & \textbf{-1214.46}                     &    M9 vs. M10                 & -0.52                   & M10 (Moderate) \\
-M10. D/P+A/B+$\delta$ all asym        & -1214.98                     &                    &                         &                         \\ \hline
-M23. IC/CD/DP+A/B+$\delta$                 & -1408.65                     &     M23 vs M24                & 21.91                   & M24(Strong)     \\
-\textbf{M24. IC/CD/DP+A/B+$\delta$ asym} & \textbf{-1386.74}                     &   M24 vs M25              & -2.78                   & M24 (Moderate) \\
-M25. IC/CD/DP+A/B+ $\delta$ all asym     & -1389.52                     &                     &                         &                         \\ \hline
-\end{tabular}
-\caption{Test for asymmetry of the hidden trait transition rates via Bayes factors.
-For all models, asymmetric transition rates between hidden trait states are preferred over models with equal rates (bold).
-Adding more complexity by assuming asymmetry in all rates within both hidden states is not preferred over models with just asymmetry between hidden states.}
-\label{supptable:asymmetry}
-\end{supptable}
-
-\begin{supptable}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Model                                    & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=log(BF(M1,M2) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
-M1.
-D/P                                  & -1238.76                                                          & M1 vs.
-M6   & 65.92           & M6 (Strong)                                                          \\
-\textbf{M6.
-D/P+$\delta$}                & \textbf{-1267.84}                                                 &             &                 &                                                                      \\ \hline
-M4.
-D/P+A/B asym                         & -1223.28                                                          & M4 vs.
-M9   & 8.81            & M9 (Moderate)                                                        \\
-\textbf{M9.
-D/P+A/B+$\delta$ asym}       & \textbf{-1214.46}                                                 &             &                 &                                                                      \\ \hline
-M16.
-ID/CD/CP                            & -1459.11                                                          & M16 vs.
-M21 & 4.41            & M21 (Moderate)                                                       \\
-\textbf{M21.
-ID/CD/CP+$\delta$}          & \textbf{-1454.68}                                                 &             &                 &                                                                      \\ \hline
-M19.
-IC/CD/CP +A/B asym                       & -1393.20                                                          & M19 vs.
-M24 & 6.46            & M24 (Moderate)                                                       \\
-\textbf{M24.
-IC/CD/CP+A/B+$\delta$ asym} & \textbf{-1386.20}                                                 &             &                 &                                                                      \\ \hline
-\end{tabular}
-\caption{Test for inclusion of a diploidization rate via Bayes factors. Models with diploidization are moderately preferred over models that do not include a diploidization rate (bold).}
-\label{supptable:testdiploidization}
-\end{supptable}

### 2019-08-31 manuscript/methods.tex
final manuscript for submission v2

@@
 \centering
-\includegraphics[width=0.5\textwidth]{states.pdf}
+\includegraphics[width=0.5\textwidth]{fig1.pdf}
 \caption{
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{alldiversificationasymhidden.pdf} 
+\includegraphics[width=\textwidth]{fig2.pdf} 
 \caption{Net diversification rates for SSE models of focal traits with or without hidden state. 
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{lumped.pdf} %lumped.pdf
+\includegraphics[width=\textwidth]{fig3.pdf} %lumped.pdf
 \caption{Comparing two and three-state models using lumped models. (A) Ploidy model only (M1) where data enter as binary $D$ and $P$. 
@@
 \begin{figure}
-    \centering 
-    \includegraphics[width=0.5\textwidth]{pathstates2} \\ [40pt] 
-        \includegraphics[width=0.9\textwidth]{pathways}
+   % \centering 
+    \includegraphics[width=0.5\textwidth]{fig4a} \\ [40pt] 
+        \includegraphics[width=0.9\textwidth]{fig4b}
     \caption{
@@
 % i.e. parameter values in figure S1 & S4 vs Table S2
+\begin{suppfigure}
+\includegraphics[width=\textwidth]{figS1.pdf}
+    \caption{Ploidy and breeding system data according to three different classifications.
+For ploidy only models, classifications with states $D$ and $P$ were used (inner circle).
+For breeding system models classifications with states $I$ and $C$ were used (middle circle).
+For ploidy and breeding system models classifications using $ID, CD, CP$ were used (outer circle).
+Data with missing information in one of the traits was classified simultaneously as two possible states, for example, diploids without breeding system $?D$ were classified as $(CD, CP)$).}
+    \label{fig:allmodels}
+\end{suppfigure}
 
-\begin{suppfigure}
+\begin{suppfigure} %figS2 all models
     \caption{ Twenty-nine models of diversification are proposed for the study of ploidy, breeding systems, and hidden states linked to the process of diversification.
@@
 The contributions of the focal trait to the diversification process can be measured by comparing the models in each of the columns.
-That is, the focal trait only models assume that speciation and extinction rates are only linked to the trait itself, the hidden trait only models assume that the diversification rates are  linked to unknown factors but not the trait of interest, and the focal trait with hidden trait models assume that both the focal trait and unknown factors are contributing to diversification.}
+That is, the focal trait only models assume that speciation and extinction rates are only linked to the trait itself, the hidden trait only models assume that the diversification rates are  linked to unknown factors but not the trait of interest, and the focal trait with hidden trait models assume that both the focal trait and unknown factors are contributing to diversification.
+(large format figure)}
     \label{fig:allmodels}
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{bisseDPnodipposteriordist.pdf}
+\includegraphics[width=\textwidth]{figS3.pdf}
 \caption{Posterior distributions for each of the parameters in the ploidy only model (M1).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{hisseDPnodipasymposteriordist.pdf}
+\includegraphics[width=\textwidth]{figS4.pdf}
 \caption{Posterior distributions for each of the parameters in the ploidy and hidden trait model (M4).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{bisseSIposteriordist.pdf}
+\includegraphics[width=\textwidth]{figS5.pdf}
 \caption{Posterior distributions for each of the parameters in the breeding system only model (M11).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{hisseSIasymposteriordist.pdf}
+\includegraphics[width=\textwidth]{figS6.pdf}
 \caption{Posterior distributions for each of the parameters in the breeding system and hidden trait model (M14).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{musseDPSInodipposteriordist.pdf}
+\includegraphics[width=\textwidth]{figS7.pdf}
 \caption{Posterior distribution for each of the parameters in the ploidy and breeding system model (M16).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{muhissenodipasymposteriordist.pdf}
+\includegraphics[width=\textwidth]{figS8.pdf}
 \caption{Posterior distribution for each of the parameters in the ploidy, breeding system, and hidden trait model (M19).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDP.pdf}
+\includegraphics[width=\textwidth]{figS9.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under  the ploidy only model (M1).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrDPAB.pdf}
+\includegraphics[width=\textwidth]{figS10.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the ploidy and hidden states model (M4).} % XXX
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIC.pdf}
+\includegraphics[width=\textwidth]{figS11.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the  breeding system only model (M11).} % XXX
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrICAB.pdf}
+\includegraphics[width=\textwidth]{figS12.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the breeding system and hidden states model (M14.} % XXX
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCP.pdf}
+\includegraphics[width=\textwidth]{figS13.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the ploidy and breeding system model (M16).} % XXX
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{asrIDCDCPAB.pdf}
+\includegraphics[width=\textwidth]{figS14.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the  ploidy, breeding systems, and hidden states model (M19).} % XXX
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{comparablesDP.pdf}
+\includegraphics[width=\textwidth]{figS15.pdf}
 \caption{Testing the addition of breeding system to ploidy models.
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{comparablesIC.pdf}
+\includegraphics[width=\textwidth]{figS16.pdf}
 \caption{Testing the addition of ploidy to breeding system models.
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{effectofasymmetry.pdf}
+\includegraphics[width=\textwidth]{figS17.pdf}
 \caption{Effect of asymmetric rates in hidden models.
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{alldiversificationasymhiddendip.pdf}
+\includegraphics[width=\textwidth]{figS18.pdf}
 \caption{Posterior distributions for the net diversification rates of the preferred models with diploidization.

### 2019-08-31 manuscript/supplement.tex
final manuscript for submission v2

--- 2019-08-31 manuscript/methods.tex
@@
-\section{Methods}
-
-\subsection{Data}
-
-%B: IMO: The data subsection would be better organized with parts, (1) ploidy, (2) breeding system, (3) tree, (4) models, (5) model selection and statistical inference.
-% E: I changed the subsectioning, but to a different arrangement: (1) Data, (2) Models, (3) Statistical inference.  With subsubsections as necessary.
-
-Chromosome number data were obtained for all Solanaceae taxa in the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were curated semi-automatically using the \mbox{CCDBcurator} R package \citep{rivero_2019}.
-CCDB contains records from original sources that have multiple complex symbol patterns denoting multivalence, or irregularites of chromosome counts.
-After a first round of automatic cleaning, we examined results by hand and corrected records as necessary.
-Our hand-curated records were also contrasted against the ploidy dataset from \citet{robertson_2011}, original references therein, and against ploidy data in the C-value DNA dataset from \citet{bennett_2005}.
-By comparing three different sources of information, we were able to code taxa as diploid, $D$, or polyploid, $P$.
-For the majority of species, ploidy was assigned according to information from the original publications included in the  C-value DNA dataset \citep{bennett_2005}.
-For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus/family, or based on SI/SC classification.
-For example, \textit{Solanum betaceum} did not have information about ploidy level, but it has 2n=24 chromosomes, and  $x=12$ is the base chromosome number of the genus \textit{Solanum} \citep{olmstead_2007}, so we assigned \textit{S.~betaceum} as diploid.
-Additionally, because of the absence of SI polyploids (explained above and below), species known to be SI could be scored as diploid.
-Species with more than one ploidy level were assigned the most frequent ploidy level recorded or the smallest ploidy in case of frequency ties.
-
-Breeding system states were scored as self-incompatible, $I$, or self-compatible, $C$, based on results curated from the literature  \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012} and original experimental crosses (B.I.
-and E.E.G., unpub.
-data).
-%IM suggestion %B:  I returned the aside about original data.
-I don't think Itay knows that we performed original crosses (many), and they are incorporated in the data, without being present in "the literature"
-%was originally: Breeding system states were scored as self-incompatible, $I$, or self-compatible, $C$, based on results curated from the literature and original experimental crosses \citep[as compiled in][]{igic_2006, goldberg_2010, robertson_2011, goldberg_2012}.
-Most species could unambiguously be coded as either $I$ or $C$ \citep{raduski_2012}.
-Following previous work, we coded any species with a functional SI system as $I$, even if SC or dioecy was also reported.
-Dioecious species without functional SI were coded as $C$.
-% E: I am trying to stick with I/C when talking about the state values, but using SI/SC when talking about the meaning of the trait (because this what people use normally). 
-%R- Okay- I was getting confused as well thanks for clarifying
-
-Resolution of taxonomic synonymy followed Solanaceae Source \citep{solsource}. 
-Hybrids and cultivars were excluded because ploidy and breeding system can be affected by artificial selection during domestication.
-Following the reasoning outlined in \citet{robertson_2011}, we closely examined the few species for which the merged ploidy and breeding system data indicated the presence of self-incompatible polyploids.
-Although SI populations frequently contain some SC individuals, and diploid populations frequently contain some polyploid individuals, in no case did we find convincing data for a naturally occurring SI polyploid population  \citep[discussed in][]{robertson_2011}.
-%B rm: The single instance of an SI polyploid individual appears to be an allopentaploid hybrid of \textit{Solanum oplocense} Hawkes x \textit{Solanum gourlayii} Hawkes, reported by \citet{camadro_1981}.
-% Under exceedingly rare circumstances, it is possible for polyploids containing multiple copies of S-loci to remain SI, so long as they express a single allele at the S-locus \citep[discussed in][]{robertson_2011}.
-Because of the resulting absence of polyploid SI populations, as well as the functional explanation for polyploidy disabling gametophytic SI systems with non-self recognition (see the Introduction), we consider only three observed character states: self-incompatible diploids $(ID)$, self-compatible diploids $(CD)$, and self-compatible polyploids $(CP)$.
-
-Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 651 species with ploidy and/or breeding system information on the tree.
-% Fixed!
-Of these, 368 had information for both states.
-The number of species in each combination of states is summarized in \cref{figure:stateclassifications}A and in \cref{suppfigure:stateclassifications}.
-We retained all 651 species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
-A total of 372 taxa without any information about breeding system or ploidy were excluded.
-%Tips without trait data might not improve point estimation of diversification parameters while increasing uncertainty in the estimates of trait linked diversification
-% R-This is very important but I have removed it because I think it is a full on paper by its own
-%Including this many more species would have prohibitively slowed our analyses, especially those implementing the most complex models.
-
-The Supplementary Information contains citations for the numerous original data sources. % FIXME (and say here how many refs?) %B: no need for an additional counting and updating headache?
-% R- There is few data that I added so I can find those references and send them to you, so it would be simply the same references that you previously had plus another 20-25 I think
-The Dryad archive contains the data and tree files used for analyses. % FIXME
-% \url{https://github.com/roszenil/solploidy}  E: we should use dryad instead, because github repos are not guaranteed to last
-%This fixme happens after publications
-
-\subsection{Models}
-
-In order to test our hypotheses about lineage diversification and trait macroevolution, we fit twenty-nine state-dependent speciation and extinction models \citep[BiSSE, MuSSE, HiSSE;][]{maddison_2007, fitzjohn_2012, beaulieu_2016}.
-SSE models contain parameters that describe per-lineage rates of speciation and extinction, specific to each character state (denoted $\lambda$ and $\mu$, respectively, with subscripts to indicate the state), along with rates of transitions between states (denoted $\rho$ for polyploidization, $\delta$ for diploidization, and $q_{IC}$ for loss of self-incompatibility).
-The full set of models and all their rate parameters are detailed in \cref{fig:allmodels}.
-Here, we summarize how each model allows us to assess whether diversification is best explained by variation in ploidy, breeding system, their combination, or some unknown factor.
-
-\subsubsection{Ploidy and diversification}
-
-We first employed a model (labeled M1), previously used by \citet{mayrose_2011}, with each species classified as diploid $(D)$ or polyploid $(P)$.
-Although this model can be powerful in studies of trait evolution, it is prone to incorrectly reporting that a trait is associated with diversification differences \citep{maddison_2015, rabosky_2015}.
-We therefore define several models that incorporate additional forms of diversification rate heterogeneity.
-The second ploidy model (M2) includes a binary hidden trait that subdivides each observed state.
-In this trait-independent model known as CID \citep{beaulieu_2016}, hidden traits can affect diversification but the observed traits do not.
-Comparing M1 and M2 allows us to test whether diversification rate heterogeneity is better explained by ploidy or by some unknown factor.
-
-We fit three models in which both ploidy and a hidden trait could influence diversification (M3--M5).
-These models differ in whether transitions between the hidden states are symmetric (M3) or asymmetric(M4), and whether the polyploidization rate depends on the hidden state(M5). %WF suggestion
-Comparing M1--M5 (as shown in \cref{table:bayesfactors}) allows us to test whether ploidy is associated with diversification differences on top of the differences potentially explained by an unknown factor.
-We further fit the analogues of these five models but including a rate parameter $\delta$ for transitions from polyploid to diploid (M6--M10).
-These comparisons allow us to assess whether our conclusions about ploidy and diversification are robust to the possibility of diploidization.
-
-\subsubsection{Breeding system and diversification}
-
-We propose five breeding system models following the same logic as the ploidy models above.
-Under the simplest breeding system and diversification model (M11), species are classified as self-incompatible $(I)$ or self-compatible $(C)$.
-This is the same model as in the analysis presented in \citet{goldberg_2010} but with an updated phylogeny \citep{sarkinen_2013} and a larger aggregated dataset.
-We then add models to allow diversification to be influenced by only a hidden trait (M12), or by both breeding system and a hidden trait (M13--M15, with varying degrees of complexity in the hidden trait transitions \cref{table:M11M15}).
-Similar models were used by \citet{freyman_2019} to study diversification in Onagraceae.
-
-Self-incompatibility is homologous in all Solanaceae species in which S-alleles have been cloned and controlled crosses performed.
-All species sampled to date possess a non-self recognition, RNase-based gametophytic self-incompatibility \citep[shared even with other euasterid families;][]{ramanauskas_2017}.
-Furthermore, species that are distantly related within this family carry closely-related alleles, with deep trans-specific polymorphism at the locus that controls the SI response \citep{ioerger_1990, igic_2006}.
-Thus, there is strong evidence in Solanaceae that the $I$ state is ancestral in the family, and that the SI mechanism was not regained.
-For all breeding system models, we estimated a transition rate from $I$ to $C$  but not the reverse ($q_{CI}=0$).
-% This is an important assumption because ancestral state reconstruction in the models might differ when polyploidy and breeding system are analyzed as independent in the Solanaceae tree.
-%
-%B: What happens if SC->SI is not fixed? Asking for a friend.  E: Surprisingly, Rosana reports, results are not turned on their head. 
-%B: Whew, I was worried.
-Er, "friend" says thanks.
-R- Your welcome, by mistake I did that model with reversibility from C to I.
-I didn't run it for as long as the rest but preliminary ancestral reconstructions seem almost identical when you add that parameter than when you don't.
-
-\subsubsection{Ploidy, breeding system, and diversification}
-
-Ploidy and breeding system might influence lineage diversification individually, but these two traits also have an intricate association (discussed in the Introduction).
-Therefore, we considered several multi-state models that investigate the contribution of both traits and the allowable transitions between them.
-
-The simplest model (M16) classifies each species as either SI diploid $(ID)$, SC diploid $(CD)$, or SC polyploid $(CP)$; recall that SI polyploids do not occur.
-Each of these states may again be associated with different rates of speciation and extinction, and the allowable transitions are loss of SI within the diploid state (from $ID$ to $CD$), loss of SI via polyploidization (from $ID$ to $CP$), and polyploidization while SC (from $CD$ to $CP$).
-% The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
-% E: Can we call it something other than q_IC?  It's a bit confusing to have the same parameter mean different things in different models.
-% R: In the lumped models I call them something different q_0 and rho_0 because they are different parameters but I don't think we should explain.
-As for the previous models of only one trait, we then allow diversification to be influenced by only a hidden trait (M17), or by ploidy, breeding system, and a hidden trait (M18--M20) with varying degrees of complexity in the hidden trait transitions
-\citep[similar to][]{caetano_2018, huang_2018}. %IM suggestion
-We also fit the analogous models but allowing for diploidization (M21--M25).
-
-\subsubsection{Lumped models}
-
-The models described so far allow us to assess the contributions of our two focal characters---ploidy and breeding system---to lineage diversification, but they do not reveal whether it is valuable to include both characters in the analysis.
-To answer this question with statistical model comparisons requires comparing the likelihood of the data given each model.
-This is impossible for the ploidy and breeding system models presented so far, however, because the data are different for the different models: they use either the $D/P$ or the $I/C$ or the $ID/CD/CP$ state spaces (see figure \cref{figure:stateclassifications} for state classifications).
-Therefore, the use of different data results in incomparable models.
-
-In order to compare fits of ploidy-only \vs breeding system-only \vs combined trait models, we use the technique of `lumping' states together \citep{tarasov_2019}.
-We use the state space of the $ID/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models.
-Lumping states requires that the transition rates from the lumped state to the singular state be equal \citep{tarasov_2019}.
-First, we lump together $ID$ and $CD$ to form the diploid state, mimicking the $D/P$ model (M26). 
-Proposing a lumped ploidy model by aggregating $ID$ and $CD$ requires forcing the rate of polyploidization from $ID$ and $CD$ to $CP$ to be equal (i.e. $\rho_0=\rho_I=\rho_c$), but also requires
-assuming that the rates of speciation and extinciton for $ID$ and $CD$ to be equal.
-Therefore, we define the new parameters $\lambda_D$ and $\mu_D$  that are the same for each of the two diploid states $ID$ and $CD$. %IM suggestion, describe more the lumping
-We used the same procedure to lump together $CD$ and $CP$ to form the self-compatible state, mimicking the $I/C$ model (M28).
-In this particular case, the rate from $CD$ to $CP$ back to $ID$ is zero and equal for both, so the model is lumpable. 
-However, to fully mimic the breeding system model, we assume that the rate of selfing is equal (i.e.
-$q_0=q_{IC}=\rho_I$) and the rates of speciation and extinction for both $CD$ and $CP$ are the same (new parameters $\lambda_C$ and $\mu_C$).
-We further add a hidden character to each of these models (M27 and M29), and then compare this group of models (\cref{table:lumped}).
-
-We do not include additional models with diploidization because this reverse ploidy transition renders the models non-lumpable.
-When including diploidization, transitions from $CP$ to $CD$ are at rate $\delta$ but transitions from $CP$ to $ID$ do not occur.
-Because $ID$ and $CD$ would be lumped to mimic the $D/P$ model, this model is non-lumpable when $\delta \ne 0$.
-Thus, we can compare models to test whether it is advantageous to include both traits, but only when ignoring diploidization.
-% E: Below was all good, but it seemed like more detail than necessary.
-% Then we define a single parameter $q_0$ going into the lumped state $(CD,CP)$ as the new rate from SI to SC.
-% This defines the lumped model M28 (\cref{figure:lumped}E) that is equivalent to M1 but comparable to M16.
-% We repeat this process to create the lumped model M26 for $D/P$ by restricting the polyploidy rate to be $rho_0$ out of the lumped state $(ID,CD)$.
-% When including diploidization, however, transitions from $CP$ to $CD$ are at rate $\delta$ but transitions from $CP$ to $ID$ do not occur.
-% Since $\delta \neq 0$, the models with diplodization cannot be lumped to be compared to the three-state models.
-
-\subsubsection{Pathways to polyploidy}
-
-Considering ploidy and breeding system together, there are two evolutionary pathways from SI diploid to SC polyploid \citep{brunet2001, robertson_2011}.
-In the one-step pathway, the $CP$ state is produced directly from the $ID$ state when whole genome duplication disables SI.
-In the two-step pathway, the $CD$ state is an intermediate: SI is first lost, and later the SC diploid undergoes polyploidization.
-We quantify the relative contribution of these pathways to polyploidy in two ways, each using the median estimates of rates from the simplest model that includes both traits (M16). 
-Our results differ from those of \citet{robertson_2011} in part because our approach relies on a dated phylogeny and models that allow for state-dependent diversification. %IM suggestion
-
-Both of our methods are based on a propagation matrix that describes flow from $ID$ to $CP$, as in \citet{robertson_2011}.
-We insert an artificial division in the $CP$ state, so that one sub-state contains the $CP$ species that arrived via the one-step pathway and the other substate contains the $CP$ species that arrived via the two-step pathway.
-We consider unidirectional change along each step of the pathway in order to separate them into clear alternatives, and because in this family there is no support for regain of SI, and no strong support for diploidization (see below).
-First, we consider only the rates of transitions between these states, placing them in the propagation matrix \myvec{Q}.
-The matrix $\myvec{P} = \exp(\myvec{Q} t)$ then provides the probabilities of changing from one state to any other state after time $t$.
-Closed-form solutions for the two pathway probabilities are provided in \citet{robertson_2011}.
-Second, we consider not only transitions between states but also diversification within each state.
-State-dependent diversification can change the relative contributions of the two pathways.
-For example, if the net diversification rate is small for $CD$, the two-step pathway will contribute relatively less.
-We therefore include the difference between speciation and extinction along the diagonal elements of the propagation matrix.
-As before, matrix exponentiation provides the relative chance of changing from one state to any other state after time $t$.
-The calculations of the propagation matrix are not probabilities because diversification changes the number of lineages as time passes. 
-We can still use ratios, however, to consider the relative contribution of each pathway, analogous to the normalized age structure in a growing population \citep{leslie_1945}.
-% E & R todo ?: for the case with net div, write out the propagation matrix and equations for pathway contributions
-% R- couldn't solve to get the exact solutions turns out it is really hard.
-
-\subsection{Statistical inference}
-
-\subsubsection{Model fitting}
-
-Parameters for each of the 29 models were coded as graphical models and Bayesian statistical inference was performed with RevBayes \citep{hoehna_2016}.
-Scripts for analyses and key results are available in Dryad. %TODO
-We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $651/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
-For all models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
-Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
-For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the models.
-Convergence and mixing of each MCMC chain was determined by ensuring the effective sample size of each parameter was over 200.
-
-We report posterior distributions for the net diversification parameters ($\lambda - \mu$)  in \cref{figure:netdivall} and full results of all parameters of the diversification models in \cref{suppfigure:DPnodip,suppfigure:DPnodipAB,suppfigure:IC,suppfigure:ICAB,suppfigure:IDCDCPnodip,suppfigure:IDCDCPnodipAB}.
-Additionally, ancestral states at each node in the phylogeny were sampled jointly during the MCMC analyses every 100 generations. 
-%R: Itay says remove but some reviewers were interested in the distribution of the states.
-I say leave we don't need to discuss.
-I'll mention briefly.
-%B: I am not sure what this comment refers to (up or down?), but I'm not touching.
-Ancestral state estimations for all models show the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes for each of the models in \cref{figure:netdivall}.
-(\cref{suppfigure:DPnodipasr,suppfigure:DPnodipABasr,suppfigure:ICasr,suppfigure:ICABasr,suppfigure:IDCDCPnodipasr,suppfigure:IDCDCPnodipABasr}).
-
-\subsubsection{Model selection key questions}
-
-We calculated the marginal log-likelihood for each of the models using fifty stepping stone steps under the methodology of \citet{xie_2010}, implemented in RevBayes \citep{hoehna_2016}.
-Each stepping stone step was found by calculating at least 500 generations of burn-in followed by a total of 1,000 MCMC steps (\cref{table:bayesfactors}).
-
-Using the marginal likelihood values, we calculated Bayes factors to answer five key biological and methodological questions:
-\begin{enumerate}
-\item Are diversification patterns only determined by hidden states and not the traits of interest?---Comparison of character independent models against hidden state (\cref{table:bayesfactors,table:M11M15,supptable:M6M10,supptable:M16M20,supptable:M21M25}).
-\item Are hidden states necessary to explain diversification rate heterogeneity?---Comparison of simple models against hidden state models (\cref{supptable:testaddhidden}).
-\item Does a second focal trait add information about the diversification process?---Comparison of lumped models against IC/CD/CP models (\cref{figure:lumped,table:lumped,suppfigure:lumpedDP,suppfigure:lumpedIC}).
-\item Are conclusions robust to assumptions about hidden state transitions?---Comparison amongst hidden states models with equal hidden rates and asymmetrical rates (\cref{supptable:asymmetry,suppfigure:asymmetric}).
-\item Is there evidence for diploidization?---Comparison amongst log-scalemodels with and without diploidization (\cref{supptable:testdiploidization,suppfigure:alldip}).
-\end{enumerate}
-
-Each model comparison is reported with a Bayes factor on the natural log-scale: the comparison between models $M_0$ and $M_1$ is $K=ln(BF(M_1,\ M_0)) = \ln[ P(\mathbf{X} | M_1) - P(\mathbf{X} | M_0)]$.
-There is `strong evidence' for $M_1$ when this value is more than 10, moderate support if the value is more than 1, and no evidence if the value is between -1 and 1.
-If the value of $K$ is negative the evidence goes towards $M_0$ \citep{kass1995}.
-
-\begin{table}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|r|c|c|c|c|l|}
-\hline
-Model                & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M2    & M3     & M4    & M5     & Evidence                                                                                     \\ \hline
-M1.
-D/P              & -1283.76                                                          & 59.90 & 49.23  & 60.48 & 58.82  & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M1\end{tabular}            \\ \hline
-\textbf{M2.
-CID D/P} & \textbf{-1223.86}                                                 &       & -10.66 & 0.579 & -1.079 & \begin{tabular}[c]{@{}l@{}}Strong preference for \\ M2 over M1 and M3\end{tabular}              \\ \hline
-M3.
-D/P+A/B          & -1234.52                                                          &       &        & 11.24 & 9.58   & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}   \\ \hline
-\textbf{M4.
-D/P+A/B asym}     & \textbf{-1223.28}                                                          &       &        &       & -1.658 & \begin{tabular}[c]{@{}l@{}}Moderate evidence for\\ only asymmetric hidden rates\end{tabular} \\\hline
-M5.
-D/P+A/B all asym & -1224.93                                                          &       &        &       &        &          \\ \hline
-\end{tabular}
-\caption{Bayes factors for ploidy only models without diplodization in log-scale.
-Numbers smaller than -1 indicate moderate preference for the model listed in the row, and values larger than 1 indicate moderate preference for the model listed in the respective column .
-Conventional threshold for `strong' preference is an absolute value larger than 10.
-The character independent model (M2, bold) is strongly preferred over the ploidy only model (M1) or a hidden state model with symmetric rates (M3).
-Moderate support exists for the model with ploidy and asymmetric hidden states (M4, bold) over the rest of the ploidy only models.}
-\label{table:bayesfactors}
-\end{table}
-
-% B: M111->M11
-\begin{table}
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|r|c|c|c|c|l|}
-\hline
-Model                      & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & M12   & M13   & M14   & M15   & Evidence                                                                                       \\ \hline
-M11.
-I/C                   & -1309.07                                                          & 41.13 & 38.59 & 61.34 & 61.40 & \begin{tabular}[c]{@{}l@{}}Every model strongly \\ preferred over M11\end{tabular}            \\ \hline
-M12.
-CID I/C               & -1267.93                                                          &       & -2.53 & 20.21 & 20.37 & \begin{tabular}[c]{@{}l@{}}Models with asymmetric rates\\ are preferred over  M12\end{tabular} \\ \hline
-M13.
-I/C+A/B               & -1270.47                                                          &       &       & 22.75 & 22.80 & \begin{tabular}[c]{@{}l@{}}Asymmetric rates strongly\\ preferred over symmetric\end{tabular}   \\ \hline
-\textbf{M14.
-I/C+A/B asym} & \textbf{-1247.72}                                                 &       &       &       & 0.05 & No evidence                                                                                    \\ \hline
-\textbf{M15.
-I/C+A/B all asym }     & \textbf{ -1247.66}                                                          &       &       &       &       &       \\ \hline
-\end{tabular}
-\caption{Bayes factors for breeding system only models in log-scale.
-Numbers smaller than -1 indicate moderate preference for the model listed on the row, and values larger than 1 indicate moderate preference for the model in the respective column.
-Conventional threshold for `strong' preference is an absolute value larger than 10.
-Moderate to strong preference exist for models M14 and M15 which include a hidden states with asymmetrical rates.}
-\label{table:M11M15}
-\end{table}
-
-% E: Could we include a figure of the traits on the tree?  At least as supp info. %FIXME
-
-\begin{table}[]
-\addtolength{\tabcolsep}{-3pt}
-\begin{tabular}{|l|c|c|c|c|}
-\hline
-Model                       & \begin{tabular}[c]{@{}l@{}}Marginal\\ log-likelihood\end{tabular} & Comparison  & K=ln(BF($M_{i}$,$M_{ii}$)) & \begin{tabular}[c]{@{}l@{}}Preferred\\ Model (Evidence)\end{tabular} \\ \hline
-M26.
-Lumped D/P             & -1463.22                                                          & M26 vs.
-M16 & 4.10            & M16 (Moderate)                                                       \\
-\textbf{M16.
-ID/CD/CP}      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
-M27.
-Lumped D/P+A/B         & -1417.67                                                          & M27 vs.
-M23 & 3.678           & M23 (Moderate)                                                       \\
-\textbf{M23.
-ID/CD/CP +A/B} & \textbf{-1414.00}                                                 &             &                 &                                                                      \\ \hline
-M28.
-Lumped I/C             & -1458.41                                                          & M28 vs.
-M16 & -0.69           & No evidence                                                          \\
-M16. ID/CD/CP      & \textbf{-1459.11}                                                 &             &                 &                                                                      \\ \hline
-M29.
-Lumped I/C+ A/B        & -1416.60                                                          & M29 vs.
-M23 & 2.60            & M23 (Moderate)                                                       \\
-\textbf{M23.
-IC/CD/CP+A/B}  & \textbf{-1414.00}                                                 &             &                 &                                                                      \\ \hline
-\end{tabular}
-\caption{Testing the addition of a focal trait to a binary state model via Bayes factors (log-scale).
-Numbers smaller than -1 indicate moderate preference for the model in the row, and values larger than 1 indicate moderate preference for the model in the respective column.
-Conventional threshold for `strong' preference is an absolute value larger than 10.
-Three-state models (M16, M23, bold) are moderately preferred over two-state models .
-Moderate evidence exists towards the inclusion of breeding systems in ploidy models (M16, M23, bold).
-Moderate to no evidence towards the  inclusion of ploidy in breeding system models is indicated by the comparisons of models M28 \vs M16 and M29 \vs M23 where the difference is the inclusion of ploidy information.}
-\label{table:lumped}
-\end{table}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.5\textwidth]{fig1.pdf}
-\caption{
-Character states used in the models.
-(A) Each species retained on the phylogeny belonged to one of five possible categories, depending on whether ploidy and/or breeding system were known. 
-Number of species in each category is indicated; for example, 70 species are self-compatible with unknown ploidy.
-Character state abbreviations are: $I$ for self-incompatible, $C$ for self-compatible, $D$ for diploid, $P$ for polyploid, $?$ for unknown.
-Because polyploidization breaks this form of self-incompatibility, self-incompatible species with unobserved ploidy ($I?$) are taken to be diploid ($ID$), and polyploid species with unobserved breeding system ($?P$) are taken to be SC ($CP$).
-(B) Category groupings into states for each model class.
-In the ploidy-only models (M1- M10), states are coded as $D$ \& $P$ when uncertain/consistent with either state; in the breeding system-only models (M11-M15) such states are coded as $C$; in the ploidy and breeding system models (M16-M29), they are coded as $CD$ \& $CP$.
-In the hidden-trait models, all species could take on either of two `hidden' character states. %IM: #33 
-Two species, \emph{Lycium californicum} and \emph{Solanum bulbocastanum}, are simultaneously $ID$ and $CP$, and by adding them the sample adds to the total of 651 taxa used for analyses.
-}\label{figure:stateclassifications}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{fig2.pdf} 
-\caption{Net diversification rates for SSE models of focal traits with or without hidden state. 
-(A) Ploidy only model (M1) showing higher net diversification linked to diploid state $D$ compared to polyploid state $P$.
-(B) Ploidy with hidden states model (M4) showing that the net diversification is higher for hidden state $B$ (lighter colors) compared to hidden state $A$ (darker colors) and both diploid and polyploid states within each hidden traits have overlapping net diversification rates.
-(C) Breeding system only model (M9) showing higher net diversification linked to self-incompatible $I$ state compared to self-compatible state $C$.
-(D) Breeding system with hidden states model (M11) showing diversification differences in both hidden states (light \vs.
-dark colors) and little to no overlapping in between self-compatible \vs self-compatible states.
-(E) Ploidy and breeding system model (M13) showing higher net diversification linked to self-incompatible diploid state $ID$ compared to both self-compatible states despite ploidy level ($CD, CP$).
-(F) Ploidy, breeding system, and hidden states model (M16) showing a similar pattern that panel (E) within each hidden state $A$ and $B$.
-For hidden state $B$ there is a larger overlap of net diversification between states $CD_B$ and $ID_B$.}
-\label{figure:netdivall}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{fig3.pdf} %lumped.pdf
-\caption{Comparing two and three-state models using lumped models. (A) Ploidy model only (M1) where data enter as binary $D$ and $P$. 
-(B) Lumped model for ploidy (M26) where data are the three-state values ($ID,CP,CD$) but results are equivalent to model M1.  
-(C) Ploidy and breeding system model (M16) where  data enter as the three-state values. Models M26 and M16 are comparable whereas M1 and M16 are not.
-(D) Breeding system only model (M11) where data are entered as binary $I$ and $C$. 
-(E) Lumped model for breeding system (M28) where data are the three-state values ($ID,CP,CD$) but results are equivalent to model M11. Model M26 can be compared with model M16 from panel (C).
-Model comparisons are done via Bayes factors and results are shown in \cref{table:lumped}}  
-\label{figure:lumped}
-\end{figure}
-
-% E: %FIXME: re-run with new rate estimates; add panel letters?
-\begin{figure}
-   % \centering 
-    \includegraphics[width=0.5\textwidth]{fig4a} \\ [40pt] 
-        \includegraphics[width=0.9\textwidth]{fig4b}
-    \caption{
-        Contributions of the two pathways to polyploidy.
-        The one-step pathway is direct ID$\rightarrow$CP transitions.
-        The two-step pathway consists of ID$\rightarrow$CD$\rightarrow$CP transitions.
-        When considering only rates of transitions among the states (ignoring the diversification rate parameters), the one-step pathway dominates on short timescales and the two-step on long timescales (left panels).
-        When also considering diversification within each state, the one-step pathway, in which polyploidization breaks down SI, dominates over any timescale (right panels).
-        The top panels show the separate contributions of each pathway.
-        The bottom panels show the proportional contribution of the one-step pathway (\ie one-step / [one-step + two-step]).
-    }
-    \label{figure:pathways}
-\end{figure}
+\subsection*{New Phytologist Supporting Information Figs S1-S18 and Tables S1-S6\\
+Article title: Interaction between ploidy, breeding system, and lineage diversification\\
+Authors: Rosana Zenil-Ferguson, J. Gordon Burleigh, William A. Freyman, Boris Igi\'c, Itay Mayrose and Emma E. Goldberg\\
+Article acceptance date: 14 August 2019}
 
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS1.pdf}
+%\includegraphics[width=\textwidth]{figS1.pdf}
     \caption{Ploidy and breeding system data according to three different classifications.
@@
 The contributions of the focal trait to the diversification process can be measured by comparing the models in each of the columns.
-That is, the focal trait only models assume that speciation and extinction rates are only linked to the trait itself, the hidden trait only models assume that the diversification rates are  linked to unknown factors but not the trait of interest, and the focal trait with hidden trait models assume that both the focal trait and unknown factors are contributing to diversification.
-(large format figure)}
+That is, the focal trait only models assume that speciation and extinction rates are only linked to the trait itself, the hidden trait only models assume that the diversification rates are  linked to unknown factors but not the trait of interest, and the focal trait with hidden trait models assume that both the focal trait and unknown factors are contributing to diversification.}
     \label{fig:allmodels}
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS3.pdf}
+%\includegraphics[width=\textwidth]{figS3.pdf}
 \caption{Posterior distributions for each of the parameters in the ploidy only model (M1).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS4.pdf}
+%\includegraphics[width=\textwidth]{figS4.pdf}
 \caption{Posterior distributions for each of the parameters in the ploidy and hidden trait model (M4).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS5.pdf}
+%\includegraphics[width=\textwidth]{figS5.pdf}
 \caption{Posterior distributions for each of the parameters in the breeding system only model (M11).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS6.pdf}
+%\includegraphics[width=\textwidth]{figS6.pdf}
 \caption{Posterior distributions for each of the parameters in the breeding system and hidden trait model (M14).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS7.pdf}
+%\includegraphics[width=\textwidth]{figS7.pdf}
 \caption{Posterior distribution for each of the parameters in the ploidy and breeding system model (M16).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS8.pdf}
+%\includegraphics[width=\textwidth]{figS8.pdf}
 \caption{Posterior distribution for each of the parameters in the ploidy, breeding system, and hidden trait model (M19).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS9.pdf}
+%\includegraphics[width=\textwidth]{figS9.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under  the ploidy only model (M1).
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS10.pdf}
+%\includegraphics[width=\textwidth]{figS10.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the ploidy and hidden states model (M4).} % XXX
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS11.pdf}
+%\includegraphics[width=\textwidth]{figS11.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the  breeding system only model (M11).} % XXX
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS12.pdf}
+%\includegraphics[width=\textwidth]{figS12.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the breeding system and hidden states model (M14.} % XXX
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS13.pdf}
+%\includegraphics[width=\textwidth]{figS13.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the ploidy and breeding system model (M16).} % XXX
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS14.pdf}
+%\includegraphics[width=\textwidth]{figS14.pdf}
 \caption{Ancestral state estimation showing the maximum \emph{a posteriori} estimates of the marginal probability distributions for each of the 650 internal nodes under the  ploidy, breeding systems, and hidden states model (M19).} % XXX
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS15.pdf}
+%\includegraphics[width=\textwidth]{figS15.pdf}
 \caption{Testing the addition of breeding system to ploidy models.
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS16.pdf}
+%\includegraphics[width=\textwidth]{figS16.pdf}
 \caption{Testing the addition of ploidy to breeding system models.
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS17.pdf}
+%\includegraphics[width=\textwidth]{figS17.pdf}
 \caption{Effect of asymmetric rates in hidden models.
@@
 \begin{suppfigure}
-\includegraphics[width=\textwidth]{figS18.pdf}
+%\includegraphics[width=\textwidth]{figS18.pdf}
 \caption{Posterior distributions for the net diversification rates of the preferred models with diploidization.

### 2019-09-05 basicdata/check.R
Rev analyses and results final

@@ -1,64 +1,0 @@
-### Compare breeding system data from 2014 compilation with the current file.
-### (Don't worry about uncertain/polymorphic species for now.)
-
-library(ape)
-phy <- read.tree("~/Dropbox/solploidy/basicdata/solcorrecttip.tre")
-#--------------------------------------------------
-# Data files from 2014
-#--------------------------------------------------
-
-dat1 <- read.csv("statesICDP.csv", as.is=T)
-dat1 <- subset(dat1, Species %in% phy$tip.label)
-
-dat1 <- subset(dat1, !(ID==CD & CD==CP)) # remove 0.33 0.33 0.33
-dat1$SI <- dat1$ID
-dat1$SC <- dat1$CD + dat1$CP
-dat1 <- subset(dat1, SI != SC) # remove 0.5 0.5
-dat1[which(!(dat1$SI %in% c(0, 1))),] # Solanum_brevicaule Solanum_candolleanum Solanum_chacoense
-dat1 <- dat1[-which(!(dat1$SI %in% c(0, 1))),] # drop further uncertainty
-table(dat1$SI)
-table(dat1$SC)
-# 331 SC, 143 SI
-subset(dat1, Species %in% c("Capsicum_schottianum", "Capsicum_scolnikianum", "Lycium_californicum")) # for comparison with dat2, below
-
-# fully cleaned
-dat1 <- data.frame(name=dat1$Species, bs1=c("SC", "SI")[dat1$SI+1])
-
-#--------------------------------------------------
-# Current data file
-#--------------------------------------------------
-
-dat2 <- read.csv("solanaceaerecords.csv", as.is=T)
-dat2 <- subset(dat2, fullname %in% phy$tip.label)
-
-dat2 <- unique(dat2[,c("fullname", "selfincomp")])
-dat2 <- subset(dat2, selfincomp != "")
-dat2[which(duplicated(dat2$fullname)),]
-subset(dat2, fullname %in% c("Capsicum_schottianum", "Lycium_californicum")) # the Capsicum was coded as SC only in dat1; the Lycium is truly polymorphic
-
-# fully cleaned
-dat2 <- subset(dat2, !(fullname %in% c("Capsicum_schottianum", "Capsicum_scolnikianum", "Lycium_californicum")))
-names(dat2) <- c("name", "bs2")
-
-#--------------------------------------------------
-# Compare
-#--------------------------------------------------
-
-dat <- merge(dat1, dat2, all=T)
-
[…]

### 2019-09-05 basicdata/datawithphylo.R

@@ -1,47 +1,0 @@
-library("ggplot2")
-library("ggtree")
-library("treeio")
-library("ggnewscale")
-#library("RevGadgets")
-library("wesanderson")
-library("dyplr")
-
-cols<-c(wes_palette("Darjeeling1"),wes_palette("Darjeeling2"))
-cols2<- c("#7b3294","#c2a5cf","#a6dba0","#008837","#ffffbf")
-cols3<-c("#ece2f0", "#67a9cf","#e31a1c","#fd8d3c","#02818a","#014636")
-
-
-soltree<-read.nexus("~/Dropbox/solploidy/basicdata/fullmatchtree.nex")
-soldata<- read.csv("~/Dropbox/solploidy/basicdata/discretematrix.csv")
-dp<-data.frame(soldata[,2:4])
-rownames(dp)<-soltree$tip.label
-
-circ <- ggtree(soltree, layout = "circular")
-
-p1 <- gheatmap(circ, dp[, "Ploidy", drop=F], offset=.8, width=.1,
-               colnames_angle=90, colnames_offset_y = .25)
-p2 <- gheatmap(p1, dp[, "BS", drop=F], offset=5, width=.1,
-               colnames_angle=90, colnames_offset_y = .25)
-p3 <- gheatmap(p2, dp[, "Ploidybs", drop=F], offset=10, width=.1,
-               colnames_angle=90, colnames_offset_y = .25)
-
-col <- c(cols2[2],cols[7],cols2[1],cols[1],cols[9],cols[9],cols2[3],cols2[4],"yellow",cols2[3],cols[6],cols[6],"grey")
-names(col) = c("C","CD","Compatible","CP","D","Diploid","I","ID","IDCP","Incompatible","P","Polyploid","Unknown")
-#D = cols[9]
-#%P=cols[6]
-
-#CD cols[7]
-#CP cols[1]
-
-#Compatible=cols2[1],
-#C= cols2[2]
-#Incompatible= cols2[3]
-#I cols2[3])
-#ID= cols2[4]
-#IDCP= "yellow"
-#Diploid= cols[9]
-#Polyploid= cols[6]
-
-pp <- p3 + scale_fill_manual(values=col)
-require(phytools)
-plotTree(soltree,type="fan", cex=0.18,fsize=0.19,label.offset= 2)

### 2019-09-05 basicdata/matrixdataprocessing.R

@@ -1,369 +1,0 @@
-library("ape")
-library("phytools")
-library("treeplyr")
-
-## Solanaceae time-tree with corrected names Sarkinen et al 2013
-solanaceae.tree<-read.tree("~/Dropbox/solploidy/basicdata/solcorrecttip.tre")
-solanaceae.allrecords<-read.csv("~/Dropbox/solploidy/basicdata/solanaceaerecords.csv",header=TRUE,sep=",", stringsAsFactors=FALSE, na.strings="")
-cultivars<-which(solanaceae.allrecords$cultivarhybrid==1)
-solanaceae.allrecords<-solanaceae.allrecords[-cultivars,]
-
-#How many species, how many genera
-species<-unique(solanaceae.allrecords$fullname)
-total.species<-length(species)# 2258 total species
-genera<-unique(solanaceae.allrecords$genus)
-total.genera<-length(genera)#100 genera
-
-# Dataset created by RZF from CCDB and independent records by EEG
-#idnumber, source,fullname, genus, species,cultivarhybrid,csomenumber,lifehistoryclean, ploidy selfincomp, ploidybyZF, namechangedfrom
-
-
-#Calculating the mode(s)
-Mode <- function(x) {
-  ux <- unique(x)
-  ux[which.max(tabulate(match(x, ux,incomparables=NA)))]
-}
-
-#Creating a very simple database with summary stats
-mode.csome<-rep(0,total.species)
-min.csome<-rep(0,total.species)
-mode.ploidy<-rep(0,total.species)
-min.ploidy<-rep(0,total.species)
-self.incomp<-rep(0,total.species)
-life.hist<-rep(0,total.species)
-
-for (i in 1:total.species){
-  index<-which(solanaceae.allrecords$fullname==species[i])
-  aux1<- solanaceae.allrecords$csomenumber[index]
-  if(all(is.na(aux1))){
-  	 mode.csome[i]<-NA
-  	 min.csome[i]<-NA  	
-  }else{
-  mode.csome[i]<-min(Mode(aux1))
-  min.csome[i]<-min(aux1,na.rm=TRUE)
-  }
-  aux2<- solanaceae.allrecords$ploidy[index]
-  if(all(is.na(aux2))){
-  	 mode.ploidy[i]<-NA
-  	 min.ploidy[i]<-NA  	
-  }else{
[…]

### 2019-09-05 basicdata/revbayesdataprocessing.R

@@ -1,58 +1,0 @@
-# Processing for Revbayes files
-library("ape")
-library("phytools")
-library("treeplyr")
-solanaceae.tree<-read.tree("~/Dropbox/solploidy/basicdata/solcorrecttip.tre")
-threestate.matrix <- read.csv("~/Dropbox/solploidy/basicdata/newstatesICDP.csv", as.is=T)
-threestate.matrix<-threestate.matrix[,-1]
-
-matched<-make.treedata(solanaceae.tree,threestate.matrix) #638 species!!!!
-write.nexus(matched$phy,file="~/Dropbox/solploidy/basicdata/fullmatchtree.nex")
-
-diploidsi<-which((matched$dat[,1]==1 & matched$dat[,2]==0 & matched$dat[,3]==0)==TRUE)# 144
-diploidsc<-which((matched$dat[,1]==0 & matched$dat[,2]==1 & matched$dat[,3]==0)==TRUE)#157
-polyploids<-which((matched$dat[,1]==0 & matched$dat[,2]==0 & matched$dat[,3]==1)==TRUE) #108
-
-diploidsna<-which((matched$dat[,1]==1 & matched$dat[,2]==1 & matched$dat[,3]==0)==TRUE) #170
-sc<-which((matched$dat[,1]==0 & matched$dat[,2]==1 & matched$dat[,3]==1)==TRUE)#70
-idcp<-which(matched$dat[,1]==1 & matched$dat[,2]==0 & matched$dat[,3]==1)#2
-
-## Data for threestate analyses
-threestate<- rep(0,dim(matched$dat)[1])
-threestate[diploidsc]="0"
-threestate[polyploids]="1"
-threestate[diploidsi]="2"
-threestate[sc]="(0 1)"
-threestate[diploidsna]<-"(0 2)"
-threestate[idcp]<-"(1 2)"
-threestate<-as.data.frame(threestate)
-row.names(threestate)<-matched$phy$tip.label
-
-write.table(threestate,file="~/Dropbox/solploidy/basicdata/threestate2.tsv",sep="\t")
-
-##binary ploidy
-binary.ploidy<-rep(0,dim(matched$dat)[1])
-binary.ploidy[diploidsc]<-"0"
-binary.ploidy[polyploids]<- "1"
-binary.ploidy[diploidsi]<-"0"
-binary.ploidy[sc]<-"?"
-binary.ploidy[diploidsna]<-"0"
[…]

### 2019-09-05 basicdata/solanaceaetreecorrection.R

@@ -1,26 +1,0 @@
-setwd('~/Dropbox/solploidy/basicdata')
-library("ape")
-library("phytools")
-library("treeplyr")
-
-## Time-tree from Sarkinen et al. 2013 solaneaceae done with Beast
-solanum.sarkinen<-read.tree(file="solanumbeast.tre")
-## Names for the solaneaceae tree, the tree above was listed with numbers which was less than ideal
-solanum.names<-read.csv("speciesnames2.csv",header=FALSE,sep=",",stringsAsFactors=FALSE)
-## Putting the names on the tip labels
-fixingtiplabels<-as.numeric(solanum.sarkinen$tip.label)
-solanum.sarkinen$tip.label<-solanum.names[fixingtiplabels,2] 
-# I double checked with their publication turns out to be exact tree shown in figure 12862_2013_2443_MOESM2_ESM.tiff
-
-#Tips to drop/ repeated species list from EEG
-remove.names<-read.csv("tipstodrop.csv",header=FALSE, sep=",",stringsAsFactors=FALSE)
-length(remove.names[,1])#55 to drop
-solanaceae.correcttips<-drop.tip(solanum.sarkinen,remove.names[,1])
-# This corrected tree has 1023 tips, it dropped 53 out 55
-write.tree(solanaceae.correcttips, file="solcorrecttip.tre")
-
-##Plot
-plotTree(solanaceae.correcttips,part=0.98,fsize=0.05, lwd=0.9)
-# The tree is approximately 30my
-h<-max(nodeHeights(solanaceae.correcttips))
-#axis(1,labels=c(136.9,68.4,0), pos=-0.03*h,at=seq(0,h,by=h/2),lwd=0.2,cex=0.2)

### 2019-09-05 rev_analyses/bisse250K/bisseDPv2.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=500mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o bisseoutput.log
+#PBS -e bisseerrorscluster.log
+#PBS -l walltime=48:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/bisse/bisseDP.Rev

### 2019-09-05 rev_analyses/bisseSInoreturn250K/bisseSInoreturn.sh

@@ -1,0 +1,12 @@
+#!/bin/bash -l
+#PBS -l pmem=500mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o bissesinoret.log
+#PBS -e bissesinoreterrorscluster.log
+#PBS -l walltime=48:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/bissesinoret/bisseSInoreturn.Rev
+

### 2019-09-05 rev_analyses/bissenodip250K/bisseDPnodip.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=500mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o bissenodipoutput.log
+#PBS -e bissenodiperrorscluster.log
+#PBS -l walltime=48:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/bissenodip/bisseDPnodiploidization.Rev

### 2019-09-05 rev_analyses/hisse250K/hisseDP.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=500mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o hisseoutput.log
+#PBS -e hisseerrorscluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb /home/eeg/shared/twostatesse/hisse/hisseDP.Rev

### 2019-09-05 rev_analyses/hisse250K/hisseDPallasym.sh

@@ -1,0 +1,12 @@
+#!/bin/bash -l
+#PBS -l pmem=500mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o hisseallasymoutput.log
+#PBS -e hisseallasymerrorscluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/hisse/hisseDPallasym.Rev
+

### 2019-09-05 rev_analyses/hisse250K/hisseDPasym.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=500mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o hisseasym.log
+#PBS -e hisseasymerrorscluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/hisse/hisseDPasym.Rev

### 2019-09-05 rev_analyses/hisseSInoreturn250K/hissesinoret.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=500mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o hisseoutput.log
+#PBS -e hisseerrorscluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/hissesinoret/hissesinoret.Rev

### 2019-09-05 rev_analyses/hisseSInoreturn250K/hissesinoretallasym.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=500mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o hisseoutputallasym.log
+#PBS -e hisseerrorsclusterallasym.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb /home/eeg/shared/twostatesse/hissesinoret/hissesinoretallasym.Rev

### 2019-09-05 rev_analyses/hisseSInoreturn250K/hissesinoretasym.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=500mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o hisseoutputas.log
+#PBS -e hisseerrorsclusteras.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb /home/eeg/shared/twostatesse/hissesinoret/hissesinoretasym.Rev

### 2019-09-05 rev_analyses/hissenodip250K/hissenodipDPv1.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=500mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o hisseoutput.log
+#PBS -e hisseerrorscluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/hissenodip/hissenodipDP.Rev

### 2019-09-05 rev_analyses/hissenodip250K/hissenodipallasym.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=500mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o hisseoutputallasym.log
+#PBS -e hisseerrorsclusterallasym.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/hissenodip/hissenodipallasym.Rev

### 2019-09-05 rev_analyses/hissenodip250K/hissenodipasymDP.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=500mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o hisseoutputasym.log
+#PBS -e hisseerrorsclusterasym.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/hissenodip/hissenodipasym.Rev

### 2019-09-05 rev_analyses/lumped/lumpedDP.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=600mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o lumpedDPrun.log
+#PBS -e lumpedDPcluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/lumped/lumpedDP.Rev

### 2019-09-05 rev_analyses/lumped/lumpedDPv2.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=600mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o lumpedDPrun2.log
+#PBS -e lumpedDPcluster2.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/lumped/lumpedDPv2.Rev

### 2019-09-05 rev_analyses/lumped/lumpedIC.sh

@@ -1,0 +1,12 @@
+#!/bin/bash -l
+#PBS -l pmem=600mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o lumpedICrun.log
+#PBS -e lumpedICcluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/lumped/lumpedIC.Rev
+

### 2019-09-05 rev_analyses/lumped/lumpedICv2.sh

@@ -1,0 +1,12 @@
+#!/bin/bash -l
+#PBS -l pmem=600mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o lumpedICrun2.log
+#PBS -e lumpedICcluster2.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/lumped/lumpedICv2.Rev
+

### 2019-09-05 rev_analyses/lumped/lumpedhiddenDP.sh

@@ -1,0 +1,12 @@
+#!/bin/bash -l
+#PBS -l pmem=600mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o lumpedhiddenDPrun.log
+#PBS -e lumpedhiddenDPcluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/lumped/lumpedhiddenDP.Rev
+

### 2019-09-05 rev_analyses/lumped/lumpedhiddenDPv2.sh

@@ -1,0 +1,12 @@
+#!/bin/bash -l
+#PBS -l pmem=600mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o lumpedhiddenDPrun2.log
+#PBS -e lumpedhiddenDPcluster2.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/lumped/lumpedhiddenDPv2.Rev
+

### 2019-09-05 rev_analyses/lumped/lumpedhiddenIC.sh

@@ -1,0 +1,12 @@
+#!/bin/bash -l
+#PBS -l pmem=600mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o lumpedhiddenICrun.log
+#PBS -e lumpedhiddenICcluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/lumped/lumpedhiddenIC.Rev
+

### 2019-09-05 rev_analyses/lumped/lumpedhiddenICv2.sh

@@ -1,0 +1,12 @@
+#!/bin/bash -l
+#PBS -l pmem=600mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o lumpedhiddenICrun2.log
+#PBS -e lumpedhiddenICcluster2.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/lumped/lumpedhiddenICv2.Rev
+

### 2019-09-05 rev_analyses/muhisse250K/muhisseallasym100K.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=600mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o muhisseallasym.log
+#PBS -e muhisseallasymerrorscluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/muhisse/muhisseallasym100K.Rev

### 2019-09-05 rev_analyses/muhisse250K/muhisseasym100K.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=600mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o muhisseasym.log
+#PBS -e muhisseaasymerrorscluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/muhisse/muhisseasym100K.Rev

### 2019-09-05 rev_analyses/muhisse250K/muhissedpsi.sh

@@ -1,0 +1,13 @@
+#!/bin/bash -l
+#PBS -l pmem=600mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o muhisse.log
+#PBS -e muhisseerrorscluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/muhisse/muhisseDPSI.Rev
+
+

### 2019-09-05 rev_analyses/muhissenodip250K/muhissedpsiv1.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=600mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o muhissenodip.log
+#PBS -e muhissenodiperrorscluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/muhissenodip/muhisseDPSInodip.Rev

### 2019-09-05 rev_analyses/muhissenodip250K/muhissenodipallasym100K.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=600mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o muhissenodipallasym.log
+#PBS -e muhissenodipallasymerrorscluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/muhissenodip/muhissenodipallasym100K.Rev

### 2019-09-05 rev_analyses/muhissenodip250K/muhissenodipasym100K.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -l
+#PBS -l pmem=600mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o muhissenodipasym.log
+#PBS -e muhissenodipasymerrorscluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/muhissenodip/muhissenodipasym100K.Rev

### 2019-09-05 rev_analyses/mussefull250K/mussedpsiv1.sh

@@ -1,0 +1,12 @@
+#!/bin/bash -l
+#PBS -l pmem=500mb
+#PBS -l nodes=1:ppn=1
+#PBS -m abe
+#PBS -M rzenil@umn.edu
+#PBS -o mussefull.log
+#PBS -e mussefullerrorscluster.log
+#PBS -l walltime=64:00:00
+module load gcc
+
+/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/mussefull/MuSSEDPSI.Rev
+

### 2019-09-05 rev_analyses/src/combine_logs.py

@@ -1,80 +1,0 @@
-#! /usr/bin/python
-
-import csv
-import itertools
-import operator
-import sys
-
-burnin = 10 # per trace
-sample_freq = 1 # number of iterations per sample
-n_runs = 20 # number of MCMC runs to combine
-models = ["sol"]
-
-csv.field_size_limit(sys.maxsize/2)
-for model in models:
-
-    print("Working on model: " + model)
-
-    print("Combining parameter traces...")
-
-    final_csv = []
-    gen = 0
-    header_done = False
-    for log in range(1, n_runs + 1):
-        with open("output/" + model + str(log) + ".log", 'r') as csvfile:
-            lines_to_skip = 0
-            csvreader = csv.reader(csvfile, delimiter="\t")
-            for j, row in enumerate(csvreader):
-                if (row[0][0] == "#"):
-                    lines_to_skip += 1
-                else:
-                    if not header_done:
-                        final_csv.append(row)
-                        header_done = True
-                    elif j - 1 - lines_to_skip > burnin:
-                        final_row = []
-                        for i, column in enumerate(row):
-                            if i == 0:
-                                final_row.append(gen)
-                                gen += sample_freq
-                            else:
-                                final_row.append(column)
-                        final_csv.append(final_row)
-
-    with open("output/combined_" + model + ".log", "w") as csvfile:
-        csvwriter = csv.writer(csvfile, delimiter="\t", lineterminator="\n")
-        for row in final_csv:
-            csvwriter.writerow(row)
-
-    print("Combining ancestral state traces...")
-
-    final_csv = []
-    gen = 0
-    header_done = False
-    for log in range(1, n_runs + 1):
-        with open("output/" + model + str(log) + "_anc.log", 'r') as csvfile:
-            lines_to_skip = 0
-            csvreader = csv.reader(csvfile, delimiter="\t")
-            for j, row in enumerate(csvreader):
-                if (row[0][0] == "#"):
-                    lines_to_skip += 1
-                else:
-                    if not header_done:
-                        final_csv.append(row)
-                        header_done = True
-                    elif j - 1 - lines_to_skip > burnin:
-                        final_row = []
-                        for i, column in enumerate(row):
-                            if i == 0:
-                                final_row.append(gen)
-                                gen += sample_freq
-                            else:
-                                final_row.append(column)
-                        final_csv.append(final_row)
-
[…]

### 2019-09-05 rev_analyses/src/generate_chromo_data.R

@@ -1,72 +1,0 @@
-
-#
-# Generate chromosome data input file for RevBayes analysis.
-#
-# Will Freyman
-#
-
-in_file = "../csomeseries/soldataalltaxa.csv"
-out_file = "data/chromo_counts.tsv"
-out_file_notes = "data/chromo_counts_notes.txt"
-
-d = read.csv(in_file, stringsAsFactors=FALSE)
-a = FALSE
-a_notes = FALSE
-max_c = 0
-min_c = 1000
-
-for (i in 1:nrow(d)) {
-
-    counts = vector()
-
-    # chromo counts are in columns 2 thru 16
-    # of original data file
-    for (j in 2:16) {
-        
-        c = d[i,j]
-        if (!is.na(c) & c %% 2 == 0) {
-            
-            c = c/2
-            counts = c(counts, c)
-            if (c > max_c)
-                max_c = c
-            if (c < min_c)
-                min_c = c
-
-        } else if (!is.na(c) & c %% 2 != 0) {
-    
-            write(paste0(d[i,1], ": dropped odd diploid chromo count = ", c),
-                  file=out_file_notes,
-                  append=a_notes)
-            a_notes = TRUE
-
-        }
-
-    }
-
-    if (length(counts) == 0) {
-
-        counts = paste0(d[i,1], "\t", "?")
-
-    } else if (length(counts) == 1) {
-        
-        counts = paste0(d[i,1], "\t", counts)
-
-    } else {
-
-        counts = paste0(counts, collapse=" ")
-        counts = paste0(d[i,1], "\t", "(", counts, ")")
-
-    }
-
-    write(counts, file=out_file, append=a)
-    a = TRUE
-
-}
-
-write(paste0("min haploid chromosome count = ", min_c),
-             file=out_file_notes,
-             append=a_notes)
-write(paste0("max haploid chromosome count = ", max_c),
-             file=out_file_notes,
-             append=a_notes)

### 2019-09-05 rev_analyses/src/run_parallel.sh

@@ -1,9 +1,0 @@
-#!/bin/bash
-
-rm -rf temp_output
-mkdir temp_output
-
-for rep in {1..20}
-do
-    echo "rep = ${rep}; seed(${rep}); source(\"src/chromoploid.Rev\");" | rb > temp_output/${rep}.out &
-done

### 2019-09-05 summarystats/pathways.R

@@ -1,149 +1,0 @@
-#--------------------------------------------------
-# method 1: P = exp(Q t)
-#-------------------------------------------------- 
-library(rmutil)  # contains mexp()
-path.probs1 <- function(t, Q)
-{
-    P <- mexp(Q * t)
-    
-    ans <- c(P[1,3], P[1,4])
-    names(ans) <- c("Wa", "Wb")
-
-    return(ans)
-}
-
-# Note: The matrix exponential function is also available in the msm library as
-# MatrixExp(), but that seems to perform poorly for large t.  So use mexp here, 
-# but also check with method 2 when assuming irreversibility.
-
-#--------------------------------------------------
-# method 2: P_14a and P_14b
-#-------------------------------------------------- 
-path.probs2 <- function(t, q13, q14, q34)
-{
-    q134 <- q13 + q14
-    e1t <- exp(-q134 * t)
-
-    ppa <- q14 / q134 * (1 - e1t)
-
-    if (abs(q34 - q13 - q14) > 1e-10)
-    {
-        ppb <- q13 / q134 * 
-              (1 - 1/(q134-q34) * (q134 * exp(-q34*t) - q34 * e1t) )
-    } else
-    {
-        ppb <- q13 / q34 * (1 - (1 + q34 * t) * exp(-q34 * t))
-    }
-
-    ans <- c(ppa, ppb)
-    names(ans) <- c("Wa", "Wb")
-
-    return(ans)
-}
-
-#--------------------------------------------------
-# usage
-#-------------------------------------------------- 
-
-# 1 = SI, D
-# 3 = SC, D
-# 4 = SC, P
-# Wa = 1 -> 4
-# Wb = 1 -> 3 -> 4
-
-# median values from MuSSE_ploidysinodip250K.log (via tracer)
-q14 <- 0.0779         # ID -> CP, rho_I
-q13 <- 0.1895         # ID -> CD, q_IC
-q34 <- 0.0188         # CD -> CP, rho_C
-q41 <- q31 <- q43 <- 0 # irreversible
-
-# also use CID estimates?
-q14 <- 0.0228         # ID -> CP, rho_I
-q13 <- 0.0703         # ID -> CD, q_IC
-q34 <- 0.0334         # CD -> CP, rho_C
[…]

### 2019-09-05 twostateSSE/bisse250K/bisseDPv1.sh

@@ -1,11 +1,0 @@
-#!/bin/bash -l
-#PBS -l pmem=500mb
-#PBS -l nodes=1:ppn=1
-#PBS -m abe
-#PBS -M rzenil@umn.edu
-#PBS -o bisseoutput.log
-#PBS -e bisseerrorscluster.log
-#PBS -l walltime=96:00:00
-module load gcc
-
-/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/bisse/bisseDP.Rev

### 2019-09-05 twostateSSE/bisseSI250K/bisseselfincompv1.sh

@@ -1,11 +1,0 @@
-#!/bin/bash -l
-#PBS -l pmem=500mb
-#PBS -l nodes=1:ppn=1
-#PBS -m abe
-#PBS -M rzenil@umn.edu
-#PBS -o bisseSIoutput.log
-#PBS -e bisseSIerrorscluster.log
-#PBS -l walltime=96:00:00
-module load gcc
-
-/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/bissesi/bisseSI.Rev

### 2019-09-05 twostateSSE/bisseSInoreturn250K/bisseSInoreturnv1.sh

@@ -1,12 +1,0 @@
-#!/bin/bash -l
-#PBS -l pmem=500mb
-#PBS -l nodes=1:ppn=1
-#PBS -m abe
-#PBS -M rzenil@umn.edu
-#PBS -o bissesinoret.log
-#PBS -e bissesinoreterrorscluster.log
-#PBS -l walltime=96:00:00
-module load gcc
-
-/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/bissesinoret/bisseSInoreturn.Rev
-

### 2019-09-05 twostateSSE/bissenodip250K/bisseDPnodipv1.sh

@@ -1,11 +1,0 @@
-#!/bin/bash -l
-#PBS -l pmem=500mb
-#PBS -l nodes=1:ppn=1
-#PBS -m abe
-#PBS -M rzenil@umn.edu
-#PBS -o bissenodipoutput.log
-#PBS -e bissenodiperrorscluster.log
-#PBS -l walltime=96:00:00
-module load gcc
-
-/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/bissenodip/bisseDPnodiploidization.Rev

### 2019-09-05 twostateSSE/bisseoutput/bisseplotresults.R

@@ -1,67 +1,0 @@
-#source("https://bioconductor.org/biocLite.R")
-#biocLite("ggtree")
-#biocLite("treeio")
-
-library(RevGadgets)
-library(ggplot2)
-# plot the ancestral states tree generated by RevBayes
-
-
-p = plot_ancestral_states("~/revbayes-master/projects/cmake/output/anc_states_summaryBiSSE_polynodip.tree", size=0.4, include_start_states=FALSE, summary_statistic="MAP",tip_label_size=0.5, tip_label_offset=0.4,tip_label_italics=TRUE,node_label_size=2,node_label_nudge_x=0.1,shoulder_label_size=2,alpha=.3,node_size_range=c(2, 6))
-print(p)
-
-
-
-
-
-# using ggtree and ggplot2 functions we can now add all sorts of details
-
-# to our figure:
-
-
-
-# let's add a time scale
-
-root_age = 10.0
-
-p = p + scale_x_continuous(breaks=c(10, 5, 0), minor_breaks=c(2.5, 7.5), labels=c("0", "5", "10"))
-
-p = p + theme(panel.grid.major = element_line(color="darkgrey", size=.2, linetype="dashed"),
-              
-              panel.grid.minor = element_line(color="grey", size=.2, linetype="dashed"),
-              
-              panel.grid.major.y = element_blank(),
-              
-              panel.grid.minor.y = element_blank(),
-              
-              legend.position="left")
-
-
-
-# annotate some of the clades
-
-p = p + geom_cladelabel(node=60, label="Isotrema", align=TRUE, offset=5.0, size=2)
-
-p = p + geom_cladelabel(node=36, label="Aristolochia", align=TRUE, offset=5.0, size=2)
-
-
-
-# add an arrow to show a possible dysploid speciation event
-
-arrow_xend = 9.65
-
-arrow_yend = 26.75
-
-p = p + geom_segment(x=arrow_xend-0.4, y=arrow_yend+0.6, xend=arrow_xend, yend=arrow_yend, 
-                     
-                     size=0.5, color="grey35", arrow=arrow(length=unit(0.2, "cm"), type="open", angle=30), inherit.aes=FALSE)
-
-
-
-# set the final dimensions and write pdf
-
-p = p + coord_cartesian(xlim = c(0, 16.5))
-
-print(p)
-
-ggsave("examples/aristolochia/data/aristolochia_ancestral_states.pdf", width = 11, height = 9)

### 2019-09-05 twostateSSE/hisse250K/hisseDPv1.sh

@@ -1,11 +1,0 @@
-#!/bin/bash -l
-#PBS -l pmem=500mb
-#PBS -l nodes=1:ppn=1
-#PBS -m abe
-#PBS -M rzenil@umn.edu
-#PBS -o hisseoutput.log
-#PBS -e hisseerrorscluster.log
-#PBS -l walltime=96:00:00
-module load gcc
-
-/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/hisse/hisseDP.Rev

### 2019-09-05 twostateSSE/hisseSInoreturn250K/hissesinoretv1.sh

@@ -1,11 +1,0 @@
-#!/bin/bash -l
-#PBS -l pmem=500mb
-#PBS -l nodes=1:ppn=1
-#PBS -m abe
-#PBS -M rzenil@umn.edu
-#PBS -o hisseoutput.log
-#PBS -e hisseerrorscluster.log
-#PBS -l walltime=96:00:00
-module load gcc
-
-/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/hissesinoret/hissesinoret.Rev

### 2019-09-05 twostateSSE/hissenodip250K/hissenodipDPv1.sh

@@ -1,11 +1,0 @@
-#!/bin/bash -l
-#PBS -l pmem=500mb
-#PBS -l nodes=1:ppn=1
-#PBS -m abe
-#PBS -M rzenil@umn.edu
-#PBS -o hisseoutput.log
-#PBS -e hisseerrorscluster.log
-#PBS -l walltime=96:00:00
-module load gcc
-
-/home/eeg/shared/revbayes/projects/cmake/rb  /home/eeg/shared/twostatesse/hissenodip/hissenodipDP.Rev

### 2019-09-05 twostateSSE/plotancstates.R

@@ -1,716 +1,0 @@
-require(colorspace)
-
-# modified from inset
-inset.revgadgets = function (tree_view, insets, width = 0.1, height = 0.1, hjust = 0, 
-                             vjust = 0, x = "node", pos = 0.5) 
-{
-  require(magrittr)
-  df <- tree_view$data[as.numeric(names(insets)), ]
-  x <- match.arg(x, c("node", "branch", "edge","parent_shoulder"))
-  if (x == "node") {
-    xx <- df$x
-  }
-  else if (x == "parent_shoulder") {
-    xx <- df$x[ match(df$parent, df$node) ]
-  }
-  else {
-    xx <- df$branch
-  }
-  yy <- df$y
-  xx <- xx - hjust
-  yy <- yy - vjust
-  for (i in seq_along(insets)) {
-    tree_view %<>% subview(insets[[i]], x = xx[i], y = yy[i], 
-                           width = width, height = height)
-  }
-  return(tree_view)
-}
-
-# modified from https://github.com/GuangchuangYu/ggtree/blob/master/R/tree-utilities.R
-getXcoord2 <- function(x, root, parent, child, len, start=0, rev=FALSE) {
-  x[root] <- start
-  x[-root] <- NA  ## only root is set to start, by default 0
-  
-  currentNode <- root
-  direction <- 1
-  if (rev == TRUE) {
-    direction <- -1
-  }
-  while(anyNA(x)) {
-    idx <- which(parent %in% currentNode)
-    newNode <- child[idx]
-    x[newNode] <- x[parent[idx]]+len[idx] * direction
-    currentNode <- newNode
-  }
-  
-  return(x)
-}
-
-# modified from https://github.com/GuangchuangYu/ggtree/blob/master/R/tree-utilities.R
-getXcoord <- function(tr) {
-  edge <- tr$edge
-  parent <- edge[,1]
-  child <- edge[,2]
-  root <- ggtree:::getRoot(tr)
-  
-  len <- tr$edge.length
-  
-  N <- ggtree:::getNodeNum(tr)
-  x <- numeric(N)
-  x <- getXcoord2(x, root, parent, child, len)
-  return(x)
-}
-
-# modified from https://github.com/GuangchuangYu/ggtree/blob/master/R/tree-utilities.R
-getYcoord <- function(tr, step=1) {
[…]
